### 2021-05-03 Manuscript/SS_GAM.tex
Create manuscript

\documentclass[AMA,STIX1COL,]{WileyNJD-v2}

% Pandoc citation processing

\articletype{Research article}

\received{2017-01-01}

\revised{2017-02-01}

\accepted{2017-03-01}

\raggedbottom

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title{A demonstration of the \LaTeX class file for Statistics in
Medicine with Rmarkdown}

\author[a]{Boyi Guo*}
\author[a]{Nengjun Yi}

\authormark{Guo and Yi}

\address[a]{Department of Biostatistics, University of Alabama at
Birmingham, Birmingham, CUSA}

\corres{*Corresponding author name, This is sample corresponding
address. \email{xxx@uab}}

\presentaddress{This is sample for present address text this is sample
for present address text}

\abstract{TODO: Change this accordingly; Lorem ipsum dolor sit amet,
consectetur adipiscing elit. Aenean ut elit odio. Donec fermentum tellus
neque, vitae fringilla orci pretium vitae. Fusce maximus finibus
facilisis. Donec ut ullamcorper turpis. Donec ut porta ipsum. Nullam
cursus mauris a sapien ornare pulvinar. Aenean malesuada molestie erat
quis mattis. Praesent scelerisque posuere faucibus. Praesent nunc nulla,
ullamcorper ut ullamcorper sed, molestie ut est. Donec consequat libero
nisi, non semper velit vulputate et. Quisque eleifend tincidunt ligula,
bibendum finibus massa cursus eget. Curabitur aliquet vehicula quam non
pulvinar. Aliquam facilisis tortor nec purus finibus, sit amet elementum
eros sodales. Ut porta porttitor vestibulum. Integer molestie, leo ut
maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit
amet dolor. Sed sed tincidunt ex. Curabitur imperdiet egestas tellus in
iaculis. Maecenas ante neque, pretium vel nisl at, lobortis lacinia
neque. In gravida elit vel volutpat imperdiet. Sed ut nulla arcu. Proin
blandit interdum ex sit amet laoreet. Phasellus efficitur, sem hendrerit
mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac
ipsum. Aenean tincidunt nisl sit amet facilisis faucibus. Donec odio
erat, bibendum eu imperdiet sed, gravida luctus turpis.}

\keywords{Class file; \LaTeX; Statist. Med.; Rmarkdown;}

\maketitle

\hypertarget{the-article-header-information}{%
\section{The Article Header
Information}\label{the-article-header-information}}

YAML header:

\begin{verbatim}
output:
  rticles::sim_article:
    keep_tex: TRUE
\end{verbatim}

Configure the YAML header including the following elements:

\begin{itemize}
\item
  \texttt{title}: Title
\item
  \texttt{author}: List of author(s) containing \texttt{name} and
  \texttt{num}
\item
  \texttt{address}: List containing \texttt{num} and \texttt{org} for
  defining \texttt{author} affiliations
\item
  \texttt{presentaddress}: Not sure what they mean with this
\item
  \texttt{corres}: Author and address for correspondence
\item
  \texttt{authormark}: Short author list for header
\item
  \texttt{received}, \texttt{revised}, \texttt{accepted}: dates of
  submission, revision, and acceptance of the manuscript
\item
  \texttt{abstract}: Limited to 250 words
\item
  \texttt{keywords}: Up to 6 keywords
\item
  \texttt{bibliography}: BibTeX \texttt{.bib} file
\item
  \texttt{classoption}: options of the \texttt{WileyNJD-v2} class
\item
  \texttt{longtable}: set to \texttt{true} to include the
  \texttt{longtable} package, used by default from \texttt{pandoc} to
  convert markdown to \LaTeX code
\end{itemize}

\hypertarget{remarks}{%
\subsection{Remarks}\label{remarks}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In \texttt{authormark} use \emph{et al.} if there are three or more
  authors.
\item
  Note the use of \texttt{num} to link names and addresses.
\item
  For submitting a double-spaced manuscript, add \texttt{doublespace} as
  an option to a \texttt{classoption} line in the YAML header:
  \texttt{classoption:\ doublespace}.
\item
  Keywords are separated by semicolons.
\end{enumerate}

\hypertarget{the-body-of-the-article}{%
\section{The Body of the Article}\label{the-body-of-the-article}}

\hypertarget{mathematics}{%
\subsection{Mathematics}\label{mathematics}}

Use mathematics in Rmarkdown as usual.

\hypertarget{figures-and-tables}{%
\subsection{Figures and Tables}\label{figures-and-tables}}

Figures are supported from R code:

\begin{verbatim}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
\end{verbatim}

\begin{figure}
\centering
\includegraphics{SS_GAM_files/figure-latex/plot-ref-1.pdf}
\caption{Fancy Caption\label{fig:plot}}
\end{figure}

\ldots and can be referenced (Figure \ref{fig:plot}) by including the
\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
\texttt{fig.cap} attribute of the R chunk:
\texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.
It is a quirky hack at the moment, see
\href{https://github.com/yihui/knitr/issues/323}{here}.

Analogously, use Rmarkdown to produce tables as usual:

\begin{verbatim}
if (!require("xtable")) install.packages("xtable")
\end{verbatim}

\begin{verbatim}
## Loading required package: xtable
\end{verbatim}

\begin{verbatim}
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
\end{verbatim}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & speed & dist \\ 
  \hline
1 & 4.00 & 2.00 \\ 
  2 & 4.00 & 10.00 \\ 
  3 & 7.00 & 4.00 \\ 
  4 & 7.00 & 22.00 \\ 
  5 & 8.00 & 16.00 \\ 
  6 & 9.00 & 10.00 \\ 
   \hline
\end{tabular}
\caption{A table} 
\label{tab:table}
\end{table}

Referenced via \ref{tab:table}. You can also use the YAML option
\texttt{header-includes} to includes custom \LaTeX packages for tables
(keep in mind that \texttt{pandoc} uses \texttt{longtables} by default,
and it is hardcoded; some things may require including the package
\texttt{longtable}). E.g., using \texttt{ctable}:

\begin{verbatim}
header-includes:
- \usepackage{ctable}
\end{verbatim}

Then, just write straight-up \LaTeX code and reference is as usual
(\texttt{\textbackslash{}ref\{tab:ctable\}}):

\begin{verbatim}
\ctable[cap = {Short caption},
        caption = {A long, long, long, long, long caption for this table.},
        label={tab:ctable},]
        {cc}
        {
        \tnote[$\ast$]{Footnote 1}
        \tnote[$\dagger$]{Other footnote}
        \tnote[b]{Mistakes are possible.}
        }{
        \FL
        COL 1\tmark[a] & COL 2\tmark[$\ast$]
        \ML
        6.92\tmark[$\dagger$] & 0.09781 \\
        6.93\tmark[$\dagger$] & 0.09901 \\
        97 & 2000
        \LL
}
\end{verbatim}

It is also possible to set the \texttt{YAML} option
\texttt{longtable:\ true} and use markdown tables (or the
\texttt{knitr::kable} function): \texttt{knitr::kable(head(cars))}
produces the same table as the \texttt{xtable} example presented before.

\hypertarget{cross-referencing}{%
\subsection{Cross-referencing}\label{cross-referencing}}

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using
\texttt{{[}@\textless{}name\textgreater{}{]}}, equivalent of
\texttt{\textbackslash{}ref\{\textless{}name\textgreater{}\}} and
\texttt{\textbackslash{}label\{\textless{}name\textgreater{}\}}). That
works well for citations in Rmarkdown, not so well for figures and
tables. In that case, it is possible to revert to standard
\LaTeX syntax.

\hypertarget{double-spacing}{%
\subsection{Double Spacing}\label{double-spacing}}

If you need to double space your document for submission please use the
\texttt{doublespace} option in the header.

\hypertarget{bibliography}{%
\section{Bibliography}\label{bibliography}}

Link a \texttt{.bib} document via the YAML header, and bibliography will
be printed at the very end (as usual). The default bibliography style is
provided by Wiley as in \texttt{WileyNJD-AMA.bst}, do not delete that
file.

Use the Rmarkdown equivalent of the \LaTeX citation system using
\texttt{{[}@\textless{}name\textgreater{}{]}}. Example:
\citep{Taylor1937}, \citep{Knupp1999, Kamm2000}.

To include all citation from the \texttt{.bib} file, add
\texttt{\textbackslash{}nocite\{*\}} before the end of the document.

\hypertarget{further-information}{%
\section{Further information}\label{further-information}}

All \LaTeX enviroments supported by the main template are supported here
as well; see the \texttt{.tex} sample file
\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm}{here}
for more details and example.

\bibliography{bibfile.bib}

\end{document}

### 2021-05-03 Manuscript/main.Rmd
Create manuscript

---
title: A demonstration of the \LaTeX class file for Statistics in Medicine with Rmarkdown
author:
- name: Boyi Guo*
  num: a
- name: Nengjun Yi
  num: a
address:
- num: a
  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, CUSA
# - num: b
  # org: Department of Applied Things, University B, City B, Country B
corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
presentaddress: This is sample for present address text this is sample for present address text
authormark: Guo and Yi
articletype: Research article
received: 2017-01-01
revised: 2017-02-01
accepted: 2017-03-01
abstract: "TODO: Change this accordingly; Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut elit odio. Donec fermentum tellus neque, vitae fringilla orci pretium vitae. Fusce maximus finibus facilisis. Donec ut ullamcorper turpis. Donec ut porta ipsum. Nullam cursus mauris a sapien ornare pulvinar. Aenean malesuada molestie erat quis mattis. Praesent scelerisque posuere faucibus. Praesent nunc nulla, ullamcorper ut ullamcorper sed, molestie ut est. Donec consequat libero nisi, non semper velit vulputate et. Quisque eleifend tincidunt ligula, bibendum finibus massa cursus eget. Curabitur aliquet vehicula quam non pulvinar. Aliquam facilisis tortor nec purus finibus, sit amet elementum eros sodales. Ut porta porttitor vestibulum. Integer molestie, leo ut maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit amet dolor. Sed sed tincidunt ex. Curabitur imperdiet egestas tellus in iaculis. Maecenas ante neque, pretium vel nisl at, lobortis lacinia neque. In gravida elit vel volutpat imperdiet. Sed ut nulla arcu. Proin blandit interdum ex sit amet laoreet. Phasellus efficitur, sem hendrerit mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac ipsum. Aenean tincidunt nisl sit amet facilisis faucibus. Donec odio erat, bibendum eu imperdiet sed, gravida luctus turpis."
keywords: Class file; \LaTeX; Statist. Med.; Rmarkdown;
bibliography: bibfile.bib
output: 
  rticles::sim_article:
    keep_tex: TRUE
---

# The Article Header Information

YAML header:

```
output:
  rticles::sim_article:
    keep_tex: TRUE
```

Configure the YAML header including the following elements:

* `title`: Title

* `author`: List of author(s) containing `name` and `num`

* `address`: List containing `num` and `org` for defining `author` affiliations

* `presentaddress`: Not sure what they mean with this

* `corres`: Author and address for correspondence

* `authormark`: Short author list for header

* `received`, `revised`, `accepted`: dates of submission, revision, and acceptance of the manuscript

* `abstract`: Limited to 250 words

* `keywords`: Up to 6 keywords

* `bibliography`: BibTeX `.bib` file

* `classoption`: options of the `WileyNJD-v2` class

* `longtable`: set to `true` to include the `longtable` package, used by default from `pandoc` to convert markdown to \LaTeX code

## Remarks

1. In `authormark` use _et al._ if there are three or more authors.

2. Note the use of `num` to link names and addresses.

3. For submitting a double-spaced manuscript, add `doublespace` as an option to a `classoption` line in the YAML header: `classoption: doublespace`.

4. Keywords are separated by semicolons.

# The Body of the Article

## Mathematics

Use mathematics in Rmarkdown as usual.

## Figures and Tables

Figures are supported from R code:

```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`. It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).

Analogously, use Rmarkdown to produce tables as usual:

```{r, results = "asis"}
if (!require("xtable")) install.packages("xtable")
xt <- xtable(head(cars), caption = "A table", label = "tab:table")
print(xt, comment = FALSE)
```

Referenced via \ref{tab:table}. You can also use the YAML option `header-includes` to includes custom \LaTeX packages for tables (keep in mind that `pandoc` uses `longtables` by default, and it is hardcoded; some things may require including the package `longtable`). E.g., using `ctable`:
```
header-includes:
- \usepackage{ctable}
```
Then, just write straight-up \LaTeX code and reference is as usual (`\ref{tab:ctable}`):
```
\ctable[cap = {Short caption},
        caption = {A long, long, long, long, long caption for this table.},
        label={tab:ctable},]
        {cc}
        {
        \tnote[$\ast$]{Footnote 1}
        \tnote[$\dagger$]{Other footnote}
        \tnote[b]{Mistakes are possible.}
        }{
        \FL
        COL 1\tmark[a] & COL 2\tmark[$\ast$]
        \ML
        6.92\tmark[$\dagger$] & 0.09781 \\
        6.93\tmark[$\dagger$] & 0.09901 \\
        97 & 2000
        \LL
}
```

It is also possible to set the `YAML` option `longtable: true` and use markdown tables (or the `knitr::kable` function): `knitr::kable(head(cars))` produces the same table as the `xtable` example presented before.

## Cross-referencing

The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`). That works well for citations in Rmarkdown, not so well for figures and tables. In that case, it is possible to revert to standard \LaTeX syntax.

## Double Spacing

If you need to double space your document for submission please use the `doublespace` option in the header.

# Bibliography

Link a `.bib` document via the YAML header, and bibliography will be printed at the very end (as usual). The default bibliography style is provided by Wiley as in `WileyNJD-AMA.bst`, do not delete that file.

Use the Rmarkdown equivalent of the \LaTeX citation system using `[@<name>]`. Example: [@Taylor1937], [@Knupp1999; @Kamm2000].

To include all citation from the `.bib` file, add `\nocite{*}` before the end of the document.

# Further information

All \LaTeX enviroments supported by the main template are supported here as well; see the `.tex` sample file [here](http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm) for more details and example.

### 2021-05-03 _targets.R
Run pipeline

@@ -1,0 +1,18 @@
+library(targets)
+library(tarchetypes)
+# This is an example _targets.R file. Every
+# {targets} pipeline needs one.
+# Use tar_script() to create _targets.R and tar_edit()
+# to open it again for editing.
+# Then, run tar_make() to run the pipeline
+# and tar_read(summary) to view the results.
+
+options(tidyverse.quiet = TRUE)
+# Set target-specific options such as packages.
+tar_option_set(packages = c("dplyr", "tidyverse", "rticles"))
+
+# End this file with a list of target objects.
+list(
+  # tar_target(data, data.frame(x = sample.int(100), y = sample.int(100))),
+  tar_render(manu, "Manuscript/main.Rmd", output_file = "SS_GAM.pdf")
+)

### 2021-05-07 Real Data/curatedBladderData.R
Add folder for real data analysis.

@@ -1,0 +1,22 @@
+# if (!requireNamespace("BiocManager", quietly = TRUE))
+#   install.packages("BiocManager")
+# 
+# BiocManager::install("curatedBladderData")
+
+library(curatedBladderData)
+
+
+data(package="curatedBladderData")
+
+data("GSE32894_eset")
+
+dat <- GSE32894_eset
+pheno_dat <- pData(dat) %>%
+  select(-uncurated_author_metadata) %>% 
+  mutate_if(is.logical, factor) %>% 
+  mutate_if(is.character, factor) %>% 
+  janitor::remove_constant() 
+summary(pheno_dat)
+
+
+# TODO: write a function that display the event rate "death rate" for all the datasets

### 2021-05-07 Manuscript/main.Rmd
Set up mansucript files

@@
 
-# The Article Header Information
-
-YAML header:
-
-```
-output:
-  rticles::sim_article:
-    keep_tex: TRUE
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-Configure the YAML header including the following elements:
+```{r child="01-Intro.Rmd"}
+```
 
-* `title`: Title
+```{r child="02-Method.Rmd"}
+```
 
-* `author`: List of author(s) containing `name` and `num`
+```{r child="03-Simulation.Rmd"}
+```
 
-* `address`: List containing `num` and `org` for defining `author` affiliations
+```{r child="04-Real_Data.Rmd"}
+```
 
-* `presentaddress`: Not sure what they mean with this
-
-* `corres`: Author and address for correspondence
-
-* `authormark`: Short author list for header
-
-* `received`, `revised`, `accepted`: dates of submission, revision, and acceptance of the manuscript
-
-* `abstract`: Limited to 250 words
-
-* `keywords`: Up to 6 keywords
-
-* `bibliography`: BibTeX `.bib` file
-
-* `classoption`: options of the `WileyNJD-v2` class
-
-* `longtable`: set to `true` to include the `longtable` package, used by default from `pandoc` to convert markdown to \LaTeX code
-
-## Remarks
-
-1. In `authormark` use _et al._ if there are three or more authors.
-
-2. Note the use of `num` to link names and addresses.
-
-3. For submitting a double-spaced manuscript, add `doublespace` as an option to a `classoption` line in the YAML header: `classoption: doublespace`.
-
-4. Keywords are separated by semicolons.
-
-# The Body of the Article
-
-## Mathematics
-
-Use mathematics in Rmarkdown as usual.
-
-## Figures and Tables
-
-Figures are supported from R code:
+```{r child="05-Discussion.Rmd"}
+```
 
@@
 ```
-
-Referenced via \ref{tab:table}.
-You can also use the YAML option `header-includes` to includes custom \LaTeX packages for tables (keep in mind that `pandoc` uses `longtables` by default, and it is hardcoded; some things may require including the package `longtable`).
-E.g., using `ctable`:
-```
-header-includes:
-- \usepackage{ctable}
-```
-Then, just write straight-up \LaTeX code and reference is as usual (`\ref{tab:ctable}`):
-```
-\ctable[cap = {Short caption},
-        caption = {A long, long, long, long, long caption for this table.},
-        label={tab:ctable},]
-        {cc}
-        {
-        \tnote[$\ast$]{Footnote 1}
-        \tnote[$\dagger$]{Other footnote}
-        \tnote[b]{Mistakes are possible.}
-        }{
-        \FL
-        COL 1\tmark[a] & COL 2\tmark[$\ast$]
-        \ML
-        6.92\tmark[$\dagger$] & 0.09781 \\
-        6.93\tmark[$\dagger$] & 0.09901 \\
-        97 & 2000
-        \LL
-}
-```
-
-It is also possible to set the `YAML` option `longtable: true` and use markdown tables (or the `knitr::kable` function): `knitr::kable(head(cars))` produces the same table as the `xtable` example presented before.
 
@@
 # Further information
-
-All \LaTeX enviroments supported by the main template are supported here as well; see the `.tex` sample file [here](http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm) for more details and example.

### 2021-05-08 Real Data/curatedBladderData.R
Understanding the data. I think the curated Ovarian data is slightly better

@@ -4,11 +4,36 @@
 # BiocManager::install("curatedBladderData")
 
 library(curatedBladderData)
+library(tidyverse)
 
 
-data(package="curatedBladderData")
+dat_lst <- data(package="curatedBladderData")
 
-data("GSE32894_eset")
+#### Display event rate for all datasets in the package ####
+data_brief <- map_dfr(dat_lst$results[,"Item"], .f = function(dat_name){
+  
+  eval(rlang::expr(data(!!dat_name)))
+  
+  dat <- eval(rlang::parse_expr(dat_name))
+  
+  peno_dat <- pData(dat)
+  
+  
+  
+  # remove loaded data
+  eval(rlang::expr(rm(dat)))
+  eval(rlang::expr(rm(!!dat_name)), envir = globalenv())
+  
+  return(table(peno_dat$vital_status) %>% c %>% t %>%
+           data.frame(dataset = dat_name))
+})
+
+
+
+
+
+
+
 
 dat <- GSE32894_eset
 pheno_dat <- pData(dat) %>%
@@ -19,4 +44,4 @@ pheno_dat <- pData(dat) %>%
 summary(pheno_dat)
 
 
-# TODO: write a function that display the event rate "death rate" for all the datasets
+# TODO: write aa function that display the event rate "death rate" for all the datasets

### 2021-05-10 Manuscript/01-Intro.Rmd
Write manuscript

# Introduction
* What is the problem
  * what is the GAM
  * expanding GAM in to high-dimensional frame work
  * Why this is important
  * Why this is hard
* Previous solutions
  * What are the methods
  * What are their problems

* Introduce the wiggliness matrix

* Our method
* The implication of our method
* Structure of the rest of the paper

### 2021-05-10 Manuscript/02-Method.Rmd
Write manuscript

# Spike-and-slab Additive Model

Following the notation in the previous section \ref{Notation}, suppose we have the model matrix $\bs X$ derived from spline functions of choice $h_j, j = 1, \dots, p.$ for the $p$ covariates that are modeled nonparametrically, the total dimension of $\bs X$ is $n \times \sum\limits_{j=1}^p k_j$. Column-wise centering and re-parameterization introduced in the GAM as random effect section[^5] are performed. The re-parameterization has two purposes. First of all the re-parameterization reshapes the columns to be linearly independent to each other, and hence, independent prior can be set to the coefficients. Secondly, the re-parameterization decomposes the model matrix $\bs X_j$ for the $j$th covariates to two sub-matrices representing the null space of the spline and penalized space of the spline,
$$
\bs X_j = \begin{bmatrix}
\bs X_j^{0} & \bs X_j^{pen}
\end{bmatrix},
$$
where $\bs X_j^{null}$ is the column space representing the null space of the spline model, and $\bs X_j^{pen}$ denotes the column spaces that need to be shrunk for smoothness of the spline. The purpose for this re-parameterization is to isolate the null space from the smoothing space, such that answering the question, if non-linear relationship is necessary, becomes easy by deciding if the corresponding coefficients of the $\bs X_j^{pen}$ are jointly zero. This strategy is widely used, previously mentioned in @Scheipl2012 and the references there in.^[Possibly select some to add here.] Hereafter, we denote the coefficients of the null space as $\bs \beta^{0}_j$ and the coefficients of the penalized space as $\bs \beta^{pen}_j$. The spline function for the $j$the covariate is 
$$
f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
$$
where the superscript $^t$ denotes matrix transpose.

Without loss of generality, we assume that the model doesn't include the the parametric part of the model $\bs A$. The coefficients of $\bs A$ can be estimated easily in the proposed EM algorithms by local update conditioning on the temporary value of the spline coefficients.

 \

## Spike and slab spline prior

A generalized spline model can be formulated as 
$$
y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
$$
The link function of the mean can be expressed in a matrix form
$$
g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
 = \beta^T X_i,
$$
where
$$
\begin{aligned}
  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
& &
  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
\end{aligned}
$$

\

### Mixture-Normal Prior
Given the full deign matrix $X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T$, the likelihood function of the generalized model can be written as
$$
p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
$$

Considering the dependency of spline bases, the grouping nature of the bases should be taken care of. Naturally, the Bayesian hierarchical models allow us easily to take account of the dependency of the spline bases as grouping information and provide more reliable estimation. Our hierarchical GAMs specify the spike-and-slab mixture normal prior on the coefficients: for the coefficients of the $j$th predictor, $\beta^0_j$ and $\beta^{pen}_{jk}$, we have
$$
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1), 
\end{aligned}
$$
where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^{pen}_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and non-linear effect of the $j$th variable respectively; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.  Thus, the priors are formulated as a mixture of the shrinkage prior normal(0, $s_0$) and the weakly informative prior normal(0, $s_1$), which are spike and slab components respectively. Like any other spike and slab priors, the spike density is to contain the minimum to zero effects, while the slab density is to allow large effects. The scale parameters $s_0$ and $s_1$ can be treated as tuning parameters, which can be optimized via cross-validation. A discussion of how to choose the scale parameters comes later.

To note here, unlike the original spike-and-slab prior where each predictor have only indicator variable, $\gamma_j$, we have two indicators associated with each predictor, $\gamma^0_j$ and $\gamma^{pen}_j$ to decide the inclusion of the linear effect and non-linear effect of the predictor respectively. It is possible to add more constraints on the model, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e. $\gamma_j = \gamma^0_j = \gamma^{pen}_j$. Such formulations, commonly referred as "all-in-all-out", has been implemented in many high dimensional Bayesian GAMs, see Yang & Narisetty (2020) and Bai (Work in progress). The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not. On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e. $\gamma^{pen}_{jk}$ of $\beta^{pen}_{jk}$ for $k = 1\dots, K_j$. This prior set-up generalized to the group spike-and-slab. The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity. The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup. Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.

We set up the hyper-prior of $\bs \gamma$ to allow local adaption of the shrinkage, using a binomial distribution. The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^{pen}_j$, shares the same probability parameter,
$$
\begin{aligned}
\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
\end{aligned}
$$

This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^{pen}$ decides the smoothness of the spline function at individual predictor level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity. However, this set-up could possibility result in a situation that is not theoretically possible: the non-linear component is selected, but the linear component is not. This can be addressed analytically by forcing to include the linear component when non-linear component is included. Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e. $\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.

 <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function. The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder. In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.  -->

We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
 The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. Specifically, we focus on a special case of beta distribution, uniform (0,1) for simplicity and convenience. To note, when the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer the hierarchical spline GAM with the spike-and-slab mixture normal prior as the ssGAM-MN.

 \

### Mixture Double Exponential Prior
One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects. In contrast, distributions with heavier tails are preferred, for example $t$ distribution with small degree of freedom and double exponential distribution. Specifically, double exponential prior is equivalent to the LASSO penalty in the frequentist models, providing a sparse solution. Meanwhile, the estimation algorithm of LASSO penalty are extremely efficient, greatly reducing the computation needs. Recent developments of spike-and-slab priors include spike-and-slab LASSO prior, i.e. spike-and-slab mixture double exponential prior. We extend the spike-and-slab LASSO prior to high-dimensional GAMs, providing computationally efficiency and sparse solution to our proposed ssGAM-MN.

The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we replace the normal distribution with double exponential distribution. The prior distributions of the coefficients $\beta^0_j, \beta^{pen}_j$ for the $j$ predictor can be formulated as
$$
\begin{aligned}
  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
\end{aligned}
$$
The hyper-priors for $\gamma^0_j, \gamma^{pen}_j, \theta_j$ are kept the same as specified in ssGAM-MN.

\

## Alforithm for fitting ssGAMs

Parsimonious computation is always encouraged in high-dimension data analysis. Normally, Bayesian methodology loses its advantages over Frequentest penalized model because of the computation cost. Previous Bayesian spline models heavily relies on the MCMC algorithm to establish posterior distribution of parameters. One of the exception is Bai (in progress), who a grouped spike-and-slab LASSO prior [@Rockova2018] to address the grouping structure of spline bases. He focused on the MAP estimator of the parameters, but failed to address the uncertainty measure of the estimates. The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the non-linear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.

In the follow section, we introduced two expectation-maximization (EM) based algorithms for fast computing, EM-IRLS and EM-coordinate Descent algorithm. Instead of simulating the posterior distribution via MCMC, we iteratively estimate the MAP. Compared to the EM-IRLS algorithm, the EM-coordinate descent provides an expedited model fitting process at the cost of model uncertainty.

\

### EM algorithms

EM algorithm is an iterative algorithm to find the maximum likelihood estimates or the Bayesian counterpart MAP estimates. It is commonly used when some necessary data to establish the likelihood function are missing. Instead, the algorithm maximizes the expectation, in respect to the "missing" data, of the likelihood function.

The recursive algorithm consists of two steps:

* E-step: to calculate the expectation of the likelihood function conditioning on some "missing" data
* M-step: to maximize the spectated likelihood function to calculate the parameter of interest

For ssGAM, we define the parameters of interest as $\Theta = {\bs \beta, \bs \theta, \phi}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
Our objective is to find the parameters $\Theta$ that maximize the posterior density function,or equivalently, the logarithm of the density function, 
$$
\begin{aligned}
& \text{argmax}_{\Theta}
\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
&= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
\end{aligned}
$$  
<!-- For simplicity, we consider the null space to be 1-dimension, and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{pen}$ from the coefficients of the the penalized space. Using the new notation, the part component $\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$. Without loos of generality, the model still work for the spline function whose dimensionality of the null space is more than 1. -->

We use the EM algorithm to find the MAP estimate of $\Theta$. Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as as the "missing data" in the EM algorithms. Naturally In the E-step, we calculate the expectation of posterior density function of $\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{t-1}$, 
$$
E_{\bs \gamma|\Theta^{t-1}}\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) .
$$
Hereafter, we use the short hand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{t-1}}(\cdot)$. In the M-step, we find the $\Theta^{t}$ that maximize $E_{\bs \gamma|\Theta^{t-1}}\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$. To note here, the log-posterior of ssGAM (up to additive constants) can be written as a two-part equation

$$ \log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\beta, \phi) + Q_2 (\gamma,\theta),$$
Where
$$ Q_1(\bs \beta, \phi) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]$$
and $$
Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .$$
$Q_1$ and $Q_2$ are respectively, the log-likelihood of model coefficients and the dispersion parameter $\bs \beta, \phi$, and latent indicators $\bs \gamma, \theta$. Conditioning on $\gamma_{jk}$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.

The E- and M- steps are iterated until the algorithm converge. Depending on the choice of the mixture priors, either mixture normal or mixture double exponential, the maximization in the M-step can take different algorithms, IRLS for uncertainty measures and coordinate descent for faster computation.

\

### EM-IRLS

The main idea of the EM-IRLS algorithm is to use the iterative re-weighted least square algorithm to find the estimate $\bs \beta, \phi$ that maximizes $E(Q_1)$. Recall the prior set-up of mixture normal prior, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$

$$
\begin{aligned}
p(\bs \beta | \bs\gamma) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],%\\
% p(\bs\gamma | \bs \theta) & = \prod\limits_{j=1}^{p}\prod\limits_{k=1}^{K_j} \theta_j^{\gamma_{jk}} (1-\theta_j)^{1-\gamma_{jk}}.
\end{aligned}
$$
and 
$$
\begin{aligned}
E(Q_1) &= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]%\\
%E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
\end{aligned}
$$
where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1$. To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv p(\gamma^{pen}_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem. The calculation for $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\beta_{jk}$ for $k=1, \dots, K_j$ and $p^0_j$ only depends on $\beta_j^0$. The calculation follows the equations below.

$$
\begin{aligned}
p_{j}^0 &= \frac{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + Pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
p_{j} &= \frac{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + Pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
\end{aligned}
$$

and $Pr(\gamma_{j}^{0} = 1|\theta_j) =Pr(\gamma_{j}^{pen} = 1|\theta_j) = \theta_j$, $Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^{pen} = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{Normal}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{Normal}(\beta|0 , s_0)$. It is trivial to show
$$
\begin{aligned}
&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{pen}_{j})  = p_{j}\\
&E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
\end{aligned}
$$

With $E({S^0}^{-1}_{j})$ and $E(S^{-1}_{j})$ known, $E(Q_1)$ can be also viewed as a $l_2$ penalized likelihood function: $E({S^0}^{-1}_{j})$ controls the shrinkage of the linear effect, informing if the predictor should be included in the model; $E(S^{-1}_{j})$ controls the smoothness of the spline function. $l_2$ penalized likelihood function can be maximize using the IWLS algorithm.  The likelihood function can be approximate by a weighted normal likelihood:
$$
  p(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
$$
    where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight, respectively.The pseudo data and the pseudo-weight are calculated by
    $$
    \begin{aligned}
    z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
    \end{aligned}
    $$
    where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function. The $l_2$ penalty, $ \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$. An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting. The estimates of $\bs \beta$ and $\phi$ can be quickly updated in the IRLS algorithm.

<!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients. Moreover, to note that, we have different shrinkage $S^{-1}_{jk}$for the coefficients $\beta_{jk}$ of the variable $x_j$, and hence, we can penalize the null space of the spline differently and allow local adaption. -->

The remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$. As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed form equation:
$$
\theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
$$

Totally, the framework of the proposed EM IRLS algorithm was summarized as follows:

1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$

2) Iterate over the E-step and M-step until convergence

E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration

M-step:

a) Update $\bs \beta^{(t)}$ using the IRLS algorithm

b) Update $\bs \theta^{(t)}$ using the closed form equation

We assess convergence by the criterion:
$|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where
$d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at
the $t$th iteration, and $\epsilon$ is a small value (say
$10^{-5}$).

\

### EM-Coordinate Descent

When the prior distribution is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$, replacing IRLS, in the M-step. Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.

Recall, the density function of spike-and-slab mixture double exponential prior can be written as
$$
f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}).
$$

This can simplify our posterior likelihood function to
$$
E(Q_1) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right]
,
$$
where the approximation of $E({S^0_j}^{-1}$ and $E(S^{-1}_{j})$ remains the same as in the EM-IRLS algorithm. The formulation of $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function, and optimized via coordinate descent algorithm. The estimates can be used to further estimates $\bs \theta$ in the M-step, following the same equation in the EM-IRLS.

Totally, the framework of the proposed EM-coordinate descent algorithm was summarized as follows:

1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$

2) Iterate over the E-step and M-step until convergence

E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration

M-step:

a) Update $\bs \beta^{(t)}$ using the coordinate descent algorithm

b) Update $\bs \theta^{(t)}$ using the closed form equation

We assess convergence by the criterion:
$|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where
$d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at
the $t$th iteration, and $\epsilon$ is a small value (say
$10^{-5}$).

 \

## Selecting Optimal Scale Values

Our proposed prior, spike and slab spline prior, requires two preset scale parameters ($s_0$, $s_1$). Hence, we would construct a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value. However, previous research on the topic^[TODO: add citation] suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value of $s_1 = 1$ that provides no or weak shrinkage. More attention is focused on examining different values of spike scale $s_0$. Instead of the 2-D grid, We consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$. Increasing the spike scale s_0 tends to include more non-zero coefficients in the model. An measure of preference, e.g. area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model. The procedure is similar to the lasso implemented in the widely used R package `glmnet`, which quickly fits the lasso model over a list of values of $\lambda$ covering the entire range, giving a sequence of models for users to choose from.

### 2021-05-10 Manuscript/main.Rmd
Write manuscript

@@
 - num: a
-  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, CUSA
+  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
 # - num: b
@@
 ---
+
+\newcommand{\bs}[1]{\boldsymbol{#1}}
 
@@
 
-## Double Spacing
+# References
+<div id="refs"></div>
 
-If you need to double space your document for submission please use the `doublespace` option in the header.
-
-# Bibliography
-
-Link a `.bib` document via the YAML header, and bibliography will be printed at the very end (as usual).
-The default bibliography style is provided by Wiley as in `WileyNJD-AMA.bst`, do not delete that file.
-
-Use the Rmarkdown equivalent of the \LaTeX citation system using `[@<name>]`. Example: [@Taylor1937], [@Knupp1999; @Kamm2000].
-
-To include all citation from the `.bib` file, add `\nocite{*}` before the end of the document.
-
-# Further information
+# Appendix

### 2021-05-10 Manuscript/main.tex
Create manuscript
Set up mansucript files
Write manuscript

--- 2021-05-03 Manuscript/SS_GAM.tex
@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \address[a]{Department of Biostatistics, University of Alabama at
-Birmingham, Birmingham, CUSA}
+Birmingham, Birmingham, USA}
 
@@
 
-\hypertarget{the-article-header-information}{%
-\section{The Article Header
-Information}\label{the-article-header-information}}
+\newcommand{\bs}[1]{\boldsymbol{#1}}
 
-YAML header:
-
-\begin{verbatim}
-output:
-  rticles::sim_article:
-    keep_tex: TRUE
-\end{verbatim}
-
-Configure the YAML header including the following elements:
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
 \begin{itemize}
+\tightlist
 \item
-  \texttt{title}: Title
+  What is the problem
 \item
-  \texttt{author}: List of author(s) containing \texttt{name} and
-  \texttt{num}
+  Previous solutions
 \item
-  \texttt{address}: List containing \texttt{num} and \texttt{org} for
-  defining \texttt{author} affiliations
+  Our method
 \item
-  \texttt{presentaddress}: Not sure what they mean with this
+  The implication of our method
 \item
-  \texttt{corres}: Author and address for correspondence
-\item
-  \texttt{authormark}: Short author list for header
-\item
-  \texttt{received}, \texttt{revised}, \texttt{accepted}: dates of
-  submission, revision, and acceptance of the manuscript
-\item
-  \texttt{abstract}: Limited to 250 words
-\item
-  \texttt{keywords}: Up to 6 keywords
-\item
-  \texttt{bibliography}: BibTeX \texttt{.bib} file
-\item
-  \texttt{classoption}: options of the \texttt{WileyNJD-v2} class
-\item
-  \texttt{longtable}: set to \texttt{true} to include the
-  \texttt{longtable} package, used by default from \texttt{pandoc} to
-  convert markdown to \LaTeX code
+  Structure of the rest of the paper
 \end{itemize}
 
-\hypertarget{remarks}{%
-\subsection{Remarks}\label{remarks}}
+\hypertarget{spike-and-slab-additive-model}{%
+\section{Spike-and-slab Additive
+Model}\label{spike-and-slab-additive-model}}
+
+Following the notation in the previous section \ref{Notation}, suppose
+we have the model matrix \(\boldsymbol{X}\) derived from spline
+functions of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates
+that are modeled nonparametrically, the total dimension of
+\(\boldsymbol{X}\) is \(n \times \sum\limits_{j=1}^p k_j\). Column-wise
+centering and re-parameterization introduced in the GAM as random effect
+section{[}\^{}5{]} are performed. The re-parameterization has two
+purposes. First of all the re-parameterization reshapes the columns to
+be linearly independent to each other, and hence, independent prior can
+be set to the coefficients. Secondly, the re-parameterization decomposes
+the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
+sub-matrices representing the null space of the spline and penalized
+space of the spline, \[
+\boldsymbol{X}_j = \begin{bmatrix}
+\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
+\end{bmatrix},
+\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
+the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
+denotes the column spaces that need to be shrunk for smoothness of the
+spline. The purpose for this re-parameterization is to isolate the null
+space from the smoothing space, such that answering the question, if
+non-linear relationship is necessary, becomes easy by deciding if the
+corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
+zero. This strategy is widely used, previously mentioned in
+\citet{Scheipl2012} and the references there in.\footnote{Possibly
+  select some to add here.} Hereafter, we denote the coefficients of the
+null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
+penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
+for the \(j\)the covariate is \[
+f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
+\] where the superscript \(^t\) denotes matrix transpose.
+
+Without loss of generality, we assume that the model doesn't include the
+the parametric part of the model \(\boldsymbol{A}\). The coefficients of
+\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
+by local update conditioning on the temporary value of the spline
+coefficients.
+
+~
+
+\hypertarget{spike-and-slab-spline-prior}{%
+\subsection{Spike and slab spline
+prior}\label{spike-and-slab-spline-prior}}
+
+A generalized spline model can be formulated as \[
+y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
+\] The link function of the mean can be expressed in a matrix form \[
+g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
+ = \beta^T X_i,
+\] where \[
+\begin{aligned}
+  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
+& &
+  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
+\end{aligned}
+\]
+
+\hfill\break
+
+\hypertarget{mixture-normal-prior}{%
+\subsubsection{Mixture-Normal Prior}\label{mixture-normal-prior}}
+
+Given the full deign matrix
+\(X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T\), the
+likelihood function of the generalized model can be written as \[
+p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
+\]
+
+Considering the dependency of spline bases, the grouping nature of the
+bases should be taken care of. Naturally, the Bayesian hierarchical
+models allow us easily to take account of the dependency of the spline
+bases as grouping information and provide more reliable estimation. Our
+hierarchical GAMs specify the spike-and-slab mixture normal prior on the
+coefficients: for the coefficients of the \(j\)th predictor,
+\(\beta^0_j\) and \(\beta^{pen}_{jk}\), we have \[
+\begin{aligned}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1), 
+\end{aligned}
+\] where \(\gamma^0_{j}\in\{0,1\}\) and
+\(\gamma^{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator variables,
+indicating if the model includes the linear effect and non-linear effect
+of the \(j\)th variable respectively; \(s_0\) and \(s_1\) are scale
+parameters, assuming \(0 < s_0 < s_1\) and given. Thus, the priors are
+formulated as a mixture of the shrinkage prior normal(0, \(s_0\)) and
+the weakly informative prior normal(0, \(s_1\)), which are spike and
+slab components respectively. Like any other spike and slab priors, the
+spike density is to contain the minimum to zero effects, while the slab
+density is to allow large effects. The scale parameters \(s_0\) and
+\(s_1\) can be treated as tuning parameters, which can be optimized via
+cross-validation. A discussion of how to choose the scale parameters
+comes later.
+
+To note here, unlike the original spike-and-slab prior where each
+predictor have only indicator variable, \(\gamma_j\), we have two
+indicators associated with each predictor, \(\gamma^0_j\) and
+\(\gamma^{pen}_j\) to decide the inclusion of the linear effect and
+non-linear effect of the predictor respectively. It is possible to add
+more constraints on the model, assuming that one indicator variable
+decides the inclusion of both the linear effect and non-linear effect,
+i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). Such formulations,
+commonly referred as ``all-in-all-out'', has been implemented in many
+high dimensional Bayesian GAMs, see Yang \& Narisetty (2020) and Bai
+(Work in progress). The ``all-in-all-out'' set-up was criticized for its
+inability to answering an important research question, if the effects
+are non-linear or not. On the other hand, the model can be much more
+flexible by allowing each dimension of the penalized space to have an
+independent indicator, i.e.~\(\gamma^{pen}_{jk}\) of
+\(\beta^{pen}_{jk}\) for \(k = 1\dots, K_j\). This prior set-up
+generalized to the group spike-and-slab. The additional flexibility
+doesn't not necessarily grand the model neither theoretical
+justification nor analytic simplicity. The spline penalty assumes a
+common smooth shrinkage \(\lambda_j\) applied to all coefficients of the
+penalized space, which contradicts to the model setup. Besides, having
+multiple indicator variables complicates the decision of if non-linear
+effects are necessary.
+
+We set up the hyper-prior of \(\boldsymbol{\gamma}\) to allow local
+adaption of the shrinkage, using a binomial distribution. The two
+indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
+\(\gamma^{pen}_j\), shares the same probability parameter, \[
+\begin{aligned}
+\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
+\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
+\end{aligned}
+\]
+
+This is to leverage the fact that the probability of selecting the bases
+of a smooth function should be similar, while allowing different penalty
+on the null space and penalty space of the spline function. The hyper
+prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
+function selection level, while that of \(\gamma_{j}^{pen}\) decides the
+smoothness of the spline function at individual predictor level.
+Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
+independently distributed for analytic simplicity. However, this set-up
+could possibility result in a situation that is not theoretically
+possible: the non-linear component is selected, but the linear component
+is not. This can be addressed analytically by forcing to include the
+linear component when non-linear component is included. Another possible
+solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
+\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).
+
+We further specify the parameter \(\theta_j\) follows a beta
+distribution with given shape parameters \(a\) and \(b\), \[
+\theta_j \sim Beta(a, b).
+\] The beta distribution is a conjugate prior for the binomial
+distribution and hence provides some computation convenience.
+Specifically, we focus on a special case of beta distribution, uniform
+(0,1) for simplicity and convenience. To note, when the variable have
+large effects in any of the bases, the parameter \(\theta_j\) will be
+estimated large, which in turn encourages the model to include the rest
+of bases, achieving the local adaption among spline bases. Hereafter, we
+refer the hierarchical spline GAM with the spike-and-slab mixture normal
+prior as the ssGAM-MN.
+
+~
+
+\hypertarget{mixture-double-exponential-prior}{%
+\subsubsection{Mixture Double Exponential
+Prior}\label{mixture-double-exponential-prior}}
+
+One of the critics received by the spike-and-slab mixture normal prior
+is that the tails of a normal distribution diminishes to zero too fast,
+which renders problem when estimating the large effects. In contrast,
+distributions with heavier tails are preferred, for example \(t\)
+distribution with small degree of freedom and double exponential
+distribution. Specifically, double exponential prior is equivalent to
+the LASSO penalty in the frequentist models, providing a sparse
+solution. Meanwhile, the estimation algorithm of LASSO penalty are
+extremely efficient, greatly reducing the computation needs. Recent
+developments of spike-and-slab priors include spike-and-slab LASSO
+prior, i.e.~spike-and-slab mixture double exponential prior. We extend
+the spike-and-slab LASSO prior to high-dimensional GAMs, providing
+computationally efficiency and sparse solution to our proposed ssGAM-MN.
+
+The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we
+replace the normal distribution with double exponential distribution.
+The prior distributions of the coefficients \(\beta^0_j, \beta^{pen}_j\)
+for the \(j\) predictor can be formulated as \[
+\begin{aligned}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
+\end{aligned}
+\] The hyper-priors for \(\gamma^0_j, \gamma^{pen}_j, \theta_j\) are
+kept the same as specified in ssGAM-MN.
+
+\hfill\break
+
+\hypertarget{alforithm-for-fitting-ssgams}{%
+\subsection{Alforithm for fitting
+ssGAMs}\label{alforithm-for-fitting-ssgams}}
+
+Parsimonious computation is always encouraged in high-dimension data
+analysis. Normally, Bayesian methodology loses its advantages over
+Frequentest penalized model because of the computation cost. Previous
+Bayesian spline models heavily relies on the MCMC algorithm to establish
+posterior distribution of parameters. One of the exception is Bai (in
+progress), who a grouped spike-and-slab LASSO prior \citep{Rockova2018}
+to address the grouping structure of spline bases. He focused on the MAP
+estimator of the parameters, but failed to address the uncertainty
+measure of the estimates. The main difference between Bai's work and the
+proposed models are two-folded: 1) the proposed models are feasible to
+answer the question if the non-linear effects are necessary for the
+predictors, while Bai's work can't due to the ``all-in-all-out''
+approach; 2) the proposed models can provide uncertainly measures in
+addition to the point estimates of the coefficients, while Bai's work
+can't not due to the model fitting algorithm.
+
+In the follow section, we introduced two expectation-maximization (EM)
+based algorithms for fast computing, EM-IRLS and EM-coordinate Descent
+algorithm. Instead of simulating the posterior distribution via MCMC, we
+iteratively estimate the MAP. Compared to the EM-IRLS algorithm, the
+EM-coordinate descent provides an expedited model fitting process at the
+cost of model uncertainty.
+
+\hfill\break
+
+\hypertarget{em-algorithms}{%
+\subsubsection{EM algorithms}\label{em-algorithms}}
+
+EM algorithm is an iterative algorithm to find the maximum likelihood
+estimates or the Bayesian counterpart MAP estimates. It is commonly used
+when some necessary data to establish the likelihood function are
+missing. Instead, the algorithm maximizes the expectation, in respect to
+the ``missing'' data, of the likelihood function.
+
+The recursive algorithm consists of two steps:
+
+\begin{itemize}
+\tightlist
+\item
+  E-step: to calculate the expectation of the likelihood function
+  conditioning on some ``missing'' data
+\item
+  M-step: to maximize the spectated likelihood function to calculate the
+  parameter of interest
+\end{itemize}
+
+For ssGAM, we define the parameters of interest as
+\(\Theta = {\boldsymbol{\beta}, \boldsymbol{\theta}, \phi}\) and
+consider the latent binary indicators \(\boldsymbol{\gamma}\) as
+nuisance parameters of the model. Our objective is to find the
+parameters \(\Theta\) that maximize the posterior density function,or
+equivalently, the logarithm of the density function, \[
+\begin{aligned}
+& \text{argmax}_{\Theta}
+\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
+&= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{aligned}
+\]\\
+
+We use the EM algorithm to find the MAP estimate of \(\Theta\). Since
+the latent binary indicators \(\boldsymbol{\gamma}\) are not of our
+primary interest, we treat them as as the ``missing data'' in the EM
+algorithms. Naturally In the E-step, we calculate the expectation of
+posterior density function of
+\(\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
+respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
+the values from previous iteration \(\Theta^{t-1}\), \[
+E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
+\] Hereafter, we use the short hand notation
+\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{t-1}}(\cdot)\). In the
+M-step, we find the \(\Theta^{t}\) that maximize
+\(E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\).
+To note here, the log-posterior of ssGAM (up to additive constants) can
+be written as a two-part equation
+
+\[ \log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\beta, \phi) + Q_2 (\gamma,\theta),\]
+Where
+\[ Q_1(\boldsymbol{\beta}, \phi) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\]
+and \[
+Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .\]
+\(Q_1\) and \(Q_2\) are respectively, the log-likelihood of model
+coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\),
+and latent indicators \(\boldsymbol{\gamma}, \theta\). Conditioning on
+\(\gamma_{jk}\), \(Q_1\) and \(Q_2\) are independent and can be
+maximized separately for \(\boldsymbol{\beta}, \phi\) and
+\(\boldsymbol{\theta}\).
+
+The E- and M- steps are iterated until the algorithm converge. Depending
+on the choice of the mixture priors, either mixture normal or mixture
+double exponential, the maximization in the M-step can take different
+algorithms, IRLS for uncertainty measures and coordinate descent for
+faster computation.
+
+\hfill\break
+
+\hypertarget{em-irls}{%
+\subsubsection{EM-IRLS}\label{em-irls}}
+
+The main idea of the EM-IRLS algorithm is to use the iterative
+re-weighted least square algorithm to find the estimate
+\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). Recall the prior
+set-up of mixture normal prior, we have the prior densities of
+\(\boldsymbol{\beta}\) conditioning on \(\boldsymbol{\gamma}\)
+
+\[
+\begin{aligned}
+p(\boldsymbol{\beta }| \boldsymbol{\gamma}) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],%\\
+% p(\bs\gamma | \bs \theta) & = \prod\limits_{j=1}^{p}\prod\limits_{k=1}^{K_j} \theta_j^{\gamma_{jk}} (1-\theta_j)^{1-\gamma_{jk}}.
+\end{aligned}
+\] and \[
+\begin{aligned}
+E(Q_1) &= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]%\\
+%E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{aligned}
+\] where \(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
+\(S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1\). To
+calculate two unknown quantities \(E({S_j^0}^{-1})\) and
+\(E(S^{-1}_j)\), the posterior probability
+\(p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
+\(p_{j} \equiv p(\gamma^{pen}_{j}=1|\Theta^{t-1})\) are necessary, which
+can be derived via Bayes' theorem. The calculation for \(p_j\) is
+slightly different from that of \(p^0_j\), as \(p_j\) depends on the
+value of \(\beta_{jk}\) for \(k=1, \dots, K_j\) and \(p^0_j\) only
+depends on \(\beta_j^0\). The calculation follows the equations below.
+
+\[
+\begin{aligned}
+p_{j}^0 &= \frac{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + Pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
+p_{j} &= \frac{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + Pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+\end{aligned}
+\]
+
+and
+\(Pr(\gamma_{j}^{0} = 1|\theta_j) =Pr(\gamma_{j}^{pen} = 1|\theta_j) = \theta_j\),
+\(Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^{pen} = 0|\theta_j) = 1-\theta_j\),
+\(f(\beta|\gamma=1, s_1) = \text{Normal}(\beta|0 , s_1)\),
+\(f(\beta|\gamma=0, s_0) = \text{Normal}(\beta|0 , s_0)\). It is trivial
+to show \[
+\begin{aligned}
+&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{pen}_{j})  = p_{j}\\
+&E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
+\end{aligned}
+\]
+
+With \(E({S^0}^{-1}_{j})\) and \(E(S^{-1}_{j})\) known, \(E(Q_1)\) can
+be also viewed as a \(l_2\) penalized likelihood function:
+\(E({S^0}^{-1}_{j})\) controls the shrinkage of the linear effect,
+informing if the predictor should be included in the model;
+\(E(S^{-1}_{j})\) controls the smoothness of the spline function.
+\(l_2\) penalized likelihood function can be maximize using the IWLS
+algorithm. The likelihood function can be approximate by a weighted
+normal likelihood: \[
+  p(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
+the pseudo-response and pseudo-weight, respectively.The pseudo data and
+the pseudo-weight are calculated by \[
+    \begin{aligned}
+    z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+    \end{aligned}
+    \] where \(\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})\) is
+the estimated linear predictor using current estimated value of
+\(\boldsymbol{\beta}\), \(L^{'}\) and \(L^{''}\) are the first
+derivative and second derivative of the log-likelihood function. The
+\(l_2\) penalty, \$
+\sum\limits\_\{j=1\}\^{}p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\$
+can be seems as a multivariate normal distribution of mean
+\(\textbf{0}\) and variance-covariance matrix
+\(\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})\),
+conditioning on the value of \(E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})\).
+An augmentation step combines the normal approximation and multivariate
+normal for \(\boldsymbol{\beta}\), similarly to Ridge regression model
+fitting. The estimates of \(\boldsymbol{\beta}\) and \(\phi\) can be
+quickly updated in the IRLS algorithm.
+
+The remaining parameters of interest \(\boldsymbol{\theta}\) can be
+updated by maximizing \(E(Q_2)\). As the beta distribution is a
+conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
+be easily updated with a closed form equation: \[
+\theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
+\]
+
+Totally, the framework of the proposed EM IRLS algorithm was summarized
+as follows:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
+\def\labelenumi{\arabic{enumi})}
 \item
-  In \texttt{authormark} use \emph{et al.} if there are three or more
-  authors.
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
 \item
-  Note the use of \texttt{num} to link names and addresses.
-\item
-  For submitting a double-spaced manuscript, add \texttt{doublespace} as
-  an option to a \texttt{classoption} line in the YAML header:
-  \texttt{classoption:\ doublespace}.
-\item
-  Keywords are separated by semicolons.
+  Iterate over the E-step and M-step until convergence
 \end{enumerate}
 
-\hypertarget{the-body-of-the-article}{%
-\section{The Body of the Article}\label{the-body-of-the-article}}
+E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
+\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
+\(\Theta^{(t-1)}\) from previous iteration
 
-\hypertarget{mathematics}{%
-\subsection{Mathematics}\label{mathematics}}
+M-step:
 
-Use mathematics in Rmarkdown as usual.
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\item
+  Update \(\boldsymbol{\beta}^{(t)}\) using the IRLS algorithm
+\item
+  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
+\end{enumerate}
 
-\hypertarget{figures-and-tables}{%
-\subsection{Figures and Tables}\label{figures-and-tables}}
+We assess convergence by the criterion:
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
+\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
+the \(t\)th iteration, and \(\epsilon\) is a small value (say
+\(10^{-5}\)).
 
-Figures are supported from R code:
+\hfill\break
+
+\hypertarget{em-coordinate-descent}{%
+\subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
+
+When the prior distribution is set to mixture double exponential,
+coordinate descent algorithm can be used to estimate
+\(\boldsymbol{b}eta\), replacing IRLS, in the M-step. Coordinate descent
+is an optimization algorithm that offers extreme computational
+advantage, and famous for its application in optimizing the \(l_1\)
+penalized likelihood function.
+
+Recall, the density function of spike-and-slab mixture double
+exponential prior can be written as \[
+f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}).
+\]
+
+This can simplify our posterior likelihood function to \[
+E(Q_1) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right]
+,
+\] where the approximation of \(E({S^0_j}^{-1}\) and \(E(S^{-1}_{j})\)
+remains the same as in the EM-IRLS algorithm. The formulation of
+\(E(Q_1)\) can be seen as a \(l_1\) penalized likelihood function, and
+optimized via coordinate descent algorithm. The estimates can be used to
+further estimates \(\boldsymbol{\theta}\) in the M-step, following the
+same equation in the EM-IRLS.
+
+Totally, the framework of the proposed EM-coordinate descent algorithm
+was summarized as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
+\item
+  Iterate over the E-step and M-step until convergence
+\end{enumerate}
+
+E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
+\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
+\(\Theta^{(t-1)}\) from previous iteration
+
+M-step:
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\item
+  Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
+  algorithm
+\item
+  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
+\end{enumerate}
+
+We assess convergence by the criterion:
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
+\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
+the \(t\)th iteration, and \(\epsilon\) is a small value (say
+\(10^{-5}\)).
+
+~
+
+\hypertarget{selecting-optimal-scale-values}{%
+\subsection{Selecting Optimal Scale
+Values}\label{selecting-optimal-scale-values}}
+
+Our proposed prior, spike and slab spline prior, requires two preset
+scale parameters (\(s_0\), \(s_1\)). Hence, we would construct a two
+dimensional grid, consists of different pairs of (\(s_0\), \(s_1\))
+value. However, previous research on the topic\footnote{TODO: add
+  citation} suggested the value of slab scale \(s_1\) have less impact
+on the final model and is recommended to be set as a generally large
+value of \(s_1 = 1\) that provides no or weak shrinkage. More attention
+is focused on examining different values of spike scale \(s_0\). Instead
+of the 2-D grid, We consider a sequence of \(L\) decreasing values
+\(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
+spike scale s\_0 tends to include more non-zero coefficients in the
+model. An measure of preference, e.g.~area under the curve (AUC), mean
+squared error, can be used to facilitate the selection of a final model.
+The procedure is similar to the lasso implemented in the widely used R
+package \texttt{glmnet}, which quickly fits the lasso model over a list
+of values of \(\lambda\) covering the entire range, giving a sequence of
+models for users to choose from.
+
+\hypertarget{simulation}{%
+\section{Simulation}\label{simulation}}
+
+\hypertarget{real-data-analysis}{%
+\section{Real Data Analysis}\label{real-data-analysis}}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 \centering
-\includegraphics{SS_GAM_files/figure-latex/plot-ref-1.pdf}
+\includegraphics{main_files/figure-latex/plot-ref-1.pdf}
 \caption{Fancy Caption\label{fig:plot}}
@@
 
-Referenced via \ref{tab:table}. You can also use the YAML option
-\texttt{header-includes} to includes custom \LaTeX packages for tables
-(keep in mind that \texttt{pandoc} uses \texttt{longtables} by default,
-and it is hardcoded; some things may require including the package
-\texttt{longtable}). E.g., using \texttt{ctable}:
-
-\begin{verbatim}
-header-includes:
-- \usepackage{ctable}
-\end{verbatim}
-
-Then, just write straight-up \LaTeX code and reference is as usual
-(\texttt{\textbackslash{}ref\{tab:ctable\}}):
-
-\begin{verbatim}
-\ctable[cap = {Short caption},
-        caption = {A long, long, long, long, long caption for this table.},
-        label={tab:ctable},]
-        {cc}
-        {
-        \tnote[$\ast$]{Footnote 1}
-        \tnote[$\dagger$]{Other footnote}
-        \tnote[b]{Mistakes are possible.}
-        }{
-        \FL
-        COL 1\tmark[a] & COL 2\tmark[$\ast$]
-        \ML
-        6.92\tmark[$\dagger$] & 0.09781 \\
-        6.93\tmark[$\dagger$] & 0.09901 \\
-        97 & 2000
-        \LL
-}
-\end{verbatim}
-
-It is also possible to set the \texttt{YAML} option
-\texttt{longtable:\ true} and use markdown tables (or the
-\texttt{knitr::kable} function): \texttt{knitr::kable(head(cars))}
-produces the same table as the \texttt{xtable} example presented before.
-
 \hypertarget{cross-referencing}{%
@@
 
-\hypertarget{double-spacing}{%
-\subsection{Double Spacing}\label{double-spacing}}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-If you need to double space your document for submission please use the
-\texttt{doublespace} option in the header.
-
-\hypertarget{bibliography}{%
-\section{Bibliography}\label{bibliography}}
-
-Link a \texttt{.bib} document via the YAML header, and bibliography will
-be printed at the very end (as usual). The default bibliography style is
-provided by Wiley as in \texttt{WileyNJD-AMA.bst}, do not delete that
-file.
-
-Use the Rmarkdown equivalent of the \LaTeX citation system using
-\texttt{{[}@\textless{}name\textgreater{}{]}}. Example:
-\citep{Taylor1937}, \citep{Knupp1999, Kamm2000}.
-
-To include all citation from the \texttt{.bib} file, add
-\texttt{\textbackslash{}nocite\{*\}} before the end of the document.
-
-\hypertarget{further-information}{%
-\section{Further information}\label{further-information}}
-
-All \LaTeX enviroments supported by the main template are supported here
-as well; see the \texttt{.tex} sample file
-\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm}{here}
-for more details and example.
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\end{CSLReferences}
 

### 2021-05-13 Manuscript/01-Intro.Rmd
Putting some contents in the introduction section

@@
 # Introduction
+
 * What is the problem
@@
 * Structure of the rest of the paper
+
+Nonlinear effect modeling has long history.
+Early work of GAMs can date back to 1979 when @Cleveland1979 proposed local smoothing for bivariate analysis.
+Many mathematical and statistical models has been proposed since, see xxx[^ TODO: insert an review citation].
+
+Many machine learning algorithms can also model nonlinear effect.
+However, a wide range of machine learning models are confined by the "black box" nature, rendering little help on understanding the etiology of diseases.
+Semi-parametric and non-parametric regression models, are more popular thanks to its interpretability and flexibility.
+Among those, generalized additive model (GAM) proposed in the seminal work of @Hastie1987 gains the most popularity thanks to its flexibility and easiness to incorporate many different smoothing functions.
+The backfitting algorithm estimates each smooth function of an additive model iteratively using partial residuals.
+The backfitting algorithm is compatible with almost any smoothing or modeling technique to be used as the nonparametric components.
+However, it is unclear how to estimate the degree of smoothness when necessary, for example for smoothing splines.
+The computation cost for cross-validating the smooth parameters can be expensive.
+@Wood2017[PP.318-320] provides more discussion on the problem.
+
+Among all the choices of smoothing function, one particular family stood out, splines.
+Splines are piece-wise polynomial functions to interpolate nonlinear signals.
+Most commonly used spline is cubic spline, thanks to its smoothness property.
+[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smooth function.
+However, the specification of knots is not a trivial task, as the the positioning of knots are influential on the performance of the model.
+As an improvement, researchers proposed smoothing splines where as many knots are used and later shrunk using penalized algorithms.
+The smoothing spline concept was first introduced [@OSullivan1986; @OSullivan1988], and later generalized and made famous by @Eilers1996, called P-spline.
+Independently, @Ruppert1999 proposed similar idea.
+Later, Bayesian GAMs were introduced as an alternative to address measurement error in covariates [@Berry2002] or local adaptive smoothing [@Lang2004].
+
+Generalized additive model (GAM) was first proposed by the seminal work: a response variable, $Y$, given a $p$-dimensional vector of covariates $\bm x$ can be expressed as the summation of smooth functions of $\bm x$, written as
+$$
+E(Y|\bm x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+$$
+where $g^{-1}$ is the inverse function of a smooth monotonic link function.
+Hence, GAM solve the model fitting problem for many distributional outcomes, such as Poisson, binomial, gamma and normal models.
+Since the appearance of GAMs, it has been applied in many area of research, such as xxx, xxx .
+
+In many area of research, high dimensional analysis becomes predominant, for example sequencing data analysis, electric health record data analysis.
+Previous analysis heavily relies on linear models.
+Recently, there are more focus putting on modeling the non-linear relationships in high dimensional data analysis.
+Hence, developing more technically matured models is imminent.
+Nevertheless expanding the GAM framework for high-dimensional applications is not a trivial task.
+The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.
+For GAM models, a smoothing function is consists of multiple basis functions.
+The basis functions of a variable is inherently grouped.
+The grouping structure should be properly taken care of for accurate estimation.
+Moreover, shrinkage is previously used solely for sparsity of the model in high dimensional data analysis, or smoothness in GAM models.
+How to organically combine the two together remains as a troubling question.
+
+#### Spike-and-slab prior
+The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior.
+It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
+$$
+ \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
+$$
+
+The most distinct feature of the SS priors family is that it is conditioned on a latent binary variable $\gamma_i$ that indicates whether the covariate $x_i$ is included in the model.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$ and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
+Here we introduce the most popular spike-and-slab prior introduced in @George1993, as  stochastic search variable selection (SSVS).
+Given a regression problem
+$$
+Y = \bm X\beta+\epsilon, 
+$$
+we can assume the prior distribution of the regression parameters $\beta$ depends on a vector of binary variables $\gamma=(\gamma_1, \dots \gamma_p)^\prime$ as follows: given $\gamma_i$, $\beta_i$ are independent distributed as a mixture priors
+$$
+\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
+$$
+The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution.
+Even the conjugate normal prior of $\bm \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bm \gamma$.
+@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bm \gamma$ for variable selection.
+The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms.
+[@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms.
+@George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.
+
+To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
+EM-based algorithms are prevalent used to speed the computation.
+@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
+The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
+The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
+Like in the spike-and slab family, the latent binary variables $\bm \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
+In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
+EMVS is also compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
+
+The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
+The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
+The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
+Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
+Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
+However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
+@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
+
+To the best of our knowledge, there are few GAM models proposed for high dimensional application.
+In common practice, model matrix of spline terms are generated using base functions of preference, and then feed to a penalized model.
+Without proper handle of the grouping nature of bases functions, the estimation of the smooth function is less accurate.
+@Bai2020Spline shows inferior performance of such practice compared to using an ad hoc high-dimensional additive model.
+
+Among the specialized high-dimensional spline methods, @Ravikumar2009 extended the original backfitting algorithm [@Hastie1987] to accommodate the high-dimensional setting.
+The sparsity and smoothness was handled separately with a soft threshold.
+The authors also demonstrated its similarity to Group Lasso.
+@Meier2009 proposed an a sparsity-smoothness penalty,
+
+$$
+J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
+$$
+which can re-formulated as a group lasso problem as well.
+However, the model can not estimate the smoothness of the spline functions correctly, as the smoothness of all spline models are controlled by the lambda_2.
+The inflexible of the model creates problems in prediction.
+
+From the Bayesian perspective, recent works that expands to the high-dimensional settings includes @Scheipl2012 and @Bai2020Spline.
+@Scheipl2012 proposed an new spike-and-slab structure prior, normal-mixture-of-inverse gamma for semi-parametric regression using MCMC.
+The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
+@Bai2020Spline proposed an novel high dimensional Bayesian GAM using the spike-and-slab group lasso prior.
+An fast fitting EM-based algorithm is proposed.
+Asymptotic theory, such as posterior contraction rates and model selection consistency are proved.
+Even though, the algorithm is flexible to cover many non-Gaussian family in a high dimensional setting, the proposed method didn't address uncertainty quantification, i.e.
+calculating credible intervals etc.
+However, most of these Bayesian GAMs still suffer a common problem, computation cost.
+These Bayesian GAMs uses MCMC to estimate the posterior distribution, which would create an intense computation load.
+
+In this dissertation, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model.
+Specifically, we will propose a new Bayesian spike-and-slab prior for the generalized additive models and Cox spline models.
+The contribution of the proposed prior comes in two-fold: 1) it supports simultaneous variable selection and function selection; 2) it provides robust estimation and prediction for the coefficients and outcomes.
+As previously described, simultaneous variable selection and its functional form selection is not a trivial problem.
+Nevertheless, the Bayesian spike-and-slab prior is a perfect solution to the problem, as the latent binary indicator can be used to for variable selection and the spike and slab densities are used to adaptive adjust the smoothness of the spline function.
+Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e.
+either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect.
+This simplifies the process of determining what effect a predictor have on the outcome, without involving hypothesis testing where the degree of freedom for non-linear effect is complicated to estimate due to the shrinkage.
+Moreover, out model inherent the the robust estimation from Bayesian models.
+
+Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
+Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
+The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
+We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
+
+The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
+We will implement the proposed models and algorithms in one of the most commonly used statistical computing environment R.
+Ancillary functions will be provided in assistance to model diagnostic, interpretation and graphing.
+The R implementation of the Bayesian Hierarchical additive model is available via a freely available R package, BHAM at (https://github.com/boyiguo1/BHAM)[https://github.com/boyiguo1/BHAM].
+
+The rest of the paper are arranged as the followings: in Chapter 32 we demonstrate the Bayesian Hierarchical generalized additive model, where we provide the setup of the proposed model and two fast-fitting algorithms.
+In Chapter 3, we Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
+Conclusion and discussions are made in Chapter 5.

### 2021-05-17 Manuscript/01-Intro.Rmd
Edit Introduction

@@
 * The implication of our method
+* Fast
 * Structure of the rest of the paper
@@
 Early work of GAMs can date back to 1979 when @Cleveland1979 proposed local smoothing for bivariate analysis.
-Many mathematical and statistical models has been proposed since, see xxx[^ TODO: insert an review citation].
+Since then, many mathematical and statistical models has been proposed, see xxx^[TODO: insert an review citation].
+With recent advancement of computing technology, many machine learning algorithms that is computationally intensive start to gainmuch popularity in outcome prediction when non-linear effects assumed.
+Such models generally have improved prediction accuracy over classic statistical models, but have a fatal flaw, lack of transparency and interpretability.
+The primary objective of biomedical studies is to understand the etiology of diseases, which requires not only the knowledge if a biomarker is predictive of the disease, but also how the biomarker cause the disease.
+"Black box" models could less likely to be helpful.
 
-Many machine learning algorithms can also model nonlinear effect.
-However, a wide range of machine learning models are confined by the "black box" nature, rendering little help on understanding the etiology of diseases.
-Semi-parametric and non-parametric regression models, are more popular thanks to its interpretability and flexibility.
+As a balance between interpretability and model flexibility, semi-parametric and non-parametric regression models are studied.
 Among those, generalized additive model (GAM) proposed in the seminal work of @Hastie1987 gains the most popularity thanks to its flexibility and easiness to incorporate many different smoothing functions.
-The backfitting algorithm estimates each smooth function of an additive model iteratively using partial residuals.
+A response variable, $Y$, that is assumed to follow some exponential family distribution $EF(\mu.
+\phi)$, can be expressed as the summation of smooth functions of a given $p$-dimensional vector of covariates $\bs x$, written as
+$$
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+$$
+where $g^{-1}$ is the inverse function of a smooth monotonic link function.
+GAM solves the model fitting problem for many distributional outcomes, such as Poisson, binomial, gamma and normal models.
+In the same article, @@Hastie1987 proposed the backfitting algorithm that estimates each smooth function of an additive model iteratively using partial residuals.
 The backfitting algorithm is compatible with almost any smoothing or modeling technique to be used as the nonparametric components.
@@
 
-Among all the choices of smoothing function, one particular family stood out, splines.
+In many area of modern biomedical research, dealing with high-dimensional data, also known as $p >> n$ problem, is common,  for example sequencing data analysis, electric health record data analysis.
+Many statistical models have been proposed for such problem, for example, penalized models and Bayesian hierarchical models.
+However, up to this point, most efforts of methodology development focus on modeling linear effect in high dimensional setting.
+It is still not clear what is the best practice on how to model non-linear effects in high-dimensional setting.
+Similar to expanding classic models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
+The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.
+For GAM models, a smoothing function is consists of multiple basis functions.
+The basis functions of a variable is inherently related.
+The grouping structure should be properly taken care of for accurate estimation.
+Moreover, shrinkage is previously used solely for sparsity of the model in high dimensional regression models, or smoothness in GAMs.
+How to organically combine the two together remains as a troubling question.
+
+<!-- Among all the choices of smoothing function, one particular family stood out, splines.
 Splines are piece-wise polynomial functions to interpolate nonlinear signals.
@@
 Later, Bayesian GAMs were introduced as an alternative to address measurement error in covariates [@Berry2002] or local adaptive smoothing [@Lang2004].
-
-Generalized additive model (GAM) was first proposed by the seminal work: a response variable, $Y$, given a $p$-dimensional vector of covariates $\bm x$ can be expressed as the summation of smooth functions of $\bm x$, written as
-$$
-E(Y|\bm x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
-$$
-where $g^{-1}$ is the inverse function of a smooth monotonic link function.
-Hence, GAM solve the model fitting problem for many distributional outcomes, such as Poisson, binomial, gamma and normal models.
-Since the appearance of GAMs, it has been applied in many area of research, such as xxx, xxx .
-
-In many area of research, high dimensional analysis becomes predominant, for example sequencing data analysis, electric health record data analysis.
-Previous analysis heavily relies on linear models.
-Recently, there are more focus putting on modeling the non-linear relationships in high dimensional data analysis.
-Hence, developing more technically matured models is imminent.
-Nevertheless expanding the GAM framework for high-dimensional applications is not a trivial task.
-The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.
-For GAM models, a smoothing function is consists of multiple basis functions.
-The basis functions of a variable is inherently grouped.
-The grouping structure should be properly taken care of for accurate estimation.
-Moreover, shrinkage is previously used solely for sparsity of the model in high dimensional data analysis, or smoothness in GAM models.
-How to organically combine the two together remains as a troubling question.
+-->
 
@@
 $$
-Y = \bm X\beta+\epsilon, 
+Y = \bs X\beta+\epsilon, 
 $$
@@
 The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution.
-Even the conjugate normal prior of $\bm \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bm \gamma$.
-@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bm \gamma$ for variable selection.
+Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
+@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection.
 The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms.
@@
 The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
-Like in the spike-and slab family, the latent binary variables $\bm \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
+Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
 In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.

### 2021-05-17 Manuscript/02-Method.Rmd
Edit Introduction

@@
 and 
-$$
-\begin{aligned}
-E(Q_1) &= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]%\\
-%E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
-\end{aligned}
-$$
+\begin{align*}
+E(Q_1) &= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^{p} \left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\\
+E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{align*}
 where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1$.
@@
     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
-The $l_2$ penalty, $ \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$.
+The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$.
 An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.

### 2021-05-17 Manuscript/main.Rmd
Edit Introduction

@@
 ---
-title: A demonstration of the \LaTeX class file for Statistics in Medicine with Rmarkdown
+title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
 author:
@@
 articletype: Research article
-received: 2017-01-01
-revised: 2017-02-01
-accepted: 2017-03-01
-abstract: "TODO: Change this accordingly; Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Aenean ut elit odio.
-Donec fermentum tellus neque, vitae fringilla orci pretium vitae.
-Fusce maximus finibus facilisis.
-Donec ut ullamcorper turpis.
-Donec ut porta ipsum.
-Nullam cursus mauris a sapien ornare pulvinar.
-Aenean malesuada molestie erat quis mattis.
-Praesent scelerisque posuere faucibus.
-Praesent nunc nulla, ullamcorper ut ullamcorper sed, molestie ut est.
-Donec consequat libero nisi, non semper velit vulputate et.
-Quisque eleifend tincidunt ligula, bibendum finibus massa cursus eget.
-Curabitur aliquet vehicula quam non pulvinar.
-Aliquam facilisis tortor nec purus finibus, sit amet elementum eros sodales.
-Ut porta porttitor vestibulum.
-Integer molestie, leo ut maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit amet dolor.
-Sed sed tincidunt ex.
-Curabitur imperdiet egestas tellus in iaculis.
-Maecenas ante neque, pretium vel nisl at, lobortis lacinia neque.
-In gravida elit vel volutpat imperdiet.
-Sed ut nulla arcu.
-Proin blandit interdum ex sit amet laoreet.
-Phasellus efficitur, sem hendrerit mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac ipsum.
-Aenean tincidunt nisl sit amet facilisis faucibus.
-Donec odio erat, bibendum eu imperdiet sed, gravida luctus turpis."
-keywords: Class file; \LaTeX; Statist. Med.; Rmarkdown;
+received: 2021-01-01
+revised: 2021-02-01
+accepted: 2021-03-01
+abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs).
+Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions.
+The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting.
+Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model.
+We propose Bayesian hierarchical generalized additive models for function selection and prediction.
+The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution.
+Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation.
+Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO.
+The software implementation of the proposed methods is freely available via the R package BHAM."
+keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
 bibliography: bibfile.bib
@@
 # References
-<div id="refs"></div>
 

### 2021-05-17 Manuscript/main.tex
Edit Introduction

@@
 % Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \begin{itemize}
+\item
+  What is the problem
+
+  \begin{itemize}
   \tightlist
   \item
-  What is the problem
+    what is the GAM
+  \item
+    expanding GAM in to high-dimensional frame work
+  \item
+    Why this is important
+  \item
+    Why this is hard
+  \end{itemize}
 \item
   Previous solutions
+
+  \begin{itemize}
+  \tightlist
+  \item
+    What are the methods
+  \item
+    What are their problems
+  \end{itemize}
+\item
+  Introduce the wiggliness matrix
 \item
@@
 \end{itemize}
+
+Nonlinear effect modeling has long history. Early work of GAMs can date
+back to 1979 when \citet{Cleveland1979} proposed local smoothing for
+bivariate analysis. Many mathematical and statistical models has been
+proposed since, see xxx{[}\^{} TODO: insert an review citation{]}.
+
+Many machine learning algorithms can also model nonlinear effect.
+However, a wide range of machine learning models are confined by the
+``black box'' nature, rendering little help on understanding the
+etiology of diseases. Semi-parametric and non-parametric regression
+models, are more popular thanks to its interpretability and flexibility.
+Among those, generalized additive model (GAM) proposed in the seminal
+work of \citet{Hastie1987} gains the most popularity thanks to its
+flexibility and easiness to incorporate many different smoothing
+functions. The backfitting algorithm estimates each smooth function of
+an additive model iteratively using partial residuals. The backfitting
+algorithm is compatible with almost any smoothing or modeling technique
+to be used as the nonparametric components. However, it is unclear how
+to estimate the degree of smoothness when necessary, for example for
+smoothing splines. The computation cost for cross-validating the smooth
+parameters can be expensive. \citet[PP.318-320]{Wood2017} provides more
+discussion on the problem.
+
+Among all the choices of smoothing function, one particular family stood
+out, splines. Splines are piece-wise polynomial functions to interpolate
+nonlinear signals. Most commonly used spline is cubic spline, thanks to
+its smoothness property. \citep[PP.197]{Wood2017} Early spline research
+focused on regression splines, where number and location of knots are
+pre-determined and ordinary least squares is used to estimate the smooth
+function. However, the specification of knots is not a trivial task, as
+the the positioning of knots are influential on the performance of the
+model. As an improvement, researchers proposed smoothing splines where
+as many knots are used and later shrunk using penalized algorithms. The
+smoothing spline concept was first introduced
+\citep{OSullivan1986, OSullivan1988}, and later generalized and made
+famous by \citet{Eilers1996}, called P-spline. Independently,
+\citet{Ruppert1999} proposed similar idea. Later, Bayesian GAMs were
+introduced as an alternative to address measurement error in covariates
+\citep{Berry2002} or local adaptive smoothing \citep{Lang2004}.
+
+Generalized additive model (GAM) was first proposed by the seminal work:
+a response variable, \(Y\), given a \(p\)-dimensional vector of
+covariates \(\boldsymbol{x}\) can be expressed as the summation of
+smooth functions of \(\boldsymbol{x}\), written as \[
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+\] where \(g^{-1}\) is the inverse function of a smooth monotonic link
+function. Hence, GAM solve the model fitting problem for many
+distributional outcomes, such as Poisson, binomial, gamma and normal
+models. Since the appearance of GAMs, it has been applied in many area
+of research, such as xxx, xxx .
+
+In many area of research, high dimensional analysis becomes predominant,
+for example sequencing data analysis, electric health record data
+analysis. Previous analysis heavily relies on linear models. Recently,
+there are more focus putting on modeling the non-linear relationships in
+high dimensional data analysis. Hence, developing more technically
+matured models is imminent. Nevertheless expanding the GAM framework for
+high-dimensional applications is not a trivial task. The difficulties
+mainly come in two-fold: grouping nature of the basis functions, and the
+balance between spline smoothness and variable sparsity. For GAM models,
+a smoothing function is consists of multiple basis functions. The basis
+functions of a variable is inherently grouped. The grouping structure
+should be properly taken care of for accurate estimation. Moreover,
+shrinkage is previously used solely for sparsity of the model in high
+dimensional data analysis, or smoothness in GAM models. How to
+organically combine the two together remains as a troubling question.
+
+\hypertarget{spike-and-slab-prior}{%
+\paragraph{Spike-and-slab prior}\label{spike-and-slab-prior}}
+
+The term ``spike-and-slab'' was first coined in \citet{Mitchell1988}
+describing the shape of the prior. It is a mixture distribution,
+comprised with a skinny spike density \(f_{spike}(\cdot)\) for weak
+signals and a flat slab density \(f_{slab}(\cdot)\) for large signals,
+mathematically \[
+ \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
+\]
+
+The most distinct feature of the SS priors family is that it is
+conditioned on a latent binary variable \(\gamma_i\) that indicates
+whether the covariate \(x_i\) is included in the model. There are
+various spike-and-slab priors depending on the choice for the spike
+density \(f_{spike}(\cdot)\) and the slab density \(f_{slab}(\cdot)\),
+for example \citet{George1993}; \citet{George1997}; \citet{Chipman1996}
+for grouped variables; \citet{Brown1998} for multivariate outcomes;
+\citet{Ishwaran2005}; \citet{Clyde2004} and reference therein. Here we
+introduce the most popular spike-and-slab prior introduced in
+\citet{George1993}, as stochastic search variable selection (SSVS).
+Given a regression problem \[
+Y = \boldsymbol{X}\beta+\epsilon, 
+\] we can assume the prior distribution of the regression parameters
+\(\beta\) depends on a vector of binary variables
+\(\gamma=(\gamma_1, \dots \gamma_p)^\prime\) as follows: given
+\(\gamma_i\), \(\beta_i\) are independent distributed as a mixture
+priors \[
+\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
+\] The prior of \(\gamma\) is traditionally assumed to be identically
+independent Bernoulli distribution. Even the conjugate normal prior of
+\(\boldsymbol{\beta}\) provides analytic simplicity, there is no closed
+form solution for the posterior density of \(\boldsymbol{\gamma}\).
+\citet{George1993} described a Gibbs sampling algorithm to approximate
+the posterior distribution of \(\boldsymbol{\gamma}\) for variable
+selection. The same authors later compared SSVS with other Bayesian
+variable selection approaches, providing some empirical suggestions on
+the advantages and disadvantages of various formulations and estimation
+algorithms. \citep{George1997} Among the various problem SSVS addresses,
+high-dimensionality is not one of them, due to the intensive computation
+needs of MCMC algorithms. \citet{George1997} suggested MCMC algorithms
+works well for medium size of predictors (p=25), which is not feasible
+for high-dimensional data, where the number of predictors easily exceeds
+100.
+
+To alleviate the computation burden, many methodology development
+studies focus on the maximum a posteriori (MAP) estimate
+\citep{Tipping2004}. EM-based algorithms are prevalent used to speed the
+computation. \citet{Rockova2014a} proposed an EM-based algorithm for a
+spike-and-slab Gaussian mixture prior, EMVS. The coefficient \(\beta_j\)
+independently follows a mixture normal distribution
+\(N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)\) for
+\(0 \leq v_0 < v_1\). The parameter \(\sigma^2\) follows a inverse gamma
+prior, \(IG(v/2, v\lambda /2)\). Like in the spike-and slab family, the
+latent binary variables \(\boldsymbol{\gamma}\) follows a Bernoulli
+prior with a hyper prior \(\theta\) follows a beta distribution, or
+simpler the uniform (0,1). In the E step, the latent variable are
+treated as the missing data, and calculate the posterior mean of
+\(\gamma\); in the M step, a ridge estimator was used to update the
+coefficients, and \(\sigma^2, \theta\) are updated accordingly. As the
+ridge regression creates small but nonzero coefficients, a further step
+is needed to select the variable: EMVS further suggests to threshold the
+inclusion probability for variable selection or generate regularization
+plot with different values of the spike scale parameter. EMVS is also
+compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary
+variable \(\gamma\) via either logistic regression or markov random
+field.
+
+The same authors \citep{Rockova2018b, Rockova2018} further extended the
+continuous spike-and-slab Gaussian prior to the spike-and-slab double
+exponential distribution, as known as spike-and-slab lasso (SSL). The
+authors also lay the theoretical discussion of comparing spike-and-slab
+from the penalty perspective and the Bayesian perspective. The SSL also
+mitigated the problem of EMVS that the weak signals are not shrink to
+zero. Yi and his group incorporated the SSL prior in the EM-IWLS
+algorithm and proposed another EM algorithm, EM-cyclic coordinate
+descent, based on the Lasso penalty SSL exhibit. Such application were
+used in generalized linear models\citep{Tang2017a}, cox models
+\citep{Tang2017} and their grouped
+counterparts\citep{Tang2018, Tang2019}. Both EM based algorithm provide
+deterministic solutions, which becomes a more popular property as
+reproducible research becomes more valued. However, while the EMVS are
+not able to provide inference of the coefficients easily, the EM-IWLS
+can easily extract the variance-covariance matrix due to the IWLS
+algorithm. \citet{Bai2020} provides an up-to-date summary of SSL methods
+and theoretical discussion.
+
+To the best of our knowledge, there are few GAM models proposed for high
+dimensional application. In common practice, model matrix of spline
+terms are generated using base functions of preference, and then feed to
+a penalized model. Without proper handle of the grouping nature of bases
+functions, the estimation of the smooth function is less accurate.
+\citet{Bai2020Spline} shows inferior performance of such practice
+compared to using an ad hoc high-dimensional additive model.
+
+Among the specialized high-dimensional spline methods,
+\citet{Ravikumar2009} extended the original backfitting algorithm
+\citep{Hastie1987} to accommodate the high-dimensional setting. The
+sparsity and smoothness was handled separately with a soft threshold.
+The authors also demonstrated its similarity to Group Lasso.
+\citet{Meier2009} proposed an a sparsity-smoothness penalty,
+
+\[
+J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
+\] which can re-formulated as a group lasso problem as well. However,
+the model can not estimate the smoothness of the spline functions
+correctly, as the smoothness of all spline models are controlled by the
+lambda\_2. The inflexible of the model creates problems in prediction.
+
+From the Bayesian perspective, recent works that expands to the
+high-dimensional settings includes \citet{Scheipl2012} and
+\citet{Bai2020Spline}. \citet{Scheipl2012} proposed an new
+spike-and-slab structure prior, normal-mixture-of-inverse gamma for
+semi-parametric regression using MCMC. The normal-mixture-of-inverse
+gamma prior is mixture of t-distribution with an additional coefficients
+mixing vector, where the vector follows a mixture of two Gaussian
+distribution. \citet{Bai2020Spline} proposed an novel high dimensional
+Bayesian GAM using the spike-and-slab group lasso prior. An fast fitting
+EM-based algorithm is proposed. Asymptotic theory, such as posterior
+contraction rates and model selection consistency are proved. Even
+though, the algorithm is flexible to cover many non-Gaussian family in a
+high dimensional setting, the proposed method didn't address uncertainty
+quantification, i.e.~calculating credible intervals etc. However, most
+of these Bayesian GAMs still suffer a common problem, computation cost.
+These Bayesian GAMs uses MCMC to estimate the posterior distribution,
+which would create an intense computation load.
+
+In this dissertation, we aim to address the challenges of non-linear
+signal modeling in high-dimensional application using Bayesian
+Hierarchical model. Specifically, we will propose a new Bayesian
+spike-and-slab prior for the generalized additive models and Cox spline
+models. The contribution of the proposed prior comes in two-fold: 1) it
+supports simultaneous variable selection and function selection; 2) it
+provides robust estimation and prediction for the coefficients and
+outcomes. As previously described, simultaneous variable selection and
+its functional form selection is not a trivial problem. Nevertheless,
+the Bayesian spike-and-slab prior is a perfect solution to the problem,
+as the latent binary indicator can be used to for variable selection and
+the spike and slab densities are used to adaptive adjust the smoothness
+of the spline function. Moreover, compared to currently available
+Bayesian spline models which mostly conduct a `all-in-all-out' strategy,
+i.e.~either including a non-linear form of a variable or excluding the
+variable completely, we offer a more flexible solution, rending one of
+three possibilities: no effect of a predictor, only linear effect of a
+predictor, or non-linear effect. This simplifies the process of
+determining what effect a predictor have on the outcome, without
+involving hypothesis testing where the degree of freedom for non-linear
+effect is complicated to estimate due to the shrinkage. Moreover, out
+model inherent the the robust estimation from Bayesian models.
+
+Given the proposed model specification, the estimation of the parameters
+can be easily achieved with MCMC algorithm. Nevertheless, considering
+the computational burden MCMC algorithms bear especially in
+high-dimensional analysis, we propose parsimonious solutions, stemming
+from previous EM-IWLS \citep{Rockova2014a}and EM-coordinate descent
+algorithms\citep{Tang2018, Tang2019}. The computational friendliness
+allows for fitting a model with different sets of initial values,
+mimicking the posterior sampling in the MCMC algorithms. We are hopeful
+that this could improve the sacrificed robustness of the estimation
+process due to the MAP estimation used by the EM algorithms in
+comparison to the model averaging approach.
+
+The third aim of the dissertation work will focus on the software
+implementation of the proposed models and algorithms. We will implement
+the proposed models and algorithms in one of the most commonly used
+statistical computing environment R. Ancillary functions will be
+provided in assistance to model diagnostic, interpretation and graphing.
+The R implementation of the Bayesian Hierarchical additive model is
+available via a freely available R package, BHAM at
+(\url{https://github.com/boyiguo1/BHAM}){[}\url{https://github.com/boyiguo1/BHAM}{]}.
+
+The rest of the paper are arranged as the followings: in Chapter 32 we
+demonstrate the Bayesian Hierarchical generalized additive model, where
+we provide the setup of the proposed model and two fast-fitting
+algorithms. In Chapter 3, we Application to real-world datasets
+demonstrates the usefulness of the proposed methodologies in Chpater 4.
+Conclusion and discussions are made in Chapter 5.
 
@@
 \end{aligned}
-\] and \[
-\begin{aligned}
-E(Q_1) &= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]%\\
-%E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
-\end{aligned}
-\] where \(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
+\] and \begin{align*}
+E(Q_1) &= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^{p} \left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\\
+E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{align*} where
+\(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
 \(S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1\). To
@@
 derivative and second derivative of the log-likelihood function. The
-\(l_2\) penalty, \$
-\sum\limits\_\{j=1\}\^{}p\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\$
+\(l_2\) penalty,
+\(\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\)
 can be seems as a multivariate normal distribution of mean
@@
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\end{CSLReferences}
-
 \bibliography{bibfile.bib}

### 2021-05-17 Simulation/Code/Main.R
Set up simulation folder

@@ -1,0 +1,243 @@
+args=(commandArgs(TRUE))
+
+# print(args)
+
+if(length(args)==0){
+  print("No arguments supplied.")
+}else{
+  for(i in 1:length(args)){
+    eval(parse(text=args[[i]]))
+  }
+}
+
+# Arguments
+# n_train <- 500
+# n_test <- 1000
+# p <- 50
+# it <- 10
+
+# Required Library
+library(mgcv)
+library(splines)
+library(BhGLM)
+library(tidyverse)
+library(rlang)
+library(cosso)
+# library(HRW)
+
+source("~/bgam_nosharing/Code/helper_func.R")
+
+# Using Array ID as seed ID
+it <- Sys.getenv('SLURM_ARRAY_TASK_ID') %>% as.numeric
+# it <- 1
+
+set.seed(it)
+
+
+# Train Data
+tmp <- sim_Bai_logistic(n_train, p)
+dat <- tmp$dat %>% data.frame
+theta <- tmp$theta
+
+# Test Data
+test_tmp <- sim_Bai_logistic(n_test, p)
+test_dat <- test_tmp$dat %>% data.frame
+test_theta <- tmp$theta
+
+
+mgcv_df <- data.frame(
+  Var = setdiff(names(dat), "y"),
+  Func = "s",
+  Args ="bs='cr', k=27"
+)
+
+
+#### mgcv_mdl ####
+
+mgcv_start <- Sys.time()
+mgcv_mdl <- tryCatch({
+  gam(create_HD_formula(y~1, spl_df = mgcv_df), 
+      data=dat %>% data.frame(), family=binomial)
+},
+error = function(err) {
+  mgcv_mdl <- NULL
+  return(NULL)
+}
+)
+mgcv_end <- Sys.time()
+
+mgcv_time <- mgcv_end - mgcv_start
+
+
+mgcv_sum <- NULL
+mgcv_train <- data.frame(deviance = NA, auc = NA, mse = NA, mae = NA, misclassification = NA)
+mgcv_test <- data.frame(deviance = NA, auc = NA, mse = NA, mae = NA, misclassification = NA)
+
+
+if(!is.null(mgcv_mdl)){
+  mgcv_sum <- summary(mgcv_mdl)$s.table[,"p-value"]
+  
+  mgcv_train <- measure.glm(mgcv_mdl$y, mgcv_mdl$fitted.values, family = binomial())
+  
+  
[…]

### 2021-05-18 Manuscript/main.tex
Add Flowchart to the table

@@
 
-\received{2017-01-01}
+\received{2021-01-01}
 
-\revised{2017-02-01}
+\revised{2021-02-01}
 
-\accepted{2017-03-01}
+\accepted{2021-03-01}
 
@@
 
-\title{A demonstration of the \LaTeX class file for Statistics in
-Medicine with Rmarkdown}
+\title{Spike-and-Slab Generalized Additive Models and Fast Algorithms
+for High-Dimensional Data}
 
@@
 
-\abstract{TODO: Change this accordingly; Lorem ipsum dolor sit amet,
-consectetur adipiscing elit. Aenean ut elit odio. Donec fermentum tellus
-neque, vitae fringilla orci pretium vitae. Fusce maximus finibus
-facilisis. Donec ut ullamcorper turpis. Donec ut porta ipsum. Nullam
-cursus mauris a sapien ornare pulvinar. Aenean malesuada molestie erat
-quis mattis. Praesent scelerisque posuere faucibus. Praesent nunc nulla,
-ullamcorper ut ullamcorper sed, molestie ut est. Donec consequat libero
-nisi, non semper velit vulputate et. Quisque eleifend tincidunt ligula,
-bibendum finibus massa cursus eget. Curabitur aliquet vehicula quam non
-pulvinar. Aliquam facilisis tortor nec purus finibus, sit amet elementum
-eros sodales. Ut porta porttitor vestibulum. Integer molestie, leo ut
-maximus aliquam, velit dui iaculis nibh, eget hendrerit purus risus sit
-amet dolor. Sed sed tincidunt ex. Curabitur imperdiet egestas tellus in
-iaculis. Maecenas ante neque, pretium vel nisl at, lobortis lacinia
-neque. In gravida elit vel volutpat imperdiet. Sed ut nulla arcu. Proin
-blandit interdum ex sit amet laoreet. Phasellus efficitur, sem hendrerit
-mattis dapibus, nunc tellus ornare nisi, nec eleifend enim nibh ac
-ipsum. Aenean tincidunt nisl sit amet facilisis faucibus. Donec odio
-erat, bibendum eu imperdiet sed, gravida luctus turpis.}
+\abstract{Spike-and-slab priors have been extensively studied for
+variable selection, and recently been extended to function selection in
+generalized additive models (GAMs). Most of the previous works on
+fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC)
+algorithms to identify the promising subsets of functions. The
+computational burden of MCMC makes these models less feasible,
+especially in high-dimensional setting. Besides, most of the current
+spike-and-slab GAM consider an ``all-in-all-out'' approach for function
+selection, rendering less flexibility to answer if non-linear effects
+are necessary in the model. We propose Bayesian hierarchical generalized
+additive models for function selection and prediction. The proposed
+models employ spike-and-slab priors, mixture normal and mixture double
+exponential, for smooth and sparse solution. Two EM based algorithms,
+EM-IWLS and EM-Coordinate Descent, are developed for accelerated
+computation. Simulation studies and real data analysis demonstrated
+improved performance against the state-of-art models, mgcv and COSSO.
+The software implementation of the proposed methods is freely available
+via the R package BHAM.}
 
-\keywords{Class file; \LaTeX; Statist. Med.; Rmarkdown;}
+\keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
+Additive Models; EM-IWLS; EM-Coordinate Decent}
 
@@
 \item
+  Fast
+\item
   Structure of the rest of the paper
@@
 back to 1979 when \citet{Cleveland1979} proposed local smoothing for
-bivariate analysis. Many mathematical and statistical models has been
-proposed since, see xxx{[}\^{} TODO: insert an review citation{]}.
+bivariate analysis. Since then, many mathematical and statistical models
+has been proposed, see xxx\footnote{TODO: insert an review citation}.
+With recent advancement of computing technology, many machine learning
+algorithms that is computationally intensive start to gainmuch
+popularity in outcome prediction when non-linear effects assumed. Such
+models generally have improved prediction accuracy over classic
+statistical models, but have a fatal flaw, lack of transparency and
+interpretability. The primary objective of biomedical studies is to
+understand the etiology of diseases, which requires not only the
+knowledge if a biomarker is predictive of the disease, but also how the
+biomarker cause the disease. ``Black box'' models could less likely to
+be helpful.
 
-Many machine learning algorithms can also model nonlinear effect.
-However, a wide range of machine learning models are confined by the
-``black box'' nature, rendering little help on understanding the
-etiology of diseases. Semi-parametric and non-parametric regression
-models, are more popular thanks to its interpretability and flexibility.
-Among those, generalized additive model (GAM) proposed in the seminal
-work of \citet{Hastie1987} gains the most popularity thanks to its
-flexibility and easiness to incorporate many different smoothing
-functions. The backfitting algorithm estimates each smooth function of
-an additive model iteratively using partial residuals. The backfitting
-algorithm is compatible with almost any smoothing or modeling technique
-to be used as the nonparametric components. However, it is unclear how
-to estimate the degree of smoothness when necessary, for example for
-smoothing splines. The computation cost for cross-validating the smooth
-parameters can be expensive. \citet[PP.318-320]{Wood2017} provides more
-discussion on the problem.
-
-Among all the choices of smoothing function, one particular family stood
-out, splines. Splines are piece-wise polynomial functions to interpolate
-nonlinear signals. Most commonly used spline is cubic spline, thanks to
-its smoothness property. \citep[PP.197]{Wood2017} Early spline research
-focused on regression splines, where number and location of knots are
-pre-determined and ordinary least squares is used to estimate the smooth
-function. However, the specification of knots is not a trivial task, as
-the the positioning of knots are influential on the performance of the
-model. As an improvement, researchers proposed smoothing splines where
-as many knots are used and later shrunk using penalized algorithms. The
-smoothing spline concept was first introduced
-\citep{OSullivan1986, OSullivan1988}, and later generalized and made
-famous by \citet{Eilers1996}, called P-spline. Independently,
-\citet{Ruppert1999} proposed similar idea. Later, Bayesian GAMs were
-introduced as an alternative to address measurement error in covariates
-\citep{Berry2002} or local adaptive smoothing \citep{Lang2004}.
-
-Generalized additive model (GAM) was first proposed by the seminal work:
-a response variable, \(Y\), given a \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\) can be expressed as the summation of
-smooth functions of \(\boldsymbol{x}\), written as \[
+As a balance between interpretability and model flexibility,
+semi-parametric and non-parametric regression models are studied. Among
+those, generalized additive model (GAM) proposed in the seminal work of
+\citet{Hastie1987} gains the most popularity thanks to its flexibility
+and easiness to incorporate many different smoothing functions. A
+response variable, \(Y\), that is assumed to follow some exponential
+family distribution \(EF(\mu. \phi)\), can be expressed as the summation
+of smooth functions of a given \(p\)-dimensional vector of covariates
+\(\boldsymbol{x}\), written as \[
 E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
 \] where \(g^{-1}\) is the inverse function of a smooth monotonic link
-function. Hence, GAM solve the model fitting problem for many
-distributional outcomes, such as Poisson, binomial, gamma and normal
-models. Since the appearance of GAMs, it has been applied in many area
-of research, such as xxx, xxx .
+function. GAM solves the model fitting problem for many distributional
+outcomes, such as Poisson, binomial, gamma and normal models. In the
+same article, @\citet{Hastie1987} proposed the backfitting algorithm
+that estimates each smooth function of an additive model iteratively
+using partial residuals. The backfitting algorithm is compatible with
+almost any smoothing or modeling technique to be used as the
+nonparametric components. However, it is unclear how to estimate the
+degree of smoothness when necessary, for example for smoothing splines.
+The computation cost for cross-validating the smooth parameters can be
+expensive. \citet[PP.318-320]{Wood2017} provides more discussion on the
+problem.
 
-In many area of research, high dimensional analysis becomes predominant,
-for example sequencing data analysis, electric health record data
-analysis. Previous analysis heavily relies on linear models. Recently,
-there are more focus putting on modeling the non-linear relationships in
-high dimensional data analysis. Hence, developing more technically
-matured models is imminent. Nevertheless expanding the GAM framework for
-high-dimensional applications is not a trivial task. The difficulties
-mainly come in two-fold: grouping nature of the basis functions, and the
-balance between spline smoothness and variable sparsity. For GAM models,
-a smoothing function is consists of multiple basis functions. The basis
-functions of a variable is inherently grouped. The grouping structure
-should be properly taken care of for accurate estimation. Moreover,
-shrinkage is previously used solely for sparsity of the model in high
-dimensional data analysis, or smoothness in GAM models. How to
-organically combine the two together remains as a troubling question.
+In many area of modern biomedical research, dealing with
+high-dimensional data, also known as \(p >> n\) problem, is common, for
+example sequencing data analysis, electric health record data analysis.
+Many statistical models have been proposed for such problem, for
+example, penalized models and Bayesian hierarchical models. However, up
+to this point, most efforts of methodology development focus on modeling
+linear effect in high dimensional setting. It is still not clear what is
+the best practice on how to model non-linear effects in high-dimensional
+setting. Similar to expanding classic models to high-dimensional models,
+expanding the GAM framework for high-dimensional applications is not a
+trivial task. The difficulties mainly come in two-fold: grouping nature
+of the basis functions, and the balance between spline smoothness and
+variable sparsity. For GAM models, a smoothing function is consists of
+multiple basis functions. The basis functions of a variable is
+inherently related. The grouping structure should be properly taken care
+of for accurate estimation. Moreover, shrinkage is previously used
+solely for sparsity of the model in high dimensional regression models,
+or smoothness in GAMs. How to organically combine the two together
+remains as a troubling question.
 
-\hypertarget{spike-and-slab-prior}{%
-\paragraph{Spike-and-slab prior}\label{spike-and-slab-prior}}
-
-The term ``spike-and-slab'' was first coined in \citet{Mitchell1988}
-describing the shape of the prior. It is a mixture distribution,
-comprised with a skinny spike density \(f_{spike}(\cdot)\) for weak
-signals and a flat slab density \(f_{slab}(\cdot)\) for large signals,
-mathematically \[
+For high-dimensional data analysis, a common Bayesian modeling approach
+is to impose a particular family of priors, the spike-and-slab (SS)
+priors on the coeffcient in a regression modelss. The term
+``spike-and-slab'' was first coined in \citet{Mitchell1988} describing
+the shape of the prior. It is a mixture distribution, comprised with a
+skinny spike density \(f_{spike}(\cdot)\) for weak signals and a flat
+slab density \(f_{slab}(\cdot)\) for large signals, mathematically \[
  \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
@@
 
-In this dissertation, we aim to address the challenges of non-linear
-signal modeling in high-dimensional application using Bayesian
-Hierarchical model. Specifically, we will propose a new Bayesian
-spike-and-slab prior for the generalized additive models and Cox spline
-models. The contribution of the proposed prior comes in two-fold: 1) it
-supports simultaneous variable selection and function selection; 2) it
-provides robust estimation and prediction for the coefficients and
-outcomes. As previously described, simultaneous variable selection and
-its functional form selection is not a trivial problem. Nevertheless,
-the Bayesian spike-and-slab prior is a perfect solution to the problem,
-as the latent binary indicator can be used to for variable selection and
-the spike and slab densities are used to adaptive adjust the smoothness
-of the spline function. Moreover, compared to currently available
-Bayesian spline models which mostly conduct a `all-in-all-out' strategy,
-i.e.~either including a non-linear form of a variable or excluding the
-variable completely, we offer a more flexible solution, rending one of
-three possibilities: no effect of a predictor, only linear effect of a
+In this article, we aim to address the challenges of non-linear signal
+modeling in high-dimensional application using Bayesian Hierarchical
+model. Specifically, we will propose a new Bayesian spike-and-slab prior
+for the generalized additive models and Cox spline models. The
+contribution of the proposed prior comes in two-fold: 1) it supports
+simultaneous variable selection and function selection; 2) it provides
+robust estimation and prediction for the coefficients and outcomes. As
+previously described, simultaneous variable selection and its functional
+form selection is not a trivial problem. Nevertheless, the Bayesian
+spike-and-slab prior is a perfect solution to the problem, as the latent
+binary indicator can be used to for variable selection and the spike and
+slab densities are used to adaptive adjust the smoothness of the spline
+function. Moreover, compared to currently available Bayesian spline
+models which mostly conduct a `all-in-all-out' strategy, i.e.~either
+including a non-linear form of a variable or excluding the variable
+completely, we offer a more flexible solution, rending one of three
+possibilities: no effect of a predictor, only linear effect of a
 predictor, or non-linear effect. This simplifies the process of
@@
 prior as the ssGAM-MN.
+
+\includegraphics[width=24.76in]{Fig/SS_prior_flow_chart}
 
@@
 
-\begin{verbatim}
-x = rnorm(10)
-y = rnorm(10)
-plot(x, y)
-\end{verbatim}
-
 \begin{figure}
@@
 Analogously, use Rmarkdown to produce tables as usual:
-
-\begin{verbatim}
-if (!require("xtable")) install.packages("xtable")
-\end{verbatim}
-
-\begin{verbatim}
-## Loading required package: xtable
-\end{verbatim}
-
-\begin{verbatim}
-xt <- xtable(head(cars), caption = "A table", label = "tab:table")
-print(xt, comment = FALSE)
-\end{verbatim}
 

### 2021-05-27 Simulation/tuning/Tuning Study Report.Rmd
Adding Simulation Study of Tuning Parameter

---
title: "Sim_tuning Report"
author: "Boyi Guo"
date: "5/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Computation environment

* Run on UAB Cheaha cluster. Find more description at [https://docs.uabgrid.uab.edu/wiki/Cheaha](https://docs.uabgrid.uab.edu/wiki/Cheaha)
* For the sake of time, the maximum number of iterations are set to 10.

# Restuls

* Poisson outcomes requires extremely more time to finish

* `bmlasso` doesn't not work well in gaussian outcomes. Similarly as we previously observed in netwrok ss studies

* Tuning parameter `s0` matters when the number of outcome grows exponentially.

### 2021-05-27 Simulation/Code/Main.R

@@ -42,7 +42,7 @@ theta <- tmp$theta
 # Test Data
 test_tmp <- sim_Bai_logistic(n_test, p)
 test_dat <- test_tmp$dat %>% data.frame
-test_theta <- tmp$theta
+test_theta <- test_tmp$theta
 
 
 mgcv_df <- data.frame(

### 2021-05-27 Manuscript/01-Intro.Rmd

@@
 
-#### Spike-and-slab prior
+For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss.
 The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior.
@@
 
-In this dissertation, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model.
+In this article, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model.
 Specifically, we will propose a new Bayesian spike-and-slab prior for the generalized additive models and Cox spline models.

### 2021-05-31 Manuscript/SS_GAM.tex
Adding Emory Card Biobank to Real Data Analysis

@@
 
-\hypertarget{the-article-header-information}{%
-\section{The Article Header
-Information}\label{the-article-header-information}}
+\newcommand{\bs}[1]{\boldsymbol{#1}}
 
-YAML header:
-
-\begin{verbatim}
-output:
-  rticles::sim_article:
-    keep_tex: TRUE
-\end{verbatim}
-
-Configure the YAML header including the following elements:
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 \item
-  \texttt{title}: Title
+  What is the problem
+
+  \begin{itemize}
+  \tightlist
   \item
-  \texttt{author}: List of author(s) containing \texttt{name} and
-  \texttt{num}
+    what is the GAM
   \item
-  \texttt{address}: List containing \texttt{num} and \texttt{org} for
-  defining \texttt{author} affiliations
+    expanding GAM in to high-dimensional frame work
   \item
-  \texttt{presentaddress}: Not sure what they mean with this
+    Why this is important
   \item
-  \texttt{corres}: Author and address for correspondence
+    Why this is hard
+  \end{itemize}
 \item
-  \texttt{authormark}: Short author list for header
+  Previous solutions
+
+  \begin{itemize}
+  \tightlist
   \item
-  \texttt{received}, \texttt{revised}, \texttt{accepted}: dates of
-  submission, revision, and acceptance of the manuscript
+    What are the methods
   \item
-  \texttt{abstract}: Limited to 250 words
+    What are their problems
+  \end{itemize}
 \item
-  \texttt{keywords}: Up to 6 keywords
+  Introduce the wiggliness matrix
 \item
-  \texttt{bibliography}: BibTeX \texttt{.bib} file
+  Our method
 \item
-  \texttt{classoption}: options of the \texttt{WileyNJD-v2} class
+  The implication of our method
 \item
-  \texttt{longtable}: set to \texttt{true} to include the
-  \texttt{longtable} package, used by default from \texttt{pandoc} to
-  convert markdown to \LaTeX code
+  Fast
+\item
+  Structure of the rest of the paper
 \end{itemize}
 
-\hypertarget{remarks}{%
-\subsection{Remarks}\label{remarks}}
+Nonlinear effect modeling has long history. Early work of GAMs can date
+back to 1979 when \citet{Cleveland1979} proposed local smoothing for
+bivariate analysis. Since then, many mathematical and statistical models
+has been proposed, see xxx\footnote{TODO: insert an review citation}.
+With recent advancement of computing technology, many machine learning
+algorithms that is computationally intensive start to gainmuch
+popularity in outcome prediction when non-linear effects assumed. Such
+models generally have improved prediction accuracy over classic
+statistical models, but have a fatal flaw, lack of transparency and
+interpretability. The primary objective of biomedical studies is to
+understand the etiology of diseases, which requires not only the
+knowledge if a biomarker is predictive of the disease, but also how the
+biomarker cause the disease. ``Black box'' models could less likely to
+be helpful.
+
+As a balance between interpretability and model flexibility,
+semi-parametric and non-parametric regression models are studied. Among
+those, generalized additive model (GAM) proposed in the seminal work of
+\citet{Hastie1987} gains the most popularity thanks to its flexibility
+and easiness to incorporate many different smoothing functions. A
+response variable, \(Y\), that is assumed to follow some exponential
+family distribution \(EF(\mu. \phi)\), can be expressed as the summation
+of smooth functions of a given \(p\)-dimensional vector of covariates
+\(\boldsymbol{x}\), written as \[
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+\] where \(g^{-1}\) is the inverse function of a smooth monotonic link
+function. GAM solves the model fitting problem for many distributional
+outcomes, such as Poisson, binomial, gamma and normal models. In the
+same article, @\citet{Hastie1987} proposed the backfitting algorithm
+that estimates each smooth function of an additive model iteratively
+using partial residuals. The backfitting algorithm is compatible with
+almost any smoothing or modeling technique to be used as the
+nonparametric components. However, it is unclear how to estimate the
+degree of smoothness when necessary, for example for smoothing splines.
+The computation cost for cross-validating the smooth parameters can be
+expensive. \citet[PP.318-320]{Wood2017} provides more discussion on the
+problem.
+
+In many area of modern biomedical research, dealing with
+high-dimensional data, also known as \(p >> n\) problem, is common, for
+example sequencing data analysis, electric health record data analysis.
+Many statistical models have been proposed for such problem, for
+example, penalized models and Bayesian hierarchical models. However, up
+to this point, most efforts of methodology development focus on modeling
+linear effect in high dimensional setting. It is still not clear what is
+the best practice on how to model non-linear effects in high-dimensional
+setting. Similar to expanding classic models to high-dimensional models,
+expanding the GAM framework for high-dimensional applications is not a
+trivial task. The difficulties mainly come in two-fold: grouping nature
+of the basis functions, and the balance between spline smoothness and
+variable sparsity. For GAM models, a smoothing function is consists of
+multiple basis functions. The basis functions of a variable is
+inherently related. The grouping structure should be properly taken care
+of for accurate estimation. Moreover, shrinkage is previously used
+solely for sparsity of the model in high dimensional regression models,
+or smoothness in GAMs. How to organically combine the two together
+remains as a troubling question.
+
+For high-dimensional data analysis, a common Bayesian modeling approach
+is to impose a particular family of priors, the spike-and-slab (SS)
+priors on the coeffcient in a regression modelss. The term
+``spike-and-slab'' was first coined in \citet{Mitchell1988} describing
+the shape of the prior. It is a mixture distribution, comprised with a
+skinny spike density \(f_{spike}(\cdot)\) for weak signals and a flat
+slab density \(f_{slab}(\cdot)\) for large signals, mathematically \[
+ \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
+\]
+
+The most distinct feature of the SS priors family is that it is
+conditioned on a latent binary variable \(\gamma_i\) that indicates
+whether the covariate \(x_i\) is included in the model. There are
+various spike-and-slab priors depending on the choice for the spike
+density \(f_{spike}(\cdot)\) and the slab density \(f_{slab}(\cdot)\),
+for example \citet{George1993}; \citet{George1997}; \citet{Chipman1996}
+for grouped variables; \citet{Brown1998} for multivariate outcomes;
+\citet{Ishwaran2005}; \citet{Clyde2004} and reference therein. Here we
+introduce the most popular spike-and-slab prior introduced in
+\citet{George1993}, as stochastic search variable selection (SSVS).
+Given a regression problem \[
+Y = \boldsymbol{X}\beta+\epsilon, 
+\] we can assume the prior distribution of the regression parameters
+\(\beta\) depends on a vector of binary variables
+\(\gamma=(\gamma_1, \dots \gamma_p)^\prime\) as follows: given
+\(\gamma_i\), \(\beta_i\) are independent distributed as a mixture
+priors \[
+\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
+\] The prior of \(\gamma\) is traditionally assumed to be identically
+independent Bernoulli distribution. Even the conjugate normal prior of
+\(\boldsymbol{\beta}\) provides analytic simplicity, there is no closed
+form solution for the posterior density of \(\boldsymbol{\gamma}\).
+\citet{George1993} described a Gibbs sampling algorithm to approximate
+the posterior distribution of \(\boldsymbol{\gamma}\) for variable
+selection. The same authors later compared SSVS with other Bayesian
+variable selection approaches, providing some empirical suggestions on
+the advantages and disadvantages of various formulations and estimation
+algorithms. \citep{George1997} Among the various problem SSVS addresses,
+high-dimensionality is not one of them, due to the intensive computation
+needs of MCMC algorithms. \citet{George1997} suggested MCMC algorithms
+works well for medium size of predictors (p=25), which is not feasible
+for high-dimensional data, where the number of predictors easily exceeds
+100.
+
+To alleviate the computation burden, many methodology development
+studies focus on the maximum a posteriori (MAP) estimate
+\citep{Tipping2004}. EM-based algorithms are prevalent used to speed the
+computation. \citet{Rockova2014a} proposed an EM-based algorithm for a
+spike-and-slab Gaussian mixture prior, EMVS. The coefficient \(\beta_j\)
+independently follows a mixture normal distribution
+\(N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)\) for
+\(0 \leq v_0 < v_1\). The parameter \(\sigma^2\) follows a inverse gamma
+prior, \(IG(v/2, v\lambda /2)\). Like in the spike-and slab family, the
+latent binary variables \(\boldsymbol{\gamma}\) follows a Bernoulli
+prior with a hyper prior \(\theta\) follows a beta distribution, or
+simpler the uniform (0,1). In the E step, the latent variable are
+treated as the missing data, and calculate the posterior mean of
+\(\gamma\); in the M step, a ridge estimator was used to update the
+coefficients, and \(\sigma^2, \theta\) are updated accordingly. As the
+ridge regression creates small but nonzero coefficients, a further step
+is needed to select the variable: EMVS further suggests to threshold the
+inclusion probability for variable selection or generate regularization
+plot with different values of the spike scale parameter. EMVS is also
+compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary
+variable \(\gamma\) via either logistic regression or markov random
+field.
+
+The same authors \citep{Rockova2018b, Rockova2018} further extended the
+continuous spike-and-slab Gaussian prior to the spike-and-slab double
+exponential distribution, as known as spike-and-slab lasso (SSL). The
+authors also lay the theoretical discussion of comparing spike-and-slab
+from the penalty perspective and the Bayesian perspective. The SSL also
+mitigated the problem of EMVS that the weak signals are not shrink to
+zero. Yi and his group incorporated the SSL prior in the EM-IWLS
+algorithm and proposed another EM algorithm, EM-cyclic coordinate
+descent, based on the Lasso penalty SSL exhibit. Such application were
+used in generalized linear models\citep{Tang2017a}, cox models
+\citep{Tang2017} and their grouped
+counterparts\citep{Tang2018, Tang2019}. Both EM based algorithm provide
+deterministic solutions, which becomes a more popular property as
+reproducible research becomes more valued. However, while the EMVS are
+not able to provide inference of the coefficients easily, the EM-IWLS
+can easily extract the variance-covariance matrix due to the IWLS
+algorithm. \citet{Bai2020} provides an up-to-date summary of SSL methods
+and theoretical discussion.
+
+To the best of our knowledge, there are few GAM models proposed for high
+dimensional application. In common practice, model matrix of spline
+terms are generated using base functions of preference, and then feed to
+a penalized model. Without proper handle of the grouping nature of bases
+functions, the estimation of the smooth function is less accurate.
+\citet{Bai2020Spline} shows inferior performance of such practice
+compared to using an ad hoc high-dimensional additive model.
+
+Among the specialized high-dimensional spline methods,
+\citet{Ravikumar2009} extended the original backfitting algorithm
+\citep{Hastie1987} to accommodate the high-dimensional setting. The
+sparsity and smoothness was handled separately with a soft threshold.
+The authors also demonstrated its similarity to Group Lasso.
+\citet{Meier2009} proposed an a sparsity-smoothness penalty,
+
+\[
+J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
+\] which can re-formulated as a group lasso problem as well. However,
+the model can not estimate the smoothness of the spline functions
+correctly, as the smoothness of all spline models are controlled by the
+lambda\_2. The inflexible of the model creates problems in prediction.
+
+From the Bayesian perspective, recent works that expands to the
+high-dimensional settings includes \citet{Scheipl2012} and
+\citet{Bai2020Spline}. \citet{Scheipl2012} proposed an new
+spike-and-slab structure prior, normal-mixture-of-inverse gamma for
+semi-parametric regression using MCMC. The normal-mixture-of-inverse
+gamma prior is mixture of t-distribution with an additional coefficients
+mixing vector, where the vector follows a mixture of two Gaussian
+distribution. \citet{Bai2020Spline} proposed an novel high dimensional
+Bayesian GAM using the spike-and-slab group lasso prior. An fast fitting
+EM-based algorithm is proposed. Asymptotic theory, such as posterior
+contraction rates and model selection consistency are proved. Even
+though, the algorithm is flexible to cover many non-Gaussian family in a
+high dimensional setting, the proposed method didn't address uncertainty
+quantification, i.e.~calculating credible intervals etc. However, most
+of these Bayesian GAMs still suffer a common problem, computation cost.
+These Bayesian GAMs uses MCMC to estimate the posterior distribution,
+which would create an intense computation load.
+
+In this article, we aim to address the challenges of non-linear signal
+modeling in high-dimensional application using Bayesian Hierarchical
+model. Specifically, we will propose a new Bayesian spike-and-slab prior
+for the generalized additive models and Cox spline models. The
+contribution of the proposed prior comes in two-fold: 1) it supports
+simultaneous variable selection and function selection; 2) it provides
+robust estimation and prediction for the coefficients and outcomes. As
+previously described, simultaneous variable selection and its functional
+form selection is not a trivial problem. Nevertheless, the Bayesian
+spike-and-slab prior is a perfect solution to the problem, as the latent
+binary indicator can be used to for variable selection and the spike and
+slab densities are used to adaptive adjust the smoothness of the spline
+function. Moreover, compared to currently available Bayesian spline
+models which mostly conduct a `all-in-all-out' strategy, i.e.~either
+including a non-linear form of a variable or excluding the variable
+completely, we offer a more flexible solution, rending one of three
+possibilities: no effect of a predictor, only linear effect of a
+predictor, or non-linear effect. This simplifies the process of
+determining what effect a predictor have on the outcome, without
+involving hypothesis testing where the degree of freedom for non-linear
+effect is complicated to estimate due to the shrinkage. Moreover, out
+model inherent the the robust estimation from Bayesian models.
+
+Given the proposed model specification, the estimation of the parameters
+can be easily achieved with MCMC algorithm. Nevertheless, considering
+the computational burden MCMC algorithms bear especially in
+high-dimensional analysis, we propose parsimonious solutions, stemming
+from previous EM-IWLS \citep{Rockova2014a}and EM-coordinate descent
+algorithms\citep{Tang2018, Tang2019}. The computational friendliness
+allows for fitting a model with different sets of initial values,
+mimicking the posterior sampling in the MCMC algorithms. We are hopeful
+that this could improve the sacrificed robustness of the estimation
+process due to the MAP estimation used by the EM algorithms in
+comparison to the model averaging approach.
+
+The third aim of the dissertation work will focus on the software
+implementation of the proposed models and algorithms. We will implement
+the proposed models and algorithms in one of the most commonly used
+statistical computing environment R. Ancillary functions will be
+provided in assistance to model diagnostic, interpretation and graphing.
+The R implementation of the Bayesian Hierarchical additive model is
+available via a freely available R package, BHAM at
+(\url{https://github.com/boyiguo1/BHAM}){[}\url{https://github.com/boyiguo1/BHAM}{]}.
+
+The rest of the paper are arranged as the followings: in Chapter 32 we
+demonstrate the Bayesian Hierarchical generalized additive model, where
+we provide the setup of the proposed model and two fast-fitting
+algorithms. In Chapter 3, we Application to real-world datasets
+demonstrates the usefulness of the proposed methodologies in Chpater 4.
+Conclusion and discussions are made in Chapter 5.
+
+\hypertarget{spike-and-slab-additive-model}{%
+\section{Spike-and-slab Additive
+Model}\label{spike-and-slab-additive-model}}
+
+Following the notation in the previous section \ref{Notation}, suppose
+we have the model matrix \(\boldsymbol{X}\) derived from spline
+functions of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates
+that are modeled nonparametrically, the total dimension of
+\(\boldsymbol{X}\) is \(n \times \sum\limits_{j=1}^p k_j\). Column-wise
+centering and re-parameterization introduced in the GAM as random effect
+section{[}\^{}5{]} are performed. The re-parameterization has two
+purposes. First of all the re-parameterization reshapes the columns to
+be linearly independent to each other, and hence, independent prior can
+be set to the coefficients. Secondly, the re-parameterization decomposes
+the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
+sub-matrices representing the null space of the spline and penalized
+space of the spline, \[
+\boldsymbol{X}_j = \begin{bmatrix}
+\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
+\end{bmatrix},
+\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
+the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
+denotes the column spaces that need to be shrunk for smoothness of the
+spline. The purpose for this re-parameterization is to isolate the null
+space from the smoothing space, such that answering the question, if
+non-linear relationship is necessary, becomes easy by deciding if the
+corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
+zero. This strategy is widely used, previously mentioned in
+\citet{Scheipl2012} and the references there in.\footnote{Possibly
+  select some to add here.} Hereafter, we denote the coefficients of the
+null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
+penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
+for the \(j\)the covariate is \[
+f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
+\] where the superscript \(^t\) denotes matrix transpose.
+
+Without loss of generality, we assume that the model doesn't include the
+the parametric part of the model \(\boldsymbol{A}\). The coefficients of
+\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
+by local update conditioning on the temporary value of the spline
+coefficients.
+
+~
+
+\hypertarget{spike-and-slab-spline-prior}{%
+\subsection{Spike and slab spline
+prior}\label{spike-and-slab-spline-prior}}
+
+A generalized spline model can be formulated as \[
+y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
+\] The link function of the mean can be expressed in a matrix form \[
+g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
+ = \beta^T X_i,
+\] where \[
+\begin{aligned}
+  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
+& &
+  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
+\end{aligned}
+\]
+
+\hfill\break
+
+\hypertarget{mixture-normal-prior}{%
+\subsubsection{Mixture-Normal Prior}\label{mixture-normal-prior}}
+
+Given the full deign matrix
+\(X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T\), the
+likelihood function of the generalized model can be written as \[
+p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
+\]
+
+Considering the dependency of spline bases, the grouping nature of the
+bases should be taken care of. Naturally, the Bayesian hierarchical
+models allow us easily to take account of the dependency of the spline
+bases as grouping information and provide more reliable estimation. Our
+hierarchical GAMs specify the spike-and-slab mixture normal prior on the
+coefficients: for the coefficients of the \(j\)th predictor,
+\(\beta^0_j\) and \(\beta^{pen}_{jk}\), we have \[
+\begin{aligned}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1), 
+\end{aligned}
+\] where \(\gamma^0_{j}\in\{0,1\}\) and
+\(\gamma^{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator variables,
+indicating if the model includes the linear effect and non-linear effect
+of the \(j\)th variable respectively; \(s_0\) and \(s_1\) are scale
+parameters, assuming \(0 < s_0 < s_1\) and given. Thus, the priors are
+formulated as a mixture of the shrinkage prior normal(0, \(s_0\)) and
+the weakly informative prior normal(0, \(s_1\)), which are spike and
+slab components respectively. Like any other spike and slab priors, the
+spike density is to contain the minimum to zero effects, while the slab
+density is to allow large effects. The scale parameters \(s_0\) and
+\(s_1\) can be treated as tuning parameters, which can be optimized via
+cross-validation. A discussion of how to choose the scale parameters
+comes later.
+
+To note here, unlike the original spike-and-slab prior where each
+predictor have only indicator variable, \(\gamma_j\), we have two
+indicators associated with each predictor, \(\gamma^0_j\) and
+\(\gamma^{pen}_j\) to decide the inclusion of the linear effect and
+non-linear effect of the predictor respectively. It is possible to add
+more constraints on the model, assuming that one indicator variable
+decides the inclusion of both the linear effect and non-linear effect,
+i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). Such formulations,
+commonly referred as ``all-in-all-out'', has been implemented in many
+high dimensional Bayesian GAMs, see Yang \& Narisetty (2020) and Bai
+(Work in progress). The ``all-in-all-out'' set-up was criticized for its
+inability to answering an important research question, if the effects
+are non-linear or not. On the other hand, the model can be much more
+flexible by allowing each dimension of the penalized space to have an
+independent indicator, i.e.~\(\gamma^{pen}_{jk}\) of
+\(\beta^{pen}_{jk}\) for \(k = 1\dots, K_j\). This prior set-up
+generalized to the group spike-and-slab. The additional flexibility
+doesn't not necessarily grand the model neither theoretical
+justification nor analytic simplicity. The spline penalty assumes a
+common smooth shrinkage \(\lambda_j\) applied to all coefficients of the
+penalized space, which contradicts to the model setup. Besides, having
+multiple indicator variables complicates the decision of if non-linear
+effects are necessary.
+
+We set up the hyper-prior of \(\boldsymbol{\gamma}\) to allow local
+adaption of the shrinkage, using a binomial distribution. The two
+indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
+\(\gamma^{pen}_j\), shares the same probability parameter, \[
+\begin{aligned}
+\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
+\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
+\end{aligned}
+\]
+
+This is to leverage the fact that the probability of selecting the bases
+of a smooth function should be similar, while allowing different penalty
+on the null space and penalty space of the spline function. The hyper
+prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
+function selection level, while that of \(\gamma_{j}^{pen}\) decides the
+smoothness of the spline function at individual predictor level.
+Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
+independently distributed for analytic simplicity. However, this set-up
+could possibility result in a situation that is not theoretically
+possible: the non-linear component is selected, but the linear component
+is not. This can be addressed analytically by forcing to include the
+linear component when non-linear component is included. Another possible
+solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
+\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).
+
+We further specify the parameter \(\theta_j\) follows a beta
+distribution with given shape parameters \(a\) and \(b\), \[
+\theta_j \sim Beta(a, b).
+\] The beta distribution is a conjugate prior for the binomial
+distribution and hence provides some computation convenience.
+Specifically, we focus on a special case of beta distribution, uniform
+(0,1) for simplicity and convenience. To note, when the variable have
+large effects in any of the bases, the parameter \(\theta_j\) will be
+estimated large, which in turn encourages the model to include the rest
+of bases, achieving the local adaption among spline bases. Hereafter, we
+refer the hierarchical spline GAM with the spike-and-slab mixture normal
+prior as the ssGAM-MN.
+
+\includegraphics{Fig/SS_prior_flow_chart.png}
+
+~
+
+\hypertarget{mixture-double-exponential-prior}{%
+\subsubsection{Mixture Double Exponential
+Prior}\label{mixture-double-exponential-prior}}
+
+One of the critics received by the spike-and-slab mixture normal prior
+is that the tails of a normal distribution diminishes to zero too fast,
+which renders problem when estimating the large effects. In contrast,
+distributions with heavier tails are preferred, for example \(t\)
+distribution with small degree of freedom and double exponential
+distribution. Specifically, double exponential prior is equivalent to
+the LASSO penalty in the frequentist models, providing a sparse
+solution. Meanwhile, the estimation algorithm of LASSO penalty are
+extremely efficient, greatly reducing the computation needs. Recent
+developments of spike-and-slab priors include spike-and-slab LASSO
+prior, i.e.~spike-and-slab mixture double exponential prior. We extend
+the spike-and-slab LASSO prior to high-dimensional GAMs, providing
+computationally efficiency and sparse solution to our proposed ssGAM-MN.
+
+The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we
+replace the normal distribution with double exponential distribution.
+The prior distributions of the coefficients \(\beta^0_j, \beta^{pen}_j\)
+for the \(j\) predictor can be formulated as \[
+\begin{aligned}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
+\end{aligned}
+\] The hyper-priors for \(\gamma^0_j, \gamma^{pen}_j, \theta_j\) are
+kept the same as specified in ssGAM-MN.
+
+\hfill\break
+
+\hypertarget{alforithm-for-fitting-ssgams}{%
+\subsection{Alforithm for fitting
+ssGAMs}\label{alforithm-for-fitting-ssgams}}
+
+Parsimonious computation is always encouraged in high-dimension data
+analysis. Normally, Bayesian methodology loses its advantages over
+Frequentest penalized model because of the computation cost. Previous
+Bayesian spline models heavily relies on the MCMC algorithm to establish
+posterior distribution of parameters. One of the exception is Bai (in
+progress), who a grouped spike-and-slab LASSO prior \citep{Rockova2018}
+to address the grouping structure of spline bases. He focused on the MAP
+estimator of the parameters, but failed to address the uncertainty
+measure of the estimates. The main difference between Bai's work and the
+proposed models are two-folded: 1) the proposed models are feasible to
+answer the question if the non-linear effects are necessary for the
+predictors, while Bai's work can't due to the ``all-in-all-out''
+approach; 2) the proposed models can provide uncertainly measures in
+addition to the point estimates of the coefficients, while Bai's work
+can't not due to the model fitting algorithm.
+
+In the follow section, we introduced two expectation-maximization (EM)
+based algorithms for fast computing, EM-IRLS and EM-coordinate Descent
+algorithm. Instead of simulating the posterior distribution via MCMC, we
+iteratively estimate the MAP. Compared to the EM-IRLS algorithm, the
+EM-coordinate descent provides an expedited model fitting process at the
+cost of model uncertainty.
+
+\hfill\break
+
+\hypertarget{em-algorithms}{%
+\subsubsection{EM algorithms}\label{em-algorithms}}
+
+EM algorithm is an iterative algorithm to find the maximum likelihood
+estimates or the Bayesian counterpart MAP estimates. It is commonly used
+when some necessary data to establish the likelihood function are
+missing. Instead, the algorithm maximizes the expectation, in respect to
+the ``missing'' data, of the likelihood function.
+
+The recursive algorithm consists of two steps:
+
+\begin{itemize}
+\tightlist
+\item
+  E-step: to calculate the expectation of the likelihood function
+  conditioning on some ``missing'' data
+\item
+  M-step: to maximize the spectated likelihood function to calculate the
+  parameter of interest
+\end{itemize}
+
+For ssGAM, we define the parameters of interest as
+\(\Theta = {\boldsymbol{\beta}, \boldsymbol{\theta}, \phi}\) and
+consider the latent binary indicators \(\boldsymbol{\gamma}\) as
+nuisance parameters of the model. Our objective is to find the
+parameters \(\Theta\) that maximize the posterior density function,or
+equivalently, the logarithm of the density function, \[
+\begin{aligned}
+& \text{argmax}_{\Theta}
+\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
+&= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{aligned}
+\]\\
+
+We use the EM algorithm to find the MAP estimate of \(\Theta\). Since
+the latent binary indicators \(\boldsymbol{\gamma}\) are not of our
+primary interest, we treat them as as the ``missing data'' in the EM
+algorithms. Naturally In the E-step, we calculate the expectation of
+posterior density function of
+\(\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
+respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
+the values from previous iteration \(\Theta^{t-1}\), \[
+E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
+\] Hereafter, we use the short hand notation
+\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{t-1}}(\cdot)\). In the
+M-step, we find the \(\Theta^{t}\) that maximize
+\(E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\).
+To note here, the log-posterior of ssGAM (up to additive constants) can
+be written as a two-part equation
+
+\[ \log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\beta, \phi) + Q_2 (\gamma,\theta),\]
+Where
+\[ Q_1(\boldsymbol{\beta}, \phi) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\]
+and \[
+Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .\]
+\(Q_1\) and \(Q_2\) are respectively, the log-likelihood of model
+coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\),
+and latent indicators \(\boldsymbol{\gamma}, \theta\). Conditioning on
+\(\gamma_{jk}\), \(Q_1\) and \(Q_2\) are independent and can be
+maximized separately for \(\boldsymbol{\beta}, \phi\) and
+\(\boldsymbol{\theta}\).
+
+The E- and M- steps are iterated until the algorithm converge. Depending
+on the choice of the mixture priors, either mixture normal or mixture
+double exponential, the maximization in the M-step can take different
+algorithms, IRLS for uncertainty measures and coordinate descent for
+faster computation.
+
+\hfill\break
+
+\hypertarget{em-irls}{%
+\subsubsection{EM-IRLS}\label{em-irls}}
+
+The main idea of the EM-IRLS algorithm is to use the iterative
+re-weighted least square algorithm to find the estimate
+\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). Recall the prior
+set-up of mixture normal prior, we have the prior densities of
+\(\boldsymbol{\beta}\) conditioning on \(\boldsymbol{\gamma}\)
+
+\[
+\begin{aligned}
+p(\boldsymbol{\beta }| \boldsymbol{\gamma}) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],%\\
+% p(\bs\gamma | \bs \theta) & = \prod\limits_{j=1}^{p}\prod\limits_{k=1}^{K_j} \theta_j^{\gamma_{jk}} (1-\theta_j)^{1-\gamma_{jk}}.
+\end{aligned}
+\] and \begin{align*}
+E(Q_1) &= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^{p} \left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\\
+E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\end{align*} where
+\(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
+\(S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1\). To
+calculate two unknown quantities \(E({S_j^0}^{-1})\) and
+\(E(S^{-1}_j)\), the posterior probability
+\(p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
+\(p_{j} \equiv p(\gamma^{pen}_{j}=1|\Theta^{t-1})\) are necessary, which
+can be derived via Bayes' theorem. The calculation for \(p_j\) is
+slightly different from that of \(p^0_j\), as \(p_j\) depends on the
+value of \(\beta_{jk}\) for \(k=1, \dots, K_j\) and \(p^0_j\) only
+depends on \(\beta_j^0\). The calculation follows the equations below.
+
+\[
+\begin{aligned}
+p_{j}^0 &= \frac{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + Pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
+p_{j} &= \frac{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + Pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+\end{aligned}
+\]
+
+and
+\(Pr(\gamma_{j}^{0} = 1|\theta_j) =Pr(\gamma_{j}^{pen} = 1|\theta_j) = \theta_j\),
+\(Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^{pen} = 0|\theta_j) = 1-\theta_j\),
+\(f(\beta|\gamma=1, s_1) = \text{Normal}(\beta|0 , s_1)\),
+\(f(\beta|\gamma=0, s_0) = \text{Normal}(\beta|0 , s_0)\). It is trivial
+to show \[
+\begin{aligned}
+&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{pen}_{j})  = p_{j}\\
+&E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
+\end{aligned}
+\]
+
+With \(E({S^0}^{-1}_{j})\) and \(E(S^{-1}_{j})\) known, \(E(Q_1)\) can
+be also viewed as a \(l_2\) penalized likelihood function:
+\(E({S^0}^{-1}_{j})\) controls the shrinkage of the linear effect,
+informing if the predictor should be included in the model;
+\(E(S^{-1}_{j})\) controls the smoothness of the spline function.
+\(l_2\) penalized likelihood function can be maximize using the IWLS
+algorithm. The likelihood function can be approximate by a weighted
+normal likelihood: \[
+  p(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
+the pseudo-response and pseudo-weight, respectively.The pseudo data and
+the pseudo-weight are calculated by \[
+    \begin{aligned}
+    z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+    \end{aligned}
+    \] where \(\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})\) is
+the estimated linear predictor using current estimated value of
+\(\boldsymbol{\beta}\), \(L^{'}\) and \(L^{''}\) are the first
+derivative and second derivative of the log-likelihood function. The
+\(l_2\) penalty,
+\(\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\)
+can be seems as a multivariate normal distribution of mean
+\(\textbf{0}\) and variance-covariance matrix
+\(\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})\),
+conditioning on the value of \(E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})\).
+An augmentation step combines the normal approximation and multivariate
+normal for \(\boldsymbol{\beta}\), similarly to Ridge regression model
+fitting. The estimates of \(\boldsymbol{\beta}\) and \(\phi\) can be
+quickly updated in the IRLS algorithm.
+
+The remaining parameters of interest \(\boldsymbol{\theta}\) can be
+updated by maximizing \(E(Q_2)\). As the beta distribution is a
+conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
+be easily updated with a closed form equation: \[
+\theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
+\]
+
+Totally, the framework of the proposed EM IRLS algorithm was summarized
+as follows:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
+\def\labelenumi{\arabic{enumi})}
 \item
-  In \texttt{authormark} use \emph{et al.} if there are three or more
-  authors.
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
 \item
-  Note the use of \texttt{num} to link names and addresses.
-\item
-  For submitting a double-spaced manuscript, add \texttt{doublespace} as
-  an option to a \texttt{classoption} line in the YAML header:
-  \texttt{classoption:\ doublespace}.
-\item
-  Keywords are separated by semicolons.
+  Iterate over the E-step and M-step until convergence
 \end{enumerate}
 
-\hypertarget{the-body-of-the-article}{%
-\section{The Body of the Article}\label{the-body-of-the-article}}
+E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
+\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
+\(\Theta^{(t-1)}\) from previous iteration
 
-\hypertarget{mathematics}{%
-\subsection{Mathematics}\label{mathematics}}
+M-step:
 
-Use mathematics in Rmarkdown as usual.
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\item
+  Update \(\boldsymbol{\beta}^{(t)}\) using the IRLS algorithm
+\item
+  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
+\end{enumerate}
 
-\hypertarget{figures-and-tables}{%
-\subsection{Figures and Tables}\label{figures-and-tables}}
+We assess convergence by the criterion:
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
+\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
+the \(t\)th iteration, and \(\epsilon\) is a small value (say
+\(10^{-5}\)).
 
-Figures are supported from R code:
+\hfill\break
 
-\begin{verbatim}
-x = rnorm(10)
-y = rnorm(10)
-plot(x, y)
-\end{verbatim}
+\hypertarget{em-coordinate-descent}{%
+\subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
+
+When the prior distribution is set to mixture double exponential,
+coordinate descent algorithm can be used to estimate
+\(\boldsymbol{b}eta\), replacing IRLS, in the M-step. Coordinate descent
+is an optimization algorithm that offers extreme computational
+advantage, and famous for its application in optimizing the \(l_1\)
+penalized likelihood function.
+
+Recall, the density function of spike-and-slab mixture double
+exponential prior can be written as \[
+f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}).
+\]
+
+This can simplify our posterior likelihood function to \[
+E(Q_1) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right]
+,
+\] where the approximation of \(E({S^0_j}^{-1}\) and \(E(S^{-1}_{j})\)
+remains the same as in the EM-IRLS algorithm. The formulation of
+\(E(Q_1)\) can be seen as a \(l_1\) penalized likelihood function, and
+optimized via coordinate descent algorithm. The estimates can be used to
+further estimates \(\boldsymbol{\theta}\) in the M-step, following the
+same equation in the EM-IRLS.
+
+Totally, the framework of the proposed EM-coordinate descent algorithm
+was summarized as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
+\item
+  Iterate over the E-step and M-step until convergence
+\end{enumerate}
+
+E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
+\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
+\(\Theta^{(t-1)}\) from previous iteration
+
+M-step:
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\item
+  Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
+  algorithm
+\item
+  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
+\end{enumerate}
+
+We assess convergence by the criterion:
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
+\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
+the \(t\)th iteration, and \(\epsilon\) is a small value (say
+\(10^{-5}\)).
+
+~
+
+\hypertarget{selecting-optimal-scale-values}{%
+\subsection{Selecting Optimal Scale
+Values}\label{selecting-optimal-scale-values}}
+
+Our proposed prior, spike and slab spline prior, requires two preset
+scale parameters (\(s_0\), \(s_1\)). Hence, we would construct a two
+dimensional grid, consists of different pairs of (\(s_0\), \(s_1\))
+value. However, previous research on the topic\footnote{TODO: add
+  citation} suggested the value of slab scale \(s_1\) have less impact
+on the final model and is recommended to be set as a generally large
+value of \(s_1 = 1\) that provides no or weak shrinkage. More attention
+is focused on examining different values of spike scale \(s_0\). Instead
+of the 2-D grid, We consider a sequence of \(L\) decreasing values
+\(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
+spike scale s\_0 tends to include more non-zero coefficients in the
+model. An measure of preference, e.g.~area under the curve (AUC), mean
+squared error, can be used to facilitate the selection of a final model.
+The procedure is similar to the lasso implemented in the widely used R
+package \texttt{glmnet}, which quickly fits the lasso model over a list
+of values of \(\lambda\) covering the entire range, giving a sequence of
+models for users to choose from.
+
+\hypertarget{simulation}{%
+\section{Simulation}\label{simulation}}
+
+\hypertarget{real-data-analysis}{%
+\section{Real Data Analysis}\label{real-data-analysis}}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\hypertarget{double-spacing}{%
-\subsection{Double Spacing}\label{double-spacing}}
-
-If you need to double space your document for submission please use the
-\texttt{doublespace} option in the header.
-
-\hypertarget{bibliography}{%
-\section{Bibliography}\label{bibliography}}
-
-Link a \texttt{.bib} document via the YAML header, and bibliography will
-be printed at the very end (as usual). The default bibliography style is
-provided by Wiley as in \texttt{WileyNJD-AMA.bst}, do not delete that
-file.
-
-Use the Rmarkdown equivalent of the \LaTeX citation system using
-\texttt{{[}@\textless{}name\textgreater{}{]}}. Example:
-\citep{Taylor1937}, \citep{Knupp1999, Kamm2000}.
-
-To include all citation from the \texttt{.bib} file, add
-\texttt{\textbackslash{}nocite\{*\}} before the end of the document.
-
-\hypertarget{further-information}{%
-\section{Further information}\label{further-information}}
-
-All \LaTeX enviroments supported by the main template are supported here
-as well; see the \texttt{.tex} sample file
-\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1097-0258/homepage/la_tex_class_file.htm}{here}
-for more details and example.
+\hypertarget{references}{%
+\section{References}\label{references}}
 

### 2021-05-31 Real_Data/Emory_Card_Biobank/Report.Rmd
Adding Emory Card Biobank to Real Data Analysis

---
title: "Emory Cardiovscular Biobank"
author: "Boyi Guo"
date: "5/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(targets)
library(readr)
library(here)
library(gtsummary)
library(tidyverse)

train_dat <- targets::tar_read(RD_ECB_train_dat)
valid_dat <- targets::tar_read(RD_ECB_valid_dat)
```

#Background

Data was retrieved from <https://datadryad.org/stash/dataset/doi:10.5061/dryad.866t1g1mt> on May 31, 2021.

For study design, please visit the article <https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0237579>

## Data Introducetion

* Two data sets: model development set (N=`r NROW(train_dat)`), validation set (N=`r NROW(valid_dat)`)

* Variable Descriptions
GENEID: subject IDs
timetodeath3yr: time to death event or censoring at three years
death3yr: death event at three years
age: age in years
male: male gender, 1=male, 0=female
BlackRace: black race, 1=black, 0=non-black
batch: batch of metabolomics profiling
Strokehx: history of stroke, 1=yes, 0=no
CABGhx: prior CABG, 1=yes, 0=no
PVDhx: peripheral artery disease, 1=yes, 0=no
eGFR_max120_lt60: estimated glomerular filtration rate less than 60 ml/min/1.73m2, 1=yes, 0=no
curr_smoking: current smoking, 1=yes, 0=no
HFhx: heart failure history, 1=yes, 0=no
HTN: hypertension, 1=yes, 0=no
DM: diabetes, 1=yes, 0=no
mzXX_tXX: intensities of metabolic features

### Covariate Summary
```{r}
list(first_set = train_dat,
     second_set = valid_dat) %>% 
  map(~select(.x, # Data
              age, male, BlackRace,
           Strokehx:DM, 
           death3yr, timetodeath3yr) %>% 
        gtsummary::tbl_summary()
      ) %>% 
tbl_merge()
```

# Primary Data Analysis of Metabolites

### 2021-05-31 Real_Data/curatedBladderData.R

@@ -1,0 +1,47 @@
+# if (!requireNamespace("BiocManager", quietly = TRUE))
+#   install.packages("BiocManager")
+# 
+# BiocManager::install("curatedBladderData")
+
+library(curatedBladderData)
+library(tidyverse)
+
+
+dat_lst <- data(package="curatedBladderData")
+
+#### Display event rate for all datasets in the package ####
+data_brief <- map_dfr(dat_lst$results[,"Item"], .f = function(dat_name){
+  
+  eval(rlang::expr(data(!!dat_name)))
+  
+  dat <- eval(rlang::parse_expr(dat_name))
+  
+  peno_dat <- pData(dat)
+  
+  
+  
+  # remove loaded data
+  eval(rlang::expr(rm(dat)))
+  eval(rlang::expr(rm(!!dat_name)), envir = globalenv())
+  
+  return(table(peno_dat$vital_status) %>% c %>% t %>%
+           data.frame(dataset = dat_name))
+})
+
+
+
+
+
+
+
+
+dat <- GSE32894_eset
+pheno_dat <- pData(dat) %>%
+  select(-uncurated_author_metadata) %>% 
+  mutate_if(is.logical, factor) %>% 
+  mutate_if(is.character, factor) %>% 
+  janitor::remove_constant() 
+summary(pheno_dat)
+
+
+# TODO: write aa function that display the event rate "death rate" for all the datasets

### 2021-05-31 Real_Data/curatedOvarianData.R

@@ -1,0 +1,24 @@
+library(tidyverse)
+library(curatedOvarianData)
+
+
+dat_lst <- data(package="curatedOvarianData")
+
+#### Display event rate for all datasets in the package ####
+data_brief <- map_dfr(dat_lst$results[,"Item"], .f = function(dat_name){
+  
+  eval(rlang::expr(data(!!dat_name)))
+  
+  dat <- eval(rlang::parse_expr(dat_name))
+  
+  peno_dat <- pData(dat)
+  
+  
+  
+  # remove loaded data
+  eval(rlang::expr(rm(dat)))
+  eval(rlang::expr(rm(!!dat_name)), envir = globalenv())
+  
+  return(table(peno_dat$vital_status) %>% c %>% t %>%
+           data.frame(dataset = dat_name))
+})

### 2021-05-31 _targets.R

@@ -1,5 +1,6 @@
 library(targets)
 library(tarchetypes)
+
 # This is an example _targets.R file. Every
 # {targets} pipeline needs one.
 # Use tar_script() to create _targets.R and tar_edit()
@@ -9,10 +10,31 @@ library(tarchetypes)
 
 options(tidyverse.quiet = TRUE)
 # Set target-specific options such as packages.
-tar_option_set(packages = c("dplyr", "tidyverse", "rticles"))
+tar_option_set(packages = c("dplyr", "tidyverse", "rticles", "gtsummary", "here"))
 
 # End this file with a list of target objects.
 list(
   # tar_target(data, data.frame(x = sample.int(100), y = sample.int(100))),
+  
+  
+  #  Real Data Analysis -----------------------------------------------------
+  #* Emory Card Biobank ####
+  tar_target(realdata_ECB_train_path,
+             "Real_Data/Emory_Card_Biobank/Data/Analysis_data_first_cohort.csv",
+             format = "file"),
+  tar_target(realdata_ECB_valid_path,
+             "Real_Data/Emory_Card_Biobank/Data/Analysis_data_second_cohort.csv",
+             format = "file"),
+  tar_target(RD_ECB_train_dat,
+             readr::read_csv(realdata_ECB_train_path)),
+  tar_target(RD_ECB_valid_dat,
+             readr::read_csv(realdata_ECB_valid_path)), 
+  tar_render(real_data_report_ECB,
+             "Real_Data/Emory_Card_Biobank/Report.Rmd"),
+  
+  
+  # Manuscript --------------------------------------------------------------
+  #* Assemble Manuscript ####
   tar_render(manu, "Manuscript/main.Rmd", output_file = "SS_GAM.pdf")
+  
 )

### 2021-05-31 Real_Data/Emory_Card_Biobank/Report.Rmd
Update ECB data analysis

@@
 
-# Primary Data Analysis of Metabolites
+## Primary Data Analysis of Metabolites
+```{r eval=F}
+train_dat %>% select(starts_with("mz")) %>% 
+  summarise_all(var)
+```
+
+# Predictive Model
+
+Outcome: `death3yr` (bin)
+Covariates: None
+Predictors: `starts_with("mz")`
+
+```{r}
+library(BHAM)
+library(BhGLM)
+
+mz_names <- train_dat %>% select(starts_with("mz")) %>% names
+
+test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
+
+sm_df <- data.frame(
+  Var = mz_names,
+ Func = "s",
+ Args ="bs='cr', k=5"
+)
+
+sm_obj <- construct_smooth_data(sm_df, train_dat)
+
+dsn_mat <- sm_obj$data
+
+# The validation data doesn't overlap with training data
+# test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = valid_dat)
+
+mdl_defaul <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial")
+
+mdl_dflt <- bmlasso_spline(dsn_mat, train_dat$death3yr, family = "binomial",
+                           group = make_group(names(dsn_mat)))
+
+measure.bh(mdl_defaul)
+
+measure.bh(mdl_dflt)
+```
+
+## Validation
+```{r eval = F}
+
+measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)
+
+measure.bh(mdl_dflt)
+```

### 2021-05-31 _targets.R
Add Simulation Results: tuning parameter study.

@@ -14,8 +14,14 @@ tar_option_set(packages = c("dplyr", "tidyverse", "rticles", "gtsummary", "here"
 
 # End this file with a list of target objects.
 list(
-  # tar_target(data, data.frame(x = sample.int(100), y = sample.int(100))),
-  
+
+  #  Simulations Studies -----------------------------------------------------
+  #* Tuning Study ####
+  tar_render(sim_tuning_report,
+             "Simulation/tuning/tuning_study_report.Rmd",
+             output_dir = "Simulation/tuning/",
+             output_file = "sim_tuning_report.html"),
+    
   
   #  Real Data Analysis -----------------------------------------------------
   #* Emory Card Biobank ####
@@ -30,7 +36,9 @@ list(
   tar_target(RD_ECB_valid_dat,
              readr::read_csv(realdata_ECB_valid_path)), 
   tar_render(real_data_report_ECB,
-             "Real_Data/Emory_Card_Biobank/Report.Rmd"),
+             "Real_Data/Emory_Card_Biobank/Report.Rmd",
+             output_dir = "Real_Data/Emory_Card_Biobank/",
+             output_file = "ECB_report.html"),
   
   
   # Manuscript --------------------------------------------------------------

### 2021-06-01 Real_Data/Emory_Card_Biobank/Report.Rmd
Addeing split validation in the model

--- 2021-05-31 Real_Data/Emory_Card_Biobank/Report.Rmd
@@
 
-# Primary Data Analysis of Metabolites
+## Primary Data Analysis of Metabolites
+```{r eval=F}
+train_dat %>% select(starts_with("mz")) %>% 
+  summarise_all(var)
+```
+
+# Predictive Model
+
+Outcome: `death3yr` (bin)
+Covariates: None
+Predictors: `starts_with("mz")`
+
+```{r data_split}
+library(tidymodels)
+
+set.seed(1)
+
+data_split <- rsample::initial_split(train_dat, prop = 3/4)
+
+train_data <- training(data_split)
+test_data <- testing(data_split)
+```
+
+## Baseline: Linear Lasso Model
+
+```{r}
+library(glmnet)
+
+lasso_mdl_cv <- cv.glmnet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial", type.measure = "auc")
+
+lasso_mdl <- glmNet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.1se)
+
+# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
+measure.bh(lasso_mdl)
+# assess.glmnet(lasso_mdl$)
+
+measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% data.matrix,
+           new.y = test_data %>% pull(death3yr))
+```
+
+```{r}
+library(BHAM)
+library(BhGLM)
+
+mz_names <- train_dat %>% select(starts_with("mz")) %>% names  
+sm_df <- data.frame(
+  Var = mz_names,
+ Func = "s",
+ Args ="bs='cr', k=5"
+)
+
+sm_obj <- construct_smooth_data(sm_df, train_data)
+
+dsn_mat <- sm_obj$data
+
+# The validation data doesn't overlap with training data
+# test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = valid_dat)
+
+mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial")
+
+mdl_dflt <- bmlasso_spline(dsn_mat, train_data$death3yr, family = "binomial",
+                           group = make_group(names(dsn_mat)))
+
+# Training Data Performance
+measure.bh(mdl_defaul)
+measure.bh(mdl_dflt)
+
+test_sm_dat <- BHAM::make_predict_dat(sm_obj$Smooth, dat=test_data)
+
+# Testing Data Performance
+measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=test_data$death3yr)
+measure.bh(mdl_dflt, new.x = test_sm_dat, new.y=test_data$death3yr)
+```
+
+# Extremely Penalized Model
+```{r}
+
+mdl_pen_extreme <- bmlasso_spline(dsn_mat, train_data$death3yr, family = "binomial",
+                           group = make_group(names(dsn_mat)),
+                           ss = c(0.001, 0.5))
+
+# Training Data Performance
+measure.bh(mdl_pen_extreme)
+
+# Testing Data Performance
+measure.bh(mdl_pen_extreme, new.x = test_sm_dat, new.y=test_data$death3yr)
+```
+
+## Validation
+
+The validation data set have mostly distinct feature from the training data set. Hence, the validation in the traditional sense does not exists.
+
+```{r eval = F}
+
+test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
+measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)
+
+measure.bh(mdl_dflt)
+```

### 2021-06-02 Manuscript/01-Intro.Rmd
* Update manuscript writing * Set up targets for sub files

@@
 It is still not clear what is the best practice on how to model non-linear effects in high-dimensional setting.
+
 Similar to expanding classic models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
@@
 
-<!-- Among all the choices of smoothing function, one particular family stood out, splines.
-Splines are piece-wise polynomial functions to interpolate nonlinear signals.
-Most commonly used spline is cubic spline, thanks to its smoothness property.
-[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smooth function.
-However, the specification of knots is not a trivial task, as the the positioning of knots are influential on the performance of the model.
-As an improvement, researchers proposed smoothing splines where as many knots are used and later shrunk using penalized algorithms.
-The smoothing spline concept was first introduced [@OSullivan1986; @OSullivan1988], and later generalized and made famous by @Eilers1996, called P-spline.
-Independently, @Ruppert1999 proposed similar idea.
-Later, Bayesian GAMs were introduced as an alternative to address measurement error in covariates [@Berry2002] or local adaptive smoothing [@Lang2004].
--->
-
-For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss.
-The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior.
-It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
-$$
- \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
-$$
-
-The most distinct feature of the SS priors family is that it is conditioned on a latent binary variable $\gamma_i$ that indicates whether the covariate $x_i$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$ and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
-Here we introduce the most popular spike-and-slab prior introduced in @George1993, as  stochastic search variable selection (SSVS).
-Given a regression problem
-$$
-Y = \bs X\beta+\epsilon, 
-$$
-we can assume the prior distribution of the regression parameters $\beta$ depends on a vector of binary variables $\gamma=(\gamma_1, \dots \gamma_p)^\prime$ as follows: given $\gamma_i$, $\beta_i$ are independent distributed as a mixture priors
-$$
-\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
-$$
-The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution.
-Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
-@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection.
-The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms.
-[@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms.
-@George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.
-
-To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
-EM-based algorithms are prevalent used to speed the computation.
-@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
-The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
-The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
-Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
-In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
-As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
-EMVS is also compatible with structured prior when dealing with grouped variables.
-The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
-
-The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
-The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
-The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
-Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
-Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
-However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
-@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
-
 To the best of our knowledge, there are few GAM models proposed for high dimensional application.
@@
 @Bai2020Spline shows inferior performance of such practice compared to using an ad hoc high-dimensional additive model.
-
 Among the specialized high-dimensional spline methods, @Ravikumar2009 extended the original backfitting algorithm [@Hastie1987] to accommodate the high-dimensional setting.
@@
 
+<!-- Among all the choices of smoothing function, one particular family stood out, splines.
+Splines are piece-wise polynomial functions to interpolate nonlinear signals.
+Most commonly used spline is cubic spline, thanks to its smoothness property.
+[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smooth function.
+However, the specification of knots is not a trivial task, as the the positioning of knots are influential on the performance of the model.
+As an improvement, researchers proposed smoothing splines where as many knots are used and later shrunk using penalized algorithms.
+The smoothing spline concept was first introduced [@OSullivan1986; @OSullivan1988], and later generalized and made famous by @Eilers1996, called P-spline.
+Independently, @Ruppert1999 proposed similar idea.
+Later, Bayesian GAMs were introduced as an alternative to address measurement error in covariates [@Berry2002] or local adaptive smoothing [@Lang2004].
+-->
+
 In this article, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model.
@@
 Moreover, out model inherent the the robust estimation from Bayesian models.
-
-Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
-Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
-The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
-We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
-
-The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
+Lastly, The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
 We will implement the proposed models and algorithms in one of the most commonly used statistical computing environment R.
 Ancillary functions will be provided in assistance to model diagnostic, interpretation and graphing.
-The R implementation of the Bayesian Hierarchical additive model is available via a freely available R package, BHAM at (https://github.com/boyiguo1/BHAM)[https://github.com/boyiguo1/BHAM].
+The R implementation of the Bayesian hierarchical additive model is available via a freely available R package, BHAM at (https://github.com/boyiguo1/BHAM)[https://github.com/boyiguo1/BHAM].
 
-The rest of the paper are arranged as the followings: in Chapter 32 we demonstrate the Bayesian Hierarchical generalized additive model, where we provide the setup of the proposed model and two fast-fitting algorithms.
+The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian Hierarchical generalized additive model, where we provide the setup of the proposed model and two fast-fitting algorithms.
 In Chapter 3, we Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.

### 2021-06-02 Manuscript/02-Method.Rmd
Add Flowchart to the table
* Update manuscript writing * Set up targets for sub files

@@
 
- \
+## Spike and Slab Prior
+
+For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss.
+The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior.
+It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
+$$
+ \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
+$$
+
+The most distinct feature of the SS priors family is that it is conditioned on a latent binary variable $\gamma_i$ that indicates whether the covariate $x_i$ is included in the model.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$ and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
+Here we introduce the most popular spike-and-slab prior introduced in @George1993, as  stochastic search variable selection (SSVS).
+Given a regression problem
+$$
+Y = \bs X\beta+\epsilon, 
+$$
+we can assume the prior distribution of the regression parameters $\beta$ depends on a vector of binary variables $\gamma=(\gamma_1, \dots \gamma_p)^\prime$ as follows: given $\gamma_i$, $\beta_i$ are independent distributed as a mixture priors
+$$
+\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
+$$
+The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution.
+Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
+@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection.
+The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms.
+[@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms.
+@George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.
+
+To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
+EM-based algorithms are prevalent used to speed the computation.
+@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
+The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
+The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
+Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
+In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
+EMVS is also compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
+
+The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
+The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
+The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
+Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
+Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
+However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
+@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
+
+ Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
+Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
+The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
+We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
 
@@
 Hereafter, we refer the hierarchical spline GAM with the spike-and-slab mixture normal prior as the ssGAM-MN.
+
+```{r}
+include_graphics("Fig/SS_prior_flow_chart.png")
+```
 

### 2021-06-02 Manuscript/05-Discussion.Rmd
* Update manuscript writing * Set up targets for sub files

# Discussion

Remaining issue:
1. Degree of freedom has yet to closely studied. As an direct impact, the functional selection can be improved when effective degree of freedom are more carefully studied
2. Future exapansion, as we were analyzing the metabolites data, we observed that metabolites pathway infromation can be useful in the predictive and integrative model can be developed based on the proposed framework. As an naive solution, the selection indicator of linear effect of the spline terms of all metabolites on the same pathway can be set to share the same hyper prior such that the group selection can be modestly achieved. Meanwhile further advancedment of the methodology regarding to the integrative analysis can consider setting the hypor priors of the selection indivator of the non-linear components conditioning on the probability of inclusion of the linear effect. Such structure is more sensible analytically.

### 2021-06-02 Manuscript/SS_GAM.tex
* Update manuscript writing * Set up targets for sub files

@@
 the best practice on how to model non-linear effects in high-dimensional
-setting. Similar to expanding classic models to high-dimensional models,
+setting.
+
+Similar to expanding classic models to high-dimensional models,
 expanding the GAM framework for high-dimensional applications is not a
@@
 remains as a troubling question.
+
+To the best of our knowledge, there are few GAM models proposed for high
+dimensional application. In common practice, model matrix of spline
+terms are generated using base functions of preference, and then feed to
+a penalized model. Without proper handle of the grouping nature of bases
+functions, the estimation of the smooth function is less accurate.
+\citet{Bai2020Spline} shows inferior performance of such practice
+compared to using an ad hoc high-dimensional additive model. Among the
+specialized high-dimensional spline methods, \citet{Ravikumar2009}
+extended the original backfitting algorithm \citep{Hastie1987} to
+accommodate the high-dimensional setting. The sparsity and smoothness
+was handled separately with a soft threshold. The authors also
+demonstrated its similarity to Group Lasso. \citet{Meier2009} proposed
+an a sparsity-smoothness penalty,
+
+\[
+J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
+\] which can re-formulated as a group lasso problem as well. However,
+the model can not estimate the smoothness of the spline functions
+correctly, as the smoothness of all spline models are controlled by the
+lambda\_2. The inflexible of the model creates problems in prediction.
+
+From the Bayesian perspective, recent works that expands to the
+high-dimensional settings includes \citet{Scheipl2012} and
+\citet{Bai2020Spline}. \citet{Scheipl2012} proposed an new
+spike-and-slab structure prior, normal-mixture-of-inverse gamma for
+semi-parametric regression using MCMC. The normal-mixture-of-inverse
+gamma prior is mixture of t-distribution with an additional coefficients
+mixing vector, where the vector follows a mixture of two Gaussian
+distribution. \citet{Bai2020Spline} proposed an novel high dimensional
+Bayesian GAM using the spike-and-slab group lasso prior. An fast fitting
+EM-based algorithm is proposed. Asymptotic theory, such as posterior
+contraction rates and model selection consistency are proved. Even
+though, the algorithm is flexible to cover many non-Gaussian family in a
+high dimensional setting, the proposed method didn't address uncertainty
+quantification, i.e.~calculating credible intervals etc. However, most
+of these Bayesian GAMs still suffer a common problem, computation cost.
+These Bayesian GAMs uses MCMC to estimate the posterior distribution,
+which would create an intense computation load.
+
+In this article, we aim to address the challenges of non-linear signal
+modeling in high-dimensional application using Bayesian Hierarchical
+model. Specifically, we will propose a new Bayesian spike-and-slab prior
+for the generalized additive models and Cox spline models. The
+contribution of the proposed prior comes in two-fold: 1) it supports
+simultaneous variable selection and function selection; 2) it provides
+robust estimation and prediction for the coefficients and outcomes. As
+previously described, simultaneous variable selection and its functional
+form selection is not a trivial problem. Nevertheless, the Bayesian
+spike-and-slab prior is a perfect solution to the problem, as the latent
+binary indicator can be used to for variable selection and the spike and
+slab densities are used to adaptive adjust the smoothness of the spline
+function. Moreover, compared to currently available Bayesian spline
+models which mostly conduct a `all-in-all-out' strategy, i.e.~either
+including a non-linear form of a variable or excluding the variable
+completely, we offer a more flexible solution, rending one of three
+possibilities: no effect of a predictor, only linear effect of a
+predictor, or non-linear effect. This simplifies the process of
+determining what effect a predictor have on the outcome, without
+involving hypothesis testing where the degree of freedom for non-linear
+effect is complicated to estimate due to the shrinkage. Moreover, out
+model inherent the the robust estimation from Bayesian models. Lastly,
+The third aim of the dissertation work will focus on the software
+implementation of the proposed models and algorithms. We will implement
+the proposed models and algorithms in one of the most commonly used
+statistical computing environment R. Ancillary functions will be
+provided in assistance to model diagnostic, interpretation and graphing.
+The R implementation of the Bayesian hierarchical additive model is
+available via a freely available R package, BHAM at
+(\url{https://github.com/boyiguo1/BHAM}){[}\url{https://github.com/boyiguo1/BHAM}{]}.
+
+The rest of the paper are arranged as the followings: in Chapter 2 we
+demonstrate the Bayesian Hierarchical generalized additive model, where
+we provide the setup of the proposed model and two fast-fitting
+algorithms. In Chapter 3, we Application to real-world datasets
+demonstrates the usefulness of the proposed methodologies in Chpater 4.
+Conclusion and discussions are made in Chapter 5.
+
+\hypertarget{spike-and-slab-additive-model}{%
+\section{Spike-and-slab Additive
+Model}\label{spike-and-slab-additive-model}}
+
+Following the notation in the previous section \ref{Notation}, suppose
+we have the model matrix \(\boldsymbol{X}\) derived from spline
+functions of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates
+that are modeled nonparametrically, the total dimension of
+\(\boldsymbol{X}\) is \(n \times \sum\limits_{j=1}^p k_j\). Column-wise
+centering and re-parameterization introduced in the GAM as random effect
+section{[}\^{}5{]} are performed. The re-parameterization has two
+purposes. First of all the re-parameterization reshapes the columns to
+be linearly independent to each other, and hence, independent prior can
+be set to the coefficients. Secondly, the re-parameterization decomposes
+the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
+sub-matrices representing the null space of the spline and penalized
+space of the spline, \[
+\boldsymbol{X}_j = \begin{bmatrix}
+\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
+\end{bmatrix},
+\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
+the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
+denotes the column spaces that need to be shrunk for smoothness of the
+spline. The purpose for this re-parameterization is to isolate the null
+space from the smoothing space, such that answering the question, if
+non-linear relationship is necessary, becomes easy by deciding if the
+corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
+zero. This strategy is widely used, previously mentioned in
+\citet{Scheipl2012} and the references there in.\footnote{Possibly
+  select some to add here.} Hereafter, we denote the coefficients of the
+null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
+penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
+for the \(j\)the covariate is \[
+f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
+\] where the superscript \(^t\) denotes matrix transpose.
+
+Without loss of generality, we assume that the model doesn't include the
+the parametric part of the model \(\boldsymbol{A}\). The coefficients of
+\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
+by local update conditioning on the temporary value of the spline
+coefficients.
+
+\hypertarget{spike-and-slab-prior}{%
+\subsection{Spike and Slab Prior}\label{spike-and-slab-prior}}
 
@@
 
-To the best of our knowledge, there are few GAM models proposed for high
-dimensional application. In common practice, model matrix of spline
-terms are generated using base functions of preference, and then feed to
-a penalized model. Without proper handle of the grouping nature of bases
-functions, the estimation of the smooth function is less accurate.
-\citet{Bai2020Spline} shows inferior performance of such practice
-compared to using an ad hoc high-dimensional additive model.
-
-Among the specialized high-dimensional spline methods,
-\citet{Ravikumar2009} extended the original backfitting algorithm
-\citep{Hastie1987} to accommodate the high-dimensional setting. The
-sparsity and smoothness was handled separately with a soft threshold.
-The authors also demonstrated its similarity to Group Lasso.
-\citet{Meier2009} proposed an a sparsity-smoothness penalty,
-
-\[
-J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
-\] which can re-formulated as a group lasso problem as well. However,
-the model can not estimate the smoothness of the spline functions
-correctly, as the smoothness of all spline models are controlled by the
-lambda\_2. The inflexible of the model creates problems in prediction.
-
-From the Bayesian perspective, recent works that expands to the
-high-dimensional settings includes \citet{Scheipl2012} and
-\citet{Bai2020Spline}. \citet{Scheipl2012} proposed an new
-spike-and-slab structure prior, normal-mixture-of-inverse gamma for
-semi-parametric regression using MCMC. The normal-mixture-of-inverse
-gamma prior is mixture of t-distribution with an additional coefficients
-mixing vector, where the vector follows a mixture of two Gaussian
-distribution. \citet{Bai2020Spline} proposed an novel high dimensional
-Bayesian GAM using the spike-and-slab group lasso prior. An fast fitting
-EM-based algorithm is proposed. Asymptotic theory, such as posterior
-contraction rates and model selection consistency are proved. Even
-though, the algorithm is flexible to cover many non-Gaussian family in a
-high dimensional setting, the proposed method didn't address uncertainty
-quantification, i.e.~calculating credible intervals etc. However, most
-of these Bayesian GAMs still suffer a common problem, computation cost.
-These Bayesian GAMs uses MCMC to estimate the posterior distribution,
-which would create an intense computation load.
-
-In this article, we aim to address the challenges of non-linear signal
-modeling in high-dimensional application using Bayesian Hierarchical
-model. Specifically, we will propose a new Bayesian spike-and-slab prior
-for the generalized additive models and Cox spline models. The
-contribution of the proposed prior comes in two-fold: 1) it supports
-simultaneous variable selection and function selection; 2) it provides
-robust estimation and prediction for the coefficients and outcomes. As
-previously described, simultaneous variable selection and its functional
-form selection is not a trivial problem. Nevertheless, the Bayesian
-spike-and-slab prior is a perfect solution to the problem, as the latent
-binary indicator can be used to for variable selection and the spike and
-slab densities are used to adaptive adjust the smoothness of the spline
-function. Moreover, compared to currently available Bayesian spline
-models which mostly conduct a `all-in-all-out' strategy, i.e.~either
-including a non-linear form of a variable or excluding the variable
-completely, we offer a more flexible solution, rending one of three
-possibilities: no effect of a predictor, only linear effect of a
-predictor, or non-linear effect. This simplifies the process of
-determining what effect a predictor have on the outcome, without
-involving hypothesis testing where the degree of freedom for non-linear
-effect is complicated to estimate due to the shrinkage. Moreover, out
-model inherent the the robust estimation from Bayesian models.
-
 Given the proposed model specification, the estimation of the parameters
@@
 comparison to the model averaging approach.
-
-The third aim of the dissertation work will focus on the software
-implementation of the proposed models and algorithms. We will implement
-the proposed models and algorithms in one of the most commonly used
-statistical computing environment R. Ancillary functions will be
-provided in assistance to model diagnostic, interpretation and graphing.
-The R implementation of the Bayesian Hierarchical additive model is
-available via a freely available R package, BHAM at
-(\url{https://github.com/boyiguo1/BHAM}){[}\url{https://github.com/boyiguo1/BHAM}{]}.
-
-The rest of the paper are arranged as the followings: in Chapter 32 we
-demonstrate the Bayesian Hierarchical generalized additive model, where
-we provide the setup of the proposed model and two fast-fitting
-algorithms. In Chapter 3, we Application to real-world datasets
-demonstrates the usefulness of the proposed methodologies in Chpater 4.
-Conclusion and discussions are made in Chapter 5.
-
-\hypertarget{spike-and-slab-additive-model}{%
-\section{Spike-and-slab Additive
-Model}\label{spike-and-slab-additive-model}}
-
-Following the notation in the previous section \ref{Notation}, suppose
-we have the model matrix \(\boldsymbol{X}\) derived from spline
-functions of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates
-that are modeled nonparametrically, the total dimension of
-\(\boldsymbol{X}\) is \(n \times \sum\limits_{j=1}^p k_j\). Column-wise
-centering and re-parameterization introduced in the GAM as random effect
-section{[}\^{}5{]} are performed. The re-parameterization has two
-purposes. First of all the re-parameterization reshapes the columns to
-be linearly independent to each other, and hence, independent prior can
-be set to the coefficients. Secondly, the re-parameterization decomposes
-the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
-sub-matrices representing the null space of the spline and penalized
-space of the spline, \[
-\boldsymbol{X}_j = \begin{bmatrix}
-\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
-\end{bmatrix},
-\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
-the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
-denotes the column spaces that need to be shrunk for smoothness of the
-spline. The purpose for this re-parameterization is to isolate the null
-space from the smoothing space, such that answering the question, if
-non-linear relationship is necessary, becomes easy by deciding if the
-corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
-zero. This strategy is widely used, previously mentioned in
-\citet{Scheipl2012} and the references there in.\footnote{Possibly
-  select some to add here.} Hereafter, we denote the coefficients of the
-null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
-penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
-for the \(j\)the covariate is \[
-f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
-\] where the superscript \(^t\) denotes matrix transpose.
-
-Without loss of generality, we assume that the model doesn't include the
-the parametric part of the model \(\boldsymbol{A}\). The coefficients of
-\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
-by local update conditioning on the temporary value of the spline
-coefficients.
-
-~
 
@@
 
+Remaining issue: 1. Degree of freedom has yet to closely studied. As an
+direct impact, the functional selection can be improved when effective
+degree of freedom are more carefully studied 2. Future exapansion, as we
+were analyzing the metabolites data, we observed that metabolites
+pathway infromation can be useful in the predictive and integrative
+model can be developed based on the proposed framework. As an naive
+solution, the selection indicator of linear effect of the spline terms
+of all metabolites on the same pathway can be set to share the same
+hyper prior such that the group selection can be modestly achieved.
+Meanwhile further advancedment of the methodology regarding to the
+integrative analysis can consider setting the hypor priors of the
+selection indivator of the non-linear components conditioning on the
+probability of inclusion of the linear effect. Such structure is more
+sensible analytically.
+
 \begin{figure}

### 2021-06-02 Manuscript/main.Rmd
Add Flowchart to the table
* Update manuscript writing * Set up targets for sub files

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+
+library(knitr)
+library(targets)
+library(here)
+
+tar_load(intro_path)
+tar_load(method_path)
+tar_load(sim_path)
+tar_load(real_data_path)
+tar_load(disc_path)
 ```
 
-```{r child="01-Intro.Rmd"}
+```{r child=here(intro_path)}
 ```
 
-```{r child="02-Method.Rmd"}
+```{r child=here(method_path)}
 ```
 
-```{r child="03-Simulation.Rmd"}
+```{r child=here(sim_path)}
 ```
 
-```{r child="04-Real_Data.Rmd"}
+```{r child=here(real_data_path)}
 ```
 
-```{r child="05-Discussion.Rmd"}
+```{r child=here(disc_path)}
 ```

### 2021-06-02 Manuscript/main.tex
* Update manuscript writing * Set up targets for sub files

@@
 
-\includegraphics[width=24.76in]{Fig/SS_prior_flow_chart}
+\includegraphics{Fig/SS_prior_flow_chart.png}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+Remaining issue: 1. Degree of freedom has yet to closely studied. As an
+direct impact, the functional selection can be improved when effective
+degree of freedom are more carefully studied 2. Future exapansion, as we
+were analyzing the metabolites data, we observed that metabolites
+pathway infromation can be useful in the predictive and integrative
+model can be developed based on the proposed framework. As an naive
+solution, the selection indicator of linear effect of the spline terms
+of all metabolites on the same pathway can be set to share the same
+hyper prior such that the group selection can be modestly achieved.
+Meanwhile further advancedment of the methodology regarding to the
+integrative analysis can consider setting the hypor priors of the
+selection indivator of the non-linear components conditioning on the
+probability of inclusion of the linear effect. Such structure is more
+sensible analytically.
 

### 2021-06-02 _targets.R

@@ -25,6 +25,7 @@ list(
   
   #  Real Data Analysis -----------------------------------------------------
   #* Emory Card Biobank ####
+  #** Load Data ####
   tar_target(realdata_ECB_train_path,
              "Real_Data/Emory_Card_Biobank/Data/Analysis_data_first_cohort.csv",
              format = "file"),
@@ -35,14 +36,47 @@ list(
              readr::read_csv(realdata_ECB_train_path)),
   tar_target(RD_ECB_valid_dat,
              readr::read_csv(realdata_ECB_valid_path)), 
-  tar_render(real_data_report_ECB,
-             "Real_Data/Emory_Card_Biobank/Report.Rmd",
-             output_dir = "Real_Data/Emory_Card_Biobank/",
-             output_file = "ECB_report.html"),
-  
+  #** Report ####
+  # tar_render(real_data_report_ECB,
+  #            "Real_Data/Emory_Card_Biobank/Report.Rmd",
+  #            output_dir = "Real_Data/Emory_Card_Biobank/",
+  #            output_file = "ECB_report.html"),
+  # 
   
   # Manuscript --------------------------------------------------------------
+
+  #* Section Paths ####
+  tar_target(intro_path,
+             "Manuscript/01-Intro.Rmd",
+             format = "file"),
+  
+  tar_target(method_path,
+             "Manuscript/02-Method.Rmd",
+             format = "file"),
+  
+  tar_target(sim_path,
+             "Manuscript/03-Simulation.Rmd",
+             format = "file"),
+  
+  tar_target(real_data_path,
+             "Manuscript/04-Real_Data.Rmd",
+             format = "file"),
+  
+  tar_target(disc_path,
+             "Manuscript/05-Discussion.Rmd",
+             format = "file"),
+  tar_target(fig_path,
+             "Manuscript/Fig/",
+             format = "file"),
+  
   #* Assemble Manuscript ####
-  tar_render(manu, "Manuscript/main.Rmd", output_file = "SS_GAM.pdf")
+  tar_render(manu, "Manuscript/main.Rmd", 
+             # params = list(
+             #   intro = intro_path,
+             #   method = method_path,
+             #   sim = sim_path,
+             #   RD = real_data_path,
+             #   disc = disc_path),
+             output_file = "SS_GAM.pdf")
   
 )

### 2021-06-02 Manuscript/00-main.Rmd
Improved the manuscript

--- 2021-06-02 Manuscript/main.Rmd
@@
 
-```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}"}
+```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}", eval = F}
 x = rnorm(10)
@@
 
-```{r, results = "asis"}
+```{r, results = "asis", eval = F}
 if (!require("xtable")) install.packages("xtable")

### 2021-06-02 Manuscript/05-Discussion.Rmd
Improved the manuscript

@@
 2.
+While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
+Our current effort resides on expanding.
+An naive approach could be convert an time-to-event outcome to a poisson outcome following xxx.
+Hence, the .
+However, it would be more efficient to , via penalized pseudo likelihood function, which is the current effort resides.
+2.
 Future exapansion, as we were analyzing the metabolites data, we observed that metabolites pathway infromation can be useful in the predictive and integrative model can be developed based on the proposed framework.

### 2021-06-02 Manuscript/SS_GAM.tex
Improved the manuscript

@@
 \section{Introduction}\label{introduction}}
-
-\begin{itemize}
-\item
-  What is the problem
-
-  \begin{itemize}
-  \tightlist
-  \item
-    what is the GAM
-  \item
-    expanding GAM in to high-dimensional frame work
-  \item
-    Why this is important
-  \item
-    Why this is hard
-  \end{itemize}
-\item
-  Previous solutions
-
-  \begin{itemize}
-  \tightlist
-  \item
-    What are the methods
-  \item
-    What are their problems
-  \end{itemize}
-\item
-  Introduce the wiggliness matrix
-\item
-  Our method
-\item
-  The implication of our method
-\item
-  Fast
-\item
-  Structure of the rest of the paper
-\end{itemize}
 
@@
 direct impact, the functional selection can be improved when effective
-degree of freedom are more carefully studied 2. Future exapansion, as we
-were analyzing the metabolites data, we observed that metabolites
-pathway infromation can be useful in the predictive and integrative
-model can be developed based on the proposed framework. As an naive
-solution, the selection indicator of linear effect of the spline terms
-of all metabolites on the same pathway can be set to share the same
-hyper prior such that the group selection can be modestly achieved.
-Meanwhile further advancedment of the methodology regarding to the
-integrative analysis can consider setting the hypor priors of the
-selection indivator of the non-linear components conditioning on the
-probability of inclusion of the linear effect. Such structure is more
-sensible analytically.
-
-\begin{figure}
-\centering
-\includegraphics{SS_GAM_files/figure-latex/plot-ref-1.pdf}
-\caption{Fancy Caption\label{fig:plot}}
-\end{figure}
+degree of freedom are more carefully studied 2. While the proposed model
+addresses a great deal of analytic problem, analyzing the time-to-event
+outcome remains unsolved. Our current effort resides on expanding. An
+naive approach could be convert an time-to-event outcome to a poisson
+outcome following xxx. Hence, the . However, it would be more efficient
+to , via penalized pseudo likelihood function, which is the current
+effort resides. 2. Future exapansion, as we were analyzing the
+metabolites data, we observed that metabolites pathway infromation can
+be useful in the predictive and integrative model can be developed based
+on the proposed framework. As an naive solution, the selection indicator
+of linear effect of the spline terms of all metabolites on the same
+pathway can be set to share the same hyper prior such that the group
+selection can be modestly achieved. Meanwhile further advancedment of
+the methodology regarding to the integrative analysis can consider
+setting the hypor priors of the selection indivator of the non-linear
+components conditioning on the probability of inclusion of the linear
+effect. Such structure is more sensible analytically.
 
@@
 Analogously, use Rmarkdown to produce tables as usual:
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrr}
-  \hline
- & speed & dist \\ 
-  \hline
-1 & 4.00 & 2.00 \\ 
-  2 & 4.00 & 10.00 \\ 
-  3 & 7.00 & 4.00 \\ 
-  4 & 7.00 & 22.00 \\ 
-  5 & 8.00 & 16.00 \\ 
-  6 & 9.00 & 10.00 \\ 
-   \hline
-\end{tabular}
-\caption{A table} 
-\label{tab:table}
-\end{table}
 

### 2021-06-02 _targets.R

@@ -65,12 +65,13 @@ list(
   tar_target(disc_path,
              "Manuscript/05-Discussion.Rmd",
              format = "file"),
-  tar_target(fig_path,
-             "Manuscript/Fig/",
-             format = "file"),
   
+  # tar_target(fig_path,
+  #            "Manuscript/Fig/",
+  #            format = "file"),
+  # 
   #* Assemble Manuscript ####
-  tar_render(manu, "Manuscript/main.Rmd", 
+  tar_render(manu, "Manuscript/00-main.Rmd", 
              # params = list(
              #   intro = intro_path,
              #   method = method_path,

### 2021-06-15 Manuscript/03-Simulation.Rmd
Adding Content

## Simulation Study
In this section, we compare the performance of the proposed models to 3 other models: COSSO, adaptive COSSO ,mgcv. COSSO is xxx method. Adaptive COSSO is another method, mgcv is another method. 
To demonstrate the correctness and feasibility of the proposed SS-GAM,
we follow the logistic regression simulation introduced in Bai (Work in
Progress). We first generate $n=500$ training data sample with number
of predictors $p=4, 10, 50, 200$ respectively, where the predictors $X \sim MVN_{n\times p}(0, I_{P})$. We simulate the outcome data from three distributions, gaussian, poisson, and binomial with the identity link, logit link $g(x) = \log(\frac{x}{1-x})$, and log link $g(x) = log(x)$ functions to mimic the common use of generalized model. The outcome is simulated via the following function
$$
\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
$$
for gaussian and binomial outcome and shrinked effects

$$
\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
$$
for stable poisson outcome. For the Gaussian outcome where dispersion are required, the dispersion parameter was set to 1. We follow the same data generating functions in Bai. In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest
covariates are inactive $f_j(x_j) = 0$ for $j = 4, \dots, p$. Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data. For mgcv and proposed models where choice of spline function are customizable, we choose to cubic regression spline with 10 basis. In COSSO and adaptive COSSO where the choice of spline are pre-determined, we set to have 10 bases. The corresponding number of parameter that we estimate with the model is 37, 82, 451 and 1801. We are run the simulation where the number of bases are increased to 25 deliberately. To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across50 simulations. The results are tabulated in Table 1, 2, 3.

%are calculated to assess out-of-sample estimation accuracy,

%We also calculated the area under the curve (AUC) of the receiver operating characteristic curve for $y_{new}$. We will also visually display the estimated smooth function with the smooth function of truth to compare if the models can infer the smooth function correctly, despite the prediction accuracy.

%When constructing the smooth function of each variables, we constructs the natural cubic spline with degree of freedom 25. In other words, the design matrix of each smooth function will have 26 columns. The total number of columns for $p=4, 10, 50$ is $105, 261, 1301$ respectively, where intercept is kept in the model when fitting. The $p=50$ can mimic the high dimensional setting as the design matrix are in high-dimensional setting ($p >> n$)

<!-- The accuracy of variable selection is measured using Matthews correlation coefficient (MCC), -->
<!-- $$ -->
<!-- MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}, -->
<!-- $$ -->
<!-- where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively. -->

%We benchmark the performance of the proposed methods with the state of art GAM package `mgcv`[TODO: add citation]. In the future steps, we will compare our proposed model with other high dimensional GAM applications.

 \

### Simulation Results

Overall, the proposed priors outperforms the state-of-art GAM model `mgcv` when the dimensionality is low, median, and high. The proposed priors predicts better than `mgcv` and infers better of the underlying smooth function. Most importantly, `mgcv` is infeasible in the high dimensional setting, i.e. when $p=50$ where the number of parameters are greater than the sample size.

#### Prediction Performace

The prediction performance is tabulated in the table below for AUC and Supplementary tables for MSE and misclassification rate. In summary, the proposed Bayesian models outperforms the benchmark model `mgcv` regardless the choice of prior or fitting algorithm. The models with IRLS fitting algorithm give slightly better performance than the the model with cyclic coordinate descent algorithm. The advantage diminishes in the high dimensional setting (p=50). This is possible because the SS normal spline prior provides a more smooth solution, which would work well in a low to median dimension setting. However, in the high dimensional setting, there are not enough sparsity in the model, i.e. SS normal prior will provides a more complex model than necessary. This is confirmed with the visualization in the following section.

```{r tab.id="sim_GAM_AUC", eval=F}
glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")

glm_spline_pred_dat$auc %>% 
        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
        flextable() %>% 
  set_caption(caption = "Averaged AUC of out-of-bag samples",
              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_AUC", pre_label = "Table "))
```

```{r  eval=F}
glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")

glm_spline_pred_dat$misclass %>% 
        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
        flextable() %>% 
  set_caption(caption = "Averaged misclassification rate of out-of-bag samples",
              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_Misclass", pre_label = "Table "))
```

```{r  eval=F}

glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")

glm_spline_pred_dat$mse %>% 
        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
        flextable() %>% 
  set_caption(caption = "Averaged mean squared area of out-of-bag samples",
              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_MSE", pre_label = "Table "))
```

#### Visualization
We plot the estimated smooth functions of the four active variables for the 100 simulation iterations, along with the 'true' smooth function. Across different settings of predictors, the patterns look similar. First of all, in all the settings, `mgcv` tends to provide more  extreme models, where in the graphs the smooth function have much higher bound than the true smooth function. Rarely the estimated smooth function from `mgcv` matches with the true smooth function. In comparison, the proposed method can interpolate the shape of the true smooth function. However, different priors and fitting algorithms have different advantages. For example, in Figure 1, we see that the SS normal distribution have problem with the tail of the smooth function, especially obvious in the estimations of the variable _x4_. However, it would have less sparse solutions, which is more likely to accurately interpolate the smooth function especially when the shape is complicated (see variable _x2_). On the contrary, the SS double exponential prior model encourages sparse solutions, which can sometime omit active predictors.

### 2021-06-15 Manuscript/05-Discussion.Rmd
Adding Content

@@
 Such structure is more sensible analytically.
+
+## Discussion
+
+In this chapter, we propose the spike-and-slab generalized additive models for high-dimensional analysis.Two fast computing algorithms, EM-IRLS and EM-coordinate descent are developed for expedited model fitting.
+EM-IRLS provides uncertainty measures, while EM-coordinate descent algorithm are extremely fast for model fitting.
+In the simulation studies, we shows the proposed model outperforms the state of the art GAM model, `mgcv` in prediction and function inference.
+Moreover, due to the model selection nature of the spike-and-slab prior, the proposed model can identify predictors that have large effects on the outcome.
+In addition, the model can be used to answer the question if non-linear effect are necessary.
+
+In the manuscript, we describe the implementation of ssGAM-NM using EM-IRLS.
+Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
+Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
+mixture t-distribution, and mixture double exponential distributions.
+@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors,
+$$
+\begin{aligned}
+\beta_i \sim N(0, \tau^2_i),\\
+\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
+\end{aligned}
+$$
+with $\tau^2$ integrated out analytically.
+Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
+One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
+$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
+The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
+[@Rockova2014a]

### 2021-06-15 Manuscript/SS_GAM.tex
Adding Content

@@
 
-\hypertarget{simulation}{%
-\section{Simulation}\label{simulation}}
+\hypertarget{simulation-study}{%
+\subsection{Simulation Study}\label{simulation-study}}
+
+In this section, we compare the performance of the proposed models to 3
+other models: COSSO, adaptive COSSO ,mgcv. COSSO is xxx method. Adaptive
+COSSO is another method, mgcv is another method. To demonstrate the
+correctness and feasibility of the proposed SS-GAM, we follow the
+logistic regression simulation introduced in Bai (Work in Progress). We
+first generate \(n=500\) training data sample with number of predictors
+\(p=4, 10, 50, 200\) respectively, where the predictors
+\(X \sim MVN_{n\times p}(0, I_{P})\). We simulate the outcome data from
+three distributions, gaussian, poisson, and binomial with the identity
+link, logit link \(g(x) = \log(\frac{x}{1-x})\), and log link
+\(g(x) = log(x)\) functions to mimic the common use of generalized
+model. The outcome is simulated via the following function \[
+\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
+\] for gaussian and binomial outcome and shrinked effects
+
+\[
+\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
+\] for stable poisson outcome. For the Gaussian outcome where dispersion
+are required, the dispersion parameter was set to 1. We follow the same
+data generating functions in Bai. In this data generating process, we
+have \(x_1, x_2, x_3, x_4\) as the active covariates, while the rest
+covariates are inactive \(f_j(x_j) = 0\) for \(j = 4, \dots, p\).
+Another set of independent sample of size
+\(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are created following the same
+data generating process, serving as the testing data. For mgcv and
+proposed models where choice of spline function are customizable, we
+choose to cubic regression spline with 10 basis. In COSSO and adaptive
+COSSO where the choice of spline are pre-determined, we set to have 10
+bases. The corresponding number of parameter that we estimate with the
+model is 37, 82, 451 and 1801. We are run the simulation where the
+number of bases are increased to 25 deliberately. To evaluate the
+performance of the models, mean squared error, mean absolute error for
+gaussian model, AUC and misclassifcation rate are calculated for
+binomial model, and deviance are calculated for Poisson model, averaging
+across50 simulations. The results are tabulated in Table 1, 2, 3.
+
+\%are calculated to assess out-of-sample estimation accuracy,
+
+\%We also calculated the area under the curve (AUC) of the receiver
+operating characteristic curve for \(y_{new}\). We will also visually
+display the estimated smooth function with the smooth function of truth
+to compare if the models can infer the smooth function correctly,
+despite the prediction accuracy.
+
+\%When constructing the smooth function of each variables, we constructs
+the natural cubic spline with degree of freedom 25. In other words, the
+design matrix of each smooth function will have 26 columns. The total
+number of columns for \(p=4, 10, 50\) is \(105, 261, 1301\)
+respectively, where intercept is kept in the model when fitting. The
+\(p=50\) can mimic the high dimensional setting as the design matrix are
+in high-dimensional setting (\(p >> n\))
+
+\%We benchmark the performance of the proposed methods with the state of
+art GAM package \texttt{mgcv}{[}TODO: add citation{]}. In the future
+steps, we will compare our proposed model with other high dimensional
+GAM applications.
+
+~
+
+\hypertarget{simulation-results}{%
+\subsubsection{Simulation Results}\label{simulation-results}}
+
+Overall, the proposed priors outperforms the state-of-art GAM model
+\texttt{mgcv} when the dimensionality is low, median, and high. The
+proposed priors predicts better than \texttt{mgcv} and infers better of
+the underlying smooth function. Most importantly, \texttt{mgcv} is
+infeasible in the high dimensional setting, i.e.~when \(p=50\) where the
+number of parameters are greater than the sample size.
+
+\hypertarget{prediction-performace}{%
+\paragraph{Prediction Performace}\label{prediction-performace}}
+
+The prediction performance is tabulated in the table below for AUC and
+Supplementary tables for MSE and misclassification rate. In summary, the
+proposed Bayesian models outperforms the benchmark model \texttt{mgcv}
+regardless the choice of prior or fitting algorithm. The models with
+IRLS fitting algorithm give slightly better performance than the the
+model with cyclic coordinate descent algorithm. The advantage diminishes
+in the high dimensional setting (p=50). This is possible because the SS
+normal spline prior provides a more smooth solution, which would work
+well in a low to median dimension setting. However, in the high
+dimensional setting, there are not enough sparsity in the model, i.e.~SS
+normal prior will provides a more complex model than necessary. This is
+confirmed with the visualization in the following section.
+
+\hypertarget{visualization}{%
+\paragraph{Visualization}\label{visualization}}
+
+We plot the estimated smooth functions of the four active variables for
+the 100 simulation iterations, along with the `true' smooth function.
+Across different settings of predictors, the patterns look similar.
+First of all, in all the settings, \texttt{mgcv} tends to provide more
+extreme models, where in the graphs the smooth function have much higher
+bound than the true smooth function. Rarely the estimated smooth
+function from \texttt{mgcv} matches with the true smooth function. In
+comparison, the proposed method can interpolate the shape of the true
+smooth function. However, different priors and fitting algorithms have
+different advantages. For example, in Figure 1, we see that the SS
+normal distribution have problem with the tail of the smooth function,
+especially obvious in the estimations of the variable \emph{x4}.
+However, it would have less sparse solutions, which is more likely to
+accurately interpolate the smooth function especially when the shape is
+complicated (see variable \emph{x2}). On the contrary, the SS double
+exponential prior model encourages sparse solutions, which can sometime
+omit active predictors.
 
@@
 effect. Such structure is more sensible analytically.
+
+\hypertarget{discussion-1}{%
+\subsection{Discussion}\label{discussion-1}}
+
+In this chapter, we propose the spike-and-slab generalized additive
+models for high-dimensional analysis.Two fast computing algorithms,
+EM-IRLS and EM-coordinate descent are developed for expedited model
+fitting. EM-IRLS provides uncertainty measures, while EM-coordinate
+descent algorithm are extremely fast for model fitting. In the
+simulation studies, we shows the proposed model outperforms the state of
+the art GAM model, \texttt{mgcv} in prediction and function inference.
+Moreover, due to the model selection nature of the spike-and-slab prior,
+the proposed model can identify predictors that have large effects on
+the outcome. In addition, the model can be used to answer the question
+if non-linear effect are necessary.
+
+In the manuscript, we describe the implementation of ssGAM-NM using
+EM-IRLS. Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
+Other spike-and-slab prior can be fitted with EM-IRLS as well,
+e.g.~mixture t-distribution, and mixture double exponential
+distributions. \citet{Yi2012} demonstrated that t prior and double
+exponential prior can be seen hierarchical normal priors, \[
+\begin{aligned}
+\beta_i \sim N(0, \tau^2_i),\\
+\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
+\end{aligned}
+\] with \(\tau^2\) integrated out analytically. Hence, the EM-IRLS can
+be easily applied to those two additional mixture priors. One of the
+advantage of using mixture \(t\) prior over mixture normal prior is that
+\(t\) distributions with low degree of freedom, e.g.~\(v=1\), have
+heavier tail than the normal distribution, which is also a special case
+of \(t\) distribution when \(v=\infty\). The heavier tails help to
+obtain posteriors that concentrated around the true nonzero coefficients
+at optimal rates. \citep{Rockova2014a}
 

### 2021-06-15 Real_Data/Emory_Card_Biobank/Report.Rmd
Adding Content

@@
 library(tidyverse)
+library(BhGLM)
 
@@
 
+## Screening
+### Linear Effect Screening
+```{r}
+# feature_name <- 
+linear_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
+  names() %>% 
+map_dfr(#feature_name[1:2],
+        .f = function(name, .dat){
+  # name <- feature_name[1]
+  y <- .dat %>% pull(death3yr)
+  x <- .dat %>% pull({{name}})
+  glm(y ~ x, family = binomial()) %>% 
+    tidy() %>% 
+    filter(term == "x") %>% 
+    select(p.value) %>% 
+    mutate(var = name,
+           p.value)
+
+},
+.dat = train_dat)
+
+lnr_scn_res <- data.frame(
+linear_screening_res,
+p.adj = p.adjust(linear_screening_res$p.value, "fdr")
+) %>% arrange(p.adj) %>% filter(p.adj<0.2) #%>% 
+  # slice_head(n = 100)
+```
+
+### GAM Screening
+```{r}
+
+library(mgcv)
+
+# feature_name <- 
+gam_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
+  names() %>% 
+map_dfr(#feature_name[1:2],
+        .f = function(name, .dat){
+  # name <- feature_name[1]
+  y <- .dat %>% pull(death3yr)
+  x <- .dat %>% pull({{name}})
+  mgcv::gam(y ~ s(x, bs = "cr", k = 5), family = binomial(),
+            data = data.frame(x, y)) %>% 
+    tidy() %>% 
+    filter(term == "s(x)") %>% 
+    select(p.value) %>% 
+    mutate(var = name,
+           p.value)
+
+},
+.dat = train_dat)
+
+gam_scn_res <- data.frame(
+gam_screening_res,
+p.adj = p.adjust(gam_screening_res$p.value, "fdr")
+) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
+  head(200)
+```
+
 # Predictive Model
@@
                     y = train_data %>% pull(death3yr), alpha = 1,
-                    family = "binomial",lambda = lasso_mdl_cv$lambda.1se)
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
 
@@
 
-measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% data.matrix,
+measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% as.matrix,
            new.y = test_data %>% pull(death3yr))
+
+lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+
+assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
+
+assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))
+
+```
+
+```{r}
+library(glmnet)
+
+lasso_mdl_cv <- cv.glmnet(x = train_data %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial", type.measure = "auc")
+
+lasso_mdl <- glmNet(x = train_data %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+
+# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
+measure.bh(lasso_mdl)
+# assess.glmnet(lasso_mdl$)
+
+measure.bh(lasso_mdl, new.x = test_data %>% select(gam_scn_res %>% pull(var)) %>% as.matrix,
+           new.y = test_data %>% pull(death3yr))
+
+lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+
+assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
+
+assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))
+
 ```
@@
 library(BHAM)
-library(BhGLM)
 
-mz_names <- train_dat %>% select(starts_with("mz")) %>% names  
+mz_names <- train_dat %>% select(gam_scn_res %>% pull(var)) %>% names  
 sm_df <- data.frame(
@@
 
-mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial")
+mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial", ss  = c(lasso_mdl$prior.scale, 0.5))
 
 mdl_dflt <- bmlasso_spline(dsn_mat, train_data$death3yr, family = "binomial",
-                           group = make_group(names(dsn_mat)))
+                           group = make_group(names(dsn_mat)),
+                           ss  = c(lasso_mdl$prior.scale, 0.5))
 
 # Training Data Performance
+cat("bmlasso model\n")
 measure.bh(mdl_defaul)
+cat("bmlasso_spline model\n")
 measure.bh(mdl_dflt)
 
+```
+
+```{r}
 test_sm_dat <- BHAM::make_predict_dat(sm_obj$Smooth, dat=test_data)
@@
 # Testing Data Performance
+cat("bmlasso model\n")
 measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=test_data$death3yr)
+cat("bmlasso_spline model\n")
 measure.bh(mdl_dflt, new.x = test_sm_dat, new.y=test_data$death3yr)

### 2021-06-15 _targets.R

@@ -37,11 +37,11 @@ list(
   tar_target(RD_ECB_valid_dat,
              readr::read_csv(realdata_ECB_valid_path)), 
   #** Report ####
-  # tar_render(real_data_report_ECB,
-  #            "Real_Data/Emory_Card_Biobank/Report.Rmd",
-  #            output_dir = "Real_Data/Emory_Card_Biobank/",
-  #            output_file = "ECB_report.html"),
-  # 
+  tar_render(real_data_report_ECB,
+             "Real_Data/Emory_Card_Biobank/Report.Rmd",
+             output_dir = "Real_Data/Emory_Card_Biobank/",
+             output_file = "ECB_report.html"),
+
   
   # Manuscript --------------------------------------------------------------
 

### 2021-06-21 Manuscript/03-Simulation.Rmd
Add more content

@@
 The corresponding number of parameter that we estimate with the model is 37, 82, 451 and 1801.
-We are run the simulation where the number of bases are increased to 25 deliberately.
-To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across50 simulations.
+We also run the simulation where the number of bases are increased to 25 deliberately.
+To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across 50 simulations.
 The results are tabulated in Table 1, 2, 3.
+
+```{r, fig.height=8}
+targets::tar_load(binom_plot)
+binom_plot
+```
+
+```{r, fig.height=8}
+targets::tar_load(gaussian_plot)
+gaussian_plot
+```
+
+```{r, fig.height=8}
+targets::tar_load(poisson_plot)
+poisson_plot
+```
 

### 2021-06-21 Manuscript/SS_GAM.tex
Add more content

@@
 bases. The corresponding number of parameter that we estimate with the
-model is 37, 82, 451 and 1801. We are run the simulation where the
+model is 37, 82, 451 and 1801. We also run the simulation where the
 number of bases are increased to 25 deliberately. To evaluate the
@@
 binomial model, and deviance are calculated for Poisson model, averaging
-across50 simulations. The results are tabulated in Table 1, 2, 3.
+across 50 simulations. The results are tabulated in Table 1, 2, 3.
+
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}
 

### 2021-06-21 Real_Data/Emory_Card_Biobank/Report.Rmd
Add more content

--- 2021-05-31 Real_Data/Emory_Card_Biobank/Report.Rmd
@@
 library(tidyverse)
+library(BhGLM)
+library(BHAM)
 
@@
 valid_dat <- targets::tar_read(RD_ECB_valid_dat)
+
+gam_scn_res <- targets::tar_read(train_gam_screen)
 ```
@@
 ### Covariate Summary
-```{r}
+```{r eval = F}
 list(first_set = train_dat,
@@
 
+## Screening
+### Linear Effect Screening
+```{r eval = F}
+# feature_name <- 
+linear_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
+  names() %>% 
+map_dfr(#feature_name[1:2],
+        .f = function(name, .dat){
+  # name <- feature_name[1]
+  y <- .dat %>% pull(death3yr)
+  x <- .dat %>% pull({{name}})
+  glm(y ~ x, family = binomial()) %>% 
+    tidy() %>% 
+    filter(term == "x") %>% 
+    select(p.value) %>% 
+    mutate(var = name,
+           p.value)
+
+},
+.dat = train_dat)
+
+lnr_scn_res <- data.frame(
+linear_screening_res,
+p.adj = p.adjust(linear_screening_res$p.value, "fdr")
+) %>% arrange(p.adj) %>% filter(p.adj<0.2) #%>% 
+  # slice_head(n = 100)
+```
+
+### GAM Screening
+```{r eval = F}
+
+library(mgcv)
+library(broom)
+
+# feature_name <- 
+gam_screening_res <- train_dat %>% select(starts_with("mz")) %>% 
+  names() %>% 
+map_dfr(#feature_name[1:2],
+        .f = function(name, .dat){
+  # name <- feature_name[1]
+  y <- .dat %>% pull(death3yr)
+  x <- .dat %>% pull({{name}})
+  mgcv::gam(y ~ s(x, bs = "cr", k = 5), family = binomial(),
+            data = data.frame(x, y)) %>% 
+    tidy() %>% 
+    filter(term == "s(x)") %>% 
+    select(p.value) %>% 
+    mutate(var = name,
+           p.value)
+
+},
+.dat = train_dat)
+
+gam_scn_res <- data.frame(
+gam_screening_res,
+p.adj = p.adjust(gam_screening_res$p.value, "fdr")
+) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
+  head(200)
+```
+
 # Predictive Model
@@
 
+```{r data_split, eval = F}
+library(tidymodels)
+
+set.seed(1)
+
+data_split <- rsample::initial_split(train_dat, prop = 3/4)
+
+train_data <- training(data_split)
+test_data <- testing(data_split)
+```
+
+## Baseline: Linear Lasso Model
+
+```{r eval = F}
+library(glmnet)
+
+train_data <- train_dat %>% select(gam_scn_res %>% pull(var), death3yr)
+
+lasso_mdl_cv <- cv.glmnet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial", type.measure = "auc")
+
+lasso_mdl <- glmNet(x = train_data %>% select(starts_with("mz")) %>% data.matrix,
+                    y = train_data %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+
+lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures
+
+# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
+# measure.bh(lasso_mdl)
+# assess.glmnet(lasso_mdl$)
+
+# measure.bh(lasso_mdl, new.x = test_data %>% select(starts_with("mz")) %>% as.matrix,
+#            new.y = test_data %>% pull(death3yr))
+
+# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
+#                     y = train_data %>% pull(death3yr), alpha = 1,
+#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+# 
+# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
+# 
+# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))
+
+```
+
+```{r}
+library(glmnet)
+
+lasso_mdl_cv <- cv.glmnet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
+                    y = train_dat %>% pull(death3yr), alpha = 1,
+                    family = "binomial", type.measure = "auc")
+
+lasso_mdl <- glmNet(x = train_dat %>% select(gam_scn_res %>% pull(var)) %>% data.matrix,
+                    y = train_dat %>% pull(death3yr), alpha = 1,
+                    family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+
+lasso_cv_res <- cv.bh(lasso_mdl, ncv = 10)$measures
+
+# class(lasso_mdl) <- c("glmNet", class(lasso_mdl))
+# measure.bh(lasso_mdl)
+# assess.glmnet(lasso_mdl$)
+
+# measure.bh(lasso_mdl, new.x = test_data %>% select(gam_scn_res %>% pull(var)) %>% as.matrix,
+#            new.y = test_data %>% pull(death3yr))
+
+# lasso_mdl <- glmnet(x = train_data %>% select(starts_with("mz")) %>% as.matrix,
+#                     y = train_data %>% pull(death3yr), alpha = 1,
+#                     family = "binomial",lambda = lasso_mdl_cv$lambda.min)
+# 
+# assess.glmnet(lasso_mdl, newx = train_data %>% select(starts_with("mz")) %>% as.matrix, newy = train_data %>% pull(death3yr))
+# 
+# assess.glmnet(lasso_mdl, newx = test_data %>% select(starts_with("mz")) %>% as.matrix, newy = test_data %>% pull(death3yr))
+
+```
+
 ```{r}
 library(BHAM)
-library(BhGLM)
 
-mz_names <- train_dat %>% select(starts_with("mz")) %>% names
-
-test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
-
+mz_names <- train_dat %>% select(gam_scn_res %>% pull(var)) %>% names  
 sm_df <- data.frame(
@@
  Func = "s",
- Args ="bs='cr', k=5"
+ Args ="bs='cr', k=10"
 )
@@
 sm_obj <- construct_smooth_data(sm_df, train_dat)
+```
+```{r}
+dsn_mat <- sm_obj$data
+```
 
-dsn_mat <- sm_obj$data
-
+```{r}
 # The validation data doesn't overlap with training data
@@
 
-mdl_defaul <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial")
+# mdl_defaul <- bmlasso(dsn_mat, train_data$death3yr, family = "binomial", ss  = c(, 0.5))
 
-mdl_dflt <- bmlasso_spline(dsn_mat, train_dat$death3yr, family = "binomial",
+bmlasso_mdl <- bmlasso(dsn_mat, train_dat$death3yr, family = "binomial",
+                           group = make_group(names(dsn_mat)))
+```
+
+```{r}
+spline_mdl <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
                            group = make_group(names(dsn_mat)))
 
-measure.bh(mdl_defaul)
+# bmlasso_cv_res <- cv.bh(bmlasso_mdl, ncv = 10)$measures
+# spline_cv_res <- cv.bh(spline_mdl, ncv = 10)$measures
 
-measure.bh(mdl_dflt)
+# Training Data Performance
+# cat("bmlasso model\n")
+# measure.bh(mdl_defaul)
+# cat("bmlasso_spline model\n")
+# measure.bh(mdl_dflt)
+
+```
+
+```{r eval = F}
+test_sm_dat <- BHAM::make_predict_dat(sm_obj$Smooth, dat=test_data)
+
+# Testing Data Performance
+cat("bmlasso model\n")
+measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=test_data$death3yr)
+cat("bmlasso_spline model\n")
+measure.bh(mdl_dflt, new.x = test_sm_dat, new.y=test_data$death3yr)
+```
+
+# Extremely Penalized Model
+```{r}
+
+mdl_pen_extreme <- bamlasso(dsn_mat, train_dat$death3yr, family = "binomial",
+                           group = make_group(names(dsn_mat)),
+                           ss = c(0.001, 0.5))
+
+# spline_ext_cv_res <- cv.bh(mdl_pen_extreme, ncv = 10)
+
+# Training Data Performance
+# measure.bh(mdl_pen_extreme)
+
+# Testing Data Performance
+# measure.bh(mdl_pen_extreme, new.x = test_sm_dat, new.y=test_data$death3yr)
 ```
@@
 ## Validation
+
+The validation data set have mostly distinct feature from the training data set. Hence, the validation in the traditional sense does not exists.
+
 ```{r eval = F}
 
+test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
 measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)

### 2021-06-21 R/functions.R

@@ -1,0 +1,169 @@
+#' Use GAM to screen the features
+#'
+#' @param dat the raw data
+#' @param nrow How many rows are selected
+#'
+#' @return
+#'
+gam_screen <- function(dat, nrow = 200){
+  dat %>% 
+    select(starts_with("mz")) %>% 
+    names() %>% 
+    map_dfr(#feature_name[1:2],
+      .f = function(name, .dat){
+        # name <- feature_name[1]
+        y <- .dat %>% pull(death3yr)
+        x <- .dat %>% pull({{name}})
+        mgcv::gam(y ~ s(x, bs = "cr", k = 10), family = binomial(),
+                  data = data.frame(x, y)) %>% 
+          tidy() %>% 
+          filter(term == "s(x)") %>% 
+          select(p.value) %>% 
+          mutate(var = name,
+                 p.value)
+        
+      },
+      .dat = dat) %>% data.frame(
+        .,
+        p.adj = p.adjust(.$p.value, "fdr")
+      ) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
+    head(nrow)
+}
+
+
+create_success_rate_table <- function(path){
+  sims <- list.files(path,full.names =TRUE)
+  
+  success_rate <- map_dfr(sims, .f = function(sim){
+    sim.df <- unglue_data(sim, 
+                          "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
+      mutate( p = as.numeric(p))
+    fls <- list.files(sim, full.names = TRUE) 
+    fls <- fls[grep(".rds", x=fls)]
+    n <- length(fls)
+    
+    data.frame(sim.df, n_success = n, path = sim)
+  }) %>% 
+    arrange(p) 
+
+  return(success_rate)  
+}
+
+
+make_sim_main_plots <- function(success_rate, dist, measures){
+  total_dat <- success_rate %>% 
+    # filter(dist==dist, n_success!=0) %>% 
+    filter(dist=={{dist}}, n_success!=0) %>%
+    pull(path) %>% 
+    map_dfr( .f = function(sim){
+      sim.df <- unglue_data(sim, 
+                            "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
+        mutate( p = as.numeric(p))
+      fls <- list.files(sim, full.names = TRUE) 
+      fls <- fls[grep(".rds", x=fls)]
+      n <- length(fls)
+      
+      
+      if(n == 0)
[…]

### 2021-06-21 _targets.R

@@ -1,6 +1,8 @@
 library(targets)
 library(tarchetypes)
 
+source("R/functions.R")
+
 # This is an example _targets.R file. Every
 # {targets} pipeline needs one.
 # Use tar_script() to create _targets.R and tar_edit()
@@ -10,18 +12,46 @@ library(tarchetypes)
 
 options(tidyverse.quiet = TRUE)
 # Set target-specific options such as packages.
-tar_option_set(packages = c("dplyr", "tidyverse", "rticles", "gtsummary", "here"))
+tar_option_set(packages = c("dplyr", "tidyverse", "rticles", "gtsummary", "here",
+                            "mgcv", "broom", "unglue", "knitr",
+                            "ggpubr"))
 
 # End this file with a list of target objects.
 list(
-
+  
   #  Simulations Studies -----------------------------------------------------
+  #* Main Study ####
+  #* 
+  #* 
+  tar_target(success_rate,
+             create_success_rate_table("Simulation/Main/Data/main_k=10/")),
+  
+  tar_target(binom_plot,
+             make_sim_main_plots(
+               success_rate,
+               "binomial",
+               list("test.deviance", "test.auc", "test.misclassification"))
+             ),
+
+  tar_target(gaussian_plot,
+             make_sim_main_plots(
+               success_rate,
+               "gaussian",
+               list("test.R2", "test.mse", "test.mae"))),
+
+  tar_target(poisson_plot,
+             make_sim_main_plots(
+               success_rate,
+               "poisson",
+               list("test.deviance", "test.mse", "test.mae"))
+             ),
+  
   #* Tuning Study ####
   tar_render(sim_tuning_report,
              "Simulation/tuning/tuning_study_report.Rmd",
              output_dir = "Simulation/tuning/",
              output_file = "sim_tuning_report.html"),
-    
+  
   
   #  Real Data Analysis -----------------------------------------------------
   #* Emory Card Biobank ####
@@ -36,15 +66,20 @@ list(
              readr::read_csv(realdata_ECB_train_path)),
   tar_target(RD_ECB_valid_dat,
              readr::read_csv(realdata_ECB_valid_path)), 
+  
+  #** GAM Screening ####
+  tar_target(train_gam_screen,
+             gam_screen(RD_ECB_train_dat)),
+  
   #** Report ####
   tar_render(real_data_report_ECB,
              "Real_Data/Emory_Card_Biobank/Report.Rmd",
[…]

### 2021-06-21 Manuscript/00-main.Rmd
1. Revise introduction section 2. Add some simulation results 3. Add coauthors 4. Bibfiles

@@
 - name: Boyi Guo*
+  num: a
+- name: D. Leann Long
+  num: a
+- name: Byron C. Jaeger
+  num: a 
+- name: AKM Fazlur Rahman
   num: a   

### 2021-06-21 Manuscript/01-Intro.Rmd
Improved the manuscript
1. Revise introduction section 2. Add some simulation results 3. Add coauthors 4. Bibfiles

@@
 
-* What is the problem
-  * what is the GAM
-  * expanding GAM in to high-dimensional frame work
-  * Why this is important
-  * Why this is hard
-* Previous solutions
-  * What are the methods
-  * What are their problems
+Using additive models to interpolate nonlinear effects has a long history.
+Early work of additive models can date back to 1979 when @Cleveland1979 proposed local smoothing for bivariate analysis.
+Since then, many mathematical and statistical models have been proposed, see @Wood2020.
+With the recent advancement of computational technology, many machine learning algorithms that are computationally intensive start to gain much popularity in prognostic modeling and functional selection, especially when complex signals, for example nonlinear effects, are assumed.
+Such models generally have improved prediction accuracy over classic statistical models.
+However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models.
+Given that the primary objective of biomedical studies is to understand the etiology of diseases, "Black box" models are less helpful as they can hardly explain how the change of biomarkers is associated with the disease.
 
-* Introduce the wiggliness matrix
-
-* Our method
-* The implication of our method
-* Fast
-* Structure of the rest of the paper
-
-Nonlinear effect modeling has long history.
-Early work of GAMs can date back to 1979 when @Cleveland1979 proposed local smoothing for bivariate analysis.
-Since then, many mathematical and statistical models has been proposed, see xxx^[TODO: insert an review citation].
-With recent advancement of computing technology, many machine learning algorithms that is computationally intensive start to gainmuch popularity in outcome prediction when non-linear effects assumed.
-Such models generally have improved prediction accuracy over classic statistical models, but have a fatal flaw, lack of transparency and interpretability.
-The primary objective of biomedical studies is to understand the etiology of diseases, which requires not only the knowledge if a biomarker is predictive of the disease, but also how the biomarker cause the disease.
-"Black box" models could less likely to be helpful.
-
-As a balance between interpretability and model flexibility, semi-parametric and non-parametric regression models are studied.
-Among those, generalized additive model (GAM) proposed in the seminal work of @Hastie1987 gains the most popularity thanks to its flexibility and easiness to incorporate many different smoothing functions.
-A response variable, $Y$, that is assumed to follow some exponential family distribution $EF(\mu.
-\phi)$, can be expressed as the summation of smooth functions of a given $p$-dimensional vector of covariates $\bs x$, written as
+Semi-parametric regression models are a more appropriate alternative to the "black-box" models, thanks to their balancing between model flexibility and interpretability.
+Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most utilized modeling tools.
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu.
+\phi)$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
 $$
 where $g^{-1}$ is the inverse function of a smooth monotonic link function.
-GAM solves the model fitting problem for many distributional outcomes, such as Poisson, binomial, gamma and normal models.
-In the same article, @@Hastie1987 proposed the backfitting algorithm that estimates each smooth function of an additive model iteratively using partial residuals.
-The backfitting algorithm is compatible with almost any smoothing or modeling technique to be used as the nonparametric components.
-However, it is unclear how to estimate the degree of smoothness when necessary, for example for smoothing splines.
-The computation cost for cross-validating the smooth parameters can be expensive.
-@Wood2017[PP.318-320] provides more discussion on the problem.
+The smooth functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
+Nevertheless, the classic GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
 
-In many area of modern biomedical research, dealing with high-dimensional data, also known as $p >> n$ problem, is common,  for example sequencing data analysis, electric health record data analysis.
+Many applications of modern biomedical research, e.g.
+sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
+There exists an extensive literature on high-dimensional linear models, via penalized models or Bayesian hierarchical models, see @Mallick2013.
+[TODO a sentence of sparsity.
+What it means in a penalized model and Bayesian model] Directly apply to additive models, spline models specifically.
+Problem, oversmoothing of the spline function.
+An more appropriate treatment would be incorporating the smooth penalty of spline functions.[TODO: a sentence about smoothness penalty]
+
+What is the smooth penalty of spline function?
+
 Many statistical models have been proposed for such problem, for example, penalized models and Bayesian hierarchical models.

### 2021-06-21 Manuscript/03-Simulation.Rmd
1. Revise introduction section 2. Add some simulation results 3. Add coauthors 4. Bibfiles

@@
-## Simulation Study
+# Simulation Study
 In this section, we compare the performance of the proposed models to 3 other models: COSSO, adaptive COSSO ,mgcv. COSSO is xxx method. Adaptive COSSO is another method, mgcv is another method. 
@@
 ```
+
+For the binomial outcome shown in Figure 1, we can see that the proposed models are consistently performed better than COSSO and adaptive COSSO.
+Mgcv slightly outperforms the proposed models in low dimension settings(p = 4, 10), but suffers from overfitting in medium dimension setting (p=50).
+In addition, in the high-dimension setting (p=200), mgcv could not make prediction due to the p > n problem.
+Among the proposed models, we see that there is modest difference between the models that shares the same indicator variable('_spline' models) and or not sharing indicator variable (modles), where the not sharing indicator variable models have slightly edge over indicator sharing model.
+Moreover, the mixture double exponential model and lasso model perform substaintially better in the medium and high settings.
+For the remaining two outcomes, the results shows similar trends, where mixture double exponential model and ss lasso model are recommended for use.
 

### 2021-06-21 Manuscript/SS_GAM.tex
1. Revise introduction section 2. Add some simulation results 3. Add coauthors 4. Bibfiles

@@
 \author[a]{Boyi Guo*}
+\author[a]{D. Leann Long}
+\author[a]{Byron C. Jaeger}
+\author[a]{AKM Fazlur Rahman}
 \author[a]{Nengjun Yi}
@@
 
-Nonlinear effect modeling has long history. Early work of GAMs can date
-back to 1979 when \citet{Cleveland1979} proposed local smoothing for
-bivariate analysis. Since then, many mathematical and statistical models
-has been proposed, see xxx\footnote{TODO: insert an review citation}.
-With recent advancement of computing technology, many machine learning
-algorithms that is computationally intensive start to gainmuch
-popularity in outcome prediction when non-linear effects assumed. Such
-models generally have improved prediction accuracy over classic
-statistical models, but have a fatal flaw, lack of transparency and
-interpretability. The primary objective of biomedical studies is to
-understand the etiology of diseases, which requires not only the
-knowledge if a biomarker is predictive of the disease, but also how the
-biomarker cause the disease. ``Black box'' models could less likely to
-be helpful.
-
-As a balance between interpretability and model flexibility,
-semi-parametric and non-parametric regression models are studied. Among
-those, generalized additive model (GAM) proposed in the seminal work of
-\citet{Hastie1987} gains the most popularity thanks to its flexibility
-and easiness to incorporate many different smoothing functions. A
-response variable, \(Y\), that is assumed to follow some exponential
-family distribution \(EF(\mu. \phi)\), can be expressed as the summation
-of smooth functions of a given \(p\)-dimensional vector of covariates
-\(\boldsymbol{x}\), written as \[
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),\qquad E[f_j(x_j)] = 0.
+Using additive models to interpolate nonlinear effects has a long
+history. Early work of additive models can date back to 1979 when
+\citet{Cleveland1979} proposed local smoothing for bivariate analysis.
+Since then, many mathematical and statistical models have been proposed,
+see \citet{Wood2020}. With the recent advancement of computational
+technology, many machine learning algorithms that are computationally
+intensive start to gain much popularity in prognostic modeling and
+functional selection, especially when complex signals, for example
+nonlinear effects, are assumed. Such models generally have improved
+prediction accuracy over classic statistical models. However, these
+models have a fatal flaw, the lack of transparency and interpretability,
+and hence are referred to as ``black box'' models. Given that the
+primary objective of biomedical studies is to understand the etiology of
+diseases, ``Black box'' models are less helpful as they can hardly
+explain how the change of biomarkers is associated with the disease.
+
+Semi-parametric regression models are a more appropriate alternative to
+the ``black-box'' models, thanks to their balancing between model
+flexibility and interpretability. Among those, generalized additive
+models (GAMs), proposed in the seminal work of \citet{Hastie1987}, grew
+to be one of the most utilized modeling tools. In a GAM, the response
+variable, \(Y\), which is assumed to follow some exponential family
+distribution \(EF(\mu. \phi)\), can be expressed as the summation of
+smooth functions, \(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional
+vector of covariates \(\boldsymbol{x}\), written as \[
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
 \] where \(g^{-1}\) is the inverse function of a smooth monotonic link
-function. GAM solves the model fitting problem for many distributional
-outcomes, such as Poisson, binomial, gamma and normal models. In the
-same article, @\citet{Hastie1987} proposed the backfitting algorithm
-that estimates each smooth function of an additive model iteratively
-using partial residuals. The backfitting algorithm is compatible with
-almost any smoothing or modeling technique to be used as the
-nonparametric components. However, it is unclear how to estimate the
-degree of smoothness when necessary, for example for smoothing splines.
-The computation cost for cross-validating the smooth parameters can be
-expensive. \citet[PP.318-320]{Wood2017} provides more discussion on the
-problem.
+function. The smooth functions can take many forms and are estimated
+using a pseudo-weighted version of the backfitting algorithm
+\citep{Breiman1985}. Nevertheless, the classic GAMs cannot fulfill the
+increasing analytic demands for high-dimensional data analysis in
+biomedical studies.
 
@@
 \hypertarget{simulation-study}{%
-\subsection{Simulation Study}\label{simulation-study}}
+\section{Simulation Study}\label{simulation-study}}
 
@@
 
+For the binomial outcome shown in Figure 1, we can see that the proposed
+models are consistently performed better than COSSO and adaptive COSSO.
+Mgcv slightly outperforms the proposed models in low dimension
+settings(p = 4, 10), but suffers from overfitting in medium dimension
+setting (p=50). In addition, in the high-dimension setting (p=200), mgcv
+could not make prediction due to the p \textgreater{} n problem. Among
+the proposed models, we see that there is modest difference between the
+models that shares the same indicator variable('\_spline' models) and or
+not sharing indicator variable (modles), where the not sharing indicator
+variable models have slightly edge over indicator sharing model.
+Moreover, the mixture double exponential model and lasso model perform
+substaintially better in the medium and high settings. For the remaining
+two outcomes, the results shows similar trends, where mixture double
+exponential model and ss lasso model are recommended for use.
+
 \%are calculated to assess out-of-sample estimation accuracy,

### 2021-06-22 Manuscript/00-main.Rmd
Improved the the itnroduction part

--- 2021-06-02 Manuscript/00-main.Rmd
@@
 - name: Boyi Guo*
+  num: a
+- name: Byron C. Jaeger
+  num: a 
+- name: AKM Fazlur Rahman
+  num: a   
+- name: D. Leann Long
   num: a  
@@
 presentaddress: This is sample for present address text this is sample for present address text
-authormark: Guo and Yi
+authormark: Guo et al.
 articletype: Research article
@@
 In that case, it is possible to revert to standard \LaTeX syntax.
-
-# References
-
-# Appendix

### 2021-06-22 Manuscript/01-Intro.Rmd
Improved the the itnroduction part

@@
 Using additive models to interpolate nonlinear effects has a long history.
-Early work of additive models can date back to 1979 when @Cleveland1979 proposed local smoothing for bivariate analysis.
+Early work of additive models can date back to 1979 when \citet{Cleveland1979} proposed local smoothing for bivariate analysis.
 Since then, many mathematical and statistical models have been proposed, see @Wood2020.
@@
 sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
-There exists an extensive literature on high-dimensional linear models, via penalized models or Bayesian hierarchical models, see @Mallick2013.
+There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see @Mallick2013.
+These models assume sparsity among the predictors, i.e.
+most predictors have small or no effects on the outcome.
+Further development focuses on addressing the grouping structure among the predictors, where AMs and GAMs can be considered special cases.
+Various penalties are imposed on the basis function coefficients: @Ravikumar2009 extended the grouped lasso [@Yuan2006] to additive models; @Huang2010 further developed adaptive grouped lasso for additive models; @Wang2007 and @Xue2009 applied grouped SCAD penalty [@Fan2001] to the additive models.
+Recent Bayesian hierarchical models are also used in the context of high-dimensional additive models.
+Various group spike-and-slab priors combining with computationally intensive MCMC algorithms [@Xu2015; @Yang2020] are proposed where the application on AMs are by-products.
+@Bai2020 was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs in @Bai2021.
+Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.
+In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.
+[@Scheipl2013]
+
+To address this problem.
+
+where majority of the efforts ware focusing on the
+
 [TODO a sentence of sparsity.

### 2021-06-22 Manuscript/SS_GAM.tex
Improved the the itnroduction part

@@
 \author[a]{Boyi Guo*}
-\author[a]{D. Leann Long}
 \author[a]{Byron C. Jaeger}
 \author[a]{AKM Fazlur Rahman}
+\author[a]{D. Leann Long}
 \author[a]{Nengjun Yi}
 
-\authormark{Guo and Yi}
+\authormark{Guo et al.}
 
@@
 
-In many area of modern biomedical research, dealing with
-high-dimensional data, also known as \(p >> n\) problem, is common, for
-example sequencing data analysis, electric health record data analysis.
+Many applications of modern biomedical research, e.g.~sequencing data
+analysis, electric health record data analysis, require special
+treatment of high-dimensionality, commonly known as \(p >> n\) problem.
+There is extensive literature on high-dimensional linear models via
+penalized models or Bayesian hierarchical models, see
+\citet{Mallick2013}. These models assume sparsity among the predictors,
+i.e.~most predictors have small or no effects on the outcome. Further
+development focuses on addressing the grouping structure among the
+predictors, where AMs and GAMs can be considered special cases. Various
+penalties are imposed on the basis function coefficients:
+\citet{Ravikumar2009} extended the grouped lasso \citep{Yuan2006} to
+additive models; \citet{Huang2010} further developed adaptive grouped
+lasso for additive models; \citet{Wang2007} and \citet{Xue2009} applied
+grouped SCAD penalty \citep{Fan2001} to the additive models. Recent
+Bayesian hierarchical models are also used in the context of
+high-dimensional additive models. Various group spike-and-slab priors
+combining with computationally intensive MCMC algorithms
+\citep{Xu2015, Yang2020} are proposed where the application on AMs are
+by-products. \citet{Bai2020} was the first to apply group spike-and-slab
+lasso prior to AMs using a fast optimization algorithm, and further
+generalized the framework to GAMs in \citet{Bai2021}. Regardless of the
+nice statistical properties, a practical and pivotal question, if
+non-linear effects exist, yet to be addressed by these methods. In
+addition, a common criticism received by these methods is that the basis
+function coefficients tend to be overly penalized due to the sparsity
+assumption directly applied on the basis function coefficients.
+\citep{Scheipl2013}
+
+To address this problem.
+
+where majority of the efforts ware focusing on the
+
+{[}TODO a sentence of sparsity. What it means in a penalized model and
+Bayesian model{]} Directly apply to additive models, spline models
+specifically. Problem, oversmoothing of the spline function. An more
+appropriate treatment would be incorporating the smooth penalty of
+spline functions.{[}TODO: a sentence about smoothness penalty{]}
+
+What is the smooth penalty of spline function?
+
 Many statistical models have been proposed for such problem, for
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
-
 \bibliography{bibfile.bib}

### 2021-06-22 Manuscript/01-Intro.Rmd
Improved introduction

@@
 Recent Bayesian hierarchical models are also used in the context of high-dimensional additive models.
-Various group spike-and-slab priors combining with computationally intensive MCMC algorithms [@Xu2015; @Yang2020] are proposed where the application on AMs are by-products.
+Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed where the application on AMs are by-products.
+@Scheipl2012 proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma, specifically for semi-parametric regression using MCMC.
+The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
 @Bai2020 was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs in @Bai2021.
 Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.
-In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.
-[@Scheipl2013]
+In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.[@Scheipl2013] <!-- It is still not clear what is the best practice on how to model non-linear effects in high-dimensional setting.
+-->
 
-To address this problem.
+To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data, where the smooth functions are imposed with a new spike-and-slab spline prior.
+Two fast computing optimization based algorithms are developed for high and ultra-high dimension settings respectively.
+The proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using mixture double exponential distribution.
+To avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
+Smoothness penalties are commonly implemented in GAMs through smoothing and penalized regression splines.
+They are quadratic norm of the coefficients, and allow locally adaptive penalties on each smoothing function.
+For the most common case of cubic splines, the smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of second derivative of the spline function,  expressed mathematically as
+$$
+  \text{pen}f_j(x) = \lambda_j \int f^{"}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j,
+$$
+where $\bs S_j$ is a known penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
+Smoothing penalties are also previously used in spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
+Moreover, incorporating the smoothness penalty allows the separation of the linear space of the smooth function, also know as the null space, from the nonlinear space.
+[TODO: descrive wood.]
 
-where majority of the efforts ware focusing on the
+Describe the differences from previous models SS GAM and Bai, taling about the computational burden.
 
-[TODO a sentence of sparsity.
-What it means in a penalized model and Bayesian model] Directly apply to additive models, spline models specifically.
-Problem, oversmoothing of the spline function.
-An more appropriate treatment would be incorporating the smooth penalty of spline functions.[TODO: a sentence about smoothness penalty]
+Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e.
+either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect.
 
-What is the smooth penalty of spline function?
-
-Many statistical models have been proposed for such problem, for example, penalized models and Bayesian hierarchical models.
-However, up to this point, most efforts of methodology development focus on modeling linear effect in high dimensional setting.
-It is still not clear what is the best practice on how to model non-linear effects in high-dimensional setting.
-
-Similar to expanding classic models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
+<!-- Similar to expanding classic models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
 The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.
@@
 How to organically combine the two together remains as a troubling question.
+-->
 
-To the best of our knowledge, there are few GAM models proposed for high dimensional application.
+<!-- To the best of our knowledge, there are few GAM models proposed for high dimensional application.
 In common practice, model matrix of spline terms are generated using base functions of preference, and then feed to a penalized model.
@@
 The authors also demonstrated its similarity to Group Lasso.
-@Meier2009 proposed an a sparsity-smoothness penalty,
+@Meier2009 proposed an a sparsity-smoothness penalty, -->
 
-$$
-J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
-$$
-which can re-formulated as a group lasso problem as well.
+<!-- $$ -->
+<!-- J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx}, -->
+<!-- $$ -->
+<!-- which can re-formulated as a group lasso problem as well.
 However, the model can not estimate the smoothness of the spline functions correctly, as the smoothness of all spline models are controlled by the lambda_2.
 The inflexible of the model creates problems in prediction.
-
-From the Bayesian perspective, recent works that expands to the high-dimensional settings includes @Scheipl2012 and @Bai2020Spline.
-@Scheipl2012 proposed an new spike-and-slab structure prior, normal-mixture-of-inverse gamma for semi-parametric regression using MCMC.
-The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
-@Bai2020Spline proposed an novel high dimensional Bayesian GAM using the spike-and-slab group lasso prior.
-An fast fitting EM-based algorithm is proposed.
-Asymptotic theory, such as posterior contraction rates and model selection consistency are proved.
-Even though, the algorithm is flexible to cover many non-Gaussian family in a high dimensional setting, the proposed method didn't address uncertainty quantification, i.e.
-calculating credible intervals etc.
-However, most of these Bayesian GAMs still suffer a common problem, computation cost.
-These Bayesian GAMs uses MCMC to estimate the posterior distribution, which would create an intense computation load.
+-->
 
@@
 
-In this article, we aim to address the challenges of non-linear signal modeling in high-dimensional application using Bayesian Hierarchical model.
-Specifically, we will propose a new Bayesian spike-and-slab prior for the generalized additive models and Cox spline models.
-The contribution of the proposed prior comes in two-fold: 1) it supports simultaneous variable selection and function selection; 2) it provides robust estimation and prediction for the coefficients and outcomes.
-As previously described, simultaneous variable selection and its functional form selection is not a trivial problem.
-Nevertheless, the Bayesian spike-and-slab prior is a perfect solution to the problem, as the latent binary indicator can be used to for variable selection and the spike and slab densities are used to adaptive adjust the smoothness of the spline function.
-Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e.
-either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect.
-This simplifies the process of determining what effect a predictor have on the outcome, without involving hypothesis testing where the degree of freedom for non-linear effect is complicated to estimate due to the shrinkage.
-Moreover, out model inherent the the robust estimation from Bayesian models.
 Lastly, The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
@@
 
-The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian Hierarchical generalized additive model, where we provide the setup of the proposed model and two fast-fitting algorithms.
-In Chapter 3, we Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
+The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian hierarchical generalized additive model, where we provide the setup of the proposed model and describe the two fast-fitting algorithms.
+In Chapter 3, we compare the proposed via monte carlo simulation.
+.Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
 Conclusion and discussions are made in Chapter 5.

### 2021-06-23 Manuscript/01-Intro.Rmd
More introduction writting

@@
 Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.
-In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.[@Scheipl2013] <!-- It is still not clear what is the best practice on how to model non-linear effects in high-dimensional setting.
--->
+Besides @Schipel2012, most proposed methods adapt an 'all-in-all-out' strategy, i.e.
+either including or excluding the variable completely, rendering no space for within group selection.
+In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.[@Scheipl2013] It would be of special interest to develop a fast yet flexible generalized additive model.
 
-To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data, where the smooth functions are imposed with a new spike-and-slab spline prior.
-Two fast computing optimization based algorithms are developed for high and ultra-high dimension settings respectively.
-The proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using mixture double exponential distribution.
-To avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
+To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
+Specifically, we impose a new bi-part spike-and-slab spline prior on the smooth functions to perform bilevel selection, where the linear and nonlinear spaces of smooth functions can be selected separately.
+The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
+Additionally, two fast computing optimization based algorithms are developed for high and ultra-high dimension settings respectively.
+Meanwhile, to avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
 Smoothness penalties are commonly implemented in GAMs through smoothing and penalized regression splines.
@@
 where $\bs S_j$ is a known penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
-Smoothing penalties are also previously used in spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
+Smoothness penalties are also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
 Moreover, incorporating the smoothness penalty allows the separation of the linear space of the smooth function, also know as the null space, from the nonlinear space.
-[TODO: descrive wood.]
+@Marra2011 introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
+The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
+[TODO: a sentence that it decompose the prior to independent priors, provide obvious computational convenience.] The computational convenience improved the scalability of the model.
 
-Describe the differences from previous models SS GAM and Bai, taling about the computational burden.
+The proposed BHAM uses a spike-and-slab prior for function selections, but it substantially differs from previous spike-and-slab based GAMs, i.e.
+the spike-and-slab GAM [@Scheipl2012] and the SB GAM [@Bai2021] in many ways.
+First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM.
+Spike-and-slab lasso priors provide computational convenience during model fitting through optimization algorithms, instead of computational intensive sampling algorithms.
+They make fitting high- and ultrahigh-dimensional model feasible without sacrificing performance in prediction and variable selection.
+Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent algorithm and EM-Iterative weighted least square algorithm for different utilities.
+The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis, see [TODO: add citations].
+Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.
+EM-IWLS complements the lack of inference potential from EM-CD in a median and high data setting.
+Furthermore, the proposed model addresses the incapability of bilevel selection in SB-GAM.
+Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), make transnational science more accessible.
 
-Moreover, compared to currently available Bayesian spline models which mostly conduct a 'all-in-all-out' strategy, i.e.
-either including a non-linear form of a variable or excluding the variable completely, we offer a more flexible solution, rending one of three possibilities: no effect of a predictor, only linear effect of a predictor, or non-linear effect.
+The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian hierarchical generalized additive model, where we provide the setup of the proposed model and describe the two fast-fitting algorithms.
+In Chapter 3, we compare the proposed via monte carlo simulation.
+.Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
+Conclusion and discussions are made in Chapter 5.
 
@@
 
-Lastly, The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
+<!-- Lastly, The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
 We will implement the proposed models and algorithms in one of the most commonly used statistical computing environment R.
-Ancillary functions will be provided in assistance to model diagnostic, interpretation and graphing.
 The R implementation of the Bayesian hierarchical additive model is available via a freely available R package, BHAM at (https://github.com/boyiguo1/BHAM)[https://github.com/boyiguo1/BHAM].
-
-The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian hierarchical generalized additive model, where we provide the setup of the proposed model and describe the two fast-fitting algorithms.
-In Chapter 3, we compare the proposed via monte carlo simulation.
-.Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
-Conclusion and discussions are made in Chapter 5.
+-->

### 2021-06-23 Manuscript/SS_GAM.tex
More introduction writting

@@
 high-dimensional additive models. Various group spike-and-slab priors
-combining with computationally intensive MCMC algorithms
-\citep{Xu2015, Yang2020} are proposed where the application on AMs are
-by-products. \citet{Bai2020} was the first to apply group spike-and-slab
-lasso prior to AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs in \citet{Bai2021}. Regardless of the
-nice statistical properties, a practical and pivotal question, if
-non-linear effects exist, yet to be addressed by these methods. In
-addition, a common criticism received by these methods is that the basis
-function coefficients tend to be overly penalized due to the sparsity
-assumption directly applied on the basis function coefficients.
-\citep{Scheipl2013}
-
-To address this problem.
-
-where majority of the efforts ware focusing on the
-
-{[}TODO a sentence of sparsity. What it means in a penalized model and
-Bayesian model{]} Directly apply to additive models, spline models
-specifically. Problem, oversmoothing of the spline function. An more
-appropriate treatment would be incorporating the smooth penalty of
-spline functions.{[}TODO: a sentence about smoothness penalty{]}
-
-What is the smooth penalty of spline function?
-
-Many statistical models have been proposed for such problem, for
-example, penalized models and Bayesian hierarchical models. However, up
-to this point, most efforts of methodology development focus on modeling
-linear effect in high dimensional setting. It is still not clear what is
-the best practice on how to model non-linear effects in high-dimensional
-setting.
-
-Similar to expanding classic models to high-dimensional models,
-expanding the GAM framework for high-dimensional applications is not a
-trivial task. The difficulties mainly come in two-fold: grouping nature
-of the basis functions, and the balance between spline smoothness and
-variable sparsity. For GAM models, a smoothing function is consists of
-multiple basis functions. The basis functions of a variable is
-inherently related. The grouping structure should be properly taken care
-of for accurate estimation. Moreover, shrinkage is previously used
-solely for sparsity of the model in high dimensional regression models,
-or smoothness in GAMs. How to organically combine the two together
-remains as a troubling question.
-
-To the best of our knowledge, there are few GAM models proposed for high
-dimensional application. In common practice, model matrix of spline
-terms are generated using base functions of preference, and then feed to
-a penalized model. Without proper handle of the grouping nature of bases
-functions, the estimation of the smooth function is less accurate.
-\citet{Bai2020Spline} shows inferior performance of such practice
-compared to using an ad hoc high-dimensional additive model. Among the
-specialized high-dimensional spline methods, \citet{Ravikumar2009}
-extended the original backfitting algorithm \citep{Hastie1987} to
-accommodate the high-dimensional setting. The sparsity and smoothness
-was handled separately with a soft threshold. The authors also
-demonstrated its similarity to Group Lasso. \citet{Meier2009} proposed
-an a sparsity-smoothness penalty,
-
-\[
-J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx},
-\] which can re-formulated as a group lasso problem as well. However,
-the model can not estimate the smoothness of the spline functions
-correctly, as the smoothness of all spline models are controlled by the
-lambda\_2. The inflexible of the model creates problems in prediction.
-
-From the Bayesian perspective, recent works that expands to the
-high-dimensional settings includes \citet{Scheipl2012} and
-\citet{Bai2020Spline}. \citet{Scheipl2012} proposed an new
-spike-and-slab structure prior, normal-mixture-of-inverse gamma for
-semi-parametric regression using MCMC. The normal-mixture-of-inverse
-gamma prior is mixture of t-distribution with an additional coefficients
-mixing vector, where the vector follows a mixture of two Gaussian
-distribution. \citet{Bai2020Spline} proposed an novel high dimensional
-Bayesian GAM using the spike-and-slab group lasso prior. An fast fitting
-EM-based algorithm is proposed. Asymptotic theory, such as posterior
-contraction rates and model selection consistency are proved. Even
-though, the algorithm is flexible to cover many non-Gaussian family in a
-high dimensional setting, the proposed method didn't address uncertainty
-quantification, i.e.~calculating credible intervals etc. However, most
-of these Bayesian GAMs still suffer a common problem, computation cost.
-These Bayesian GAMs uses MCMC to estimate the posterior distribution,
-which would create an intense computation load.
-
-In this article, we aim to address the challenges of non-linear signal
-modeling in high-dimensional application using Bayesian Hierarchical
-model. Specifically, we will propose a new Bayesian spike-and-slab prior
-for the generalized additive models and Cox spline models. The
-contribution of the proposed prior comes in two-fold: 1) it supports
-simultaneous variable selection and function selection; 2) it provides
-robust estimation and prediction for the coefficients and outcomes. As
-previously described, simultaneous variable selection and its functional
-form selection is not a trivial problem. Nevertheless, the Bayesian
-spike-and-slab prior is a perfect solution to the problem, as the latent
-binary indicator can be used to for variable selection and the spike and
-slab densities are used to adaptive adjust the smoothness of the spline
-function. Moreover, compared to currently available Bayesian spline
-models which mostly conduct a `all-in-all-out' strategy, i.e.~either
-including a non-linear form of a variable or excluding the variable
-completely, we offer a more flexible solution, rending one of three
-possibilities: no effect of a predictor, only linear effect of a
-predictor, or non-linear effect. This simplifies the process of
-determining what effect a predictor have on the outcome, without
-involving hypothesis testing where the degree of freedom for non-linear
-effect is complicated to estimate due to the shrinkage. Moreover, out
-model inherent the the robust estimation from Bayesian models. Lastly,
-The third aim of the dissertation work will focus on the software
-implementation of the proposed models and algorithms. We will implement
-the proposed models and algorithms in one of the most commonly used
-statistical computing environment R. Ancillary functions will be
-provided in assistance to model diagnostic, interpretation and graphing.
-The R implementation of the Bayesian hierarchical additive model is
-available via a freely available R package, BHAM at
-(\url{https://github.com/boyiguo1/BHAM}){[}\url{https://github.com/boyiguo1/BHAM}{]}.
+combining with computationally intensive Markov chain Monte Carlo (MCMC)
+algorithms \citep{Xu2015, Yang2020} are proposed where the application
+on AMs are by-products. \citet{Scheipl2012} proposed an ad hoc
+spike-and-slab structure prior, normal-mixture-of-inverse gamma,
+specifically for semi-parametric regression using MCMC. The
+normal-mixture-of-inverse gamma prior is mixture of t-distribution with
+an additional coefficients mixing vector, where the vector follows a
+mixture of two Gaussian distribution. \citet{Bai2020} was the first to
+apply group spike-and-slab lasso prior to AMs using a fast optimization
+algorithm, and further generalized the framework to GAMs in
+\citet{Bai2021}. Regardless of the nice statistical properties, a
+practical and pivotal question, if non-linear effects exist, yet to be
+addressed by these methods. Besides \citet{Schipel2012}, most proposed
+methods adapt an `all-in-all-out' strategy, i.e.~either including or
+excluding the variable completely, rendering no space for within group
+selection. In addition, a common criticism received by these methods is
+that the basis function coefficients tend to be overly penalized due to
+the sparsity assumption directly applied on the basis function
+coefficients.\citep{Scheipl2013} It would be of special interest to
+develop a fast yet flexible generalized additive model.
+
+To address these challenges, we propose a novel Bayesian hierarchical
+generalized additive model (BHAM) for high dimensional data analysis.
+Specifically, we impose a new bi-part spike-and-slab spline prior on the
+smooth functions to perform bilevel selection, where the linear and
+nonlinear spaces of smooth functions can be selected separately. The
+prior setup encourages a flexible solution, rending one of three
+possibilities: no effect, only linear effect, or nonlinear effect of a
+predictor. Additionally, two fast computing optimization based
+algorithms are developed for high and ultra-high dimension settings
+respectively. Meanwhile, to avoid the over-penalization of the basis
+function coefficients, smoothness penalties are incorporated in the
+model via re-parameterization. Smoothness penalties are commonly
+implemented in GAMs through smoothing and penalized regression splines.
+They are quadratic norm of the coefficients, and allow locally adaptive
+penalties on each smoothing function. For the most common case of cubic
+splines, the smoothness penalty conditioning on a smoothness penalty
+parameter \(\lambda_j\) is a function of the integration of second
+derivative of the spline function, expressed mathematically as \[
+  \text{pen}f_j(x) = \lambda_j \int f^{"}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j,
+\] where \(\boldsymbol{S}_j\) is a known penalty matrix, and
+\(\boldsymbol{\beta}_j\) are the basis function coefficients. Smoothness
+penalties are also previously used in the spike-and-slab GAM
+\citep{Scheipl2012} and the sparsity-smoothness penalty
+\citep{Meier2009}. Moreover, incorporating the smoothness penalty allows
+the separation of the linear space of the smooth function, also know as
+the null space, from the nonlinear space. \citet{Marra2011} introduced a
+re-parameterization of the basis function matrix based on the
+decomposition of the smoothness penalty matrix \(\boldsymbol{S}\). The
+re-parameterization allows different penalties imposed on the null and
+nonlinear spaces separately, and further allows the sparsity in linear
+space and smoothness in nonlinear effect interpolation. {[}TODO: a
+sentence that it decompose the prior to independent priors, provide
+obvious computational convenience.{]} The computational convenience
+improved the scalability of the model.
+
+The proposed BHAM uses a spike-and-slab prior for function selections,
+but it substantially differs from previous spike-and-slab based GAMs,
+i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB GAM
+\citep{Bai2021} in many ways. First of all, the proposed spike-and-slab
+spline prior is a spike-and-slab lasso type prior using independent
+mixture double exponential distribution, compared to spike-and-slab GAM.
+Spike-and-slab lasso priors provide computational convenience during
+model fitting through optimization algorithms, instead of computational
+intensive sampling algorithms. They make fitting high- and
+ultrahigh-dimensional model feasible without sacrificing performance in
+prediction and variable selection. Secondly, the optimization algorithms
+introduced in this article include EM-Coordinate Descent algorithm and
+EM-Iterative weighted least square algorithm for different utilities.
+The applications of the two aforementioned algorithm are fruitful in
+Bayesian hierarchical models in high-dimensional data analysis, see
+{[}TODO: add citations{]}. Even though SB-GAM with a group
+spike-and-slab lasso prior also uses EM-CD for model fitting, the
+proposed EM-IWLS provides uncertainty measures that could be further
+used for confidence bound calculation and hypothesis testing. EM-IWLS
+complements the lack of inference potential from EM-CD in a median and
+high data setting. Furthermore, the proposed model addresses the
+incapability of bilevel selection in SB-GAM. Lastly, the R
+implementation of the Bayesian hierarchical additive model is available
+publicly via BHAM at \url{https://github.com/boyiguo1/BHAM}, make
+transnational science more accessible.
 
 The rest of the paper are arranged as the followings: in Chapter 2 we
-demonstrate the Bayesian Hierarchical generalized additive model, where
-we provide the setup of the proposed model and two fast-fitting
-algorithms. In Chapter 3, we Application to real-world datasets
-demonstrates the usefulness of the proposed methodologies in Chpater 4.
-Conclusion and discussions are made in Chapter 5.
+demonstrate the Bayesian hierarchical generalized additive model, where
+we provide the setup of the proposed model and describe the two
+fast-fitting algorithms. In Chapter 3, we compare the proposed via monte
+carlo simulation. .Application to real-world datasets demonstrates the
+usefulness of the proposed methodologies in Chpater 4. Conclusion and
+discussions are made in Chapter 5.
 

### 2021-06-27 Manuscript/01-Intro.Rmd
Revise method section

@@
 For the most common case of cubic splines, the smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of second derivative of the spline function, expressed mathematically as
-$$
-  \text{pen}f_j(x) = \lambda_j \int f^{"}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j,
-$$
+\begin{equation}\label{eq:smoothpen}
+  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{"}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
+\end{equation}
 where $\bs S_j$ is a known penalty matrix, and $\bs \beta_j$ are the basis function coefficients.

### 2021-06-27 Manuscript/02-Method.Rmd
Revise method section

@@
-# Spike-and-slab Additive Model
+# Bayesian Hierarchical Additive Models
 
-Following the notation in the previous section \ref{Notation}, suppose we have the model matrix $\bs X$ derived from spline functions of choice $h_j, j = 1, \dots, p.$ for the $p$ covariates that are modeled nonparametrically, the total dimension of $\bs X$ is $n \times \sum\limits_{j=1}^p k_j$.
+Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model 
+$$
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
+$$
+with the smoothing function $f_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
+the design matrix derived from the smoothing function $f_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
+The dimension of the design matrix depends on the choice of the smoothing function.
+For example, a cubic spline function with $k$ knots produces a $k+2$ elements column vector.
+Its corresponding smoothness penalty matrix is denoted as $\bs S_j$.
+In all, the GAMs can be formulated in a matrix form
+$$
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+$$
+where $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^t$.
+To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
+$f_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+
+To encourage proper smoothness of the smoothing functions, we adopt the idea of the smoothing penalty from smoothing spline models.
+The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the frequentist framework.
+However, direct integration of the smoothness penalty into the sparsity penalty is not obvious.
+Given that the smoothness penalty matrix is, @Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.
+
  Column-wise centering and re-parameterization introduced in the GAM as random effect section[^5] are performed.
@@
 $$
-f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
+f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen}.
 $$
-where the superscript $^t$ denotes matrix transpose.
 
-Without loss of generality, we assume that the model doesn't include the the parametric part of the model $\bs A$.
-The coefficients of $\bs A$ can be estimated easily in the proposed EM algorithms by local update conditioning on the temporary value of the spline coefficients.
-
-## Spike and Slab Prior
+## Spike and Slab Spline Prior
 

### 2021-06-27 Manuscript/SS_GAM.tex
Revise method section

@@
 parameter \(\lambda_j\) is a function of the integration of second
-derivative of the spline function, expressed mathematically as \[
-  \text{pen}f_j(x) = \lambda_j \int f^{"}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j,
-\] where \(\boldsymbol{S}_j\) is a known penalty matrix, and
+derivative of the spline function, expressed mathematically as
+\begin{equation}\label{eq:smoothpen}
+  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{"}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
+\end{equation} where \(\boldsymbol{S}_j\) is a known penalty matrix, and
 \(\boldsymbol{\beta}_j\) are the basis function coefficients. Smoothness
@@
 
-\hypertarget{spike-and-slab-additive-model}{%
-\section{Spike-and-slab Additive
-Model}\label{spike-and-slab-additive-model}}
-
-Following the notation in the previous section \ref{Notation}, suppose
-we have the model matrix \(\boldsymbol{X}\) derived from spline
-functions of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates
-that are modeled nonparametrically, the total dimension of
-\(\boldsymbol{X}\) is \(n \times \sum\limits_{j=1}^p k_j\). Column-wise
-centering and re-parameterization introduced in the GAM as random effect
+\hypertarget{bayesian-hierarchical-additive-models}{%
+\section{Bayesian Hierarchical Additive
+Models}\label{bayesian-hierarchical-additive-models}}
+
+Following the GLM notation introduced in Section \label{introduction},
+we have a generalized additive model \[
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
+\] with the smoothing function \(f_j(x_j)\) of the variable
+\(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
+matrix derived from the smoothing function \(f_j(x_j)\), is denoted
+\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
+design matrix depends on the choice of the smoothing function. For
+example, a cubic spline function with \(k\) knots produces a \(k+2\)
+elements column vector. Its corresponding smoothness penalty matrix is
+denoted as \(\boldsymbol{S}_j\). In all, the GAMs can be formulated in a
+matrix form \[
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+\] where \(\boldsymbol{\beta}_j\) denotes the basis function
+coefficients for the \(j\)th variable, and the matrix transposing
+operation is denoted with a superscript \(^t\). To note, the proposed
+model can include parametric forms of variables in the model, treating
+the parametric function as a special case of the smoothing function,
+e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix defined as
+\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
+
+To encourage proper smoothness of the smoothing functions, we adopt the
+idea of the smoothing penalty from smoothing spline models. The basic
+idea is that set up a smoothing penalty, depending on the smoothness
+penalty matrix via the penalty introduced in Eq \ref{eq:smoothpen} , in
+the posterior density function, or least squared and likelihood function
+in the frequentist framework. However, direct integration of the
+smoothness penalty with the sparsity penalty is not straightforward.
+
+have the model matrix \(\boldsymbol{X}\) derived from spline functions
+of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates that are
+modeled nonparametrically, the total dimension of \(\boldsymbol{X}\) is
+\(n \times \sum\limits_{j=1}^p k_j\). Column-wise centering and
+re-parameterization introduced in the GAM as random effect
 section{[}\^{}5{]} are performed. The re-parameterization has two
@@
 
-\hypertarget{spike-and-slab-prior}{%
-\subsection{Spike and Slab Prior}\label{spike-and-slab-prior}}
+\hypertarget{spike-and-slab-spline-prior}{%
+\subsection{Spike and Slab Spline
+Prior}\label{spike-and-slab-spline-prior}}
 
@@
 
-\hypertarget{spike-and-slab-spline-prior}{%
+\hypertarget{spike-and-slab-spline-prior-1}{%
 \subsection{Spike and slab spline
-prior}\label{spike-and-slab-spline-prior}}
+prior}\label{spike-and-slab-spline-prior-1}}
 

### 2021-06-29 Manuscript/02-Method.Rmd
Revise Method Section

@@
 To encourage proper smoothness of the smoothing functions, we adopt the idea of the smoothing penalty from smoothing spline models.
-The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the frequentist framework.
+The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework.
 However, direct integration of the smoothness penalty into the sparsity penalty is not obvious.
-Given that the smoothness penalty matrix is, @Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.
+Given that the smoothness penalty matrix is most often symmetric and positive semi-definite [^ we limit the discussion to univariate smoothing function.
+However, the penalty matrix could possibly contain multiple matrices, e.g.
+tensor product smoother.], the smoothness penalty matrix can be integrated to the design matrix $\bs X$ via re-parameterization.
+@Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.
 

### 2021-06-29 Manuscript/SS_GAM.tex
Revise Method Section

@@
 idea of the smoothing penalty from smoothing spline models. The basic
-idea is that set up a smoothing penalty, depending on the smoothness
-penalty matrix via the penalty introduced in Eq \ref{eq:smoothpen} , in
-the posterior density function, or least squared and likelihood function
-in the frequentist framework. However, direct integration of the
-smoothness penalty with the sparsity penalty is not straightforward.
-
-have the model matrix \(\boldsymbol{X}\) derived from spline functions
-of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates that are
-modeled nonparametrically, the total dimension of \(\boldsymbol{X}\) is
-\(n \times \sum\limits_{j=1}^p k_j\). Column-wise centering and
-re-parameterization introduced in the GAM as random effect
-section{[}\^{}5{]} are performed. The re-parameterization has two
-purposes. First of all the re-parameterization reshapes the columns to
-be linearly independent to each other, and hence, independent prior can
-be set to the coefficients. Secondly, the re-parameterization decomposes
-the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
-sub-matrices representing the null space of the spline and penalized
-space of the spline, \[
+idea is to set up a smoothing penalty, depending on the smoothness
+penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density
+function, or least squared and likelihood function in the penalized
+framework. However, direct integration of the smoothness penalty into
+the sparsity penalty is not obvious. Given that the smoothness penalty
+matrix is most often symmetric and positive semi-definite {[}\^{} we
+limit the discusstion to univariate smoothing function. However, the
+penalty matrix could possibly contain multiple matrices, e.g.~tensor
+product smoother.{]}, the smoothness penalty matrix can be integrated to
+the design matrix \(\boldsymbol{X}\) via re-parameterization.
+\citet{Marra2011} introduced a re-parameterization of the basis function
+matrix \(\boldsymbol{X}\) based on the decomposition of the smoothness
+penalty matrix \(\boldsymbol{S}\).
+
+Column-wise centering and re-parameterization introduced in the GAM as
+random effect section{[}\^{}5{]} are performed. The re-parameterization
+has two purposes. First of all the re-parameterization reshapes the
+columns to be linearly independent to each other, and hence, independent
+prior can be set to the coefficients. Secondly, the re-parameterization
+decomposes the model matrix \(\boldsymbol{X}_j\) for the \(j\)th
+covariates to two sub-matrices representing the null space of the spline
+and penalized space of the spline, \[
 \boldsymbol{X}_j = \begin{bmatrix}
@@
 for the \(j\)the covariate is \[
-f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
-\] where the superscript \(^t\) denotes matrix transpose.
-
-Without loss of generality, we assume that the model doesn't include the
-the parametric part of the model \(\boldsymbol{A}\). The coefficients of
-\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
-by local update conditioning on the temporary value of the spline
-coefficients.
+f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen}.
+\]
 

### 2021-06-29 Manuscript/01-Intro.Rmd
* Revise the notation section of the Method * Correct typo and grammar in intro

@@
 The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
-Additionally, two fast computing optimization based algorithms are developed for high and ultra-high dimension settings respectively.
+Additionally, two fast computing optimization-based algorithms are developed for high and ultra-high dimension settings respectively.
 Meanwhile, to avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
 Smoothness penalties are commonly implemented in GAMs through smoothing and penalized regression splines.
-They are quadratic norm of the coefficients, and allow locally adaptive penalties on each smoothing function.
-For the most common case of cubic splines, the smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of second derivative of the spline function, expressed mathematically as
+They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
+For the most common cases of cubic splines, the smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
 \begin{equation}\label{eq:smoothpen}
@@
 \end{equation}
-where $\bs S_j$ is a known penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
+where $\bs S_j$ is a known smoothness penalty matrix, and $\bs \beta_j$ is the basis function coefficients.
 Smoothness penalties are also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
@@
 The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
-[TODO: a sentence that it decompose the prior to independent priors, provide obvious computational convenience.] The computational convenience improved the scalability of the model.
+^[TODO: a sentence that it decompose the prior to independent priors, provide obvious computational convenience.] The computational convenience improved the scalability of the model.
 
@@
 Spike-and-slab lasso priors provide computational convenience during model fitting through optimization algorithms, instead of computational intensive sampling algorithms.
-They make fitting high- and ultrahigh-dimensional model feasible without sacrificing performance in prediction and variable selection.
+They make fitting high- and ultrahigh-dimensional models feasible without sacrificing performance in prediction and variable selection.
 Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent algorithm and EM-Iterative weighted least square algorithm for different utilities.
-The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis, see [TODO: add citations].
+The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis, see ^[TODO: add citations].
 Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.

### 2021-06-29 Manuscript/02-Method.Rmd
* Revise the notation section of the Method * Correct typo and grammar in intro

@@
 
-To encourage proper smoothness of the smoothing functions, we adopt the idea of the smoothing penalty from smoothing spline models.
+To encourage proper smoothness of the smoothing functions, we adopt the idea of smoothing penalties from smoothing spline models.
 The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework.
-However, direct integration of the smoothness penalty into the sparsity penalty is not obvious.
-Given that the smoothness penalty matrix is most often symmetric and positive semi-definite [^ we limit the discussion to univariate smoothing function.
+However, the direct integration of smoothness penalties into sparsity penalties is not obvious.
+Given the smoothness penalty matrix is most often symmetric and positive semi-definite ^[ we limit the discussion to univariate smoothing function.
 However, the penalty matrix could possibly contain multiple matrices, e.g.
-tensor product smoother.], the smoothness penalty matrix can be integrated to the design matrix $\bs X$ via re-parameterization.
-@Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.
+tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization.
+The re-parameterization follows @Marra2011, in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function.
+We further write the orthonormal matrix $\bs U$, containing the eigenvectors as column, as an (row) augmentation of two sub-matrices $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$, where $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
+the penalized space, and $\bs U^0$ contains the eigenvectors for the null space.
+We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$.
+An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see @Wood2006), but I am not sure if it can be generalized to all smoothing functions].
+For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter.
+Similarly, the spline function can be expressed in the matrix form
+$$
+f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = {\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp ,
+$$
 
- Column-wise centering and re-parameterization introduced in the GAM as random effect section[^5] are performed.
-The re-parameterization has two purposes.
-First of all the re-parameterization reshapes the columns to be linearly independent to each other, and hence, independent prior can be set to the coefficients.
-Secondly, the re-parameterization decomposes the model matrix $\bs X_j$ for the $j$th covariates to two sub-matrices representing the null space of the spline and penalized space of the spline,
-$$
-\bs X_j = \begin{bmatrix}
-\bs X_j^{0} & \bs X_j^{pen}
-\end{bmatrix},
-$$
-where $\bs X_j^{null}$ is the column space representing the null space of the spline model, and $\bs X_j^{pen}$ denotes the column spaces that need to be shrunk for smoothness of the spline.
-The purpose for this re-parameterization is to isolate the null space from the smoothing space, such that answering the question, if non-linear relationship is necessary, becomes easy by deciding if the corresponding coefficients of the $\bs X_j^{pen}$ are jointly zero.
-This strategy is widely used, previously mentioned in @Scheipl2012 and the references there in.^[Possibly select some to add here.] Hereafter, we denote the coefficients of the null space as $\bs \beta^{0}_j$ and the coefficients of the penalized space as $\bs \beta^{pen}_j$.
-The spline function for the $j$the covariate is
-$$
-f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen}.
-$$
+The re-parameterization sets up the foundation of the proposed model, which has three-fold benefits.
+First of all, as previously mentioned, the re-parameterization integrated the smoothness penalty into the design matrix, and hence the penalty can be expressed in the posterior density function implicitly in addition to the sparse penalty for variable and function selection.
+Secondly, the eigendecomposition of the smooth penalty isolates the linear space of a smooth function from the non-linear space, improving the feasibility of bi-level variable/function selection and estimation inference.
+The eigendecomposition also introduced the linearly independent column spaces of the smooth function matrix, which allows imposing independent priors on the coefficients.
+This reduces the computational complexity that multivariate priors induces, and greatly broadens the choices of priors and further model choices.
+Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of penalty priors, from column-specific penalties to a unified penalty.
+This step is also necessary for the proposed spike-and-slab spline prior.
+Conventionally, the spike-and-slab priors require scaling of predictors ^[TODO: insert a citation.]
 

### 2021-06-29 Manuscript/SS_GAM.tex
* Revise the notation section of the Method * Correct typo and grammar in intro

@@
 \newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\tp}{\text{pen}}
 
@@
 possibilities: no effect, only linear effect, or nonlinear effect of a
-predictor. Additionally, two fast computing optimization based
+predictor. Additionally, two fast computing optimization-based
 algorithms are developed for high and ultra-high dimension settings
@@
 implemented in GAMs through smoothing and penalized regression splines.
-They are quadratic norm of the coefficients, and allow locally adaptive
-penalties on each smoothing function. For the most common case of cubic
+They are quadratic norms of the coefficients and allow locally adaptive
+penalties on each smoothing function. For the most common cases of cubic
 splines, the smoothness penalty conditioning on a smoothness penalty
-parameter \(\lambda_j\) is a function of the integration of second
+parameter \(\lambda_j\) is a function of the integration of the second
 derivative of the spline function, expressed mathematically as
@@
   \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{"}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known penalty matrix, and
-\(\boldsymbol{\beta}_j\) are the basis function coefficients. Smoothness
-penalties are also previously used in the spike-and-slab GAM
+\end{equation} where \(\boldsymbol{S}_j\) is a known smoothness penalty
+matrix, and \(\boldsymbol{\beta}_j\) is the basis function coefficients.
+Smoothness penalties are also previously used in the spike-and-slab GAM
 \citep{Scheipl2012} and the sparsity-smoothness penalty
@@
 nonlinear spaces separately, and further allows the sparsity in linear
-space and smoothness in nonlinear effect interpolation. {[}TODO: a
-sentence that it decompose the prior to independent priors, provide
-obvious computational convenience.{]} The computational convenience
+space and smoothness in nonlinear effect interpolation. \footnote{TODO:
+  a sentence that it decompose the prior to independent priors, provide
+  obvious computational convenience.} The computational convenience
 improved the scalability of the model.
@@
 intensive sampling algorithms. They make fitting high- and
-ultrahigh-dimensional model feasible without sacrificing performance in
+ultrahigh-dimensional models feasible without sacrificing performance in
 prediction and variable selection. Secondly, the optimization algorithms
@@
 Bayesian hierarchical models in high-dimensional data analysis, see
-{[}TODO: add citations{]}. Even though SB-GAM with a group
+\footnote{TODO: add citations}. Even though SB-GAM with a group
 spike-and-slab lasso prior also uses EM-CD for model fitting, the
@@
 To encourage proper smoothness of the smoothing functions, we adopt the
-idea of the smoothing penalty from smoothing spline models. The basic
-idea is to set up a smoothing penalty, depending on the smoothness
-penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density
+idea of smoothing penalties from smoothing spline models. The basic idea
+is to set up a smoothing penalty, depending on the smoothness penalty
+matrix via Equation \ref{eq:smoothpen}, in the posterior density
 function, or least squared and likelihood function in the penalized
-framework. However, direct integration of the smoothness penalty into
-the sparsity penalty is not obvious. Given that the smoothness penalty
-matrix is most often symmetric and positive semi-definite {[}\^{} we
-limit the discusstion to univariate smoothing function. However, the
-penalty matrix could possibly contain multiple matrices, e.g.~tensor
-product smoother.{]}, the smoothness penalty matrix can be integrated to
-the design matrix \(\boldsymbol{X}\) via re-parameterization.
-\citet{Marra2011} introduced a re-parameterization of the basis function
-matrix \(\boldsymbol{X}\) based on the decomposition of the smoothness
-penalty matrix \(\boldsymbol{S}\).
-
-Column-wise centering and re-parameterization introduced in the GAM as
-random effect section{[}\^{}5{]} are performed. The re-parameterization
-has two purposes. First of all the re-parameterization reshapes the
-columns to be linearly independent to each other, and hence, independent
-prior can be set to the coefficients. Secondly, the re-parameterization
-decomposes the model matrix \(\boldsymbol{X}_j\) for the \(j\)th
-covariates to two sub-matrices representing the null space of the spline
-and penalized space of the spline, \[
-\boldsymbol{X}_j = \begin{bmatrix}
-\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
-\end{bmatrix},
-\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
-the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
-denotes the column spaces that need to be shrunk for smoothness of the
-spline. The purpose for this re-parameterization is to isolate the null
-space from the smoothing space, such that answering the question, if
-non-linear relationship is necessary, becomes easy by deciding if the
-corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
-zero. This strategy is widely used, previously mentioned in
-\citet{Scheipl2012} and the references there in.\footnote{Possibly
-  select some to add here.} Hereafter, we denote the coefficients of the
-null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
-penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
-for the \(j\)the covariate is \[
-f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen}.
+framework. However, the direct integration of smoothness penalties into
+sparsity penalties is not obvious. Given the smoothness penalty matrix
+is most often symmetric and positive semi-definite \footnote{ we limit
+  the discussion to univariate smoothing function. However, the penalty
+  matrix could possibly contain multiple matrices, e.g.~tensor product
+  smoother.}, it can be integrated into the design matrix
+\(\boldsymbol{X}\) via a re-parameterization. The re-parameterization
+follows \citet{Marra2011}, in which eigen-decomposed the penalty matrix
+\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\),
+where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
+arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
+\(0\) elements on the diagonal, associated with the null space of the
+smooth function. We further write the orthonormal matrix
+\(\boldsymbol{U}\), containing the eigenvectors as column, as an (row)
+augmentation of two sub-matrices
+\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\),
+where \(\boldsymbol{U}^{\text{pen}}\) contains the eigenvectors (as
+columns) for the non-zero eigenvalues, i.e.~the penalized space, and
+\(\boldsymbol{U}^0\) contains the eigenvectors for the null space. We
+multiply the basis function matrix \(\boldsymbol{X}\) with the
+orthonormal matrix \(\boldsymbol{U}\) for the new design matrix
+\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{\text{pen}} \end{bmatrix}\).
+An additional scaling step is imposed on \(\boldsymbol{X}^{\text{pen}}\)
+by the non-zero eigenvalues of \(\boldsymbol{D}\) such that the new
+basis function matrix \(\boldsymbol{X}^\ast\) is scale-invariant
+\footnote{ I know this work for tensor spline (see \citet{Wood2006}),
+  but I am not sure if it can be generalized to all smoothing functions}.
+For notation convenience, we denote
+\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{\text{pen}} \end{bmatrix}\)
+as the after re-parameterized basis function matrix for the \(j\)th
+variable hereafter. Similarly, the spline function can be expressed in
+the matrix form \[
+f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = {\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
 \]
+
+The re-parameterization sets up the foundation of the proposed model,
+which has three-fold benefits. First of all, as previously mentioned,
+the re-parameterization integrated the smoothness penalty into the
+design matrix, and hence the penalty can be expressed in the posterior
+density function implicitly in addition to the sparse penalty for
+variable and function selection. Secondly, the eigendecomposition of the
+smooth penalty isolates the linear space of a smooth function from the
+non-linear space, improving the feasibility of bi-level
+variable/function selection and estimation inference. The
+eigendecomposition also introduced the linearly independent column
+spaces of the smooth function matrix, which allows imposing independent
+priors on the coefficients. This reduces the computational complexity
+that multivariate priors induces, and greatly broadens the choices of
+priors and further model choices. Last but not least, the scaling step
+of \(\boldsymbol{X}^\text{pen}\) further simplifies the choice of
+penalty priors, from column-specific penalties to a unified penalty.
+This step is also necessary for the proposed spike-and-slab spline
+prior. Conventionally, the spike-and-slab priors require scaling of
+predictors \footnote{TODO: insert a citation.}
 

### 2021-07-02 Manuscript/00-main.Rmd
* Revise the notation section of the Method * Correct typo and grammar in intro
Revise Method section

@@
 - name: Byron C. Jaeger
-  num: a 
+  num: b 
 - name: AKM Fazlur Rahman
@@
   org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
-# - num: b
-  # org: Department of Applied Things, University B, City B, Country B
+- num: b
+  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
 corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"

### 2021-07-02 Manuscript/01-Intro.Rmd
Revise Method section

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most utilized modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu.
-\phi)$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$^[TODO: explain $\phi$], can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$

### 2021-07-02 Manuscript/02-Method.Rmd
Revise Method section

@@
 
-Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model 
-$$
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
-$$
+Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model and its matrix form,
+\begin{equation}\label{eq:gam}
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+\end{equation}
 with the smoothing function $f_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
@@
 Its corresponding smoothness penalty matrix is denoted as $\bs S_j$.
-In all, the GAMs can be formulated in a matrix form
-$$
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
-$$
-where $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^t$.
+$\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$.
 To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
@@
 $$
-f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = {\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp ,
+f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
 $$
+and the generalized additive model in Equation \ref{eq:gam} be
+\begin{equation}\label{eq:gam-repa}
+E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp)\right].
+\end{equation}
 
@@
 
-## Spike and Slab Spline Prior
+## Spike-and-Slab Smooth Priors
 
-For high-dimensional data analysis, a common Bayesian modeling approach is to impose a particular family of priors, the spike-and-slab (SS) priors on the coeffcient in a regression modelss.
-The term "spike-and-slab" was first coined in @Mitchell1988 describing the shape of the prior.
-It is a mixture distribution, comprised with a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
+The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection.
+We defer to @Bai2021Review for an in-depth introduction to spike-and-slab priors.
+To summary, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
 $$
- \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
+ \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
 $$
+The most distinct feature of the SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
 
-The most distinct feature of the SS priors family is that it is conditioned on a latent binary variable $\gamma_i$ that indicates whether the covariate $x_i$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$ and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
-Here we introduce the most popular spike-and-slab prior introduced in @George1993, as  stochastic search variable selection (SSVS).
-Given a regression problem
-$$
-Y = \bs X\beta+\epsilon, 
-$$
-we can assume the prior distribution of the regression parameters $\beta$ depends on a vector of binary variables $\gamma=(\gamma_1, \dots \gamma_p)^\prime$ as follows: given $\gamma_i$, $\beta_i$ are independent distributed as a mixture priors
-$$
-\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
-$$
-The prior of $\gamma$ is traditionally assumed to be identically independent Bernoulli distribution.
-Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
+<!-- Comments -->
+<!-- Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
 @George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection.
@@
 @George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.
+-->
 
-To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
-EM-based algorithms are prevalent used to speed the computation.
-@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
-The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
-The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
-Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
-In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
-As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
-EMVS is also compatible with structured prior when dealing with grouped variables.
-The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
+The major criticism of early spike-and-slab priors receives is being computationally prohibitive.
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a], ^[TODO: add more citation].
+Particularly, the development of Spike-and-slab Lasso [@Rockova2018b; @Rockova2018] substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis ^[TODO: add Dr.
+Yi's papers].
+The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+\begin{equation} \label{eq:ssl}
+\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
+\end{equation}
+The SSL also mitigated the problem of EMVS where the weak signals are not shrink to zero, favored in high-dimensional data analysis.
+We noticed that @Bai2021 introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
+@Bai2021 takes an "all-in-all-out" strategy for function selection.
 
-The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
-The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
-The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
-Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
-Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
-However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
-@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
-
- Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
-Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
-The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
-We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
-
-## Spike and slab spline prior
-
-A generalized spline model can be formulated as 
-$$
-y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
-$$
-The link function of the mean can be expressed in a matrix form
-$$
-g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
- = \beta^T X_i,
-$$
-where
+### Spike-and-Slab Lasso Prior
+We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear effect modeling with bi-level selection.
+The proposed prior extends from the spike-and-slab lasso prior described in Equation \ref{eq:ssl}.
+Given the reparameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}^T$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$.
+Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,
 $$
 \begin{aligned}
-  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
-& &
-  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), 
+\end{aligned}
+$$
+where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the non-linear effect of the $j$th variable respectively.
+; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
+These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation.
+A discussion of how to choose the scale parameters comes in Section \ref{selecting-optimal-scale-values}.
+To note, this prior differs from previous group spike-and-slab lasso priors[TODO: add Dr.
+Yi's citation], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively.
+It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection.
+Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e.
+$\gamma_j = \gamma^0_j = \gamma^{pen}_j$.
+This converges^[Is "becomes" better?] to the SB-GAM [@Bai2021].
+
+The reparameterization introduced in Section \ref{bayesian-hierarchical-additive-models} grants the validity of the proposed prior.
+First of all, the smooth function bases are linear dependent to each other based on the definitions, and ask for extra attention.
+The eigeondecomposition remedies the problem.
+Hence our prior can be set to be conditionally independent.
+Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables, where scaling is mandate.
+
+The rest of the hierchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution.
+The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^{pen}_j$, shares the same probability parameter,
+$$
+\begin{aligned}
+\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
+\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
 \end{aligned}
@@
 
-\
+This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function.
+The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^{pen}$ decides the smoothness of the spline function at basis function level.
+Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity.
+We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
+$$
+\theta_j \sim Beta(a, b).
+$$
+The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience.
+For simplicity, we focus on a special case of beta distribution, uniform (0,1).
+When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
+Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL.
 
-### Mixture-Normal Prior
-Given the full deign matrix $X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T$, the likelihood function of the generalized model can be written as
-$$
-p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
-$$
+\newpage
+
+```{r}
+include_graphics("Fig/SS_prior_flow_chart.png")
+```
+
+### Other Priors
 
@@
 
-```{r}
-include_graphics("Fig/SS_prior_flow_chart.png")
-```
-
- \
-
-### Mixture Double Exponential Prior
 One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects.
@@
 Specifically, double exponential prior is equivalent to the LASSO penalty in the frequentist models, providing a sparse solution.
-Meanwhile, the estimation algorithm of LASSO penalty are extremely efficient, greatly reducing the computation needs.
-Recent developments of spike-and-slab priors include spike-and-slab LASSO prior, i.e.
-spike-and-slab mixture double exponential prior.
-We extend the spike-and-slab LASSO prior to high-dimensional GAMs, providing computationally efficiency and sparse solution to our proposed ssGAM-MN.
 
-The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we replace the normal distribution with double exponential distribution.
-The prior distributions of the coefficients $\beta^0_j, \beta^{pen}_j$ for the $j$ predictor can be formulated as
-$$
-\begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
-  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
-\end{aligned}
-$$
-The hyper-priors for $\gamma^0_j, \gamma^{pen}_j, \theta_j$ are kept the same as specified in ssGAM-MN.
+## Algorithms for Fitting BHAMs
 
-\
+To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
+EM-based algorithms are prevalent used to speed the computation.
+@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
+The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
+The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
+Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
+In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
+EMVS is also compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
 
-## Alforithm for fitting ssGAMs
+The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
+The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
+The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
+Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
+Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
+However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
+@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
+
+ Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
+Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
+The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
+We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
 
@@
 Compared to the EM-IRLS algorithm, the EM-coordinate descent provides an expedited model fitting process at the cost of model uncertainty.
-
-\
 

### 2021-07-02 Manuscript/05-Discussion.Rmd
Revise Method section

@@
 
+Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity.
+However, this set-up could possibility result in a situation that is not theoretically possible: the non-linear component is selected, but the linear component is not.
+This can be addressed analytically by forcing to include the linear component when non-linear component is included.
+Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e.
+$\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.
+
 ## Discussion
@@
 In addition, the model can be used to answer the question if non-linear effect are necessary.
+
+  <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function.
+The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder.
+In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.
+-->
+
+The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not.
+On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e.
+$\gamma^{pen}_{jk}$ of $\beta^{pen}_{jk}$ for $k = 1\dots, K_j$.
+This prior set-up generalized to the group spike-and-slab.
+The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity.
+The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup.
+Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
 

### 2021-07-02 Manuscript/SS_GAM.tex
Revise Method section

@@
 \author[a]{Boyi Guo*}
-\author[a]{Byron C. Jaeger}
+\author[b]{Byron C. Jaeger}
 \author[a]{AKM Fazlur Rahman}
@@
 Birmingham, Birmingham, USA}
+\address[b]{Department of Biostatistics and Data Science, Wake Forest
+University, Winston-Salem, USA}
 
@@
 variable, \(Y\), which is assumed to follow some exponential family
-distribution \(EF(\mu. \phi)\), can be expressed as the summation of
-smooth functions, \(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional
-vector of covariates \(\boldsymbol{x}\), written as \[
+distribution \(EF(\mu, \phi)\)\footnote{TODO: explain \(\phi\)}, can be
+expressed as the summation of smooth functions,
+\(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector of
+covariates \(\boldsymbol{x}\), written as \[
 E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
@@
 Following the GLM notation introduced in Section \label{introduction},
-we have a generalized additive model \[
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
-\] with the smoothing function \(f_j(x_j)\) of the variable
+we have a generalized additive model and its matrix form,
+\begin{equation}\label{eq:gam}
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+\end{equation} with the smoothing function \(f_j(x_j)\) of the variable
 \(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
@@
 elements column vector. Its corresponding smoothness penalty matrix is
-denoted as \(\boldsymbol{S}_j\). In all, the GAMs can be formulated in a
-matrix form \[
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
-\] where \(\boldsymbol{\beta}_j\) denotes the basis function
-coefficients for the \(j\)th variable, and the matrix transposing
-operation is denoted with a superscript \(^t\). To note, the proposed
-model can include parametric forms of variables in the model, treating
-the parametric function as a special case of the smoothing function,
-e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix defined as
-\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
+denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\) denotes the
+basis function coefficients for the \(j\)th variable, and the matrix
+transposing operation is denoted with a superscript \(^T\). To note, the
+proposed model can include parametric forms of variables in the model,
+treating the parametric function as a special case of the smoothing
+function, e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix
+defined as \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
 
@@
 the matrix form \[
-f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = {\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
-\]
+f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
+\] and the generalized additive model in Equation \ref{eq:gam} be
+\begin{equation}\label{eq:gam-repa}
+E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen})\right].
+\end{equation}
 
@@
 
-\hypertarget{spike-and-slab-spline-prior}{%
-\subsection{Spike and Slab Spline
-Prior}\label{spike-and-slab-spline-prior}}
-
-For high-dimensional data analysis, a common Bayesian modeling approach
-is to impose a particular family of priors, the spike-and-slab (SS)
-priors on the coeffcient in a regression modelss. The term
-``spike-and-slab'' was first coined in \citet{Mitchell1988} describing
-the shape of the prior. It is a mixture distribution, comprised with a
-skinny spike density \(f_{spike}(\cdot)\) for weak signals and a flat
-slab density \(f_{slab}(\cdot)\) for large signals, mathematically \[
- \beta_i|\gamma_i \sim (1-\gamma_i)f_{spike}(\beta_i) + \gamma_if_{slab}(\beta_i).
-\]
-
-The most distinct feature of the SS priors family is that it is
-conditioned on a latent binary variable \(\gamma_i\) that indicates
-whether the covariate \(x_i\) is included in the model. There are
-various spike-and-slab priors depending on the choice for the spike
-density \(f_{spike}(\cdot)\) and the slab density \(f_{slab}(\cdot)\),
-for example \citet{George1993}; \citet{George1997}; \citet{Chipman1996}
-for grouped variables; \citet{Brown1998} for multivariate outcomes;
-\citet{Ishwaran2005}; \citet{Clyde2004} and reference therein. Here we
-introduce the most popular spike-and-slab prior introduced in
-\citet{George1993}, as stochastic search variable selection (SSVS).
-Given a regression problem \[
-Y = \boldsymbol{X}\beta+\epsilon, 
-\] we can assume the prior distribution of the regression parameters
-\(\beta\) depends on a vector of binary variables
-\(\gamma=(\gamma_1, \dots \gamma_p)^\prime\) as follows: given
-\(\gamma_i\), \(\beta_i\) are independent distributed as a mixture
-priors \[
-\beta_i | \gamma_i \sim (1-\gamma_i)N(0, \sigma^2 v_0) + \gamma_iN(0,\sigma^2 v_1).
-\] The prior of \(\gamma\) is traditionally assumed to be identically
-independent Bernoulli distribution. Even the conjugate normal prior of
-\(\boldsymbol{\beta}\) provides analytic simplicity, there is no closed
-form solution for the posterior density of \(\boldsymbol{\gamma}\).
-\citet{George1993} described a Gibbs sampling algorithm to approximate
-the posterior distribution of \(\boldsymbol{\gamma}\) for variable
-selection. The same authors later compared SSVS with other Bayesian
-variable selection approaches, providing some empirical suggestions on
-the advantages and disadvantages of various formulations and estimation
-algorithms. \citep{George1997} Among the various problem SSVS addresses,
-high-dimensionality is not one of them, due to the intensive computation
-needs of MCMC algorithms. \citet{George1997} suggested MCMC algorithms
-works well for medium size of predictors (p=25), which is not feasible
-for high-dimensional data, where the number of predictors easily exceeds
-100.
-
-To alleviate the computation burden, many methodology development
-studies focus on the maximum a posteriori (MAP) estimate
-\citep{Tipping2004}. EM-based algorithms are prevalent used to speed the
-computation. \citet{Rockova2014a} proposed an EM-based algorithm for a
-spike-and-slab Gaussian mixture prior, EMVS. The coefficient \(\beta_j\)
-independently follows a mixture normal distribution
-\(N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)\) for
-\(0 \leq v_0 < v_1\). The parameter \(\sigma^2\) follows a inverse gamma
-prior, \(IG(v/2, v\lambda /2)\). Like in the spike-and slab family, the
-latent binary variables \(\boldsymbol{\gamma}\) follows a Bernoulli
-prior with a hyper prior \(\theta\) follows a beta distribution, or
-simpler the uniform (0,1). In the E step, the latent variable are
-treated as the missing data, and calculate the posterior mean of
-\(\gamma\); in the M step, a ridge estimator was used to update the
-coefficients, and \(\sigma^2, \theta\) are updated accordingly. As the
-ridge regression creates small but nonzero coefficients, a further step
-is needed to select the variable: EMVS further suggests to threshold the
-inclusion probability for variable selection or generate regularization
-plot with different values of the spike scale parameter. EMVS is also
-compatible with structured prior when dealing with grouped variables.
-The group structure can be imposed on the prior of the latent binary
-variable \(\gamma\) via either logistic regression or markov random
-field.
-
-The same authors \citep{Rockova2018b, Rockova2018} further extended the
-continuous spike-and-slab Gaussian prior to the spike-and-slab double
-exponential distribution, as known as spike-and-slab lasso (SSL). The
-authors also lay the theoretical discussion of comparing spike-and-slab
-from the penalty perspective and the Bayesian perspective. The SSL also
-mitigated the problem of EMVS that the weak signals are not shrink to
-zero. Yi and his group incorporated the SSL prior in the EM-IWLS
-algorithm and proposed another EM algorithm, EM-cyclic coordinate
-descent, based on the Lasso penalty SSL exhibit. Such application were
-used in generalized linear models\citep{Tang2017a}, cox models
-\citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithm provide
-deterministic solutions, which becomes a more popular property as
-reproducible research becomes more valued. However, while the EMVS are
-not able to provide inference of the coefficients easily, the EM-IWLS
-can easily extract the variance-covariance matrix due to the IWLS
-algorithm. \citet{Bai2020} provides an up-to-date summary of SSL methods
-and theoretical discussion.
-
-Given the proposed model specification, the estimation of the parameters
-can be easily achieved with MCMC algorithm. Nevertheless, considering
-the computational burden MCMC algorithms bear especially in
-high-dimensional analysis, we propose parsimonious solutions, stemming
-from previous EM-IWLS \citep{Rockova2014a}and EM-coordinate descent
-algorithms\citep{Tang2018, Tang2019}. The computational friendliness
-allows for fitting a model with different sets of initial values,
-mimicking the posterior sampling in the MCMC algorithms. We are hopeful
-that this could improve the sacrificed robustness of the estimation
-process due to the MAP estimation used by the EM algorithms in
-comparison to the model averaging approach.
-
-\hypertarget{spike-and-slab-spline-prior-1}{%
-\subsection{Spike and slab spline
-prior}\label{spike-and-slab-spline-prior-1}}
-
-A generalized spline model can be formulated as \[
-y_i \sim EF(\mu_i, \phi), \quad g(\mu_i) = \sum\limits_{j=1}^p f_j(x_{ij}), \quad i = 1, \dots, n.
-\] The link function of the mean can be expressed in a matrix form \[
-g(\mu_i) = \sum\limits_{j=1}^p\left[f_j^0(x_{ij}) + f_j^{pen}(x_{ij})\right] = \sum\limits_{j=1}^p\left[{\beta_j^0}^T X_{ij}^0 + {\beta_j^{pen}}^T X_{ij}^{pen}\right]
- = \beta^T X_i,
-\] where \[
+\hypertarget{spike-and-slab-smooth-priors}{%
+\subsection{Spike-and-Slab Smooth
+Priors}\label{spike-and-slab-smooth-priors}}
+
+The family of spike-and-slab (SS) priors regression models dominates
+Bayesian high-dimensional analysis for its utility in outcome prediction
+and variable selection. We defer to \citet{Bai2021Review} for an
+in-depth introduction to spike-and-slab priors. To summary,
+spike-and-slab priors are a family of mixture distributions that
+comprises a skinny spike density \(f_{spike}(\cdot)\) for weak signals
+and a flat slab density \(f_{slab}(\cdot)\) for large signals,
+mathematically \[
+ \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
+\] The most distinct feature of the SS priors is that it is conditioned
+on a latent binary variable \(\gamma\) that indicates whether the
+variable \(x\) is included in the model. There are various
+spike-and-slab priors depending on the choice for the spike density
+\(f_{spike}(\cdot)\)\footnote{TODO: should I change to notation of \(f\)
+  to not confound with the smooth functions?} and the slab density
+\(f_{slab}(\cdot)\), for example \citet{George1993}; \citet{George1997};
+\citet{Chipman1996} for grouped variables; \citet{Brown1998} for
+multivariate outcomes; \citet{Ishwaran2005}; \citet{Clyde2004} and
+reference therein.
+
+The major criticism of early spike-and-slab priors receives is being
+computationally prohibitive. \citep{Bai2021Review} Since then, many
+studies focus on alleviating the computational burden that sampling
+algorithms bear, which include EMVS \citep{Rockova2014a}, \footnote{TODO:
+  add more citation}. Particularly, the development of Spike-and-slab
+Lasso \citep{Rockova2018b, Rockova2018} substantially improves the
+scalability of SS models, setting up the theoretical foundation for
+generalized models in -omics data analysis \footnote{TODO: add Dr.~Yi's
+  papers}. The spike-and-slab lasso prior is composed of two double
+exponential distributions with mean 0 and different dispersion
+parameters, \(0 < s_0 < s_1\), mathematically,
+\begin{equation} \label{eq:ssl}
+\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
+\end{equation} The SSL also mitigated the problem of EMVS where the weak
+signals are not shrink to zero, favored in high-dimensional data
+analysis. We noticed that \citet{Bai2021} introduces a spike-and-slab
+group prior in the GAM framework, where the densities of the spike and
+slab components take the group lasso density \citep{Xu2015}.
+\citet{Bai2021} takes an ``all-in-all-out'' strategy for function
+selection.
+
+\hypertarget{spike-and-slab-lasso-prior}{%
+\subsubsection{Spike-and-Slab Lasso
+Prior}\label{spike-and-slab-lasso-prior}}
+
+We introduce a novel prior for GAMs, particularly for high-dimensional
+nonlinear effect modeling with bi-level selection. The proposed prior
+extends from the spike-and-slab lasso prior described in Equation
+\ref{eq:ssl}. Given the reparameterized deign matrix
+\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^\text{pen}\end{bmatrix}^T\)
+for the \(j\)th variable, we impose a two-part SSL prior to the
+coefficients \(\boldsymbol{\beta}_j\). Specifically, we impose
+independent group priors on the null space coefficients and on the
+penalized space coefficients respectively, \[
 \begin{aligned}
-  X_i = \begin{bmatrix}X^0_{i1}\\X^{pen}_{i1}\\\vdots \\X^0_{ip}\\X^{pen}_{ip}\end{bmatrix}
-& &
-  \beta = \begin{bmatrix}\beta^0_1\\\beta^{pen}_{1}\\\vdots \\\beta^0_{p}\\\beta^{pen}_{p}\end{bmatrix}
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1), 
+\end{aligned}
+\] where \(\gamma^0_{j}\in\{0,1\}\) and
+\(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator
+variables, indicating if the model includes the linear effect and the
+non-linear effect of the \(j\)th variable respectively. ; \(s_0\) and
+\(s_1\) are scale parameters, assuming \(0 < s_0 < s_1\) and given.
+These scale parameters \(s_0\) and \(s_1\) can be treated as tuning
+parameters and optimized via cross-validation. A discussion of how to
+choose the scale parameters comes in Section
+\ref{selecting-optimal-scale-values}. To note, this prior differs from
+previous group spike-and-slab lasso priors{[}TODO: add Dr.~Yi's
+citation{]}, where the \(\boldsymbol{\beta}^0_j\) and
+\(\boldsymbol{\beta}^\text{pen}_j\) shares the same indicator variables
+\(\gamma_j^0\), \(\gamma_j^\text{pen}\) respectively. It is possible to
+relax the assumption such that each coefficient \(\beta\) to have its
+own latent indicator \(\gamma\), but at the cost ofcomplicating the
+bi-level function selection. Conversely, it is possible to add a more
+restrictive assumption on the priors, assuming that one indicator
+variable decides the inclusion of both the linear effect and non-linear
+effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). This
+converges\footnote{Is ``becomes'' better?} to the SB-GAM
+\citep{Bai2021}.
+
+The reparameterization introduced in Section
+\ref{bayesian-hierarchical-additive-models} grants the validity of the
+proposed prior. First of all, the smooth function bases are linear
+dependent to each other based on the definitions, and ask for extra
+attention. The eigeondecomposition remedies the problem. Hence our prior
+can be set to be conditionally independent. Secondly, the eigenvalue
+scaling provides a panacea to allow unified scale parameters for all
+bases of all smoothing functions of variables, where scaling is mandate.
+
+The rest of the hierchical prior follows the traditional spike-and-slab
+lasso piror: we set up the hyper-prior of \(\boldsymbol{\gamma}\), to
+allow local adaption of the shrinkage, using a bernoulli distribution.
+The two indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
+\(\gamma^{pen}_j\), shares the same probability parameter, \[
+\begin{aligned}
+\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
+\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
 \end{aligned}
@@
 
-\hfill\break
-
-\hypertarget{mixture-normal-prior}{%
-\subsubsection{Mixture-Normal Prior}\label{mixture-normal-prior}}
-
-Given the full deign matrix
-\(X = \begin{pmatrix} X_1 & X_2 & \dots & X_n \end{pmatrix}^T\), the
-likelihood function of the generalized model can be written as \[
-p(y|X\beta, \phi) = \prod\limits^n_{i=1}p(y_i|\beta^T X_i, \phi).
-\]
+This is to leverage the fact that the probability of selecting the bases
+of a smooth function should be similar, while allowing different penalty
+on the null space and penalty space of the spline function. The hyper
+prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
+function selection level, while that of \(\gamma_{j}^{pen}\) decides the
+smoothness of the spline function at basis function level. Meanwhile, we
+specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are independently
+distributed for analytic simplicity. We further specify the parameter
+\(\theta_j\) follows a beta distribution with given shape parameters
+\(a\) and \(b\), \[
+\theta_j \sim Beta(a, b).
+\] The beta distribution is a conjugate prior for the binomial
+distribution and hence provides some computation convenience. For
+simplicity, we focus on a special case of beta distribution, uniform
+(0,1). When the variable have large effects in any of the bases, the
+parameter \(\theta_j\) will be estimated large, which in turn encourages
+the model to include the rest of bases, achieving the local adaption
+among spline bases. Hereafter, we refer Bayesian hierarchical
+generalized additive models with the spike-and-slab lasso prior as the
+BHAM-SSL.
+
+\newpage
+
+\includegraphics{Fig/SS_prior_flow_chart.png}
+
+\hypertarget{other-priors}{%
+\subsubsection{Other Priors}\label{other-priors}}
 
@@
 
-\includegraphics{Fig/SS_prior_flow_chart.png}
-
-~
-
-\hypertarget{mixture-double-exponential-prior}{%
-\subsubsection{Mixture Double Exponential
-Prior}\label{mixture-double-exponential-prior}}
-
 One of the critics received by the spike-and-slab mixture normal prior
@@
 the LASSO penalty in the frequentist models, providing a sparse
-solution. Meanwhile, the estimation algorithm of LASSO penalty are
-extremely efficient, greatly reducing the computation needs. Recent
-developments of spike-and-slab priors include spike-and-slab LASSO
-prior, i.e.~spike-and-slab mixture double exponential prior. We extend
-the spike-and-slab LASSO prior to high-dimensional GAMs, providing
-computationally efficiency and sparse solution to our proposed ssGAM-MN.
-
-The set-up of ssGAM-LASSO model is similar to ssGAM-MN, except we
-replace the normal distribution with double exponential distribution.
-The prior distributions of the coefficients \(\beta^0_j, \beta^{pen}_j\)
-for the \(j\) predictor can be formulated as \[
-\begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
-  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1).
-\end{aligned}
-\] The hyper-priors for \(\gamma^0_j, \gamma^{pen}_j, \theta_j\) are
-kept the same as specified in ssGAM-MN.
-
-\hfill\break
-
-\hypertarget{alforithm-for-fitting-ssgams}{%
-\subsection{Alforithm for fitting
-ssGAMs}\label{alforithm-for-fitting-ssgams}}
+solution.
+
+\hypertarget{algorithms-for-fitting-bhams}{%
+\subsection{Algorithms for Fitting
+BHAMs}\label{algorithms-for-fitting-bhams}}
+
+To alleviate the computation burden, many methodology development
+studies focus on the maximum a posteriori (MAP) estimate
+\citep{Tipping2004}. EM-based algorithms are prevalent used to speed the
+computation. \citet{Rockova2014a} proposed an EM-based algorithm for a
+spike-and-slab Gaussian mixture prior, EMVS. The coefficient \(\beta_j\)
+independently follows a mixture normal distribution
+\(N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)\) for
+\(0 \leq v_0 < v_1\). The parameter \(\sigma^2\) follows a inverse gamma
+prior, \(IG(v/2, v\lambda /2)\). Like in the spike-and slab family, the
+latent binary variables \(\boldsymbol{\gamma}\) follows a Bernoulli
+prior with a hyper prior \(\theta\) follows a beta distribution, or
+simpler the uniform (0,1). In the E step, the latent variable are
+treated as the missing data, and calculate the posterior mean of
+\(\gamma\); in the M step, a ridge estimator was used to update the
+coefficients, and \(\sigma^2, \theta\) are updated accordingly. As the
+ridge regression creates small but nonzero coefficients, a further step
+is needed to select the variable: EMVS further suggests to threshold the
+inclusion probability for variable selection or generate regularization
+plot with different values of the spike scale parameter. EMVS is also
+compatible with structured prior when dealing with grouped variables.
+The group structure can be imposed on the prior of the latent binary
+variable \(\gamma\) via either logistic regression or markov random
+field.
+
+The same authors \citep{Rockova2018b, Rockova2018} further extended the
+continuous spike-and-slab Gaussian prior to the spike-and-slab double
+exponential distribution, as known as spike-and-slab lasso (SSL). The
+authors also lay the theoretical discussion of comparing spike-and-slab
+from the penalty perspective and the Bayesian perspective. The SSL also
+mitigated the problem of EMVS that the weak signals are not shrink to
+zero. Yi and his group incorporated the SSL prior in the EM-IWLS
+algorithm and proposed another EM algorithm, EM-cyclic coordinate
+descent, based on the Lasso penalty SSL exhibit. Such application were
+used in generalized linear models\citep{Tang2017a}, cox models
+\citep{Tang2017} and their grouped
+counterparts\citep{Tang2018, Tang2019}. Both EM based algorithm provide
+deterministic solutions, which becomes a more popular property as
+reproducible research becomes more valued. However, while the EMVS are
+not able to provide inference of the coefficients easily, the EM-IWLS
+can easily extract the variance-covariance matrix due to the IWLS
+algorithm. \citet{Bai2020} provides an up-to-date summary of SSL methods
+and theoretical discussion.
+
+Given the proposed model specification, the estimation of the parameters
+can be easily achieved with MCMC algorithm. Nevertheless, considering
+the computational burden MCMC algorithms bear especially in
+high-dimensional analysis, we propose parsimonious solutions, stemming
+from previous EM-IWLS \citep{Rockova2014a}and EM-coordinate descent
+algorithms\citep{Tang2018, Tang2019}. The computational friendliness
+allows for fitting a model with different sets of initial values,
+mimicking the posterior sampling in the MCMC algorithms. We are hopeful
+that this could improve the sacrificed robustness of the estimation
+process due to the MAP estimation used by the EM algorithms in
+comparison to the model averaging approach.
 
@@
 cost of model uncertainty.
-
-\hfill\break
 
@@
 
+Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
+independently distributed for analytic simplicity. However, this set-up
+could possibility result in a situation that is not theoretically
+possible: the non-linear component is selected, but the linear component
+is not. This can be addressed analytically by forcing to include the
+linear component when non-linear component is included. Another possible
+solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
+\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).
+
 \hypertarget{discussion-1}{%
@@
 if non-linear effect are necessary.
+
+The ``all-in-all-out'' set-up was criticized for its inability to
+answering an important research question, if the effects are non-linear
+or not. On the other hand, the model can be much more flexible by
+allowing each dimension of the penalized space to have an independent
+indicator, i.e.~\(\gamma^{pen}_{jk}\) of \(\beta^{pen}_{jk}\) for
+\(k = 1\dots, K_j\). This prior set-up generalized to the group
+spike-and-slab. The additional flexibility doesn't not necessarily grand
+the model neither theoretical justification nor analytic simplicity. The
+spline penalty assumes a common smooth shrinkage \(\lambda_j\) applied
+to all coefficients of the penalized space, which contradicts to the
+model setup. Besides, having multiple indicator variables complicates
+the decision of if non-linear effects are necessary.
 

### 2021-07-05 Manuscript/02-Method.Rmd
Finish revising method section.

@@
 Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,
-$$
-\begin{aligned}
+
+\begin{align}\label{eq:bham_ssl}
   \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
   \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), 
-\end{aligned}
-$$
+\end{align}
+
 where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the non-linear effect of the $j$th variable respectively.
@@
 Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e.
-$\gamma_j = \gamma^0_j = \gamma^{pen}_j$.
+$\gamma_j = \gamma^0_j = \gamma^\tp_j$.
 This converges^[Is "becomes" better?] to the SB-GAM [@Bai2021].
@@
 The rest of the hierchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution.
-The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^{pen}_j$, shares the same probability parameter,
+The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter,
 $$
@@
 \gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
+\gamma_{j}^\tp | \theta_j &\sim Bin(\gamma^\tp_{j}|1, \theta_j).
 \end{aligned}
@@
 This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function.
-The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^{pen}$ decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity.
+The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
+Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
 We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
@@
 
-\newpage
-
-```{r}
+```{r fig.height=4, fig.width=6}
+# TODO: Refactor the graph to a latex flowchart
 include_graphics("Fig/SS_prior_flow_chart.png")
@@
 
-Considering the dependency of spline bases, the grouping nature of the bases should be taken care of.
-Naturally, the Bayesian hierarchical models allow us easily to take account of the dependency of the spline bases as grouping information and provide more reliable estimation.
-Our hierarchical GAMs specify the spike-and-slab mixture normal prior on the coefficients: for the coefficients of the $j$th predictor, $\beta^0_j$ and $\beta^{pen}_{jk}$, we have
+With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularizations and mixture normal prior for spike-and-slab
+regularization.
+These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
+Here we focus on the mixture normal prior as a demonstration of applying continous spike-and-slab prior in additive models.
+^[TODO: enter a sentence introducing EMVS and spike-and-slab mixture normal prior].
+
+A spike-and-slab mixture normal prior in additive models can be expressed as 
 $$
 \begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
-  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1), 
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\nonumber\\
+  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1).
 \end{aligned}
 $$
-where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^{pen}_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and non-linear effect of the $j$th variable respectively; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
-Thus, the priors are formulated as a mixture of the shrinkage prior normal(0, $s_0$) and the weakly informative prior normal(0, $s_1$), which are spike and slab components respectively.
-Like any other spike and slab priors, the spike density is to contain the minimum to zero effects, while the slab density is to allow large effects.
-The scale parameters $s_0$ and $s_1$ can be treated as tuning parameters, which can be optimized via cross-validation.
-A discussion of how to choose the scale parameters comes later.
-
-To note here, unlike the original spike-and-slab prior where each predictor have only indicator variable, $\gamma_j$, we have two indicators associated with each predictor, $\gamma^0_j$ and $\gamma^{pen}_j$ to decide the inclusion of the linear effect and non-linear effect of the predictor respectively.
-It is possible to add more constraints on the model, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e.
-$\gamma_j = \gamma^0_j = \gamma^{pen}_j$.
-Such formulations, commonly referred as "all-in-all-out", has been implemented in many high dimensional Bayesian GAMs, see Yang & Narisetty (2020) and Bai (Work in progress).
-The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not.
-On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e.
-$\gamma^{pen}_{jk}$ of $\beta^{pen}_{jk}$ for $k = 1\dots, K_j$.
-This prior set-up generalized to the group spike-and-slab.
-The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity.
-The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup.
-Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
-
-We set up the hyper-prior of $\bs \gamma$ to allow local adaption of the shrinkage, using a binomial distribution.
-The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^{pen}_j$, shares the same probability parameter,
-$$
-\begin{aligned}
-\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
-\end{aligned}
-$$
-
-This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function.
-The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^{pen}$ decides the smoothness of the spline function at individual predictor level.
-Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity.
-However, this set-up could possibility result in a situation that is not theoretically possible: the non-linear component is selected, but the linear component is not.
-This can be addressed analytically by forcing to include the linear component when non-linear component is included.
-Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e.
-$\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.
-
- <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function.
-The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder.
-In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.
--->
-
-We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
-$$
-\theta_j \sim Beta(a, b).
-$$
- The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience.
-Specifically, we focus on a special case of beta distribution, uniform (0,1) for simplicity and convenience.
-To note, when the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
-Hereafter, we refer the hierarchical spline GAM with the spike-and-slab mixture normal prior as the ssGAM-MN.
-
+Similar to the BHAM-SSL prior in Equation \ref{eq:bham_ssl}, $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
 One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects.
-In contrast, distributions with heavier tails are preferred, for example $t$ distribution with small degree of freedom and double exponential distribution.
-Specifically, double exponential prior is equivalent to the LASSO penalty in the frequentist models, providing a sparse solution.
+Distributions with heavier tails can be used as an alternative, for example mixture $t$ distribution with small degree of freedom.
 
@@
 
-To alleviate the computation burden, many methodology development studies focus on the maximum a posteriori (MAP) estimate [@Tipping2004].
-EM-based algorithms are prevalent used to speed the computation.
-@Rockova2014a proposed an EM-based algorithm for a spike-and-slab Gaussian mixture prior, EMVS.
-The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
+Given the proposed model specifications, these models can be easily fitted with MCMC algorithms.
+Nevertheless,the computational burden MCMC algorithms bear creates scalability issues.
+George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with continous spike-and-slab priors, and suggested MCMC algorithms works well for medium size of predictors ($p$=25).
+However, it is not feasible to high-dimensionality data analysis where the number of predictors easily exceeds 100.
+Sepcific to additive models, each predictor would expand to mutliple new "predictors" via basis functions, creating greater computational demands.
+@Scheipl2013 demonstrated the computational demands of MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
+Hence, we feel compelled to develop scalable and parsimonious algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
+
+Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures.
+The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a].
+@Rockova2014a proposed an expectation-maximization (EM) based algorithm to fit models that use continous mixture normal priors.
+<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
 The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
 Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).
-In the E step, the latent variable are treated as the missing data, and calculate the posterior mean of $\gamma$; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
-As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
+-->
+In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
 EMVS is also compatible with structured prior when dealing with grouped variables.
 The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
-
-The same authors [@Rockova2018b; @Rockova2018] further extended the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
+-->
+The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coodirnate descent (CD) algoirthm for fitting SSL models.
+<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
 The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
 The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
-Yi and his group incorporated the SSL prior in the EM-IWLS algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
-Such application were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithm provide deterministic solutions, which becomes a more popular property as reproducible research becomes more valued.
-However, while the EMVS are not able to provide inference of the coefficients easily, the EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm.
+-->
+Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
+Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results.
 @Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
 
- Given the proposed model specification, the estimation of the parameters can be easily achieved with MCMC algorithm.
-Nevertheless, considering the computational burden MCMC algorithms bear especially in high-dimensional analysis, we propose parsimonious solutions, stemming from previous EM-IWLS [@Rockova2014a]and EM-coordinate descent algorithms[@Tang2018; @Tang2019].
-The computational friendliness allows for fitting a model with different sets of initial values, mimicking the posterior sampling in the MCMC algorithms.
-We are hopeful that this could improve the sacrificed robustness of the estimation process due to the MAP estimation used by the EM algorithms in comparison to the model averaging approach.
+In the follow section, we expand the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
+To note, the two proposed algorithm provides different utilities.
+The EM-CD algorithm is specifically for fitting BHAM-SSL with an expediated performance, recommending to use in high and ultra-high dimensional setting.
+SB-GAM[@Bai2020; @Bai2021] also used EM-CD algorithm.
+The main difference between proposed EM-CD algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior.
+In comparison, our prior is pairwise independent requiring no speical treatment.
+A sepcifica concern of EM-CD algorithm is that it provides no information for inference.
+In the contrast, the EM-IWLS provides a complement to the problem.
+The EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm, and used to construct Wald type inference.
+Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS pirors but also bayesian ridge and lasso priors.
 
-Parsimonious computation is always encouraged in high-dimension data analysis.
-Normally, Bayesian methodology loses its advantages over Frequentest penalized model because of the computation cost.
-Previous Bayesian spline models heavily relies on the MCMC algorithm to establish posterior distribution of parameters.
-One of the exception is Bai (in progress), who a grouped spike-and-slab LASSO prior [@Rockova2018] to address the grouping structure of spline bases.
-He focused on the MAP estimator of the parameters, but failed to address the uncertainty measure of the estimates.
-The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the non-linear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
-
-In the follow section, we introduced two expectation-maximization (EM) based algorithms for fast computing, EM-IRLS and EM-coordinate Descent algorithm.
+<!-- , stemming from previous EM-coordinate descent algorithms[@Tang2018; @Tang2019] and EM-IWLS [@Rockova2014a].
 Instead of simulating the posterior distribution via MCMC, we iteratively estimate the MAP.
 Compared to the EM-IRLS algorithm, the EM-coordinate descent provides an expedited model fitting process at the cost of model uncertainty.
+-->
+
+<!-- The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the non-linear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
+One of the exception is Bai (in progress), who a grouped spike-and-slab LASSO prior [@Rockova2018] to address the grouping structure of spline bases.
+-->
 
@@
 
-For ssGAM, we define the parameters of interest as $\Theta = {\bs \beta, \bs \theta, \phi}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
-Our objective is to find the parameters $\Theta$ that maximize the posterior density function,or equivalently, the logarithm of the density function, 
+For BHAMs, we define the parameters of interest as $\Theta = {\bs \beta, \bs \theta, \phi}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
+Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
 $$
@@
 $$  
-<!-- For simplicity, we consider the null space to be 1-dimension, and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{pen}$ from the coefficients of the the penalized space.
+For simplicity, we consider the null space to be 1-dimension.
+
+<!-- , and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{pen}$ from the coefficients of the the penalized space.
 Using the new notation, the part component $\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
@@
 We use the EM algorithm to find the MAP estimate of $\Theta$.
-Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as as the "missing data" in the EM algorithms.
+Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithms.
 Naturally In the E-step, we calculate the expectation of posterior density function of $\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{t-1}$,
@@
 Hereafter, we use the short hand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{t-1}}(\cdot)$.
-In the M-step, we find the $\Theta^{t}$ that maximize $E_{\bs \gamma|\Theta^{t-1}}\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
-To note here, the log-posterior of ssGAM (up to additive constants) can be written as a two-part equation
+In the M-step, we find the $\Theta^{t}$ that maximize $E\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
+To note here, the log-posterior density of BHAMs (up to additive constants) can be written as a two-part equation
 
@@
 Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .$$
-$Q_1$ and $Q_2$ are respectively, the log-likelihood of model coefficients and the dispersion parameter $\bs \beta, \phi$, and latent indicators $\bs \gamma, \theta$.
+$Q_1$ and $Q_2$ are, respectively, the log-likelihoods of model coefficients and the dispersion parameter $\bs \beta, \phi$, and latent indicators $\bs \gamma, \theta$.
 Conditioning on $\gamma_{jk}$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
@@
 The E- and M- steps are iterated until the algorithm converge.
-Depending on the choice of the mixture priors, either mixture normal or mixture double exponential, the maximization in the M-step can take different algorithms, IRLS for uncertainty measures and coordinate descent for faster computation.
+Depending on the choice of the mixture priors, the maximization in the M-step can take different algorithms, CD for faster computation and IWLS for uncertainty measures.
 
-\
+### EM-Coordinate Descent
 
-### EM-IRLS
+When the prior distribution is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$, replacing IRLS, in the M-step.
+Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
 
-The main idea of the EM-IRLS algorithm is to use the iterative re-weighted least square algorithm to find the estimate $\bs \beta, \phi$ that maximizes $E(Q_1)$.
-Recall the prior set-up of mixture normal prior, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$
-
+Recall, the density function of spike-and-slab mixture double exponential prior can be written as
 $$
-\begin{aligned}
-p(\bs \beta | \bs\gamma) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],%\\
-% p(\bs\gamma | \bs \theta) & = \prod\limits_{j=1}^{p}\prod\limits_{k=1}^{K_j} \theta_j^{\gamma_{jk}} (1-\theta_j)^{1-\gamma_{jk}}.
-\end{aligned}
+f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
 $$
-and 
-\begin{align*}
-E(Q_1) &= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^{p} \left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\\
-E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
-\end{align*}
-where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1$.
-To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv p(\gamma^{pen}_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem.
+and $E(Q_1)$ can be expressed as a likelihood function with $l_1$ penalty
+\begin{equation}\label{eq:Q1_CD}
+E(Q_1) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+\end{equation}
+where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
+To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv p(\gamma^\tp_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem.
 The calculation for $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\beta_{jk}$ for $k=1, \dots, K_j$ and $p^0_j$ only depends on $\beta_j^0$.
@@
 
-and $Pr(\gamma_{j}^{0} = 1|\theta_j) =Pr(\gamma_{j}^{pen} = 1|\theta_j) = \theta_j$, $Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^{pen} = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{Normal}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{Normal}(\beta|0 , s_0)$.
+and $Pr(\gamma_{j}^{0} = 1|\theta_j) = Pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
@@
 
-With $E({S^0}^{-1}_{j})$ and $E(S^{-1}_{j})$ known, $E(Q_1)$ can be also viewed as a $l_2$ penalized likelihood function: $E({S^0}^{-1}_{j})$ controls the shrinkage of the linear effect, informing if the predictor should be included in the model; $E(S^{-1}_{j})$ controls the smoothness of the spline function.
-$l_2$ penalized likelihood function can be maximize using the IWLS algorithm.
-The likelihood function can be approximate by a weighted normal likelihood:
-$$
-  p(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
-$$
-    where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight, respectively.The pseudo data and the pseudo-weight are calculated by
-    $$
-    \begin{aligned}
-    z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
-    \end{aligned}
-    $$
-    where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
-The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$.
-An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.
-The estimates of $\bs \beta$ and $\phi$ can be quickly updated in the IRLS algorithm.
-
-<!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients.
-Moreover, to note that, we have different shrinkage $S^{-1}_{jk}$for the coefficients $\beta_{jk}$ of the variable $x_j$, and hence, we can penalize the null space of the spline differently and allow local adaption.
--->
-
+As the formulation of $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function, it can be optimized via coordinate descent algorithm.
+After the $\bs \beta$ are updated with the coordinate descent algorithm, we update the hypor parameter $\bs \theta$.
 The remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
 As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed form equation:
-$$
+
+\begin{equation}\label{eq:update_theta}
 \theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
-$$
-
-Totally, the framework of the proposed EM IRLS algorithm was summarized as follows:
-
-1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$
-
-2) Iterate over the E-step and M-step until convergence
-
-E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration
-
-M-step:
-
-a) Update $\bs \beta^{(t)}$ using the IRLS algorithm
-
-b) Update $\bs \theta^{(t)}$ using the closed form equation
-
-We assess convergence by the criterion:
-$|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where
-$d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at
-the $t$th iteration, and $\epsilon$ is a small value (say
-$10^{-5}$).
-
-\
-
-### EM-Coordinate Descent
-
-When the prior distribution is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$, replacing IRLS, in the M-step.
-Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
-
-Recall, the density function of spike-and-slab mixture double exponential prior can be written as
-$$
-f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}).
-$$
-
-This can simplify our posterior likelihood function to
-$$
-E(Q_1) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right]
-,
-$$
-where the approximation of $E({S^0_j}^{-1}$ and $E(S^{-1}_{j})$ remains the same as in the EM-IRLS algorithm.
-The formulation of $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function, and optimized via coordinate descent algorithm.
-The estimates can be used to further estimates $\bs \theta$ in the M-step, following the same equation in the EM-IRLS.
+\end{equation}
 
@@
 
-b) Update $\bs \theta^{(t)}$ using the closed form equation
+b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta}
 
-We assess convergence by the criterion:
-$|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where
-$d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at
-the $t$th iteration, and $\epsilon$ is a small value (say
-$10^{-5}$).
+c) Update the dispersion parameter $\phi$ if exists.
 
- \
+We assess convergence by the criterion: $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at the $t$th iteration, and $\epsilon$ is a small value (say $10^{-5}$).
+
+### EM-IRLS
+
+Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
+It can fit the models with a wider range of priors, e.g.
+mixture $t$ prior, including mixture normal prior as an special case, and mixture double exponential prior.
+The idea is that the mixture priors can be expressed as a hierarchical normal distribution with various forms of disperssion distributions.
+[@Yi2012] Given the hiearchical normal prior of the coefficeints, a genalized model can be fitted using linear approximation and least square algorithm.
+For simplicity, we demonstrate using EM-IWLS algorithm to fit mixture normal priors, with the potential of generalization following @Yi2012.
+
+Recall the prior set-up of mixture normal prior, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$
+
+$$
+\begin{aligned}
+p(\bs \beta | \bs\gamma) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],
+\end{aligned}
+$$
+where $S_{j}^0$ and $S_{j}$ are defined the same as in Equation \ref{eq:Q1_CD}.
+Similarly, $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$ are derived via Bayes' theorem, conditioning on $p^0_j$ and $p_j$.
+We re-write $E(Q_1)$ as a $l_2$ penalized likelihood function, and maximize it using the IWLS algorithm.
+The likelihood function can be approximate by a weighted normal likelihood:
+$$
+  p(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+$$
+    where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight, respectively.The pseudo data and the pseudo-weight are calculated by
+$$
+\begin{aligned}
+z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+\end{aligned}
+$$
+
+  To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an Hessian matrix is yielded as a by-product of the model fitting.
+The Hessian matrix can be further used to construct variance-covariacne matrix of coefficient estimates, and in turn used for statistical inferences.
+
+<!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
+The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$.
+An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.
+The estimates of $\bs \beta$ and $\phi$ can be quickly updated in the IRLS algorithm.
+-->
+
+<!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients.
+Moreover, to note that, we have different shrinkage $S^{-1}_{jk}$for the coefficients $\beta_{jk}$ of the variable $x_j$, and hence, we can penalize the null space of the spline differently and allow local adaption.
+-->
+
+<!-- Totally, the framework of the proposed EM IRLS algorithm was summarized as follows: -->
+
+<!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
+
+<!-- 2) Iterate over the E-step and M-step until convergence -->
+
+<!-- E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
+
+<!-- M-step: -->
+
+<!-- a) Update $\bs \beta^{(t)}$ using the IRLS algorithm -->
+
+<!-- b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta} -->
+
+<!-- We assess convergence by the criterion: -->
+<!-- $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where -->
+<!-- $d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at -->
+<!-- the $t$th iteration, and $\epsilon$ is a small value (say -->
+<!-- $10^{-5}$). -->
 
@@
 
-Our proposed prior, spike and slab spline prior, requires two preset scale parameters ($s_0$, $s_1$).
+Our proposed models, BHAM, require two preset scale parameters ($s_0$, $s_1$).
 Hence, we would construct a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
-However, previous research on the topic^[TODO: add citation] suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value of $s_1 = 1$ that provides no or weak shrinkage.
+However, previous research on the topic^[TODO: add citations] suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value of $s_1 = 1$ that provides no or weak shrinkage.
 More attention is focused on examining different values of spike scale $s_0$.

### 2021-07-05 Manuscript/SS_GAM.tex
Finish revising method section.

@@
 independent group priors on the null space coefficients and on the
-penalized space coefficients respectively, \[
-\begin{aligned}
+penalized space coefficients respectively,
+
+\begin{align}\label{eq:bham_ssl}
   \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
   \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1), 
-\end{aligned}
-\] where \(\gamma^0_{j}\in\{0,1\}\) and
+\end{align}
+
+where \(\gamma^0_{j}\in\{0,1\}\) and
 \(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator
@@
 variable decides the inclusion of both the linear effect and non-linear
-effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). This
+effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^\text{pen}_j\). This
 converges\footnote{Is ``becomes'' better?} to the SB-GAM
@@
 The two indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
-\(\gamma^{pen}_j\), shares the same probability parameter, \[
+\(\gamma^\text{pen}_j\), shares the same probability parameter, \[
 \begin{aligned}
 \gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
+\gamma_{j}^\text{pen}| \theta_j &\sim Bin(\gamma^\text{pen}_{j}|1, \theta_j).
 \end{aligned}
@@
 prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
-function selection level, while that of \(\gamma_{j}^{pen}\) decides the
-smoothness of the spline function at basis function level. Meanwhile, we
-specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are independently
-distributed for analytic simplicity. We further specify the parameter
-\(\theta_j\) follows a beta distribution with given shape parameters
-\(a\) and \(b\), \[
+function selection level, while that of \(\gamma_{j}^\text{pen}\)
+decides the smoothness of the spline function at basis function level.
+Meanwhile, we specify that \(\gamma_{j}^0\) and
+\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
+simplicity. We further specify the parameter \(\theta_j\) follows a beta
+distribution with given shape parameters \(a\) and \(b\), \[
 \theta_j \sim Beta(a, b).
@@
 
-\newpage
-
 \includegraphics{Fig/SS_prior_flow_chart.png}
@@
 
-Considering the dependency of spline bases, the grouping nature of the
-bases should be taken care of. Naturally, the Bayesian hierarchical
-models allow us easily to take account of the dependency of the spline
-bases as grouping information and provide more reliable estimation. Our
-hierarchical GAMs specify the spike-and-slab mixture normal prior on the
-coefficients: for the coefficients of the \(j\)th predictor,
-\(\beta^0_j\) and \(\beta^{pen}_{jk}\), we have \[
+With the re-parameterization step of the basis function matrix \(X\), it
+is possible to generalized the SSL prior to other priors, for example
+normal priors for ridge-type regularizations and mixture normal prior
+for spike-and-slab regularization. These priors would work better in low
+and medium dimensional settings where the sparse assumption is not
+necessary. Here we focus on the mixture normal prior as a demonstration
+of applying continous spike-and-slab prior in additive models.
+\footnote{TODO: enter a sentence introducing EMVS and spike-and-slab
+  mixture normal prior}.
+
+A spike-and-slab mixture normal prior in additive models can be
+expressed as \[
 \begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
-  \beta^{pen}_{jk} | \gamma^{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1), 
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\nonumber\\
+  \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1).
 \end{aligned}
-\] where \(\gamma^0_{j}\in\{0,1\}\) and
-\(\gamma^{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator variables,
-indicating if the model includes the linear effect and non-linear effect
-of the \(j\)th variable respectively; \(s_0\) and \(s_1\) are scale
-parameters, assuming \(0 < s_0 < s_1\) and given. Thus, the priors are
-formulated as a mixture of the shrinkage prior normal(0, \(s_0\)) and
-the weakly informative prior normal(0, \(s_1\)), which are spike and
-slab components respectively. Like any other spike and slab priors, the
-spike density is to contain the minimum to zero effects, while the slab
-density is to allow large effects. The scale parameters \(s_0\) and
-\(s_1\) can be treated as tuning parameters, which can be optimized via
-cross-validation. A discussion of how to choose the scale parameters
-comes later.
-
-To note here, unlike the original spike-and-slab prior where each
-predictor have only indicator variable, \(\gamma_j\), we have two
-indicators associated with each predictor, \(\gamma^0_j\) and
-\(\gamma^{pen}_j\) to decide the inclusion of the linear effect and
-non-linear effect of the predictor respectively. It is possible to add
-more constraints on the model, assuming that one indicator variable
-decides the inclusion of both the linear effect and non-linear effect,
-i.e.~\(\gamma_j = \gamma^0_j = \gamma^{pen}_j\). Such formulations,
-commonly referred as ``all-in-all-out'', has been implemented in many
-high dimensional Bayesian GAMs, see Yang \& Narisetty (2020) and Bai
-(Work in progress). The ``all-in-all-out'' set-up was criticized for its
-inability to answering an important research question, if the effects
-are non-linear or not. On the other hand, the model can be much more
-flexible by allowing each dimension of the penalized space to have an
-independent indicator, i.e.~\(\gamma^{pen}_{jk}\) of
-\(\beta^{pen}_{jk}\) for \(k = 1\dots, K_j\). This prior set-up
-generalized to the group spike-and-slab. The additional flexibility
-doesn't not necessarily grand the model neither theoretical
-justification nor analytic simplicity. The spline penalty assumes a
-common smooth shrinkage \(\lambda_j\) applied to all coefficients of the
-penalized space, which contradicts to the model setup. Besides, having
-multiple indicator variables complicates the decision of if non-linear
-effects are necessary.
-
-We set up the hyper-prior of \(\boldsymbol{\gamma}\) to allow local
-adaption of the shrinkage, using a binomial distribution. The two
-indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
-\(\gamma^{pen}_j\), shares the same probability parameter, \[
-\begin{aligned}
-\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^{pen} | \theta_j &\sim Bin(\gamma^{pen}_{j}|1, \theta_j).
-\end{aligned}
-\]
-
-This is to leverage the fact that the probability of selecting the bases
-of a smooth function should be similar, while allowing different penalty
-on the null space and penalty space of the spline function. The hyper
-prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
-function selection level, while that of \(\gamma_{j}^{pen}\) decides the
-smoothness of the spline function at individual predictor level.
-Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
-independently distributed for analytic simplicity. However, this set-up
-could possibility result in a situation that is not theoretically
-possible: the non-linear component is selected, but the linear component
-is not. This can be addressed analytically by forcing to include the
-linear component when non-linear component is included. Another possible
-solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
-\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).
-
-We further specify the parameter \(\theta_j\) follows a beta
-distribution with given shape parameters \(a\) and \(b\), \[
-\theta_j \sim Beta(a, b).
-\] The beta distribution is a conjugate prior for the binomial
-distribution and hence provides some computation convenience.
-Specifically, we focus on a special case of beta distribution, uniform
-(0,1) for simplicity and convenience. To note, when the variable have
-large effects in any of the bases, the parameter \(\theta_j\) will be
-estimated large, which in turn encourages the model to include the rest
-of bases, achieving the local adaption among spline bases. Hereafter, we
-refer the hierarchical spline GAM with the spike-and-slab mixture normal
-prior as the ssGAM-MN.
-
-One of the critics received by the spike-and-slab mixture normal prior
-is that the tails of a normal distribution diminishes to zero too fast,
-which renders problem when estimating the large effects. In contrast,
-distributions with heavier tails are preferred, for example \(t\)
-distribution with small degree of freedom and double exponential
-distribution. Specifically, double exponential prior is equivalent to
-the LASSO penalty in the frequentist models, providing a sparse
-solution.
+\] Similar to the BHAM-SSL prior in Equation \ref{eq:bham_ssl},
+\(0 < s_0 < s_1\) are tuning parameters and can be optimized via
+cross-validation. One of the critics received by the spike-and-slab
+mixture normal prior is that the tails of a normal distribution
+diminishes to zero too fast, which renders problem when estimating the
+large effects. Distributions with heavier tails can be used as an
+alternative, for example mixture \(t\) distribution with small degree of
+freedom.
 
@@
 
-To alleviate the computation burden, many methodology development
-studies focus on the maximum a posteriori (MAP) estimate
-\citep{Tipping2004}. EM-based algorithms are prevalent used to speed the
-computation. \citet{Rockova2014a} proposed an EM-based algorithm for a
-spike-and-slab Gaussian mixture prior, EMVS. The coefficient \(\beta_j\)
-independently follows a mixture normal distribution
-\(N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)\) for
-\(0 \leq v_0 < v_1\). The parameter \(\sigma^2\) follows a inverse gamma
-prior, \(IG(v/2, v\lambda /2)\). Like in the spike-and slab family, the
-latent binary variables \(\boldsymbol{\gamma}\) follows a Bernoulli
-prior with a hyper prior \(\theta\) follows a beta distribution, or
-simpler the uniform (0,1). In the E step, the latent variable are
-treated as the missing data, and calculate the posterior mean of
-\(\gamma\); in the M step, a ridge estimator was used to update the
-coefficients, and \(\sigma^2, \theta\) are updated accordingly. As the
-ridge regression creates small but nonzero coefficients, a further step
-is needed to select the variable: EMVS further suggests to threshold the
-inclusion probability for variable selection or generate regularization
-plot with different values of the spike scale parameter. EMVS is also
-compatible with structured prior when dealing with grouped variables.
-The group structure can be imposed on the prior of the latent binary
-variable \(\gamma\) via either logistic regression or markov random
-field.
-
-The same authors \citep{Rockova2018b, Rockova2018} further extended the
-continuous spike-and-slab Gaussian prior to the spike-and-slab double
-exponential distribution, as known as spike-and-slab lasso (SSL). The
-authors also lay the theoretical discussion of comparing spike-and-slab
-from the penalty perspective and the Bayesian perspective. The SSL also
-mitigated the problem of EMVS that the weak signals are not shrink to
-zero. Yi and his group incorporated the SSL prior in the EM-IWLS
-algorithm and proposed another EM algorithm, EM-cyclic coordinate
-descent, based on the Lasso penalty SSL exhibit. Such application were
-used in generalized linear models\citep{Tang2017a}, cox models
-\citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithm provide
+Given the proposed model specifications, these models can be easily
+fitted with MCMC algorithms. Nevertheless,the computational burden MCMC
+algorithms bear creates scalability issues. George and
+McCulloch\citep{George1997} examined the computation speed for various
+MCMC algorithms with continous spike-and-slab priors, and suggested MCMC
+algorithms works well for medium size of predictors (\(p\)=25). However,
+it is not feasible to high-dimensionality data analysis where the number
+of predictors easily exceeds 100. Sepcific to additive models, each
+predictor would expand to mutliple new ``predictors'' via basis
+functions, creating greater computational demands. \citet{Scheipl2013}
+demonstrated the computational demands of MCMC algorithm for fitting
+spike-and-slab additive models grow exponentially as \(p\) increases
+modestly via monte carlo simulation studies. Hence, we feel compelled to
+develop scalable and parsimonious algorithms for fitting Bayesian
+hierarchical additive models in high-dimensional settings.
+
+Alternative to the sampling algorithms for fitting Bayesian models,
+optimization algorithms focus on the maximum a posteriori (MAP)
+estimates and speed up the process at the cost of uncertainty measures.
+The earliest work for fitting spike-and-slab models using optimization
+algorithm is EMVS \citep{Rockova2014a}. \citet{Rockova2014a} proposed an
+expectation-maximization (EM) based algorithm to fit models that use
+continous mixture normal priors. In the E step, the latent variables
+\(\gamma\) are treated as the missing data, and calculate the posterior
+mean of the density function; in the M step, a ridge estimator was used
+to update the coefficients, and \(\sigma^2, \theta\) are updated
+accordingly. The same authors \citep{Rockova2018b, Rockova2018} further
+combined the EM algorithm with coodirnate descent (CD) algoirthm for
+fitting SSL models. Yi and his group incorporated the SSL prior in the
+EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another
+EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty
+SSL exhibit. Such applications were used in generalized linear
+models\citep{Tang2017a}, cox models \citep{Tang2017} and their grouped
+counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
 deterministic solutions, which becomes a more popular property as
-reproducible research becomes more valued. However, while the EMVS are
-not able to provide inference of the coefficients easily, the EM-IWLS
-can easily extract the variance-covariance matrix due to the IWLS
-algorithm. \citet{Bai2020} provides an up-to-date summary of SSL methods
-and theoretical discussion.
-
-Given the proposed model specification, the estimation of the parameters
-can be easily achieved with MCMC algorithm. Nevertheless, considering
-the computational burden MCMC algorithms bear especially in
-high-dimensional analysis, we propose parsimonious solutions, stemming
-from previous EM-IWLS \citep{Rockova2014a}and EM-coordinate descent
-algorithms\citep{Tang2018, Tang2019}. The computational friendliness
-allows for fitting a model with different sets of initial values,
-mimicking the posterior sampling in the MCMC algorithms. We are hopeful
-that this could improve the sacrificed robustness of the estimation
-process due to the MAP estimation used by the EM algorithms in
-comparison to the model averaging approach.
-
-Parsimonious computation is always encouraged in high-dimension data
-analysis. Normally, Bayesian methodology loses its advantages over
-Frequentest penalized model because of the computation cost. Previous
-Bayesian spline models heavily relies on the MCMC algorithm to establish
-posterior distribution of parameters. One of the exception is Bai (in
-progress), who a grouped spike-and-slab LASSO prior \citep{Rockova2018}
-to address the grouping structure of spline bases. He focused on the MAP
-estimator of the parameters, but failed to address the uncertainty
-measure of the estimates. The main difference between Bai's work and the
-proposed models are two-folded: 1) the proposed models are feasible to
-answer the question if the non-linear effects are necessary for the
-predictors, while Bai's work can't due to the ``all-in-all-out''
-approach; 2) the proposed models can provide uncertainly measures in
-addition to the point estimates of the coefficients, while Bai's work
-can't not due to the model fitting algorithm.
-
-In the follow section, we introduced two expectation-maximization (EM)
-based algorithms for fast computing, EM-IRLS and EM-coordinate Descent
-algorithm. Instead of simulating the posterior distribution via MCMC, we
-iteratively estimate the MAP. Compared to the EM-IRLS algorithm, the
-EM-coordinate descent provides an expedited model fitting process at the
-cost of model uncertainty.
+reproducible results. \citet{Bai2020} provides an up-to-date summary of
+SSL methods and theoretical discussion.
+
+In the follow section, we expand the two EM-based algorithms, EM-CD and
+EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithm
+provides different utilities. The EM-CD algorithm is specifically for
+fitting BHAM-SSL with an expediated performance, recommending to use in
+high and ultra-high dimensional setting. SB-GAM\citep{Bai2020, Bai2021}
+also used EM-CD algorithm. The main difference between proposed EM-CD
+algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for
+their group prior. In comparison, our prior is pairwise independent
+requiring no speical treatment. A sepcifica concern of EM-CD algorithm
+is that it provides no information for inference. In the contrast, the
+EM-IWLS provides a complement to the problem. The EM-IWLS can easily
+extract the variance-covariance matrix due to the IWLS algorithm, and
+used to construct Wald type inference. Moreover, the EM-IWLS is a more
+general model fitting algorithm that can be used for fitting not only
+SSL and continuous SS pirors but also bayesian ridge and lasso priors.
 
@@
 
-For ssGAM, we define the parameters of interest as
+For BHAMs, we define the parameters of interest as
 \(\Theta = {\boldsymbol{\beta}, \boldsymbol{\theta}, \phi}\) and
@@
 nuisance parameters of the model. Our objective is to find the
-parameters \(\Theta\) that maximize the posterior density function,or
+parameters \(\Theta\) that maximize the posterior density function, or
 equivalently, the logarithm of the density function, \[
@@
 \]\\
+For simplicity, we consider the null space to be 1-dimension.
 
@@
 the latent binary indicators \(\boldsymbol{\gamma}\) are not of our
-primary interest, we treat them as as the ``missing data'' in the EM
+primary interest, we treat them as the ``missing data'' in the EM
 algorithms. Naturally In the E-step, we calculate the expectation of
@@
 M-step, we find the \(\Theta^{t}\) that maximize
-\(E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\).
-To note here, the log-posterior of ssGAM (up to additive constants) can
-be written as a two-part equation
+\(E\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). To
+note here, the log-posterior density of BHAMs (up to additive constants)
+can be written as a two-part equation
 
@@
 Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .\]
-\(Q_1\) and \(Q_2\) are respectively, the log-likelihood of model
+\(Q_1\) and \(Q_2\) are, respectively, the log-likelihoods of model
 coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\),
@@
 The E- and M- steps are iterated until the algorithm converge. Depending
-on the choice of the mixture priors, either mixture normal or mixture
-double exponential, the maximization in the M-step can take different
-algorithms, IRLS for uncertainty measures and coordinate descent for
-faster computation.
-
-\hfill\break
-
-\hypertarget{em-irls}{%
-\subsubsection{EM-IRLS}\label{em-irls}}
-
-The main idea of the EM-IRLS algorithm is to use the iterative
-re-weighted least square algorithm to find the estimate
-\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). Recall the prior
-set-up of mixture normal prior, we have the prior densities of
-\(\boldsymbol{\beta}\) conditioning on \(\boldsymbol{\gamma}\)
-
-\[
-\begin{aligned}
-p(\boldsymbol{\beta }| \boldsymbol{\gamma}) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],%\\
-% p(\bs\gamma | \bs \theta) & = \prod\limits_{j=1}^{p}\prod\limits_{k=1}^{K_j} \theta_j^{\gamma_{jk}} (1-\theta_j)^{1-\gamma_{jk}}.
-\end{aligned}
-\] and \begin{align*}
-E(Q_1) &= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^{p} \left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\\
-E(Q_2) &= \sum\limits_{j=1}^{p}\sum\limits_{k=1}^{K_j} \left[ E(\gamma_{jk})\log \theta_j + (1-E(\gamma_{jk})) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
-\end{align*} where
+on the choice of the mixture priors, the maximization in the M-step can
+take different algorithms, CD for faster computation and IWLS for
+uncertainty measures.
+
+\hypertarget{em-coordinate-descent}{%
+\subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
+
+When the prior distribution is set to mixture double exponential,
+coordinate descent algorithm can be used to estimate
+\(\boldsymbol{b}eta\), replacing IRLS, in the M-step. Coordinate descent
+is an optimization algorithm that offers extreme computational
+advantage, and famous for its application in optimizing the \(l_1\)
+penalized likelihood function.
+
+Recall, the density function of spike-and-slab mixture double
+exponential prior can be written as \[
+f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
+\] and \(E(Q_1)\) can be expressed as a likelihood function with \(l_1\)
+penalty \begin{equation}\label{eq:Q1_CD}
+E(Q_1) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+\end{equation} where
 \(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S_{j} = (1-\gamma^{pen}_{j}) s_0 + \gamma^{pen}_{j} s_1\). To
-calculate two unknown quantities \(E({S_j^0}^{-1})\) and
+\(S_{j} = (1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1\).
+To calculate two unknown quantities \(E({S_j^0}^{-1})\) and
 \(E(S^{-1}_j)\), the posterior probability
 \(p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
-\(p_{j} \equiv p(\gamma^{pen}_{j}=1|\Theta^{t-1})\) are necessary, which
-can be derived via Bayes' theorem. The calculation for \(p_j\) is
+\(p_{j} \equiv p(\gamma^\text{pen}_{j}=1|\Theta^{t-1})\) are necessary,
+which can be derived via Bayes' theorem. The calculation for \(p_j\) is
 slightly different from that of \(p^0_j\), as \(p_j\) depends on the
@@
 and
-\(Pr(\gamma_{j}^{0} = 1|\theta_j) =Pr(\gamma_{j}^{pen} = 1|\theta_j) = \theta_j\),
-\(Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^{pen} = 0|\theta_j) = 1-\theta_j\),
-\(f(\beta|\gamma=1, s_1) = \text{Normal}(\beta|0 , s_1)\),
-\(f(\beta|\gamma=0, s_0) = \text{Normal}(\beta|0 , s_0)\). It is trivial
-to show \[
+\(Pr(\gamma_{j}^{0} = 1|\theta_j) = Pr(\gamma_{j}^\text{pen}= 1|\theta_j) = \theta_j\),
+\(Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^\text{pen}= 0|\theta_j) = 1-\theta_j\),
+\(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
+\(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
+show \[
 \begin{aligned}
@@
 
-With \(E({S^0}^{-1}_{j})\) and \(E(S^{-1}_{j})\) known, \(E(Q_1)\) can
-be also viewed as a \(l_2\) penalized likelihood function:
-\(E({S^0}^{-1}_{j})\) controls the shrinkage of the linear effect,
-informing if the predictor should be included in the model;
-\(E(S^{-1}_{j})\) controls the smoothness of the spline function.
-\(l_2\) penalized likelihood function can be maximize using the IWLS
-algorithm. The likelihood function can be approximate by a weighted
-normal likelihood: \[
-  p(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
-\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
-the pseudo-response and pseudo-weight, respectively.The pseudo data and
-the pseudo-weight are calculated by \[
-    \begin{aligned}
-    z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
-    \end{aligned}
-    \] where \(\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})\) is
-the estimated linear predictor using current estimated value of
-\(\boldsymbol{\beta}\), \(L^{'}\) and \(L^{''}\) are the first
-derivative and second derivative of the log-likelihood function. The
-\(l_2\) penalty,
-\(\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]\)
-can be seems as a multivariate normal distribution of mean
-\(\textbf{0}\) and variance-covariance matrix
-\(\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})\),
-conditioning on the value of \(E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})\).
-An augmentation step combines the normal approximation and multivariate
-normal for \(\boldsymbol{\beta}\), similarly to Ridge regression model
-fitting. The estimates of \(\boldsymbol{\beta}\) and \(\phi\) can be
-quickly updated in the IRLS algorithm.
-
-The remaining parameters of interest \(\boldsymbol{\theta}\) can be
-updated by maximizing \(E(Q_2)\). As the beta distribution is a
-conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
-be easily updated with a closed form equation: \[
+As the formulation of \(E(Q_1)\) can be seen as a \(l_1\) penalized
+likelihood function, it can be optimized via coordinate descent
+algorithm. After the \(\boldsymbol{\beta}\) are updated with the
+coordinate descent algorithm, we update the hypor parameter
+\(\boldsymbol{\theta}\). The remaining parameters of interest
+\(\boldsymbol{\theta}\) can be updated by maximizing \(E(Q_2)\). As the
+beta distribution is a conjugate prior for Bernoulli distribution,
+\(\boldsymbol{\theta}\) can be easily updated with a closed form
+equation:
+
+\begin{equation}\label{eq:update_theta}
 \theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
-\]
-
-Totally, the framework of the proposed EM IRLS algorithm was summarized
-as follows:
+\end{equation}
+
+Totally, the framework of the proposed EM-coordinate descent algorithm
+was summarized as follows:
 
@@
 \item
-  Update \(\boldsymbol{\beta}^{(t)}\) using the IRLS algorithm
+  Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
+  algorithm
 \item
-  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
+  Update \(\boldsymbol{\theta}^{(t)}\) using Equation
+  \ref{eq:update_theta}
+\item
+  Update the dispersion parameter \(\phi\) if exists.
 \end{enumerate}
@@
 
-\hfill\break
-
-\hypertarget{em-coordinate-descent}{%
-\subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
-
-When the prior distribution is set to mixture double exponential,
-coordinate descent algorithm can be used to estimate
-\(\boldsymbol{b}eta\), replacing IRLS, in the M-step. Coordinate descent
-is an optimization algorithm that offers extreme computational
-advantage, and famous for its application in optimizing the \(l_1\)
-penalized likelihood function.
-
-Recall, the density function of spike-and-slab mixture double
-exponential prior can be written as \[
-f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}).
+\hypertarget{em-irls}{%
+\subsubsection{EM-IRLS}\label{em-irls}}
+
+Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative
+EM-based algorithm where iterative weighted least square algorithm are
+used to find the estimate \(\boldsymbol{\beta}, \phi\) that maximizes
+\(E(Q_1)\), instead of the coordinate descent algorithm. It can fit the
+models with a wider range of priors, e.g.~mixture \(t\) prior, including
+mixture normal prior as an special case, and mixture double exponential
+prior. The idea is that the mixture priors can be expressed as a
+hierarchical normal distribution with various forms of disperssion
+distributions. \citep{Yi2012} Given the hiearchical normal prior of the
+coefficeints, a genalized model can be fitted using linear approximation
+and least square algorithm. For simplicity, we demonstrate using EM-IWLS
+algorithm to fit mixture normal priors, with the potential of
+generalization following \citet{Yi2012}.
+
+Recall the prior set-up of mixture normal prior, we have the prior
+densities of \(\boldsymbol{\beta}\) conditioning on
+\(\boldsymbol{\gamma}\)
+
+\[
+\begin{aligned}
+p(\boldsymbol{\beta }| \boldsymbol{\gamma}) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],
+\end{aligned}
+\] where \(S_{j}^0\) and \(S_{j}\) are defined the same as in Equation
+\ref{eq:Q1_CD}. Similarly, \(E({S_j^0}^{-1})\) and \(E(S^{-1}_j)\) are
+derived via Bayes' theorem, conditioning on \(p^0_j\) and \(p_j\). We
+re-write \(E(Q_1)\) as a \(l_2\) penalized likelihood function, and
+maximize it using the IWLS algorithm. The likelihood function can be
+approximate by a weighted normal likelihood: \[
+  p(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
+the pseudo-response and pseudo-weight, respectively.The pseudo data and
+the pseudo-weight are calculated by \[
+\begin{aligned}
+z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+\end{aligned}
 \]
 
-This can simplify our posterior likelihood function to \[
-E(Q_1) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right]
-,
-\] where the approximation of \(E({S^0_j}^{-1}\) and \(E(S^{-1}_{j})\)
-remains the same as in the EM-IRLS algorithm. The formulation of
-\(E(Q_1)\) can be seen as a \(l_1\) penalized likelihood function, and
-optimized via coordinate descent algorithm. The estimates can be used to
-further estimates \(\boldsymbol{\theta}\) in the M-step, following the
-same equation in the EM-IRLS.
-
-Totally, the framework of the proposed EM-coordinate descent algorithm
-was summarized as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
-  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
-  \(\boldsymbol{\theta}\). For example, we can initialize
-  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
-  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
-\item
-  Iterate over the E-step and M-step until convergence
-\end{enumerate}
-
-E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
-\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
-\(\Theta^{(t-1)}\) from previous iteration
-
-M-step:
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi})}
-\item
-  Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
-  algorithm
-\item
-  Update \(\boldsymbol{\theta}^{(t)}\) using the closed form equation
-\end{enumerate}
-
-We assess convergence by the criterion:
-\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
-\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
-the \(t\)th iteration, and \(\epsilon\) is a small value (say
-\(10^{-5}\)).
-
-~
+To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an
+Hessian matrix is yielded as a by-product of the model fitting. The
+Hessian matrix can be further used to construct variance-covariacne
+matrix of coefficient estimates, and in turn used for statistical
+inferences.
 
@@
 
-Our proposed prior, spike and slab spline prior, requires two preset
-scale parameters (\(s_0\), \(s_1\)). Hence, we would construct a two
-dimensional grid, consists of different pairs of (\(s_0\), \(s_1\))
-value. However, previous research on the topic\footnote{TODO: add
-  citation} suggested the value of slab scale \(s_1\) have less impact
-on the final model and is recommended to be set as a generally large
-value of \(s_1 = 1\) that provides no or weak shrinkage. More attention
-is focused on examining different values of spike scale \(s_0\). Instead
-of the 2-D grid, We consider a sequence of \(L\) decreasing values
+Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
+\(s_1\)). Hence, we would construct a two dimensional grid, consists of
+different pairs of (\(s_0\), \(s_1\)) value. However, previous research
+on the topic\footnote{TODO: add citation} suggested the value of slab
+scale \(s_1\) have less impact on the final model and is recommended to
+be set as a generally large value of \(s_1 = 1\) that provides no or
+weak shrinkage. More attention is focused on examining different values
+of spike scale \(s_0\). Instead of the 2-D grid, We consider a sequence
+of \(L\) decreasing values
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the

### 2021-07-20 Manuscript/01-Intro.Rmd
Correct AMA in-text citation

--- 2021-06-29 Manuscript/01-Intro.Rmd
@@
 # Introduction
+\label{sec:intro}
 
 Using additive models to interpolate nonlinear effects has a long history.
-Early work of additive models can date back to 1979 when \citet{Cleveland1979} proposed local smoothing for bivariate analysis.
-Since then, many mathematical and statistical models have been proposed, see @Wood2020.
-With the recent advancement of computational technology, many machine learning algorithms that are computationally intensive start to gain much popularity in prognostic modeling and functional selection, especially when complex signals, for example nonlinear effects, are assumed.
+Early work of additive models can date back to 1979 when Cleveland [@Cleveland1979] proposed local smoothing for bivariate analysis.
+Since then, many mathematical and statistical models have been proposed.
+[@Wood2020] With the recent advancement of computational technology, many machine learning algorithms that are computationally intensive start to gain much popularity in prognostic modeling and functional selection, especially when complex signals, for example nonlinear effects, are assumed.
 Such models generally have improved prediction accuracy over classic statistical models.
-However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models.
+However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models [@Breiman2001].
 Given that the primary objective of biomedical studies is to understand the etiology of diseases, "Black box" models are less helpful as they can hardly explain how the change of biomarkers is associated with the disease.
@@
 Semi-parametric regression models are a more appropriate alternative to the "black-box" models, thanks to their balancing between model flexibility and interpretability.
-Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most utilized modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu.
-\phi)$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
@@
 
-Many applications of modern biomedical research, e.g.
+Many modern biomedical research, e.g.
 sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
-There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see @Mallick2013.
+There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review.
 These models assume sparsity among the predictors, i.e.
@@
 Further development focuses on addressing the grouping structure among the predictors, where AMs and GAMs can be considered special cases.
-Various penalties are imposed on the basis function coefficients: @Ravikumar2009 extended the grouped lasso [@Yuan2006] to additive models; @Huang2010 further developed adaptive grouped lasso for additive models; @Wang2007 and @Xue2009 applied grouped SCAD penalty [@Fan2001] to the additive models.
+Various penalties are imposed on the basis function coefficients: Ravikumar et al.
+[@Ravikumar2009] extended the grouped lasso [@Yuan2006] to additive models; Huang et al.
+[@Huang2010] further developed adaptive grouped lasso for additive models; Wang et al.
+[@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to the additive models.
 Recent Bayesian hierarchical models are also used in the context of high-dimensional additive models.
-Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed where the application on AMs are by-products.
-@Scheipl2012 proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma, specifically for semi-parametric regression using MCMC.
+Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.
+Scheipl et al.
+[@Scheipl2012] proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma, specifically for semi-parametric regression using MCMC.
 The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
-@Bai2020 was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs in @Bai2021.
+Bai et al.
+[@Bai2020] was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
 Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.
-Besides @Schipel2012, most proposed methods adapt an 'all-in-all-out' strategy, i.e.
+Besides Scheipl et al.
+[@Scheipl2012], most proposed methods adapt an 'all-in-all-out' strategy, i.e.
 either including or excluding the variable completely, rendering no space for within group selection.
@@
 The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
-Additionally, two fast computing optimization-based algorithms are developed for high and ultra-high dimension settings respectively.
+Additionally, two fast computing optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
 Meanwhile, to avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
@@
 Moreover, incorporating the smoothness penalty allows the separation of the linear space of the smooth function, also know as the null space, from the nonlinear space.
-@Marra2011 introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
+Marra and Wood [@Marra2011] introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
 The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
-^[TODO: a sentence that it decompose the prior to independent priors, provide obvious computational convenience.] The computational convenience improved the scalability of the model.
+In addition, the priors of the linear and nonlinear effects are independent, creating computational convenience and further improving the scalability of the model.
 
@@
 First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM.
-Spike-and-slab lasso priors provide computational convenience during model fitting through optimization algorithms, instead of computational intensive sampling algorithms.
-They make fitting high- and ultrahigh-dimensional models feasible without sacrificing performance in prediction and variable selection.
+Spike-and-slab lasso priors provide computational convenience during model fitting through optimization algorithms versus computational intensive sampling algorithms.
+They make fitting high- and ultrahigh-dimensional models more feasible without sacrificing performance in prediction and variable selection.
 Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent algorithm and EM-Iterative weighted least square algorithm for different utilities.
-The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis, see ^[TODO: add citations].
-Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.
-EM-IWLS complements the lack of inference potential from EM-CD in a median and high data setting.
+The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis.
+[@Yi2012; @Rockova2014a; @Rockova2018] Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.
+EM-IWLS complements the lack of inference potential from EM-CD in a median and high dimensional setting.
 Furthermore, the proposed model addresses the incapability of bilevel selection in SB-GAM.
@@
 Conclusion and discussions are made in Chapter 5.
+^[TODO: rewrite this sentence.]
 

### 2021-07-20 Manuscript/02-Method.Rmd
Correct AMA in-text citation

@@
 # Bayesian Hierarchical Additive Models
+\label{sec:BHAM}
 
-Following the GLM notation introduced in Section \label{introduction}, we have a generalized additive model and its matrix form,
+Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model and its matrix form,
 \begin{equation}\label{eq:gam}
@@
 tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization.
-The re-parameterization follows @Marra2011, in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+The re-parameterization follows Marra and  Wood [@Marra2011], in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
 To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function.
@@
 We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$.
-An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see @Wood2006), but I am not sure if it can be generalized to all smoothing functions].
+An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see [@Wood2006]), but I am not sure if it can be generalized to all smoothing functions].
 For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter.
@@
 The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection.
-We defer to @Bai2021Review for an in-depth introduction to spike-and-slab priors.
+We defer to Bai et al.
+[@Bai2021Review] for an in-depth introduction to spike-and-slab priors.
 To summary, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
@@
 The most distinct feature of the SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, for example @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004 and reference therein.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
+[@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.
 
@@
 The major criticism of early spike-and-slab priors receives is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a], ^[TODO: add more citation].
-Particularly, the development of Spike-and-slab Lasso [@Rockova2018b; @Rockova2018] substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis ^[TODO: add Dr.
-Yi's papers].
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and Spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
+Particularly, the development of Spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
 The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
@@
 The SSL also mitigated the problem of EMVS where the weak signals are not shrink to zero, favored in high-dimensional data analysis.
-We noticed that @Bai2021 introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
-@Bai2021 takes an "all-in-all-out" strategy for function selection.
+We noticed that Bai [@Bai2021] introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
+Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.
 
@@
 These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation.
-A discussion of how to choose the scale parameters comes in Section \ref{selecting-optimal-scale-values}.
-To note, this prior differs from previous group spike-and-slab lasso priors[TODO: add Dr.
-Yi's citation], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively.
+A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}.
+To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively.
 It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection.
@@
 
-The reparameterization introduced in Section \ref{bayesian-hierarchical-additive-models} grants the validity of the proposed prior.
+The reparameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
 First of all, the smooth function bases are linear dependent to each other based on the definitions, and ask for extra attention.
@@
 
-The rest of the hierchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution.
+The rest of the hierarchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution.
 The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter,
@@
 
-With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularizations and mixture normal prior for spike-and-slab
+With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab
 regularization.
@@
 Sepcific to additive models, each predictor would expand to mutliple new "predictors" via basis functions, creating greater computational demands.
-@Scheipl2013 demonstrated the computational demands of MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
+Scheipl et al.
+[@Scheipl2013] demonstrated the computational demands of MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
 Hence, we feel compelled to develop scalable and parsimonious algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
@@
 The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a].
-@Rockova2014a proposed an expectation-maximization (EM) based algorithm to fit models that use continous mixture normal priors.
+@Rockova2014a proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
 <!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
@@
 -->
-The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coodirnate descent (CD) algoirthm for fitting SSL models.
+The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
 <!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
@@
 Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results.
-@Bai2020 provides an up-to-date summary of SSL methods and theoretical discussion.
+Bai et al.
+[@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.
 
@@
 To note, the two proposed algorithm provides different utilities.
-The EM-CD algorithm is specifically for fitting BHAM-SSL with an expediated performance, recommending to use in high and ultra-high dimensional setting.
+The EM-CD algorithm is specifically for fitting BHAM-SSL with an expedited performance, recommending to use in high and ultra-high dimensional setting.
 SB-GAM[@Bai2020; @Bai2021] also used EM-CD algorithm.
 The main difference between proposed EM-CD algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior.
-In comparison, our prior is pairwise independent requiring no speical treatment.
-A sepcifica concern of EM-CD algorithm is that it provides no information for inference.
+In comparison, our prior is pairwise independent requiring no special treatment.
+A specific concern of EM-CD algorithm is that it provides no information for inference.
 In the contrast, the EM-IWLS provides a complement to the problem.
 The EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm, and used to construct Wald type inference.
-Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS pirors but also bayesian ridge and lasso priors.
+Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Bayesian ridge and lasso priors.
 
@@
 The idea is that the mixture priors can be expressed as a hierarchical normal distribution with various forms of disperssion distributions.
-[@Yi2012] Given the hiearchical normal prior of the coefficeints, a genalized model can be fitted using linear approximation and least square algorithm.
-For simplicity, we demonstrate using EM-IWLS algorithm to fit mixture normal priors, with the potential of generalization following @Yi2012.
+[@Yi2012] Given the hierarchical normal prior of the coefficients, a generalized model can be fitted using linear approximation and least square algorithm.
+For simplicity, we demonstrate using EM-IWLS algorithm to fit mixture normal priors, with the potential of generalization following Yi and Ma [@Yi2012].
 
@@
 ## Selecting Optimal Scale Values
-
+\label{sec:tune}
 Our proposed models, BHAM, require two preset scale parameters ($s_0$, $s_1$).

### 2021-07-20 Manuscript/SS_GAM.tex
Correct AMA in-text citation

@@
 
+\label{sec:intro}
+
 Using additive models to interpolate nonlinear effects has a long
 history. Early work of additive models can date back to 1979 when
-\citet{Cleveland1979} proposed local smoothing for bivariate analysis.
-Since then, many mathematical and statistical models have been proposed,
-see \citet{Wood2020}. With the recent advancement of computational
+Cleveland \citep{Cleveland1979} proposed local smoothing for bivariate
+analysis. Since then, many mathematical and statistical models have been
+proposed. \citep{Wood2020} With the recent advancement of computational
 technology, many machine learning algorithms that are computationally
@@
 models have a fatal flaw, the lack of transparency and interpretability,
-and hence are referred to as ``black box'' models. Given that the
-primary objective of biomedical studies is to understand the etiology of
-diseases, ``Black box'' models are less helpful as they can hardly
-explain how the change of biomarkers is associated with the disease.
+and hence are referred to as ``black box'' models \citep{Breiman2001}.
+Given that the primary objective of biomedical studies is to understand
+the etiology of diseases, ``Black box'' models are less helpful as they
+can hardly explain how the change of biomarkers is associated with the
+disease.
 
@@
 flexibility and interpretability. Among those, generalized additive
-models (GAMs), proposed in the seminal work of \citet{Hastie1987}, grew
-to be one of the most utilized modeling tools. In a GAM, the response
-variable, \(Y\), which is assumed to follow some exponential family
-distribution \(EF(\mu, \phi)\)\footnote{TODO: explain \(\phi\)}, can be
-expressed as the summation of smooth functions,
-\(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\), written as \[
+models (GAMs), proposed in the seminal work of Hastie and Tibshirani
+\citep{Hastie1987}, grew to be one of the most popular modeling tools.
+In a GAM, the response variable, \(Y\), which is assumed to follow some
+exponential family distribution \(EF(\mu, \phi)\) with mean \(\mu\) and
+dispersion \(\phi\), can be expressed as the summation of smooth
+functions, \(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector
+of covariates \(\boldsymbol{x}\), written as \[
  E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
@@
 
-Many applications of modern biomedical research, e.g.~sequencing data
-analysis, electric health record data analysis, require special
-treatment of high-dimensionality, commonly known as \(p >> n\) problem.
-There is extensive literature on high-dimensional linear models via
-penalized models or Bayesian hierarchical models, see
-\citet{Mallick2013}. These models assume sparsity among the predictors,
-i.e.~most predictors have small or no effects on the outcome. Further
-development focuses on addressing the grouping structure among the
-predictors, where AMs and GAMs can be considered special cases. Various
-penalties are imposed on the basis function coefficients:
-\citet{Ravikumar2009} extended the grouped lasso \citep{Yuan2006} to
-additive models; \citet{Huang2010} further developed adaptive grouped
-lasso for additive models; \citet{Wang2007} and \citet{Xue2009} applied
-grouped SCAD penalty \citep{Fan2001} to the additive models. Recent
-Bayesian hierarchical models are also used in the context of
-high-dimensional additive models. Various group spike-and-slab priors
-combining with computationally intensive Markov chain Monte Carlo (MCMC)
-algorithms \citep{Xu2015, Yang2020} are proposed where the application
-on AMs are by-products. \citet{Scheipl2012} proposed an ad hoc
-spike-and-slab structure prior, normal-mixture-of-inverse gamma,
-specifically for semi-parametric regression using MCMC. The
-normal-mixture-of-inverse gamma prior is mixture of t-distribution with
-an additional coefficients mixing vector, where the vector follows a
-mixture of two Gaussian distribution. \citet{Bai2020} was the first to
-apply group spike-and-slab lasso prior to AMs using a fast optimization
-algorithm, and further generalized the framework to GAMs in
-\citet{Bai2021}. Regardless of the nice statistical properties, a
-practical and pivotal question, if non-linear effects exist, yet to be
-addressed by these methods. Besides \citet{Schipel2012}, most proposed
-methods adapt an `all-in-all-out' strategy, i.e.~either including or
-excluding the variable completely, rendering no space for within group
-selection. In addition, a common criticism received by these methods is
-that the basis function coefficients tend to be overly penalized due to
-the sparsity assumption directly applied on the basis function
-coefficients.\citep{Scheipl2013} It would be of special interest to
-develop a fast yet flexible generalized additive model.
+Many modern biomedical research, e.g.~sequencing data analysis, electric
+health record data analysis, require special treatment of
+high-dimensionality, commonly known as \(p >> n\) problem. There is
+extensive literature on high-dimensional linear models via penalized
+models or Bayesian hierarchical models, see Mallick and Yi
+\citep{Mallick2013} for review. These models assume sparsity among the
+predictors, i.e.~most predictors have small or no effects on the
+outcome. Further development focuses on addressing the grouping
+structure among the predictors, where AMs and GAMs can be considered
+special cases. Various penalties are imposed on the basis function
+coefficients: Ravikumar et al. \citep{Ravikumar2009} extended the
+grouped lasso \citep{Yuan2006} to additive models; Huang et al.
+\citep{Huang2010} further developed adaptive grouped lasso for additive
+models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
+respectively applied grouped SCAD penalty \citep{Fan2001} to the
+additive models. Recent Bayesian hierarchical models are also used in
+the context of high-dimensional additive models. Various group
+spike-and-slab priors combining with computationally intensive Markov
+chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
+proposed, where the application on AMs are by-products. Scheipl et al.
+\citep{Scheipl2012} proposed an ad hoc spike-and-slab structure prior,
+normal-mixture-of-inverse gamma, specifically for semi-parametric
+regression using MCMC. The normal-mixture-of-inverse gamma prior is
+mixture of t-distribution with an additional coefficients mixing vector,
+where the vector follows a mixture of two Gaussian distribution. Bai et
+al. \citep{Bai2020} was the first to apply group spike-and-slab lasso
+prior to AMs using a fast optimization algorithm, and further
+generalized the framework to GAMs\citep{Bai2021}. Regardless of the nice
+statistical properties, a practical and pivotal question, if non-linear
+effects exist, yet to be addressed by these methods. Besides Scheipl et
+al. \citep{Scheipl2012}, most proposed methods adapt an `all-in-all-out'
+strategy, i.e.~either including or excluding the variable completely,
+rendering no space for within group selection. In addition, a common
+criticism received by these methods is that the basis function
+coefficients tend to be overly penalized due to the sparsity assumption
+directly applied on the basis function coefficients.\citep{Scheipl2013}
+It would be of special interest to develop a fast yet flexible
+generalized additive model.
 
@@
 predictor. Additionally, two fast computing optimization-based
-algorithms are developed for high and ultra-high dimension settings
-respectively. Meanwhile, to avoid the over-penalization of the basis
-function coefficients, smoothness penalties are incorporated in the
-model via re-parameterization. Smoothness penalties are commonly
-implemented in GAMs through smoothing and penalized regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive
-penalties on each smoothing function. For the most common cases of cubic
-splines, the smoothness penalty conditioning on a smoothness penalty
-parameter \(\lambda_j\) is a function of the integration of the second
-derivative of the spline function, expressed mathematically as
+algorithms are developed for high and
+ultrahigh-dimensional\citep{Fan2009} settings respectively. Meanwhile,
+to avoid the over-penalization of the basis function coefficients,
+smoothness penalties are incorporated in the model via
+re-parameterization. Smoothness penalties are commonly implemented in
+GAMs through smoothing and penalized regression splines. They are
+quadratic norms of the coefficients and allow locally adaptive penalties
+on each smoothing function. For the most common cases of cubic splines,
+the smoothness penalty conditioning on a smoothness penalty parameter
+\(\lambda_j\) is a function of the integration of the second derivative
+of the spline function, expressed mathematically as
 \begin{equation}\label{eq:smoothpen}
@@
 the separation of the linear space of the smooth function, also know as
-the null space, from the nonlinear space. \citet{Marra2011} introduced a
-re-parameterization of the basis function matrix based on the
-decomposition of the smoothness penalty matrix \(\boldsymbol{S}\). The
-re-parameterization allows different penalties imposed on the null and
-nonlinear spaces separately, and further allows the sparsity in linear
-space and smoothness in nonlinear effect interpolation. \footnote{TODO:
-  a sentence that it decompose the prior to independent priors, provide
-  obvious computational convenience.} The computational convenience
-improved the scalability of the model.
+the null space, from the nonlinear space. Marra and Wood
+\citep{Marra2011} introduced a re-parameterization of the basis function
+matrix based on the decomposition of the smoothness penalty matrix
+\(\boldsymbol{S}\). The re-parameterization allows different penalties
+imposed on the null and nonlinear spaces separately, and further allows
+the sparsity in linear space and smoothness in nonlinear effect
+interpolation. In addition, the priors of the linear and nonlinear
+effects are independent, creating computational convenience and further
+improving the scalability of the model.
 
@@
 Spike-and-slab lasso priors provide computational convenience during
-model fitting through optimization algorithms, instead of computational
+model fitting through optimization algorithms versus computational
 intensive sampling algorithms. They make fitting high- and
-ultrahigh-dimensional models feasible without sacrificing performance in
-prediction and variable selection. Secondly, the optimization algorithms
-introduced in this article include EM-Coordinate Descent algorithm and
-EM-Iterative weighted least square algorithm for different utilities.
-The applications of the two aforementioned algorithm are fruitful in
-Bayesian hierarchical models in high-dimensional data analysis, see
-\footnote{TODO: add citations}. Even though SB-GAM with a group
-spike-and-slab lasso prior also uses EM-CD for model fitting, the
+ultrahigh-dimensional models more feasible without sacrificing
+performance in prediction and variable selection. Secondly, the
+optimization algorithms introduced in this article include EM-Coordinate
+Descent algorithm and EM-Iterative weighted least square algorithm for
+different utilities. The applications of the two aforementioned
+algorithm are fruitful in Bayesian hierarchical models in
+high-dimensional data analysis.
+\citep{Yi2012, Rockova2014a, Rockova2018} Even though SB-GAM with a
+group spike-and-slab lasso prior also uses EM-CD for model fitting, the
 proposed EM-IWLS provides uncertainty measures that could be further
@@
 complements the lack of inference potential from EM-CD in a median and
-high data setting. Furthermore, the proposed model addresses the
+high dimensional setting. Furthermore, the proposed model addresses the
 incapability of bilevel selection in SB-GAM. Lastly, the R
@@
 usefulness of the proposed methodologies in Chpater 4. Conclusion and
-discussions are made in Chapter 5.
+discussions are made in Chapter 5. \footnote{TODO: rewrite this
+  sentence.}
 
@@
 
-Following the GLM notation introduced in Section \label{introduction},
-we have a generalized additive model and its matrix form,
+\label{sec:BHAM}
+
+Following the GLM notation introduced in Section \ref{sec:intro}, we
+have a generalized additive model and its matrix form,
 \begin{equation}\label{eq:gam}
@@
 \(\boldsymbol{X}\) via a re-parameterization. The re-parameterization
-follows \citet{Marra2011}, in which eigen-decomposed the penalty matrix
+follows Marra and Wood \citep{Marra2011}, in which eigen-decomposed the
+penalty matrix
 \(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\),
@@
 basis function matrix \(\boldsymbol{X}^\ast\) is scale-invariant
-\footnote{ I know this work for tensor spline (see \citet{Wood2006}),
+\footnote{ I know this work for tensor spline (see \citep{Wood2006}),
   but I am not sure if it can be generalized to all smoothing functions}.
@@
 Bayesian high-dimensional analysis for its utility in outcome prediction
-and variable selection. We defer to \citet{Bai2021Review} for an
-in-depth introduction to spike-and-slab priors. To summary,
+and variable selection. We defer to Bai et al. \citep{Bai2021Review} for
+an in-depth introduction to spike-and-slab priors. To summary,
 spike-and-slab priors are a family of mixture distributions that
@@
   to not confound with the smooth functions?} and the slab density
-\(f_{slab}(\cdot)\), for example \citet{George1993}; \citet{George1997};
-\citet{Chipman1996} for grouped variables; \citet{Brown1998} for
-multivariate outcomes; \citet{Ishwaran2005}; \citet{Clyde2004} and
-reference therein.
+\(f_{slab}(\cdot)\), see George and McCulloch
+\citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
+variables; Brown et al. \citep{Brown1998} for multivariate outcomes;
+Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
+\citep{Clyde2004} and reference therein.
 
@@
 studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS \citep{Rockova2014a}, \footnote{TODO:
-  add more citation}. Particularly, the development of Spike-and-slab
-Lasso \citep{Rockova2018b, Rockova2018} substantially improves the
+algorithms bear, which include EMVS \citep{Rockova2014a} and
+Spike-and-slab Lasso \citep{Rockova2018b, Rockova2018}. Particularly,
+the development of Spike-and-slab Lasso substantially improves the
 scalability of SS models, setting up the theoretical foundation for
-generalized models in -omics data analysis \footnote{TODO: add Dr.~Yi's
-  papers}. The spike-and-slab lasso prior is composed of two double
-exponential distributions with mean 0 and different dispersion
-parameters, \(0 < s_0 < s_1\), mathematically,
-\begin{equation} \label{eq:ssl}
+generalized models in -omics data analysis
+\citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The spike-and-slab
+lasso prior is composed of two double exponential distributions with
+mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
+mathematically, \begin{equation} \label{eq:ssl}
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
@@
 signals are not shrink to zero, favored in high-dimensional data
-analysis. We noticed that \citet{Bai2021} introduces a spike-and-slab
-group prior in the GAM framework, where the densities of the spike and
-slab components take the group lasso density \citep{Xu2015}.
-\citet{Bai2021} takes an ``all-in-all-out'' strategy for function
-selection.
+analysis. We noticed that Bai \citep{Bai2021} introduces a
+spike-and-slab group prior in the GAM framework, where the densities of
+the spike and slab components take the group lasso density
+\citep{Xu2015}. Bai \citep{Bai2021} takes an ``all-in-all-out'' strategy
+for function selection.
 
@@
 parameters and optimized via cross-validation. A discussion of how to
-choose the scale parameters comes in Section
-\ref{selecting-optimal-scale-values}. To note, this prior differs from
-previous group spike-and-slab lasso priors{[}TODO: add Dr.~Yi's
-citation{]}, where the \(\boldsymbol{\beta}^0_j\) and
+choose the scale parameters comes in Section \ref{sec:tune}. To note,
+this prior differs from previous group spike-and-slab lasso priors
+\citep{Tang2018, Tang2019}, where the \(\boldsymbol{\beta}^0_j\) and
 \(\boldsymbol{\beta}^\text{pen}_j\) shares the same indicator variables
@@
 
-The reparameterization introduced in Section
-\ref{bayesian-hierarchical-additive-models} grants the validity of the
-proposed prior. First of all, the smooth function bases are linear
-dependent to each other based on the definitions, and ask for extra
-attention. The eigeondecomposition remedies the problem. Hence our prior
-can be set to be conditionally independent. Secondly, the eigenvalue
-scaling provides a panacea to allow unified scale parameters for all
-bases of all smoothing functions of variables, where scaling is mandate.
+The reparameterization introduced in Section \ref{sec:BHAM} grants the
+validity of the proposed prior. First of all, the smooth function bases
+are linear dependent to each other based on the definitions, and ask for
+extra attention. The eigeondecomposition remedies the problem. Hence our
+prior can be set to be conditionally independent. Secondly, the
+eigenvalue scaling provides a panacea to allow unified scale parameters
+for all bases of all smoothing functions of variables, where scaling is
+mandate.
 
-The rest of the hierchical prior follows the traditional spike-and-slab
-lasso piror: we set up the hyper-prior of \(\boldsymbol{\gamma}\), to
-allow local adaption of the shrinkage, using a bernoulli distribution.
-The two indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
-\(\gamma^\text{pen}_j\), shares the same probability parameter, \[
+The rest of the hierarchical prior follows the traditional
+spike-and-slab lasso piror: we set up the hyper-prior of
+\(\boldsymbol{\gamma}\), to allow local adaption of the shrinkage, using
+a bernoulli distribution. The two indicators of the \(j\)th predictor,
+\(\gamma^{0}_j\) and \(\gamma^\text{pen}_j\), shares the same
+probability parameter, \[
 \begin{aligned}
@@
 is possible to generalized the SSL prior to other priors, for example
-normal priors for ridge-type regularizations and mixture normal prior
-for spike-and-slab regularization. These priors would work better in low
-and medium dimensional settings where the sparse assumption is not
+normal priors for ridge-type regularization and mixture normal prior for
+spike-and-slab regularization. These priors would work better in low and
+medium dimensional settings where the sparse assumption is not
 necessary. Here we focus on the mixture normal prior as a demonstration
@@
 predictor would expand to mutliple new ``predictors'' via basis
-functions, creating greater computational demands. \citet{Scheipl2013}
-demonstrated the computational demands of MCMC algorithm for fitting
-spike-and-slab additive models grow exponentially as \(p\) increases
-modestly via monte carlo simulation studies. Hence, we feel compelled to
-develop scalable and parsimonious algorithms for fitting Bayesian
-hierarchical additive models in high-dimensional settings.
+functions, creating greater computational demands. Scheipl et al.
+\citep{Scheipl2013} demonstrated the computational demands of MCMC
+algorithm for fitting spike-and-slab additive models grow exponentially
+as \(p\) increases modestly via monte carlo simulation studies. Hence,
+we feel compelled to develop scalable and parsimonious algorithms for
+fitting Bayesian hierarchical additive models in high-dimensional
+settings.
 
@@
 expectation-maximization (EM) based algorithm to fit models that use
-continous mixture normal priors. In the E step, the latent variables
+continuous mixture normal priors. In the E step, the latent variables
 \(\gamma\) are treated as the missing data, and calculate the posterior
@@
 accordingly. The same authors \citep{Rockova2018b, Rockova2018} further
-combined the EM algorithm with coodirnate descent (CD) algoirthm for
+combined the EM algorithm with coordinate descent (CD) algorithm for
 fitting SSL models. Yi and his group incorporated the SSL prior in the
@@
 deterministic solutions, which becomes a more popular property as
-reproducible results. \citet{Bai2020} provides an up-to-date summary of
-SSL methods and theoretical discussion.
+reproducible results. Bai et al. \citep{Bai2020} provides an up-to-date
+summary of SSL methods and theoretical discussion.
 
@@
 provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM-SSL with an expediated performance, recommending to use in
+fitting BHAM-SSL with an expedited performance, recommending to use in
 high and ultra-high dimensional setting. SB-GAM\citep{Bai2020, Bai2021}
@@
 their group prior. In comparison, our prior is pairwise independent
-requiring no speical treatment. A sepcifica concern of EM-CD algorithm
-is that it provides no information for inference. In the contrast, the
+requiring no special treatment. A specific concern of EM-CD algorithm is
+that it provides no information for inference. In the contrast, the
 EM-IWLS provides a complement to the problem. The EM-IWLS can easily
@@
 general model fitting algorithm that can be used for fitting not only
-SSL and continuous SS pirors but also bayesian ridge and lasso priors.
+SSL and continuous SS priors but also Bayesian ridge and lasso priors.
 
@@
 hierarchical normal distribution with various forms of disperssion
-distributions. \citep{Yi2012} Given the hiearchical normal prior of the
-coefficeints, a genalized model can be fitted using linear approximation
-and least square algorithm. For simplicity, we demonstrate using EM-IWLS
-algorithm to fit mixture normal priors, with the potential of
-generalization following \citet{Yi2012}.
+distributions. \citep{Yi2012} Given the hierarchical normal prior of the
+coefficients, a generalized model can be fitted using linear
+approximation and least square algorithm. For simplicity, we demonstrate
+using EM-IWLS algorithm to fit mixture normal priors, with the potential
+of generalization following Yi and Ma \citep{Yi2012}.
 
@@
 
-Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
-\(s_1\)). Hence, we would construct a two dimensional grid, consists of
-different pairs of (\(s_0\), \(s_1\)) value. However, previous research
-on the topic\footnote{TODO: add citation} suggested the value of slab
-scale \(s_1\) have less impact on the final model and is recommended to
-be set as a generally large value of \(s_1 = 1\) that provides no or
-weak shrinkage. More attention is focused on examining different values
-of spike scale \(s_0\). Instead of the 2-D grid, We consider a sequence
-of \(L\) decreasing values
+\label{sec:tune} Our proposed models, BHAM, require two preset scale
+parameters (\(s_0\), \(s_1\)). Hence, we would construct a two
+dimensional grid, consists of different pairs of (\(s_0\), \(s_1\))
+value. However, previous research on the topic\footnote{TODO: add
+  citations} suggested the value of slab scale \(s_1\) have less impact
+on the final model and is recommended to be set as a generally large
+value of \(s_1 = 1\) that provides no or weak shrinkage. More attention
+is focused on examining different values of spike scale \(s_0\). Instead
+of the 2-D grid, We consider a sequence of \(L\) decreasing values
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the

### 2021-07-24 Manuscript/03-Simulation.Rmd
Revise the simulation section

@@
 # Simulation Study
-In this section, we compare the performance of the proposed models to 3 other models: COSSO, adaptive COSSO ,mgcv. COSSO is xxx method. Adaptive COSSO is another method, mgcv is another method. 
-To demonstrate the correctness and feasibility of the proposed SS-GAM,
-we follow the logistic regression simulation introduced in Bai (Work in
-Progress). We first generate $n=500$ training data sample with number
-of predictors $p=4, 10, 50, 200$ respectively, where the predictors $X \sim MVN_{n\times p}(0, I_{P})$.
-We simulate the outcome data from three distributions, gaussian, poisson, and binomial with the identity link, logit link $g(x) = \log(\frac{x}{1-x})$, and log link $g(x) = log(x)$ functions to mimic the common use of generalized model.
-The outcome is simulated via the following function
+
+In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021].
+COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
+Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component are different for extra flexibility.
+Generalized additive models with automatic smoothing, hereafter \textit{mgcv}, is one of the most popular models for non-linear effect interpolation and prediction.
+SB-GAM is the first spike-and-slab lasso GAM.
+We implemented COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \text{sparseGAM 1.0}.
+COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence used the default choices.
+Both mgcv and proposed models allow customized smoothing functions and we chose the cubic regression spline.
+We controlled the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
+5-fold cross-validation were used for COSSO models, SB-GAM and the proposed models for tuning parameter selection based on the default selection criteria.
+20 default candidates were examined for SB-GAM and the proposed models which allow user-specification of tuning candidates.
+All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
+
+Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al.
+[@Ravikumar2009], stochastic search term selection for GAM [@Scheipl2012].
+However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increasing number of predictors.
+[@Scheipl2013]
+
+## Monte Carlo Simulation Study
+We follow the data generating process described in Bai [@Bai2021].
+We first generated $n=500$ training data points with $p=4, 10, 50, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
+We then simulated the outcome $y$ from three distribution distributions, gaussian, binomial, and poisson with the identity link, logit link $g(x) = \log(\frac{x}{1-x})$, and log link $g(x) = log(x)$.
+The mean of each outcome were simulated via the following function
 $$
@@
 $$
-for gaussian and binomial outcome and shrinked effects
+for gaussian and binomial outcomes and shrinked effects
 
@@
 $$
-for stable poisson outcome.
-For the Gaussian outcome where dispersion are required, the dispersion parameter was set to 1.
-We follow the same data generating functions in Bai.
-In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest
-covariates are inactive $f_j(x_j) = 0$ for $j = 4, \dots, p$.
+for  poisson outcomes.
+Gaussian outcomes required specification of dispersion, where we set the dispersion parameter to be 1.
+In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest covariates are inactive, i.e.
+$f_j(x_j) = 0$ for $j = 4, \dots, p$.
 Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data.
-For mgcv and proposed models where choice of spline function are customizable, we choose to cubic regression spline with 10 basis.
-In COSSO and adaptive COSSO where the choice of spline are pre-determined, we set to have 10 bases.
-The corresponding number of parameter that we estimate with the model is 37, 82, 451 and 1801.
-We also run the simulation where the number of bases are increased to 25 deliberately.
+The dimensionality of the smooth functions are controlled as k=10.
+Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.
+<!-- We also run the simulation where the number of bases are increased to 25 deliberately. -->
+
 To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across 50 simulations.

### 2021-07-24 Manuscript/SS_GAM.tex
Revise the simulation section

@@
 The earliest work for fitting spike-and-slab models using optimization
-algorithm is EMVS \citep{Rockova2014a}. \citet{Rockova2014a} proposed an
-expectation-maximization (EM) based algorithm to fit models that use
-continuous mixture normal priors. In the E step, the latent variables
-\(\gamma\) are treated as the missing data, and calculate the posterior
-mean of the density function; in the M step, a ridge estimator was used
-to update the coefficients, and \(\sigma^2, \theta\) are updated
-accordingly. The same authors \citep{Rockova2018b, Rockova2018} further
-combined the EM algorithm with coordinate descent (CD) algorithm for
-fitting SSL models. Yi and his group incorporated the SSL prior in the
-EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another
-EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty
-SSL exhibit. Such applications were used in generalized linear
-models\citep{Tang2017a}, cox models \citep{Tang2017} and their grouped
+algorithm is EMVS \citep{Rockova2014a}. Ro\{\v{c}\}kov\{'\{a\}\} and
+George\citep{Rockova2014a} proposed an expectation-maximization (EM)
+based algorithm to fit models that use continuous mixture normal priors.
+In the E step, the latent variables \(\gamma\) are treated as the
+missing data, and calculate the posterior mean of the density function;
+in the M step, a ridge estimator was used to update the coefficients,
+and \(\sigma^2, \theta\) are updated accordingly. The same authors
+\citep{Rockova2018b, Rockova2018} further combined the EM algorithm with
+coordinate descent (CD) algorithm for fitting SSL models. Yi and his
+group incorporated the SSL prior in the EM-Iterative Weighted Least
+Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic
+coordinate descent, based on the Lasso penalty SSL exhibit. Such
+applications were used in generalized linear models\citep{Tang2017a},
+cox models \citep{Tang2017} and their grouped
 counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
@@
 
-In this section, we compare the performance of the proposed models to 3
-other models: COSSO, adaptive COSSO ,mgcv. COSSO is xxx method. Adaptive
-COSSO is another method, mgcv is another method. To demonstrate the
-correctness and feasibility of the proposed SS-GAM, we follow the
-logistic regression simulation introduced in Bai (Work in Progress). We
-first generate \(n=500\) training data sample with number of predictors
-\(p=4, 10, 50, 200\) respectively, where the predictors
+In this section, we compare the performance of the proposed models to
+four alternative models: component selection and smoothing operator
+(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
+generalized additive models with automatic smoothing \citep{Wood2011},
+SB-GAM \citep{Bai2021}. COSSO is one of the earliest smoothing spline
+models that consider sparsity-smoothness penalty. Adaptive COSSO
+improved upon COSSO by using adaptive weight for penalties such that the
+penalty of each functional component are different for extra
+flexibility. Generalized additive models with automatic smoothing,
+hereafter \textit{mgcv}, is one of the most popular models for
+non-linear effect interpolation and prediction. SB-GAM is the first
+spike-and-slab lasso GAM. We implemented COSSO and adaptive COSSO with R
+package \texttt{cosso} \citep{cosso} 2.1-1, generalized additive models
+with automatic smoothing with R package \texttt{mgcv} 1.8-31, SB-GAM
+with R package \text{sparseGAM} 1.0. All computation was conducted on a
+high-performance 64-bit Linux platform with 48 cores of 2.70GHz
+eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
+3.6.2 \citep{R}.
+
+Other related methods for high-dimensional GAMs also exist, notably the
+methods of sparse additive models by Ravikumar et al.
+\citep{Ravikumar2009}, stochastic search term selection for GAM
+\citep{Scheipl2012}. However, we exclude these methods from current
+simulation study because of demonstrated inferior predictive performance
+compared to mgcv and scalability issues with increasing number of
+predictors. \citep{Scheipl2013}
+
+\hypertarget{monte-carlo-simulation-study}{%
+\subsection{Monte Carlo Simulation
+Study}\label{monte-carlo-simulation-study}}
+
+xxx method. Adaptive COSSO is another method, mgcv is another method. To
+demonstrate the correctness and feasibility of the proposed SS-GAM, we
+follow the logistic regression simulation introduced in Bai (Work in
+Progress). We first generate \(n=500\) training data sample with number
+of predictors \(p=4, 10, 50, 200\) respectively, where the predictors
 \(X \sim MVN_{n\times p}(0, I_{P})\). We simulate the outcome data from

### 2021-07-25 Manuscript/00-main.Rmd
started flow chart
Added Flowchart of Prior

@@
     keep_tex: TRUE
+header-includes: 
+  - \usepackage{tikz}
+  - \usepackage{pgfplots}
 ---
 
+<!-- \usepackage{tikz} -->
+\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
 \newcommand{\bs}[1]{\boldsymbol{#1}}

### 2021-07-25 Manuscript/02-Method.Rmd
Revise the simulation section
Added Flowchart of Prior

@@
 Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL.
+\\
 
-```{r fig.height=4, fig.width=6}
+```{r fig.height=4, fig.width=6, eval = F}
 # TODO: Refactor the graph to a latex flowchart
@@
 ```
+
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
+&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
+&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+};
+
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+\node (penTheta) [outCome] { $\theta_{j}^\tp\sim Beta(a, b)$ };\\
+\node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j^\tp)$ };\\
+\node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+};
+
+\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
+%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+
+\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta^0_1, \bs \beta^\tp_1, \dots,\beta^0_j, \bs \beta^\tp_j , \dots,\beta^0_p, \bs \beta^\tp_p) $};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
+
+%Lines
+
+\draw[->] (linearTheta) -- (linearGamma);
+\draw[->] (penTheta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (linearTheta);
+\draw[->] (ab) -- (penTheta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+% Now that the diagram has been drawn, background rectangles
+% can be fitted to its elements. This requires the TikZ
+% libraries "fit" and "background".
+% Control input and measurement are labeled. These labels have
+% not been translated to English as "Measurement" instead of
+% "Messung" would not look good due to it being too long a word.
+\begin{pgfonlayer}{background}
+    \node [background,
+                fit=(linearTheta) (linearBeta),
+                label=above:Null Space:] {};
+    \node [background,
+                 fit=(penTheta) (penBeta),
+                label=above:Penalized Space:] {};
+
+\end{pgfonlayer}
+\end{tikzpicture}
 
@@
 The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a].
-@Rockova2014a proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
+Ro{\v{c}}kov{\'{a}} and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
 <!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.

### 2021-07-25 Manuscript/SS_GAM.tex
Added Flowchart of Prior

@@
 % Pandoc citation processing
+
+\usepackage{tikz}
+\usepackage{pgfplots}
 
@@
 
+\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
@@
 generalized additive models with the spike-and-slab lasso prior as the
-BHAM-SSL.
-
-\includegraphics{Fig/SS_prior_flow_chart.png}
+BHAM-SSL. \textbackslash{}
+
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
+&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
+&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+};
+
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+\node (penTheta) [outCome] { $\theta_{j}^\text{pen}\sim Beta(a, b)$ };\\
+\node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j^\text{pen})$ };\\
+\node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+};
+
+\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
+%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^\text{pen}_1, \dots,\beta^0_j, \boldsymbol{\beta}^\text{pen}_j , \dots,\beta^0_p, \boldsymbol{\beta}^\text{pen}_p) $};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
+
+%Lines
+
+\draw[->] (linearTheta) -- (linearGamma);
+\draw[->] (penTheta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (linearTheta);
+\draw[->] (ab) -- (penTheta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+% Now that the diagram has been drawn, background rectangles
+% can be fitted to its elements. This requires the TikZ
+% libraries "fit" and "background".
+% Control input and measurement are labeled. These labels have
+% not been translated to English as "Measurement" instead of
+% "Messung" would not look good due to it being too long a word.
+\begin{pgfonlayer}{background}
+    \node [background,
+                fit=(linearTheta) (linearBeta),
+                label=above:Null Space:] {};
+    \node [background,
+                 fit=(penTheta) (penBeta),
+                label=above:Penalized Space:] {};
+
+\end{pgfonlayer}
+\end{tikzpicture}
 
@@
 spike-and-slab lasso GAM. We implemented COSSO and adaptive COSSO with R
-package \texttt{cosso} \citep{cosso} 2.1-1, generalized additive models
-with automatic smoothing with R package \texttt{mgcv} 1.8-31, SB-GAM
-with R package \text{sparseGAM} 1.0. All computation was conducted on a
-high-performance 64-bit Linux platform with 48 cores of 2.70GHz
-eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
-3.6.2 \citep{R}.
+package \texttt{cosso 2.1-1}, generalized additive models with automatic
+smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package
+\text{sparseGAM 1.0}. COSSO models and SB-GAM do not provide flexibility
+to define smoothing functions, and hence used the default choices. Both
+mgcv and proposed models allow customized smoothing functions and we
+chose the cubic regression spline. We controlled the dimensionality of
+each smoothing function, 10 bases, for all different choices of
+smoothing functions. 5-fold cross-validation were used for COSSO models,
+SB-GAM and the proposed models for tuning parameter selection based on
+the default selection criteria. 20 default candidates were examined for
+SB-GAM and the proposed models which allow user-specification of tuning
+candidates. All computation was conducted on a high-performance 64-bit
+Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
+processors and 24G of RAM per core and R 3.6.2 \citep{R}.
 
@@
 
-xxx method. Adaptive COSSO is another method, mgcv is another method. To
-demonstrate the correctness and feasibility of the proposed SS-GAM, we
-follow the logistic regression simulation introduced in Bai (Work in
-Progress). We first generate \(n=500\) training data sample with number
-of predictors \(p=4, 10, 50, 200\) respectively, where the predictors
-\(X \sim MVN_{n\times p}(0, I_{P})\). We simulate the outcome data from
-three distributions, gaussian, poisson, and binomial with the identity
-link, logit link \(g(x) = \log(\frac{x}{1-x})\), and log link
-\(g(x) = log(x)\) functions to mimic the common use of generalized
-model. The outcome is simulated via the following function \[
+We follow the data generating process described in Bai \citep{Bai2021}.
+We first generated \(n=500\) training data points with
+\(p=4, 10, 50, 200\) predictors respectively, where the predictors \(X\)
+are simulated from a multivariate normal distribution
+\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulated the outcome
+\(y\) from three distribution distributions, gaussian, binomial, and
+poisson with the identity link, logit link
+\(g(x) = \log(\frac{x}{1-x})\), and log link \(g(x) = log(x)\). The mean
+of each outcome were simulated via the following function \[
 \mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for gaussian and binomial outcome and shrinked effects
+\] for gaussian and binomial outcomes and shrinked effects
 
@@
 \mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
-\] for stable poisson outcome. For the Gaussian outcome where dispersion
-are required, the dispersion parameter was set to 1. We follow the same
-data generating functions in Bai. In this data generating process, we
-have \(x_1, x_2, x_3, x_4\) as the active covariates, while the rest
-covariates are inactive \(f_j(x_j) = 0\) for \(j = 4, \dots, p\).
-Another set of independent sample of size
-\(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are created following the same
-data generating process, serving as the testing data. For mgcv and
-proposed models where choice of spline function are customizable, we
-choose to cubic regression spline with 10 basis. In COSSO and adaptive
-COSSO where the choice of spline are pre-determined, we set to have 10
-bases. The corresponding number of parameter that we estimate with the
-model is 37, 82, 451 and 1801. We also run the simulation where the
-number of bases are increased to 25 deliberately. To evaluate the
-performance of the models, mean squared error, mean absolute error for
-gaussian model, AUC and misclassifcation rate are calculated for
-binomial model, and deviance are calculated for Poisson model, averaging
-across 50 simulations. The results are tabulated in Table 1, 2, 3.
+\] for poisson outcomes. Gaussian outcomes required specification of
+dispersion, where we set the dispersion parameter to be 1. In this data
+generating process, we have \(x_1, x_2, x_3, x_4\) as the active
+covariates, while the rest covariates are inactive,
+i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
+independent sample of size \(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are
+created following the same data generating process, serving as the
+testing data. The dimensionality of the smooth functions are controlled
+as k=10. Hence, the corresponding number of coefficient parameters to be
+estimated is 37, 82, 451 and 1801.
+
+To evaluate the performance of the models, mean squared error, mean
+absolute error for gaussian model, AUC and misclassifcation rate are
+calculated for binomial model, and deviance are calculated for Poisson
+model, averaging across 50 simulations. The results are tabulated in
+Table 1, 2, 3.
 

### 2021-07-27 Manuscript/00-main.Rmd
Revise conclusion

@@
 ```
-
-## Cross-referencing
-
-The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
-for figures, tables, equations, etc., is encouraged (using `[@<name>]`, equivalent of `\ref{<name>}` and `\label{<name>}`).
-That works well for citations in Rmarkdown, not so well for figures and tables.
-In that case, it is possible to revert to standard \LaTeX syntax.

### 2021-07-27 Manuscript/05-Discussion.Rmd
Revise conclusion

@@
 # Discussion
+In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.
+Meanwhile, we introduced two optimization based algorithms for model fitting.
+The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner, but also provide naive inferential supports for many commonly used hierarchical priors in the analyses that the sparse assumption is weak or not necessary.
+Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.
+Compared another newly developed high-dimensional Bayesian GAM that emphasize computational efficiency [@Bai2021], the proposed model provides flexibility to address bi-level functional selection and inferential support while maintaining the same level of prediction accuracy.
+
+For transnational science purpose, the
+
+Compare to previous Bayesian GAMs, the proposed model greatly increases the scalability to be applied in high-dimensioanl data analyses, while addressing the bi-level selection problem [@Bai2021].
+
+the implementation of ssGAM-NM using EM-IRLS.
+Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
+Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
+mixture t-distribution, and mixture double exponential distributions.
+@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors,
+$$
+\begin{aligned}
+\beta_i \sim N(0, \tau^2_i),\\
+\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
+\end{aligned}
+$$
+with $\tau^2$ integrated out analytically.
+Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
+One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
+$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
+The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
+[@Rockova2014a]
+
+In this chapter, we propose the spike-and-slab generalized additive models for high-dimensional analysis.Two fast computing algorithms, EM-IRLS and EM-coordinate descent are developed for expedited model fitting.
+EM-IRLS provides uncertainty measures, while EM-coordinate descent algorithm are extremely fast for model fitting.
+In the simulation studies, we shows the proposed model outperforms the state of the art GAM model, `mgcv` in prediction and function inference.
+Moreover, due to the model selection nature of the spike-and-slab prior, the proposed model can identify predictors that have large effects on the outcome.
+In addition, the model can be used to answer the question if non-linear effect are necessary.
 
@@
 
-In this chapter, we propose the spike-and-slab generalized additive models for high-dimensional analysis.Two fast computing algorithms, EM-IRLS and EM-coordinate descent are developed for expedited model fitting.
-EM-IRLS provides uncertainty measures, while EM-coordinate descent algorithm are extremely fast for model fitting.
-In the simulation studies, we shows the proposed model outperforms the state of the art GAM model, `mgcv` in prediction and function inference.
-Moreover, due to the model selection nature of the spike-and-slab prior, the proposed model can identify predictors that have large effects on the outcome.
-In addition, the model can be used to answer the question if non-linear effect are necessary.
-
   <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function.
@@
 Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
-
-In the manuscript, we describe the implementation of ssGAM-NM using EM-IRLS.
-Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
-Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
-mixture t-distribution, and mixture double exponential distributions.
-@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors,
-$$
-\begin{aligned}
-\beta_i \sim N(0, \tau^2_i),\\
-\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
-\end{aligned}
-$$
-with $\tau^2$ integrated out analytically.
-Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
-One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
-$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
-The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
-[@Rockova2014a]

### 2021-07-27 Manuscript/00-main.Rmd
Finish revising the conclusion part

@@
 
-...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`.
+<!-- ...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`.
 It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).
+-->
 
-Analogously, use Rmarkdown to produce tables as usual:
+<!-- Analogously, use Rmarkdown to produce tables as usual: -->
 

### 2021-07-27 Manuscript/SS_GAM.tex
Finish revising the conclusion part

@@
 
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-7-1.pdf}
+
 \includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}
@@
 \includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}
 
@@
 
-Remaining issue: 1. Degree of freedom has yet to closely studied. As an
-direct impact, the functional selection can be improved when effective
-degree of freedom are more carefully studied 2. While the proposed model
-addresses a great deal of analytic problem, analyzing the time-to-event
-outcome remains unsolved. Our current effort resides on expanding. An
-naive approach could be convert an time-to-event outcome to a poisson
-outcome following xxx. Hence, the . However, it would be more efficient
-to , via penalized pseudo likelihood function, which is the current
-effort resides. 2. Future exapansion, as we were analyzing the
-metabolites data, we observed that metabolites pathway infromation can
-be useful in the predictive and integrative model can be developed based
-on the proposed framework. As an naive solution, the selection indicator
-of linear effect of the spline terms of all metabolites on the same
-pathway can be set to share the same hyper prior such that the group
-selection can be modestly achieved. Meanwhile further advancedment of
-the methodology regarding to the integrative analysis can consider
-setting the hypor priors of the selection indivator of the non-linear
-components conditioning on the probability of inclusion of the linear
-effect. Such structure is more sensible analytically.
-
-Meanwhile, we specify that \(\gamma_{j}^0\) and \(\gamma_{j}^{pen}\) are
-independently distributed for analytic simplicity. However, this set-up
-could possibility result in a situation that is not theoretically
-possible: the non-linear component is selected, but the linear component
-is not. This can be addressed analytically by forcing to include the
-linear component when non-linear component is included. Another possible
-solution is to impose an dependent structure of \(\gamma_{j}^{pen}\) on
-\(\gamma_{j^{0}}\), i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\).
-
-\hypertarget{discussion-1}{%
-\subsection{Discussion}\label{discussion-1}}
-
-In this chapter, we propose the spike-and-slab generalized additive
-models for high-dimensional analysis.Two fast computing algorithms,
-EM-IRLS and EM-coordinate descent are developed for expedited model
-fitting. EM-IRLS provides uncertainty measures, while EM-coordinate
-descent algorithm are extremely fast for model fitting. In the
-simulation studies, we shows the proposed model outperforms the state of
-the art GAM model, \texttt{mgcv} in prediction and function inference.
-Moreover, due to the model selection nature of the spike-and-slab prior,
-the proposed model can identify predictors that have large effects on
-the outcome. In addition, the model can be used to answer the question
-if non-linear effect are necessary.
-
-The ``all-in-all-out'' set-up was criticized for its inability to
-answering an important research question, if the effects are non-linear
-or not. On the other hand, the model can be much more flexible by
-allowing each dimension of the penalized space to have an independent
-indicator, i.e.~\(\gamma^{pen}_{jk}\) of \(\beta^{pen}_{jk}\) for
-\(k = 1\dots, K_j\). This prior set-up generalized to the group
-spike-and-slab. The additional flexibility doesn't not necessarily grand
-the model neither theoretical justification nor analytic simplicity. The
-spline penalty assumes a common smooth shrinkage \(\lambda_j\) applied
-to all coefficients of the penalized space, which contradicts to the
-model setup. Besides, having multiple indicator variables complicates
-the decision of if non-linear effects are necessary.
-
-In the manuscript, we describe the implementation of ssGAM-NM using
-EM-IRLS. Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
-Other spike-and-slab prior can be fitted with EM-IRLS as well,
-e.g.~mixture t-distribution, and mixture double exponential
-distributions. \citet{Yi2012} demonstrated that t prior and double
-exponential prior can be seen hierarchical normal priors, \[
-\begin{aligned}
-\beta_i \sim N(0, \tau^2_i),\\
-\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
-\end{aligned}
-\] with \(\tau^2\) integrated out analytically. Hence, the EM-IRLS can
-be easily applied to those two additional mixture priors. One of the
-advantage of using mixture \(t\) prior over mixture normal prior is that
-\(t\) distributions with low degree of freedom, e.g.~\(v=1\), have
-heavier tail than the normal distribution, which is also a special case
-of \(t\) distribution when \(v=\infty\). The heavier tails help to
-obtain posteriors that concentrated around the true nonzero coefficients
-at optimal rates. \citep{Rockova2014a}
-
-\ldots and can be referenced (Figure \ref{fig:plot}) by including the
-\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
-\texttt{fig.cap} attribute of the R chunk:
-\texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.
-It is a quirky hack at the moment, see
-\href{https://github.com/yihui/knitr/issues/323}{here}.
-
-Analogously, use Rmarkdown to produce tables as usual:
-
-\hypertarget{cross-referencing}{%
-\subsection{Cross-referencing}\label{cross-referencing}}
-
-The use of the Rmarkdown equivalent of the \LaTeX cross-reference system
-for figures, tables, equations, etc., is encouraged (using
-\texttt{{[}@\textless{}name\textgreater{}{]}}, equivalent of
-\texttt{\textbackslash{}ref\{\textless{}name\textgreater{}\}} and
-\texttt{\textbackslash{}label\{\textless{}name\textgreater{}\}}). That
-works well for citations in Rmarkdown, not so well for figures and
-tables. In that case, it is possible to revert to standard
-\LaTeX syntax.
+In the paper, we described a novel generalized additive model using
+Bayesian hierarchical priors, particularly the proposed spike-and-slab
+spline prior for bi-level functional selection. Meanwhile, we introduced
+two optimization based algorithms for model fitting. The algorithms can
+be easily scale up to address high-dimensional data analysis in a
+computational efficient manner, but also provide naive inferential
+supports for many commonly used hierarchical priors,
+e.g.~t-distribution, double exponential distribution, mixture t
+distribution, and mixture double exponential distribution, in the
+analyses that the sparse assumption is weak or not necessary. Via
+simulations, we demonstrated that the proposed model provides as good,
+if not better, prediction performance compared to some state-of-the-art
+non-linear modeling devices.
+
+The proposed model shares many commonality with an high-dimensional
+Bayesian GAM, SB-GAM \citep{Bai2021}, independently developed around the
+same time of this work. Both frameworks emphasize computational
+efficiency by deploying group spike-and-slab lasso type of priors and
+optimization-based fast and scalable algorithms. Bai provided the
+theoretical proof for the consistency of variable selection using group
+spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the
+bi-level functional selection and the model inference. The proposed
+model provides solutions to these remaining questions while maintaining
+the same level of prediction accuracy. Moreover, the proposed model
+renders additional flexibility to model specification, allowing various
+choices of smoothing functions and degrees of freedom.
+
+For transnational science purpose, we implemented the proposed model in
+a R package \texttt{BHAM}, deposited at
+\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
+smoothing function specification, we deploy the same programming grammar
+as in the state-of-the-art package \texttt{mgcv}, in contrast to
+previous tools where smoothing functions are limited to the default
+ones. Ancillary functions are provided for model specification in
+high-dimensional settings, curve plotting and functional selection.
+
+There are some improvements possible for the proposed models. First of
+all, the proposed model achieves a bi-level selection via the two-part
+spike-and-slab spline prior. Nevertheless, this set-up could result in a
+situation that is not theoretically sound: the non-linear component is
+selected, but the linear component is not. We currently address it
+analytically by forcing to include the linear component when non-linear
+component is included. Another possible solution is to impose an
+dependent structure of \(\gamma_{j}^{pen}\) on \(\gamma_{j^{0}}\),
+i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\). Secondly, while the
+proposed model addresses a great deal of analytic problem, analyzing the
+time-to-event outcome remains unsolved. Our current effort resides on
+expanding GAM to Cox proportional hazard models. An naive approach would
+be convert an time-to-event outcome to a poisson outcome following
+Whitehead \citep{Whitehead1980}. However, it would be more efficient to
+directly fitting Cox models via penalized pseudo likelihood function.
+Last but not least, we plan to expand the proposed model to integrate
+external biology information. The main motivation for the integrative
+model is that biologically informed groupping of weak effects increases
+the power of detecting true associations between features and the
+outcome \citep{Peterson2016}, and stabilizes the analysis results for
+reproducibility purpose. Such integration can be achieved by setting up
+a structural hyperprior on the inclusion indicator of the smoothing
+function null space \(\boldsymbol{\gamma}^0\). The similar strategy has
+been used in Ferrari and Dunson \citep{Ferrari2020}.
 

### 2021-07-27 R/functions.R
Setup targets use tflow

@@ -31,22 +31,43 @@ gam_screen <- function(dat, nrow = 200){
 }
 
 
-create_success_rate_table <- function(path){
-  sims <- list.files(path,full.names =TRUE)
-  
-  success_rate <- map_dfr(sims, .f = function(sim){
-    sim.df <- unglue_data(sim, 
-                          "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
-      mutate( p = as.numeric(p))
-    fls <- list.files(sim, full.names = TRUE) 
-    fls <- fls[grep(".rds", x=fls)]
-    n <- length(fls)
-    
-    data.frame(sim.df, n_success = n, path = sim)
-  }) %>% 
-    arrange(p) 
+create_success_rate_table <- function(path_list){
 
-  return(success_rate)  
+  # sims <- list.files(path,full.names =TRUE)
+  # browser()
+  unglue_data(path_list[str_ends(path_list, ".rds")],
+              "{path}/it_{}") %>% 
+    group_by(path) %>% 
+    summarize(n_success = n()) %>% 
+    unglue_unnest(
+      col = path,
+      patterns = "{}/bgam_{study}_-dis_{dist}-p_{p}",
+      remove = FALSE
+    ) %>% 
+    mutate( p = as.numeric(p)) %>% 
+    arrange(dist, p) %>% 
+    select(dist, p, n_success, path)
+  #   head()
+  # 
+  # 
+  # success_rate <- map_dfr(path_list, 
+  #                         .f = function(sim_file){
+  #                           browser()
+  #   sim.df <- unglue_data(sim_file, 
+  #                         "{}/bgam_{study}_-dis_{dist}-p_{p}/{}", 
+  #                         kee) %>% 
+  #     mutate( p = as.numeric(p)) %>% 
+  #     group_by(dist)
+  #     summarize()
+  #   fls <- list.files(sim_file, full.names = TRUE) 
+  #   fls <- fls[grep(".rds", x=fls)]
+  #   n <- length(fls)
+  #   
+  #   data.frame(sim.df, n_success = n, path = sim)
+  # }) %>% 
+  #   arrange(p) 
+
+  # return(success_rate)  
 }
 
 
@@ -113,57 +134,3 @@ make_sim_main_plots <- function(success_rate, dist, measures){
   ggarrange(plotlist = ., ncol = 1)
 }
 
-make_sim_table <- function(success_rate, dist#, measures
-                           ){
[…]

### 2021-07-27 packages.R

@@ -1,0 +1,21 @@
+## library() calls go here
+library(conflicted)
+library(dotenv)
+library(targets)
+library(tarchetypes)
+
+
+library(dplyr)
+library(tidyverse)
+library(rticles)
+library(gtsummary)
+library(here)
+library(mgcv)
+library(broom)
+library(unglue)
+library(knitr)
+library(ggpubr)
+library(xtable)
+
+
+conflict_prefer("filter", "dplyr")

### 2021-07-27 Manuscript/00-main.Rmd
Creating Result Tables

@@
 library(here)
+library(xtable)
 
@@
 tar_load(disc_path)
+
+tar_load(tab_binom)
+tar_load(tab_gaus)
 ```

### 2021-07-27 Manuscript/SS_GAM.tex
Creating Result Tables

@@
 
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrllllll}
+  \hline
+ & p & bglm\_spline\_de & blasso\_spline & cosso & acosso & mgcv & SB\_GAM \\ 
+  \hline
+1 & 4.00 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
+  2 & 10.00 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
+  3 & 50.00 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
+  4 & 200.00 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & NaN (NA) & 0.92 (0.01) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample area under the curve measure over for binomial outcome over 50 iterations} 
+\label{tab:bin_auc}
+\end{table}
+
 To evaluate the performance of the models, mean squared error, mean
@@
 
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-7-1.pdf}
-
 \includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}
@@
 \includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}
 

### 2021-07-27 R/Sim_Res/make_sim_main_table.R

@@ -1,0 +1,77 @@
+make_sim_main_table <- function(success_rate, dist, section = "test" , measures){
+  # browser()
+  
+  total_dat <- success_rate %>% 
+    # filter(dist==dist, n_success!=0) %>% 
+    filter(dist=={{dist}}, n_success!=0) %>%
+    pull(path) %>% 
+    map_dfr( .f = function(sim){
+      
+      
+      sim.df <- unglue_data(sim, 
+                            "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
+        mutate( p = as.numeric(p))
+      fls <- list.files(sim, full.names = TRUE) 
+      fls <- fls[grep(".rds", x=fls)]
+      n <- length(fls)
+      
+      
+      if(n == 0)
+        return(data.frame(NULL))
+      
+      # .file <- fls[1]
+      
+      ret <- fls %>%
+        map_dfr(.f = function(.file){
+          it <- unglue_data(.file, "{whatever}/it_{it}.rds") %>% pull(it) %>% as.numeric
+          ret <- data.frame(
+            it = it,
+            read_rds(.file) %>% 
+              data.frame() %>% 
+              rownames_to_column("method")
+            
+          )
+        }) %>%
+        arrange(it) %>% 
+        mutate(p = sim.df$p)#  %>% 
+      
+    }) %>% 
+    mutate(method = factor(
+      method,
+      levels = c("bglm_t", "bglm_de", "blasso",
+                 "bglm_t_group", "bglm_de_group", "blasso_group",
+                 "bglm_spline_t", "bglm_spline_de", "blasso_spline",
+                 "cosso", "acosso", "mgcv", "SB_GAM"
+      )
+    )
+    )
+  
+  total_dat %>% 
+    select(p, method, 
+           {{measures}} := paste(section, measures, sep=".")) %>% #head
+    filter(method %in% c("bglm_spline_de", "blasso_spline",
+                         "cosso", "acosso", "mgcv", "SB_GAM")) %>% 
+    group_by(p, method) %>% 
+    summarize(
+      across({{measures}}, 
+             list(mean = mean, sd = sd), na.rm = T,
+             .names = "{.fn}")#,
+              # n_miss = sum(is.na(auc))
+    ) %>% 
+    ungroup() %>% 
+    transmute(
+      p, method,
+      {{measures}} := sprintf("%.2f (%.2f)", mean, sd)#,
[…]

### 2021-07-27 _targets.R

@@ -1,56 +1,59 @@
-library(targets)
-library(tarchetypes)
+## Load your packages, e.g. library(targets).
+source("./packages.R")
 
-source("R/functions.R")
+## Load your R files
+lapply(list.files("./R", full.names = TRUE, recursive = TRUE), source)
 
-# This is an example _targets.R file. Every
-# {targets} pipeline needs one.
-# Use tar_script() to create _targets.R and tar_edit()
-# to open it again for editing.
-# Then, run tar_make() to run the pipeline
-# and tar_read(summary) to view the results.
-
-options(tidyverse.quiet = TRUE)
-# Set target-specific options such as packages.
-tar_option_set(packages = c("dplyr", "tidyverse", "rticles", "gtsummary", "here",
-                            "mgcv", "broom", "unglue", "knitr",
-                            "ggpubr"))
-
-# End this file with a list of target objects.
-list(
-  
+## tar_plan supports drake-style targets and also tar_target()
+tar_plan(
   #  Simulations Studies -----------------------------------------------------
   #* Main Study ####
-  #* 
-  #* 
-  tar_target(success_rate,
-             create_success_rate_table("Simulation/Main/Data/main_k=10/")),
   
-  tar_target(binom_plot,
-             make_sim_main_plots(
-               success_rate,
-               "binomial",
-               list("test.deviance", "test.auc", "test.misclassification"))
-             ),
-
-  tar_target(gaussian_plot,
-             make_sim_main_plots(
-               success_rate,
-               "gaussian",
-               list("test.R2", "test.mse", "test.mae"))),
-
-  tar_target(poisson_plot,
-             make_sim_main_plots(
-               success_rate,
-               "poisson",
-               list("test.deviance", "test.mse", "test.mae"))
-             ),
+  tar_files(sim_main_path,
+            list.files("./Simulation/main", recursive = T, full.names = T)),
   
-  #* Tuning Study ####
-  tar_render(sim_tuning_report,
-             "Simulation/tuning/tuning_study_report.Rmd",
-             output_dir = "Simulation/tuning/",
-             output_file = "sim_tuning_report.html"),
+  success_rate = create_success_rate_table(sim_main_path),
+  
+  tab_binom = make_sim_main_table(success_rate,
+                                  dist = "binomial",
+                                  measures = "auc"),
+  
+  tab_gaus = make_sim_main_table(success_rate,
+                                 dist = "gaussian",
[…]

### 2021-07-27 Manuscript/05-Discussion.Rmd
Run pipeline

--- 2021-07-02 Manuscript/05-Discussion.Rmd
@@
 # Discussion
+In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.
+Meanwhile, we introduced two optimization based algorithms for model fitting.
+The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner, but also provide naive inferential supports for many commonly used hierarchical priors, e.g.
+t-distribution, double exponential distribution, mixture t distribution, and mixture double exponential distribution, in the analyses that the sparse assumption is weak or not necessary.
+Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.
 
-Remaining issue:
-1. Degree of freedom has yet to closely studied. As an direct impact, the functional selection can be improved when effective degree of freedom are more carefully studied
-2.
-While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
-Our current effort resides on expanding.
-An naive approach could be convert an time-to-event outcome to a poisson outcome following xxx.
-Hence, the .
-However, it would be more efficient to , via penalized pseudo likelihood function, which is the current effort resides.
-2.
-Future exapansion, as we were analyzing the metabolites data, we observed that metabolites pathway infromation can be useful in the predictive and integrative model can be developed based on the proposed framework.
-As an naive solution, the selection indicator of linear effect of the spline terms of all metabolites on the same pathway can be set to share the same hyper prior such that the group selection can be modestly achieved.
-Meanwhile further advancedment of the methodology regarding to the integrative analysis can consider setting the hypor priors of the selection indivator of the non-linear components conditioning on the probability of inclusion of the linear effect.
-Such structure is more sensible analytically.
+The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work.
+Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based fast and scalable algorithms.
+Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior.
+Nevertheless, SB-GAM fails to address the bi-level functional selection and the model inference.
+The proposed model provides solutions to these remaining questions while maintaining the same level of prediction accuracy.
+Moreover, the proposed model renders additional flexibility to model specification, allowing various choices of smoothing functions and degrees of freedom.
 
-Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^{pen}$ are independently distributed for analytic simplicity.
-However, this set-up could possibility result in a situation that is not theoretically possible: the non-linear component is selected, but the linear component is not.
-This can be addressed analytically by forcing to include the linear component when non-linear component is included.
+For transnational science purpose, we implemented the proposed model in a R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
+To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones.
+Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.
+
+<!-- the implementation of ssGAM-NM using EM-IRLS.
+Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
+Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
+mixture t-distribution, and mixture double exponential distributions.
+@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors, -->
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- \beta_i \sim N(0, \tau^2_i),\\ -->
+<!-- \tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2) -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- with $\tau^2$ integrated out analytically.
+Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
+One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
+$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
+The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
+[@Rockova2014a] -->
+
+There are some improvements possible for the proposed models.
+First of all, the proposed model achieves a bi-level selection via the two-part spike-and-slab spline prior.
+Nevertheless, this set-up could result in a situation that is not theoretically sound: the non-linear component is selected, but the linear component is not.
+We currently address it analytically by forcing to include the linear component when non-linear component is included.
 Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e.
 $\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.
-
-## Discussion
-
-In this chapter, we propose the spike-and-slab generalized additive models for high-dimensional analysis.Two fast computing algorithms, EM-IRLS and EM-coordinate descent are developed for expedited model fitting.
-EM-IRLS provides uncertainty measures, while EM-coordinate descent algorithm are extremely fast for model fitting.
-In the simulation studies, we shows the proposed model outperforms the state of the art GAM model, `mgcv` in prediction and function inference.
-Moreover, due to the model selection nature of the spike-and-slab prior, the proposed model can identify predictors that have large effects on the outcome.
-In addition, the model can be used to answer the question if non-linear effect are necessary.
+Secondly, while the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
+Our current effort resides on expanding GAM to Cox proportional hazard models.
+An naive approach would be convert an time-to-event outcome to a poisson outcome following Whitehead [@Whitehead1980].
+However, it would be more efficient to directly fitting Cox models via penalized pseudo likelihood function.
+Last but not least, we plan to expand the proposed model to integrate external biology information.
+The main motivation for the integrative model is that biologically informed groupping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
+Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.
+The similar strategy has been used in Ferrari and Dunson [@Ferrari2020].
 
@@
 
-The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not.
+<!-- The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not.
 On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e.
@@
 Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
-
-In the manuscript, we describe the implementation of ssGAM-NM using EM-IRLS.
-Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
-Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
-mixture t-distribution, and mixture double exponential distributions.
-@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors,
-$$
-\begin{aligned}
-\beta_i \sim N(0, \tau^2_i),\\
-\tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2)
-\end{aligned}
-$$
-with $\tau^2$ integrated out analytically.
-Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
-One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
-$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
-The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
-[@Rockova2014a]
+-->

### 2021-07-27 R/functions.R

@@ -1,0 +1,136 @@
+#' Use GAM to screen the features
+#'
+#' @param dat the raw data
+#' @param nrow How many rows are selected
+#'
+#' @return
+#'
+gam_screen <- function(dat, nrow = 200){
+  dat %>% 
+    select(starts_with("mz")) %>% 
+    names() %>% 
+    map_dfr(#feature_name[1:2],
+      .f = function(name, .dat){
+        # name <- feature_name[1]
+        y <- .dat %>% pull(death3yr)
+        x <- .dat %>% pull({{name}})
+        mgcv::gam(y ~ s(x, bs = "cr", k = 10), family = binomial(),
+                  data = data.frame(x, y)) %>% 
+          tidy() %>% 
+          filter(term == "s(x)") %>% 
+          select(p.value) %>% 
+          mutate(var = name,
+                 p.value)
+        
+      },
+      .dat = dat) %>% data.frame(
+        .,
+        p.adj = p.adjust(.$p.value, "fdr")
+      ) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
+    head(nrow)
+}
+
+
+create_success_rate_table <- function(path_list){
+
+  # sims <- list.files(path,full.names =TRUE)
+  # browser()
+  unglue_data(path_list[str_ends(path_list, ".rds")],
+              "{path}/it_{}") %>% 
+    group_by(path) %>% 
+    summarize(n_success = n()) %>% 
+    unglue_unnest(
+      col = path,
+      patterns = "{}/bgam_{study}_-dis_{dist}-p_{p}",
+      remove = FALSE
+    ) %>% 
+    mutate( p = as.numeric(p)) %>% 
+    arrange(dist, p) %>% 
+    select(dist, p, n_success, path)
+  #   head()
+  # 
+  # 
+  # success_rate <- map_dfr(path_list, 
+  #                         .f = function(sim_file){
+  #                           browser()
+  #   sim.df <- unglue_data(sim_file, 
+  #                         "{}/bgam_{study}_-dis_{dist}-p_{p}/{}", 
+  #                         kee) %>% 
+  #     mutate( p = as.numeric(p)) %>% 
+  #     group_by(dist)
+  #     summarize()
+  #   fls <- list.files(sim_file, full.names = TRUE) 
+  #   fls <- fls[grep(".rds", x=fls)]
+  #   n <- length(fls)
+  #   
[…]

### 2021-07-27 Simulation/Code/Main.R

@@ -1,0 +1,243 @@
+args=(commandArgs(TRUE))
+
+# print(args)
+
+if(length(args)==0){
+  print("No arguments supplied.")
+}else{
+  for(i in 1:length(args)){
+    eval(parse(text=args[[i]]))
+  }
+}
+
+# Arguments
+# n_train <- 500
+# n_test <- 1000
+# p <- 50
+# it <- 10
+
+# Required Library
+library(mgcv)
+library(splines)
+library(BhGLM)
+library(tidyverse)
+library(rlang)
+library(cosso)
+# library(HRW)
+
+source("~/bgam_nosharing/Code/helper_func.R")
+
+# Using Array ID as seed ID
+it <- Sys.getenv('SLURM_ARRAY_TASK_ID') %>% as.numeric
+# it <- 1
+
+set.seed(it)
+
+
+# Train Data
+tmp <- sim_Bai_logistic(n_train, p)
+dat <- tmp$dat %>% data.frame
+theta <- tmp$theta
+
+# Test Data
+test_tmp <- sim_Bai_logistic(n_test, p)
+test_dat <- test_tmp$dat %>% data.frame
+test_theta <- test_tmp$theta
+
+
+mgcv_df <- data.frame(
+  Var = setdiff(names(dat), "y"),
+  Func = "s",
+  Args ="bs='cr', k=27"
+)
+
+
+#### mgcv_mdl ####
+
+mgcv_start <- Sys.time()
+mgcv_mdl <- tryCatch({
+  gam(create_HD_formula(y~1, spl_df = mgcv_df), 
+      data=dat %>% data.frame(), family=binomial)
+},
+error = function(err) {
+  mgcv_mdl <- NULL
+  return(NULL)
+}
+)
+mgcv_end <- Sys.time()
+
+mgcv_time <- mgcv_end - mgcv_start
+
+
+mgcv_sum <- NULL
+mgcv_train <- data.frame(deviance = NA, auc = NA, mse = NA, mae = NA, misclassification = NA)
+mgcv_test <- data.frame(deviance = NA, auc = NA, mse = NA, mae = NA, misclassification = NA)
+
+
+if(!is.null(mgcv_mdl)){
+  mgcv_sum <- summary(mgcv_mdl)$s.table[,"p-value"]
+  
+  mgcv_train <- measure.glm(mgcv_mdl$y, mgcv_mdl$fitted.values, family = binomial())
+  
+  
[…]

### 2021-07-27 _targets.R

@@ -1,18 +1,121 @@
-library(targets)
-library(tarchetypes)
-# This is an example _targets.R file. Every
-# {targets} pipeline needs one.
-# Use tar_script() to create _targets.R and tar_edit()
-# to open it again for editing.
-# Then, run tar_make() to run the pipeline
-# and tar_read(summary) to view the results.
+## Load your packages, e.g. library(targets).
+source("./packages.R")
 
-options(tidyverse.quiet = TRUE)
-# Set target-specific options such as packages.
-tar_option_set(packages = c("dplyr", "tidyverse", "rticles"))
+## Load your R files
+lapply(list.files("./R", full.names = TRUE, recursive = TRUE), source)
 
-# End this file with a list of target objects.
-list(
-  # tar_target(data, data.frame(x = sample.int(100), y = sample.int(100))),
-  tar_render(manu, "Manuscript/main.Rmd", output_file = "SS_GAM.pdf")
+## tar_plan supports drake-style targets and also tar_target()
+tar_plan(
+  #  Simulations Studies -----------------------------------------------------
+  #* Main Study ####
+  
+  tar_files(sim_main_path,
+            list.files("./Simulation/main", recursive = T, full.names = T)),
+  
+  success_rate = create_success_rate_table(sim_main_path),
+  
+  tab_binom = make_sim_main_table(success_rate,
+                                  dist = "binomial",
+                                  measures = "auc"),
+  
+  tab_gaus = make_sim_main_table(success_rate,
+                                 dist = "gaussian",
+                                 measures = "R2"),
+  
+  tab_pois = make_sim_main_table(success_rate,
+                                 dist = "poisson",
+                                 measures = "mse"),
+  
+  
+  
+  # 
+  # tar_target(binom_plot,
+  #            make_sim_main_plots(
+  #              success_rate,
+  #              "binomial",
+  #              list("test.deviance", "test.auc", "test.misclassification"))
+  #            ),
+  # 
+  # tar_target(gaussian_plot,
+  #            make_sim_main_plots(
+  #              success_rate,
+  #              "gaussian",
+  #              list("test.R2", "test.mse", "test.mae"))),
+  # 
+  # tar_target(poisson_plot,
+  #            make_sim_main_plots(
+  #              success_rate,
+  #              "poisson",
[…]

### 2021-07-27 _targets_packages.R

@@ -1,3 +1,0 @@
-# Generated by targets::tar_renv(): do not edit by hand
-library(dplyr)
-library(tidyverse)

### 2021-07-31 Manuscript/02-Method.Rmd
Finish revising the conclusion part
Run pipeline

@@
 When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL.
+Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
 \\
 
-```{r fig.height=4, fig.width=6, eval = F}
-# TODO: Refactor the graph to a latex flowchart
-include_graphics("Fig/SS_prior_flow_chart.png")
-```
-
+\begin{figure}
+\centering
 \begin{tikzpicture} [
@@
 
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic mnodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
 ### Other Priors
@@
 These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
-Here we focus on the mixture normal prior as a demonstration of applying continous spike-and-slab prior in additive models.
+Here we focus on the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in additive models.
 ^[TODO: enter a sentence introducing EMVS and spike-and-slab mixture normal prior].

### 2021-08-02 Manuscript/03-Simulation.Rmd
Run pipeline
Update simulation Section

@@
 
-To evaluate the performance of the models, mean squared error, mean absolute error for gaussian model, AUC and misclassifcation rate are calculated for binomial model, and deviance are calculated for Poisson model, averaging across 50 simulations.
-The results are tabulated in Table 1, 2, 3.
+```{r, results = "asis"}
+tab_gaus %>% 
+  xtable(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations", label = "tab:gaus") %>% 
+  print(comment = FALSE)
+```
+
+```{r, results = "asis"}
+tab_binom %>% 
+  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
+  print(comment = FALSE)
+```
+
+To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.^[is it better to provide a box plot or a table.]
 
@@
 
-For the binomial outcome shown in Figure 1, we can see that the proposed models are consistently performed better than COSSO and adaptive COSSO.
-Mgcv slightly outperforms the proposed models in low dimension settings(p = 4, 10), but suffers from overfitting in medium dimension setting (p=50).
-In addition, in the high-dimension setting (p=200), mgcv could not make prediction due to the p > n problem.
-Among the proposed models, we see that there is modest difference between the models that shares the same indicator variable('_spline' models) and or not sharing indicator variable (modles), where the not sharing indicator variable models have slightly edge over indicator sharing model.
-Moreover, the mixture double exponential model and lasso model perform substaintially better in the medium and high settings.
-For the remaining two outcomes, the results shows similar trends, where mixture double exponential model and ss lasso model are recommended for use.
+The predictive performances have a consistent pattern across the three distributions of outcomes.
+Across all the scenarios, COSSO and adaptive COSSO have the least favourable performance among the applicable methods examined.
+To note, mgcv doesn't support high-dimensional analysis when the number of coefficients are greater than the sample size.
+mgcv predicts well when $p$ is small or moderate (p = 4, 10), and deteriorate when the number of predictors increase.
+Among the three fast-computing Bayesian hierarchical models, the proposed models, bglm_spline_de and blasso_spline ^[need to change the name] predicts better than SB-GAM when the dimension of are small, moderate, and high.
+However, in the hyper high dimensional case, SB-GAM has better performance.
+Among the proposed methods, bglm_spline_de performs consistently better than blasso_spline.
 
-%are calculated to assess out-of-sample estimation accuracy,
+<!-- %are calculated to assess out-of-sample estimation accuracy, -->
 
-%We also calculated the area under the curve (AUC) of the receiver operating characteristic curve for $y_{new}$.
+<!-- We also calculated the area under the curve (AUC) of the receiver operating characteristic curve for $y_{new}$.
 We will also visually display the estimated smooth function with the smooth function of truth to compare if the models can infer the smooth function correctly, despite the prediction accuracy.
-
-%When constructing the smooth function of each variables, we constructs the natural cubic spline with degree of freedom 25.
-In other words, the design matrix of each smooth function will have 26 columns.
-The total number of columns for $p=4, 10, 50$ is $105, 261, 1301$ respectively, where intercept is kept in the model when fitting.
-The $p=50$ can mimic the high dimensional setting as the design matrix are in high-dimensional setting ($p >> n$)
+-->
 
@@
 
-%We benchmark the performance of the proposed methods with the state of art GAM package `mgcv`[TODO: add citation].
-In the future steps, we will compare our proposed model with other high dimensional GAM applications.
-
  \
@@
 
-Overall, the proposed priors outperforms the state-of-art GAM model `mgcv` when the dimensionality is low, median, and high.
+<!-- Overall, the proposed priors outperforms the state-of-art GAM model `mgcv` when the dimensionality is low, median, and high.
 The proposed priors predicts better than `mgcv` and infers better of the underlying smooth function.
@@
 when $p=50$ where the number of parameters are greater than the sample size.
+-->
 
-#### Prediction Performace
+<!-- #### Prediction Performace -->
 
-The prediction performance is tabulated in the table below for AUC and Supplementary tables for MSE and misclassification rate.
+<!-- The prediction performance is tabulated in the table below for AUC and Supplementary tables for MSE and misclassification rate.
 In summary, the proposed Bayesian models outperforms the benchmark model `mgcv` regardless the choice of prior or fitting algorithm.
@@
 This is confirmed with the visualization in the following section.
+-->
 
@@
 
-#### Visualization
-We plot the estimated smooth functions of the four active variables for the 100 simulation iterations, along with the 'true' smooth function.
+<!-- #### Visualization -->
+<!-- We plot the estimated smooth functions of the four active variables for the 100 simulation iterations, along with the 'true' smooth function.
 Across different settings of predictors, the patterns look similar.
@@
 On the contrary, the SS double exponential prior model encourages sparse solutions, which can sometime omit active predictors.
+-->

### 2021-08-02 _targets.R

@@ -121,17 +121,34 @@ tar_plan(
   
 # ** bgam --------------------------------------------------------------
 
+  base_mdl = lm(out_HOMA_PC~cov_Age + cov_Sex + cov_Race + cov_Triglycerides_Baseline + cov_Weight_PC,
+                data = ful_dat),
+# mean(base_mdl$residuals^2)
+
+  # data.frame(ful_dat$out_HOMA_PC, bgam_mdl$y, bgam_mdl$linear.predictors, base_mdl$fitted.values) %>% head()
+
   bgam_mdl = make_bgam_mdl(ful_dat),
-  
+  # mean((bgam_mdl$y - bgam_mdl$linear.predictors)^2),  
+
+
   bgam_dml_tun = tune.bgam(bgam_mdl, nfolds = 5, s0 = seq(0.005, 0.1, 0.01)),
   
-  cv_bai_mdl = cv.SBGAM(y = ful_dat$out_HOMA_PC, X = ful_dat %>% select(starts_with("mb_")) %>% as.matrix,  df=5, family = "gaussian", a = 1, b = 1,
-                       max.iter=100, tol = 1e-6),
+  # bgam_mdl = make_bgam_mdl(ful_dat),
+  
 
 
-  bai_mdl = SBGAM(ful_dat$out_HOMA_PC, X = ful_dat %>% select(starts_with("mb_")) %>% as.matrix, df=5, family = "gaussian" ,
-                 lambda0 = cv_bai_mdl$lambda0.min, a = 1, b = 1,
-                 max.iter=100, tol = 1e-6, print.iter=FALSE),
+  # cv_bai_mdl = cv.SBGAM(y = ful_dat$out_HOMA_PC, X = ful_dat %>% select(starts_with("mb_")) %>% as.matrix,  df=5, family = "gaussian", a = 1, b = 1,
+  #                      max.iter=100, tol = 1e-6),
+
+
+  # bai_mdl = SBGAM(ful_dat$out_HOMA_PC, X = ful_dat %>% select(starts_with("mb_")) %>% as.matrix, df=5, family = "gaussian" ,
+  #                lambda0 = cv_bai_mdl$lambda0.min, a = 1, b = 1,
+  #                max.iter=100, tol = 1e-6, print.iter=FALSE),
+  # 
+  # cv_bai_mdl_final = cv.SBGAM(y = ful_dat$out_HOMA_PC, X = ful_dat %>% select(starts_with("mb_")) %>% as.matrix, 
[…]

### 2021-08-02 Manuscript/SS_GAM.tex
Run pipeline

@@
 generalized additive models with the spike-and-slab lasso prior as the
-BHAM-SSL. \textbackslash{}
-
+BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
+\textbackslash{}
+
+\begin{figure}
+\centering
 \begin{tikzpicture} [
@@
 
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic mnodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
 \hypertarget{other-priors}{%
@@
 necessary. Here we focus on the mixture normal prior as a demonstration
-of applying continous spike-and-slab prior in additive models.
+of applying continuous spike-and-slab prior in additive models.
 \footnote{TODO: enter a sentence introducing EMVS and spike-and-slab
@@
   \hline
+1 & 4.00 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
+  2 & 10.00 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
+  3 & 50.00 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
+  4 & 200.00 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & NaN (NA) & 0.81 (0.05) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample area under the curve measure over for binomial outcome over 50 iterations} 
+\label{tab:gaus}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrllllll}
+  \hline
+ & p & bglm\_spline\_de & blasso\_spline & cosso & acosso & mgcv & SB\_GAM \\ 
+  \hline
 1 & 4.00 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
@@
 
-To evaluate the performance of the models, mean squared error, mean
-absolute error for gaussian model, AUC and misclassifcation rate are
-calculated for binomial model, and deviance are calculated for Poisson
+To evaluate the predictive performance of the models, \(R^2\) for
+Gaussian model, area under the curve (AUC) are calculated for binomial
 model, averaging across 50 simulations. The results are tabulated in
-Table 1, 2, 3.
-
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-8-1.pdf}
+Table \ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to
+  provide a box plot or a table.}
 
@@
 \includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-11-1.pdf}
 

### 2021-08-02 Manuscript/04-Real_Data.Rmd
Run pipeline

# Metabolomics Data Analysis

We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study ^[TODO: add citation:https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0240764] on the association between metabolic biomarkers and weight loss. The data contained 765 metaboliltes in baseline plasma collected from three separate studies, profiled using liquid chromatography mass spectrometry. Quality control and natural log transformation were performed during metabolites data preparation. The outcome of interest was percent change in insulin resistance, and hence modeled using a Gaussian model adjusting for covariates including age, sex, race, baseline clinical triglycerides, and percent change in weight over the follow-up time period. 731 individuals were included in the initial studies, and reduced to `r nrow(ful_dat)` after removing missing values. The descriptive statistics were tabulated in Table \ref{tab:tbl_one}.

```{r, results = "asis", eval = F}
tbl_one %>% 
  xtable(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies", label = "tab:tbl_one") %>% 
  print(comment = FALSE)
```

```{r}
tbl_one %>% 
  as_flex_table() %>% 
  flextable::autofit() %>% 
  flextable::set_caption(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies\\label{tab:tbl_one}")
```

### 2021-08-02 Manuscript/SS_GAM.tex
Run pipeline

@@
 \usepackage{pgfplots}
+\usepackage{multirow}
+\usepackage{multicol}
+\usepackage{colortbl}
+\usepackage{hhline}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{hyperref}
 
@@
 \end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measure over for binomial outcome over 50 iterations} 
+\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations} 
 \label{tab:gaus}
@@
 \end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measure over for binomial outcome over 50 iterations} 
+\caption{The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations} 
 \label{tab:bin_auc}
@@
 
-To evaluate the predictive performance of the models, \(R^2\) for
-Gaussian model, area under the curve (AUC) are calculated for binomial
-model, averaging across 50 simulations. The results are tabulated in
-Table \ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to
-  provide a box plot or a table.}
+To evaluate the predictive performance of the models, the statistics
+caluclated based on the testing dataset, \(R^2\) for Gaussian model,
+area under the curve (AUC) for binomial model, are calculated averaging
+across 50 simulations. The results are tabulated in Table
+\ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to provide a
+  box plot or a table.}
 
@@
 
-For the binomial outcome shown in Figure 1, we can see that the proposed
-models are consistently performed better than COSSO and adaptive COSSO.
-Mgcv slightly outperforms the proposed models in low dimension
-settings(p = 4, 10), but suffers from overfitting in medium dimension
-setting (p=50). In addition, in the high-dimension setting (p=200), mgcv
-could not make prediction due to the p \textgreater{} n problem. Among
-the proposed models, we see that there is modest difference between the
-models that shares the same indicator variable('\_spline' models) and or
-not sharing indicator variable (modles), where the not sharing indicator
-variable models have slightly edge over indicator sharing model.
-Moreover, the mixture double exponential model and lasso model perform
-substaintially better in the medium and high settings. For the remaining
-two outcomes, the results shows similar trends, where mixture double
-exponential model and ss lasso model are recommended for use.
-
-\%are calculated to assess out-of-sample estimation accuracy,
-
-\%We also calculated the area under the curve (AUC) of the receiver
-operating characteristic curve for \(y_{new}\). We will also visually
-display the estimated smooth function with the smooth function of truth
-to compare if the models can infer the smooth function correctly,
-despite the prediction accuracy.
-
-\%When constructing the smooth function of each variables, we constructs
-the natural cubic spline with degree of freedom 25. In other words, the
-design matrix of each smooth function will have 26 columns. The total
-number of columns for \(p=4, 10, 50\) is \(105, 261, 1301\)
-respectively, where intercept is kept in the model when fitting. The
-\(p=50\) can mimic the high dimensional setting as the design matrix are
-in high-dimensional setting (\(p >> n\))
-
-\%We benchmark the performance of the proposed methods with the state of
-art GAM package \texttt{mgcv}{[}TODO: add citation{]}. In the future
-steps, we will compare our proposed model with other high dimensional
-GAM applications.
+The predictive performances have a consistent pattern across the three
+distributions of outcomes. Across all the scenarios, COSSO and adaptive
+COSSO have the least favourable performance among the applicable methods
+examined. To note, mgcv doesn't support high-dimensional analysis when
+the number of coefficients are greater than the sample size. mgcv
+predicts well when \(p\) is small or moderate (p = 4, 10), and
+deteriorate when the number of predictors increase. Among the three
+fast-computing Bayesian hierarchical models, the proposed models,
+bglm\_spline\_de and blasso\_spline \footnote{need to change the name}
+predicts better than SB-GAM when the dimension of are small, moderate,
+and high. However, in the hyper high dimensional case, SB-GAM has better
+performance. Among the proposed methods, bglm\_spline\_de performs
+consistently better than blasso\_spline.
 
@@
 
-Overall, the proposed priors outperforms the state-of-art GAM model
-\texttt{mgcv} when the dimensionality is low, median, and high. The
-proposed priors predicts better than \texttt{mgcv} and infers better of
-the underlying smooth function. Most importantly, \texttt{mgcv} is
-infeasible in the high dimensional setting, i.e.~when \(p=50\) where the
-number of parameters are greater than the sample size.
-
-\hypertarget{prediction-performace}{%
-\paragraph{Prediction Performace}\label{prediction-performace}}
-
-The prediction performance is tabulated in the table below for AUC and
-Supplementary tables for MSE and misclassification rate. In summary, the
-proposed Bayesian models outperforms the benchmark model \texttt{mgcv}
-regardless the choice of prior or fitting algorithm. The models with
-IRLS fitting algorithm give slightly better performance than the the
-model with cyclic coordinate descent algorithm. The advantage diminishes
-in the high dimensional setting (p=50). This is possible because the SS
-normal spline prior provides a more smooth solution, which would work
-well in a low to median dimension setting. However, in the high
-dimensional setting, there are not enough sparsity in the model, i.e.~SS
-normal prior will provides a more complex model than necessary. This is
-confirmed with the visualization in the following section.
-
-\hypertarget{visualization}{%
-\paragraph{Visualization}\label{visualization}}
-
-We plot the estimated smooth functions of the four active variables for
-the 100 simulation iterations, along with the `true' smooth function.
-Across different settings of predictors, the patterns look similar.
-First of all, in all the settings, \texttt{mgcv} tends to provide more
-extreme models, where in the graphs the smooth function have much higher
-bound than the true smooth function. Rarely the estimated smooth
-function from \texttt{mgcv} matches with the true smooth function. In
-comparison, the proposed method can interpolate the shape of the true
-smooth function. However, different priors and fitting algorithms have
-different advantages. For example, in Figure 1, we see that the SS
-normal distribution have problem with the tail of the smooth function,
-especially obvious in the estimations of the variable \emph{x4}.
-However, it would have less sparse solutions, which is more likely to
-accurately interpolate the smooth function especially when the shape is
-complicated (see variable \emph{x2}). On the contrary, the SS double
-exponential prior model encourages sparse solutions, which can sometime
-omit active predictors.
-
-\hypertarget{real-data-analysis}{%
-\section{Real Data Analysis}\label{real-data-analysis}}
+\hypertarget{metabolomics-data-analysis}{%
+\section{Metabolomics Data Analysis}\label{metabolomics-data-analysis}}
+
+We use the proposed models BHAM-SSL to analyze metabolomics data from a
+recently published study \footnote{TODO: add
+  citation:\url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0240764}}
+on the association between metabolic biomarkers and weight loss. The
+data contained 765 metaboliltes in baseline plasma collected from three
+separate studies, profiled using liquid chromatography mass
+spectrometry. Quality control and natural log transformation were
+performed during metabolites data preparation. The outcome of interest
+was percent change in insulin resistance, and hence modeled using a
+Gaussian model adjusting for covariates including age, sex, race,
+baseline clinical triglycerides, and percent change in weight over the
+follow-up time period. 731 individuals were included in the initial
+studies, and reduced to 315 after removing missing values. The
+descriptive statistics were tabulated in Table \ref{tab:tbl_one}.
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{2.14in}|p{1.44in}|p{1.56in}|p{1.75in}|p{1.39in}}
+
+\caption{The descriptive statistics of the sample in the metabolomics data analysis stratified by studies\label{tab:tbl_one}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Characteristic}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Overall,\ N\ =\ 315}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CBD,\ N\ =\ 17}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{STRRIDEPD,\ N\ =\ 56}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM,\ N\ =\ 242}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{000000}\cline{1-5}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Characteristic}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Overall,\ N\ =\ 315}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CBD,\ N\ =\ 17}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{STRRIDEPD,\ N\ =\ 56}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM,\ N\ =\ 242}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{000000}\cline{1-5}\endhead
+
+\multicolumn{5}{!{\color[HTML]{FFFFFF}\vrule width 0pt}>{\raggedright}p{\dimexpr 8.28in+8\tabcolsep+4\arrayrulewidth}!{\color[HTML]{FFFFFF}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Median\ (IQR);\ n\ (\%)}}\fontsize{11}{11}\selectfont{\textsuperscript{\textcolor[HTML]{000000}{1}}}} \\
+
+\endfoot
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{cov\_Age}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{56\ (51,\ 62)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{39\ (34,\ 47)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{60\ (55,\ 66)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{56\ (51,\ 62)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{cov\_Sex}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{F}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{188\ (60\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{14\ (82\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{33\ (59\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{141\ (58\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{M}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{127\ (40\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3\ (18\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{23\ (41\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{101\ (42\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{cov\_Race}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{AA}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{109\ (35\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{8\ (47\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{9\ (16\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{92\ (38\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{EA}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{203\ (64\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{9\ (53\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{45\ (80\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{149\ (62\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{O}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3\ (1.0\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0\ (0\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2\ (3.6\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1\ (0.4\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{cov\_Triglycerides\_Baseline}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{106\ (76,\ 154)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{118\ (91,\ 170)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{104\ (75,\ 150)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{106\ (76,\ 154)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.14in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{cov\_Weight\_PC}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.44in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-7.6\ (-10.9,\ -5.6)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.56in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-29.9\ (-37.0,\ -25.1)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-3.1\ (-5.7,\ -0.6)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-7.8\ (-10.8,\ -6.2)}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{000000}\cline{1-5}
+
+\end{longtable}
 

### 2021-08-02 packages.R

@@ -17,6 +17,7 @@ library(knitr)
 library(ggpubr)
 library(xtable)
 library(janitor)
+library(flextable)
 
 
 library(BHAM)

### 2021-08-06 Manuscript/02-Method.Rmd
Run pipeline

@@
 \matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
-&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
-&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
 };
-
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-\node (penTheta) [outCome] { $\theta_{j}^\tp\sim Beta(a, b)$ };\\
   \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j^\tp)$ };\\
@@
 
-\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
-%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-
 \node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta^0_1, \bs \beta^\tp_1, \dots,\beta^0_j, \bs \beta^\tp_j , \dots,\beta^0_p, \bs \beta^\tp_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
 \node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
 
-%Lines
-
-\draw[->] (linearTheta) -- (linearGamma);
-\draw[->] (penTheta) -- (penGamma);
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
 \draw[->] (linearGamma) -- (linearBeta) ;
 \draw[->] (penGamma) -- (penBeta);
-\draw[->] (ab) -- (linearTheta);
-\draw[->] (ab) -- (penTheta);
+\draw[->] (ab) -- (Theta);
 \draw[->] (s) -- (linearBeta) ;
@@
 
-% Now that the diagram has been drawn, background rectangles
-% can be fitted to its elements. This requires the TikZ
-% libraries "fit" and "background".
-% Control input and measurement are labeled. These labels have
-% not been translated to English as "Measurement" instead of
-% "Messung" would not look good due to it being too long a word.
 \begin{pgfonlayer}{background}
   \node [background,
-                fit=(linearTheta) (linearBeta),
+   fit=(linearGamma) (linearBeta),
    label=above:Null Space:] {};
   \node [background,
-                 fit=(penTheta) (penBeta),
+    fit=(penGamma) (penBeta),
     label=above:Penalized Space:] {};
+\end{pgfonlayer}
 
-\end{pgfonlayer}
 \end{tikzpicture}

### 2021-08-06 Manuscript/SS_GAM.tex
Run pipeline

@@
 \matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
-&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
-&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
 };
-
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-\node (penTheta) [outCome] { $\theta_{j}^\text{pen}\sim Beta(a, b)$ };\\
   \node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j^\text{pen})$ };\\
@@
 
-\node (ab)[staticCompo] at ($(linearTheta)!0.5!(penTheta)$) {$(a, b)$};
-%\node (PenaltyPrior) [PriorBoarder,  right = 1cm of ab] {part 2};
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-
 \node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^\text{pen}_1, \dots,\beta^0_j, \boldsymbol{\beta}^\text{pen}_j , \dots,\beta^0_p, \boldsymbol{\beta}^\text{pen}_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
 \node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
 
-%Lines
-
-\draw[->] (linearTheta) -- (linearGamma);
-\draw[->] (penTheta) -- (penGamma);
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
 \draw[->] (linearGamma) -- (linearBeta) ;
 \draw[->] (penGamma) -- (penBeta);
-\draw[->] (ab) -- (linearTheta);
-\draw[->] (ab) -- (penTheta);
+\draw[->] (ab) -- (Theta);
 \draw[->] (s) -- (linearBeta) ;
@@
 
-% Now that the diagram has been drawn, background rectangles
-% can be fitted to its elements. This requires the TikZ
-% libraries "fit" and "background".
-% Control input and measurement are labeled. These labels have
-% not been translated to English as "Measurement" instead of
-% "Messung" would not look good due to it being too long a word.
 \begin{pgfonlayer}{background}
   \node [background,
-                fit=(linearTheta) (linearBeta),
+   fit=(linearGamma) (linearBeta),
    label=above:Null Space:] {};
   \node [background,
-                 fit=(penTheta) (penBeta),
+    fit=(penGamma) (penBeta),
     label=above:Penalized Space:] {};
-
 \end{pgfonlayer}
+
 \end{tikzpicture}

### 2021-08-19 _targets.R
Run pipeline

@@ -107,7 +107,11 @@ tar_plan(
   ful_dat_raw = left_join(cov_dat, mb_dat),
   
   ful_dat = ful_dat_raw %>% 
-    filter(complete.cases(.)),
+    filter(complete.cases(.)) %>% 
+    filter(out_HOMA_PC < 100,
+           cov_Race != "O"),
+  train_dat = ful_dat %>% filter(Study == "WLM"),
+  test_dat = ful_dat %>% filter(Study != "WLM"),
   
   # * Analysis --------------------------------------------------------------
 
@@ -121,11 +125,26 @@ tar_plan(
   
 # ** bgam --------------------------------------------------------------
 
+  ## Linear Regression
   base_mdl = lm(out_HOMA_PC~cov_Age + cov_Sex + cov_Race + cov_Triglycerides_Baseline + cov_Weight_PC,
                 data = ful_dat),
 # mean(base_mdl$residuals^2)
 
   # data.frame(ful_dat$out_HOMA_PC, bgam_mdl$y, bgam_mdl$linear.predictors, base_mdl$fitted.values) %>% head()
+
+  ## Lasso Model
+  cv_lasso_mdl = cv.glmnet(x = train_dat %>%
+                             select(starts_with("cov"), starts_with("mb")) %>%
+                             data.matrix(),
+            y = train_dat$out_HOMA_PC,
+            penalty.factor = c(rep(0, 5), rep(1, 484))),
+
+  lasso_mdl = glmnet(x = train_dat %>%
+                       select(starts_with("cov"), starts_with("mb")) %>%
+                       data.matrix(),
+                     y = train_dat$out_HOMA_PC,
+                     lambda = cv_lasso_mdl$lambda.min,
+                     penalty.factor = c(rep(0, 5), rep(1, 484))),
 
   bgam_mdl = make_bgam_mdl(ful_dat),
   # mean((bgam_mdl$y - bgam_mdl$linear.predictors)^2),  

### 2021-08-19 Manuscript/03-Simulation.Rmd
Working on real_data analysis

@@
 
-```{r, fig.height=8}
+```{r, fig.height=8, eval = FALSE}
 targets::tar_load(binom_plot)
@@
 
-```{r, fig.height=8}
+```{r, fig.height=8, eval = FALSE}
 targets::tar_load(gaussian_plot)
@@
 
-```{r, fig.height=8}
+```{r, fig.height=8, eval = FALSE}
 targets::tar_load(poisson_plot)

### 2021-08-19 Real_Data/Emory_Card_Biobank/Report.Rmd
Working on real_data analysis

@@
 train_dat <- targets::tar_read(RD_ECB_train_dat)
-valid_dat <- targets::tar_read(RD_ECB_valid_dat)
-
 gam_scn_res <- targets::tar_read(train_gam_screen)
@@
 
-* Two data sets: model development set (N=`r NROW(train_dat)`), validation set (N=`r NROW(valid_dat)`)
+* Two data sets: model development set (N=`r NROW(train_dat)`)
 
@@
 ```
-
-## Validation
-
-The validation data set have mostly distinct feature from the training data set. Hence, the validation in the traditional sense does not exists.
-
-```{r eval = F}
-
-test_mz_names <- valid_dat  %>% select(starts_with("mz")) %>% names
-measure.bh(mdl_defaul, new.x = test_sm_dat, new.y=valid_dat$death3yr)
-
-measure.bh(mdl_dflt)
-```

### 2021-08-19 R/make_bgam_dat.R
Run pipeline

@@ -1,0 +1,45 @@
+make_bgam_dat <- function(train_dat, test_dat = NULL){
+  
+  # Create sm dat -----------------------------------------------------------
+  mz_names <- train_dat %>% names
+  mz_names <- mz_names[grep("mb_", mz_names, fixed = TRUE)] 
+  
+  
+  sm_df <- data.frame(
+    Var = mz_names,
+    Func = "s",
+    Args ="bs='cr', k=5"
+  )
+  
+  sm_obj <- construct_smooth_data(sm_df, train_dat)
+  
+  dsn_mat <- sm_obj$data
+  spl_pt_name <- names(dsn_mat)
+  
+  cov_mat <- model.matrix(~ -1 +cov_Age + cov_Sex + cov_Race + cov_Triglycerides_Baseline + cov_Weight_PC,
+                          train_dat %>% select(starts_with("cov_")))
+  dsn_mat <- data.frame(
+    cov_mat,
+    dsn_mat)
+  
+  spl_grp <- make_group(spl_pt_name)
+  
+  ret <- list(train_dat = dsn_mat,
+              group = spl_grp)
+  
+  # Create test dat ---------------------------------------------------------
+  if(!is.null(test_dat)){
+    
+    test_sm_dat <- make_predict_dat(sm_obj$Smooth, dat = test_dat)
+    test_cov_mat <- model.matrix(~ -1 +cov_Age + cov_Sex + cov_Race + cov_Triglycerides_Baseline + cov_Weight_PC,
+                                 test_dat %>% select(starts_with("cov_")))
+    test_dsn_mat <- data.frame(
+      test_cov_mat,
+      test_sm_dat
+    )
+    
+    ret$test_dat <- test_dsn_mat
+  }
+  
+  return(ret)
+}

### 2021-08-19 _targets.R

@@ -60,27 +60,27 @@ tar_plan(
   #  Real Data Analysis -----------------------------------------------------
   # #* Emory Card Biobank ####
   # #** Load Data ####
-  # tar_target(realdata_ECB_train_path,
-  #            "Real_Data/Emory_Card_Biobank/Data/Analysis_data_first_cohort.csv",
-  #            format = "file"),
+  tar_target(realdata_ECB_train_path,
+             "Real_Data/Emory_Card_Biobank/Data/Analysis_data_first_cohort.csv",
+             format = "file"),
   # tar_target(realdata_ECB_valid_path,
   #            "Real_Data/Emory_Card_Biobank/Data/Analysis_data_second_cohort.csv",
   #            format = "file"),
-  # tar_target(RD_ECB_train_dat,
-  #            readr::read_csv(realdata_ECB_train_path)),
+  tar_target(RD_ECB_train_dat,
+             readr::read_csv(realdata_ECB_train_path)),
   # tar_target(RD_ECB_valid_dat,
   #            readr::read_csv(realdata_ECB_valid_path)), 
   # 
   # #** GAM Screening ####
-  # tar_target(train_gam_screen,
-  #            gam_screen(RD_ECB_train_dat)),
-  # 
+  tar_target(train_gam_screen,
+             gam_screen(RD_ECB_train_dat)),
+
   # #** Report ####
-  # tar_render(real_data_report_ECB,
-  #            "Real_Data/Emory_Card_Biobank/Report.Rmd",
-  #            output_dir = "Real_Data/Emory_Card_Biobank/",
-  #            output_file = "ECB_report.html"),
-  
+  tar_render(real_data_report_ECB,
+             "Real_Data/Emory_Card_Biobank/Report.Rmd",
+             output_dir = "Real_Data/Emory_Card_Biobank/",
+             output_file = "ECB_report.html"),
+
   
   # Dataset 2 ---------------------------------------------------------------
   
@@ -95,7 +95,7 @@ tar_plan(
     clean_meta_dat(),
   
   tar_target(cov_path,
-             "./Real_Data/Copy of Adult_Obesity_Nontargeted_Metabolomics_Sample_Metadata.csv",
+             "./Real_Data/Adult_Obesity_Nontargeted_Metabolomics_Sample_Metadata.csv",
              format = "file"),
[…]

### 2021-08-22 Manuscript/01-Intro.Rmd
Run pipeline

@@
 
-Using additive models to interpolate nonlinear effects has a long history.
+Many modern biomedical research, e.g.
+sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
+There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review.
+These models are built upon a restrictive and unrealistic assumption, linearity.
+In classical statistical modelling, many modeling strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
+For example, variable categorization is a simple and common practice in epidemiology, but suffers from power and interpretation issues.
+More complex models to address nonlinear effects include random forest and other so-called "black box" models [@Breiman2001].
+These models are not particularly attractive in biomedical studies, as they lack the transparency to understand the etiology of diseases.
+In addition, how to generalized these models to the high-dimensional setting remains unclear.
+
+<!-- Using additive models to interpolate nonlinear effects has a long history.
 Early work of additive models can date back to 1979 when Cleveland [@Cleveland1979] proposed local smoothing for bivariate analysis.
@@
 Such models generally have improved prediction accuracy over classic statistical models.
-However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models [@Breiman2001].
-Given that the primary objective of biomedical studies is to understand the etiology of diseases, "Black box" models are less helpful as they can hardly explain how the change of biomarkers is associated with the disease.
+-->
 
-Semi-parametric regression models are a more appropriate alternative to the "black-box" models, thanks to their balancing between model flexibility and interpretability.
+<!-- However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models [@Breiman2001].
+Given that the primary objective of biomedical studies is to understand , "Black box" models are less helpful as they can hardly explain how the change of biomarkers is associated with the disease.
+-->
+
+Semi-parametric regression models are an appropriate alternative to the "black-box" models thanks to their balance between model flexibility and interpretability.
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
@@
 $$
-where $g^{-1}$ is the inverse function of a smooth monotonic link function.
+where $g^{-1}$ is the inverse function of a monotonic link function.
 The smooth functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
-Nevertheless, the classic GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
+Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
 
-Many modern biomedical research, e.g.
-sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
-There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review.
-These models assume sparsity among the predictors, i.e.
+There exists some proposals to generalized the classical GAM to accommodate high-dimensional applications.
+The regularized models that are primarily designed for linear models are used to fit GAMs by accounting for the group structure introduced when expanding smoothing functions.
+<!-- These models assume sparsity among the predictors, i.e.
 most predictors have small or no effects on the outcome.
 Further development focuses on addressing the grouping structure among the predictors, where AMs and GAMs can be considered special cases.
-Various penalties are imposed on the basis function coefficients: Ravikumar et al.
+Various penalties are imposed on the basis function coefficients:  -->
+Ravikumar et al.
 [@Ravikumar2009] extended the grouped lasso [@Yuan2006] to additive models; Huang et al.
@@
 [@Bai2020] was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
+
+TODO: emphasize it doesn't accounting for smoothing penalty, and possibly lead to overshinkage, resulting wrong estimation of the smoothing function and prediction.
+
 Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.

### 2021-08-31 Manuscript/01-Intro.Rmd
Update Real Data Analysis

@@
 Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.
-Scheipl et al.
-[@Scheipl2012] proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma, specifically for semi-parametric regression using MCMC.
-The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
 Bai et al.
 [@Bai2020] was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
-
-TODO: emphasize it doesn't accounting for smoothing penalty, and possibly lead to overshinkage, resulting wrong estimation of the smoothing function and prediction.
-
-Regardless of the nice statistical properties, a practical and pivotal question, if non-linear effects exist, yet to be addressed by these methods.
-Besides Scheipl et al.
-[@Scheipl2012], most proposed methods adapt an 'all-in-all-out' strategy, i.e.
-either including or excluding the variable completely, rendering no space for within group selection.
-In addition, a common criticism received by these methods is that the basis function coefficients tend to be overly penalized due to the sparsity assumption directly applied on the basis function coefficients.[@Scheipl2013] It would be of special interest to develop a fast yet flexible generalized additive model.
+As these methods focus on addressing the sparsity, it is very likely to overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
+[@Scheipl2013] In addition, the proposed methods adapt an 'all-in-all-out' strategy, i.e.
+either including or excluding the variable completely, rendering no space for bi-level selection.
+Scheipl et al.
+[@Scheipl2012] proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma prior, that address the previous challenges.
+But the model fitting relies on computational intensive MCMC algorithmic and creates scalability concern.
+ <!-- specifically for semi-parametric regression using MCMC.
+The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
+-->
+It would be of special interest to develop a fast, flexible and accurate generalized additive model.
 
 To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we impose a new bi-part spike-and-slab spline prior on the smooth functions to perform bilevel selection, where the linear and nonlinear spaces of smooth functions can be selected separately.
-The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
-Additionally, two fast computing optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
-Meanwhile, to avoid the over-penalization of the basis function coefficients, smoothness penalties are incorporated in the model via re-parameterization.
+Specifically, we incorporate smoothness penalties in the model via re-parameterization of the smoothing function to avoid over-penalizing basis function coefficients.
 Smoothness penalties are commonly implemented in GAMs through smoothing and penalized regression splines.
 They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
-For the most common cases of cubic splines, the smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
+The smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
 \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{"}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
+  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
 \end{equation}
@@
 Smoothness penalties are also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-Moreover, incorporating the smoothness penalty allows the separation of the linear space of the smooth function, also know as the null space, from the nonlinear space.
-Marra and Wood [@Marra2011] introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
+Moreover, incorporating the smoothness penalty allows the separation of the linear space of smooth functions, also know as the null space, from the nonlinear space.
+We then impose a new bi-part spike-and-slab spline prior on the smooth functions for bi-level selection such that the linear and nonlinear spaces of smooth functions can be selected separately.
+The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
+In addition, two fast computing optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
+
+<!-- Marra and Wood [@Marra2011] introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
 The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
-In addition, the priors of the linear and nonlinear effects are independent, creating computational convenience and further improving the scalability of the model.
+-->
+ <!-- In addition, the priors of the linear and nonlinear effects are independent, creating computational convenience and further improving the scalability of the model. -->
 
@@
 EM-IWLS complements the lack of inference potential from EM-CD in a median and high dimensional setting.
-Furthermore, the proposed model addresses the incapability of bilevel selection in SB-GAM.
+Furthermore, the proposed model addresses the incapability of bi-level selection in SB-GAM.
 Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), make transnational science more accessible.
@@
 -->
-
-<!-- Lastly, The third aim of the dissertation work will focus on the software implementation of the proposed models and algorithms.
-We will implement the proposed models and algorithms in one of the most commonly used statistical computing environment R.
-The R implementation of the Bayesian hierarchical additive model is available via a freely available R package, BHAM at (https://github.com/boyiguo1/BHAM)[https://github.com/boyiguo1/BHAM].
--->

### 2021-08-31 packages.R

@@ -23,6 +23,7 @@ library(flextable)
 library(BHAM)
 library(BhGLM)
 library(sparseGAM)
+library(cosso)
 library(glmnet)
 
 conflict_prefer("filter", "dplyr")

### 2021-08-31 Manuscript/04-Real_Data.Rmd
Run pipeline

@@
 
+In this section, we apply the proposed models BHAM-SSL to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
+
+## Emory Cardiovarcular Biobank
+We use the proposed models BHAM-SSL to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
+The dataset is publicly available via \textit{Dryad} [@Mehta2020_data].
+It contains in total of 776 subjects from two cohorts.
+As there is a large number of non-overlapping features among the two cohorts, we use the cohort with larger sample size (N=454).
+There were initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
+Hence, we performed a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and chose the the top 200 features with smallest p-values.
+5-Knot spline additive models for binary outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
+10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+Out-of-bag samples were used for prediction performance evaluation, where deviance, area under the ROC curve (AUC), mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
+
+```{r}
+
+bam_mdl_res <- bamlasso_tune %>% 
+  arrange(desc(auc)) %>% 
+  slice(1) %>% 
+  select(-s0)
+
+rbind(
+  'BHAM-SSL' = bam_mdl_res,
+  'SB-GAM' = bai_cv_res %>% t %>%  data.frame()
+) %>% 
+  flextable() %>% 
+  flextable::autofit() %>% 
+  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}")
+```
+
+## Weight Loss Maintenance Cohort 
 We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study ^[TODO: add citation:https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0240764] on the association between metabolic biomarkers and weight loss.
@@
 
-```{r, results = "asis", eval = F}
-tbl_one %>% 
-  xtable(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies", label = "tab:tbl_one") %>% 
-  print(comment = FALSE)
-```
-
 ```{r}

### 2021-09-03 Manuscript/00-main.Rmd
Run pipeline
Update Real Data Analysis
Update draft

@@
   org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "*Corresponding author name, This is sample corresponding address. \\email{xxx@uab}"
+corres: "Nengjun Yi, epartment of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{nyi@uab.edu}"
 presentaddress: This is sample for present address text this is sample for present address text
@@
 tar_load(tab_gaus)
+
+tar_load(ful_dat)
+tar_load(tbl_one)
+tar_load(bamlasso_tune)
+tar_load(bai_cv_res)
 ```

### 2021-09-03 Manuscript/01-Intro.Rmd
Run pipeline

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smooth functions, $f_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smooth functions, $B_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
- E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
+ E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
 $$
@@
 Bai et al.
-[@Bai2020] was the first to apply group spike-and-slab lasso prior to AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
+[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian additive models(AMs) using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
 As these methods focus on addressing the sparsity, it is very likely to overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
@@
 To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothness penalties in the model via re-parameterization of the smoothing function to avoid over-penalizing basis function coefficients.
-Smoothness penalties are commonly implemented in GAMs through smoothing and penalized regression splines.
+Specifically, we incorporate smoothness penalties in the model via re-parameterization of the smoothing function to avoid overly shrinking basis function coefficients.
+Smoothness penalties are commonly implemented in GAMs through smoothing regression splines.
 They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
@@
 
-The proposed BHAM uses a spike-and-slab prior for function selections, but it substantially differs from previous spike-and-slab based GAMs, i.e.
-the spike-and-slab GAM [@Scheipl2012] and the SB GAM [@Bai2021] in many ways.
+The proposed framework, BHAM differs from previous spike-and-slab based GAMs, i.e.
+the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in many ways.
 First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM.
-Spike-and-slab lasso priors provide computational convenience during model fitting through optimization algorithms versus computational intensive sampling algorithms.
+Spike-and-slab lasso priors provide computational convenience during model fitting by using optimization algorithms instead of computational intensive sampling algorithms.
 They make fitting high- and ultrahigh-dimensional models more feasible without sacrificing performance in prediction and variable selection.

### 2021-09-03 Manuscript/02-Method.Rmd
Run pipeline

@@
 \begin{equation}\label{eq:gam}
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
 \end{equation}
-with the smoothing function $f_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
-the design matrix derived from the smoothing function $f_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
+with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
+the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
 The dimension of the design matrix depends on the choice of the smoothing function.
@@
 To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
-$f_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+$B_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
 
@@
 For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter.
-Similarly, the spline function can be expressed in the matrix form
+Similarly, a spline function can be expressed in the matrix form
 $$
@@
 $$
-and the generalized additive model in Equation \ref{eq:gam} be
+and the generalized additive model in Equation (\ref{eq:gam}) be
 \begin{equation}\label{eq:gam-repa}
@@
 \matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
+  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j) $ };\\
   \node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j^\tp)$ };\\
+  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j)$ };\\
   \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\

### 2021-09-03 Manuscript/04-Real_Data.Rmd
Run pipeline

@@
 
-bam_mdl_res <- bamlasso_tune %>% 
-  arrange(desc(auc)) %>% 
-  slice(1) %>% 
-  select(-s0)
-
 rbind(
-  'BHAM-SSL' = bam_mdl_res,
-  'SB-GAM' = bai_cv_res %>% t %>%  data.frame()
+  'BHAM-SSL' = bamlasso_tune %>% slice_min(mse) %>% select(-s0),
+  'SB-GAM' = bai_cv_res %>%  slice_min(mse) %>% select(-lambda)
 ) %>% 
+  rownames_to_column("Methods") %>% 
   flextable() %>% 

### 2021-09-06 R/summary_cv_SBGAM.R
Add new cv per-fold function

@@ -6,3 +6,27 @@ summary_cv_SBGAM <- function(cv.SBGAM_output, family = "gaussian"){
            }) %>% 
      data.frame(lambda = cv.SBGAM_output$lambda0,.)
 }
+
+
+summary_cv_SBGAM_per_fold <- function(cv.SBGAM_output, family = "gaussian"){
+  browser()
+  ret <- map_dfr(1:length(cv.SBGAM_output$lambda0),
+          .f = function(i){
+            map_dfr(unique(cv.SBGAM_output$folds), .f = function(j){
+              idx <- which(cv.SBGAM_output$folds==j)
+              .y <- cv.SBGAM_output$y.new[idx]
+              pred.y <- cv.SBGAM_output$prevalid.mu[idx,i]
+              measure.glm(.y, pred.y, family = family) %>% 
+                t %>% data.frame()
+            }, .id = "folds")
+          }, .id = "lambda_idx") %>% 
+    mutate(
+      lambda0 = cv.SBGAM_output$lambda0[lambda_idx %>% as.numeric] %>% factor
+    )
+  # ggplot(data = ret) +
+  #   geom_violin(aes(x = lambda0, y = auc))
+  return(ret)
+  
+  # %>% 
+  #   data.frame(lambda = cv.SBGAM_output$lambda0,.)
+}

### 2021-09-06 Manuscript/00-main.Rmd
Manuscript revision

@@
   org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "Nengjun Yi, epartment of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{nyi@uab.edu}"
+corres: "Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{nyi@uab.edu}"
 presentaddress: This is sample for present address text this is sample for present address text
@@
 accepted: 2021-03-01
-abstract: "Spike-and-slab priors have been extensively studied for variable selection, and recently been extended to function selection in generalized additive models (GAMs).
-Most of the previous works on fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC) algorithms to identify the promising subsets of functions.
-The computational burden of MCMC makes these models less feasible, especially in high-dimensional setting.
-Besides, most of the current spike-and-slab GAM consider an “all-in-all-out” approach for function selection, rendering less flexibility to answer if non-linear effects are necessary in the model.
-We propose Bayesian hierarchical generalized additive models for function selection and prediction.
-The proposed models employ spike-and-slab priors, mixture normal and mixture double exponential, for smooth and sparse solution.
-Two EM based algorithms, EM-IWLS and EM-Coordinate Descent, are developed for accelerated computation.
-Simulation studies and real data analysis demonstrated improved performance against the state-of-art models, mgcv and COSSO.
-The software implementation of the proposed methods is freely available via the R package BHAM."
+abstract: "There are many proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional (p>>n) problem using group penalties and priors.
+These models rarely smooth the nonlinear effect properly because of the sparse regularization.
+It leads to excess shrinkage on coefficients and inaccurate prediction.
+Besides, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
+While some Bayesian models can address the shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution.
+For smooth and sparse solutions, the proposed models employ bi-part spike-and-slab spline priors, including mixture double exponential and mixture normal priors.
+Two optimization-based algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for expedited computation.
+Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-art models, mgcv, COSSO and sparse Bayesian GAM.
+The software implementation of the proposed models is freely available via an R package BHAM."
 keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
@@
 library(targets)
-library(here)
+# library(here)
 library(xtable)
@@
 
-tar_load(ful_dat)
-tar_load(tbl_one)
-tar_load(bamlasso_tune)
-tar_load(bai_cv_res)
+tar_load(WLM_train_dat)
+# tar_load(tbl_one)
+tar_load(ECB_bamlasso_cv)
+tar_load(ECB_SBGAM_cv)
+tar_load(WLM_bamlasso_cv)
+tar_load(WLM_SBGAM_cv)
 ```
 
-```{r child=here(intro_path)}
+```{r child="01-Intro.Rmd"}
 ```
 
-```{r child=here(method_path)}
+```{r child="02-Method.Rmd"}
 ```
 
-```{r child=here(sim_path)}
+```{r child="03-Simulation.Rmd"}
 ```
 
-```{r child=here(real_data_path)}
+```{r child="04-Real_Data.Rmd"}
 ```
 
-```{r child=here(disc_path)}
+```{r child="05-Discussion.Rmd"}
 ```
-
-```{r plot-ref, fig.cap = "Fancy Caption\\label{fig:plot}", eval = F}
-x = rnorm(10)
-y = rnorm(10)
-plot(x, y)
-```
-
-<!-- ...and can be referenced (Figure \ref{fig:plot}) by including the `\\label{}` tag in the `fig.cap` attribute of the R chunk: `fig.cap = "Fancy Caption\\label{fig:plot}"`.
-It is a quirky hack at the moment, see [here](https://github.com/yihui/knitr/issues/323).
--->
-
-<!-- Analogously, use Rmarkdown to produce tables as usual: -->
-
-```{r, results = "asis", eval = F}
-if (!require("xtable")) install.packages("xtable")
-xt <- xtable(head(cars), caption = "A table", label = "tab:table")
-print(xt, comment = FALSE)
-```

### 2021-09-06 Manuscript/01-Intro.Rmd
Manuscript revision

@@
 These models are built upon a restrictive and unrealistic assumption, linearity.
-In classical statistical modelling, many modeling strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
+In classical statistical modelling, many strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
 For example, variable categorization is a simple and common practice in epidemiology, but suffers from power and interpretation issues.
@@
 
-Semi-parametric regression models are an appropriate alternative to the "black-box" models thanks to their balance between model flexibility and interpretability.
+Nonparametric regression models are appropriate alternatives to the "black-box" models thanks to their balance between model flexibility and interpretability.
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smooth functions, $B_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smoothing functions, $B_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
@@
 $$
-where $g^{-1}$ is the inverse function of a monotonic link function.
-The smooth functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
+where $g^{-1}$ is the inverse of a monotonic link function.
+The smoothing functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
 Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
 
-There exists some proposals to generalized the classical GAM to accommodate high-dimensional applications.
-The regularized models that are primarily designed for linear models are used to fit GAMs by accounting for the group structure introduced when expanding smoothing functions.
+There exists some proposals to generalize the classical GAM to accommodate high-dimensional applications.
+The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smoothing functions.
 <!-- These models assume sparsity among the predictors, i.e.
@@
 Ravikumar et al.
-[@Ravikumar2009] extended the grouped lasso [@Yuan2006] to additive models; Huang et al.
+[@Ravikumar2009] extended the grouped lasso [@Yuan2006] to additive models (AMs); Huang et al.
 [@Huang2010] further developed adaptive grouped lasso for additive models; Wang et al.
-[@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to the additive models.
+[@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
 Recent Bayesian hierarchical models are also used in the context of high-dimensional additive models.
@@
 Bai et al.
-[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian additive models(AMs) using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
+[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
 As these methods focus on addressing the sparsity, it is very likely to overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
-[@Scheipl2013] In addition, the proposed methods adapt an 'all-in-all-out' strategy, i.e.
+[@Scheipl2013] In addition, these methods adapt an 'all-in-all-out' strategy, i.e.
 either including or excluding the variable completely, rendering no space for bi-level selection.
@@
 But the model fitting relies on computational intensive MCMC algorithmic and creates scalability concern.
- <!-- specifically for semi-parametric regression using MCMC.
-The normal-mixture-of-inverse gamma prior is mixture of t-distribution with an additional coefficients mixing vector, where the vector follows a mixture of two Gaussian distribution.
--->
-It would be of special interest to develop a fast, flexible and accurate generalized additive model.
+It would be of special interest to develop a fast, flexible and accurate generalized additive model framework.
 
 To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothness penalties in the model via re-parameterization of the smoothing function to avoid overly shrinking basis function coefficients.
-Smoothness penalties are commonly implemented in GAMs through smoothing regression splines.
+Specifically, we incorporate smoothing penalties in the model via re-parameterization of the smoothing function to avoid overly shrinking basis function coefficients.
+Smoothing penalties are commonly implemented in the classical GAMs through smoothing regression splines.
 They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
-The smoothness penalty conditioning on a smoothness penalty parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
+A smoothing penalty conditioning on a smoothing parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
 \begin{equation}\label{eq:smoothpen}
@@
 \end{equation}
-where $\bs S_j$ is a known smoothness penalty matrix, and $\bs \beta_j$ is the basis function coefficients.
-Smoothness penalties are also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-Moreover, incorporating the smoothness penalty allows the separation of the linear space of smooth functions, also know as the null space, from the nonlinear space.
-We then impose a new bi-part spike-and-slab spline prior on the smooth functions for bi-level selection such that the linear and nonlinear spaces of smooth functions can be selected separately.
+where $\bs S_j$ is a known smoothing penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
+Smoothing penalties re also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
+Moreover, incorporating the smoothing penalty allows the separation of the linear space of smoothing functions, also know as the null space, from the nonlinear space.
+We then impose a new bi-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
 The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
@@
 
-The rest of the paper are arranged as the followings: in Chapter 2 we demonstrate the Bayesian hierarchical generalized additive model, where we provide the setup of the proposed model and describe the two fast-fitting algorithms.
-In Chapter 3, we compare the proposed via monte carlo simulation.
-.Application to real-world datasets demonstrates the usefulness of the proposed methodologies in Chpater 4.
-Conclusion and discussions are made in Chapter 5.
-^[TODO: rewrite this sentence.]
+In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the two fast-fitting algorithms.
+In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
+Analyses of two metabolomics datasets using the proposed model are presented in Section \ref{sec:real_data}.
+Conclusion and discussions are given in Section \ref{sec:concl}.
 
@@
 In common practice, model matrix of spline terms are generated using base functions of preference, and then feed to a penalized model.
-Without proper handle of the grouping nature of bases functions, the estimation of the smooth function is less accurate.
+Without proper handle of the grouping nature of bases functions, the estimation of the smoothing function is less accurate.
 @Bai2020Spline shows inferior performance of such practice compared to using an ad hoc high-dimensional additive model.
@@
 Most commonly used spline is cubic spline, thanks to its smoothness property.
-[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smooth function.
+[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smoothing function.
 However, the specification of knots is not a trivial task, as the the positioning of knots are influential on the performance of the model.

### 2021-09-06 Manuscript/02-Method.Rmd
Manuscript revision

@@
 First of all, as previously mentioned, the re-parameterization integrated the smoothness penalty into the design matrix, and hence the penalty can be expressed in the posterior density function implicitly in addition to the sparse penalty for variable and function selection.
-Secondly, the eigendecomposition of the smooth penalty isolates the linear space of a smooth function from the non-linear space, improving the feasibility of bi-level variable/function selection and estimation inference.
+Secondly, the eigendecomposition of the smooth penalty isolates the linear space of a smooth function from the nonlinear space, improving the feasibility of bi-level variable/function selection and estimation inference.
 The eigendecomposition also introduced the linearly independent column spaces of the smooth function matrix, which allows imposing independent priors on the coefficients.
@@
 
-where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the non-linear effect of the $j$th variable respectively.
+where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
 ; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
@@
 It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection.
-Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and non-linear effect, i.e.
+Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
 $\gamma_j = \gamma^0_j = \gamma^\tp_j$.
@@
 
-<!-- The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the non-linear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
+<!-- The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the nonlinear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
 One of the exception is Bai (in progress), who a grouped spike-and-slab LASSO prior [@Rockova2018] to address the grouping structure of spline bases.

### 2021-09-06 Manuscript/03-Simulation.Rmd
Manuscript revision

@@
 # Simulation Study
+\label{sec:sim}
 
@@
 Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component are different for extra flexibility.
-Generalized additive models with automatic smoothing, hereafter \textit{mgcv}, is one of the most popular models for non-linear effect interpolation and prediction.
+Generalized additive models with automatic smoothing, hereafter \textit{mgcv}, is one of the most popular models for nonlinear effect interpolation and prediction.
 SB-GAM is the first spike-and-slab lasso GAM.

### 2021-09-06 Manuscript/04-Real_Data.Rmd
Manuscript revision

--- 2021-08-31 Manuscript/04-Real_Data.Rmd
@@
 # Metabolomics Data Analysis
+\label{sec:real_data}
 
@@
 ```{r}
-
-bam_mdl_res <- bamlasso_tune %>% 
-  arrange(desc(auc)) %>% 
-  slice(1) %>% 
-  select(-s0)
-
 rbind(
-  'BHAM-SSL' = bam_mdl_res,
-  'SB-GAM' = bai_cv_res %>% t %>%  data.frame()
+  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
 ) %>% 
+  rownames_to_column("Methods") %>% 
   flextable() %>% 
@@
 ## Weight Loss Maintenance Cohort 
-We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study ^[TODO: add citation:https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0240764] on the association between metabolic biomarkers and weight loss.
-The data contained 765 metaboliltes in baseline plasma collected from three separate studies, profiled using liquid chromatography mass spectrometry.
+We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
+In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
+In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
 Quality control and natural log transformation were performed during metabolites data preparation.
-The outcome of interest was percent change in insulin resistance, and hence modeled using a Gaussian model adjusting for covariates including age, sex, race, baseline clinical triglycerides, and percent change in weight over the follow-up time period.
-731 individuals were included in the initial studies, and reduced to `r nrow(ful_dat)` after removing missing values.
-The descriptive statistics were tabulated in Table \ref{tab:tbl_one}.
+The outcome of interest was standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
+After removing missing datapoints and addressing outliers in the data, there were `r nrow(WLM_train_dat)` remaining in the analysis.
+5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
+10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
 
 ```{r}
-tbl_one %>% 
-  as_flex_table() %>% 
+rbind(
+  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
+) %>% 
+  rownames_to_column("Methods") %>% 
+  flextable() %>% 
   flextable::autofit() %>% 
-  flextable::set_caption(caption = "The descriptive statistics of the sample in the metabolomics data analysis stratified by studies\\label{tab:tbl_one}")
+  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
 ```

### 2021-09-06 Manuscript/05-Discussion.Rmd
Manuscript revision

@@
 # Discussion
+\label{sec:concl}
 In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.

### 2021-09-06 Manuscript/SS_GAM.tex
Manuscript revision

--- 2021-08-02 Manuscript/SS_GAM.tex
@@
 
-\corres{*Corresponding author name, This is sample corresponding
-address. \email{xxx@uab}}
+\corres{Nengjun Yi, Department of Biostatistics, University of Alabama
+at Birmingham, Birmingham, USA. \email{nyi@uab.edu}}
 
@@
 
-\abstract{Spike-and-slab priors have been extensively studied for
-variable selection, and recently been extended to function selection in
-generalized additive models (GAMs). Most of the previous works on
-fitting spike-and-slab GAM relies on Markov chain Monte Carlo (MCMC)
-algorithms to identify the promising subsets of functions. The
-computational burden of MCMC makes these models less feasible,
-especially in high-dimensional setting. Besides, most of the current
-spike-and-slab GAM consider an ``all-in-all-out'' approach for function
-selection, rendering less flexibility to answer if non-linear effects
-are necessary in the model. We propose Bayesian hierarchical generalized
-additive models for function selection and prediction. The proposed
-models employ spike-and-slab priors, mixture normal and mixture double
-exponential, for smooth and sparse solution. Two EM based algorithms,
-EM-IWLS and EM-Coordinate Descent, are developed for accelerated
-computation. Simulation studies and real data analysis demonstrated
-improved performance against the state-of-art models, mgcv and COSSO.
-The software implementation of the proposed methods is freely available
-via the R package BHAM.}
+\abstract{There are many proposals that extend the classical generalized
+additive models (GAMs) to accommodate the high-dimensional
+(p\textgreater\textgreater n) problem using group penalties and priors.
+These models rarely smooth the nonlinear effect properly because of the
+sparse regularization. It leads to excess shrinkage on coefficients and
+inaccurate prediction. Besides, most of these GAMs consider an
+``all-in-all-out'' approach for function selection, rendering them
+difficult to answer if nonlinear effects are necessary. While some
+Bayesian models can address the shortcomings, using Markov chain Monte
+Carlo (MCMC) algorithms for model fitting creates a new challenge,
+scalability. Hence, we propose Bayesian hierarchical generalized
+additive models as a solution. For smooth and sparse solutions, the
+proposed models employ bi-part spike-and-slab spline priors, including
+mixture double exponential and mixture normal priors. Two
+optimization-based algorithms, EM-Coordinate Descent and EM-Iterative
+Weighted Least Squares, are developed for expedited computation.
+Simulation studies and metabolomics data analyses demonstrate improved
+performance against state-of-art models, mgcv, COSSO and sparse Bayesian
+GAM. The software implementation of the proposed models is freely
+available via an R package BHAM.}
 
@@
 \label{sec:intro}
-
-Using additive models to interpolate nonlinear effects has a long
-history. Early work of additive models can date back to 1979 when
-Cleveland \citep{Cleveland1979} proposed local smoothing for bivariate
-analysis. Since then, many mathematical and statistical models have been
-proposed. \citep{Wood2020} With the recent advancement of computational
-technology, many machine learning algorithms that are computationally
-intensive start to gain much popularity in prognostic modeling and
-functional selection, especially when complex signals, for example
-nonlinear effects, are assumed. Such models generally have improved
-prediction accuracy over classic statistical models. However, these
-models have a fatal flaw, the lack of transparency and interpretability,
-and hence are referred to as ``black box'' models \citep{Breiman2001}.
-Given that the primary objective of biomedical studies is to understand
-the etiology of diseases, ``Black box'' models are less helpful as they
-can hardly explain how the change of biomarkers is associated with the
-disease.
-
-Semi-parametric regression models are a more appropriate alternative to
-the ``black-box'' models, thanks to their balancing between model
-flexibility and interpretability. Among those, generalized additive
-models (GAMs), proposed in the seminal work of Hastie and Tibshirani
-\citep{Hastie1987}, grew to be one of the most popular modeling tools.
-In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution \(EF(\mu, \phi)\) with mean \(\mu\) and
-dispersion \(\phi\), can be expressed as the summation of smooth
-functions, \(f_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector
-of covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)),
-\] where \(g^{-1}\) is the inverse function of a smooth monotonic link
-function. The smooth functions can take many forms and are estimated
-using a pseudo-weighted version of the backfitting algorithm
-\citep{Breiman1985}. Nevertheless, the classic GAMs cannot fulfill the
-increasing analytic demands for high-dimensional data analysis in
-biomedical studies.
 
@@
 models or Bayesian hierarchical models, see Mallick and Yi
-\citep{Mallick2013} for review. These models assume sparsity among the
-predictors, i.e.~most predictors have small or no effects on the
-outcome. Further development focuses on addressing the grouping
-structure among the predictors, where AMs and GAMs can be considered
-special cases. Various penalties are imposed on the basis function
-coefficients: Ravikumar et al. \citep{Ravikumar2009} extended the
-grouped lasso \citep{Yuan2006} to additive models; Huang et al.
+\citep{Mallick2013} for review. These models are built upon a
+restrictive and unrealistic assumption, linearity. In classical
+statistical modelling, many strategies and models are proposed to relax
+the linearity assumption with various degrees of complexity. For
+example, variable categorization is a simple and common practice in
+epidemiology, but suffers from power and interpretation issues. More
+complex models to address nonlinear effects include random forest and
+other so-called ``black box'' models \citep{Breiman2001}. These models
+are not particularly attractive in biomedical studies, as they lack the
+transparency to understand the etiology of diseases. In addition, how to
+generalized these models to the high-dimensional setting remains
+unclear.
+
+Nonparametric regression models are appropriate alternatives to the
+``black-box'' models thanks to their balance between model flexibility
+and interpretability. Among those, generalized additive models (GAMs),
+proposed in the seminal work of Hastie and Tibshirani
+\citep{Hastie1987}, grew to be one of the most popular modeling tools.
+In a GAM, the response variable, \(Y\), which is assumed to follow some
+exponential family distribution \(EF(\mu, \phi)\) with mean \(\mu\) and
+dispersion \(\phi\), can be expressed as the summation of smoothing
+functions, \(B_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector
+of covariates \(\boldsymbol{x}\), written as \[
+ E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
+\] where \(g^{-1}\) is the inverse of a monotonic link function. The
+smoothing functions can take many forms and are estimated using a
+pseudo-weighted version of the backfitting algorithm
+\citep{Breiman1985}. Nevertheless, the classical GAMs cannot fulfill the
+increasing analytic demands for high-dimensional data analysis in
+biomedical studies.
+
+There exists some proposals to generalize the classical GAM to
+accommodate high-dimensional applications. The regularized models,
+branching out from regularized linear models, are used to fit GAMs by
+accounting for the group structure introduced when expanding smoothing
+functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
+lasso \citep{Yuan2006} to additive models; Huang et al.
 \citep{Huang2010} further developed adaptive grouped lasso for additive
@@
 chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Scheipl et al.
-\citep{Scheipl2012} proposed an ad hoc spike-and-slab structure prior,
-normal-mixture-of-inverse gamma, specifically for semi-parametric
-regression using MCMC. The normal-mixture-of-inverse gamma prior is
-mixture of t-distribution with an additional coefficients mixing vector,
-where the vector follows a mixture of two Gaussian distribution. Bai et
-al. \citep{Bai2020} was the first to apply group spike-and-slab lasso
-prior to AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs\citep{Bai2021}. Regardless of the nice
-statistical properties, a practical and pivotal question, if non-linear
-effects exist, yet to be addressed by these methods. Besides Scheipl et
-al. \citep{Scheipl2012}, most proposed methods adapt an `all-in-all-out'
-strategy, i.e.~either including or excluding the variable completely,
-rendering no space for within group selection. In addition, a common
-criticism received by these methods is that the basis function
-coefficients tend to be overly penalized due to the sparsity assumption
-directly applied on the basis function coefficients.\citep{Scheipl2013}
-It would be of special interest to develop a fast yet flexible
-generalized additive model.
+proposed, where the application on AMs are by-products. Bai et al.
+\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
+to Gaussian additive models(AMs) using a fast optimization algorithm,
+and further generalized the framework to GAMs\citep{Bai2021}. As these
+methods focus on addressing the sparsity, it is very likely to overly
+penalize the basis function coefficients and produce inaccurate
+predictions and curve interpolation, particularly when complex signals
+are assumed and large number of knots are used. \citep{Scheipl2013} In
+addition, the proposed methods adapt an `all-in-all-out' strategy,
+i.e.~either including or excluding the variable completely, rendering no
+space for bi-level selection. Scheipl et al. \citep{Scheipl2012}
+proposed an ad hoc spike-and-slab structure prior,
+normal-mixture-of-inverse gamma prior, that address the previous
+challenges. But the model fitting relies on computational intensive MCMC
+algorithmic and creates scalability concern. It would be of special
+interest to develop a fast, flexible and accurate generalized additive
+model.
 
@@
 generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we impose a new bi-part spike-and-slab spline prior on the
-smooth functions to perform bilevel selection, where the linear and
-nonlinear spaces of smooth functions can be selected separately. The
-prior setup encourages a flexible solution, rending one of three
-possibilities: no effect, only linear effect, or nonlinear effect of a
-predictor. Additionally, two fast computing optimization-based
-algorithms are developed for high and
-ultrahigh-dimensional\citep{Fan2009} settings respectively. Meanwhile,
-to avoid the over-penalization of the basis function coefficients,
-smoothness penalties are incorporated in the model via
-re-parameterization. Smoothness penalties are commonly implemented in
-GAMs through smoothing and penalized regression splines. They are
+Specifically, we incorporate smoothness penalties in the model via
+re-parameterization of the smoothing function to avoid overly shrinking
+basis function coefficients. Smoothness penalties are commonly
+implemented in GAMs through smoothing regression splines. They are
 quadratic norms of the coefficients and allow locally adaptive penalties
-on each smoothing function. For the most common cases of cubic splines,
-the smoothness penalty conditioning on a smoothness penalty parameter
-\(\lambda_j\) is a function of the integration of the second derivative
-of the spline function, expressed mathematically as
-\begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{"}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
+on each smoothing function. The smoothness penalty conditioning on a
+smoothness penalty parameter \(\lambda_j\) is a function of the
+integration of the second derivative of the spline function, expressed
+mathematically as \begin{equation}\label{eq:smoothpen}
+  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
 \end{equation} where \(\boldsymbol{S}_j\) is a known smoothness penalty
@@
 \citep{Meier2009}. Moreover, incorporating the smoothness penalty allows
-the separation of the linear space of the smooth function, also know as
-the null space, from the nonlinear space. Marra and Wood
-\citep{Marra2011} introduced a re-parameterization of the basis function
-matrix based on the decomposition of the smoothness penalty matrix
-\(\boldsymbol{S}\). The re-parameterization allows different penalties
-imposed on the null and nonlinear spaces separately, and further allows
-the sparsity in linear space and smoothness in nonlinear effect
-interpolation. In addition, the priors of the linear and nonlinear
-effects are independent, creating computational convenience and further
-improving the scalability of the model.
-
-The proposed BHAM uses a spike-and-slab prior for function selections,
-but it substantially differs from previous spike-and-slab based GAMs,
-i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB GAM
+the separation of the linear space of smoothing functions, also know as
+the null space, from the nonlinear space. We then impose a new bi-part
+spike-and-slab spline prior on the smoothing functions for bi-level
+selection such that the linear and nonlinear spaces of smoothing
+functions can be selected separately. The prior setup encourages a
+flexible solution, rending one of three possibilities: no effect, only
+linear effect, or nonlinear effect of a predictor. In addition, two fast
+computing optimization-based algorithms are developed for high and
+ultrahigh-dimensional\citep{Fan2009} settings respectively.
+
+The proposed framework, BHAM differs from previous spike-and-slab based
+GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
 \citep{Bai2021} in many ways. First of all, the proposed spike-and-slab
@@
 Spike-and-slab lasso priors provide computational convenience during
-model fitting through optimization algorithms versus computational
+model fitting by using optimization algorithms instead of computational
 intensive sampling algorithms. They make fitting high- and
@@
 high dimensional setting. Furthermore, the proposed model addresses the
-incapability of bilevel selection in SB-GAM. Lastly, the R
+incapability of bi-level selection in SB-GAM. Lastly, the R
 implementation of the Bayesian hierarchical additive model is available
@@
 
-The rest of the paper are arranged as the followings: in Chapter 2 we
-demonstrate the Bayesian hierarchical generalized additive model, where
-we provide the setup of the proposed model and describe the two
-fast-fitting algorithms. In Chapter 3, we compare the proposed via monte
-carlo simulation. .Application to real-world datasets demonstrates the
-usefulness of the proposed methodologies in Chpater 4. Conclusion and
-discussions are made in Chapter 5. \footnote{TODO: rewrite this
-  sentence.}
+In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical
+generalized additive model, introduce the proposed spike-and-slab spline
+priors, and describe the two fast-fitting algorithms. In Section
+\ref{sec:sim}, we compare the proposed framework to state-of-art models,
+mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
+Analyses of two metabolomics datasets using the proposed model are
+presented in Section \ref{sec:real_data}. Conclusion and discussions are
+given in Section \ref{sec:concl}.
 
@@
 \begin{equation}\label{eq:gam}
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}f_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
-\end{equation} with the smoothing function \(f_j(x_j)\) of the variable
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
 \(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
-matrix derived from the smoothing function \(f_j(x_j)\), is denoted
+matrix derived from the smoothing function \(B_j(x_j)\), is denoted
 \(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
@@
 treating the parametric function as a special case of the smoothing
-function, e.g.~\(f_j(x_j) = x_j\) with the smoothness penalty matrix
+function, e.g.~\(B_j(x_j) = x_j\) with the smoothness penalty matrix
 defined as \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
@@
 as the after re-parameterized basis function matrix for the \(j\)th
-variable hereafter. Similarly, the spline function can be expressed in
-the matrix form \[
+variable hereafter. Similarly, a spline function can be expressed in the
+matrix form \[
 f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
-\] and the generalized additive model in Equation \ref{eq:gam} be
+\] and the generalized additive model in Equation (\ref{eq:gam}) be
 \begin{equation}\label{eq:gam-repa}
@@
 smooth penalty isolates the linear space of a smooth function from the
-non-linear space, improving the feasibility of bi-level
-variable/function selection and estimation inference. The
-eigendecomposition also introduced the linearly independent column
-spaces of the smooth function matrix, which allows imposing independent
-priors on the coefficients. This reduces the computational complexity
-that multivariate priors induces, and greatly broadens the choices of
-priors and further model choices. Last but not least, the scaling step
-of \(\boldsymbol{X}^\text{pen}\) further simplifies the choice of
-penalty priors, from column-specific penalties to a unified penalty.
-This step is also necessary for the proposed spike-and-slab spline
-prior. Conventionally, the spike-and-slab priors require scaling of
-predictors \footnote{TODO: insert a citation.}
+nonlinear space, improving the feasibility of bi-level variable/function
+selection and estimation inference. The eigendecomposition also
+introduced the linearly independent column spaces of the smooth function
+matrix, which allows imposing independent priors on the coefficients.
+This reduces the computational complexity that multivariate priors
+induces, and greatly broadens the choices of priors and further model
+choices. Last but not least, the scaling step of
+\(\boldsymbol{X}^\text{pen}\) further simplifies the choice of penalty
+priors, from column-specific penalties to a unified penalty. This step
+is also necessary for the proposed spike-and-slab spline prior.
+Conventionally, the spike-and-slab priors require scaling of predictors
+\footnote{TODO: insert a citation.}
 
@@
 variables, indicating if the model includes the linear effect and the
-non-linear effect of the \(j\)th variable respectively. ; \(s_0\) and
+nonlinear effect of the \(j\)th variable respectively. ; \(s_0\) and
 \(s_1\) are scale parameters, assuming \(0 < s_0 < s_1\) and given.
@@
 restrictive assumption on the priors, assuming that one indicator
-variable decides the inclusion of both the linear effect and non-linear
+variable decides the inclusion of both the linear effect and nonlinear
 effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^\text{pen}_j\). This
@@
 \matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-&\node (linearTheta) [outCome] { $\theta_j^{0}\sim Beta(a, b)$ };\\
-&\node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j^0) $ };\\
-&\node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
 };
-
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-\node (penTheta) [outCome] { $\theta_{j}^\text{pen}\sim Beta(a, b)$ };\\
-\node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j^\text{pen})$ };\\
+  \node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j)$ };\\
   \node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
@@
 
+\label{sec:sim}
+
 In this section, we compare the performance of the proposed models to
@@
 flexibility. Generalized additive models with automatic smoothing,
-hereafter \textit{mgcv}, is one of the most popular models for
-non-linear effect interpolation and prediction. SB-GAM is the first
-spike-and-slab lasso GAM. We implemented COSSO and adaptive COSSO with R
-package \texttt{cosso 2.1-1}, generalized additive models with automatic
+hereafter \textit{mgcv}, is one of the most popular models for nonlinear
+effect interpolation and prediction. SB-GAM is the first spike-and-slab
+lasso GAM. We implemented COSSO and adaptive COSSO with R package
+\texttt{cosso 2.1-1}, generalized additive models with automatic
 smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package
@@
 
-To evaluate the predictive performance of the models, \(R^2\) for
-Gaussian model, area under the curve (AUC) are calculated for binomial
-model, averaging across 50 simulations. The results are tabulated in
-Table \ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to
-  provide a box plot or a table.}
-
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-\includegraphics{SS_GAM_files/figure-latex/unnamed-chunk-11-1.pdf}
-
-For the binomial outcome shown in Figure 1, we can see that the proposed
-models are consistently performed better than COSSO and adaptive COSSO.
-Mgcv slightly outperforms the proposed models in low dimension
-settings(p = 4, 10), but suffers from overfitting in medium dimension
-setting (p=50). In addition, in the high-dimension setting (p=200), mgcv
-could not make prediction due to the p \textgreater{} n problem. Among
-the proposed models, we see that there is modest difference between the
-models that shares the same indicator variable('\_spline' models) and or
-not sharing indicator variable (modles), where the not sharing indicator
-variable models have slightly edge over indicator sharing model.
-Moreover, the mixture double exponential model and lasso model perform
-substaintially better in the medium and high settings. For the remaining
-two outcomes, the results shows similar trends, where mixture double
-exponential model and ss lasso model are recommended for use.
-
-\%are calculated to assess out-of-sample estimation accuracy,
-
-\%We also calculated the area under the curve (AUC) of the receiver
-operating characteristic curve for \(y_{new}\). We will also visually
-display the estimated smooth function with the smooth function of truth
-to compare if the models can infer the smooth function correctly,
-despite the prediction accuracy.
-
-\%When constructing the smooth function of each variables, we constructs
-the natural cubic spline with degree of freedom 25. In other words, the
-design matrix of each smooth function will have 26 columns. The total
-number of columns for \(p=4, 10, 50\) is \(105, 261, 1301\)
-respectively, where intercept is kept in the model when fitting. The
-\(p=50\) can mimic the high dimensional setting as the design matrix are
-in high-dimensional setting (\(p >> n\))
-
-\%We benchmark the performance of the proposed methods with the state of
-art GAM package \texttt{mgcv}{[}TODO: add citation{]}. In the future
-steps, we will compare our proposed model with other high dimensional
-GAM applications.
+To evaluate the predictive performance of the models, the statistics
+caluclated based on the testing dataset, \(R^2\) for Gaussian model,
+area under the curve (AUC) for binomial model, are calculated averaging
+across 50 simulations. The results are tabulated in Table
+\ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to provide a
+  box plot or a table.}
+
+The predictive performances have a consistent pattern across the three
+distributions of outcomes. Across all the scenarios, COSSO and adaptive
+COSSO have the least favourable performance among the applicable methods
+examined. To note, mgcv doesn't support high-dimensional analysis when
+the number of coefficients are greater than the sample size. mgcv
+predicts well when \(p\) is small or moderate (p = 4, 10), and
+deteriorate when the number of predictors increase. Among the three
+fast-computing Bayesian hierarchical models, the proposed models,
+bglm\_spline\_de and blasso\_spline \footnote{need to change the name}
+predicts better than SB-GAM when the dimension of are small, moderate,
+and high. However, in the hyper high dimensional case, SB-GAM has better
+performance. Among the proposed methods, bglm\_spline\_de performs
+consistently better than blasso\_spline.
 
@@
 
-Overall, the proposed priors outperforms the state-of-art GAM model
-\texttt{mgcv} when the dimensionality is low, median, and high. The
-proposed priors predicts better than \texttt{mgcv} and infers better of
-the underlying smooth function. Most importantly, \texttt{mgcv} is
-infeasible in the high dimensional setting, i.e.~when \(p=50\) where the
-number of parameters are greater than the sample size.
-
-\hypertarget{prediction-performace}{%
-\paragraph{Prediction Performace}\label{prediction-performace}}
-
-The prediction performance is tabulated in the table below for AUC and
-Supplementary tables for MSE and misclassification rate. In summary, the
-proposed Bayesian models outperforms the benchmark model \texttt{mgcv}
-regardless the choice of prior or fitting algorithm. The models with
-IRLS fitting algorithm give slightly better performance than the the
-model with cyclic coordinate descent algorithm. The advantage diminishes
-in the high dimensional setting (p=50). This is possible because the SS
-normal spline prior provides a more smooth solution, which would work
-well in a low to median dimension setting. However, in the high
-dimensional setting, there are not enough sparsity in the model, i.e.~SS
-normal prior will provides a more complex model than necessary. This is
-confirmed with the visualization in the following section.
-
-\hypertarget{visualization}{%
-\paragraph{Visualization}\label{visualization}}
-
-We plot the estimated smooth functions of the four active variables for
-the 100 simulation iterations, along with the `true' smooth function.
-Across different settings of predictors, the patterns look similar.
-First of all, in all the settings, \texttt{mgcv} tends to provide more
-extreme models, where in the graphs the smooth function have much higher
-bound than the true smooth function. Rarely the estimated smooth
-function from \texttt{mgcv} matches with the true smooth function. In
-comparison, the proposed method can interpolate the shape of the true
-smooth function. However, different priors and fitting algorithms have
-different advantages. For example, in Figure 1, we see that the SS
-normal distribution have problem with the tail of the smooth function,
-especially obvious in the estimations of the variable \emph{x4}.
-However, it would have less sparse solutions, which is more likely to
-accurately interpolate the smooth function especially when the shape is
-complicated (see variable \emph{x2}). On the contrary, the SS double
-exponential prior model encourages sparse solutions, which can sometime
-omit active predictors.
-
-\hypertarget{real-data-analysis}{%
-\section{Real Data Analysis}\label{real-data-analysis}}
+\hypertarget{metabolomics-data-analysis}{%
+\section{Metabolomics Data Analysis}\label{metabolomics-data-analysis}}
+
+\label{sec:real_data}
+
+In this section, we apply the proposed models BHAM-SSL to analyze two
+real-world metabolomics datasets where the outcomes are binary and
+continuous respectively. We demonstrate the improved prediction
+performance compared to the other Bayesian hierarchical additive model,
+SB-GAM \citep{Bai2021}.
+
+\hypertarget{emory-cardiovarcular-biobank}{%
+\subsection{Emory Cardiovarcular
+Biobank}\label{emory-cardiovarcular-biobank}}
+
+We use the proposed models BHAM-SSL to analyze a metabolic dataset from
+a recently published research \citep{Mehta2020} studying plasma
+metabolomic profile on the three-year all-cause mortality among patients
+undergoing cardiac catheterization. The dataset is publicly available
+via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
+subjects from two cohorts. As there is a large number of non-overlapping
+features among the two cohorts, we use the cohort with larger sample
+size (N=454). There were initially 6796 features in the dataset, which
+is too large to be practically meaningful to analyze. Hence, we
+performed a univariate screening procedure on the features, via GAM
+implemented in \texttt{mgcv}, and chose the the top 200 features with
+smallest p-values. 5-Knot spline additive models for binary outcome are
+constructed using two different models, the proposed BHAM-SSL and the
+SB-GAM. 10-Fold cross-validation(CV) are used to choose the optimal
+tuning parameters of each framework with respect to the default
+selection criterion implemented in the software. Out-of-bag samples were
+used for prediction performance evaluation, where deviance, area under
+the ROC curve (AUC), mean squared error (MSE) defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
+misclassication error defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) were
+calculated and tabulated in Table \ref{tab:ECB_res}.
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{1.07in}|p{0.96in}|p{0.67in}|p{0.67in}|p{0.67in}|p{1.39in}}
+
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\label{tab:ECB_res}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{455.686}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.740}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.159}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.285}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.214}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,230.063}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.706}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.208}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.257}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.240}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
+
+\end{longtable}
+
+\hypertarget{weight-loss-maintenance-cohort}{%
+\subsection{Weight Loss Maintenance
+Cohort}\label{weight-loss-maintenance-cohort}}
+
+We use the proposed models BHAM-SSL to analyze metabolomics data from a
+recently published study \citep{Bihlmeyer2021} on the association
+between metabolic biomarkers and weight loss, where the dataset is
+publically available \citep{Bihlmeyer2021_data}. In this analysis, we
+primarily focused on the analysis of one of the three studies included,
+weight loss maintenance (WLM) cohort \citep{Svetkey2008}, due to the
+drastically different intervention effects. In the dataset, 765
+metaboliltes in baseline plasma collected were profiled using liquid
+chromatography mass spectrometry. Quality control and natural log
+transformation were performed during metabolites data preparation. The
+outcome of interest was standardized percent change in insulin
+resistance, and hence modeled using a Gaussian model. After removing
+missing datapoints and addressing outliers in the data, there were 237
+remaining in the analysis. 5-Knot spline additive models for the
+gaussian outcome are constructed using two different models, the
+proposed BHAM-SSL and the SB-GAM. 10-Fold cross-validation(CV) are used
+to choose the optimal tuning parameters of each framework with respect
+to the default selection criterion implemented in the software.
+Out-of-bag samples were used for prediction performance evaluation,
+where deviance, mean squared error (MSE) defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
+abosolute error (MAE)defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) were calculated and
+tabulated in Table \ref{tab:WLM_res}.
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{1.07in}|p{0.90in}|p{0.67in}|p{0.67in}|p{0.67in}}
+
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\label{tab:WLM_res}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{669.375}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.077}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.924}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.745}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{666.826}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.026}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.976}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.766}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
+
+\end{longtable}
 
@@
 
-In the paper, we described a novel generalized additive model using
-Bayesian hierarchical priors, particularly the proposed spike-and-slab
-spline prior for bi-level functional selection. Meanwhile, we introduced
-two optimization based algorithms for model fitting. The algorithms can
-be easily scale up to address high-dimensional data analysis in a
-computational efficient manner, but also provide naive inferential
-supports for many commonly used hierarchical priors,
-e.g.~t-distribution, double exponential distribution, mixture t
-distribution, and mixture double exponential distribution, in the
-analyses that the sparse assumption is weak or not necessary. Via
-simulations, we demonstrated that the proposed model provides as good,
-if not better, prediction performance compared to some state-of-the-art
-non-linear modeling devices.
+\label{sec:concl} In the paper, we described a novel generalized
+additive model using Bayesian hierarchical priors, particularly the
+proposed spike-and-slab spline prior for bi-level functional selection.
+Meanwhile, we introduced two optimization based algorithms for model
+fitting. The algorithms can be easily scale up to address
+high-dimensional data analysis in a computational efficient manner, but
+also provide naive inferential supports for many commonly used
+hierarchical priors, e.g.~t-distribution, double exponential
+distribution, mixture t distribution, and mixture double exponential
+distribution, in the analyses that the sparse assumption is weak or not
+necessary. Via simulations, we demonstrated that the proposed model
+provides as good, if not better, prediction performance compared to some
+state-of-the-art non-linear modeling devices.
 

### 2021-09-06 _targets.R

@@ -93,14 +93,14 @@ tar_plan(
   
   ECB_SBGAM_cv = summary_cv_SBGAM(ECB_SBGAM_cv_raw, family = "binomial"),
   
-  # ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 32, a = 1, b = 1),
+  ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 24, a = 1, b = 1),
 
-  #   tar_target(ECB_SBGAM_insample_msr,
-  #            measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial")),
+    tar_target(ECB_SBGAM_insample_msr,
+             measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial")),
   # 
   # train_msr = data.frame("method" = c("BHAM-SSL", "SB-GAM"),
   #                        rbind(ECB_bamlasso_insample_msr ,ECB_SBGAM_insample_msr)),
-  
+
   
   #* Weight Loss Maintainace Study---------------------------------------------------------------
   
@@ -148,11 +148,12 @@ tar_plan(
                               group = WLM_bgam_dat$group),
   
   WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw,s0 = seq(0.005, 0.1, 0.01)),
-  # measure.bh(bgam_mdl),
+
   
-  # bglm_mdl_fnl = bamlasso(bgam_dat$train_dat, train_dat$out_HOMA_std, family = "gaussian",
-  #                         ss=c(0.005, 0.05),
-  #                         group = bgam_dat$group)
+  bglm_mdl_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
+                          ss=c(0.025, 0.05),
+                          group = WLM_bgam_dat$group),
+  # measure.bh(bglm_mdl_fnl),
   
   # ** SB_GAM ---------------------------------------------------------------
   
@@ -170,7 +171,7 @@ tar_plan(
     df=5, family = "gaussian", a = 1, b = 1,
     max.iter=100, tol = 1e-6, lambda0 = 58),
   
-  # measure.glm(ful_dat$out_HOMA_std, bai_mdl_fnl$mu.pred, family = "gaussian"),  
[…]

### 2021-09-06 Manuscript/01-Intro.Rmd
Finish revise section 1

@@
 [@Wood2020] With the recent advancement of computational technology, many machine learning algorithms that are computationally intensive start to gain much popularity in prognostic modeling and functional selection, especially when complex signals, for example nonlinear effects, are assumed.
-Such models generally have improved prediction accuracy over classic statistical models.
+Such models generally have improved prediction accuracy over classical statistical models.
 -->
@@
 
-The proposed framework, BHAM differs from previous spike-and-slab based GAMs, i.e.
+The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e.
 the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in many ways.
@@
 In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
-Analyses of two metabolomics datasets using the proposed model are presented in Section \ref{sec:real_data}.
+Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.
 Conclusion and discussions are given in Section \ref{sec:concl}.
 
-<!-- Similar to expanding classic models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
+<!-- Similar to expanding classical models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
 The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.

### 2021-09-07 Manuscript/02-Method.Rmd
Finish Revising Section 2.2

@@
 The dimension of the design matrix depends on the choice of the smoothing function.
-For example, a cubic spline function with $k$ knots produces a $k+2$ elements column vector.
-Its corresponding smoothness penalty matrix is denoted as $\bs S_j$.
+For example, a cubic spline function with $k$ knots produces a $k+3$ elements column vector.
+Its corresponding smoothing penalty matrix is denoted as $\bs S_j$.
 $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$.
 To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
-$B_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+$B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
 
-To encourage proper smoothness of the smoothing functions, we adopt the idea of smoothing penalties from smoothing spline models.
-The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework.
-However, the direct integration of smoothness penalties into sparsity penalties is not obvious.
-Given the smoothness penalty matrix is most often symmetric and positive semi-definite ^[ we limit the discussion to univariate smoothing function.
-However, the penalty matrix could possibly contain multiple matrices, e.g.
-tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization.
-The re-parameterization follows Marra and  Wood [@Marra2011], in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
-To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function.
-We further write the orthonormal matrix $\bs U$, containing the eigenvectors as column, as an (row) augmentation of two sub-matrices $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$, where $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
-the penalized space, and $\bs U^0$ contains the eigenvectors for the null space.
+To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models.
+The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function, or least squared and likelihood function in the penalized framework.
+However, the direct integration of smoothing penalties into sparsity penalties is not obvious.
+Marra and  Wood [@Marra2011] proposed a re-parameterization procedure to accommodate the smoothing penalty implicitly.
+Given the smoothing penalty matrix is most often symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smoothing function.
+We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$ containing the eigenvectors as columns in cthe orresponding order.
+That is $\bs U^0$ contains the eigenvectors  with zero eigenvalues for the null space and $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
+the penalized space.
 We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$.
-An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see [@Wood2006]), but I am not sure if it can be generalized to all smoothing functions].
-For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter.
-Similarly, a spline function can be expressed in the matrix form
+An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
+With slight abuse of the notation, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
+A spline function can be expressed in the matrix form
 $$
-f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
+B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
 $$
@@
 \begin{equation}\label{eq:gam-repa}
-E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp)\right].
+E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\bs {\beta_j^0}^T \bs X_j^0 + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right].
 \end{equation}
 
-The re-parameterization sets up the foundation of the proposed model, which has three-fold benefits.
-First of all, as previously mentioned, the re-parameterization integrated the smoothness penalty into the design matrix, and hence the penalty can be expressed in the posterior density function implicitly in addition to the sparse penalty for variable and function selection.
-Secondly, the eigendecomposition of the smooth penalty isolates the linear space of a smooth function from the nonlinear space, improving the feasibility of bi-level variable/function selection and estimation inference.
-The eigendecomposition also introduced the linearly independent column spaces of the smooth function matrix, which allows imposing independent priors on the coefficients.
-This reduces the computational complexity that multivariate priors induces, and greatly broadens the choices of priors and further model choices.
-Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of penalty priors, from column-specific penalties to a unified penalty.
-This step is also necessary for the proposed spike-and-slab spline prior.
-Conventionally, the spike-and-slab priors require scaling of predictors ^[TODO: insert a citation.]
+The re-parameterization step sets up the foundation of the proposed model and provides three-fold benefits.
+First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection.
+Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear from the nonlinear space, improving the feasibility of bi-level function selection and inference.
+The eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
+This reduces the computational complexity compared to using a multivariate priors, and greatly broadens the choices of priors and further model choices.
+Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of priors, from column-specific priors to a unified prior.
 
@@
 [@Bai2021Review] for an in-depth introduction to spike-and-slab priors.
-To summary, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
+To summarize, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{\text{spike}}(\cdot)$ for weak signals and a flat slab density $f_{\text{slab}}(\cdot)$ for strong signals, mathematically
 $$
- \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
+ \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
 $$
-The most distinct feature of the SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
+The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
 [@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.
@@
 
-The major criticism of early spike-and-slab priors receives is being computationally prohibitive.
+The major criticism of early spike-and-slab priors receive is being computationally prohibitive.
 [@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and Spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
@@
 \end{equation}
-The SSL also mitigated the problem of EMVS where the weak signals are not shrink to zero, favored in high-dimensional data analysis.
-We noticed that Bai [@Bai2021] introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
+The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is prefered in high-dimensional data analysis.
+We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
 Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.
@@
 ### Spike-and-Slab Lasso Prior
-We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear effect modeling with bi-level selection.
-The proposed prior extends from the spike-and-slab lasso prior described in Equation \ref{eq:ssl}.
-Given the reparameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}^T$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$.
+We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear modeling with bi-level selection.
+The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
+Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$.
 Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,
@@
 
-where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
-; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
+where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variable and vectors, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
+$s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
 These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation.
@@
 To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively.
-It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection.
-Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
+It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
 $\gamma_j = \gamma^0_j = \gamma^\tp_j$.
-This converges^[Is "becomes" better?] to the SB-GAM [@Bai2021].
+This converges to the SB-GAM [@Bai2021].
+Conversely, it is also possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost of complicating the bi-level function selection.
+This reduces the proposed prior to the classic spike-and-slab lasso prior.
 
-The reparameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
-First of all, the smooth function bases are linear dependent to each other based on the definitions, and ask for extra attention.
-The eigeondecomposition remedies the problem.
-Hence our prior can be set to be conditionally independent.
-Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables, where scaling is mandate.
+The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
+First of all, the smoothing function bases are linear dependent and ask for extra attention.
+The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
+Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables.
 
-The rest of the hierarchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution.
-The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter,
+The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors of $\bs \gamma$ to allow local adaption of the shrinkage using a Bernoulli distribution (binomial distribution of one trial).
+The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
 $$
@@
 $$
-
-This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function.
+This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the null space and penalty space of the smoothing function.
 The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
@@
 
-With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab
-regularization.
+With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
 These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
 Here we focus on the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in additive models.
-^[TODO: enter a sentence introducing EMVS and spike-and-slab mixture normal prior].
 
@@
 The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a].
-Ro{\v{c}}kov{\'{a}} and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
+Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
 <!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.

### 2021-09-07 Manuscript/02-Method.Rmd
Finish revise Section 2

@@
 \begin{align}\label{eq:bham_ssl}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1) \nonumber \\
   \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), 
@@
 Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
-\\
 
@@
   \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
 };
@@
 
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic mnodes, rectangles and are deterministic nodes. }
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
 \label{fig:SSprior}
@@
 $$
-Similar to the BHAM-SSL prior in Equation \ref{eq:bham_ssl}, $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
+Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
 One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects.
@@
 
-Given the proposed model specifications, these models can be easily fitted with MCMC algorithms.
-Nevertheless,the computational burden MCMC algorithms bear creates scalability issues.
+The proposed models can be easily fitted with MCMC algorithms.
+Nevertheless, the computational burden MCMC algorithms bear creates scalability issues.
 George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with continous spike-and-slab priors, and suggested MCMC algorithms works well for medium size of predictors ($p$=25).
-However, it is not feasible to high-dimensionality data analysis where the number of predictors easily exceeds 100.
-Sepcific to additive models, each predictor would expand to mutliple new "predictors" via basis functions, creating greater computational demands.
+However, it is not feasible to high-dimensional data analysis where the number of predictors easily exceeds 100.
+Specific to additive models, each predictor would expand to multiple new "predictors" via basis functions, creating greater computational demands.
 Scheipl et al.
-[@Scheipl2013] demonstrated the computational demands of MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
+[@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
 Hence, we feel compelled to develop scalable and parsimonious algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
 
-Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures.
-The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a].
+As alternatives to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of uncertainty measures.
+The earlier work for fitting spike-and-slab models using optimization algorithms includes EMVS [@Rockova2014a].
 Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
@@
 -->
-In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
 <!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
@@
 -->
-Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit.
-Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results.
-Bai et al.
-[@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.
+Yi and his group independently developed the EM-Iterative Weighted Least Square(IWLS)  and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithms provide deterministic solutions, which becomes a more popular property for reproducible research
+
+<!-- Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion. -->
 
@@
 The main difference between proposed EM-CD algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior.
-In comparison, our prior is pairwise independent requiring no special treatment.
+In comparison, the proposed prior is pairwise independent requiring no special treatment.
 A specific concern of EM-CD algorithm is that it provides no information for inference.
-In the contrast, the EM-IWLS provides a complement to the problem.
-The EM-IWLS can easily extract the variance-covariance matrix due to the IWLS algorithm, and used to construct Wald type inference.
+In contrast, the EM-IWLS can extract the variance-covariance matrix thanks to the IWLS algorithm and construct Wald type inference.
 Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Bayesian ridge and lasso priors.
@@
 
-EM algorithm is an iterative algorithm to find the maximum likelihood estimates or the Bayesian counterpart MAP estimates.
+EM algorithm is an iterative algorithm to find MAP estimates or the maximum likelihood estimates.
 It is commonly used when some necessary data to establish the likelihood function are missing.
-Instead, the algorithm maximizes the expectation, in respect to the "missing" data, of the likelihood function.
+Instead, the algorithm maximizes the expectation, with respect to the "missing" data, of the likelihood function.
 
@@
 
-For BHAMs, we define the parameters of interest as $\Theta = {\bs \beta, \bs \theta, \phi}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
 Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
@@
 & \text{argmax}_{\Theta}
-\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
-&= \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
+&= \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
 \end{aligned}
 $$  
-For simplicity, we consider the null space to be 1-dimension.
+For simplicity, we consider the null space to be 1-dimension, which holds true for many smoothing functions.
 
@@
 We use the EM algorithm to find the MAP estimate of $\Theta$.
-Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithms.
-Naturally In the E-step, we calculate the expectation of posterior density function of $\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{t-1}$,
+Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithm.
+Naturally, in the E-step, we calculate the expectation of posterior density function of $\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{t-1}$,
 $$
-E_{\bs \gamma|\Theta^{t-1}}\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) .
+E_{\bs \gamma|\Theta^{t-1}}\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) .
 $$
-Hereafter, we use the short hand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{t-1}}(\cdot)$.
+Hereafter, we use the shorthand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{t-1}}(\cdot)$.
 In the M-step, we find the $\Theta^{t}$ that maximize $E\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
@@
 
-$$ \log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\beta, \phi) + Q_2 (\gamma,\theta),$$
+$$ \log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\bs \beta, \phi) + Q_2 (\bs \gamma,\bs \theta),$$
 Where
-$$ Q_1(\bs \beta, \phi) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]$$
+$$ Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]$$
 and $$
-Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .$$
-$Q_1$ and $Q_2$ are, respectively, the log-likelihoods of model coefficients and the dispersion parameter $\bs \beta, \phi$, and latent indicators $\bs \gamma, \theta$.
-Conditioning on $\gamma_{jk}$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
+Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .$$
+$Q_1$ and $Q_2$ are, respectively, the log-likelihoods of model coefficients and the dispersion parameter $\bs \beta, \phi$ and latent indicators $\bs \gamma, \theta$.
+Conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
 
 The E- and M- steps are iterated until the algorithm converge.
-Depending on the choice of the mixture priors, the maximization in the M-step can take different algorithms, CD for faster computation and IWLS for uncertainty measures.
+Depending on the choice of the mixture priors, the maximization in the M-step can take different algorithms, CD for fast computation and IWLS for uncertainty measures.
 
@@
 
-When the prior distribution is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$, replacing IRLS, in the M-step.
+When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$ in the M-step.
 Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
 
-Recall, the density function of spike-and-slab mixture double exponential prior can be written as
+To recall, the density function of spike-and-slab mixture double exponential prior can be written as
 $$
@@
 \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log p(\textbf{y}|\bs \beta, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
 \end{equation}
 where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
-To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv p(\gamma^\tp_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem.
-The calculation for $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\beta_{jk}$ for $k=1, \dots, K_j$ and $p^0_j$ only depends on $\beta_j^0$.
+To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv \pr(\gamma^\tp_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem.
+The calculation of $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\bs \beta_{j}$ and $p^0_j$ only depends on $\beta_j^0$.
 The calculation follows the equations below.
@@
 \begin{aligned}
-p_{j}^0 &= \frac{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + Pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + Pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+p_{j}^0 &= \frac{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
+p_{j} &= \frac{\pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{\pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + \pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
 \end{aligned}
@@
 
-and $Pr(\gamma_{j}^{0} = 1|\theta_j) = Pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = \pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
@@
 
-a) Update $\bs \beta^{(t)}$ using the coordinate descent algorithm
+a) Update $\bs \beta^{(t)}$ using the coordinate descent algorithm with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
 
-b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta}
+b) Update $\bs \theta^{(t)}$ using Equation (\ref{eq:update_theta})
 
@@
 
-We assess convergence by the criterion: $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at the $t$th iteration, and $\epsilon$ is a small value (say $10^{-5}$).
+We assess convergence by the criterion: $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $d^{(t)} = -2\log f(\textbf{y}| \textbf{X}, \bs\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at the $t$th iteration, and $\epsilon$ is a small value (say $10^{-5}$).
 
@@
 
-Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
-It can fit the models with a wider range of priors, e.g.
+Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
+It also can fit the models with a wider range of priors, e.g.
 mixture $t$ prior, including mixture normal prior as an special case, and mixture double exponential prior.
-The idea is that the mixture priors can be expressed as a hierarchical normal distribution with various forms of disperssion distributions.
-[@Yi2012] Given the hierarchical normal prior of the coefficients, a generalized model can be fitted using linear approximation and least square algorithm.
-For simplicity, we demonstrate using EM-IWLS algorithm to fit mixture normal priors, with the potential of generalization following Yi and Ma [@Yi2012].
+The idea is that the mixture priors can be expressed as a hierarchical normal distribution with various forms of dispersion distributions.
+[@Yi2012] Given the hierarchical normal prior of the coefficients, a generalized model can be fitted using linear approximation and least squares algorithms.
+For simplicity, we demonstrate using EM-IWLS algorithm to fit continuous spike-and slab priors.
 
-Recall the prior set-up of mixture normal prior, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$
+Recall the prior set-up of continuous spike-and slab priors as a mixture normal distribution, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$
 
@@
 \begin{aligned}
-p(\bs \beta | \bs\gamma) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],
+f( \beta | \gamma, s_0, s_1) &\propto S^{1/2} \exp(-S^{-1} \beta^2/2),
 \end{aligned}
@@
 $$
-  p(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+  f(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
 $$
-    where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight, respectively.The pseudo data and the pseudo-weight are calculated by
+    where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively.The pseudo data and the pseudo-weight are calculated by
 $$
 \begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i).
 \end{aligned}
 $$
-
-  To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an Hessian matrix is yielded as a by-product of the model fitting.
-The Hessian matrix can be further used to construct variance-covariacne matrix of coefficient estimates, and in turn used for statistical inferences.
+In each step, we use least squares algorithms to solve the weight normal linear regression that augments the multivariate normal prior of $\bs \beta$.
+  To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an Hessian matrix can be derived.
+The Hessian matrix can be further used to construct variance-covariance matrix of coefficient estimates and in turn used for statistical inferences.
 
@@
 Our proposed models, BHAM, require two preset scale parameters ($s_0$, $s_1$).
-Hence, we would construct a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
-However, previous research on the topic^[TODO: add citations] suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value of $s_1 = 1$ that provides no or weak shrinkage.
-More attention is focused on examining different values of spike scale $s_0$.
+Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding a criteria of preference.
+This would be achieved by constructing a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
+However, previous research suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value, e.g.
+$s_1 = 1$, that provides no or weak shrinkage.
+[@Rockova2018] As a result, we focus on examining different values of spike scale $s_0$.
 Instead of the 2-D grid, We consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
-Increasing the spike scale s_0 tends to include more non-zero coefficients in the model.
+Increasing the spike scale $s0$ tends to include more non-zero coefficients in the model.
 An measure of preference, e.g.

### 2021-09-11 Manuscript/00-main.Rmd
Finish Revising Section 2.2
Finish revise Section 2
Finish Revision

@@
 ---
-title: Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data
+title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
 author:
@@
 accepted: 2021-03-01
-abstract: "There are many proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional (p>>n) problem using group penalties and priors.
+abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional (p>>n) problem using group penalties and priors.
 These models rarely smooth the nonlinear effect properly because of the sparse regularization.
@@
 <!-- \usepackage{tikz} -->
+\pgfplotsset{compat=1.18}
 \usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
@@
 \newcommand{\tp}{\text{pen}}
+\newcommand{\pr}{\text{Pr}}
 

### 2021-09-11 Manuscript/01-Intro.Rmd
Finish Revising Section 2.2
Finish Revision

@@
 These models are built upon a restrictive and unrealistic assumption, linearity.
-In classical statistical modelling, many strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
+In classical statistical modeling, many strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
 For example, variable categorization is a simple and common practice in epidemiology, but suffers from power and interpretation issues.
@@
 Scheipl et al.
-[@Scheipl2012] proposed an ad hoc spike-and-slab structure prior, normal-mixture-of-inverse gamma prior, that address the previous challenges.
+[@Scheipl2012] proposed a spike-and-slab structure prior, normal-mixture-of-inverse gamma prior, that address the previous challenges.
 But the model fitting relies on computational intensive MCMC algorithmic and creates scalability concern.
@@
 where $\bs S_j$ is a known smoothing penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
-Smoothing penalties re also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
+Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
 Moreover, incorporating the smoothing penalty allows the separation of the linear space of smoothing functions, also know as the null space, from the nonlinear space.
@@
 They make fitting high- and ultrahigh-dimensional models more feasible without sacrificing performance in prediction and variable selection.
-Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent algorithm and EM-Iterative weighted least square algorithm for different utilities.
+Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm for different utilities.
 The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis.
@@
 In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the two fast-fitting algorithms.
-In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
+In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
 Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.

### 2021-09-11 Manuscript/03-Simulation.Rmd
Finish Revising Section 2.2
Finish Revision

@@
 SB-GAM is the first spike-and-slab lasso GAM.
-We implemented COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \text{sparseGAM 1.0}.
+We implemented COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \texttt{sparseGAM 1.0}.
 COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence used the default choices.
@@
 [@Ravikumar2009], stochastic search term selection for GAM [@Scheipl2012].
-However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increasing number of predictors.
+However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increasednumber of predictors.
 [@Scheipl2013]
@@
 We first generated $n=500$ training data points with $p=4, 10, 50, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
-We then simulated the outcome $y$ from three distribution distributions, gaussian, binomial, and poisson with the identity link, logit link $g(x) = \log(\frac{x}{1-x})$, and log link $g(x) = log(x)$.
+We then simulated the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
 The mean of each outcome were simulated via the following function
@@
 $$
-for gaussian and binomial outcomes and shrinked effects
-
-$$
-\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
-$$
-for  poisson outcomes.
+for gaussian and binomial outcomes.
+<!-- and shrinked effects -->
+<!-- $$ -->
+<!-- \mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3)) -->
+<!-- $$ -->
+<!-- for  poisson outcomes.  -->
 Gaussian outcomes required specification of dispersion, where we set the dispersion parameter to be 1.
@@
 Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data.
-The dimensionality of the smooth functions are controlled as k=10.
+The dimensionality of the smoothing functions are controlled as k=10.
 Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.
@@
 ```{r, results = "asis"}
+tab_gaus$mgcv[4] <- "-"
 tab_gaus %>% 
-  xtable(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations", label = "tab:gaus") %>% 
-  print(comment = FALSE)
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>% 
+  xtable(
+    align = "cccccccc",
+    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
+    label = "tab:gaus") %>% 
+  print(comment = FALSE, include.rownames = FALSE)
+  # flextable(
+  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
+  # ) %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
 ```
@@
 ```{r, results = "asis"}
+tab_binom$mgcv[4] <- "-"
+
 tab_binom %>% 
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>%  
   xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
-  print(comment = FALSE)
+  print(comment = FALSE, include.rownames = FALSE)
 ```
@@
 To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
-The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.^[is it better to provide a box plot or a table.]
+The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
 
@@
 The predictive performances have a consistent pattern across the three distributions of outcomes.
-Across all the scenarios, COSSO and adaptive COSSO have the least favourable performance among the applicable methods examined.
+Across all the scenarios, COSSO and adaptive COSSO have the least favorable performance among the applicable methods examined.
 To note, mgcv doesn't support high-dimensional analysis when the number of coefficients are greater than the sample size.
@@
 
-### Simulation Results
+<!-- ### Simulation Results -->
 

### 2021-09-11 Manuscript/05-Discussion.Rmd
Finish Revision

@@
 Last but not least, we plan to expand the proposed model to integrate external biology information.
-The main motivation for the integrative model is that biologically informed groupping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
+The main motivation for the integrative model is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
 Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.

### 2021-09-11 Manuscript/SS_GAM.tex
Finish Revision

@@
 
-\title{Spike-and-Slab Generalized Additive Models and Fast Algorithms
-for High-Dimensional Data}
+\title{Spike-and-Slab Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data}
 
@@
 
+\pgfplotsset{compat=1.18}
 \usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
@@
 restrictive and unrealistic assumption, linearity. In classical
-statistical modelling, many strategies and models are proposed to relax
+statistical modeling, many strategies and models are proposed to relax
 the linearity assumption with various degrees of complexity. For
@@
 accommodate high-dimensional applications. The regularized models,
-branching out from regularized linear models, are used to fit GAMs by
-accounting for the group structure introduced when expanding smoothing
+branching out from group regularized linear models, are used to fit GAMs
+by accounting for the structure introduced when expanding smoothing
 functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
-lasso \citep{Yuan2006} to additive models; Huang et al.
+lasso \citep{Yuan2006} to additive models (AMs); Huang et al.
 \citep{Huang2010} further developed adaptive grouped lasso for additive
 models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
-respectively applied grouped SCAD penalty \citep{Fan2001} to the
-additive models. Recent Bayesian hierarchical models are also used in
-the context of high-dimensional additive models. Various group
-spike-and-slab priors combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Bai et al.
-\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
-to Gaussian additive models(AMs) using a fast optimization algorithm,
-and further generalized the framework to GAMs\citep{Bai2021}. As these
-methods focus on addressing the sparsity, it is very likely to overly
-penalize the basis function coefficients and produce inaccurate
-predictions and curve interpolation, particularly when complex signals
-are assumed and large number of knots are used. \citep{Scheipl2013} In
-addition, the proposed methods adapt an `all-in-all-out' strategy,
-i.e.~either including or excluding the variable completely, rendering no
-space for bi-level selection. Scheipl et al. \citep{Scheipl2012}
-proposed an ad hoc spike-and-slab structure prior,
+respectively applied grouped SCAD penalty \citep{Fan2001} to additive
+models. Recent Bayesian hierarchical models are also used in the context
+of high-dimensional additive models. Various group spike-and-slab priors
+combining with computationally intensive Markov chain Monte Carlo (MCMC)
+algorithms \citep{Xu2015, Yang2020} are proposed, where the application
+on AMs are by-products. Bai et al. \citep{Bai2020} was the first to
+apply group spike-and-slab lasso prior to Gaussian AMs using a fast
+optimization algorithm, and further generalized the framework to
+GAMs\citep{Bai2021}. As these methods focus on addressing the sparsity,
+it is very likely to overly penalize the basis function coefficients and
+produce inaccurate predictions and curve interpolation, particularly
+when complex signals are assumed and large number of knots are used.
+\citep{Scheipl2013} In addition, these methods adapt an `all-in-all-out'
+strategy, i.e.~either including or excluding the variable completely,
+rendering no space for bi-level selection. Scheipl et al.
+\citep{Scheipl2012} proposed an ad hoc spike-and-slab structure prior,
 normal-mixture-of-inverse gamma prior, that address the previous
@@
 interest to develop a fast, flexible and accurate generalized additive
-model.
+model framework.
 
@@
 generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothness penalties in the model via
+Specifically, we incorporate smoothing penalties in the model via
 re-parameterization of the smoothing function to avoid overly shrinking
-basis function coefficients. Smoothness penalties are commonly
-implemented in GAMs through smoothing regression splines. They are
-quadratic norms of the coefficients and allow locally adaptive penalties
-on each smoothing function. The smoothness penalty conditioning on a
-smoothness penalty parameter \(\lambda_j\) is a function of the
-integration of the second derivative of the spline function, expressed
+basis function coefficients. Smoothing penalties are commonly
+implemented in the classical GAMs through smoothing regression splines.
+They are quadratic norms of the coefficients and allow locally adaptive
+penalties on each smoothing function. A smoothing penalty conditioning
+on a smoothing parameter \(\lambda_j\) is a function of the integration
+of the second derivative of the spline function, expressed
 mathematically as \begin{equation}\label{eq:smoothpen}
   \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known smoothness penalty
-matrix, and \(\boldsymbol{\beta}_j\) is the basis function coefficients.
-Smoothness penalties are also previously used in the spike-and-slab GAM
-\citep{Scheipl2012} and the sparsity-smoothness penalty
-\citep{Meier2009}. Moreover, incorporating the smoothness penalty allows
-the separation of the linear space of smoothing functions, also know as
-the null space, from the nonlinear space. We then impose a new bi-part
-spike-and-slab spline prior on the smoothing functions for bi-level
-selection such that the linear and nonlinear spaces of smoothing
-functions can be selected separately. The prior setup encourages a
-flexible solution, rending one of three possibilities: no effect, only
-linear effect, or nonlinear effect of a predictor. In addition, two fast
-computing optimization-based algorithms are developed for high and
-ultrahigh-dimensional\citep{Fan2009} settings respectively.
-
-The proposed framework, BHAM differs from previous spike-and-slab based
+\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
+matrix, and \(\boldsymbol{\beta}_j\) are the basis function
+coefficients. Smoothing penalties re also previously used in the
+spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
+penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
+allows the separation of the linear space of smoothing functions, also
+know as the null space, from the nonlinear space. We then impose a new
+bi-part spike-and-slab spline prior on the smoothing functions for
+bi-level selection such that the linear and nonlinear spaces of
+smoothing functions can be selected separately. The prior setup
+encourages a flexible solution, rending one of three possibilities: no
+effect, only linear effect, or nonlinear effect of a predictor. In
+addition, two fast computing optimization-based algorithms are developed
+for high and ultrahigh-dimensional\citep{Fan2009} settings respectively.
+
+The proposed framework, BHAM, differs from previous spike-and-slab based
 GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
@@
 mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
-Analyses of two metabolomics datasets using the proposed model are
-presented in Section \ref{sec:real_data}. Conclusion and discussions are
-given in Section \ref{sec:concl}.
+Analyses of two metabolomics datasets are presented in Section
+\ref{sec:real_data}. Conclusion and discussions are given in Section
+\ref{sec:concl}.
 
@@
 design matrix depends on the choice of the smoothing function. For
-example, a cubic spline function with \(k\) knots produces a \(k+2\)
-elements column vector. Its corresponding smoothness penalty matrix is
+example, a cubic spline function with \(k\) knots produces a \(k+3\)
+elements column vector. Its corresponding smoothing penalty matrix is
 denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\) denotes the
@@
 treating the parametric function as a special case of the smoothing
-function, e.g.~\(B_j(x_j) = x_j\) with the smoothness penalty matrix
+function, e.g.~\(B_j(x_j) = x_j\) with the smoothing penalty matrix
 defined as \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
 
-To encourage proper smoothness of the smoothing functions, we adopt the
-idea of smoothing penalties from smoothing spline models. The basic idea
-is to set up a smoothing penalty, depending on the smoothness penalty
-matrix via Equation \ref{eq:smoothpen}, in the posterior density
-function, or least squared and likelihood function in the penalized
-framework. However, the direct integration of smoothness penalties into
-sparsity penalties is not obvious. Given the smoothness penalty matrix
-is most often symmetric and positive semi-definite \footnote{ we limit
-  the discussion to univariate smoothing function. However, the penalty
-  matrix could possibly contain multiple matrices, e.g.~tensor product
-  smoother.}, it can be integrated into the design matrix
-\(\boldsymbol{X}\) via a re-parameterization. The re-parameterization
-follows Marra and Wood \citep{Marra2011}, in which eigen-decomposed the
+To encourage proper smoothing of the functions, we adopt the idea of
+smoothing penalties from smoothing spline models. The basic idea is to
+set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}),
+in the prior density function, or least squared and likelihood function
+in the penalized framework. However, the direct integration of smoothing
+penalties into sparsity penalties is not obvious. Marra and Wood
+\citep{Marra2011} proposed a re-parameterization procedure to
+accommodate the smoothing penalty implicitly. Given the smoothing
+penalty matrix is most often symmetric and positive semi-definite for
+univariate smoothing functions, we apply eigendecomposition on the
 penalty matrix
-\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\),
+\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
 where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
@@
 \(0\) elements on the diagonal, associated with the null space of the
-smooth function. We further write the orthonormal matrix
-\(\boldsymbol{U}\), containing the eigenvectors as column, as an (row)
-augmentation of two sub-matrices
-\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\),
-where \(\boldsymbol{U}^{\text{pen}}\) contains the eigenvectors (as
-columns) for the non-zero eigenvalues, i.e.~the penalized space, and
-\(\boldsymbol{U}^0\) contains the eigenvectors for the null space. We
-multiply the basis function matrix \(\boldsymbol{X}\) with the
-orthonormal matrix \(\boldsymbol{U}\) for the new design matrix
+smoothing function. We further write the orthonormal matrix
+\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\)
+containing the eigenvectors as columns in cthe orresponding order. That
+is \(\boldsymbol{U}^0\) contains the eigenvectors with zero eigenvalues
+for the null space and \(\boldsymbol{U}^{\text{pen}}\) contains the
+eigenvectors (as columns) for the non-zero eigenvalues, i.e.~the
+penalized space. We multiply the basis function matrix
+\(\boldsymbol{X}\) with the orthonormal matrix \(\boldsymbol{U}\) for
+the new design matrix
 \(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{\text{pen}} \end{bmatrix}\).
@@
 by the non-zero eigenvalues of \(\boldsymbol{D}\) such that the new
-basis function matrix \(\boldsymbol{X}^\ast\) is scale-invariant
-\footnote{ I know this work for tensor spline (see \citep{Wood2006}),
-  but I am not sure if it can be generalized to all smoothing functions}.
-For notation convenience, we denote
+basis function matrix \(\boldsymbol{X}^\ast\) can receive uniform
+penalty on each of its dimensions. With slight abuse of the notation, we
+denote
 \(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{\text{pen}} \end{bmatrix}\)
-as the after re-parameterized basis function matrix for the \(j\)th
-variable hereafter. Similarly, a spline function can be expressed in the
-matrix form \[
-f_j(x_j) = f_j^0(x_j) + f_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
+as the basis function matrix for the \(j\)th variable after the
+re-parameterization. A spline function can be expressed in the matrix
+form \[
+B_j(x_j) = B_j^0(x_j) + B_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
 \] and the generalized additive model in Equation (\ref{eq:gam}) be
 \begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \boldsymbol{X}_j^0 + {\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen})\right].
+E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + {\boldsymbol{\beta}_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen})\right].
 \end{equation}
 
-The re-parameterization sets up the foundation of the proposed model,
-which has three-fold benefits. First of all, as previously mentioned,
-the re-parameterization integrated the smoothness penalty into the
-design matrix, and hence the penalty can be expressed in the posterior
-density function implicitly in addition to the sparse penalty for
-variable and function selection. Secondly, the eigendecomposition of the
-smooth penalty isolates the linear space of a smooth function from the
-nonlinear space, improving the feasibility of bi-level variable/function
-selection and estimation inference. The eigendecomposition also
-introduced the linearly independent column spaces of the smooth function
-matrix, which allows imposing independent priors on the coefficients.
-This reduces the computational complexity that multivariate priors
-induces, and greatly broadens the choices of priors and further model
-choices. Last but not least, the scaling step of
-\(\boldsymbol{X}^\text{pen}\) further simplifies the choice of penalty
-priors, from column-specific penalties to a unified penalty. This step
-is also necessary for the proposed spike-and-slab spline prior.
-Conventionally, the spike-and-slab priors require scaling of predictors
-\footnote{TODO: insert a citation.}
+The re-parameterization step sets up the foundation of the proposed
+model and provides three-fold benefits. First of all, the
+re-parameterization integrates the smoothing penalty into the design
+matrix, and encourages models to properly smooth the nonlinear function
+in addition to the sparse penalty for function selection. Secondly, the
+eigendecomposition of the smoothing penalty allows the isolation of the
+linear from the nonlinear space, improving the feasibility of bi-level
+function selection and inference. The eigendecomposition facilitates the
+construction of orthonormal design matrix, which makes imposing
+independent priors on the coefficients possible. This reduces the
+computational complexity compared to using a multivariate priors, and
+greatly broadens the choices of priors and further model choices. Last
+but not least, the scaling step of \(\boldsymbol{X}^\text{pen}\) further
+simplifies the choice of priors, from column-specific priors to a
+unified prior.
 
@@
 and variable selection. We defer to Bai et al. \citep{Bai2021Review} for
-an in-depth introduction to spike-and-slab priors. To summary,
+an in-depth introduction to spike-and-slab priors. To summarize,
 spike-and-slab priors are a family of mixture distributions that
-comprises a skinny spike density \(f_{spike}(\cdot)\) for weak signals
-and a flat slab density \(f_{slab}(\cdot)\) for large signals,
-mathematically \[
- \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
-\] The most distinct feature of the SS priors is that it is conditioned
-on a latent binary variable \(\gamma\) that indicates whether the
-variable \(x\) is included in the model. There are various
-spike-and-slab priors depending on the choice for the spike density
-\(f_{spike}(\cdot)\)\footnote{TODO: should I change to notation of \(f\)
-  to not confound with the smooth functions?} and the slab density
-\(f_{slab}(\cdot)\), see George and McCulloch
+comprises a skinny spike density \(f_{\text{spike}}(\cdot)\) for weak
+signals and a flat slab density \(f_{\text{slab}}(\cdot)\) for strong
+signals, mathematically \[
+ \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
+\] The most distinct feature of SS priors is that it is conditioned on a
+latent binary variable \(\gamma\) that indicates whether the variable
+\(x\) is included in the model. There are various spike-and-slab priors
+depending on the choice for the spike density
+\(f_{\text{spike}}(\cdot)\) and the slab density
+\(f_{\text{slab}}(\cdot)\), see George and McCulloch
 \citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
@@
 
-The major criticism of early spike-and-slab priors receives is being
+The major criticism of early spike-and-slab priors receive is being
 computationally prohibitive. \citep{Bai2021Review} Since then, many
@@
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
-\end{equation} The SSL also mitigated the problem of EMVS where the weak
-signals are not shrink to zero, favored in high-dimensional data
-analysis. We noticed that Bai \citep{Bai2021} introduces a
-spike-and-slab group prior in the GAM framework, where the densities of
-the spike and slab components take the group lasso density
-\citep{Xu2015}. Bai \citep{Bai2021} takes an ``all-in-all-out'' strategy
-for function selection.
+\end{equation} The SSL also mitigates the problem of EMVS where the weak
+signals are not shrink to zero, and hence is prefered in
+high-dimensional data analysis. We notice that Bai \citep{Bai2021} is
+the first to apply spike-and-slab lasso prior in the GAM framework,
+where the densities of the spike and slab components take the group
+lasso density \citep{Xu2015}. Bai \citep{Bai2021} takes an
+``all-in-all-out'' strategy for function selection.
 
@@
 We introduce a novel prior for GAMs, particularly for high-dimensional
-nonlinear effect modeling with bi-level selection. The proposed prior
-extends from the spike-and-slab lasso prior described in Equation
-\ref{eq:ssl}. Given the reparameterized deign matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^\text{pen}\end{bmatrix}^T\)
+nonlinear modeling with bi-level selection. The proposed prior extends
+from the spike-and-slab lasso prior described in Equation
+(\ref{eq:ssl}). Given the re-parameterized deign matrix
+\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^\text{pen}\end{bmatrix}\)
 for the \(j\)th variable, we impose a two-part SSL prior to the
@@
 \begin{align}\label{eq:bham_ssl}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\\
+  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1) \nonumber \\
   \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1), 
@@
 where \(\gamma^0_{j}\in\{0,1\}\) and
-\(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator
-variables, indicating if the model includes the linear effect and the
-nonlinear effect of the \(j\)th variable respectively. ; \(s_0\) and
+\(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator variable
+and vectors, indicating if the model includes the linear effect and the
+nonlinear effect of the \(j\)th variable respectively. \(s_0\) and
 \(s_1\) are scale parameters, assuming \(0 < s_0 < s_1\) and given.
@@
 \(\gamma_j^0\), \(\gamma_j^\text{pen}\) respectively. It is possible to
-relax the assumption such that each coefficient \(\beta\) to have its
-own latent indicator \(\gamma\), but at the cost ofcomplicating the
-bi-level function selection. Conversely, it is possible to add a more
-restrictive assumption on the priors, assuming that one indicator
-variable decides the inclusion of both the linear effect and nonlinear
-effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^\text{pen}_j\). This
-converges\footnote{Is ``becomes'' better?} to the SB-GAM
-\citep{Bai2021}.
-
-The reparameterization introduced in Section \ref{sec:BHAM} grants the
-validity of the proposed prior. First of all, the smooth function bases
-are linear dependent to each other based on the definitions, and ask for
-extra attention. The eigeondecomposition remedies the problem. Hence our
-prior can be set to be conditionally independent. Secondly, the
-eigenvalue scaling provides a panacea to allow unified scale parameters
-for all bases of all smoothing functions of variables, where scaling is
-mandate.
+add a more restrictive assumption on the priors, assuming that one
+indicator variable decides the inclusion of both the linear effect and
+nonlinear effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^\text{pen}_j\).
+This converges to the SB-GAM \citep{Bai2021}. Conversely, it is also
+possible to relax the assumption such that each coefficient \(\beta\) to
+have its own latent indicator \(\gamma\), but at the cost of
+complicating the bi-level function selection. This reduces the proposed
+prior to the classic spike-and-slab lasso prior.
+
+The re-parameterization introduced in Section \ref{sec:BHAM} grants the
+validity of the proposed prior. First of all, the smoothing function
+bases are linear dependent and ask for extra attention. The
+eigeondecomposition remedies the problem and hence our prior can be set
+to be conditionally independent. Secondly, the eigenvalue scaling
+provides a panacea to allow unified scale parameters for all bases of
+all smoothing functions of variables.
 
 The rest of the hierarchical prior follows the traditional
-spike-and-slab lasso piror: we set up the hyper-prior of
-\(\boldsymbol{\gamma}\), to allow local adaption of the shrinkage, using
-a bernoulli distribution. The two indicators of the \(j\)th predictor,
-\(\gamma^{0}_j\) and \(\gamma^\text{pen}_j\), shares the same
-probability parameter, \[
+spike-and-slab lasso prior: we set up hyper-priors of
+\(\boldsymbol{\gamma}\) to allow local adaption of the shrinkage using a
+Bernoulli distribution (binomial distribution of one trial). The two
+indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
+\(\gamma^\text{pen}_j\), shares the same probability parameter
+\(\theta_j\), \[
 \begin{aligned}
@@
 \end{aligned}
-\]
-
-This is to leverage the fact that the probability of selecting the bases
-of a smooth function should be similar, while allowing different penalty
-on the null space and penalty space of the spline function. The hyper
-prior of \(\gamma_{j}^{0}\) decides the sparsity of the model at the
-function selection level, while that of \(\gamma_{j}^\text{pen}\)
-decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that \(\gamma_{j}^0\) and
+\] This is to leverage the fact that the probability of selecting the
+bases of a smoothing function should be similar, while allowing
+different penalty on the null space and penalty space of the smoothing
+function. The hyper prior of \(\gamma_{j}^{0}\) decides the sparsity of
+the model at the function selection level, while that of
+\(\gamma_{j}^\text{pen}\) decides the smoothness of the spline function
+at basis function level. Meanwhile, we specify that \(\gamma_{j}^0\) and
 \(\gamma_{j}^\text{pen}\) are independently distributed for analytic
@@
 BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
-\textbackslash{}
 
@@
   \node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1)$};\\
 };
@@
 of applying continuous spike-and-slab prior in additive models.
-\footnote{TODO: enter a sentence introducing EMVS and spike-and-slab
-  mixture normal prior}.
 
@@
 \end{aligned}
-\] Similar to the BHAM-SSL prior in Equation \ref{eq:bham_ssl},
+\] Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}),
 \(0 < s_0 < s_1\) are tuning parameters and can be optimized via
@@
 
-Given the proposed model specifications, these models can be easily
-fitted with MCMC algorithms. Nevertheless,the computational burden MCMC
-algorithms bear creates scalability issues. George and
-McCulloch\citep{George1997} examined the computation speed for various
-MCMC algorithms with continous spike-and-slab priors, and suggested MCMC
-algorithms works well for medium size of predictors (\(p\)=25). However,
-it is not feasible to high-dimensionality data analysis where the number
-of predictors easily exceeds 100. Sepcific to additive models, each
-predictor would expand to mutliple new ``predictors'' via basis
-functions, creating greater computational demands. Scheipl et al.
-\citep{Scheipl2013} demonstrated the computational demands of MCMC
-algorithm for fitting spike-and-slab additive models grow exponentially
-as \(p\) increases modestly via monte carlo simulation studies. Hence,
-we feel compelled to develop scalable and parsimonious algorithms for
-fitting Bayesian hierarchical additive models in high-dimensional
-settings.
-
-Alternative to the sampling algorithms for fitting Bayesian models,
+The proposed models can be easily fitted with MCMC algorithms.
+Nevertheless, the computational burden MCMC algorithms bear creates
+scalability issues. George and McCulloch\citep{George1997} examined the
+computation speed for various MCMC algorithms with continous
+spike-and-slab priors, and suggested MCMC algorithms works well for
+medium size of predictors (\(p\)=25). However, it is not feasible to
+high-dimensional data analysis where the number of predictors easily
+exceeds 100. Specific to additive models, each predictor would expand to
+multiple new ``predictors'' via basis functions, creating greater
+computational demands. Scheipl et al. \citep{Scheipl2013} demonstrated
+the computational demands of a MCMC algorithm for fitting spike-and-slab
+additive models grow exponentially as \(p\) increases modestly via monte
+carlo simulation studies. Hence, we feel compelled to develop scalable
+and parsimonious algorithms for fitting Bayesian hierarchical additive
+models in high-dimensional settings.
+
+As alternatives to the sampling algorithms for fitting Bayesian models,
 optimization algorithms focus on the maximum a posteriori (MAP)
-estimates and speed up the process at the cost of uncertainty measures.
-The earliest work for fitting spike-and-slab models using optimization
-algorithm is EMVS \citep{Rockova2014a}. Ro\{\v{c}\}kov\{'\{a\}\} and
-George\citep{Rockova2014a} proposed an expectation-maximization (EM)
-based algorithm to fit models that use continuous mixture normal priors.
-In the E step, the latent variables \(\gamma\) are treated as the
-missing data, and calculate the posterior mean of the density function;
-in the M step, a ridge estimator was used to update the coefficients,
-and \(\sigma^2, \theta\) are updated accordingly. The same authors
-\citep{Rockova2018b, Rockova2018} further combined the EM algorithm with
-coordinate descent (CD) algorithm for fitting SSL models. Yi and his
-group incorporated the SSL prior in the EM-Iterative Weighted Least
-Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic
-coordinate descent, based on the Lasso penalty SSL exhibit. Such
-applications were used in generalized linear models\citep{Tang2017a},
-cox models \citep{Tang2017} and their grouped
+estimates and speed up the model fitting process at the cost of
+uncertainty measures. The earlier work for fitting spike-and-slab models
+using optimization algorithms includes EMVS \citep{Rockova2014a}.
+Rockova and George\citep{Rockova2014a} proposed an
+expectation-maximization (EM) based algorithm to fit models that use
+continuous mixture normal priors. In the E step, the latent binary
+indicators \(\gamma\)s are treated as the missing data, and the
+posterior means are calculated conditioning on the current value of
+other parameters; in the M step, a ridge estimator was used to update
+the coefficients, and \(\sigma^2, \theta\) are updated accordingly. The
+same authors \citep{Rockova2018b, Rockova2018} further combined the EM
+algorithm with coordinate descent (CD) algorithm for fitting SSL models.
+Yi and his group independently developed the EM-Iterative Weighted Least
+Square(IWLS) and EM-cyclic coordinate descent algorithms to fit models
+with broader class of priors, SSL included.\citep{Yi2019} These
+algorithms were implemented for generalized linear
+models\citep{Tang2017a}, cox models \citep{Tang2017} and their grouped
 counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a more popular property as
-reproducible results. Bai et al. \citep{Bai2020} provides an up-to-date
-summary of SSL methods and theoretical discussion.
+deterministic solutions, which becomes a more popular property for
+reproducible research
 
@@
 algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for
-their group prior. In comparison, our prior is pairwise independent
-requiring no special treatment. A specific concern of EM-CD algorithm is
-that it provides no information for inference. In the contrast, the
-EM-IWLS provides a complement to the problem. The EM-IWLS can easily
-extract the variance-covariance matrix due to the IWLS algorithm, and
-used to construct Wald type inference. Moreover, the EM-IWLS is a more
-general model fitting algorithm that can be used for fitting not only
-SSL and continuous SS priors but also Bayesian ridge and lasso priors.
+their group prior. In comparison, the proposed prior is pairwise
+independent requiring no special treatment. A specific concern of EM-CD
+algorithm is that it provides no information for inference. In contrast,
+the EM-IWLS can extract the variance-covariance matrix thanks to the
+IWLS algorithm and construct Wald type inference. Moreover, the EM-IWLS
+is a more general model fitting algorithm that can be used for fitting
+not only SSL and continuous SS priors but also Bayesian ridge and lasso
+priors.
 
@@
 
-EM algorithm is an iterative algorithm to find the maximum likelihood
-estimates or the Bayesian counterpart MAP estimates. It is commonly used
-when some necessary data to establish the likelihood function are
-missing. Instead, the algorithm maximizes the expectation, in respect to
-the ``missing'' data, of the likelihood function.
+EM algorithm is an iterative algorithm to find MAP estimates or the
+maximum likelihood estimates. It is commonly used when some necessary
+data to establish the likelihood function are missing. Instead, the
+algorithm maximizes the expectation, with respect to the ``missing''
+data, of the likelihood function.
 
@@
 For BHAMs, we define the parameters of interest as
-\(\Theta = {\boldsymbol{\beta}, \boldsymbol{\theta}, \phi}\) and
+\(\Theta = \{\boldsymbol{\beta}, \boldsymbol{\theta}, \phi\}\) and
 consider the latent binary indicators \(\boldsymbol{\gamma}\) as
@@
 & \text{argmax}_{\Theta}
-\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j)
+\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
+&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
 \end{aligned}
 \]\\
-For simplicity, we consider the null space to be 1-dimension.
+For simplicity, we consider the null space to be 1-dimension, which
+holds true for many smoothing functions.
 
@@
 primary interest, we treat them as the ``missing data'' in the EM
-algorithms. Naturally In the E-step, we calculate the expectation of
+algorithm. Naturally, in the E-step, we calculate the expectation of
 posterior density function of
@@
 the values from previous iteration \(\Theta^{t-1}\), \[
-E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
-\] Hereafter, we use the short hand notation
+E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
+\] Hereafter, we use the shorthand notation
 \(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{t-1}}(\cdot)\). In the
@@
 
-\[ \log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\beta, \phi) + Q_2 (\gamma,\theta),\]
+\[ \log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
 Where
-\[ Q_1(\boldsymbol{\beta}, \phi) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\]
+\[ Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\]
 and \[
-Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log p(\theta_j) .\]
+Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .\]
 \(Q_1\) and \(Q_2\) are, respectively, the log-likelihoods of model
-coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\),
+coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\)
 and latent indicators \(\boldsymbol{\gamma}, \theta\). Conditioning on
-\(\gamma_{jk}\), \(Q_1\) and \(Q_2\) are independent and can be
+\(\boldsymbol{\gamma}\), \(Q_1\) and \(Q_2\) are independent and can be
 maximized separately for \(\boldsymbol{\beta}, \phi\) and
@@
 on the choice of the mixture priors, the maximization in the M-step can
-take different algorithms, CD for faster computation and IWLS for
+take different algorithms, CD for fast computation and IWLS for
 uncertainty measures.
@@
 
-When the prior distribution is set to mixture double exponential,
-coordinate descent algorithm can be used to estimate
-\(\boldsymbol{b}eta\), replacing IRLS, in the M-step. Coordinate descent
-is an optimization algorithm that offers extreme computational
-advantage, and famous for its application in optimizing the \(l_1\)
-penalized likelihood function.
-
-Recall, the density function of spike-and-slab mixture double
+When the prior distribution of the coefficients is set to mixture double
+exponential, coordinate descent algorithm can be used to estimate
+\(\boldsymbol{b}eta\) in the M-step. Coordinate descent is an
+optimization algorithm that offers extreme computational advantage, and
+famous for its application in optimizing the \(l_1\) penalized
+likelihood function.
+
+To recall, the density function of spike-and-slab mixture double
 exponential prior can be written as \[
@@
 penalty \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log p(\textbf{y}|\boldsymbol{\beta}, \phi) + \log p(\phi) + \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
 \end{equation} where
@@
 \(E(S^{-1}_j)\), the posterior probability
-\(p^0_{j} \equiv p(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
-\(p_{j} \equiv p(\gamma^\text{pen}_{j}=1|\Theta^{t-1})\) are necessary,
-which can be derived via Bayes' theorem. The calculation for \(p_j\) is
-slightly different from that of \(p^0_j\), as \(p_j\) depends on the
-value of \(\beta_{jk}\) for \(k=1, \dots, K_j\) and \(p^0_j\) only
-depends on \(\beta_j^0\). The calculation follows the equations below.
+\(p^0_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
+\(p_{j} \equiv \text{Pr}(\gamma^\text{pen}_{j}=1|\Theta^{t-1})\) are
+necessary, which can be derived via Bayes' theorem. The calculation of
+\(p_j\) is slightly different from that of \(p^0_j\), as \(p_j\) depends
+on the value of \(\boldsymbol{\beta}_{j}\) and \(p^0_j\) only depends on
+\(\beta_j^0\). The calculation follows the equations below.
 
@@
 \begin{aligned}
-p_{j}^0 &= \frac{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{Pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + Pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{Pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + Pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+p_{j}^0 &= \frac{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \text{Pr}(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
+p_{j} &= \frac{\text{Pr}(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{\text{Pr}(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + \text{Pr}(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
 \end{aligned}
@@
 and
-\(Pr(\gamma_{j}^{0} = 1|\theta_j) = Pr(\gamma_{j}^\text{pen}= 1|\theta_j) = \theta_j\),
-\(Pr(\gamma_{j}^{0} = 0|\theta_j) = Pr(\gamma_{j}^\text{pen}= 0|\theta_j) = 1-\theta_j\),
+\(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \text{Pr}(\gamma_{j}^\text{pen}= 1|\theta_j) = \theta_j\),
+\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = \text{Pr}(\gamma_{j}^\text{pen}= 0|\theta_j) = 1-\theta_j\),
 \(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
@@
   Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
-  algorithm
+  algorithm with the penalized likelihood function in Equation
+  (\ref{eq:Q1_CD})
 \item
   Update \(\boldsymbol{\theta}^{(t)}\) using Equation
-  \ref{eq:update_theta}
+  (\ref{eq:update_theta})
 \item
@@
 \(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
-\(d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})\) is the estimate of deviance at
-the \(t\)th iteration, and \(\epsilon\) is a small value (say
-\(10^{-5}\)).
+\(d^{(t)} = -2\log f(\textbf{y}| \textbf{X}, \boldsymbol{\beta}^{(t)},\phi^{(t)})\)
+is the estimate of deviance at the \(t\)th iteration, and \(\epsilon\)
+is a small value (say \(10^{-5}\)).
 
@@
 EM-based algorithm where iterative weighted least square algorithm are
-used to find the estimate \(\boldsymbol{\beta}, \phi\) that maximizes
-\(E(Q_1)\), instead of the coordinate descent algorithm. It can fit the
-models with a wider range of priors, e.g.~mixture \(t\) prior, including
-mixture normal prior as an special case, and mixture double exponential
-prior. The idea is that the mixture priors can be expressed as a
-hierarchical normal distribution with various forms of disperssion
+used to find the estimate of \(\boldsymbol{\beta}, \phi\) that maximizes
+\(E(Q_1)\), instead of the coordinate descent algorithm. It also can fit
+the models with a wider range of priors, e.g.~mixture \(t\) prior,
+including mixture normal prior as an special case, and mixture double
+exponential prior. The idea is that the mixture priors can be expressed
+as a hierarchical normal distribution with various forms of dispersion
 distributions. \citep{Yi2012} Given the hierarchical normal prior of the
 coefficients, a generalized model can be fitted using linear
-approximation and least square algorithm. For simplicity, we demonstrate
-using EM-IWLS algorithm to fit mixture normal priors, with the potential
-of generalization following Yi and Ma \citep{Yi2012}.
-
-Recall the prior set-up of mixture normal prior, we have the prior
-densities of \(\boldsymbol{\beta}\) conditioning on
-\(\boldsymbol{\gamma}\)
+approximation and least squares algorithms. For simplicity, we
+demonstrate using EM-IWLS algorithm to fit continuous spike-and slab
+priors.
+
+Recall the prior set-up of continuous spike-and slab priors as a mixture
+normal distribution, we have the prior densities of
+\(\boldsymbol{\beta}\) conditioning on \(\boldsymbol{\gamma}\)
 
@@
 \begin{aligned}
-p(\boldsymbol{\beta }| \boldsymbol{\gamma}) &\propto \prod\limits_{j=1}^{p}\left[{S^0_j}^{-1/2}\exp(-1/2({\beta^0_{j}}^2/S^0_{j}))\prod\limits_{k=1}^{K_j}S_{j}^{-1/2}\exp(-1/2(\beta_{jk}^2/S_{j}))\right],
+f( \beta | \gamma, s_0, s_1) &\propto S^{1/2} \exp(-S^{-1} \beta^2/2),
 \end{aligned}
@@
 approximate by a weighted normal likelihood: \[
-  p(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+  f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
 \] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
-the pseudo-response and pseudo-weight, respectively.The pseudo data and
+the pseudo-response and pseudo-weight respectively.The pseudo data and
 the pseudo-weight are calculated by \[
 \begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}((y_i|\hat\eta_i))}{L^{''}((y_i|\hat\eta_i))}& w_i &= - L^{''}((y_i|\hat\eta_i)),
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i).
 \end{aligned}
-\]
-
-To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an
-Hessian matrix is yielded as a by-product of the model fitting. The
-Hessian matrix can be further used to construct variance-covariacne
-matrix of coefficient estimates, and in turn used for statistical
-inferences.
+\] In each step, we use least squares algorithms to solve the weight
+normal linear regression that augments the multivariate normal prior of
+\(\boldsymbol{\beta}\). To note, as the EM-IWLS algorithm is based on
+the IWLS algorithm, an Hessian matrix can be derived. The Hessian matrix
+can be further used to construct variance-covariance matrix of
+coefficient estimates and in turn used for statistical inferences.
 
@@
 \label{sec:tune} Our proposed models, BHAM, require two preset scale
-parameters (\(s_0\), \(s_1\)). Hence, we would construct a two
-dimensional grid, consists of different pairs of (\(s_0\), \(s_1\))
-value. However, previous research on the topic\footnote{TODO: add
-  citations} suggested the value of slab scale \(s_1\) have less impact
-on the final model and is recommended to be set as a generally large
-value of \(s_1 = 1\) that provides no or weak shrinkage. More attention
-is focused on examining different values of spike scale \(s_0\). Instead
-of the 2-D grid, We consider a sequence of \(L\) decreasing values
+parameters (\(s_0\), \(s_1\)). Hence, we need to find the optimal values
+for the scale parameters such that the model reaches its best prediction
+performance regarding a criteria of preference. This would be achieved
+by constructing a two dimensional grid, consists of different pairs of
+(\(s_0\), \(s_1\)) value. However, previous research suggested the value
+of slab scale \(s_1\) have less impact on the final model and is
+recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
+provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
+on examining different values of spike scale \(s_0\). Instead of the 2-D
+grid, We consider a sequence of \(L\) decreasing values
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
-spike scale s\_0 tends to include more non-zero coefficients in the
+spike scale \(s0\) tends to include more non-zero coefficients in the
 model. An measure of preference, e.g.~area under the curve (AUC), mean
@@
 smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package
-\text{sparseGAM 1.0}. COSSO models and SB-GAM do not provide flexibility
-to define smoothing functions, and hence used the default choices. Both
-mgcv and proposed models allow customized smoothing functions and we
-chose the cubic regression spline. We controlled the dimensionality of
-each smoothing function, 10 bases, for all different choices of
-smoothing functions. 5-fold cross-validation were used for COSSO models,
-SB-GAM and the proposed models for tuning parameter selection based on
-the default selection criteria. 20 default candidates were examined for
-SB-GAM and the proposed models which allow user-specification of tuning
-candidates. All computation was conducted on a high-performance 64-bit
-Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
-processors and 24G of RAM per core and R 3.6.2 \citep{R}.
+\texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
+flexibility to define smoothing functions, and hence used the default
+choices. Both mgcv and proposed models allow customized smoothing
+functions and we chose the cubic regression spline. We controlled the
+dimensionality of each smoothing function, 10 bases, for all different
+choices of smoothing functions. 5-fold cross-validation were used for
+COSSO models, SB-GAM and the proposed models for tuning parameter
+selection based on the default selection criteria. 20 default candidates
+were examined for SB-GAM and the proposed models which allow
+user-specification of tuning candidates. All computation was conducted
+on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz
+eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
+3.6.2 \citep{R}.
 
@@
 simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increasing number of
+compared to mgcv and scalability issues with increasednumber of
 predictors. \citep{Scheipl2013}
@@
 \(\text{MVN}_{n\times p}(0, I_{P})\). We then simulated the outcome
-\(y\) from three distribution distributions, gaussian, binomial, and
-poisson with the identity link, logit link
-\(g(x) = \log(\frac{x}{1-x})\), and log link \(g(x) = log(x)\). The mean
-of each outcome were simulated via the following function \[
+\(y\) from two distributions, Gaussian and binomial with the identity
+link and logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The
+mean of each outcome were simulated via the following function \[
 \mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for gaussian and binomial outcomes and shrinked effects
-
-\[
-\mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3))
-\] for poisson outcomes. Gaussian outcomes required specification of
-dispersion, where we set the dispersion parameter to be 1. In this data
-generating process, we have \(x_1, x_2, x_3, x_4\) as the active
-covariates, while the rest covariates are inactive,
+\] for gaussian and binomial outcomes. Gaussian outcomes required
+specification of dispersion, where we set the dispersion parameter to be
+1. In this data generating process, we have \(x_1, x_2, x_3, x_4\) as
+the active covariates, while the rest covariates are inactive,
 i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
@@
 created following the same data generating process, serving as the
-testing data. The dimensionality of the smooth functions are controlled
-as k=10. Hence, the corresponding number of coefficient parameters to be
-estimated is 37, 82, 451 and 1801.
+testing data. The dimensionality of the smoothing functions are
+controlled as k=10. Hence, the corresponding number of coefficient
+parameters to be estimated is 37, 82, 451 and 1801.
 
@@
 \centering
-\begin{tabular}{rrllllll}
+\begin{tabular}{ccccccc}
   \hline
- & p & bglm\_spline\_de & blasso\_spline & cosso & acosso & mgcv & SB\_GAM \\ 
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
   \hline
-1 & 4.00 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-  2 & 10.00 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
-  3 & 50.00 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
-  4 & 200.00 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & NaN (NA) & 0.81 (0.05) \\ 
+  4 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   50 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
+  200 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & - & 0.81 (0.05) \\ 
    \hline
@@
 \centering
-\begin{tabular}{rrllllll}
+\begin{tabular}{rllllll}
   \hline
- & p & bglm\_spline\_de & blasso\_spline & cosso & acosso & mgcv & SB\_GAM \\ 
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
   \hline
-1 & 4.00 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
-  2 & 10.00 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-  3 & 50.00 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  4 & 200.00 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & NaN (NA) & 0.92 (0.01) \\ 
+  4 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
+   10 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
+   50 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
+  200 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
    \hline
@@
 across 50 simulations. The results are tabulated in Table
-\ref{tab:gaus}, \ref{tab:bin_auc}.\footnote{is it better to provide a
-  box plot or a table.}
+\ref{tab:gaus}, \ref{tab:bin_auc}.
 
@@
 distributions of outcomes. Across all the scenarios, COSSO and adaptive
-COSSO have the least favourable performance among the applicable methods
+COSSO have the least favorable performance among the applicable methods
 examined. To note, mgcv doesn't support high-dimensional analysis when
@@
 ~
-
-\hypertarget{simulation-results}{%
-\subsubsection{Simulation Results}\label{simulation-results}}
 
@@
 external biology information. The main motivation for the integrative
-model is that biologically informed groupping of weak effects increases
+model is that biologically informed grouping of weak effects increases
 the power of detecting true associations between features and the

### 2021-09-11 R/summary_cv_SBGAM.R

@@ -9,7 +9,7 @@ summary_cv_SBGAM <- function(cv.SBGAM_output, family = "gaussian"){
 
 
 summary_cv_SBGAM_per_fold <- function(cv.SBGAM_output, family = "gaussian"){
-  browser()
+  # browser()
   ret <- map_dfr(1:length(cv.SBGAM_output$lambda0),
           .f = function(i){
             map_dfr(unique(cv.SBGAM_output$folds), .f = function(j){

### 2021-09-11 _targets.R

@@ -92,11 +92,16 @@ tar_plan(
                        a = 1, b = 1),
   
   ECB_SBGAM_cv = summary_cv_SBGAM(ECB_SBGAM_cv_raw, family = "binomial"),
+  # ECB_SBGAM_cv_per_fold = summary_cv_SBGAM_per_fold(ECB_SBGAM_cv_raw, family = "binomial"),
+  # ggplot(ECB_SBGAM_cv_per_fold)+
+  #   geom_violin(aes(x = lambda0, y = deviance)),
   
   ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 24, a = 1, b = 1),
 
     tar_target(ECB_SBGAM_insample_msr,
              measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial")),
+  
+  
   # 
   # train_msr = data.frame("method" = c("BHAM-SSL", "SB-GAM"),
   #                        rbind(ECB_bamlasso_insample_msr ,ECB_SBGAM_insample_msr)),
@@ -164,6 +169,10 @@ tar_plan(
     max.iter=100, tol = 1e-6, nlambda0 = 10, nfolds = 10),
   
   WLM_SBGAM_cv = summary_cv_SBGAM(WLM_SBGAM_cv_raw),
+  # WLM_SBGAM_cv_per_fold = summary_cv_SBGAM_per_fold(WLM_SBGAM_cv_raw),
+  # 
+  # ggplot(WLM_SBGAM_cv_per_fold %>% filter(lambda0!=23.7))+
+  #   geom_violin(aes(x = lambda0, y = deviance)),
   
   WLM_SBGAM_fnl = SBGAM(
     y = WLM_train_dat$out_HOMA_std,

### 2021-09-13 Manuscript/02-Method.Rmd
Revisions based on Dr. Yi's comment

@@
 E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+\end{equation}
+and the data distribution
+\begin{equation}
+f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,)
 \end{equation}
@@
 
-The proposed models can be easily fitted with MCMC algorithms.
-Nevertheless, the computational burden MCMC algorithms bear creates scalability issues.
-George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with continous spike-and-slab priors, and suggested MCMC algorithms works well for medium size of predictors ($p$=25).
+The proposed models can be fitted with MCMC algorithms.
+Nevertheless, the computational burden of MCMC algorithms creates scalability issues.
+George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with continous spike-and-slab priors, and suggested MCMC algorithms works well for medium size ($p$=25) of predictors with only linear effects .
 However, it is not feasible to high-dimensional data analysis where the number of predictors easily exceeds 100.
@@
 [@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
-Hence, we feel compelled to develop scalable and parsimonious algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
+Hence, we feel compelled to develop scalable algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
 
@@
 -->
-Yi and his group independently developed the EM-Iterative Weighted Least Square(IWLS)  and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Yi and his group independently developed the EM-Iterative Weighted Least Square(IWLS)  and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
 Both EM based algorithms provide deterministic solutions, which becomes a more popular property for reproducible research
@@
 In the follow section, we expand the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
-To note, the two proposed algorithm provides different utilities.
+To note, the two proposed algorithms provides different utilities.
 The EM-CD algorithm is specifically for fitting BHAM-SSL with an expedited performance, recommending to use in high and ultra-high dimensional setting.
@@
 In contrast, the EM-IWLS can extract the variance-covariance matrix thanks to the IWLS algorithm and construct Wald type inference.
-Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Bayesian ridge and lasso priors.
+Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
 
@@
 * E-step: to calculate the expectation of the likelihood function conditioning on some "missing" data
-* M-step: to maximize the spectated likelihood function to calculate the parameter of interest
+* M-step: to maximize the expected likelihood function to calculate the parameter of interest
 
@@
 
-When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate $\bs beta$ in the M-step.
+When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate the parameters in the M-step.
 Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
@@
 
-a) Update $\bs \beta^{(t)}$ using the coordinate descent algorithm with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
+a) Update $\bs \beta^{(t)}$ and the dispersion parameter $\phi$, if exists, using the coordinate descent algorithm with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
 
 b) Update $\bs \theta^{(t)}$ using Equation (\ref{eq:update_theta})
-
-c) Update the dispersion parameter $\phi$ if exists.
 
@@
 Instead of the 2-D grid, We consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
-Increasing the spike scale $s0$ tends to include more non-zero coefficients in the model.
-An measure of preference, e.g.
-area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
+Increasing the spike scale $s_0$ tends to include more non-zero coefficients in the model.
+An measure of preference calculated with cross-validations, e.g.
+deviance, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
 The procedure is similar to the lasso implemented in the widely used R package `glmnet`, which quickly fits the lasso model over a list of values of $\lambda$ covering the entire range, giving a sequence of models for users to choose from.

### 2021-09-13 Manuscript/03-Simulation.Rmd
Revisions based on Dr. Yi's comment

@@
 5-fold cross-validation were used for COSSO models, SB-GAM and the proposed models for tuning parameter selection based on the default selection criteria.
-20 default candidates were examined for SB-GAM and the proposed models which allow user-specification of tuning candidates.
+20 default candidates of tuning parameters ($s0$ in BHAM, $\lambda_0$ in SB-GAM) were examined for SB-GAM and the proposed models that allow user-specification of tuning candidates.
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
@@
 mgcv predicts well when $p$ is small or moderate (p = 4, 10), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the proposed models, bglm_spline_de and blasso_spline ^[need to change the name] predicts better than SB-GAM when the dimension of are small, moderate, and high.
+Among the three fast-computing Bayesian hierarchical models, the proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of are small, moderate, and high.
 However, in the hyper high dimensional case, SB-GAM has better performance.
-Among the proposed methods, bglm_spline_de performs consistently better than blasso_spline.
+Among the proposed methods, BHAM-IWLS performs consistently better than BHAM-CD.
 

### 2021-09-13 Manuscript/04-Real_Data.Rmd
Revisions based on Dr. Yi's comment

@@
 
-In this section, we apply the proposed models BHAM-SSL to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+In this section, we apply the proposed models BHAM-SSL, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].

### 2021-09-13 Manuscript/00-main.Rmd
Replace all ^\tp with *

@@
   num: a  
-- name: Nengjun Yi
+- name: Nengjun Yi*
   num: a
@@
   org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{nyi@uab.edu}"
+corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu} and  \\email{nyi@uab.edu}"
 presentaddress: This is sample for present address text this is sample for present address text
@@
 \newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\tp}{\text{pen}}
+\newcommand{\tp}{*}
 \newcommand{\pr}{\text{Pr}}

### 2021-09-13 Manuscript/02-Method.Rmd
Replace all ^\tp with *

@@
 \end{equation}
-and the data distribution
+and the data distribution is
 \begin{equation}
-f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,)
+f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,\phi),\nonumber
 \end{equation}
@@
 Given the smoothing penalty matrix is most often symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
-To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smoothing function.
-We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$ containing the eigenvectors as columns in cthe orresponding order.
-That is $\bs U^0$ contains the eigenvectors  with zero eigenvalues for the null space and $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
-the penalized space.
-We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$.
-An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
-With slight abuse of the notation, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
+To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the linear space of the smoothing function.
+We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\tp}\end{bmatrix}$ containing the eigenvectors as columns in the corresponding order.
+That is $\bs U^0$ contains the eigenvectors with zero eigenvalues for the linear space and $\bs U^{\tp}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
+the non-linear space.
+We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\tp} \end{bmatrix}$.
+An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
+With slight abuse of the notation, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
 A spline function can be expressed in the matrix form
@@
 $$
-The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model.
+The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma \in \{0,1\}$ that indicates whether the variable $x$ is included in the model.
 There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
@@
 The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+\begin{equation} 
+\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
+\end{equation}
+Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as a double exponential density,
 \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
+\beta | \gamma \sim DE(0, (1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1), 0 < s_0 < s_1,
 \end{equation}
-The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is prefered in high-dimensional data analysis.
+with the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$.
+The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is preferred in high-dimensional data analysis.
 We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
@@
 Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$.
-Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,
+Specifically, we impose independent group priors on the linear space coefficients and on the nonlinear space coefficients respectively,
 
@@
 $$
-This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the null space and penalty space of the smoothing function.
+This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
 The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
@@
 \node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\bs \beta^T \bs X_i), \phi)$};
 
@@
    fit=(linearGamma) (linearBeta),
-   label=above:Null Space:] {};
+   label=above:Linear Space:] {};
   \node [background,
     fit=(penGamma) (penBeta),
-    label=above:Penalized Space:] {};
+    label=above:Nonlinear Space:] {};
 \end{pgfonlayer}
@@
 -->
-The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
+The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm for fitting SSL models.
 <!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
@@
 -->
-Yi and his group independently developed the EM-Iterative Weighted Least Square(IWLS)  and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
 Both EM based algorithms provide deterministic solutions, which becomes a more popular property for reproducible research
@@
 \log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
+&= \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
 \end{aligned}
 $$  
-For simplicity, we consider the null space to be 1-dimension, which holds true for many smoothing functions.
+For simplicity, we consider the linear space to be 1-dimension, which holds true for many smoothing functions.
 
-<!-- , and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{pen}$ from the coefficients of the the penalized space.
-Using the new notation, the part component $\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{pen}_{jk}|\gamma^{pen}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
+<!-- , and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{*}$ from the coefficients of the the penalized space.
+Using the new notation, the part component $\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{*}_{jk}|\gamma^{*}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
 Without loos of generality, the model still work for the spline function whose dimensionality of the null space is more than 1.
@@
 Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithm.
-Naturally, in the E-step, we calculate the expectation of posterior density function of $\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{t-1}$,
+Naturally, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
 $$
-E_{\bs \gamma|\Theta^{t-1}}\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) .
+E_{\bs \gamma|\Theta^{(t-1)}}\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) .
 $$
-Hereafter, we use the shorthand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{t-1}}(\cdot)$.
-In the M-step, we find the $\Theta^{t}$ that maximize $E\log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
+Hereafter, we use the shorthand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{(t-1)}}(\cdot)$.
+In the M-step, we find the $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
 To note here, the log-posterior density of BHAMs (up to additive constants) can be written as a two-part equation
 
-$$ \log p(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\bs \beta, \phi) + Q_2 (\bs \gamma,\bs \theta),$$
+$$ \log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\bs \beta, \phi) + Q_2 (\bs \gamma,\bs \theta),$$
 Where
-$$ Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]$$
+$$ Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{jk})\right]$$
 and $$
-Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .$$
+Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .$$
 $Q_1$ and $Q_2$ are, respectively, the log-likelihoods of model coefficients and the dispersion parameter $\bs \beta, \phi$ and latent indicators $\bs \gamma, \theta$.
@@
 $$
-f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
+f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
 $$
@@
 where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
-To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{t-1})$ and $p_{j} \equiv \pr(\gamma^\tp_{j}=1|\Theta^{t-1})$ are necessary, which can be derived via Bayes' theorem.
+To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j} \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
 The calculation of $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\bs \beta_{j}$ and $p^0_j$ only depends on $\beta_j^0$.
@@
 p_{j}^0 &= \frac{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{\pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{\pr(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + \pr(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+p_{j} &= \frac{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) }{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) + \pr(\gamma^{\tp}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=0, s_0)}
 \end{aligned}
@@
 It is trivial to show
-$$
-\begin{aligned}
-&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{pen}_{j}) = p_{j}\\
+\begin{align}\label{eq:exp_scale}
+&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{\tp}_{j}) = p_{j}\nonumber\\
 &E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
-\end{aligned}
-$$
+\end{align}
 
@@
 
-E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration
+E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{\tp}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration
 
@@
 
-### EM-IRLS
+### EM-IWLS
 
-Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
+Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
 It also can fit the models with a wider range of priors, e.g.
 mixture $t$ prior, including mixture normal prior as an special case, and mixture double exponential prior.
-The idea is that the mixture priors can be expressed as a hierarchical normal distribution with various forms of dispersion distributions.
-[@Yi2012] Given the hierarchical normal prior of the coefficients, a generalized model can be fitted using linear approximation and least squares algorithms.
-For simplicity, we demonstrate using EM-IWLS algorithm to fit continuous spike-and slab priors.
+The basic idea is to formulate the priors as hierarchical normal distributions and augment with the weighted normal likelihood of the generalized model, such that the model is feasible to fit via least squares algorithms.
+[@Yi2012] The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.
 
-Recall the prior set-up of continuous spike-and slab priors as a mixture normal distribution, we have the prior densities of $\bs \beta$ conditioning on $\bs \gamma$
+A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variances, following gamma distributions, integrated out:
+\begin{align*}
+  \beta &\sim N(0, \tau^2)\\
+  \tau^2 & \sim Gamma(1, 1/(2S^2)), 
+\end{align*}
+where $S = (1-\gamma)s_0 + \gamma s_1$ for spike-and-slab lasso priors. Now $Q_1$ needs to account for the hyper prior of $\tau^2$ and is expressed as 
+\begin{equation}\label{eq:Q1_IWLS}
+Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|{\tau^0}^2_j) + \log f({\tau^0}^2_j| S_j^0)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
+\end{equation}
+As $\bs \tau^2$ are not of primary interest, we treat them as the "missing" data besides the latent indicators $\bs \gamma$, and hence need to construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$.
+To note, unlike the coefficients of the non-linear terms $\beta^\tp_{jk}$ which share the same latent indicator $\gamma^\tp_j$, $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
+Primarily, four quantities need to be calculate to formulate $E(Q_1), E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$.
+As neither $E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived via Bayes' theorem as in Equation (\ref{eq:exp_scale}).
+Shown in Yi and Ma [@Yi2012], the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution, and hence
+\begin{align*}
+ E({\tau^0}^{-2}_{j}|S_j^0, \beta^0_j) ={S_j^0}^{-1}/|\beta^0_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
+\end{align*}
+where $S_j^0$ and $S_j^\tp$ are replaced by the expectation.
 
+To simplify $E(Q_1)$ by removing constant additive terms, we have 
+\begin{equation}\label{eq:Q1_IWLS}
+E(Q_1(\bs \beta, \phi)) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)E({\tau_j^0}^{-2}}{2\phi} +\sum\limits_{k=1}^{K_j} \frac{({\beta_{jk}^\tp}^2)E({\tau_{jk}^\tp}^{-2}}{2\phi}\right],
+\end{equation}
+where $\frac{({\beta_{jk}^\tp}^2)E({\tau_{jk}^\tp}^{-2}}{2\phi}$ is the kernel of a normal density with mean 0 and variance $E({\tau_{jk}^\tp}^{-2})\phi$.
+
+Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration of the algorithms by the weighted normal likelihood:
 $$
-\begin{aligned}
-f( \beta | \gamma, s_0, s_1) &\propto S^{1/2} \exp(-S^{-1} \beta^2/2),
-\end{aligned}
-$$
-where $S_{j}^0$ and $S_{j}$ are defined the same as in Equation \ref{eq:Q1_CD}.
-Similarly, $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$ are derived via Bayes' theorem, conditioning on $p^0_j$ and $p_j$.
-We re-write $E(Q_1)$ as a $l_2$ penalized likelihood function, and maximize it using the IWLS algorithm.
-The likelihood function can be approximate by a weighted normal likelihood:
-$$
-  f(\textbf{y}|\bs \beta, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+  f(\textbf{y}_i|\bs \beta, \phi) \approx Normal(\textbf{z}_i|X_i\beta, w_i^{-1}\phi)
 $$
@@
 \begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i).
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
 \end{aligned}
 $$
-In each step, we use least squares algorithms to solve the weight normal linear regression that augments the multivariate normal prior of $\bs \beta$.
+where $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$.
+With $z\sim N(X\beta, \Sigma\phi)$ and $\beta \sim N(0, \phi\Sigma_{\tau^2})$, we can augment the two normal distributions and iterative solve for $\bs \beta$ and $\phi$.
   To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an Hessian matrix can be derived.
@@
 
-<!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
+<!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ [for $\bs \beta$ and $\bs \theta$.
+For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
 

### 2021-09-13 Manuscript/03-Simulation.Rmd
Replace all ^\tp with *

@@
 5-fold cross-validation were used for COSSO models, SB-GAM and the proposed models for tuning parameter selection based on the default selection criteria.
-20 default candidates of tuning parameters ($s0$ in BHAM, $\lambda_0$ in SB-GAM) were examined for SB-GAM and the proposed models that allow user-specification of tuning candidates.
+20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) were examined for SB-GAM and the proposed models that allow user-specification of tuning candidates.
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
@@
 [@Ravikumar2009], stochastic search term selection for GAM [@Scheipl2012].
-However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increasednumber of predictors.
+However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increased number of predictors.
 [@Scheipl2013]
@@
 Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data.
-The dimensionality of the smoothing functions are controlled as k=10.
-Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.
+We generated 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
+<!-- The dimensional of the smoothing functions are controlled as k=10. Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.  -->
 <!-- We also run the simulation where the number of bases are increased to 25 deliberately. -->

### 2021-09-13 Manuscript/SS_GAM.tex
Replace all ^\tp with *

@@
 \author[a]{D. Leann Long}
-\author[a]{Nengjun Yi}
+\author[a]{Nengjun Yi*}
 
@@
 
-\corres{Nengjun Yi, Department of Biostatistics, University of Alabama
-at Birmingham, Birmingham, USA. \email{nyi@uab.edu}}
+\corres{Boyi Guo and Nengjun Yi, Department of Biostatistics, University
+of Alabama at Birmingham, Birmingham, USA. \email{boyiguo1@uab.edu} and
+\email{nyi@uab.edu}}
 
@@
 
-\abstract{There are many proposals that extend the classical generalized
+\abstract{There are proposals that extend the classical generalized
 additive models (GAMs) to accommodate the high-dimensional
@@
 rendering no space for bi-level selection. Scheipl et al.
-\citep{Scheipl2012} proposed an ad hoc spike-and-slab structure prior,
+\citep{Scheipl2012} proposed a spike-and-slab structure prior,
 normal-mixture-of-inverse gamma prior, that address the previous
@@
 matrix, and \(\boldsymbol{\beta}_j\) are the basis function
-coefficients. Smoothing penalties re also previously used in the
+coefficients. Smoothing penalties were also previously used in the
 spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
@@
 optimization algorithms introduced in this article include EM-Coordinate
-Descent algorithm and EM-Iterative weighted least square algorithm for
-different utilities. The applications of the two aforementioned
-algorithm are fruitful in Bayesian hierarchical models in
+Descent (EM-CD) algorithm and EM-Iterative weighted least square
+(EM-IWLS) algorithm for different utilities. The applications of the two
+aforementioned algorithm are fruitful in Bayesian hierarchical models in
 high-dimensional data analysis.
@@
 \ref{sec:sim}, we compare the proposed framework to state-of-art models,
-mgcv, COSSO and sparse Bayesian GAM via monte carlo simulation studies.
+mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
 Analyses of two metabolomics datasets are presented in Section
@@
 E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+\end{equation} and the data distribution is \begin{equation}
+f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,\phi),\nonumber
 \end{equation} with smoothing functions \(B_j(x_j)\) of the variable
@@
 arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
-\(0\) elements on the diagonal, associated with the null space of the
+\(0\) elements on the diagonal, associated with the linear space of the
 smoothing function. We further write the orthonormal matrix
-\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{\text{pen}}\end{bmatrix}\)
-containing the eigenvectors as columns in cthe orresponding order. That
+\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
+containing the eigenvectors as columns in the corresponding order. That
 is \(\boldsymbol{U}^0\) contains the eigenvectors with zero eigenvalues
-for the null space and \(\boldsymbol{U}^{\text{pen}}\) contains the
+for the linear space and \(\boldsymbol{U}^{*}\) contains the
 eigenvectors (as columns) for the non-zero eigenvalues, i.e.~the
-penalized space. We multiply the basis function matrix
+non-linear space. We multiply the basis function matrix
 \(\boldsymbol{X}\) with the orthonormal matrix \(\boldsymbol{U}\) for
 the new design matrix
-\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{\text{pen}} \end{bmatrix}\).
-An additional scaling step is imposed on \(\boldsymbol{X}^{\text{pen}}\)
-by the non-zero eigenvalues of \(\boldsymbol{D}\) such that the new
-basis function matrix \(\boldsymbol{X}^\ast\) can receive uniform
-penalty on each of its dimensions. With slight abuse of the notation, we
-denote
-\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{\text{pen}} \end{bmatrix}\)
+\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
+An additional scaling step is imposed on \(\boldsymbol{X}^{*}\) by the
+non-zero eigenvalues of \(\boldsymbol{D}\) such that the new basis
+function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
+each of its dimensions. With slight abuse of the notation, we denote
+\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
 as the basis function matrix for the \(j\)th variable after the
@@
 form \[
-B_j(x_j) = B_j^0(x_j) + B_j^\text{pen}(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen},
+B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
 \] and the generalized additive model in Equation (\ref{eq:gam}) be
 \begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + {\boldsymbol{\beta}_j^\text{pen}}^T \boldsymbol{X}_j^\text{pen})\right].
+E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right].
 \end{equation}
@@
 greatly broadens the choices of priors and further model choices. Last
-but not least, the scaling step of \(\boldsymbol{X}^\text{pen}\) further
+but not least, the scaling step of \(\boldsymbol{X}^*\) further
 simplifies the choice of priors, from column-specific priors to a
@@
 \] The most distinct feature of SS priors is that it is conditioned on a
-latent binary variable \(\gamma\) that indicates whether the variable
-\(x\) is included in the model. There are various spike-and-slab priors
-depending on the choice for the spike density
+latent binary variable \(\gamma \in \{0,1\}\) that indicates whether the
+variable \(x\) is included in the model. There are various
+spike-and-slab priors depending on the choice for the spike density
 \(f_{\text{spike}}(\cdot)\) and the slab density
@@
 mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
-mathematically, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
-\end{equation} The SSL also mitigates the problem of EMVS where the weak
-signals are not shrink to zero, and hence is prefered in
-high-dimensional data analysis. We notice that Bai \citep{Bai2021} is
-the first to apply spike-and-slab lasso prior in the GAM framework,
-where the densities of the spike and slab components take the group
-lasso density \citep{Xu2015}. Bai \citep{Bai2021} takes an
+mathematically, \begin{equation} 
+\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
+\end{equation} Given that both double exponential distributions have a
+mean of 0 and the latent indicator \(\gamma\) can only take the value of
+0 or 1, the mixture double exponential distribution can be formulated as
+a double exponential density, \begin{equation} \label{eq:ssl}
+\beta | \gamma \sim DE(0, (1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1), 0 < s_0 < s_1,
+\end{equation} with the scale parameter
+\(S = (1-\gamma)s_0 + \gamma s_1\). The SSL also mitigates the problem
+of EMVS where the weak signals are not shrink to zero, and hence is
+preferred in high-dimensional data analysis. We notice that Bai
+\citep{Bai2021} is the first to apply spike-and-slab lasso prior in the
+GAM framework, where the densities of the spike and slab components take
+the group lasso density \citep{Xu2015}. Bai \citep{Bai2021} takes an
 ``all-in-all-out'' strategy for function selection.
@@
 (\ref{eq:ssl}). Given the re-parameterized deign matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^\text{pen}\end{bmatrix}\)
+\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^*\end{bmatrix}\)
 for the \(j\)th variable, we impose a two-part SSL prior to the
 coefficients \(\boldsymbol{\beta}_j\). Specifically, we impose
-independent group priors on the null space coefficients and on the
-penalized space coefficients respectively,
+independent group priors on the linear space coefficients and on the
+nonlinear space coefficients respectively,
 
@@
   \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1) \nonumber \\
-  \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1), 
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\sim DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), 
 \end{align}
 
-where \(\gamma^0_{j}\in\{0,1\}\) and
-\(\gamma^\text{pen}_{j}\in \{0,1\}^{K_j}\) are latent indicator variable
-and vectors, indicating if the model includes the linear effect and the
-nonlinear effect of the \(j\)th variable respectively. \(s_0\) and
-\(s_1\) are scale parameters, assuming \(0 < s_0 < s_1\) and given.
-These scale parameters \(s_0\) and \(s_1\) can be treated as tuning
-parameters and optimized via cross-validation. A discussion of how to
-choose the scale parameters comes in Section \ref{sec:tune}. To note,
-this prior differs from previous group spike-and-slab lasso priors
-\citep{Tang2018, Tang2019}, where the \(\boldsymbol{\beta}^0_j\) and
-\(\boldsymbol{\beta}^\text{pen}_j\) shares the same indicator variables
-\(\gamma_j^0\), \(\gamma_j^\text{pen}\) respectively. It is possible to
-add a more restrictive assumption on the priors, assuming that one
-indicator variable decides the inclusion of both the linear effect and
-nonlinear effect, i.e.~\(\gamma_j = \gamma^0_j = \gamma^\text{pen}_j\).
-This converges to the SB-GAM \citep{Bai2021}. Conversely, it is also
-possible to relax the assumption such that each coefficient \(\beta\) to
-have its own latent indicator \(\gamma\), but at the cost of
-complicating the bi-level function selection. This reduces the proposed
-prior to the classic spike-and-slab lasso prior.
+where \(\gamma^0_{j}\in\{0,1\}\) and \(\gamma^*_{j}\in \{0,1\}^{K_j}\)
+are latent indicator variable and vectors, indicating if the model
+includes the linear effect and the nonlinear effect of the \(j\)th
+variable respectively. \(s_0\) and \(s_1\) are scale parameters,
+assuming \(0 < s_0 < s_1\) and given. These scale parameters \(s_0\) and
+\(s_1\) can be treated as tuning parameters and optimized via
+cross-validation. A discussion of how to choose the scale parameters
+comes in Section \ref{sec:tune}. To note, this prior differs from
+previous group spike-and-slab lasso priors \citep{Tang2018, Tang2019},
+where the \(\boldsymbol{\beta}^0_j\) and \(\boldsymbol{\beta}^*_j\)
+shares the same indicator variables \(\gamma_j^0\), \(\gamma_j^*\)
+respectively. It is possible to add a more restrictive assumption on the
+priors, assuming that one indicator variable decides the inclusion of
+both the linear effect and nonlinear effect,
+i.e.~\(\gamma_j = \gamma^0_j = \gamma^*_j\). This converges to the
+SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
+assumption such that each coefficient \(\beta\) to have its own latent
+indicator \(\gamma\), but at the cost of complicating the bi-level
+function selection. This reduces the proposed prior to the classic
+spike-and-slab lasso prior.
 
@@
 indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
-\(\gamma^\text{pen}_j\), shares the same probability parameter
-\(\theta_j\), \[
+\(\gamma^*_j\), shares the same probability parameter \(\theta_j\), \[
 \begin{aligned}
 \gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^\text{pen}| \theta_j &\sim Bin(\gamma^\text{pen}_{j}|1, \theta_j).
+\gamma_{j}^*| \theta_j &\sim Bin(\gamma^*_{j}|1, \theta_j).
 \end{aligned}
@@
 bases of a smoothing function should be similar, while allowing
-different penalty on the null space and penalty space of the smoothing
-function. The hyper prior of \(\gamma_{j}^{0}\) decides the sparsity of
-the model at the function selection level, while that of
-\(\gamma_{j}^\text{pen}\) decides the smoothness of the spline function
-at basis function level. Meanwhile, we specify that \(\gamma_{j}^0\) and
-\(\gamma_{j}^\text{pen}\) are independently distributed for analytic
-simplicity. We further specify the parameter \(\theta_j\) follows a beta
+different penalty on the linear space and non-linear space of the
+smoothing function. The hyper prior of \(\gamma_{j}^{0}\) decides the
+sparsity of the model at the function selection level, while that of
+\(\gamma_{j}^*\) decides the smoothness of the spline function at basis
+function level. Meanwhile, we specify that \(\gamma_{j}^0\) and
+\(\gamma_{j}^*\) are independently distributed for analytic simplicity.
+We further specify the parameter \(\theta_j\) follows a beta
 distribution with given shape parameters \(a\) and \(b\), \[
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^\text{pen}\sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^\text{pen}\sim  DE(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1)$};\\
+  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
 };
@@
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^\text{pen}_1, \dots,\beta^0_j, \boldsymbol{\beta}^\text{pen}_j , \dots,\beta^0_p, \boldsymbol{\beta}^\text{pen}_p) $};
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^*_1, \dots,\beta^0_j, \boldsymbol{\beta}^*_j , \dots,\beta^0_p, \boldsymbol{\beta}^*_p) $};
 \node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
 \node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g(\eta_i))$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
 
@@
   \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\nonumber\\
-  \beta^\text{pen}_{jk} | \gamma^\text{pen}_{j},s_0,s_1 &\sim N(0,(1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1).
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\sim N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1).
 \end{aligned}
@@
 
-The proposed models can be easily fitted with MCMC algorithms.
-Nevertheless, the computational burden MCMC algorithms bear creates
-scalability issues. George and McCulloch\citep{George1997} examined the
-computation speed for various MCMC algorithms with continous
-spike-and-slab priors, and suggested MCMC algorithms works well for
-medium size of predictors (\(p\)=25). However, it is not feasible to
+The proposed models can be fitted with MCMC algorithms. Nevertheless,
+the computational burden of MCMC algorithms creates scalability issues.
+George and McCulloch\citep{George1997} examined the computation speed
+for various MCMC algorithms with continous spike-and-slab priors, and
+suggested MCMC algorithms works well for medium size (\(p\)=25) of
+predictors with only linear effects . However, it is not feasible to
 high-dimensional data analysis where the number of predictors easily
@@
 carlo simulation studies. Hence, we feel compelled to develop scalable
-and parsimonious algorithms for fitting Bayesian hierarchical additive
-models in high-dimensional settings.
+algorithms for fitting Bayesian hierarchical additive models in
+high-dimensional settings.
 
@@
 same authors \citep{Rockova2018b, Rockova2018} further combined the EM
-algorithm with coordinate descent (CD) algorithm for fitting SSL models.
-Yi and his group independently developed the EM-Iterative Weighted Least
-Square(IWLS) and EM-cyclic coordinate descent algorithms to fit models
-with broader class of priors, SSL included.\citep{Yi2019} These
-algorithms were implemented for generalized linear
-models\citep{Tang2017a}, cox models \citep{Tang2017} and their grouped
+algorithm with coordinate descent algorithm for fitting SSL models. Yi
+and his group independently developed the EM-Iterative Weighted Least
+Square and EM-cyclic coordinate descent algorithms to fit models with
+broader class of priors, SSL included.\citep{Yi2019} These algorithms
+were implemented for generalized linear models\citep{Tang2017a}, Cox
+proportional hazards models \citep{Tang2017} and their grouped
 counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
@@
 In the follow section, we expand the two EM-based algorithms, EM-CD and
-EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithm
+EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
 provides different utilities. The EM-CD algorithm is specifically for
@@
 is a more general model fitting algorithm that can be used for fitting
-not only SSL and continuous SS priors but also Bayesian ridge and lasso
-priors.
+not only SSL and continuous SS priors but also Student's t-priors and
+double exponential priors.
 
@@
 \item
-  M-step: to maximize the spectated likelihood function to calculate the
+  M-step: to maximize the expected likelihood function to calculate the
   parameter of interest
@@
 \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
+&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
 \end{aligned}
 \]\\
-For simplicity, we consider the null space to be 1-dimension, which
+For simplicity, we consider the linear space to be 1-dimension, which
 holds true for many smoothing functions.
@@
 posterior density function of
-\(\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
+\(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
 respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
-the values from previous iteration \(\Theta^{t-1}\), \[
-E_{\boldsymbol{\gamma}|\Theta^{t-1}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
+the values from previous iteration \(\Theta^{(t-1)}\), \[
+E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
 \] Hereafter, we use the shorthand notation
-\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{t-1}}(\cdot)\). In the
-M-step, we find the \(\Theta^{t}\) that maximize
-\(E\log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). To
+\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}(\cdot)\). In the
+M-step, we find the \(\Theta^{(t)}\) that maximize
+\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). To
 note here, the log-posterior density of BHAMs (up to additive constants)
@@
 
-\[ \log p(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
+\[ \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
 Where
-\[ Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{pen}_{jk}|\gamma^{pen}_{jk})\right]\]
+\[ Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
 and \[
-Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{pen})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{pen}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .\]
+Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .\]
 \(Q_1\) and \(Q_2\) are, respectively, the log-likelihoods of model
@@
 When the prior distribution of the coefficients is set to mixture double
-exponential, coordinate descent algorithm can be used to estimate
-\(\boldsymbol{b}eta\) in the M-step. Coordinate descent is an
-optimization algorithm that offers extreme computational advantage, and
-famous for its application in optimizing the \(l_1\) penalized
-likelihood function.
+exponential, coordinate descent algorithm can be used to estimate the
+parameters in the M-step. Coordinate descent is an optimization
+algorithm that offers extreme computational advantage, and famous for
+its application in optimizing the \(l_1\) penalized likelihood function.
 
@@
 exponential prior can be written as \[
-f(\beta|\gamma, s_0, s_1) = \frac{1}{(1-\gamma)s_0 + \gamma s_1}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
+f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
 \] and \(E(Q_1)\) can be expressed as a likelihood function with \(l_1\)
@@
 \(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S_{j} = (1-\gamma^\text{pen}_{j}) s_0 + \gamma^\text{pen}_{j} s_1\).
-To calculate two unknown quantities \(E({S_j^0}^{-1})\) and
-\(E(S^{-1}_j)\), the posterior probability
-\(p^0_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{t-1})\) and
-\(p_{j} \equiv \text{Pr}(\gamma^\text{pen}_{j}=1|\Theta^{t-1})\) are
-necessary, which can be derived via Bayes' theorem. The calculation of
-\(p_j\) is slightly different from that of \(p^0_j\), as \(p_j\) depends
-on the value of \(\boldsymbol{\beta}_{j}\) and \(p^0_j\) only depends on
+\(S_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
+unknown quantities \(E({S_j^0}^{-1})\) and \(E(S^{-1}_j)\), the
+posterior probability
+\(p^0_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
+\(p_{j} \equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are necessary,
+which can be derived via Bayes' theorem. The calculation of \(p_j\) is
+slightly different from that of \(p^0_j\), as \(p_j\) depends on the
+value of \(\boldsymbol{\beta}_{j}\) and \(p^0_j\) only depends on
 \(\beta_j^0\). The calculation follows the equations below.
@@
 p_{j}^0 &= \frac{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \text{Pr}(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{\text{Pr}(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) }{\text{Pr}(\gamma^{pen}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=1, s_1) + \text{Pr}(\gamma^{pen}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{pen}_{j}=0, s_0)}
+p_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
 \end{aligned}
@@
 and
-\(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \text{Pr}(\gamma_{j}^\text{pen}= 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = \text{Pr}(\gamma_{j}^\text{pen}= 0|\theta_j) = 1-\theta_j\),
+\(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j\),
+\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = \text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta_j\),
 \(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
 \(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
-show \[
-\begin{aligned}
-&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{pen}_{j}) = p_{j}\\
+show \begin{align}\label{eq:exp_scale}
+&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{*}_{j}) = p_{j}\nonumber\\
 &E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
-\end{aligned}
-\]
+\end{align}
 
@@
 
-E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{pen}_{j})\)and
+E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{*}_{j})\)and
 \(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
@@
 \item
-  Update \(\boldsymbol{\beta}^{(t)}\) using the coordinate descent
-  algorithm with the penalized likelihood function in Equation
-  (\ref{eq:Q1_CD})
+  Update \(\boldsymbol{\beta}^{(t)}\) and the dispersion parameter
+  \(\phi\), if exists, using the coordinate descent algorithm with the
+  penalized likelihood function in Equation (\ref{eq:Q1_CD})
 \item
@@
   (\ref{eq:update_theta})
-\item
-  Update the dispersion parameter \(\phi\) if exists.
 \end{enumerate}
@@
 
-\hypertarget{em-irls}{%
-\subsubsection{EM-IRLS}\label{em-irls}}
-
-Similar to the EM-CD algorithm, the EM-IRLS algorithm is an iterative
+\hypertarget{em-iwls}{%
+\subsubsection{EM-IWLS}\label{em-iwls}}
+
+Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative
 EM-based algorithm where iterative weighted least square algorithm are
@@
 including mixture normal prior as an special case, and mixture double
-exponential prior. The idea is that the mixture priors can be expressed
-as a hierarchical normal distribution with various forms of dispersion
-distributions. \citep{Yi2012} Given the hierarchical normal prior of the
-coefficients, a generalized model can be fitted using linear
-approximation and least squares algorithms. For simplicity, we
-demonstrate using EM-IWLS algorithm to fit continuous spike-and slab
-priors.
-
-Recall the prior set-up of continuous spike-and slab priors as a mixture
-normal distribution, we have the prior densities of
-\(\boldsymbol{\beta}\) conditioning on \(\boldsymbol{\gamma}\)
-
-\[
-\begin{aligned}
-f( \beta | \gamma, s_0, s_1) &\propto S^{1/2} \exp(-S^{-1} \beta^2/2),
-\end{aligned}
-\] where \(S_{j}^0\) and \(S_{j}\) are defined the same as in Equation
-\ref{eq:Q1_CD}. Similarly, \(E({S_j^0}^{-1})\) and \(E(S^{-1}_j)\) are
-derived via Bayes' theorem, conditioning on \(p^0_j\) and \(p_j\). We
-re-write \(E(Q_1)\) as a \(l_2\) penalized likelihood function, and
-maximize it using the IWLS algorithm. The likelihood function can be
-approximate by a weighted normal likelihood: \[
-  f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}|X\beta, \phi\Sigma)
+exponential prior. The basic idea is to formulate the priors as
+hierarchical normal distributions and augment with the weighted normal
+likelihood of the generalized model, such that the model is feasible to
+fit via least squares algorithms. \citep{Yi2012} The primary purpose of
+the EM algorithm here is mitigate the computational burden associate
+with the hierarchical normal formulation.
+
+A double exponential prior, \(\beta|S \sim DE(0, S)\) can be formulated
+as a hierarchical normal prior with unknown variances, following gamma
+distributions, integrated out: \begin{align*}
+  \beta &\sim N(0, \tau^2)\\
+  \tau^2 & \sim Gamma(1, 1/(2S^2)), 
+\end{align*} where \(S = (1-\gamma)s_0 + \gamma s_1\) for spike-and-slab
+lasso priors. Now \(Q_1\) needs to account for the hyper prior of
+\(\tau^2\) and is expressed as \begin{equation}\label{eq:Q1_IWLS}
+Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|{\tau^0}^2_j) + \log f({\tau^0}^2_j| S_j^0)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
+\end{equation} As \(\boldsymbol{\tau}^2\) are not of primary interest,
+we treat them as the ``missing'' data besides the latent indicators
+\(\boldsymbol{\gamma}\), and hence need to construct the expectation
+\(E_{\boldsymbol{\gamma}, \boldsymbol{\tau}^2|\Theta^{(t-1)}}(Q_1)\). To
+note, unlike the coefficients of the non-linear terms \(\beta^*_{jk}\)
+which share the same latent indicator \(\gamma^*_j\), \(\tau^2_{jk}\) is
+coefficient specific for \(\beta^*_{jk}\). Primarily, four quantities
+need to be calculate to formulate
+\(E(Q_1), E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\).
+As neither \(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)\) nor
+\(E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on \(\tau^2\)s,
+they can be derived via Bayes' theorem as in Equation
+(\ref{eq:exp_scale}). Shown in Yi and Ma \citep{Yi2012}, the conditional
+posterior density of \(\tau^{-2}\) is an inverse Gaussian distribution,
+and hence \begin{align*}
+ E({\tau^0}^{-2}_{j}|S_j^0, \beta^0_j) ={S_j^0}^{-1}/|\beta^0_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
+\end{align*} where \(S_j^0\) and \(S_j^*\) are replaced by the
+expectation.
+
+To simplify \(E(Q_1)\) by removing constant additive terms, we have
+\begin{equation}\label{eq:Q1_IWLS}
+E(Q_1(\boldsymbol{\beta}, \phi)) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)E({\tau_j^0}^{-2}}{2\phi} +\sum\limits_{k=1}^{K_j} \frac{({\beta_{jk}^*}^2)E({\tau_{jk}^*}^{-2}}{2\phi}\right],
+\end{equation} where
+\(\frac{({\beta_{jk}^*}^2)E({\tau_{jk}^*}^{-2}}{2\phi}\) is the kernel
+of a normal density with mean 0 and variance
+\(E({\tau_{jk}^*}^{-2})\phi\).
+
+Meanwhile, following the classical IWLS, we can approximate the
+generalized model likelihood at each iteration of the algorithms by the
+weighted normal likelihood: \[
+  f(\textbf{y}_i|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}_i|X_i\beta, w_i^{-1}\phi)
 \] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
@@
 \begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i).
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
 \end{aligned}
-\] In each step, we use least squares algorithms to solve the weight
-normal linear regression that augments the multivariate normal prior of
-\(\boldsymbol{\beta}\). To note, as the EM-IWLS algorithm is based on
-the IWLS algorithm, an Hessian matrix can be derived. The Hessian matrix
-can be further used to construct variance-covariance matrix of
-coefficient estimates and in turn used for statistical inferences.
+\] where \(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
+\(L^{''}(y_i|\hat\eta_i, \hat \phi)\) are the first and second
+derivative of the log density,
+\(\log f(\textbf{y}_i|\boldsymbol{\beta}, \phi)\). With
+\(z\sim N(X\beta, \Sigma\phi)\) and
+\(\beta \sim N(0, \phi\Sigma_{\tau^2})\), we can augment the two normal
+distributions and iterative solve for \(\boldsymbol{\beta}\) and
+\(\phi\). To note, as the EM-IWLS algorithm is based on the IWLS
+algorithm, an Hessian matrix can be derived. The Hessian matrix can be
+further used to construct variance-covariance matrix of coefficient
+estimates and in turn used for statistical inferences.
 
@@
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
-spike scale \(s0\) tends to include more non-zero coefficients in the
-model. An measure of preference, e.g.~area under the curve (AUC), mean
-squared error, can be used to facilitate the selection of a final model.
-The procedure is similar to the lasso implemented in the widely used R
-package \texttt{glmnet}, which quickly fits the lasso model over a list
-of values of \(\lambda\) covering the entire range, giving a sequence of
+spike scale \(s_0\) tends to include more non-zero coefficients in the
+model. An measure of preference calculated with cross-validations,
+e.g.~deviance, area under the curve (AUC), mean squared error, can be
+used to facilitate the selection of a final model. The procedure is
+similar to the lasso implemented in the widely used R package
+\texttt{glmnet}, which quickly fits the lasso model over a list of
+values of \(\lambda\) covering the entire range, giving a sequence of
 models for users to choose from.
@@
 selection based on the default selection criteria. 20 default candidates
-were examined for SB-GAM and the proposed models which allow
+of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) were
+examined for SB-GAM and the proposed models that allow
 user-specification of tuning candidates. All computation was conducted
@@
 simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increasednumber of
+compared to mgcv and scalability issues with increased number of
 predictors. \citep{Scheipl2013}
@@
 created following the same data generating process, serving as the
-testing data. The dimensionality of the smoothing functions are
-controlled as k=10. Hence, the corresponding number of coefficient
-parameters to be estimated is 37, 82, 451 and 1801.
+testing data. We generated 50 independent pairs of training and testing
+datasets to evaluate the prediction performance of the chosen models,
+where training datasets are used to fit the models and testing datasets
+used to calculate assessment measures.
 
@@
 fast-computing Bayesian hierarchical models, the proposed models,
-bglm\_spline\_de and blasso\_spline \footnote{need to change the name}
-predicts better than SB-GAM when the dimension of are small, moderate,
-and high. However, in the hyper high dimensional case, SB-GAM has better
-performance. Among the proposed methods, bglm\_spline\_de performs
-consistently better than blasso\_spline.
+BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of
+are small, moderate, and high. However, in the hyper high dimensional
+case, SB-GAM has better performance. Among the proposed methods,
+BHAM-IWLS performs consistently better than BHAM-CD.
 
@@
 
-In this section, we apply the proposed models BHAM-SSL to analyze two
-real-world metabolomics datasets where the outcomes are binary and
-continuous respectively. We demonstrate the improved prediction
-performance compared to the other Bayesian hierarchical additive model,
-SB-GAM \citep{Bai2021}.
+In this section, we apply the proposed models BHAM-SSL, fitted with the
+EM-CD algorithm, to analyze two real-world metabolomics datasets where
+the outcomes are binary and continuous respectively. We demonstrate the
+improved prediction performance compared to the other Bayesian
+hierarchical additive model, SB-GAM \citep{Bai2021}.
 
@@
 component is included. Another possible solution is to impose an
-dependent structure of \(\gamma_{j}^{pen}\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j\). Secondly, while the
+dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
+i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, while the
 proposed model addresses a great deal of analytic problem, analyzing the

### 2021-09-13 _targets.R

@@ -46,6 +46,7 @@ tar_plan(
   tar_target(ECB_train_path,
              "Real_Data/Emory_Card_Biobank/Data/Analysis_data_first_cohort.csv",
              format = "file"),
+  
   tar_target(ECB_train_dat,
              readr::read_csv(ECB_train_path)),
   
@@ -205,7 +206,10 @@ tar_plan(
   
   #* Assemble Manuscript ####
   tar_render(manu, "Manuscript/00-main.Rmd",
-             output_file = "SS_GAM.pdf")
+             output_file = "SS_GAM.pdf"),
+  
+  # tar_render(manu, "Manuscript/00-main_word.Rmd",
+  #            output_file = "SS_GAM.pdf"), 
   
 )
 

### 2021-09-21 Manuscript/02-Method.Rmd
Revise IWLS section

@@
 
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where iterative weighted least square algorithm are used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$, instead of the coordinate descent algorithm.
-It also can fit the models with a wider range of priors, e.g.
-mixture $t$ prior, including mixture normal prior as an special case, and mixture double exponential prior.
-The basic idea is to formulate the priors as hierarchical normal distributions and augment with the weighted normal likelihood of the generalized model, such that the model is feasible to fit via least squares algorithms.
-[@Yi2012] The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.
+Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
+The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
+In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary.
 
-A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variances, following gamma distributions, integrated out:
+<!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
+
+A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2 \sim Gamma(1, 1/(2S^2))$ integrated out:
 \begin{align*}
@@
 \end{align*}
-where $S = (1-\gamma)s_0 + \gamma s_1$ for spike-and-slab lasso priors. Now $Q_1$ needs to account for the hyper prior of $\tau^2$ and is expressed as 
+For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
+The change in the prior formulation in turns leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyper prior of $\tau^2$:
 \begin{equation}\label{eq:Q1_IWLS}
@@
 \end{equation}
-As $\bs \tau^2$ are not of primary interest, we treat them as the "missing" data besides the latent indicators $\bs \gamma$, and hence need to construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$.
-To note, unlike the coefficients of the non-linear terms $\beta^\tp_{jk}$ which share the same latent indicator $\gamma^\tp_j$, $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
-Primarily, four quantities need to be calculate to formulate $E(Q_1), E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$.
-As neither $E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived via Bayes' theorem as in Equation (\ref{eq:exp_scale}).
-Shown in Yi and Ma [@Yi2012], the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution, and hence
+Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
+To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
+$E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
+As neither $E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
+On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
+$E({\tau^0}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
 \begin{align*}
@@
 \end{align*}
-where $S_j^0$ and $S_j^\tp$ are replaced by the expectation.
+where $S_j^0$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
+\begin{equation}\label{eq:EQ1_IWLS}
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)}{2\phi E{\tau_j^0}^{2}} +\sum\limits_{k=1}^{K_j} \frac{{\beta_{jk}^\tp}^2}{2\phi E({\tau_{jk}^\tp}^{2})}\right].
+\end{equation}
+$\frac{\beta^2}{2\phi E(\tau^{2})}$ can be seen as the kernel of a normal density with mean 0 and variance $\phi E(\tau^{2})$.
 
-To simplify $E(Q_1)$ by removing constant additive terms, we have 
-\begin{equation}\label{eq:Q1_IWLS}
-E(Q_1(\bs \beta, \phi)) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)E({\tau_j^0}^{-2}}{2\phi} +\sum\limits_{k=1}^{K_j} \frac{({\beta_{jk}^\tp}^2)E({\tau_{jk}^\tp}^{-2}}{2\phi}\right],
-\end{equation}
-where $\frac{({\beta_{jk}^\tp}^2)E({\tau_{jk}^\tp}^{-2}}{2\phi}$ is the kernel of a normal density with mean 0 and variance $E({\tau_{jk}^\tp}^{-2})\phi$.
-
-Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration of the algorithms by the weighted normal likelihood:
+Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood:
 $$
-  f(\textbf{y}_i|\bs \beta, \phi) \approx Normal(\textbf{z}_i|X_i\beta, w_i^{-1}\phi)
+  f(\textbf{y}|\bs \beta, \phi) \approx MVN(\textbf{z}|X\beta, \Sigma w_i^{-1}\phi)
 $$

### 2021-09-21 Manuscript/00-main_word.Rmd
Add word version

--- 2021-09-13 Manuscript/00-main.Rmd
@@
 ---
-title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
-author:
-- name: Boyi Guo*
-  num: a
-- name: Byron C. Jaeger
-  num: b 
-- name: AKM Fazlur Rahman
-  num: a   
-- name: D. Leann Long
-  num: a  
-- name: Nengjun Yi*
-  num: a
-address:
-- num: a
-  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
-- num: b
-  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu} and  \\email{nyi@uab.edu}"
-presentaddress: This is sample for present address text this is sample for present address text
-authormark: Guo et al.
-articletype: Research article
-received: 2021-01-01
-revised: 2021-02-01
-accepted: 2021-03-01
+title: "Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data"
+
 abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional (p>>n) problem using group penalties and priors.
@@
 output: 
-  rticles::sim_article:
-    keep_tex: TRUE
+  word_document:
+    reference_docx: template.docx
 header-includes: 
@@
 
+\newpage
+
 ```{r child="01-Intro.Rmd"}

### 2021-09-21 Manuscript/02-Method.Rmd
Add word version

@@
 where $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$.
-With $z\sim N(X\beta, \Sigma\phi)$ and $\beta \sim N(0, \phi\Sigma_{\tau^2})$, we can augment the two normal distributions and iterative solve for $\bs \beta$ and $\phi$.
-  To note, as the EM-IWLS algorithm is based on the IWLS algorithm, an Hessian matrix can be derived.
-The Hessian matrix can be further used to construct variance-covariance matrix of coefficient estimates and in turn used for statistical inferences.
+With $z\sim N(X\beta, \Sigma\phi)$ and $\beta \sim N(0, \phi\Sigma_{\tau^2})$, we can augment the two normal distributions and derive the estimates for $\bs \beta$ and $\phi$ in each step in closed form equations.
+To note, an estimate of the variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences.
 

### 2021-09-21 Manuscript/SS_GAM.tex
Add word version

@@
 Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative
-EM-based algorithm where iterative weighted least square algorithm are
-used to find the estimate of \(\boldsymbol{\beta}, \phi\) that maximizes
-\(E(Q_1)\), instead of the coordinate descent algorithm. It also can fit
-the models with a wider range of priors, e.g.~mixture \(t\) prior,
-including mixture normal prior as an special case, and mixture double
-exponential prior. The basic idea is to formulate the priors as
-hierarchical normal distributions and augment with the weighted normal
-likelihood of the generalized model, such that the model is feasible to
-fit via least squares algorithms. \citep{Yi2012} The primary purpose of
-the EM algorithm here is mitigate the computational burden associate
-with the hierarchical normal formulation.
+EM-based algorithm where the iterative weighted least squares algorithm
+is used to find the estimate of \(\boldsymbol{\beta}, \phi\) that
+maximizes \(E(Q_1)\). The iterative weighted least squares algorithm was
+originally proposed to fit the classical generalized linear models, and
+generalized to fit some Bayesian hierarchical models.\citep{Gelman2013}
+Yi and Ma \citep{Yi2012} formulated Student's t-distribution and double
+exponential distribution as hierarchical normal distributions such that
+generalized linear models with shrinkage priors can be easily fitted
+using IWLS in combination with EM algorithm. In this work, we adapt the
+EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the
+full potential to adopt Student's t, double exponential and mixture
+Student's t priors when sparse assumption is not necessary.
 
 A double exponential prior, \(\beta|S \sim DE(0, S)\) can be formulated
-as a hierarchical normal prior with unknown variances, following gamma
-distributions, integrated out: \begin{align*}
+as a hierarchical normal prior with unknown variance
+\(\tau^2 \sim Gamma(1, 1/(2S^2))\) integrated out: \begin{align*}
   \beta &\sim N(0, \tau^2)\\
   \tau^2 & \sim Gamma(1, 1/(2S^2)), 
-\end{align*} where \(S = (1-\gamma)s_0 + \gamma s_1\) for spike-and-slab
-lasso priors. Now \(Q_1\) needs to account for the hyper prior of
-\(\tau^2\) and is expressed as \begin{equation}\label{eq:Q1_IWLS}
+\end{align*} For the mixture double exponential priors, we can define
+the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following
+Equation (\ref{eq:ssl}). The change in the prior formulation in turns
+leads to the change in the log posterior density function, as \(Q_1\)
+needs to account for the hyper prior of \(\tau^2\):
+\begin{equation}\label{eq:Q1_IWLS}
 Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|{\tau^0}^2_j) + \log f({\tau^0}^2_j| S_j^0)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
-\end{equation} As \(\boldsymbol{\tau}^2\) are not of primary interest,
-we treat them as the ``missing'' data besides the latent indicators
-\(\boldsymbol{\gamma}\), and hence need to construct the expectation
-\(E_{\boldsymbol{\gamma}, \boldsymbol{\tau}^2|\Theta^{(t-1)}}(Q_1)\). To
-note, unlike the coefficients of the non-linear terms \(\beta^*_{jk}\)
-which share the same latent indicator \(\gamma^*_j\), \(\tau^2_{jk}\) is
-coefficient specific for \(\beta^*_{jk}\). Primarily, four quantities
-need to be calculate to formulate
-\(E(Q_1), E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\).
-As neither \(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)\) nor
+\end{equation} Since \(\boldsymbol{\tau}^2\) are not of our primary
+interest, we treat them as the ``missing'' data in addition to the
+latent indicators \(\boldsymbol{\gamma}\), and hence construct the
+expectation
+\(E_{\boldsymbol{\gamma}, \boldsymbol{\tau}^2|\Theta^{(t-1)}}(Q_1)\) in
+the E-step. To note, unlike the same latent indicator \(\gamma^*_j\)
+which is shared by the coefficients of the non-linear terms
+\(\beta^*_{jk}\) for \(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is
+coefficient specific for \(\beta^*_{jk}\).
+\(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
+needs to be calculated to formulate \(E(Q_1)\). As neither
+\(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)\) nor
 \(E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on \(\tau^2\)s,
-they can be derived via Bayes' theorem as in Equation
-(\ref{eq:exp_scale}). Shown in Yi and Ma \citep{Yi2012}, the conditional
-posterior density of \(\tau^{-2}\) is an inverse Gaussian distribution,
-and hence \begin{align*}
+they can be derived using Equation (\ref{eq:exp_scale}) via Bayes'
+Theorem. On the other hand, \(\tau^{2}\) following a gamma distribution
+is a conjugate prior for the normal variance, and the conditional
+posterior density of \(\tau^{-2}\) is an inverse Gaussian distribution.
+\(E({\tau^0}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
+using the closed form equation \begin{align*}
  E({\tau^0}^{-2}_{j}|S_j^0, \beta^0_j) ={S_j^0}^{-1}/|\beta^0_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
 \end{align*} where \(S_j^0\) and \(S_j^*\) are replaced by the
-expectation.
-
-To simplify \(E(Q_1)\) by removing constant additive terms, we have
-\begin{equation}\label{eq:Q1_IWLS}
-E(Q_1(\boldsymbol{\beta}, \phi)) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)E({\tau_j^0}^{-2}}{2\phi} +\sum\limits_{k=1}^{K_j} \frac{({\beta_{jk}^*}^2)E({\tau_{jk}^*}^{-2}}{2\phi}\right],
-\end{equation} where
-\(\frac{({\beta_{jk}^*}^2)E({\tau_{jk}^*}^{-2}}{2\phi}\) is the kernel
-of a normal density with mean 0 and variance
-\(E({\tau_{jk}^*}^{-2})\phi\).
+expectation and \(\beta\)s are replaced with \(\beta^{(t-1)}\). With
+simplification (up to constant additive terms), we have
+\begin{equation}\label{eq:EQ1_IWLS}
+E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)}{2\phi E{\tau_j^0}^{2}} +\sum\limits_{k=1}^{K_j} \frac{{\beta_{jk}^*}^2}{2\phi E({\tau_{jk}^*}^{2})}\right].
+\end{equation} \(\frac{\beta^2}{2\phi E(\tau^{2})}\) can be seen as the
+kernel of a normal density with mean 0 and variance
+\(\phi E(\tau^{2})\).
 
 Meanwhile, following the classical IWLS, we can approximate the
-generalized model likelihood at each iteration of the algorithms by the
-weighted normal likelihood: \[
-  f(\textbf{y}_i|\boldsymbol{\beta}, \phi) \approx Normal(\textbf{z}_i|X_i\beta, w_i^{-1}\phi)
+generalized model likelihood at each iteration with a weighted normal
+likelihood: \[
+  f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx MVN(\textbf{z}|X\beta, \Sigma w_i^{-1}\phi)
 \] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
@@
 \(\beta \sim N(0, \phi\Sigma_{\tau^2})\), we can augment the two normal
-distributions and iterative solve for \(\boldsymbol{\beta}\) and
-\(\phi\). To note, as the EM-IWLS algorithm is based on the IWLS
-algorithm, an Hessian matrix can be derived. The Hessian matrix can be
-further used to construct variance-covariance matrix of coefficient
-estimates and in turn used for statistical inferences.
+distributions and derive the estimates for \(\boldsymbol{\beta}\) and
+\(\phi\) in each step in closed form equations. To note, an estimate of
+the variance-covariance matrix can be derived in the EM-IWLS algorithm
+and in turn can be used for statistical inferences.
 

### 2021-09-21 _targets.R

@@ -208,8 +208,8 @@ tar_plan(
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
-  # tar_render(manu, "Manuscript/00-main_word.Rmd",
-  #            output_file = "SS_GAM.pdf"), 
+  tar_render(manu_word, "Manuscript/00-main_word.Rmd",
+             output_file = "SS_GAM.docx"),
   
 )
 

### 2021-09-23 Manuscript/01-Intro.Rmd
Revise the manuscript

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be expressed as the summation of smoothing functions, $B_j, j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
@@
 $$
-where $g^{-1}$ is the inverse of a monotonic link function.
+where $g^{-1}(\cdot)$ is the inverse of a monotonic link function.
 The smoothing functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
@@
 \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
 \end{equation}
@@
 Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-Moreover, incorporating the smoothing penalty allows the separation of the linear space of smoothing functions, also know as the null space, from the nonlinear space.
+Moreover, incorporating the smoothing penalty allows the separation of the linear space of smoothing functions from the nonlinear space.
 We then impose a new bi-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
 The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
-In addition, two fast computing optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
+In addition, two scalable optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
 
@@
 Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm for different utilities.
-The applications of the two aforementioned algorithm are fruitful in Bayesian hierarchical models in high-dimensional data analysis.
+The applications of the two aforementioned algorithms are fruitful in Bayesian hierarchical models for high-dimensional data.
 [@Yi2012; @Rockova2014a; @Rockova2018] Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.
@@
 Furthermore, the proposed model addresses the incapability of bi-level selection in SB-GAM.
-Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), make transnational science more accessible.
+Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
 

### 2021-09-26 Manuscript/00-main.Rmd

@@
 accepted: 2021-03-01
-abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional (p>>n) problem using group penalties and priors.
+abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional ($p>>n$) problem using group penalties and priors.
 These models rarely smooth the nonlinear effect properly because of the sparse regularization.
@@
 \newcommand{\pr}{\text{Pr}}
+\newcommand{\repa}{\text{repa}}
+\newcommand{\simiid}{\overset{\text{iid}}{\sim}}
 

### 2021-09-26 Manuscript/02-Method.Rmd

@@
 \begin{equation}
-f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,\phi),\nonumber
+f(\bs y| \bs \beta, \phi) = \prod\limits^n_{i=1}f(y_i|\bs \beta, \phi),\nonumber
 \end{equation}
@@
 Its corresponding smoothing penalty matrix is denoted as $\bs S_j$.
-$\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$.
+$\bs \beta_j$ denotes the basis function coefficients for the $j$th variable such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
+With slight abuse of notation, we denote vectors and matrices in bold fonts $\bs \beta, \bs X$ with conformable dimensions, where scalar and random variables are denoted in with unbold fonts $\beta, X$.
+The matrix transposing operation is denoted with a superscript $^T$.
 To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
@@
 To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models.
-The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function, or least squared and likelihood function in the penalized framework.
-However, the direct integration of smoothing penalties into sparsity penalties is not obvious.
+The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
+However, the direct integration of smoothing penalty with sparsity penalty is not obvious.
 Marra and  Wood [@Marra2011] proposed a re-parameterization procedure to accommodate the smoothing penalty implicitly.
-Given the smoothing penalty matrix is most often symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
-To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the linear space of the smoothing function.
-We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\tp}\end{bmatrix}$ containing the eigenvectors as columns in the corresponding order.
-That is $\bs U^0$ contains the eigenvectors with zero eigenvalues for the linear space and $\bs U^{\tp}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
+Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
+For the most popular smoothing function, cubic splines, the dimension of the linear space is one.
+Hereafter, we focus on discussing a uni-dimensional linear space for simplicity; however, it generalizes easily to the cases where the linear space is multidimensional.
+We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix} \bs U^0 : \bs U^{\tp}\end{bmatrix}$ containing the eigenvectors as columns in the corresponding order to $\bs D$.
+That is, $\bs U$ contains the eigenvectors $U^0$ with zero eigenvalues for the linear space and $\bs U^{\tp}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
 the non-linear space.
-We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\tp} \end{bmatrix}$.
+We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix ${\bs X}^\repa = \bs X \bs U \equiv \begin{bmatrix} X^0 : \bs X^{\tp} \end{bmatrix}$.
 An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
-With slight abuse of the notation, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
+With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
 A spline function can be expressed in the matrix form
 $$
-B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
+B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \beta_j X^0_j + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
 $$
@@
 \begin{equation}\label{eq:gam-repa}
-E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\bs {\beta_j^0}^T \bs X_j^0 + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right].
+E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\beta_j X^0_j + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right],
 \end{equation}
+where the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of non-linear space.
 
@@
 
-The major criticism of early spike-and-slab priors receive is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and Spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
-Particularly, the development of Spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
-The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+The major criticism of early spike-and-slab models is being computationally prohibitive.
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
+Particularly, the development of spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+The spike-and-slab Lasso (SSL) prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
 \begin{equation} 
@@
 \end{equation}
-Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as a double exponential density,
+Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as one single double exponential density,
 \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, (1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1), 0 < s_0 < s_1,
+\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
 \end{equation}
@@
 ### Spike-and-Slab Lasso Prior
+
 We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear modeling with bi-level selection.
 The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
-Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$.
+Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
 Specifically, we impose independent group priors on the linear space coefficients and on the nonlinear space coefficients respectively,
-
 \begin{align}\label{eq:bham_ssl}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1) \nonumber \\
-  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), 
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
+  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j
 \end{align}
-
-where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variable and vectors, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
+where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
 $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
@@
 A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}.
-To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively.
+To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
 It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
-$\gamma_j = \gamma^0_j = \gamma^\tp_j$.
+$\gamma_j = \gamma^\tp_j$.
 This converges to the SB-GAM [@Bai2021].
-Conversely, it is also possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost of complicating the bi-level function selection.
+Conversely, it is also possible to relax the assumption such that each coefficient $\beta_{jk} \in \bs \beta^\tp_j$ has its own latent indicator $\gamma_{jk}$, but at the cost of complicating the bi-level function selection.
 This reduces the proposed prior to the classic spike-and-slab lasso prior.
@@
 
-The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors of $\bs \gamma$ to allow local adaption of the shrinkage using a Bernoulli distribution (binomial distribution of one trial).
-The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
+The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
+The two indicators of the $j$th predictor, $\gamma_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
 $$
 \begin{aligned}
-\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^\tp | \theta_j &\sim Bin(\gamma^\tp_{j}|1, \theta_j).
+&\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
+&\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j).
 \end{aligned}
@@
 This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
-The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
+The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
+Meanwhile, we specify that $\gamma_{j}$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
 We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
@@
 The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience.
-For simplicity, we focus on a special case of beta distribution, uniform (0,1).
+For simplicity, we focus on a special case of beta distribution, uniform (0,1), i.e.
+$a = 1, b = 1$.
 When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
@@
 \matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j^0 \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j^0 \sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)$};\\
+  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
 };
@@
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta^0_1, \bs \beta^\tp_1, \dots,\beta^0_j, \bs \beta^\tp_j , \dots,\beta^0_p, \bs \beta^\tp_p) $};
+\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta_1, \bs \beta^\tp_1, \dots,\beta_j, \bs \beta^\tp_j , \dots,\beta_p, \bs \beta^\tp_p) $};
 \node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
@@
 
-With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
+With the re-parameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
 These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
-Here we focus on the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in additive models.
+Here we elaborate the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in BHAM.
 
-A spike-and-slab mixture normal prior in additive models can be expressed as 
-$$
-\begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\nonumber\\
-  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\sim N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1).
-\end{aligned}
-$$
+A spike-and-slab mixture normal spline prior can be expressed as 
+\begin{align*}
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
+  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j.
+\end{align*}
 Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
-One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which renders problem when estimating the large effects.
-Distributions with heavier tails can be used as an alternative, for example mixture $t$ distribution with small degree of freedom.
+One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
+Distributions with heavier tails can be used as an alternative, for example mixture Student's $t$ distribution with small degree of freedom.
 
@@
 Nevertheless, the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with continous spike-and-slab priors, and suggested MCMC algorithms works well for medium size ($p$=25) of predictors with only linear effects .
-However, it is not feasible to high-dimensional data analysis where the number of predictors easily exceeds 100.
+George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with spike-and-slab mixture normal priors, and suggested MCMC algorithms works well for medium size ($p$=25) of predictors with only linear effects.
+However, it is not feasible for high-dimensional data analysis where the number of predictors easily exceeds 100.
 Specific to additive models, each predictor would expand to multiple new "predictors" via basis functions, creating greater computational demands.
 Scheipl et al.
-[@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab additive models grow exponentially as $p$ increases modestly via monte carlo simulation studies.
+[@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab GAM grow exponentially as $p$ increases modestly via simulation studies.
 Hence, we feel compelled to develop scalable algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
@@
 -->
-In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly.
+In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, followed by updating $\phi, \theta$.
 <!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
@@
 -->
-The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm for fitting SSL models.
+The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm to fit SSL models.
 <!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
@@
 Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithms provide deterministic solutions, which becomes a more popular property for reproducible research
+Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research
 
@@
 
-In the follow section, we expand the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
+In this section, we extend the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
 To note, the two proposed algorithms provides different utilities.
 The EM-CD algorithm is specifically for fitting BHAM-SSL with an expedited performance, recommending to use in high and ultra-high dimensional setting.
-SB-GAM[@Bai2020; @Bai2021] also used EM-CD algorithm.
-The main difference between proposed EM-CD algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior.
-In comparison, the proposed prior is pairwise independent requiring no special treatment.
 A specific concern of EM-CD algorithm is that it provides no information for inference.
-In contrast, the EM-IWLS can extract the variance-covariance matrix thanks to the IWLS algorithm and construct Wald type inference.
+In contrast, the EM-IWLS can extract the variance-covariance matrix and be used to construct Wald type inference.
 Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
+To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
+The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
 
@@
 It is commonly used when some necessary data to establish the likelihood function are missing.
-Instead, the algorithm maximizes the expectation, with respect to the "missing" data, of the likelihood function.
+Instead of maximizing the the likelihood function, the algorithm maximizes the expectation of the likelihood function with respect to the "missing" data.
 
@@
 
-* E-step: to calculate the expectation of the likelihood function conditioning on some "missing" data
-* M-step: to maximize the expected likelihood function to calculate the parameter of interest
+* E-step: to calculate the expectation of the posterior density function with respect to some "missing" data
+* M-step: to maximize the expectation derived in the E-step to calculate the parameter of interest
 
-For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data.
 Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
@@
 \log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
+&= \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{j})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j),
 \end{aligned}
 $$  
-For simplicity, we consider the linear space to be 1-dimension, which holds true for many smoothing functions.
+where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(1,1) density.
 
-<!-- , and rewrite $\beta^0_j$ as $\beta_{j0}$, and hence removing the superscript $^{*}$ from the coefficients of the the penalized space.
-Using the new notation, the part component $\log p(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{*}_{jk}|\gamma^{*}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
+<!-- , and rewrite $\beta_j$ as $\beta_{j0}$, and hence removing the superscript $^{*}$ from the coefficients of the the penalized space.
+Using the new notation, the part component $\log p(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{*}_{jk}|\gamma^{*}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
 Without loos of generality, the model still work for the spline function whose dimensionality of the null space is more than 1.
@@
 In the M-step, we find the $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
+The E- and M- steps are iterated until the algorithm converge.
+
 To note here, the log-posterior density of BHAMs (up to additive constants) can be written as a two-part equation
-
 $$ \log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) = Q_1(\bs \beta, \phi) + Q_2 (\bs \gamma,\bs \theta),$$
-Where
-$$ Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{jk})\right]$$
-and $$
-Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .$$
-$Q_1$ and $Q_2$ are, respectively, the log-likelihoods of model coefficients and the dispersion parameter $\bs \beta, \phi$ and latent indicators $\bs \gamma, \theta$.
-Conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
-
-The E- and M- steps are iterated until the algorithm converge.
-Depending on the choice of the mixture priors, the maximization in the M-step can take different algorithms, CD for fast computation and IWLS for uncertainty measures.
+where
+$$ Q_1\equiv Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{jk})\right]$$ and
+$$Q_2 \equiv Q_2(\bs\gamma,\bs\theta) = \sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{\tp})\log \theta_j + (2-\gamma_j-\gamma_{j}^{\tp}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j).$$
+$Q_1$ and $Q_2$ are respectively the log posterior density of the coefficients $\bs \beta$ and the log posterior density of the probability parameters $\bs \theta$ conditioning on $\bs \gamma$.
+Meanwhile, conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
+Depending on the choice of coefficient priors, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via CD algorithm or IWLS algorithm in each iteration.
+Maximization of $E(Q_2)$ can be solved via closed form equations following the beta-binomial conjugate relationship.
 
 ### EM-Coordinate Descent
+\label{sec:EMCD}
 
@@
 \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{\tp}}^{-1}_{j})|\beta_{jk}|\right],
 \end{equation}
-where $S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
-To calculate two unknown quantities $E({S_j^0}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p^0_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j} \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
-The calculation of $p_j$ is slightly different from that of $p^0_j$, as $p_j$ depends on the value of $\bs \beta_{j}$ and $p^0_j$ only depends on $\beta_j^0$.
+where $S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S^\tp_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
+To calculate two unknown quantities $E({S_j}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
+The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the value of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
 The calculation follows the equations below.
-
-$$
-\begin{aligned}
-p_{j}^0 &= \frac{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\pr(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \pr(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) }{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) + \pr(\gamma^{\tp}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=0, s_0)}
-\end{aligned}
-$$
-
+\begin{align*}
+p_{j} &= \frac{\pr(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\pr(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \pr(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
+p^\tp_{j} &= \frac{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) }{\pr(\gamma^{\tp}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=1, s_1) + \pr(\gamma^{\tp}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{\tp}_{j}=0, s_0)}
+\end{align*}
 and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = \pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
@@
 \begin{align}\label{eq:exp_scale}
-&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{\tp}_{j}) = p_{j}\nonumber\\
-&E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
+&E(\gamma_{j})  = p_{j} & &E(\gamma^{\tp}_{j}) = p_{j}^{\tp}\nonumber\\
+&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^\tp}^{-1}_{j}) = \frac{1-p_{j}^{\tp}}{s_0} + \frac{p_{j}^{\tp}}{s_1}.
 \end{align}
-
-As the formulation of $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function, it can be optimized via coordinate descent algorithm.
-After the $\bs \beta$ are updated with the coordinate descent algorithm, we update the hypor parameter $\bs \theta$.
-The remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
-As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed form equation:
-
+After replacing withe calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm [@Friedman2010].
+Independently, the remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
+As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed form equation,
 \begin{equation}\label{eq:update_theta}
-\theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
+\theta_j = \frac{p_j + p^\tp_{j} + a - 1 }{a + b}.
 \end{equation}
 
-Totally, the framework of the proposed EM-coordinate descent algorithm was summarized as follows:
+Totally, the proposed EM-coordinate descent algorithm is summarized as follows:
 
@@
 
-E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{\tp}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration
+    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E({S}^{-1}_{j})$, $E({S^\tp}^{-1}_{j})$ with estimates of $\Theta^{(t-1)}$ from previous iteration
 
@@
 
-a) Update $\bs \beta^{(t)}$ and the dispersion parameter $\phi$, if exists, using the coordinate descent algorithm with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
+    a) Update $\bs \beta^{(t)}$, and the dispersion parameter $\phi^{(t)}$ if exists, using the coordinate descent algorithm with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
 
@@
 
-A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2 \sim Gamma(1, 1/(2S^2))$ integrated out:
+A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
 \begin{align*}
-  \beta &\sim N(0, \tau^2)\\
-  \tau^2 & \sim Gamma(1, 1/(2S^2)), 
+  \beta|\tau^2 &\sim N(0, \tau^2)\\
+  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
 \end{align*}
 For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
-The change in the prior formulation in turns leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyper prior of $\tau^2$:
+The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
 \begin{equation}\label{eq:Q1_IWLS}
-Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|{\tau^0}^2_j) + \log f({\tau^0}^2_j| S_j^0)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
+Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
 \end{equation}
@@
 To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
-$E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
-As neither $E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
+$E({S_j}^{-1}|\beta_j, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
 On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
-$E({\tau^0}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
+$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
 \begin{align*}
- E({\tau^0}^{-2}_{j}|S_j^0, \beta^0_j) ={S_j^0}^{-1}/|\beta^0_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
+ E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
 \end{align*}
-where $S_j^0$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
+where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
 \begin{equation}\label{eq:EQ1_IWLS}
-E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)}{2\phi E{\tau_j^0}^{2}} +\sum\limits_{k=1}^{K_j} \frac{{\beta_{jk}^\tp}^2}{2\phi E({\tau_{jk}^\tp}^{2})}\right].
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
 \end{equation}
-$\frac{\beta^2}{2\phi E(\tau^{2})}$ can be seen as the kernel of a normal density with mean 0 and variance $\phi E(\tau^{2})$.
+<!-- $E(Q_1)$ can be seen as a $l_2$ penalized likelihood function where the regularization parameter $\lambda = 2E({\tau}^{-2})$. Equivalently, we have  -->
+$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$, and we can formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal,
+$$
+\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
+$$
 
@@
 $$
-  f(\textbf{y}|\bs \beta, \phi) \approx MVN(\textbf{z}|X\beta, \Sigma w_i^{-1}\phi)
+f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
 $$
-where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively.The pseudo data and the pseudo-weight are calculated by
+where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. The pseudo-response and the pseudo-weight are calculated by
 $$
@@
 $$
-where $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$.
-With $z\sim N(X\beta, \Sigma\phi)$ and $\beta \sim N(0, \phi\Sigma_{\tau^2})$, we can augment the two normal distributions and derive the estimates for $\bs \beta$ and $\phi$ in each step in closed form equations.
-To note, an estimate of the variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences.
+where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
+
+With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
+We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*}
+& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
+  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
+  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
+\end{align*}
+ such that 
+$$
+\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
+$$
+Using the least squares estimators to update $\bs\beta$ and $\phi$, we have 
+\begin{align*}
+& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
+\end{align*}
+To note, the variance-covariance matrixof the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
+$$
+  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
+$$
+The rest of the algorithm remains the same as described in Section \ref{sec:EMCD}.
 
 <!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
-The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S^0_j}^{-1}){\beta^0_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S^0}_1^{-1}), E(S_1^{-1}), \dots, E({S_p^0}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S^0_1}^{-1}), \dots E({S^0_p}^{-1})$.
+The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S_j}^{-1}){\beta_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S}_1^{-1}), E(S_1^{-1}), \dots, E({S_p}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S_1}^{-1}), \dots E({S_p}^{-1})$.
 An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.
@@
 
-<!-- E-step: calculate $E(\gamma^0_{j})$, $E(\gamma^{pen}_{j})$and $E({S^0}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
+<!-- E-step: calculate $E(\gamma_{j})$, $E(\gamma^{pen}_{j})$and $E({S}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
 
@@
 Increasing the spike scale $s_0$ tends to include more non-zero coefficients in the model.
-An measure of preference calculated with cross-validations, e.g.
+A measure of preference calculated with cross-validations, e.g.
 deviance, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
-The procedure is similar to the lasso implemented in the widely used R package `glmnet`, which quickly fits the lasso model over a list of values of $\lambda$ covering the entire range, giving a sequence of models for users to choose from.
+The procedure is similar to the Lasso implemented in the widely used R package `glmnet`, which quickly fits Lasso models over a list of values of regularization parameters $\lambda$ , giving a sequence of models for users to choose from.

### 2021-09-26 Manuscript/SS_GAM.tex

@@
 \abstract{There are proposals that extend the classical generalized
-additive models (GAMs) to accommodate the high-dimensional
-(p\textgreater\textgreater n) problem using group penalties and priors.
-These models rarely smooth the nonlinear effect properly because of the
-sparse regularization. It leads to excess shrinkage on coefficients and
-inaccurate prediction. Besides, most of these GAMs consider an
-``all-in-all-out'' approach for function selection, rendering them
-difficult to answer if nonlinear effects are necessary. While some
-Bayesian models can address the shortcomings, using Markov chain Monte
-Carlo (MCMC) algorithms for model fitting creates a new challenge,
-scalability. Hence, we propose Bayesian hierarchical generalized
-additive models as a solution. For smooth and sparse solutions, the
-proposed models employ bi-part spike-and-slab spline priors, including
-mixture double exponential and mixture normal priors. Two
-optimization-based algorithms, EM-Coordinate Descent and EM-Iterative
-Weighted Least Squares, are developed for expedited computation.
-Simulation studies and metabolomics data analyses demonstrate improved
-performance against state-of-art models, mgcv, COSSO and sparse Bayesian
-GAM. The software implementation of the proposed models is freely
-available via an R package BHAM.}
+additive models (GAMs) to accommodate the high-dimensional (\(p>>n\))
+problem using group penalties and priors. These models rarely smooth the
+nonlinear effect properly because of the sparse regularization. It leads
+to excess shrinkage on coefficients and inaccurate prediction. Besides,
+most of these GAMs consider an ``all-in-all-out'' approach for function
+selection, rendering them difficult to answer if nonlinear effects are
+necessary. While some Bayesian models can address the shortcomings,
+using Markov chain Monte Carlo (MCMC) algorithms for model fitting
+creates a new challenge, scalability. Hence, we propose Bayesian
+hierarchical generalized additive models as a solution. For smooth and
+sparse solutions, the proposed models employ bi-part spike-and-slab
+spline priors, including mixture double exponential and mixture normal
+priors. Two optimization-based algorithms, EM-Coordinate Descent and
+EM-Iterative Weighted Least Squares, are developed for expedited
+computation. Simulation studies and metabolomics data analyses
+demonstrate improved performance against state-of-art models, mgcv,
+COSSO and sparse Bayesian GAM. The software implementation of the
+proposed models is freely available via an R package BHAM.}
 
@@
 exponential family distribution \(EF(\mu, \phi)\) with mean \(\mu\) and
-dispersion \(\phi\), can be expressed as the summation of smoothing
-functions, \(B_j, j = 1, \dots, p\), of a given \(p\)-dimensional vector
-of covariates \(\boldsymbol{x}\), written as \[
+dispersion \(\phi\), can be modeled with the summation of smoothing
+functions, \(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional
+vector of covariates \(\boldsymbol{x}\), written as \[
  E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
-\] where \(g^{-1}\) is the inverse of a monotonic link function. The
-smoothing functions can take many forms and are estimated using a
+\] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
+The smoothing functions can take many forms and are estimated using a
 pseudo-weighted version of the backfitting algorithm
@@
 mathematically as \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[f_j(x)\right] = \lambda_j \int f^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
 \end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
@@
 penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of smoothing functions, also
-know as the null space, from the nonlinear space. We then impose a new
-bi-part spike-and-slab spline prior on the smoothing functions for
-bi-level selection such that the linear and nonlinear spaces of
-smoothing functions can be selected separately. The prior setup
-encourages a flexible solution, rending one of three possibilities: no
-effect, only linear effect, or nonlinear effect of a predictor. In
-addition, two fast computing optimization-based algorithms are developed
-for high and ultrahigh-dimensional\citep{Fan2009} settings respectively.
+allows the separation of the linear space of smoothing functions from
+the nonlinear space. We then impose a new bi-part spike-and-slab spline
+prior on the smoothing functions for bi-level selection such that the
+linear and nonlinear spaces of smoothing functions can be selected
+separately. The prior setup encourages a flexible solution, rending one
+of three possibilities: no effect, only linear effect, or nonlinear
+effect of a predictor. In addition, two scalable optimization-based
+algorithms are developed for high and
+ultrahigh-dimensional\citep{Fan2009} settings respectively.
 
@@
 (EM-IWLS) algorithm for different utilities. The applications of the two
-aforementioned algorithm are fruitful in Bayesian hierarchical models in
-high-dimensional data analysis.
-\citep{Yi2012, Rockova2014a, Rockova2018} Even though SB-GAM with a
-group spike-and-slab lasso prior also uses EM-CD for model fitting, the
-proposed EM-IWLS provides uncertainty measures that could be further
-used for confidence bound calculation and hypothesis testing. EM-IWLS
-complements the lack of inference potential from EM-CD in a median and
-high dimensional setting. Furthermore, the proposed model addresses the
-incapability of bi-level selection in SB-GAM. Lastly, the R
-implementation of the Bayesian hierarchical additive model is available
-publicly via BHAM at \url{https://github.com/boyiguo1/BHAM}, make
-transnational science more accessible.
+aforementioned algorithms are fruitful in Bayesian hierarchical models
+for high-dimensional data. \citep{Yi2012, Rockova2014a, Rockova2018}
+Even though SB-GAM with a group spike-and-slab lasso prior also uses
+EM-CD for model fitting, the proposed EM-IWLS provides uncertainty
+measures that could be further used for confidence bound calculation and
+hypothesis testing. EM-IWLS complements the lack of inference potential
+from EM-CD in a median and high dimensional setting. Furthermore, the
+proposed model addresses the incapability of bi-level selection in
+SB-GAM. Lastly, the R implementation of the Bayesian hierarchical
+additive model is available publicly via BHAM at
+\url{https://github.com/boyiguo1/BHAM}, making transnational science
+more accessible.
 
@@
 \end{equation} and the data distribution is \begin{equation}
-f(y|X\beta, \phi) = \prod\limits^n_{i=1}f(y_i|X_i,\phi),\nonumber
+f(\boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f(y_i|\boldsymbol{\beta}, \phi),\nonumber
 \end{equation} with smoothing functions \(B_j(x_j)\) of the variable
@@
 denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\) denotes the
-basis function coefficients for the \(j\)th variable, and the matrix
-transposing operation is denoted with a superscript \(^T\). To note, the
-proposed model can include parametric forms of variables in the model,
-treating the parametric function as a special case of the smoothing
-function, e.g.~\(B_j(x_j) = x_j\) with the smoothing penalty matrix
-defined as \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
+basis function coefficients for the \(j\)th variable such that
+\(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
+abuse of notation, we denote vectors and matrices in bold fonts
+\(\boldsymbol{\beta}, \boldsymbol{X}\) with conformable dimensions,
+where scalar and random variables are denoted in with unbold fonts
+\(\beta, X\). The matrix transposing operation is denoted with a
+superscript \(^T\). To note, the proposed model can include parametric
+forms of variables in the model, treating the parametric function as a
+special case of the smoothing function, e.g.~\(B_j(x_j) = x_j\) with the
+smoothing penalty matrix defined as
+\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
 
@@
 set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}),
-in the prior density function, or least squared and likelihood function
-in the penalized framework. However, the direct integration of smoothing
-penalties into sparsity penalties is not obvious. Marra and Wood
+in the prior density function. However, the direct integration of
+smoothing penalty with sparsity penalty is not obvious. Marra and Wood
 \citep{Marra2011} proposed a re-parameterization procedure to
 accommodate the smoothing penalty implicitly. Given the smoothing
-penalty matrix is most often symmetric and positive semi-definite for
-univariate smoothing functions, we apply eigendecomposition on the
-penalty matrix
+penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
+semi-definite for univariate smoothing functions, we apply
+eigendecomposition on the penalty matrix
 \(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
@@
 arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
-\(0\) elements on the diagonal, associated with the linear space of the
-smoothing function. We further write the orthonormal matrix
-\(\boldsymbol{U} \equiv \begin{bmatrix}\boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
-containing the eigenvectors as columns in the corresponding order. That
-is \(\boldsymbol{U}^0\) contains the eigenvectors with zero eigenvalues
-for the linear space and \(\boldsymbol{U}^{*}\) contains the
-eigenvectors (as columns) for the non-zero eigenvalues, i.e.~the
-non-linear space. We multiply the basis function matrix
-\(\boldsymbol{X}\) with the orthonormal matrix \(\boldsymbol{U}\) for
-the new design matrix
-\(\boldsymbol{X}^\ast = \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{X}^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
+elements of zeros on the diagonal, where the zeros are associated with
+the linear space of the smoothing function. For the most popular
+smoothing function, cubic splines, the dimension of the linear space is
+one. Hereafter, we focus on discussing a uni-dimensional linear space
+for simplicity; however, it generalizes easily to the cases where the
+linear space is multidimensional. We further write the orthonormal
+matrix
+\(\boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
+containing the eigenvectors as columns in the corresponding order to
+\(\boldsymbol{D}\). That is, \(\boldsymbol{U}\) contains the
+eigenvectors \(U^0\) with zero eigenvalues for the linear space and
+\(\boldsymbol{U}^{*}\) contains the eigenvectors (as columns) for the
+non-zero eigenvalues, i.e.~the non-linear space. We multiply the basis
+function matrix \(\boldsymbol{X}\) with the orthonormal matrix
+\(\boldsymbol{U}\) for the new design matrix
+\({\boldsymbol{X}}^\text{repa}= \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} X^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
 An additional scaling step is imposed on \(\boldsymbol{X}^{*}\) by the
@@
 function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
-each of its dimensions. With slight abuse of the notation, we denote
-\(\boldsymbol{X}_j \equiv \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
+each of its dimensions. With slight abuse of the notation, we drop the
+superscript \(^\text{repa}\) and denote
+\(\boldsymbol{X}_j \equiv \begin{bmatrix} X_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
 as the basis function matrix for the \(j\)th variable after the
@@
 form \[
-B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
+B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \beta_j X^0_j + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
 \] and the generalized additive model in Equation (\ref{eq:gam}) be
 \begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) =  g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\boldsymbol{\beta_j^0}^T \boldsymbol{X}_j^0 + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right].
-\end{equation}
+E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
+\end{equation} where the coefficients
+\(\boldsymbol{\beta}_j \equiv \begin{bmatrix} \beta_j : \boldsymbol{\beta}^*_j \end{bmatrix}\)
+is an augmentation of the coefficient scalar \(\beta_j\) of linear space
+and the coefficient vector \(\boldsymbol{\beta}^*_j\) of non-linear
+space.
 
@@
 
-The major criticism of early spike-and-slab priors receive is being
+The major criticism of early spike-and-slab models is being
 computationally prohibitive. \citep{Bai2021Review} Since then, many
@@
 algorithms bear, which include EMVS \citep{Rockova2014a} and
-Spike-and-slab Lasso \citep{Rockova2018b, Rockova2018}. Particularly,
-the development of Spike-and-slab Lasso substantially improves the
+spike-and-slab Lasso \citep{Rockova2018b, Rockova2018}. Particularly,
+the development of spike-and-slab Lasso substantially improves the
 scalability of SS models, setting up the theoretical foundation for
@@
 \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The spike-and-slab
-lasso prior is composed of two double exponential distributions with
-mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
+Lasso (SSL) prior is composed of two double exponential distributions
+with mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
 mathematically, \begin{equation} 
@@
 0 or 1, the mixture double exponential distribution can be formulated as
-a double exponential density, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, (1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1), 0 < s_0 < s_1,
+one single double exponential density, \begin{equation} \label{eq:ssl}
+\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
 \end{equation} with the scale parameter
@@
 (\ref{eq:ssl}). Given the re-parameterized deign matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} \boldsymbol{X}_j^0 : \boldsymbol{X}_j^*\end{bmatrix}\)
+\(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
 for the \(j\)th variable, we impose a two-part SSL prior to the
-coefficients \(\boldsymbol{\beta}_j\). Specifically, we impose
-independent group priors on the linear space coefficients and on the
-nonlinear space coefficients respectively,
-
+coefficients
+\(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
+Specifically, we impose independent group priors on the linear space
+coefficients and on the nonlinear space coefficients respectively,
 \begin{align}\label{eq:bham_ssl}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim DE(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1) \nonumber \\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\sim DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), 
-\end{align}
-
-where \(\gamma^0_{j}\in\{0,1\}\) and \(\gamma^*_{j}\in \{0,1\}^{K_j}\)
-are latent indicator variable and vectors, indicating if the model
-includes the linear effect and the nonlinear effect of the \(j\)th
-variable respectively. \(s_0\) and \(s_1\) are scale parameters,
-assuming \(0 < s_0 < s_1\) and given. These scale parameters \(s_0\) and
-\(s_1\) can be treated as tuning parameters and optimized via
-cross-validation. A discussion of how to choose the scale parameters
-comes in Section \ref{sec:tune}. To note, this prior differs from
-previous group spike-and-slab lasso priors \citep{Tang2018, Tang2019},
-where the \(\boldsymbol{\beta}^0_j\) and \(\boldsymbol{\beta}^*_j\)
-shares the same indicator variables \(\gamma_j^0\), \(\gamma_j^*\)
-respectively. It is possible to add a more restrictive assumption on the
-priors, assuming that one indicator variable decides the inclusion of
-both the linear effect and nonlinear effect,
-i.e.~\(\gamma_j = \gamma^0_j = \gamma^*_j\). This converges to the
-SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
-assumption such that each coefficient \(\beta\) to have its own latent
-indicator \(\gamma\), but at the cost of complicating the bi-level
-function selection. This reduces the proposed prior to the classic
-spike-and-slab lasso prior.
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j
+\end{align} where \(\gamma_{j}\in\{0,1\}\) and
+\(\gamma^*_{j}\in \{0,1\}\) are two latent indicator variables,
+indicating if the model includes the linear effect and the nonlinear
+effect of the \(j\)th variable respectively. \(s_0\) and \(s_1\) are
+scale parameters, assuming \(0 < s_0 < s_1\) and given. These scale
+parameters \(s_0\) and \(s_1\) can be treated as tuning parameters and
+optimized via cross-validation. A discussion of how to choose the scale
+parameters comes in Section \ref{sec:tune}. To note, this prior differs
+from previous group spike-and-slab lasso priors
+\citep{Tang2018, Tang2019}, as the \(\beta_j\) and
+\(\boldsymbol{\beta}^*_j\) have different indicator variables
+\(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to add a more
+restrictive assumption on the priors, assuming that one indicator
+variable decides the inclusion of both the linear effect and nonlinear
+effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the SB-GAM
+\citep{Bai2021}. Conversely, it is also possible to relax the assumption
+such that each coefficient \(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has
+its own latent indicator \(\gamma_{jk}\), but at the cost of
+complicating the bi-level function selection. This reduces the proposed
+prior to the classic spike-and-slab lasso prior.
 
@@
 The rest of the hierarchical prior follows the traditional
-spike-and-slab lasso prior: we set up hyper-priors of
-\(\boldsymbol{\gamma}\) to allow local adaption of the shrinkage using a
-Bernoulli distribution (binomial distribution of one trial). The two
-indicators of the \(j\)th predictor, \(\gamma^{0}_j\) and
+spike-and-slab lasso prior: we set up hyper-priors on
+\(\gamma_j, \gamma^*_j\) to allow local adaption of the shrinkage using
+a Bernoulli distribution, written as binomial distribution of one trial.
+The two indicators of the \(j\)th predictor, \(\gamma_j\) and
 \(\gamma^*_j\), shares the same probability parameter \(\theta_j\), \[
 \begin{aligned}
-\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
-\gamma_{j}^*| \theta_j &\sim Bin(\gamma^*_{j}|1, \theta_j).
+&\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
+&\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j).
 \end{aligned}
@@
 different penalty on the linear space and non-linear space of the
-smoothing function. The hyper prior of \(\gamma_{j}^{0}\) decides the
+smoothing function. The hyper prior of \(\gamma_{j}\) decides the
 sparsity of the model at the function selection level, while that of
 \(\gamma_{j}^*\) decides the smoothness of the spline function at basis
-function level. Meanwhile, we specify that \(\gamma_{j}^0\) and
+function level. Meanwhile, we specify that \(\gamma_{j}\) and
 \(\gamma_{j}^*\) are independently distributed for analytic simplicity.
@@
 simplicity, we focus on a special case of beta distribution, uniform
-(0,1). When the variable have large effects in any of the bases, the
-parameter \(\theta_j\) will be estimated large, which in turn encourages
-the model to include the rest of bases, achieving the local adaption
-among spline bases. Hereafter, we refer Bayesian hierarchical
-generalized additive models with the spike-and-slab lasso prior as the
-BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
+(0,1), i.e.~\(a = 1, b = 1\). When the variable have large effects in
+any of the bases, the parameter \(\theta_j\) will be estimated large,
+which in turn encourages the model to include the rest of bases,
+achieving the local adaption among spline bases. Hereafter, we refer
+Bayesian hierarchical generalized additive models with the
+spike-and-slab lasso prior as the BHAM-SSL, and visually presented in
+Figure \ref{fig:SSprior}.
 
@@
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta^0_1, \boldsymbol{\beta}^*_1, \dots,\beta^0_j, \boldsymbol{\beta}^*_j , \dots,\beta^0_p, \boldsymbol{\beta}^*_p) $};
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
 \node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
@@
 
-With the re-parameterization step of the basis function matrix \(X\), it
-is possible to generalized the SSL prior to other priors, for example
-normal priors for ridge-type regularization and mixture normal prior for
-spike-and-slab regularization. These priors would work better in low and
-medium dimensional settings where the sparse assumption is not
-necessary. Here we focus on the mixture normal prior as a demonstration
-of applying continuous spike-and-slab prior in additive models.
-
-A spike-and-slab mixture normal prior in additive models can be
-expressed as \[
-\begin{aligned}
-  \beta^0_{j} |\gamma^0_{j},s_0,s_1 &\sim N(0,(1-\gamma^0_{j}) s_0 + \gamma^0_{j} s_1)\nonumber\\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\sim N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1).
-\end{aligned}
-\] Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}),
-\(0 < s_0 < s_1\) are tuning parameters and can be optimized via
-cross-validation. One of the critics received by the spike-and-slab
-mixture normal prior is that the tails of a normal distribution
-diminishes to zero too fast, which renders problem when estimating the
-large effects. Distributions with heavier tails can be used as an
-alternative, for example mixture \(t\) distribution with small degree of
-freedom.
+With the re-parameterization step of the basis function matrix
+\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
+priors, for example normal priors for ridge-type regularization and
+mixture normal prior for spike-and-slab regularization. These priors
+would work better in low and medium dimensional settings where the
+sparse assumption is not necessary. Here we elaborate the mixture normal
+prior as a demonstration of applying continuous spike-and-slab prior in
+BHAM.
+
+A spike-and-slab mixture normal spline prior can be expressed as
+\begin{align*}
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
+\end{align*} Similar to the BHAM-SSL prior in Equation
+(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
+optimized via cross-validation. One of the critics received by the
+spike-and-slab mixture normal prior is that the tails of a normal
+distribution diminishes to zero too fast, which causes problems when
+estimating the large effects. Distributions with heavier tails can be
+used as an alternative, for example mixture Student's \(t\) distribution
+with small degree of freedom.
 
@@
 George and McCulloch\citep{George1997} examined the computation speed
-for various MCMC algorithms with continous spike-and-slab priors, and
-suggested MCMC algorithms works well for medium size (\(p\)=25) of
-predictors with only linear effects . However, it is not feasible to
+for various MCMC algorithms with spike-and-slab mixture normal priors,
+and suggested MCMC algorithms works well for medium size (\(p\)=25) of
+predictors with only linear effects. However, it is not feasible for
 high-dimensional data analysis where the number of predictors easily
@@
 the computational demands of a MCMC algorithm for fitting spike-and-slab
-additive models grow exponentially as \(p\) increases modestly via monte
-carlo simulation studies. Hence, we feel compelled to develop scalable
-algorithms for fitting Bayesian hierarchical additive models in
-high-dimensional settings.
+GAM grow exponentially as \(p\) increases modestly via simulation
+studies. Hence, we feel compelled to develop scalable algorithms for
+fitting Bayesian hierarchical additive models in high-dimensional
+settings.
 
@@
 other parameters; in the M step, a ridge estimator was used to update
-the coefficients, and \(\sigma^2, \theta\) are updated accordingly. The
-same authors \citep{Rockova2018b, Rockova2018} further combined the EM
-algorithm with coordinate descent algorithm for fitting SSL models. Yi
-and his group independently developed the EM-Iterative Weighted Least
-Square and EM-cyclic coordinate descent algorithms to fit models with
-broader class of priors, SSL included.\citep{Yi2019} These algorithms
-were implemented for generalized linear models\citep{Tang2017a}, Cox
+the coefficients, followed by updating \(\phi, \theta\). The same
+authors \citep{Rockova2018b, Rockova2018} further combined the EM
+algorithm with coordinate descent algorithm to fit SSL models. Yi and
+his group independently developed the EM-Iterative Weighted Least Square
+and EM-cyclic coordinate descent algorithms to fit models with broader
+class of priors, SSL included.\citep{Yi2019} These algorithms were
+implemented for generalized linear models\citep{Tang2017a}, Cox
 proportional hazards models \citep{Tang2017} and their grouped
 counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a more popular property for
+deterministic solutions, which becomes a popular property for
 reproducible research
 
-In the follow section, we expand the two EM-based algorithms, EM-CD and
+In this section, we extend the two EM-based algorithms, EM-CD and
 EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
@@
 fitting BHAM-SSL with an expedited performance, recommending to use in
-high and ultra-high dimensional setting. SB-GAM\citep{Bai2020, Bai2021}
-also used EM-CD algorithm. The main difference between proposed EM-CD
-algorithm and SB-GAM is that SB-GAM uses a blocked CD algorithm for
-their group prior. In comparison, the proposed prior is pairwise
-independent requiring no special treatment. A specific concern of EM-CD
+high and ultra-high dimensional setting. A specific concern of EM-CD
 algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can extract the variance-covariance matrix thanks to the
-IWLS algorithm and construct Wald type inference. Moreover, the EM-IWLS
-is a more general model fitting algorithm that can be used for fitting
-not only SSL and continuous SS priors but also Student's t-priors and
-double exponential priors.
+the EM-IWLS can extract the variance-covariance matrix and be used to
+construct Wald type inference. Moreover, the EM-IWLS is a more general
+model fitting algorithm that can be used for fitting not only SSL and
+continuous SS priors but also Student's t-priors and double exponential
+priors. To note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD
+algorithm. The main difference between proposed EM-CD algorithm and that
+in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group
+prior, while the proposed prior is pairwise independent requiring no
+special treatment in the CD algorithm.
 
@@
 maximum likelihood estimates. It is commonly used when some necessary
-data to establish the likelihood function are missing. Instead, the
-algorithm maximizes the expectation, with respect to the ``missing''
-data, of the likelihood function.
+data to establish the likelihood function are missing. Instead of
+maximizing the the likelihood function, the algorithm maximizes the
+expectation of the likelihood function with respect to the ``missing''
+data.
 
@@
 \item
-  E-step: to calculate the expectation of the likelihood function
-  conditioning on some ``missing'' data
+  E-step: to calculate the expectation of the posterior density function
+  with respect to some ``missing'' data
 \item
-  M-step: to maximize the expected likelihood function to calculate the
-  parameter of interest
+  M-step: to maximize the expectation derived in the E-step to calculate
+  the parameter of interest
 \end{itemize}
@@
 consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model. Our objective is to find the
-parameters \(\Theta\) that maximize the posterior density function, or
-equivalently, the logarithm of the density function, \[
+nuisance parameters of the model, in other words the ``missing'' data.
+Our objective is to find the parameters \(\Theta\) that maximize the
+posterior density function, or equivalently, the logarithm of the
+density function, \[
 \begin{aligned}
@@
 \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j)
+&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{j})\right]\\
+& +\sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j),
 \end{aligned}
 \]\\
-For simplicity, we consider the linear space to be 1-dimension, which
-holds true for many smoothing functions.
+where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
+distribution and \(f(\theta)\) is the Beta(1,1) density.
 
@@
 M-step, we find the \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). To
-note here, the log-posterior density of BHAMs (up to additive constants)
-can be written as a two-part equation
-
+\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The E-
+and M- steps are iterated until the algorithm converge.
+
+To note here, the log-posterior density of BHAMs (up to additive
+constants) can be written as a two-part equation
 \[ \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
-Where
-\[ Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|\gamma^0_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
-and \[
-Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma^0_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma^0_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j) .\]
-\(Q_1\) and \(Q_2\) are, respectively, the log-likelihoods of model
-coefficients and the dispersion parameter \(\boldsymbol{\beta}, \phi\)
-and latent indicators \(\boldsymbol{\gamma}, \theta\). Conditioning on
+where
+\[ Q_1\equiv Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
+and
+\[Q_2 \equiv Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j).\]
+\(Q_1\) and \(Q_2\) are respectively the log posterior density of the
+coefficients \(\boldsymbol{\beta}\) and the log posterior density of the
+probability parameters \(\boldsymbol{\theta}\) conditioning on
+\(\boldsymbol{\gamma}\). Meanwhile, conditioning on
 \(\boldsymbol{\gamma}\), \(Q_1\) and \(Q_2\) are independent and can be
 maximized separately for \(\boldsymbol{\beta}, \phi\) and
-\(\boldsymbol{\theta}\).
-
-The E- and M- steps are iterated until the algorithm converge. Depending
-on the choice of the mixture priors, the maximization in the M-step can
-take different algorithms, CD for fast computation and IWLS for
-uncertainty measures.
+\(\boldsymbol{\theta}\). Depending on the choice of coefficient priors,
+\(Q_1\) can be treated as penalized likelihood function and maximization
+of \(E(Q_1)\) can be solved via CD algorithm or IWLS algorithm in each
+iteration. Maximization of \(E(Q_2)\) can be solved via closed form
+equations following the beta-binomial conjugate relationship.
 
@@
 \subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
+
+\label{sec:EMCD}
 
@@
 penalty \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[E({S^0_j}^{-1})|\beta^0_j|+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})|\beta_{jk}|\right],
+E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
 \end{equation} where
-\(S_{j}^0 = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
-unknown quantities \(E({S_j^0}^{-1})\) and \(E(S^{-1}_j)\), the
-posterior probability
-\(p^0_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
-\(p_{j} \equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are necessary,
-which can be derived via Bayes' theorem. The calculation of \(p_j\) is
-slightly different from that of \(p^0_j\), as \(p_j\) depends on the
-value of \(\boldsymbol{\beta}_{j}\) and \(p^0_j\) only depends on
-\(\beta_j^0\). The calculation follows the equations below.
-
-\[
-\begin{aligned}
-p_{j}^0 &= \frac{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) }{\text{Pr}(\gamma_{j}^0 = 1|\theta_j)f(\beta_{j}^0|\gamma_{j}^0=1, s_1) + \text{Pr}(\gamma_{j}^0 = 0|\theta_j)f(\beta^0_{j}|\gamma^0_{j}=0, s_0)}\\
-p_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{aligned}
-\]
-
-and
+\(S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
+\(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
+unknown quantities \(E({S_j}^{-1})\) and \(E(S^{-1}_j)\), the posterior
+probability \(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\)
+and \(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
+necessary, which can be derived via Bayes' theorem. The calculation of
+\(p_j^*\) is slightly different from that of \(p_j\), as \(p_j^*\)
+depends on the value of the vector \(\boldsymbol{\beta}^*_{j}\) and
+\(p_j\) only depends on the scalar \(\beta_j\). The calculation follows
+the equations below. \begin{align*}
+p_{j} &= \frac{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \text{Pr}(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
+p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
+\end{align*} and
 \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j\),
@@
 show \begin{align}\label{eq:exp_scale}
-&E(\gamma^0_{j})  = p^0_{j} & &E(\gamma^{*}_{j}) = p_{j}\nonumber\\
-&E({S^0}^{-1}_{j}) = \frac{1-p^{0}_{j}}{s_0} + \frac{p^{0}_{j}}{s_1} & &E(S^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1}.
-\end{align}
-
-As the formulation of \(E(Q_1)\) can be seen as a \(l_1\) penalized
-likelihood function, it can be optimized via coordinate descent
-algorithm. After the \(\boldsymbol{\beta}\) are updated with the
-coordinate descent algorithm, we update the hypor parameter
-\(\boldsymbol{\theta}\). The remaining parameters of interest
-\(\boldsymbol{\theta}\) can be updated by maximizing \(E(Q_2)\). As the
-beta distribution is a conjugate prior for Bernoulli distribution,
-\(\boldsymbol{\theta}\) can be easily updated with a closed form
-equation:
-
+&E(\gamma_{j})  = p_{j} & &E(\gamma^{*}_{j}) = p_{j}^{*}\nonumber\\
+&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^*}^{-1}_{j}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
+\end{align} After replacing withe calculated quantities, \(E(Q_1)\) can
+be seen as a \(l_1\) penalized likelihood function with the
+regularization parameter \(\lambda = E(S^{-1})\), and hence be optimized
+via coordinate descent algorithm \citep{Friedman2010}. Independently,
+the remaining parameters of interest \(\boldsymbol{\theta}\) can be
+updated by maximizing \(E(Q_2)\). As the beta distribution is a
+conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
+be easily updated with a closed form equation,
 \begin{equation}\label{eq:update_theta}
-\theta_j = \frac{p^0_j + p_{j} + a - 1 }{a + b}.
+\theta_j = \frac{p_j + p^*_{j} + a - 1 }{a + b}.
 \end{equation}
 
-Totally, the framework of the proposed EM-coordinate descent algorithm
-was summarized as follows:
+Totally, the proposed EM-coordinate descent algorithm is summarized as
+follows:
 
@@
   Iterate over the E-step and M-step until convergence
-\end{enumerate}
-
-E-step: calculate \(E(\gamma^0_{j})\), \(E(\gamma^{*}_{j})\)and
-\(E({S^0}^{-1}_{j})\) , \(E({S}^{-1}_{j})\)with estimates of
+
+  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
+  \(E({S}^{-1}_{j})\), \(E({S^*}^{-1}_{j})\) with estimates of
   \(\Theta^{(t-1)}\) from previous iteration
@@
   \begin{enumerate}
-\def\labelenumi{\alph{enumi})}
+  \def\labelenumii{\alph{enumii})}
   \item
-  Update \(\boldsymbol{\beta}^{(t)}\) and the dispersion parameter
-  \(\phi\), if exists, using the coordinate descent algorithm with the
-  penalized likelihood function in Equation (\ref{eq:Q1_CD})
+    Update \(\boldsymbol{\beta}^{(t)}\), and the dispersion parameter
+    \(\phi^{(t)}\) if exists, using the coordinate descent algorithm
+    with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
   \item
@@
     (\ref{eq:update_theta})
+  \end{enumerate}
 \end{enumerate}
@@
 A double exponential prior, \(\beta|S \sim DE(0, S)\) can be formulated
-as a hierarchical normal prior with unknown variance
-\(\tau^2 \sim Gamma(1, 1/(2S^2))\) integrated out: \begin{align*}
-  \beta &\sim N(0, \tau^2)\\
-  \tau^2 & \sim Gamma(1, 1/(2S^2)), 
+as a hierarchical normal prior with unknown variance \(\tau^2\)
+integrated out: \begin{align*}
+  \beta|\tau^2 &\sim N(0, \tau^2)\\
+  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
 \end{align*} For the mixture double exponential priors, we can define
 the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following
-Equation (\ref{eq:ssl}). The change in the prior formulation in turns
+Equation (\ref{eq:ssl}). The change in the prior formulation in turn
 leads to the change in the log posterior density function, as \(Q_1\)
-needs to account for the hyper prior of \(\tau^2\):
+needs to account for the hyperprior of \(\tau^2\):
 \begin{equation}\label{eq:Q1_IWLS}
-Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) + \sum\limits_{j=1}^p\left[\log f(\beta^0_j|{\tau^0}^2_j) + \log f({\tau^0}^2_j| S_j^0)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
+Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
 \end{equation} Since \(\boldsymbol{\tau}^2\) are not of our primary
@@
 coefficient specific for \(\beta^*_{jk}\).
-\(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau^0}^2_{j}|S_j^0, \beta^0_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
+\(E({S_j}^{-1}|\beta_j, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
 needs to be calculated to formulate \(E(Q_1)\). As neither
-\(E({S_j^0}^{-1}|\beta_j^0, s_0, s_1)\) nor
+\(E({S_j}^{-1}|\beta_j, s_0, s_1)\) nor
 \(E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on \(\tau^2\)s,
@@
 posterior density of \(\tau^{-2}\) is an inverse Gaussian distribution.
-\(E({\tau^0}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
+\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
 using the closed form equation \begin{align*}
- E({\tau^0}^{-2}_{j}|S_j^0, \beta^0_j) ={S_j^0}^{-1}/|\beta^0_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
-\end{align*} where \(S_j^0\) and \(S_j^*\) are replaced by the
-expectation and \(\beta\)s are replaced with \(\beta^{(t-1)}\). With
-simplification (up to constant additive terms), we have
+ E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
+\end{align*} where \(S_j\) and \(S_j^*\) are replaced by the expectation
+and \(\beta\)s are replaced with \(\beta^{(t-1)}\). With simplification
+(up to constant additive terms), we have
 \begin{equation}\label{eq:EQ1_IWLS}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \log f(\phi) - \sum\limits_{j=1}^p\left[ \frac{({\beta_j^0}^2)}{2\phi E{\tau_j^0}^{2}} +\sum\limits_{k=1}^{K_j} \frac{{\beta_{jk}^*}^2}{2\phi E({\tau_{jk}^*}^{2})}\right].
-\end{equation} \(\frac{\beta^2}{2\phi E(\tau^{2})}\) can be seen as the
-kernel of a normal density with mean 0 and variance
-\(\phi E(\tau^{2})\).
+E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^*}^{-2})}{\beta_{jk}^*}^2\right].
+\end{equation} \(2E({\tau}^{-2})\beta^2\) can be seen as the kernel of a
+normal density with mean 0 and variance \(E(\tau^{2})\), and we can
+formulate the coefficients \(\boldsymbol{\beta}\) as a multivariate
+normal distribution with means \(\boldsymbol{0}\) and variance
+covariance matrix \(\boldsymbol{\Sigma}_{\tau^2}\), where
+\(\boldsymbol{\Sigma}_{\tau^2}\) is a diagonal matrix with
+\(E(\tau^2)\)s on the diagonal, \[
+\boldsymbol{\beta }\sim \text{MVN}(0, \boldsymbol{\Sigma}_{\tau^2}).
+\]
 
@@
 likelihood: \[
-  f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx MVN(\textbf{z}|X\beta, \Sigma w_i^{-1}\phi)
+f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx \text{MVN}(\textbf{z}|\boldsymbol{X} \boldsymbol{\beta}, \phi\boldsymbol{\Sigma })
 \] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
-the pseudo-response and pseudo-weight respectively.The pseudo data and
-the pseudo-weight are calculated by \[
+the pseudo-response and pseudo-weight respectively. The pseudo-response
+and the pseudo-weight are calculated by \[
 \begin{aligned}
@@
 \end{aligned}
-\] where \(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
+\] where \(\hat\eta_i = (\boldsymbol{X} {\hat{\boldsymbol{\beta}}})_i\),
+\(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
 \(L^{''}(y_i|\hat\eta_i, \hat \phi)\) are the first and second
 derivative of the log density,
-\(\log f(\textbf{y}_i|\boldsymbol{\beta}, \phi)\). With
-\(z\sim N(X\beta, \Sigma\phi)\) and
-\(\beta \sim N(0, \phi\Sigma_{\tau^2})\), we can augment the two normal
-distributions and derive the estimates for \(\boldsymbol{\beta}\) and
-\(\phi\) in each step in closed form equations. To note, an estimate of
-the variance-covariance matrix can be derived in the EM-IWLS algorithm
-and in turn can be used for statistical inferences.
+\(\log f(\textbf{y}_i|\boldsymbol{\beta}, \phi)\) with respect to
+\(\eta_i\).
+
+With
+\(\boldsymbol{z}\sim \text{MVN}(\boldsymbol{X} \boldsymbol{\beta}, \phi \boldsymbol{\Sigma})\)
+and
+\(\boldsymbol{\beta }\sim \text{MVN}(0, \phi \boldsymbol{\Sigma}_{\tau^2})\),
+we can augment the two multivariate normal distributions and update the
+estimates for \(\boldsymbol{\beta}\) and \(\phi\) via least squares in
+each iteration of the EM algorithm. We create the augmented response,
+augmented data, and augmented variance-covariance matrix following
+\begin{align*}
+& \boldsymbol{z}_* = \begin{bmatrix} \boldsymbol{z}\\ \boldsymbol{0}\end{bmatrix} &&
+  \boldsymbol{X}_* = \begin{bmatrix} \boldsymbol{X} \\ \boldsymbol{I} \end{bmatrix} &&
+  \boldsymbol{\Sigma}_* = \begin{bmatrix} \boldsymbol{\Sigma }& \boldsymbol{0}  \\ \boldsymbol{0} & \boldsymbol{\Sigma}_{\tau^2}/\phi \end{bmatrix}, &
+\end{align*} such that \[
+\boldsymbol{z}_* \sim \text{MVN}(\boldsymbol{X}_* \boldsymbol{\beta }, \phi \Sigma_*).
+\] Using the least squares estimators to update \(\boldsymbol{\beta}\)
+and \(\phi\), we have \begin{align*}
+& \boldsymbol{\beta}^{(t)} = (\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{z}_* && \phi^{(t)} = \frac{1}{n}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)})^T\boldsymbol{\Sigma}^{-1}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)}).&
+\end{align*} To note, the variance-covariance matrixof the coefficient
+estimates variance-covariance matrix can be derived in the EM-IWLS
+algorithm and in turn can be used for statistical inferences, \[
+  \text{Var}(\boldsymbol{\beta}^{(t)}) = (\boldsymbol{X}_*^T\boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\phi^{(t)}.
+\] The rest of the algorithm remains the same as described in Section
+\ref{sec:EMCD}.
 
@@
 spike scale \(s_0\) tends to include more non-zero coefficients in the
-model. An measure of preference calculated with cross-validations,
+model. A measure of preference calculated with cross-validations,
 e.g.~deviance, area under the curve (AUC), mean squared error, can be
 used to facilitate the selection of a final model. The procedure is
-similar to the lasso implemented in the widely used R package
-\texttt{glmnet}, which quickly fits the lasso model over a list of
-values of \(\lambda\) covering the entire range, giving a sequence of
-models for users to choose from.
+similar to the Lasso implemented in the widely used R package
+\texttt{glmnet}, which quickly fits Lasso models over a list of values
+of regularization parameters \(\lambda\) , giving a sequence of models
+for users to choose from.
 

### 2021-09-27 Manuscript/01-Intro.Rmd
Revision

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution, denoted by $EF(\mu, \phi)$, with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
- E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
+ E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
 $$

### 2021-09-27 Manuscript/SS_GAM.tex
Revision

@@
 In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution \(EF(\mu, \phi)\) with mean \(\mu\) and
-dispersion \(\phi\), can be modeled with the summation of smoothing
-functions, \(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional
-vector of covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
+exponential family distribution, denoted by \(EF(\mu, \phi)\), with mean
+\(\mu\) and dispersion \(\phi\), can be modeled with the summation of
+smoothing functions, \(B_j(\cdot), j = 1, \dots, p\), of a given
+\(p\)-dimensional vector of covariates \(\boldsymbol{x}\), written as \[
+ E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
 \] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
@@
 
-\hypertarget{bayesian-hierarchical-additive-models}{%
-\section{Bayesian Hierarchical Additive
-Models}\label{bayesian-hierarchical-additive-models}}
+\hypertarget{bayesian-hierarchical-additive-models-bhams}{%
+\section{Bayesian Hierarchical Additive Models
+(BHAMs)}\label{bayesian-hierarchical-additive-models-bhams}}
 
@@
 \begin{equation}\label{eq:gam}
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
 \end{equation} and the data distribution is \begin{equation}
-f(\boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f(y_i|\boldsymbol{\beta}, \phi),\nonumber
+f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
 \end{equation} with smoothing functions \(B_j(x_j)\) of the variable
@@
 \begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) = g^{-1}(a + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
+E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
 \end{equation} where the coefficients

### 2021-09-28 Manuscript/00-main.Rmd
Add computation time for real data analyses in the manuscript.

@@
 tar_load(WLM_SBGAM_cv)
+tar_load(tbl_real_time)
 ```

### 2021-09-28 Manuscript/00-main_word.Rmd
Add computation time for real data analyses in the manuscript.

@@
 \newcommand{\tp}{*}
+\newcommand{\repa}{\text{repa}}
 \newcommand{\pr}{\text{Pr}}

### 2021-09-28 Manuscript/04-Real_Data.Rmd
Add computation time for real data analyses in the manuscript.

@@
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
+Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
 
-## Emory Cardiovarcular Biobank
+```{r}
+tbl_real_time %>% 
+  flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
+```
+
+## Emory Cardiovascular Biobank
 We use the proposed models BHAM-SSL to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.

### 2021-09-28 Manuscript/SS_GAM.tex
Add computation time for real data analyses in the manuscript.

@@
 improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}.
+hierarchical additive model, SB-GAM \citep{Bai2021}. Computation time
+for cross-validation and optimal model fitting steps are recorded and
+presented in Table \ref{tab:time_real_data}.
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\caption{Model fitting time in seconds for the two real data analysis, including computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) as the summation of the two steps.\label{tab:time_real_data}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{152.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{154.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,143.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{8.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,151.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\end{longtable}
 

### 2021-09-28 R/real_data_time.R

@@ -1,0 +1,47 @@
+create_real_data_time_tbl <- function(...){
+  
+
+  
+  # Real Work
+  time_tbl_raw <- tar_meta() %>% 
+    select(name, seconds) %>%
+    filter(name %in% c("ECB_bamlasso_cv", "ECB_bamlasso_fnl", "ECB_SBGAM_cv_raw", "ECB_SBGAM_fnl",
+                       "WLM_bamlasso_cv","WLM_bamlasso_fnl", "WLM_SBGAM_cv_raw", "WLM_SBGAM_fnl")) %>% 
+    unglue::unglue_unnest(name,
+                          "{data}_{model}_{step}", remove = FALSE) %>% 
+    arrange(data, model, step) %>%
+    pivot_wider(id_cols = data,
+                names_from = c("model", "step"),
+                values_from = "seconds") %>% 
+    rename(SBGAM_cv = SBGAM_cv_raw) %>% 
+    transmute(
+      data, bamlasso_cv, bamlasso_fnl, 
+      bamlasso_total = bamlasso_cv + bamlasso_fnl,
+      SBGAM_cv, SBGAM_fnl,
+      SBGAM_total = SBGAM_cv + SBGAM_fnl) 
+  
+  time_tbl_header <- time_tbl_raw %>% 
+    colnames %>% data.frame(col_keys = .) %>% 
+    unglue::unglue_unnest(
+      col_keys,"{line2}_{line3}",
+      na = "Data",
+      remove = FALSE) %>% 
+    mutate(line2 = recode(line2, bamlasso = "BHAM-CD",
+                          SBGAM = "SB-GAM"),
+           line3 = recode(line3, cv = "CV", fnl = "Final",
+                          total = "Total"))
+  
+  flextable(time_tbl_raw) %>% 
+    colformat_double(digits = 1) %>% 
+    set_header_df(mapping = time_tbl_header) %>% 
+    set_header_labels(bamlasso = "BHAM-CD",
+                      SBGAM = "SB-GAM",
+                      cv = "CV",
+                      fnl = "Final",
+                      total = "Total") %>% 
+    merge_h(part = "header", i = 1) %>% 
+    merge_v(part = "header", j = 1) %>% 
+    theme_booktabs %>% 
+    align(align = "center", part = "all")
+  
+}

### 2021-09-28 _targets.R

@@ -8,7 +8,7 @@ tar_option_set(
                "broom", "unglue", "knitr", "ggpubr", "xtable", "janitor", 
                "flextable", "BHAM", "BhGLM", "sparseGAM", "cosso", "glmnet"
   ),
-  imports = c("BHAM", "BhGLM", "sparseGAM")
+  imports = c( "BHAM", "BhGLM", "sparseGAM")
 )
 
 ## Load your R files
@@ -156,7 +156,7 @@ tar_plan(
   WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw,s0 = seq(0.005, 0.1, 0.01)),
 
   
-  bglm_mdl_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
+  WLM_bamlasso_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
                           ss=c(0.025, 0.05),
                           group = WLM_bgam_dat$group),
   # measure.bh(bglm_mdl_fnl),
@@ -203,6 +203,17 @@ tar_plan(
   tar_target(disc_path,
              "Manuscript/05-Discussion.Rmd",
              format = "file"),
+  
+  tar_target(tbl_real_time,
+             create_real_data_time_tbl(  # Introduce dependency
+               ECB_bamlasso_cv,
+               ECB_bamlasso_fnl,
+               ECB_SBGAM_cv_raw,
+               ECB_SBGAM_fnl,
+               WLM_bamlasso_cv,
+               WLM_bamlasso_fnl,
+               WLM_SBGAM_cv_raw,
+               WLM_SBGAM_fnl)),
   
   #* Assemble Manuscript ####
   tar_render(manu, "Manuscript/00-main.Rmd",

### 2021-09-28 Manuscript/00-main.Rmd
Add Variable selection results in the work flow.
* Change the corresponding author

@@
   org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu} and  \\email{nyi@uab.edu}"
+corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu},  \\email{nyi@uab.edu}"
 presentaddress: This is sample for present address text this is sample for present address text

### 2021-09-28 Manuscript/SS_GAM.tex
Add Variable selection results in the work flow.
* Change the corresponding author

@@
 \corres{Boyi Guo and Nengjun Yi, Department of Biostatistics, University
-of Alabama at Birmingham, Birmingham, USA. \email{boyiguo1@uab.edu} and
+of Alabama at Birmingham, Birmingham, USA. \email{boyiguo1@uab.edu},
 \email{nyi@uab.edu}}
@@
 
-\caption{Model fitting time in seconds for the two real data analysis, including computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) as the summation of the two steps.\label{tab:time_real_data}}\\
+\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
 
@@
 
-\hypertarget{emory-cardiovarcular-biobank}{%
-\subsection{Emory Cardiovarcular
-Biobank}\label{emory-cardiovarcular-biobank}}
+\hypertarget{emory-cardiovascular-biobank}{%
+\subsection{Emory Cardiovascular
+Biobank}\label{emory-cardiovascular-biobank}}
 

### 2021-09-28 _targets.R

@@ -83,6 +83,7 @@ tar_plan(
                           ss = c(0.095, 0.5)),
   
   ECB_bamlasso_insample_msr = measure.bh(ECB_bamlasso_fnl),
+  ECB_bamlasso_var = bamlasso_var_selection(ECB_bamlasso_fnl),
   # TODO: Use cv.bh to get the measures.
   
   # ** SB-GAM ---------------------------------------------------------------
@@ -101,6 +102,8 @@ tar_plan(
 
     tar_target(ECB_SBGAM_insample_msr,
              measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial")),
+  
+  ECB_SBGAM_var = (ECB_cov %>% colnames())[ECB_SBGAM_fnl$classifications!=0],
   
   
   # 
@@ -160,6 +163,7 @@ tar_plan(
                           ss=c(0.025, 0.05),
                           group = WLM_bgam_dat$group),
   # measure.bh(bglm_mdl_fnl),
+  WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),
   
   # ** SB_GAM ---------------------------------------------------------------
   
@@ -180,6 +184,7 @@ tar_plan(
     X = WLM_train_dat %>% select(starts_with("mb_")) %>% as.matrix, 
     df=5, family = "gaussian", a = 1, b = 1,
     max.iter=100, tol = 1e-6, lambda0 = 58),
+  WLM_SBGAM_vars = (WLM_train_dat %>% select(starts_with("mb_")) %>% as.matrix %>% colnames())[WLM_SBGAM_fnl$classifications!=0],
   
   # measure.glm(WLM_train_dat$out_HOMA_std, WLM_SBGAM_fnl$mu.pred, family = "gaussian"),  
   

### 2021-09-29 Manuscript/00-main.Rmd
Add ENAR formating

@@
 abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional ($p>>n$) problem using group penalties and priors.
-These models rarely smooth the nonlinear effect properly because of the sparse regularization.
-It leads to excess shrinkage on coefficients and inaccurate prediction.
-Besides, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
-While some Bayesian models can address the shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution.
-For smooth and sparse solutions, the proposed models employ bi-part spike-and-slab spline priors, including mixture double exponential and mixture normal priors.
-Two optimization-based algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for expedited computation.
-Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-art models, mgcv, COSSO and sparse Bayesian GAM.
+These models rarely smooth the nonlinear effect properly as the sparse regularization leads to excess shrinkage on coefficients and inaccurate prediction.
+Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
+While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we propose a novel spike-and-slab spline prior for smoothing functions, where a reparameterization step facilitates smoothing penalty and separate selection of linear and nonlinear spaces.
+Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed to fit the proposed model.
+Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM.
 The software implementation of the proposed models is freely available via an R package BHAM."
-keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent
+keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility
 bibliography: bibfile.bib
@@
 library(xtable)
+library(flextable)
+
+set_flextable_defaults(fonts_ignore=TRUE)
 

### 2021-09-29 Manuscript/00-main_ENAR.Rmd
Add ENAR formating

--- 2021-09-29 Manuscript/00-main.Rmd
@@
 title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
-author:
-- name: Boyi Guo*
-  num: a
-- name: Byron C. Jaeger
-  num: b 
-- name: AKM Fazlur Rahman
-  num: a   
-- name: D. Leann Long
-  num: a  
-- name: Nengjun Yi*
-  num: a
-address:
-- num: a
-  org: Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA
-- num: b
-  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
-corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu},  \\email{nyi@uab.edu}"
-presentaddress: This is sample for present address text this is sample for present address text
-authormark: Guo et al.
-articletype: Research article
-received: 2021-01-01
-revised: 2021-02-01
-accepted: 2021-03-01
-abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional ($p>>n$) problem using group penalties and priors.
-These models rarely smooth the nonlinear effect properly as the sparse regularization leads to excess shrinkage on coefficients and inaccurate prediction.
+
+# author: 
+# - name: Author 1
+  # affiliation: Department of YYY, University of XXX
+  # email: abc@def
+
+# - name: Author 2
+  # affiliation: Department of ZZZ, University of WWW
+  # email: djf@wef
+
+keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility
+
+abstract: |
+  Many proposed high-dimensional generalized additive models (GAMs) use sparse regularization, and tend to overly smooth important nonlinear functions.
+The excess shrinkage on coefficients can in turn produce inaccurate predictions.
 Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
-While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution: we propose a novel spike-and-slab spline prior for smoothing functions, where a reparameterization step facilitates smoothing penalty and separate selection of linear and nonlinear spaces.
-Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed to fit the proposed model.
+While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing function.
+A novel spike-and-slab spline prior is used to select the components of smoothing functions.
+Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for different utilities.
 Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM.
-The software implementation of the proposed models is freely available via an R package BHAM."
-keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility
+The software implementation is freely available via an R package BHAM.
+
 bibliography: bibfile.bib
-output: 
-  rticles::sim_article:
-    keep_tex: TRUE
+output: rticles::biometrics_article
+# month: "`r format(Sys.Date(), '%b')`"
+# year: "`r format(Sys.Date(), '%Y')`"
+referee: true # Papers submitted to Biometrics should ALWAYS be prepared
+              # using the referee option!!!! Turn off only to preview
+              # two column-format
 header-includes: 
@@
   - \usepackage{pgfplots}
+  - \usepackage{amsfonts}
+  # - \usepackage[modulo,pagewise,displaymath,mathlines]{lineno}
 ---
@@
 <!-- \usepackage{tikz} -->
+<!-- \linenumbers -->
 \pgfplotsset{compat=1.18}

### 2021-09-29 Manuscript/02-Method.Rmd
Revision
Add ENAR formating

@@
-# Bayesian Hierarchical Additive Models
+# Bayesian Hierarchical Additive Models (BHAMs)
 \label{sec:BHAM}
@@
 \begin{equation}\label{eq:gam}
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
 \end{equation}
@@
 \begin{equation}
-f(\bs y| \bs \beta, \phi) = \prod\limits^n_{i=1}f(y_i|\bs \beta, \phi),\nonumber
+f(\bs Y = \bs y| \bs \beta, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\bs \beta, \phi),\nonumber
 \end{equation}
@@
 \begin{equation}\label{eq:gam-repa}
-E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\beta_j X^0_j + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right],
+E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right],
 \end{equation}
@@
 where $S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S^\tp_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
-To calculate two unknown quantities $E({S_j}^{-1})$ and $E(S^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
+To calculate two unknown quantities $E({S_j}^{-1})$ and $E({S^\tp}^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
 The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the value of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
@@
 To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
-$E({S_j}^{-1}|\beta_j, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
-As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
+$E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
 On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.

### 2021-09-29 Manuscript/SS_GAM.tex
Add ENAR formating

@@
 problem using group penalties and priors. These models rarely smooth the
-nonlinear effect properly because of the sparse regularization. It leads
-to excess shrinkage on coefficients and inaccurate prediction. Besides,
-most of these GAMs consider an ``all-in-all-out'' approach for function
+nonlinear effect properly as the sparse regularization leads to excess
+shrinkage on coefficients and inaccurate prediction. Moreover, most of
+these GAMs consider an ``all-in-all-out'' approach for function
 selection, rendering them difficult to answer if nonlinear effects are
-necessary. While some Bayesian models can address the shortcomings,
+necessary. While some Bayesian models can address these shortcomings,
 using Markov chain Monte Carlo (MCMC) algorithms for model fitting
 creates a new challenge, scalability. Hence, we propose Bayesian
-hierarchical generalized additive models as a solution. For smooth and
-sparse solutions, the proposed models employ bi-part spike-and-slab
-spline priors, including mixture double exponential and mixture normal
-priors. Two optimization-based algorithms, EM-Coordinate Descent and
-EM-Iterative Weighted Least Squares, are developed for expedited
-computation. Simulation studies and metabolomics data analyses
-demonstrate improved performance against state-of-art models, mgcv,
-COSSO and sparse Bayesian GAM. The software implementation of the
-proposed models is freely available via an R package BHAM.}
+hierarchical generalized additive models as a solution: we propose a
+novel spike-and-slab spline prior for smoothing functions, where a
+reparameterization step facilitates smoothing penalty and separate
+selection of linear and nonlinear spaces. Two scalable and deterministic
+algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
+Squares, are developed to fit the proposed model. Simulation studies and
+metabolomics data analyses demonstrate improved performance against
+state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM. The
+software implementation of the proposed models is freely available via
+an R package BHAM.}
 
 \keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
-Additive Models; EM-IWLS; EM-Coordinate Decent}
+Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility}
 
@@
 transparency to understand the etiology of diseases. In addition, how to
-generalized these models to the high-dimensional setting remains
-unclear.
+generalize these models to the high-dimensional setting remains unclear.
 
@@
 \(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
-unknown quantities \(E({S_j}^{-1})\) and \(E(S^{-1}_j)\), the posterior
-probability \(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\)
-and \(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
+unknown quantities \(E({S_j}^{-1})\) and \(E({S^*}^{-1}_j)\), the
+posterior probability
+\(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
+\(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
 necessary, which can be derived via Bayes' theorem. The calculation of
@@
 coefficient specific for \(\beta^*_{jk}\).
-\(E({S_j}^{-1}|\beta_j, s_0, s_1), E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
+\(E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
 needs to be calculated to formulate \(E(Q_1)\). As neither
 \(E({S_j}^{-1}|\beta_j, s_0, s_1)\) nor
-\(E(S^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on \(\tau^2\)s,
-they can be derived using Equation (\ref{eq:exp_scale}) via Bayes'
-Theorem. On the other hand, \(\tau^{2}\) following a gamma distribution
-is a conjugate prior for the normal variance, and the conditional
-posterior density of \(\tau^{-2}\) is an inverse Gaussian distribution.
-\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
-using the closed form equation \begin{align*}
+\(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
+\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}) via
+Bayes' Theorem. On the other hand, \(\tau^{2}\) following a gamma
+distribution is a conjugate prior for the normal variance, and the
+conditional posterior density of \(\tau^{-2}\) is an inverse Gaussian
+distribution. \(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are
+calculated using the closed form equation \begin{align*}
  E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,

### 2021-09-29 _targets.R

@@ -224,8 +224,8 @@ tar_plan(
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
-  tar_render(manu_word, "Manuscript/00-main_word.Rmd",
-             output_file = "SS_GAM.docx"),
+  tar_render(manu_ENAR, "Manuscript/00-main_ENAR.Rmd",
+             output_file = "SS_GAM-ENAR.pdf"),
   
 )
 

### 2021-09-30 Manuscript/00-main.Rmd
Arrange position of figures following Stat in Med Author guideline

@@
 keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility
-bibliography: bibfile.bib
+# bibliography: bibfile.bib
 output: 
@@
 ```
+
+\bibliography{bibfile.bib}
+
+\newpage
+
+\begin{figure}
+\centering
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
+};
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
+};
+
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta_1, \bs \beta^\tp_1, \dots,\beta_j, \bs \beta^\tp_j , \dots,\beta_p, \bs \beta^\tp_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\bs \beta^T \bs X_i), \phi)$};
+
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (Theta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+\begin{pgfonlayer}{background}
+  \node [background,
+   fit=(linearGamma) (linearBeta),
+   label=above:Linear Space:] {};
+  \node [background,
+    fit=(penGamma) (penBeta),
+    label=above:Nonlinear Space:] {};
+\end{pgfonlayer}
+
+\end{tikzpicture}
+
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
+```{r, results = "asis"}
+tab_gaus$mgcv[4] <- "-"
+tab_gaus %>% 
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>% 
+  xtable(
+    align = "cccccccc",
+    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
+    label = "tab:gaus") %>% 
+  print(comment = FALSE, include.rownames = FALSE)
+  # flextable(
+  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
+  # ) %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
+```
+
+```{r, results = "asis"}
+tab_binom$mgcv[4] <- "-"
+
+tab_binom %>% 
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>%  
+  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
+  print(comment = FALSE, include.rownames = FALSE)
+```
+
+```{r}
+tbl_real_time %>% 
+  flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
+```
+
+```{r}
+rbind(
+  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
+) %>% 
+  rownames_to_column("Methods") %>% 
+  flextable() %>% 
+  flextable::autofit() %>% 
+  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}")
+```
+
+```{r}
+rbind(
+  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
+) %>% 
+  rownames_to_column("Methods") %>% 
+  flextable() %>% 
+  flextable::autofit() %>% 
+  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
+```

### 2021-09-30 Manuscript/02-Method.Rmd
Arrange position of figures following Stat in Med Author guideline

@@
 Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
-
-\begin{figure}
-\centering
-\begin{tikzpicture} [
-staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
-PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
-background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
-]
-
-\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
-};
-\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
-};
-
-\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta_1, \bs \beta^\tp_1, \dots,\beta_j, \bs \beta^\tp_j , \dots,\beta_p, \bs \beta^\tp_p) $};
-\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
-\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\bs \beta^T \bs X_i), \phi)$};
-
-\draw[->] (Theta) -- (linearGamma);
-\draw[->] (Theta) -- (penGamma);
-\draw[->] (linearGamma) -- (linearBeta) ;
-\draw[->] (penGamma) -- (penBeta);
-\draw[->] (ab) -- (Theta);
-\draw[->] (s) -- (linearBeta) ;
-\draw[->] (s) -- (penBeta);
-\draw[->] (linearBeta) -- (Beta);
-\draw[->] (penBeta) -- (Beta);
-\draw[->] (Beta) --  (Y);
-
-\begin{pgfonlayer}{background}
-  \node [background,
-   fit=(linearGamma) (linearBeta),
-   label=above:Linear Space:] {};
-  \node [background,
-    fit=(penGamma) (penBeta),
-    label=above:Nonlinear Space:] {};
-\end{pgfonlayer}
-
-\end{tikzpicture}
-
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
-\end{figure}
+\begin{center}
+Figure \ref{fig:SSprior} here
+\end{center}
 

### 2021-09-30 Manuscript/03-Simulation.Rmd
Arrange position of figures following Stat in Med Author guideline

@@
 
-```{r, results = "asis"}
-tab_gaus$mgcv[4] <- "-"
-tab_gaus %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>% 
-  xtable(
-    align = "cccccccc",
-    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
-    label = "tab:gaus") %>% 
-  print(comment = FALSE, include.rownames = FALSE)
-  # flextable(
-  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
-  # ) %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
-```
-
-```{r, results = "asis"}
-tab_binom$mgcv[4] <- "-"
-
-tab_binom %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>%  
-  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
-  print(comment = FALSE, include.rownames = FALSE)
-```
-
-To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+To evaluate the  predictive performance of the models, the statistics calculated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
 The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
-
+\begin{center}
+Table \ref{tab:gaus} here\\
+Table \ref{tab:bin_auc} here\\
+\end{center}
 ```{r, fig.height=8, eval = FALSE}

### 2021-09-30 Manuscript/04-Real_Data.Rmd
Arrange position of figures following Stat in Med Author guideline

@@
 Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
-
-```{r}
-tbl_real_time %>% 
-  flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
-```
+\begin{center}
+Table \ref{tab:time_real_data} here
+\end{center}
 
@@
 Out-of-bag samples were used for prediction performance evaluation, where deviance, area under the ROC curve (AUC), mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
-
-```{r}
-rbind(
-  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
-  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
-) %>% 
-  rownames_to_column("Methods") %>% 
-  flextable() %>% 
-  flextable::autofit() %>% 
-  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}")
-```
+\begin{center}
+Table \ref{tab:ECB_res} here
+\end{center}
 
@@
 Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
-
-```{r}
-rbind(
-  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
-  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
-) %>% 
-  rownames_to_column("Methods") %>% 
-  flextable() %>% 
-  flextable::autofit() %>% 
-  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
-```
+\begin{center}
+Table \ref{tab:WLM_res} here
+\end{center}

### 2021-09-30 Manuscript/SS_GAM.tex
Arrange position of figures following Stat in Med Author guideline

@@
 
-\begin{figure}
-\centering
-\begin{tikzpicture} [
-staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
-PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
-background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
-]
-
-\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
-};
-\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
-};
-
-\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
-\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
-\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
-
-\draw[->] (Theta) -- (linearGamma);
-\draw[->] (Theta) -- (penGamma);
-\draw[->] (linearGamma) -- (linearBeta) ;
-\draw[->] (penGamma) -- (penBeta);
-\draw[->] (ab) -- (Theta);
-\draw[->] (s) -- (linearBeta) ;
-\draw[->] (s) -- (penBeta);
-\draw[->] (linearBeta) -- (Beta);
-\draw[->] (penBeta) -- (Beta);
-\draw[->] (Beta) --  (Y);
-
-\begin{pgfonlayer}{background}
-  \node [background,
-   fit=(linearGamma) (linearBeta),
-   label=above:Linear Space:] {};
-  \node [background,
-    fit=(penGamma) (penBeta),
-    label=above:Nonlinear Space:] {};
-\end{pgfonlayer}
-
-\end{tikzpicture}
-
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
-\end{figure}
+\begin{center}
+Figure \ref{fig:SSprior} here
+\end{center}
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{ccccccc}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
-  200 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & - & 0.81 (0.05) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations} 
-\label{tab:gaus}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations} 
-\label{tab:bin_auc}
-\end{table}
-
 To evaluate the predictive performance of the models, the statistics
-caluclated based on the testing dataset, \(R^2\) for Gaussian model,
+calculated based on the testing dataset, \(R^2\) for Gaussian model,
 area under the curve (AUC) for binomial model, are calculated averaging
@@
 \ref{tab:gaus}, \ref{tab:bin_auc}.
+
+\begin{center}
+Table \ref{tab:gaus} here\\
+Table \ref{tab:bin_auc} here\\
+\end{center}
 
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
-
-\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{152.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{154.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,143.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{8.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,151.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
-
-\end{longtable}
+\begin{center}
+Table \ref{tab:time_real_data} here
+\end{center}
 
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{1.07in}|p{0.96in}|p{0.67in}|p{0.67in}|p{0.67in}|p{1.39in}}
-
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\label{tab:ECB_res}}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{455.686}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.740}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.159}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.285}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.214}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,230.063}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.706}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.208}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.257}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.240}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
-
-\end{longtable}
+\begin{center}
+Table \ref{tab:ECB_res} here
+\end{center}
 
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{1.07in}|p{0.90in}|p{0.67in}|p{0.67in}|p{0.67in}}
-
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\label{tab:WLM_res}}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{669.375}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.077}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.924}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.745}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{666.826}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.026}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.976}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.766}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
-
-\end{longtable}
+\begin{center}
+Table \ref{tab:WLM_res} here
+\end{center}
 
@@
 
+\newpage
+
+\begin{figure}
+\centering
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
+};
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
+};
+
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
+
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (Theta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+\begin{pgfonlayer}{background}
+  \node [background,
+   fit=(linearGamma) (linearBeta),
+   label=above:Linear Space:] {};
+  \node [background,
+    fit=(penGamma) (penBeta),
+    label=above:Nonlinear Space:] {};
+\end{pgfonlayer}
+
+\end{tikzpicture}
+
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{ccccccc}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   50 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
+  200 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & - & 0.81 (0.05) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations} 
+\label{tab:gaus}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllllll}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
+   10 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
+   50 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
+  200 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations} 
+\label{tab:bin_auc}
+\end{table}
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{152.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{154.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,143.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{8.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,151.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\end{longtable}
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{1.07in}|p{0.96in}|p{0.67in}|p{0.67in}|p{0.67in}|p{1.39in}}
+
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\label{tab:ECB_res}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{455.686}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.740}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.159}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.285}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.214}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,230.063}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.706}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.208}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.257}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.240}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
+
+\end{longtable}
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{1.07in}|p{0.90in}|p{0.67in}|p{0.67in}|p{0.67in}}
+
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\label{tab:WLM_res}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{669.375}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.077}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.924}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.745}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{666.826}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.026}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.976}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.766}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
+
+\end{longtable}
+
 \end{document}

### 2021-09-30 Manuscript/03-Simulation.Rmd
Fix the table and figure placement

@@
 
-To evaluate the  predictive performance of the models, the statistics calculated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+```{r, results = "asis"}
+tab_gaus$mgcv[4] <- "-"
+tab_gaus %>% 
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>% 
+  xtable(
+    align = "cccccccc",
+    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
+    label = "tab:gaus") %>% 
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+  # flextable(
+  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
+  # ) %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
+```
+
+```{r, results = "asis"}
+tab_binom$mgcv[4] <- "-"
+
+tab_binom %>% 
+  mutate(p = as.integer(p)) %>% 
+  rename(
+    "P" = p,
+    "BHAM-IWLS" = bglm_spline_de,
+    "BHAM-CD" = blasso_spline,
+    "COSSO" = cosso,
+    "Adaptive COSSO" = acosso,
+    "SB-GAM" = SB_GAM
+  ) %>%  
+  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+```
+
+To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
 The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
+
 ```{r, fig.height=8, eval = FALSE}

### 2021-09-30 Manuscript/04-Real_Data.Rmd
Fix the table and figure placement

@@
 Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
+
+```{r results = "asis"}
+tbl_real_time %>% 
+  xtable(
+    caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.",
+    label= "tab:time_real_data"
+  ) %>% 
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+
+  # flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
+```
 
@@
 Out-of-bag samples were used for prediction performance evaluation, where deviance, area under the ROC curve (AUC), mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
+
+```{r results = "asis"}
+rbind(
+  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
+) %>% 
+  rownames_to_column("Methods") %>% 
+  xtable(
+    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation",
+    label = "tab:ECB_res"
+  ) %>% 
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+
+  # flextable::flextable() %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}") %>% 
+  # flextable_to_rmd()
+```
 
@@
 Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}
+
+```{r results = "asis"}
+rbind(
+  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
+) %>% 
+  rownames_to_column("Methods")  %>% 
+  xtable(
+    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
+    label = "tab:WLM_res"
+  ) %>% 
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+
+  # flextable() %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
+```

### 2021-09-30 Manuscript/SS_GAM.tex
Fix the table and figure placement

--- 2021-09-29 Manuscript/SS_GAM.tex
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{1.07in}|p{0.96in}|p{0.67in}|p{0.67in}|p{0.67in}|p{1.39in}}
-
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\label{tab:ECB_res}}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{auc}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{misclassification}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{455.686}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.740}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.159}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.285}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.214}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.96in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,230.063}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.706}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.208}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.257}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 1.39in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.240}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-6}
-
-\end{longtable}
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrrr}
+  \hline
+Methods & deviance & auc & mse & mae & misclassification \\ 
+  \hline
+BHAM-SSL & 455.69 & 0.74 & 0.16 & 0.28 & 0.21 \\ 
+  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.26 & 0.24 \\ 
+   \hline
+\end{tabular}
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation} 
+\label{tab:ECB_res}
+\end{table}
 
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{1.07in}|p{0.90in}|p{0.67in}|p{0.67in}|p{0.67in}}
-
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\label{tab:WLM_res}}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Methods}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{deviance}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{R2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mae}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{669.375}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.077}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.924}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.745}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{666.826}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.026}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.976}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.766}}} \\
-
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-5}
-
-\end{longtable}
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Methods & deviance & R2 & mse & mae \\ 
+  \hline
+BHAM-SSL & 669.38 & 0.08 & 0.92 & 0.74 \\ 
+  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
+   \hline
+\end{tabular}
+\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation} 
+\label{tab:WLM_res}
+\end{table}
 

### 2021-09-30 R/real_data_time.R

@@ -1,4 +1,4 @@
-create_real_data_time_tbl <- function(...){
+create_real_data_time_tbl <- function(raw = FALSE, ...){
   
 
   
@@ -20,6 +20,8 @@ create_real_data_time_tbl <- function(...){
       SBGAM_cv, SBGAM_fnl,
       SBGAM_total = SBGAM_cv + SBGAM_fnl) 
   
+  if(raw) return(time_tbl_raw)
+  
   time_tbl_header <- time_tbl_raw %>% 
     colnames %>% data.frame(col_keys = .) %>% 
     unglue::unglue_unnest(
@@ -30,6 +32,8 @@ create_real_data_time_tbl <- function(...){
                           SBGAM = "SB-GAM"),
            line3 = recode(line3, cv = "CV", fnl = "Final",
                           total = "Total"))
+  
+
   
   flextable(time_tbl_raw) %>% 
     colformat_double(digits = 1) %>% 

### 2021-09-30 _targets.R

@@ -210,7 +210,8 @@ tar_plan(
              format = "file"),
   
   tar_target(tbl_real_time,
-             create_real_data_time_tbl(  # Introduce dependency
+             create_real_data_time_tbl( 
+               raw=TRUE,# Introduce dependency
                ECB_bamlasso_cv,
                ECB_bamlasso_fnl,
                ECB_SBGAM_cv_raw,
@@ -221,8 +222,8 @@ tar_plan(
                WLM_SBGAM_fnl)),
   
   #* Assemble Manuscript ####
-  tar_render(manu, "Manuscript/00-main.Rmd",
-             output_file = "SS_GAM.pdf"),
+  # tar_render(manu, "Manuscript/00-main.Rmd",
+  #            output_file = "SS_GAM.pdf"),
   
   tar_render(manu_ENAR, "Manuscript/00-main_ENAR.Rmd",
              output_file = "SS_GAM-ENAR.pdf"),

### 2021-10-01 Manuscript/00-main_ENAR.Rmd
Fix the table and figure placement
Revise for ENAR submission

@@
 ```
-
 ```{r child="02-Method.Rmd"}
@@
 ```
+
+\appendix
+\section{}
+\subsection{A.1 Derivation of $E({\tau^2}^{-1})$\label{ap:IWLS}}
+
+\subsection{A.2 Weight Loss Maintenance Cohort Data Analysis\label{ap:WLM}}
+We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
+In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
+In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
+Quality control and natural log transformation were performed during metabolites data preparation.
+The outcome of interest was standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
+After removing missing datapoints and addressing outliers in the data, there were `r nrow(WLM_train_dat)` remaining in the analysis.
+5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
+10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
+
+```{r results = "asis"}
+rbind(
+  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
+) %>%
+  rownames_to_column("Methods")  %>%
+  xtable(
+    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
+    label = "tab:WLM_res"
+  ) %>%
+  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+
+  # flextable() %>%
+  # flextable::autofit() %>%
+  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
+```
+
+\newpage

### 2021-10-01 Manuscript/01-Intro.Rmd
Add ENAR formating
Revise for ENAR submission

--- 2021-09-23 Manuscript/01-Intro.Rmd
@@
 sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
-There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review.
+There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see @Mallick2013 for review.
 These models are built upon a restrictive and unrealistic assumption, linearity.
@@
 These models are not particularly attractive in biomedical studies, as they lack the transparency to understand the etiology of diseases.
-In addition, how to generalized these models to the high-dimensional setting remains unclear.
+In addition, how to generalize these models to the high-dimensional setting remains unclear.
 
@@
 Nonparametric regression models are appropriate alternatives to the "black-box" models thanks to their balance between model flexibility and interpretability.
-Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution $EF(\mu, \phi)$ with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
-$$
- E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)),
-$$
-where $g^{-1}(\cdot)$ is the inverse of a monotonic link function.
+Among those, generalized additive models (GAMs), proposed in the seminal work of @Hastie1987, grew to be one of the most popular modeling tools.
+<!-- In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution, denoted by $EF(\mu, \phi)$, with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as  -->
+<!-- $$ -->
+<!--  E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)), -->
+<!-- $$ -->
+<!-- where $g^{-1}(\cdot)$ is the inverse of a monotonic link function.
 The smoothing functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
 Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
-
+-->
 There exists some proposals to generalize the classical GAM to accommodate high-dimensional applications.
@@
 To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via re-parameterization of the smoothing function to avoid overly shrinking basis function coefficients.
-Smoothing penalties are commonly implemented in the classical GAMs through smoothing regression splines.
+Specifically, we incorporate smoothing penalties in the model via reparameterization of the smoothing function to avoid overly shrinking basis function coefficients.
+<!-- Smoothing penalties are commonly implemented in the classical GAMs through smoothing regression splines.
 They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
-A smoothing penalty conditioning on a smoothing parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
-\begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
-\end{equation}
-where $\bs S_j$ is a known smoothing penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
+A smoothing penalty conditioning on a smoothing parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as -->
+<!-- \begin{equation}\label{eq:smoothpen} -->
+<!--   \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j , -->
+<!-- \end{equation} -->
+<!-- where $\bs S_j$ is a known smoothing penalty matrix, and $\bs \beta_j$ are the basis function coefficients.  -->
 Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
@@
 In addition, two scalable optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
-
-<!-- Marra and Wood [@Marra2011] introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
-The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
+<!-- Marra and Wood [@Marra2011] introduced a reparameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
+The reparameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
 -->
  <!-- In addition, the priors of the linear and nonlinear effects are independent, creating computational convenience and further improving the scalability of the model. -->
-
-The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e.
+<!-- The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e.
 the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in many ways.
@@
 Furthermore, the proposed model addresses the incapability of bi-level selection in SB-GAM.
+-->
 Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
@@
 In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
-Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.
+Analysis of metabolomics datasets are presented in Section \ref{sec:real_data}.
 Conclusion and discussions are given in Section \ref{sec:concl}.

### 2021-10-01 Manuscript/02-Method.Rmd
Fix the table and figure placement
Revise for ENAR submission

--- 2021-09-29 Manuscript/02-Method.Rmd
@@
 
-Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model and its matrix form,
+A generalized additive model can be formulated in its matrix form as,
 \begin{equation}\label{eq:gam}
@@
 \end{equation}
-with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
-the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
-The dimension of the design matrix depends on the choice of the smoothing function.
-For example, a cubic spline function with $k$ knots produces a $k+3$ elements column vector.
-Its corresponding smoothing penalty matrix is denoted as $\bs S_j$.
-$\bs \beta_j$ denotes the basis function coefficients for the $j$th variable such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
-With slight abuse of notation, we denote vectors and matrices in bold fonts $\bs \beta, \bs X$ with conformable dimensions, where scalar and random variables are denoted in with unbold fonts $\beta, X$.
+with the disperse parameter $\phi$ and smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix is denoted $\bs X_j$ for the variable $x_j$
+<!-- . The dimension of the design matrix depends on the choice of the smoothing function. For example, a cubic spline function with $k$ knots produces a $k+3$ elements column vector. Its  -->
+with its smoothing penalty matrix denoted as $\bs S_j$.
+$\bs \beta_j$ denotes the basis function coefficients such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
+With slight abuse of notation, we denote vectors and matrices in bold fonts $\bs \beta, \bs X$ of conformable dimensions, where scalar and random variables are denoted in with unbold fonts $\beta, X$.
 The matrix transposing operation is denoted with a superscript $^T$.
-To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
+<!-- To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
 $B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+-->
 
 To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models. 
-The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
+<!-- The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
 However, the direct integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and  Wood [@Marra2011] proposed a re-parameterization procedure to accommodate the smoothing penalty implicitly.
-Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
-To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
+-->
+While it is hard to directly integrate smoothing penalty with sparsity penalty, @Marra2011 proposed a reparameterization to implement smoothing penalty in the design matrix.
+Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where eigenvectors and eigenvalues are in the matrices $\bs U$ and $\bs D$ respectively.
+The zero eigenvalue and its corresponding eigenvector span the linear space of the smoothing function, which allows us to separate the linear space from the smoothing function.
+By multiplying the design matrix $X$ and eigenvector matrix $U$ and proper scaling with eigenvalues, we can have a new design matrix such that the smoothing function of variable $x_j$ can be written as
+$$
+B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \beta_j X^0_j + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
+$$
+where $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable; the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of non-linear space.
+
+<!-- To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
 For the most popular smoothing function, cubic splines, the dimension of the linear space is one.
@@
 An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
-With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
-A spline function can be expressed in the matrix form
-$$
-B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \beta_j X^0_j + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
-$$
-and the generalized additive model in Equation (\ref{eq:gam}) be
-\begin{equation}\label{eq:gam-repa}
-E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right],
-\end{equation}
-where the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of non-linear space.
+With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the reparameterization.
+A spline function can be expressed in the matrix form -->
+<!-- $$ -->
+<!-- B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \beta_j X^0_j + \bs {\beta_j^\tp}^T \bs X_j^\tp , -->
+<!-- $$ -->
+<!-- and the generalized additive model in Equation (\ref{eq:gam}) be -->
+<!-- \begin{equation}\label{eq:gam-repa} -->
+<!-- E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right], -->
+<!-- \end{equation} -->
+<!-- where the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of non-linear space.
+-->
 
-The re-parameterization step sets up the foundation of the proposed model and provides three-fold benefits.
-First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection.
+<!-- The reparameterization step sets up the foundation of the proposed model and provides three-fold benefits.
+First of all, the reparameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection.
 Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear from the nonlinear space, improving the feasibility of bi-level function selection and inference.
@@
 Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of priors, from column-specific priors to a unified prior.
+-->
 
-## Spike-and-Slab Smooth Priors
+## Spike-and-Slab Spline Priors
 
 The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection. 
-We defer to Bai et al.
+<!-- We defer to Bai et al.
 [@Bai2021Review] for an in-depth introduction to spike-and-slab priors.
-To summarize, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{\text{spike}}(\cdot)$ for weak signals and a flat slab density $f_{\text{slab}}(\cdot)$ for strong signals, mathematically
-$$
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-$$
-The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma \in \{0,1\}$ that indicates whether the variable $x$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
-[@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.
-
+To summarize, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{\text{spike}}(\cdot)$ for weak signals and a flat slab density $f_{\text{slab}}(\cdot)$ for strong signals, mathematically -->
+<!-- $$ -->
+<!--  \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta). -->
+<!-- $$ -->
+<!-- The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma \in \{0,1\}$ that indicates whether the variable $x$ is included in the model.
+There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see  -->
+The development includes @George1993; @George1997; @Chipman1996 for grouped variables; @Brown1998 for multivariate outcomes; @Ishwaran2005; @Clyde2004; @Rockova2014a for optimization-based model fitting algorithms and reference therein.
 <!-- Comments -->
@@
 -->
+<!-- However, the major criticism of early spike-and-slab models is being computationally prohibitive.
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
+-->
+Particularly, the development of spike-and-slab Lasso (SSL) [@Rockova2018b; @Rockova2018] substantially improves the scalability of SS models by using coordinate-wise implementations for posterior exploration.
+Yi and his group later developed generalized and cox SSL models for high-dimensional -omics data analysis.[@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+\begin{align} \label{eq:ssl}
+\beta | \gamma &\sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber\\
+\text{or} \qquad \beta | \gamma &\sim DE(0, S), 0 < s_0 < s_1,
+\end{align}
+<!-- Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as one single double exponential density, -->
+<!-- It can be aslso expressed in the form of a congregated double exponential distribution -->
+<!-- \begin{equation} \label{eq:ssl} -->
+<!-- \end{equation} -->
+with the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$.
 
-The major criticism of early spike-and-slab models is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
-Particularly, the development of spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
-The spike-and-slab Lasso (SSL) prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
-\begin{equation} 
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation}
-Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as one single double exponential density,
-\begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
-\end{equation}
-with the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$.
-The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is preferred in high-dimensional data analysis.
+<!-- The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is preferred in high-dimensional data analysis.
 We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
 Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.
+-->
 
-### Spike-and-Slab Lasso Prior
-
-We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear modeling with bi-level selection.
-The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
+<!-- ### Spike-and-Slab Lasso Prior -->
+We introduce a novel SSL type prior of GAMs for high-dimensional nonlinear modeling with bi-level selection. 
+<!-- The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
 Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
-Specifically, we impose independent group priors on the linear space coefficients and on the nonlinear space coefficients respectively,
+-->
+Specifically, we respectively impose independent group priors on the linear space coefficient $\beta_j$ and on the nonlinear space coefficients $\bs \beta^\tp_j$ of $K_j$ elements, 
 \begin{align}\label{eq:bham_ssl}
@@
 A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}.
-To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
+<!-- To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
 It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
@@
 This reduces the proposed prior to the classic spike-and-slab lasso prior.
-
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
+-->
+<!-- The reparameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
 First of all, the smoothing function bases are linear dependent and ask for extra attention.
@@
 Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables.
+-->
 
-The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
+The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution.
 The two indicators of the $j$th predictor, $\gamma_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
@@
 When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
+Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
 
@@
 \centering
+\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
 \begin{tikzpicture} [
@@
 \end{tikzpicture}
-
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
 \end{figure}
 
-### Other Priors
+<!-- ### Other Priors -->
 
-With the re-parameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
+<!-- With the reparameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
 These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
 Here we elaborate the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in BHAM.
+-->
 
-A spike-and-slab mixture normal spline prior can be expressed as 
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j.
-\end{align*}
-Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
+<!-- A spike-and-slab mixture normal spline prior can be expressed as  -->
+<!-- \begin{align*} -->
+<!--   \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\ -->
+<!--   \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j. -->
+<!-- \end{align*} -->
+<!-- Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
 One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
 Distributions with heavier tails can be used as an alternative, for example mixture Student's $t$ distribution with small degree of freedom.
+-->
 
@@
 
-The proposed models can be fitted with MCMC algorithms.
+<!-- The proposed models can be fitted with MCMC algorithms.
 Nevertheless, the computational burden of MCMC algorithms creates scalability issues.
@@
 Hence, we feel compelled to develop scalable algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
+-->
 
-As alternatives to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of uncertainty measures.
+<!-- As alternatives to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of uncertainty measures.
 The earlier work for fitting spike-and-slab models using optimization algorithms includes EMVS [@Rockova2014a].
 Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
-<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
+-->
+<!-- <!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
 The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$.
@@
 -->
-In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, followed by updating $\phi, \theta$.
-<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
+<!-- In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, followed by updating $\phi, \theta$.
+-->
+<!-- <!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
 EMVS is also compatible with structured prior when dealing with grouped variables.
@@
 -->
-The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm to fit SSL models.
-<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
+<!-- The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm to fit SSL models. -->
+<!-- <!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
 The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
@@
 -->
-Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research
+<!-- Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research  -->
 
@@
 
-In this section, we extend the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
-To note, the two proposed algorithms provides different utilities.
-The EM-CD algorithm is specifically for fitting BHAM-SSL with an expedited performance, recommending to use in high and ultra-high dimensional setting.
+In this section, we introduce the two EM-based algorithms, EM-Coordinate Descent (EM-CD) and EM-Iterative Weighted Least Squares (EM-IWLS) algorithms, to fit BHAMs.
+The two proposed algorithms provides different utilities: the EM-CD algorithm is specifically to fit BHAM with an expedited computation, recommended to use in high and ultra-high dimensional setting.
 A specific concern of EM-CD algorithm is that it provides no information for inference.
 In contrast, the EM-IWLS can extract the variance-covariance matrix and be used to construct Wald type inference.
-Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
+<!-- Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
 To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
 The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
-
+-->
 <!-- , stemming from previous EM-coordinate descent algorithms[@Tang2018; @Tang2019] and EM-IWLS [@Rockova2014a].
@@
 -->
-
 <!-- The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the nonlinear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
@@
 -->
-
-### EM algorithms
+<!-- ### EM algorithms -->
 
@@
 Instead of maximizing the the likelihood function, the algorithm maximizes the expectation of the likelihood function with respect to the "missing" data.
-
-The recursive algorithm consists of two steps:
-
-* E-step: to calculate the expectation of the posterior density function with respect to some "missing" data
-* M-step: to maximize the expectation derived in the E-step to calculate the parameter of interest
-
-For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data.
+<!-- The recursive algorithm consists of two steps:  -->
+<!-- * E-step: to calculate the expectation of the posterior density function with respect to some "missing" data -->
+<!-- * M-step: to maximize the expectation derived in the E-step to calculate the parameter of interest -->
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model.
 Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
@@
 \begin{aligned}
-& \text{argmax}_{\Theta}
-\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X}) \\
+ \text{argmax}_{\Theta}
+\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})
 &= \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{\tp}_{jk}|\gamma^{\tp}_{j})\right]\\
@@
 Meanwhile, conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
-Depending on the choice of coefficient priors, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via CD algorithm or IWLS algorithm in each iteration.
+In the optimization step, $Q_1$ can be treated as penalized likelihood function and solved via CD algorithm or IWLS algorithm.
 Maximization of $E(Q_2)$ can be solved via closed form equations following the beta-binomial conjugate relationship.
 
-### EM-Coordinate Descent
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate the parameters in the M-step.
-Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
-
+<!-- ### EM-Coordinate Descent -->
+<!-- \label{sec:EMCD} -->
+<!-- \vspace{\baselineskip} -->
+\textit{EM-Coordinate Descent\label{sec:EMCD}} When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate the parameters in the M-step.
+Coordinate descent [@Friedman2010] is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
 To recall, the density function of spike-and-slab mixture double exponential prior can be written as
@@
 $$
-and $E(Q_1)$ can be expressed as a likelihood function with $l_1$ penalty
+and $E(Q_1)$ can be expressed as a $l_1$ penalized likelihood function 
 \begin{equation}\label{eq:Q1_CD}
@@
 \end{align}
-After replacing withe calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm [@Friedman2010].
+After replacing withe calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm .
 Independently, the remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
@@
 
-### EM-IWLS
+<!-- ### EM-IWLS -->
+<!-- \vspace{\baselineskip} -->
+ \textit{EM-Iterative Weighted Least Squares} The iterative weighted least squares algorithm is a classifical algorithm to fit generalized linear models.
+Here we combine IWLS and EM algorithm to fit BHAM.
 
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
-The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
-In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary.
-
+ <!-- , and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
+-->
+<!-- In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary.
+-->
 <!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
-
-A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
+Following @Yi2012, a double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
 \begin{align*}
-  \beta|\tau^2 &\sim N(0, \tau^2)\\
-  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
+  \beta|\tau^2 \sim N(0, \tau^2), \quad \tau^2|S \sim Gamma(1, 1/(2S^2)), 
 \end{align*}
-For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
-The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
-\begin{equation}\label{eq:Q1_IWLS}
-Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
-\end{equation}
-Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
+<!-- For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).  -->
+<!-- The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$: -->
+<!-- \begin{equation}\label{eq:Q1_IWLS} -->
+<!-- Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
+-->
+<!-- \end{equation} -->
+<!-- Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
 To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
@@
 On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
-$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
-\begin{align*}
- E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
-\end{align*}
-where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
+$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation -->
+<!-- \begin{align*} -->
+<!--  E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|, -->
+<!-- \end{align*} -->
+<!-- where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. -->
+We can write the write $E(Q_1)$ as the following up to constant additive terms
+<!-- With simplification (up to constant additive terms), we have -->
 \begin{equation}\label{eq:EQ1_IWLS}
@@
 \end{equation}
+with unknown quantities $E({\tau_j}^{-2})$ and $E({\tau_{jk}^\tp}^{-2})$ for $k = 1, \dots, K_j$. 
 <!-- $E(Q_1)$ can be seen as a $l_2$ penalized likelihood function where the regularization parameter $\lambda = 2E({\tau}^{-2})$. Equivalently, we have  -->
-$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$, and we can formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal,
-$$
-\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
-$$
-
-Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood:
-$$
-f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
-$$
-where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. The pseudo-response and the pseudo-weight are calculated by
-$$
-\begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
-\end{aligned}
-$$
-where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
-
-With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
-We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*}
-& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
-  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
-  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
-\end{align*}
- such that 
-$$
-\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
-$$
-Using the least squares estimators to update $\bs\beta$ and $\phi$, we have 
-\begin{align*}
-& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
-\end{align*}
-To note, the variance-covariance matrixof the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
-$$
-  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
-$$
-The rest of the algorithm remains the same as described in Section \ref{sec:EMCD}.
-
+$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$.
+That allows to formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal.
+The calculation of $E(\tau^2)$s is similar to the calculation of unknown quantities $E(S^{-1})$ in EM-CD via Bayes' theorem.
+We defer the detailed calculation to Appendix A.1.
+In each iteration of the EM-algorithm, we can approximate the generalized model likelihood with a weighted multivariate normal density following the classical IWLS, and augment the multivariate normal density of the coefficients $\bs \beta$.
+Least squares algorithm is used to derive the per-iteration estimate of the coefficients.
+The rest of the steps follows EM-CD algorithm.
+<!-- $$ -->
+<!-- \bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}). -->
+<!-- $$ -->
+<!-- Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood: -->
+<!-- $$ -->
+<!-- f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma ) -->
+<!-- $$ -->
+<!-- where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. -->
+<!-- The pseudo-response and the pseudo-weight are calculated by -->
+<!-- $$ -->
+<!-- \begin{aligned} -->
+<!-- z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i), -->
+<!-- \end{aligned} -->
+<!-- $$ -->
+<!-- where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
+-->
+<!-- With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
+We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*} -->
+<!-- & \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} && -->
+<!--   \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} && -->
+<!--   \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, & -->
+<!-- \end{align*} -->
+<!--  such that  -->
+<!-- $$ -->
+<!-- \bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*). -->
+<!-- $$ -->
+<!-- Using the least squares estimators to update $\bs\beta$ and $\phi$, we have  -->
+<!-- \begin{align*} -->
+<!-- & \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).& -->
+<!-- \end{align*} -->
+<!-- To note, the variance-covariance matrixof the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences, -->
+<!-- $$ -->
+<!--   \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}. -->
+<!-- $$ -->
+<!-- The rest of the algorithm remains the same as described in Section \ref{sec:EMCD}. -->
 <!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
@@
 -->
-
 <!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients.
@@
 -->
-
 <!-- Totally, the framework of the proposed EM IRLS algorithm was summarized as follows: -->
-
 <!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ [for $\bs \beta$ and $\bs \theta$.
 For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
-
 <!-- 2) Iterate over the E-step and M-step until convergence -->
-
 <!-- E-step: calculate $E(\gamma_{j})$, $E(\gamma^{pen}_{j})$and $E({S}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
-
 <!-- M-step: -->
-
 <!-- a) Update $\bs \beta^{(t)}$ using the IRLS algorithm -->
-
 <!-- b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta} -->
-
 <!-- We assess convergence by the criterion: -->
@@
 deviance, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
-The procedure is similar to the Lasso implemented in the widely used R package `glmnet`, which quickly fits Lasso models over a list of values of regularization parameters $\lambda$ , giving a sequence of models for users to choose from.
+<!-- The procedure is similar to the Lasso implemented in the widely used R package `glmnet`, which quickly fits Lasso models over a list of values of regularization parameters $\lambda$ , giving a sequence of models for users to choose from.
+-->

### 2021-10-01 Manuscript/03-Simulation.Rmd
Revise for ENAR submission

@@
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
-
-Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al.
+<!-- Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al.
 [@Ravikumar2009], stochastic search term selection for GAM [@Scheipl2012].
 However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increased number of predictors.
-[@Scheipl2013]
+[@Scheipl2013] -->
 
-## Monte Carlo Simulation Study
+<!-- ## Monte Carlo Simulation Study -->
 We follow the data generating process described in Bai [@Bai2021].
@@
   # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
-```
 
-```{r, results = "asis"}
 tab_binom$mgcv[4] <- "-"

### 2021-10-01 Manuscript/04-Real_Data.Rmd
Revise for ENAR submission

@@
 
-In this section, we apply the proposed models BHAM-SSL, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+We apply the proposed models BHAM, fitted with the EM-CD algorithm ,to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+We present the analysis results for the Emory Cardiovascular Biobank data [@Mehta2020], and include the analysis of Weight Loss Maintenance Cohort  data in Appendix A.2.
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
-Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
+Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}, where we see the proposed model are more computationally efficient.
 
@@
 
-## Weight Loss Maintenance Cohort 
-We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
+<!-- ## Weight Loss Maintenance Cohort  -->
+<!-- We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
 In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
@@
 Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
+-->
 
-```{r results = "asis"}
-rbind(
-  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
-  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
-) %>% 
-  rownames_to_column("Methods")  %>% 
-  xtable(
-    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
-    label = "tab:WLM_res"
-  ) %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
+<!-- ```{r results = "asis"} -->
+<!-- rbind( -->
+<!--   'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0), -->
+<!--   'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda) -->
+<!-- ) %>% -->
+<!--   rownames_to_column("Methods")  %>% -->
+<!--   xtable( -->
+<!--     caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation", -->
+<!--     label = "tab:WLM_res" -->
+<!--   ) %>% -->
+<!--   print(comment = FALSE, include.rownames = FALSE, caption.placement = "top") -->
 
-  # flextable() %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
-```
+<!--   # flextable() %>% -->
+<!--   # flextable::autofit() %>% -->
+<!--   # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}") -->
+<!-- ``` -->

### 2021-10-01 Manuscript/05-Discussion.Rmd
Replace all ^\tp with *
Revise for ENAR submission

@@
 
-The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work.
+<!-- The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work.
 Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based fast and scalable algorithms.
@@
 Moreover, the proposed model renders additional flexibility to model specification, allowing various choices of smoothing functions and degrees of freedom.
+-->
 
@@
 We currently address it analytically by forcing to include the linear component when non-linear component is included.
-Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e.
-$\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.
+Another possible solution is to impose an dependent structure of $\gamma_{j}^\tp$ on $\gamma_{j^{0}}$, i.e.
+$\gamma_j^\tp|\gamma_{j}^{0}, \theta_j$.
 Secondly, while the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
 Our current effort resides on expanding GAM to Cox proportional hazard models.
-An naive approach would be convert an time-to-event outcome to a poisson outcome following Whitehead [@Whitehead1980].
+An naive approach would be convert an time-to-event outcome to a poisson outcome following @Whitehead1980.
 However, it would be more efficient to directly fitting Cox models via penalized pseudo likelihood function.
@@
 Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.
-The similar strategy has been used in Ferrari and Dunson [@Ferrari2020].
+The similar strategy has been used in @Ferrari2020.
 
@@
 On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e.
-$\gamma^{pen}_{jk}$ of $\beta^{pen}_{jk}$ for $k = 1\dots, K_j$.
+$\gamma^\tp_{jk}$ of $\beta^\tp_{jk}$ for $k = 1\dots, K_j$.
 This prior set-up generalized to the group spike-and-slab.

### 2021-10-01 Manuscript/00-main_ENAR.Rmd
ENAR Submission Version

@@
 While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing function.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing functions.
 A novel spike-and-slab spline prior is used to select the components of smoothing functions.
@@
 \section{}
-\subsection{A.1 Derivation of $E({\tau^2}^{-1})$\label{ap:IWLS}}
+\subsection{A.1 EM-Iterative Weighted Least Squares Algorithm\label{ap:IWLS}}
+Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
+The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
+In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary.
+<!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
+A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
+\begin{align*}
+  \beta|\tau^2 &\sim N(0, \tau^2)\\
+  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
+\end{align*}
+For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
+The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
+\begin{equation}\label{eq:Q1_IWLS}
+Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
+\end{equation}
+Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
+To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
+$E({S_j}^{-1}|\beta_j, s_0, s_1), E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E(S^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
+On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
+$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
+\begin{align*}
+ E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
+\end{align*}
+where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
+\begin{equation}
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
+\end{equation}
+<!-- $E(Q_1)$ can be seen as a $l_2$ penalized likelihood function where the regularization parameter $\lambda = 2E({\tau}^{-2})$. Equivalently, we have  -->
+$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$, and we can formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal,
+$$
+\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
+$$
+Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood:
+$$
+f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
+$$
+where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. The pseudo-response and the pseudo-weight are calculated by
+$$
+\begin{aligned}
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
+\end{aligned}
+$$
+where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
+With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
+We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*}
+& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
+  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
+  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
+\end{align*}
+ such that 
+$$
+\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
+$$
+Using the least squares estimators to update $\bs\beta$ and $\phi$, we have 
+\begin{align*}
+& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
+\end{align*}
+To note, the variance-covariance matrixof the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
+$$
+  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
+$$
+The rest of the algorithm remains the same as described in Section \ref{sec:EMCD}.
 
 \subsection{A.2 Weight Loss Maintenance Cohort Data Analysis\label{ap:WLM}}
-We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
+We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
 In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
@@
 After removing missing datapoints and addressing outliers in the data, there were `r nrow(WLM_train_dat)` remaining in the analysis.
-5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
+5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM and the SB-GAM.
 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
@@
 rbind(
-  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'BHAM' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
   'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
@@
   rownames_to_column("Methods")  %>% 
+  select(-mae) %>% 
+  rename(Deviance = deviance,
+         MSE = mse)%>% 
   xtable(
-    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
+    caption = "Prediction performance of of optimal BHAM and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$(R2) and mean squared error (MSE)",
     label = "tab:WLM_res"

### 2021-10-01 Manuscript/01-Intro.Rmd
ENAR Submission Version

@@
 Scheipl et al.
-[@Scheipl2012] proposed a spike-and-slab structure prior, normal-mixture-of-inverse gamma prior, that address the previous challenges.
+[@Scheipl2012] proposed a spike-and-slab structure prior that resolves the previous problems.
 But the model fitting relies on computational intensive MCMC algorithmic and creates scalability concern.
@@
 -->
-Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
+Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via \texttt{BHAM} at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
 

### 2021-10-01 Manuscript/02-Method.Rmd
ENAR Submission Version

@@
 
-To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models. 
+To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models [@Wood2017].
 <!-- The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
@@
 While it is hard to directly integrate smoothing penalty with sparsity penalty, @Marra2011 proposed a reparameterization to implement smoothing penalty in the design matrix.
-Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where eigenvectors and eigenvalues are in the matrices $\bs U$ and $\bs D$ respectively.
+Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where eigenvectors and eigenvalues are arranged in the matrices $\bs U$ and $\bs D$ respectively.
 The zero eigenvalue and its corresponding eigenvector span the linear space of the smoothing function, which allows us to separate the linear space from the smoothing function.
-By multiplying the design matrix $X$ and eigenvector matrix $U$ and proper scaling with eigenvalues, we can have a new design matrix such that the smoothing function of variable $x_j$ can be written as
+By multiplying the design matrix $X$ and eigenvector matrix $U$ and properly scaling with eigenvalues, we can have a new design matrix such that the smoothing function of variable $x_j$ can be written as
 $$
@@
 Particularly, the development of spike-and-slab Lasso (SSL) [@Rockova2018b; @Rockova2018] substantially improves the scalability of SS models by using coordinate-wise implementations for posterior exploration.
-Yi and his group later developed generalized and cox SSL models for high-dimensional -omics data analysis.[@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
-The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+Yi and his group later developed generalized and cox SSL models for high-dimensional -omics data analysis.[@Tang2017a; @Tang2017; @Tang2018; @Tang2019] The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
 \begin{align} \label{eq:ssl}
@@
 -->
-Specifically, we respectively impose independent group priors on the linear space coefficient $\beta_j$ and on the nonlinear space coefficients $\bs \beta^\tp_j$ of $K_j$ elements, 
+Specifically, we impose independent group priors on the linear space coefficient $\beta_j$ and on the nonlinear space coefficients $\bs \beta^\tp_j$ of $K_j$ elements, 
 \begin{align}\label{eq:bham_ssl}
@@
 \end{align}
-where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
+where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear and nonlinear effects of the $j$th variable respectively.
 $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
@@
 Meanwhile, we specify that $\gamma_{j}$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
+The discussion on how to improve the hierarchy of $\gamma_{j}$ and $\gamma_{j}^\tp$ is located in Section \ref{sec:concl}.
 We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
@@
 When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
+Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM, and is visually presented in Figure \ref{fig:SSprior}.
 
@@
 \centering
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\caption{Directed acyclic grpah of BHAM model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
 \label{fig:SSprior}
+\scalebox{0.8}{%
 \begin{tikzpicture} [
@@
 \end{tikzpicture}
+}
 \end{figure}
@@
 The two proposed algorithms provides different utilities: the EM-CD algorithm is specifically to fit BHAM with an expedited computation, recommended to use in high and ultra-high dimensional setting.
-A specific concern of EM-CD algorithm is that it provides no information for inference.
-In contrast, the EM-IWLS can extract the variance-covariance matrix and be used to construct Wald type inference.
+A shortcoming of EM-CD algorithm is that it provides no information for uncertainty inference.
+In contrast, the EM-IWLS can extract the variance-covariance matrix of the coefficients and be used to construct Wald type inference.
 <!-- Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
@@
 
-We use the EM algorithm to find the MAP estimate of $\Theta$.
-Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithm.
-Naturally, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
+As the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithm.
+In the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
 $$
@@
 In the M-step, we find the $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
-The E- and M- steps are iterated until the algorithm converge.
+The E- and M- steps are iterated until the algorithm converges.
 
@@
 $Q_1$ and $Q_2$ are respectively the log posterior density of the coefficients $\bs \beta$ and the log posterior density of the probability parameters $\bs \theta$ conditioning on $\bs \gamma$.
-Meanwhile, conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
-In the optimization step, $Q_1$ can be treated as penalized likelihood function and solved via CD algorithm or IWLS algorithm.
+They are conditionally independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
+In the optimization step, $E(Q_1)$ can be treated as penalized likelihood function and solved via CD algorithm or IWLS algorithm.
 Maximization of $E(Q_2)$ can be solved via closed form equations following the beta-binomial conjugate relationship.
@@
 The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the value of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
-The calculation follows the equations below.
+The calculation follows the equations below,
 \begin{align*}
@@
 \end{align*}
-and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = \pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+where  $\pr(\gamma_{j}^{0} = 1|\theta_j) = \pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = \pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
@@
 \end{align}
-After replacing withe calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm .
+After substituting the calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm.
 Independently, the remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
@@
 <!-- \vspace{\baselineskip} -->
- \textit{EM-Iterative Weighted Least Squares} The iterative weighted least squares algorithm is a classifical algorithm to fit generalized linear models.
-Here we combine IWLS and EM algorithm to fit BHAM.
+ \textit{EM-Iterative Weighted Least Squares} The iterative weighted least squares algorithm is a classic algorithm to fit generalized linear models. Here we combine IWLS and EM algorithm to fit BHAM.
 
@@
 $2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$.
-That allows to formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal.
+That allows to formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix of $E(\tau^2)$s.
 The calculation of $E(\tau^2)$s is similar to the calculation of unknown quantities $E(S^{-1})$ in EM-CD via Bayes' theorem.
@@
 In each iteration of the EM-algorithm, we can approximate the generalized model likelihood with a weighted multivariate normal density following the classical IWLS, and augment the multivariate normal density of the coefficients $\bs \beta$.
-Least squares algorithm is used to derive the per-iteration estimate of the coefficients.
+Least squares algorithm is used to derive the estimate of the coefficients in each iteration.
 The rest of the steps follows EM-CD algorithm.
@@
 \label{sec:tune}
-Our proposed models, BHAM, require two preset scale parameters ($s_0$, $s_1$).
-Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding a criteria of preference.
-This would be achieved by constructing a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
-However, previous research suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value, e.g.
+Our proposed models, BHAMs, require two preset scale parameters ($s_0$, $s_1$).
+Hence, we need to find the optimal values such that the model reaches its best prediction performance regarding a criteria of preference.
+This is achieved by constructing a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
+However, previous research suggested the value of slab scale $s_1$ have less impact on the final model and to be set as a generally large value, e.g.
 $s_1 = 1$, that provides no or weak shrinkage.
 [@Rockova2018] As a result, we focus on examining different values of spike scale $s_0$.
-Instead of the 2-D grid, We consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
+Instead of the 2-D grid, We consider a sequence of $L$ increasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
 Increasing the spike scale $s_0$ tends to include more non-zero coefficients in the model.

### 2021-10-01 Manuscript/03-Simulation.Rmd
ENAR Submission Version

@@
 
-In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021].
+In this section, we conduct Monte Carlo simulation to compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021].
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
@@
 COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence used the default choices.
-Both mgcv and proposed models allow customized smoothing functions and we chose the cubic regression spline.
+Both mgcv and proposed models allow customized smoothing functions and we chose the cubic regression spline in this study.
 We controlled the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
@@
 We follow the data generating process described in Bai [@Bai2021].
-We first generated $n=500$ training data points with $p=4, 10, 50, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
-We then simulated the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
+We first generat $n=500$ training data points with $p=4, 10, 50, 200$ predictors respectively, where the predictors $\bs X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
+We then simulat the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
 The mean of each outcome were simulated via the following function
 $$
-\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
+\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3)).
 $$
-for gaussian and binomial outcomes.
+<!-- for gaussian and binomial outcomes. -->
 <!-- and shrinked effects -->
@@
 Gaussian outcomes required specification of dispersion, where we set the dispersion parameter to be 1.
-In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest covariates are inactive, i.e.
-$f_j(x_j) = 0$ for $j = 4, \dots, p$.
+In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active predictors, while the rest predictors are inactive, i.e.
+$B_j(x_j) = 0$ for $j = 4, \dots, p$.
 Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data.
-We generated 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
+We generated 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit and tune the models and testing datasets are used to calculate assessment measures.
 <!-- The dimensional of the smoothing functions are controlled as k=10. Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.  -->
 <!-- We also run the simulation where the number of bases are increased to 25 deliberately. -->
+
+To evaluate the  predictive performance of the models, the statistics calculated out of the testing datasets, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
 
@@
 
-To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
-The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
-
 ```{r, fig.height=8, eval = FALSE}
@@
 
-The predictive performances have a consistent pattern across the three distributions of outcomes.
+The predictive performances have a consistent pattern across the two distributions of outcomes.
 Across all the scenarios, COSSO and adaptive COSSO have the least favorable performance among the applicable methods examined.
 To note, mgcv doesn't support high-dimensional analysis when the number of coefficients are greater than the sample size.
-mgcv predicts well when $p$ is small or moderate (p = 4, 10), and deteriorate when the number of predictors increase.
+mgcv predicts well when $p$ is small or moderate (p = 4, 10), and its performance deteriorates when the number of predictors increase.
 Among the three fast-computing Bayesian hierarchical models, the proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of are small, moderate, and high.
-However, in the hyper high dimensional case, SB-GAM has better performance.
-Among the proposed methods, BHAM-IWLS performs consistently better than BHAM-CD.
+However, when $p$ increase to 200 for the extremely sparse-signal setting, SB-GAM has better performance.
+Among the proposed methods, BHAM-IWLS performs consistently better than BHAM-CD at the cost of computation time.
 

### 2021-10-01 Manuscript/04-Real_Data.Rmd
ENAR Submission Version

@@
 We apply the proposed models BHAM, fitted with the EM-CD algorithm ,to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
-We present the analysis results for the Emory Cardiovascular Biobank data [@Mehta2020], and include the analysis of Weight Loss Maintenance Cohort  data in Appendix A.2.
-We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
+We present the analysis result for the Emory Cardiovascular Biobank data [@Mehta2020], and include the analysis of Weight Loss Maintenance Cohort  data [@Bihlmeyer2021] in Appendix A.2.
+We observe a improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
 Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}, where we see the proposed model are more computationally efficient.
 
-```{r results = "asis"}
+\begin{table}[ht]
+\centering
+\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.}
+\label{tab:time_real_data}
+\begin{tabular}{lcccccc}
+  \hline
+  & & BHAM-CD & & & SB-GAM &\\
+  \cmidrule{2-4}  \cmidrule{5-7}
+Data & CV & Final & Total & CV & Final & Total \\ 
+  \hline
+ECB & 225.23 & 3.02 & 228.25 & 3506.39 & 34.36 & 3540.75 \\ 
+  WLM & 1207.98 & 8.03 & 1216.01 & 3116.00 & 32.68 & 3148.68 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+```{r eval = F, results = "asis"}
 tbl_real_time %>% 
@@
 ## Emory Cardiovascular Biobank
-We use the proposed models BHAM-SSL to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
+We use the proposed models BHAM to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
 The dataset is publicly available via \textit{Dryad} [@Mehta2020_data].
@@
 Hence, we performed a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and chose the the top 200 features with smallest p-values.
-5-Knot spline additive models for binary outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
-10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+5-Knot spline additive models for binary outcome are constructed using two different models, the proposed BHAM and the SB-GAM.
+10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the softwares.
 Out-of-bag samples were used for prediction performance evaluation, where deviance, area under the ROC curve (AUC), mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
@@
 rbind(
-  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'BHAM' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
   'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
@@
   rownames_to_column("Methods") %>% 
+  select(-mae) %>% 
+  rename(Deviance = deviance,
+         AUC = auc,
+         MSE = mse,
+         Misclass = misclassification) %>% 
   xtable(
-    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation",
+    caption = "Prediction performance of optimal BHAM and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), mean squared error (MSE), and misclassification (Misclass)",
     label = "tab:ECB_res"

### 2021-10-01 Manuscript/05-Discussion.Rmd
ENAR Submission Version

@@
 \label{sec:concl}
-In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.
-Meanwhile, we introduced two optimization based algorithms for model fitting.
+In this article, we describe a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.
+Meanwhile, we introduce two optimization-based algorithms for model fitting.
 The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner, but also provide naive inferential supports for many commonly used hierarchical priors, e.g.
 t-distribution, double exponential distribution, mixture t distribution, and mixture double exponential distribution, in the analyses that the sparse assumption is weak or not necessary.
-Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.
+Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art nonlinear modeling tools.
 

### 2021-10-13 Manuscript/SS_GAM.tex
Add BHAM IWLS model

--- 2021-09-30 Manuscript/SS_GAM.tex
@@
 
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{152.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{154.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,143.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{8.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1,151.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{5.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{488.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
 
@@
 
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{669.375}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.077}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.924}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.745}}} \\
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 1.07in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-SSL}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.9in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{665.628}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.065}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.936}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.67in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.754}}} \\
 

### 2021-10-13 _targets.R

@@ -148,22 +148,31 @@ tar_plan(
   #               include = starts_with("cov_")) %>% 
   #   add_overall(),
   
-  # ** BHAM --------------------------------------------------------------
+  # ** BHAM-CD --------------------------------------------------------------
   
   ## Prepare dat
   WLM_bgam_dat = make_bgam_dat(WLM_train_dat, test_dat = WLM_test_dat),
   WLM_bamlasso_raw = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
-                              ss=c(0.005, 0.05),
+                              ss=c(0.005, 0.5),
                               group = WLM_bgam_dat$group),
   
-  WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw,s0 = seq(0.005, 0.1, 0.01)),
+  WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw, s0 = seq(0.005, 0.1, 0.01)),
 
   
   WLM_bamlasso_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
-                          ss=c(0.025, 0.05),
+                          ss=c(0.025, 0.5),
                           group = WLM_bgam_dat$group),
   # measure.bh(bglm_mdl_fnl),
   WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),
+  
+  # ** BHAM-IWLS --------------------------------------------------------------
+  WLM_bgam_raw = bgam(out_HOMA_std~.-out_HOMA_std, 
+                      data = data.frame(out_HOMA_std =  WLM_train_dat$out_HOMA_std, WLM_bgam_dat$train_dat),
+                      family = "gaussian",
+                              prior = mde(),
+                              group = WLM_bgam_dat$group),
+  
+  WLM_bgam_cv = tune.bgam(WLM_bgam_raw,s0 = seq(0.005, 0.1, 0.01)),
   
   # ** SB_GAM ---------------------------------------------------------------
   

### 2021-10-13 _targets.R
Remove Unnecessary files

@@ -233,8 +233,8 @@ tar_plan(
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
-  tar_render(manu_ENAR, "Manuscript/00-main_ENAR.Rmd",
-             output_file = "SS_GAM-ENAR.pdf"),
+  # tar_render(manu_ENAR, "Manuscript/00-main_ENAR.Rmd",
+  #            output_file = "SS_GAM-ENAR.pdf"),
   
 )
 

### 2021-10-13 Manuscript/00-main.Rmd
Revise based on Comments

@@
 - num: b
-  org: Department of Biostatistics and Data Science, Wake Forest University, Winston-Salem, USA
+  org: Department of Biostatistics and Data Science, Wake Forest School of Medicine, Winston-Salem, USA
 corres: "Boyi Guo and Nengjun Yi, Department of Biostatistics, University of Alabama at Birmingham, Birmingham, USA. \\email{boyiguo1@uab.edu},  \\email{nyi@uab.edu}"
@@
 accepted: 2021-03-01
-abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate the high-dimensional ($p>>n$) problem using group penalties and priors.
-These models rarely smooth the nonlinear effect properly as the sparse regularization leads to excess shrinkage on coefficients and inaccurate prediction.
+abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate high-dimensional data ($p>>n$) using group sparse regularization.
+However, the sparse reuglarization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance.
 Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
 While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution: we propose a novel spike-and-slab spline prior for smoothing functions, where a reparameterization step facilitates smoothing penalty and separate selection of linear and nonlinear spaces.
-Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed to fit the proposed model.
-Simulation studies and metabolomics data analyses demonstrate improved performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing function.
+A novel spike-and-slab spline prior is used to select the components of smoothing functions.
+Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for different utilities.
+Simulation studies and metabolomics data analyses demonstrate improved predictive or computational performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM.
 The software implementation of the proposed models is freely available via an R package BHAM."
@@
 
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\caption{Directed acyclic graph of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
 \label{fig:SSprior}
@@
 
+\newpage
+
 ```{r, results = "asis"}
 tab_gaus$mgcv[4] <- "-"
+tab_gaus$mgcv[5] <- "-"
 tab_gaus %>% 
@@
   # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
+  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.\\label{tab:gaus}")
 ```
+
+\newpage
 
@@
 tab_binom$mgcv[4] <- "-"
-
+tab_binom$mgcv[5] <- "-"
 tab_binom %>% 
@@
   ) %>%  
-  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
+  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations.", label = "tab:bin_auc") %>% 
   print(comment = FALSE, include.rownames = FALSE)
 ```
+
+\newpage
 
@@
 
-```{r}
+\newpage
+
+```{r results = "asis"}
 rbind(
@@
   rownames_to_column("Methods") %>%
-  flextable() %>% 
-  flextable::autofit() %>% 
-  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}")
+  select(-mae) %>% 
+  rename(Deviance = deviance,
+         AUC = auc,
+         Brier = mse,
+         Misclass = misclassification) %>% 
+  xtable(
+    caption = "Prediction performance of optimal BHAM and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
+    label = "tab:ECB_res"
+  ) %>% 
+  print(comment = FALSE, include.rownames = FALSE#, caption.placement = "top"
+        )
 ```
 
-```{r}
+\newpage
+
+```{r results = "asis"}
 rbind(
@@
   rownames_to_column("Methods") %>%
-  flextable() %>% 
-  flextable::autofit() %>% 
-  flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
+  rename(
+    Deviance = deviance,
+    MSE = mse,
+    MAE = mae,
+    `$R^2$` = R2
+  ) %>% 
+    xtable(
+    caption = "Prediction performance of of optimal BHAM and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
+    label = "tab:WLM_res"
+  ) %>%
+  print(comment = FALSE, include.rownames = FALSE, sanitize.text.function=function(x){x}#, caption.placement = "top"
+        )
+  # flextable() %>% 
+  # flextable::autofit() %>% 
+  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
 ```

### 2021-10-13 Manuscript/01-Intro.Rmd
Revise based on Comments

--- 2021-09-27 Manuscript/01-Intro.Rmd
@@
 More complex models to address nonlinear effects include random forest and other so-called "black box" models [@Breiman2001].
-These models are not particularly attractive in biomedical studies, as they lack the transparency to understand the etiology of diseases.
-In addition, how to generalized these models to the high-dimensional setting remains unclear.
+These models are useful for statistical prediction but do not estimate parameters relevant to the data generation process that one can draw inferences from.
+In addition, how to generalize these models to the high-dimensional setting remains unclear.
 
@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution, denoted by $EF(\mu, \phi)$, with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
 $$
@@
 [@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
-As these methods focus on addressing the sparsity, it is very likely to overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
+Focus on addressing the sparsity, these methods can overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
 [@Scheipl2013] In addition, these methods adapt an 'all-in-all-out' strategy, i.e.
@@
 The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e.
-the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in many ways.
+the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in three ways.
 First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM.
@@
 The applications of the two aforementioned algorithms are fruitful in Bayesian hierarchical models for high-dimensional data.
-[@Yi2012; @Rockova2014a; @Rockova2018] Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides uncertainty measures that could be further used for confidence bound calculation and hypothesis testing.
+[@Yi2012; @Rockova2014a; @Rockova2018] Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides variance-covariance matrix of the coefficients, which has the potential to provide uncertainty inference.
 EM-IWLS complements the lack of inference potential from EM-CD in a median and high dimensional setting.
 Furthermore, the proposed model addresses the incapability of bi-level selection in SB-GAM.
-Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
+Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
 
 In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the two fast-fitting algorithms.
-In Section \ref{sec:sim}, we compare the proposed framework to state-of-art models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
+In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
 Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.

### 2021-10-13 Manuscript/02-Method.Rmd
Revise based on Comments

--- 2021-09-30 Manuscript/02-Method.Rmd
@@
 
-The re-parameterization step sets up the foundation of the proposed model and provides three-fold benefits.
+The re-parameterization step sets up the foundation of the proposed model and provides three benefits.
 First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection.
@@
 
-## Spike-and-Slab Smooth Priors
+## Spike-and-Slab Spline Priors
 
@@
 
-### Spike-and-Slab Lasso Prior
+### Two-part Spike-and-Slab Lasso Prior
 
@@
 The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
-Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
+Given the re-parameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
 Specifically, we impose independent group priors on the linear space coefficients and on the nonlinear space coefficients respectively,
@@
 The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
-First of all, the smoothing function bases are linear dependent and ask for extra attention.
+First of all, the smoothing function bases are linear dependent and necessitate extra attention.
 The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
@@
 
-To recall, the density function of spike-and-slab mixture double exponential prior can be written as
+The density function of spike-and-slab mixture double exponential prior can be written as
 $$
@@
 The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
-In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary.
+In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior
+<!-- , with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary -->
+.
 
@@
 $E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
-As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}) via Bayes' Theorem.
-On the other hand, $\tau^{2}$ following a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}).
+On the other hand, $\tau^{2}$ follows a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
 $E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation

### 2021-10-13 Manuscript/03-Simulation.Rmd
Revise based on Comments

--- 2021-09-30 Manuscript/03-Simulation.Rmd
@@
 We follow the data generating process described in Bai [@Bai2021].
-We first generated $n=500$ training data points with $p=4, 10, 50, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
+We first generated $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
 We then simulated the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
@@
 
-```{r, results = "asis"}
-tab_gaus$mgcv[4] <- "-"
-tab_gaus %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>% 
-  xtable(
-    align = "cccccccc",
-    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
-    label = "tab:gaus") %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
-  # flextable(
-  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
-  # ) %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations\\label{tab:gaus}")
-```
-
-```{r, results = "asis"}
-tab_binom$mgcv[4] <- "-"
-
-tab_binom %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>%  
-  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations", label = "tab:bin_auc") %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
-```
-
-To evaluate the  predictive performance of the models, the statistics caluclated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+To evaluate the  predictive performance of the models, the statistics calculated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
 The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
-
+\begin{center}
+Table \ref{tab:gaus} here\\
+Table \ref{tab:bin_auc} here\\
+\end{center}
 ```{r, fig.height=8, eval = FALSE}
@@
 
-<!-- Overall, the proposed priors outperforms the state-of-art GAM model `mgcv` when the dimensionality is low, median, and high.
+<!-- Overall, the proposed priors outperforms the state-of-the-art GAM model `mgcv` when the dimensionality is low, median, and high.
 The proposed priors predicts better than `mgcv` and infers better of the underlying smooth function.

### 2021-10-13 Manuscript/04-Real_Data.Rmd
Revise based on Comments

--- 2021-09-30 Manuscript/04-Real_Data.Rmd
@@
 Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
-
-```{r results = "asis"}
-tbl_real_time %>% 
-  xtable(
-    caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.",
-    label= "tab:time_real_data"
-  ) %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
-
-  # flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
-```
+\begin{center}
+Table \ref{tab:time_real_data} here
+\end{center}
 
@@
 Hence, we performed a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and chose the the top 200 features with smallest p-values.
-5-Knot spline additive models for binary outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
+We used 5-knot spline additive models for binary outcome using two different models, the proposed BHAM-SSL and the SB-GAM.
 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, area under the ROC curve (AUC), mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
-
-```{r results = "asis"}
-rbind(
-  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
-  'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
-) %>% 
-  rownames_to_column("Methods") %>% 
-  xtable(
-    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation",
-    label = "tab:ECB_res"
-  ) %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
-
-  # flextable::flextable() %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation\\label{tab:ECB_res}") %>% 
-  # flextable_to_rmd()
-```
+Out-of-bag samples were used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
+\begin{center}
+Table \ref{tab:ECB_res} here
+\end{center}
 
 ## Weight Loss Maintenance Cohort 
-We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publically available [@Bihlmeyer2021_data].
+We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
 In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
@@
 The outcome of interest was standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
-After removing missing datapoints and addressing outliers in the data, there were `r nrow(WLM_train_dat)` remaining in the analysis.
+After removing missing datapoints and addressing outliers in the data, there were $p$=`r nrow(WLM_train_dat)` features remaining in the analysis.
 5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean abosolute error (MAE)defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
-
-```{r results = "asis"}
-rbind(
-  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
-  'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
-) %>% 
-  rownames_to_column("Methods")  %>% 
-  xtable(
-    caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation",
-    label = "tab:WLM_res"
-  ) %>% 
-  print(comment = FALSE, include.rownames = FALSE, caption.placement = "top")
-
-  # flextable() %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
-```
+Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
+\begin{center}
+Table \ref{tab:WLM_res} here
+\end{center}

### 2021-10-13 Manuscript/05-Discussion.Rmd
Revise based on Comments

--- 2021-09-11 Manuscript/05-Discussion.Rmd
@@
 
-For transnational science purpose, we implemented the proposed model in a R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
+For translational science purpose, we implemented the proposed model in a R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
 To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones.
@@
 Nevertheless, this set-up could result in a situation that is not theoretically sound: the non-linear component is selected, but the linear component is not.
-We currently address it analytically by forcing to include the linear component when non-linear component is included.
-Another possible solution is to impose an dependent structure of $\gamma_{j}^{pen}$ on $\gamma_{j^{0}}$, i.e.
-$\gamma_j^{pen}|\gamma_{j}^{0}, \theta_j$.
+We currently address it analytically by including the linear component in the model when non-linear component is selected.
+Another possible solution is to impose a dependent structure of $\gamma_{j}^\tp$ on $\gamma_{j^{0}}$, i.e.
+$\gamma_j^\tp|\gamma_{j}^{0}, \theta_j$.
 Secondly, while the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
 Our current effort resides on expanding GAM to Cox proportional hazard models.
-An naive approach would be convert an time-to-event outcome to a poisson outcome following Whitehead [@Whitehead1980].
-However, it would be more efficient to directly fitting Cox models via penalized pseudo likelihood function.
+An naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].
+However, it would be more efficient to directly fit Cox models via penalized pseudo likelihood function [@Simon2011].
 Last but not least, we plan to expand the proposed model to integrate external biology information.

### 2021-10-13 R/Sim_Res/make_sim_main_table.R

@@ -20,13 +20,13 @@ make_sim_main_table <- function(success_rate, dist, section = "test" , measures)
         return(data.frame(NULL))
       
       # .file <- fls[1]
-      
       ret <- fls %>%
         map_dfr(.f = function(.file){
           it <- unglue_data(.file, "{whatever}/it_{it}.rds") %>% pull(it) %>% as.numeric
           ret <- data.frame(
             it = it,
             read_rds(.file) %>% 
+              `[[`(section)%>% 
               data.frame() %>% 
               rownames_to_column("method")
             
@@ -47,8 +47,9 @@ make_sim_main_table <- function(success_rate, dist, section = "test" , measures)
     )
   
   total_dat %>% 
-    select(p, method, 
-           {{measures}} := paste(section, measures, sep=".")) %>% #head
+    select(p, method, {{measures}}
+           # {{measures}} := paste(section, measures, sep=".")
+           ) %>% #head
     filter(method %in% c("bglm_spline_de", "blasso_spline",
                          "cosso", "acosso", "mgcv", "SB_GAM")) %>% 
     group_by(p, method) %>% 

### 2021-10-13 _targets.R

@@ -22,18 +22,18 @@ tar_plan(
   
   #  Simulations Studies -----------------------------------------------------
   #* Main Simulation Study ####
-  # tar_files(sim_main_path,
-  #           list.files("./Simulation/main", recursive = T, full.names = T)),
+  tar_files(sim_main_path,
+            list.files("~/Downloads/main/", recursive = T, full.names = T)),
+
+  success_rate = create_success_rate_table(sim_main_path),
   # 
-  # success_rate = create_success_rate_table(sim_main_path),
+  tab_binom = make_sim_main_table(success_rate,
+                                  dist = "binomial",
+                                  measures = "auc"),
   # 
-  # tab_binom = make_sim_main_table(success_rate,
-  #                                 dist = "binomial",
-  #                                 measures = "auc"),
-  # 
-  # tab_gaus = make_sim_main_table(success_rate,
-  #                                dist = "gaussian",
-  #                                measures = "R2"),
+  tab_gaus = make_sim_main_table(success_rate,
+                                 dist = "gaussian",
+                                 measures = "R2"),
   # 
   # tab_pois = make_sim_main_table(success_rate,
   #                                dist = "poisson",

### 2021-10-14 Manuscript/00-main.Rmd
Revise Manuscript

@@
 tar_load(tbl_real_time)
+tar_load(tab_time)
 ```
@@
     align = "cccccccc",
-    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations",
+    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.",
     label = "tab:gaus") %>% 
@@
   # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.\\label{tab:gaus}")
+  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measures for Gaussian outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+\\label{tab:gaus}")
 ```
@@
   ) %>%  
-  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations.", label = "tab:bin_auc") %>% 
+  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.", label = "tab:bin_auc") %>%
   print(comment = FALSE, include.rownames = FALSE)
+```
+
+\newpage
+```{r}
+tab_time %>% 
+  select(dist, p, `BHAM-IWLS`,`BHAM-CD`, cosso, acosso, mgcv, `SB-GAM`) %>% 
+  rename(Distribution = dist,
+         P = p,
+         COSSO = cosso,
+         `Adaptive COSSO` = acosso) %>%
+
+  flextable::flextable() %>% 
+    merge_v(j = 1) %>% 
+  flextable::set_caption(caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.\\label{tab:time_sim}")
 ```
@@
 rbind(
-  'BHAM-SSL' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'BHAM-CD' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
   'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
@@
   xtable(
-    caption = "Prediction performance of optimal BHAM and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
+    caption = "Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
     label = "tab:ECB_res"
@@
 rbind(
-  'BHAM-SSL' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'BHAM-CD' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
   'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
@@
     xtable(
-    caption = "Prediction performance of of optimal BHAM and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
+    caption = "Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
     label = "tab:WLM_res"

### 2021-10-14 Manuscript/02-Method.Rmd
Revise Manuscript

@@
 
-Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model and its matrix form,
+Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model with link function $g(\cdot)$ and linear predictor
 \begin{equation}\label{eq:gam}
-E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j,
 \end{equation}
-and the data distribution is
+with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The outcome $Y$ follows a exponential family distribution with density function $f(y)$, mean $\mu = g^{-1}(\eta)$ and dispersion parameter $\phi$, and the data distribution is
 \begin{equation}
@@
 \end{equation}
-with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e.
+ The basis function matrix, i.e.
 the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
@@
 
-The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection.
+The family of spike-and-slab (SS) priors regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
 We defer to Bai et al.
@@
 [@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.
+Two of most popular spike and slab priors are spike-and-slab normal prior [@George1993] and spike-and-slab double exponential prior[@Rockova2018].
 
@@
 The major criticism of early spike-and-slab models is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS based on spike-and-slab normal prior[@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
 Particularly, the development of spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
@@
 A specific concern of EM-CD algorithm is that it provides no information for inference.
-In contrast, the EM-IWLS can extract the variance-covariance matrix and be used to construct Wald type inference.
+In contrast, the EM-IWLS can estimate the variance-covariance matrix of the coefficients.
 Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
@@
 where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(1,1) density.
+We choose non-informative prior for the intercept $\beta_0$ and the dispersion parameter $\phi$; for example, $f(\beta_0|\tau_0^2)=N(0,\tau_0^2)$ with $\tau^2_0$ set to a large value and $f(\log \phi) \propto 1$.
 
@@
 $$
-The rest of the algorithm remains the same as described in Section \ref{sec:EMCD}.
+
+Totally, the proposed EM-iterative weighted least square algorithm is summarized as follows:
+
+1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$
+
+2) Iterate over the E-step and M-step until convergence
+
+    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E(\tau^{-2}_{j})$, $E({\tau^\tp}^{-2}_{jk})$ with estimates of $\Theta^{(t-1)}$ from the previous iteration
+
+    M-step:
+
+    a) Based on the current value of $\beta$, calculate the pseudo-data $z_i^{(t)}$ and the pseudo-weights $w_i^{(t)}$
+    b) Update $\bs \beta^{(t)}$ by runing the augmented weighted least squared
+    c) If $\phi$ is present, update $\phi$
+
+Similar to EM-CD, we assess convergence by the criterion, $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $\epsilon$ is a small value (say $10^{-5}$).
 

### 2021-10-14 Manuscript/04-Real_Data.Rmd
Revise Manuscript

@@
 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
+Out-of-bag samples were used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated.
+BHAM-CD obtained superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (Table \ref{tab:ECB_res}).
 \begin{center}
@@
 10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
+Out-of-bag samples were used for prediction performance evaluation, where deviance, $R^2$, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated.
+BHAM-CD obtained superior R2, MSE, and MAE in the out-of-bag samples compared to SB-GAM (Table \ref{tab:WLM_res}).
 \begin{center}

### 2021-10-14 R/Sim_Res/make_time_table.R

@@ -1,0 +1,72 @@
+make_time_table <- function(success_rate){
+  total_dat <- success_rate %>% 
+    # filter(dist==dist, n_success!=0) %>% 
+    # filter(dist=={{dist}}, n_success!=0) %>%
+    pull(path) %>% 
+    map_dfr( .f = function(sim){
+      
+      
+      sim.df <- unglue_data(sim, 
+                            "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
+        mutate( p = as.numeric(p))
+      fls <- list.files(sim, full.names = TRUE) 
+      fls <- fls[grep(".rds", x=fls)]
+      n <- length(fls)
+      
+      
+      if(n == 0)
+        return(data.frame(NULL))
+      
+      # .file <- fls[1]
+      ret <- fls %>%
+        map_dfr(.f = function(.file){
+          it <- unglue_data(.file, "{whatever}/it_{it}.rds") %>% pull(it) %>% as.numeric
+          ret <- data.frame(
+            it = it,
+            read_rds(.file) %>% 
+              `[[`("time")%>% 
+              data.frame() %>% 
+              rownames_to_column("method")
+            
+          )
+        }) %>%
+        arrange(it) %>% 
+        mutate(p = sim.df$p,
+               dist = sim.df$dist)#  %>% 
+      
+    }) %>% 
+    select(dist, p, it, method, elapsed) %>%
+    pivot_wider(names_from = method, values_from = elapsed) %>% 
+    transmute(dist = case_when(dist=="binomial"~"Binomial",
+                               dist=="gaussian"~"Gaussian"),
+              p, it, mgcv, cosso, acosso,
+              `BHAM-CD` = blasso_spline_cv + blasso_spline_final,
+              `BHAM-IWLS` = bglm_spline_de_cv + bglm_spline_de_final,
+              `SB-GAM` = bai_cv + bai_final) %>% 
+    group_by(dist, p) %>%  
+    summarize(
+      across(mgcv:`SB-GAM`, ~sprintf("%.2f (%.2f)", mean(.x), sd(.x))
+      )
+    ) %>% 
+    ungroup()
+  
+  total_dat[total_dat$p %in% c(100,200), "mgcv"] <- "-"
+  total_dat
+  
+  # 
+  # 
+  # 
+  # 
+  # 
+  # 
+  # %>% 
+  #   mutate(method = factor(
+  #     method,
[…]

### 2021-10-14 _targets.R

@@ -34,6 +34,7 @@ tar_plan(
   tab_gaus = make_sim_main_table(success_rate,
                                  dist = "gaussian",
                                  measures = "R2"),
+  tab_time = make_time_table(success_rate),
   # 
   # tab_pois = make_sim_main_table(success_rate,
   #                                dist = "poisson",

### 2021-10-15 Manuscript/00-main.Rmd
Revise abstract

@@
 Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
-While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo (MCMC) algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution: we incorporate the smoothing penalty that can separate linear and nonlinear spaces of smoothing function.
-A novel spike-and-slab spline prior is used to select the components of smoothing functions.
+While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability.
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we consider the smoothing penalty for proper shrinkage of curve interpolation and separation of smoothing function linear and nonlinear spaces.
+A novel spike-and-slab spline prior is proposed to select components of smoothing functions.
 Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for different utilities.

### 2021-10-15 Manuscript/01-Intro.Rmd
Revise Section 01

@@
 [@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
-Recent Bayesian hierarchical models are also used in the context of high-dimensional additive models.
+Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models.
 Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.
@@
 Scheipl et al.
-[@Scheipl2012] proposed a spike-and-slab structure prior, normal-mixture-of-inverse gamma prior, that address the previous challenges.
-But the model fitting relies on computational intensive MCMC algorithmic and creates scalability concern.
+[@Scheipl2012] proposed a spike-and-slab structure prior that address the previous challenges.
+But the model fitting relies on computational intensive MCMC algorithms and creates scalability concern.
 It would be of special interest to develop a fast, flexible and accurate generalized additive model framework.
@@
 Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-Moreover, incorporating the smoothing penalty allows the separation of the linear space of smoothing functions from the nonlinear space.
-We then impose a new bi-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
-The prior setup encourages a flexible solution, rending one of three possibilities: no effect, only linear effect, or nonlinear effect of a predictor.
-In addition, two scalable optimization-based algorithms are developed for high and ultrahigh-dimensional[@Fan2009] settings respectively.
+Moreover, incorporating the smoothing penalty allows the separation of the linear space of a smoothing function from the nonlinear space.
+We then impose a new two-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
+The prior setup encourages a flexible solution, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
+In addition, two scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented in an publicly available R pacakge \textt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
 
@@
 the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in three ways.
-First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM.
-Spike-and-slab lasso priors provide computational convenience during model fitting by using optimization algorithms instead of computational intensive sampling algorithms.
-They make fitting high- and ultrahigh-dimensional models more feasible without sacrificing performance in prediction and variable selection.
-Secondly, the optimization algorithms introduced in this article include EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm for different utilities.
-The applications of the two aforementioned algorithms are fruitful in Bayesian hierarchical models for high-dimensional data.
-[@Yi2012; @Rockova2014a; @Rockova2018] Even though SB-GAM with a group spike-and-slab lasso prior also uses EM-CD for model fitting, the proposed EM-IWLS provides variance-covariance matrix of the coefficients, which has the potential to provide uncertainty inference.
-EM-IWLS complements the lack of inference potential from EM-CD in a median and high dimensional setting.
-Furthermore, the proposed model addresses the incapability of bi-level selection in SB-GAM.
-Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via BHAM at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
+First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
+Spike-and-slab lasso priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms.
+They make fitting high-dimensional models more feasible without sacrificing performance in prediction and variable selection.
+Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-CD algorithm to fit the model.
+While both methods use the combination of expectation maximization algorithm and coordient descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.
+The proposed model sets up independent priors among basis function coefficients after the re-parameterization step, which provides some advantage in computation.
+In addition, we offer an alternative algorithm, EM-IWLS, that can provide variance-covariance matrix of the coefficients.
+Last but not least, the proposed model addresses the incapability of bi-level selection in SB-GAM.
 

### 2021-10-17 Manuscript/SS_GAM.tex
Upload tex file

--- 2021-09-30 Manuscript/SS_GAM.tex
@@
 \address[b]{Department of Biostatistics and Data Science, Wake Forest
-University, Winston-Salem, USA}
+School of Medicine, Winston-Salem, USA}
 
@@
 \abstract{There are proposals that extend the classical generalized
-additive models (GAMs) to accommodate the high-dimensional (\(p>>n\))
-problem using group penalties and priors. These models rarely smooth the
-nonlinear effect properly as the sparse regularization leads to excess
-shrinkage on coefficients and inaccurate prediction. Moreover, most of
-these GAMs consider an ``all-in-all-out'' approach for function
-selection, rendering them difficult to answer if nonlinear effects are
-necessary. While some Bayesian models can address these shortcomings,
-using Markov chain Monte Carlo (MCMC) algorithms for model fitting
-creates a new challenge, scalability. Hence, we propose Bayesian
-hierarchical generalized additive models as a solution: we propose a
-novel spike-and-slab spline prior for smoothing functions, where a
-reparameterization step facilitates smoothing penalty and separate
-selection of linear and nonlinear spaces. Two scalable and deterministic
-algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
-Squares, are developed to fit the proposed model. Simulation studies and
-metabolomics data analyses demonstrate improved performance against
-state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM. The
-software implementation of the proposed models is freely available via
-an R package BHAM.}
+additive models (GAMs) to accommodate high-dimensional data (\(p>>n\))
+using group sparse regularization. However, the sparse reuglarization
+may induce excess shrinkage when estimating smoothing functions,
+damaging predictive performance. Moreover, most of these GAMs consider
+an ``all-in-all-out'' approach for function selection, rendering them
+difficult to answer if nonlinear effects are necessary. While some
+Bayesian models can address these shortcomings, using Markov chain Monte
+Carlo (MCMC) algorithms for model fitting creates a new challenge,
+scalability. Hence, we propose Bayesian hierarchical generalized
+additive models as a solution: we incorporate the smoothing penalty that
+can separate linear and nonlinear spaces of smoothing function. A novel
+spike-and-slab spline prior is used to select the components of
+smoothing functions. Two scalable and deterministic algorithms,
+EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are
+developed for different utilities. Simulation studies and metabolomics
+data analyses demonstrate improved predictive or computational
+performance against state-of-the-art models, mgcv, COSSO and sparse
+Bayesian GAM. The software implementation of the proposed models is
+freely available via an R package BHAM.}
 
@@
 other so-called ``black box'' models \citep{Breiman2001}. These models
-are not particularly attractive in biomedical studies, as they lack the
-transparency to understand the etiology of diseases. In addition, how to
-generalize these models to the high-dimensional setting remains unclear.
+are useful for statistical prediction but do not estimate parameters
+relevant to the data generation process that one can draw inferences
+from. In addition, how to generalize these models to the
+high-dimensional setting remains unclear.
 
@@
 In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution, denoted by \(EF(\mu, \phi)\), with mean
-\(\mu\) and dispersion \(\phi\), can be modeled with the summation of
-smoothing functions, \(B_j(\cdot), j = 1, \dots, p\), of a given
-\(p\)-dimensional vector of covariates \(\boldsymbol{x}\), written as \[
+exponential family distribution with mean \(\mu\) and dispersion
+\(\phi\), can be modeled with the summation of smoothing functions,
+\(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional vector of
+covariates \(\boldsymbol{x}\), written as \[
  E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
@@
 optimization algorithm, and further generalized the framework to
-GAMs\citep{Bai2021}. As these methods focus on addressing the sparsity,
-it is very likely to overly penalize the basis function coefficients and
-produce inaccurate predictions and curve interpolation, particularly
-when complex signals are assumed and large number of knots are used.
-\citep{Scheipl2013} In addition, these methods adapt an `all-in-all-out'
-strategy, i.e.~either including or excluding the variable completely,
-rendering no space for bi-level selection. Scheipl et al.
-\citep{Scheipl2012} proposed a spike-and-slab structure prior,
-normal-mixture-of-inverse gamma prior, that address the previous
-challenges. But the model fitting relies on computational intensive MCMC
-algorithmic and creates scalability concern. It would be of special
-interest to develop a fast, flexible and accurate generalized additive
-model framework.
+GAMs\citep{Bai2021}. Focus on addressing the sparsity, these methods can
+overly penalize the basis function coefficients and produce inaccurate
+predictions and curve interpolation, particularly when complex signals
+are assumed and large number of knots are used. \citep{Scheipl2013} In
+addition, these methods adapt an `all-in-all-out' strategy, i.e.~either
+including or excluding the variable completely, rendering no space for
+bi-level selection. Scheipl et al. \citep{Scheipl2012} proposed a
+spike-and-slab structure prior, normal-mixture-of-inverse gamma prior,
+that address the previous challenges. But the model fitting relies on
+computational intensive MCMC algorithmic and creates scalability
+concern. It would be of special interest to develop a fast, flexible and
+accurate generalized additive model framework.
 
@@
 GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in many ways. First of all, the proposed spike-and-slab
+\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
 spline prior is a spike-and-slab lasso type prior using independent
@@
 Even though SB-GAM with a group spike-and-slab lasso prior also uses
-EM-CD for model fitting, the proposed EM-IWLS provides uncertainty
-measures that could be further used for confidence bound calculation and
-hypothesis testing. EM-IWLS complements the lack of inference potential
-from EM-CD in a median and high dimensional setting. Furthermore, the
-proposed model addresses the incapability of bi-level selection in
-SB-GAM. Lastly, the R implementation of the Bayesian hierarchical
-additive model is available publicly via BHAM at
-\url{https://github.com/boyiguo1/BHAM}, making transnational science
+EM-CD for model fitting, the proposed EM-IWLS provides
+variance-covariance matrix of the coefficients, which has the potential
+to provide uncertainty inference. EM-IWLS complements the lack of
+inference potential from EM-CD in a median and high dimensional setting.
+Furthermore, the proposed model addresses the incapability of bi-level
+selection in SB-GAM. Lastly, the R implementation of the Bayesian
+hierarchical additive model is available publicly via BHAM at
+\url{https://github.com/boyiguo1/BHAM}, making translational science
 more accessible.
@@
 priors, and describe the two fast-fitting algorithms. In Section
-\ref{sec:sim}, we compare the proposed framework to state-of-art models,
-mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
-Analyses of two metabolomics datasets are presented in Section
+\ref{sec:sim}, we compare the proposed framework to state-of-the-art
+models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation
+studies. Analyses of two metabolomics datasets are presented in Section
 \ref{sec:real_data}. Conclusion and discussions are given in Section
@@
 Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model and its matrix form,
-\begin{equation}\label{eq:gam}
-E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
-\end{equation} and the data distribution is \begin{equation}
+have a generalized additive model with link function \(g(\cdot)\) and
+linear predictor \begin{equation}\label{eq:gam}
+\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
+\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
+\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
+distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
+and dispersion parameter \(\phi\), and the data distribution is
+\begin{equation}
 f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
-matrix derived from the smoothing function \(B_j(x_j)\), is denoted
+\end{equation} The basis function matrix, i.e.~the design matrix derived
+from the smoothing function \(B_j(x_j)\), is denoted
 \(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
@@
 The re-parameterization step sets up the foundation of the proposed
-model and provides three-fold benefits. First of all, the
-re-parameterization integrates the smoothing penalty into the design
-matrix, and encourages models to properly smooth the nonlinear function
-in addition to the sparse penalty for function selection. Secondly, the
-eigendecomposition of the smoothing penalty allows the isolation of the
-linear from the nonlinear space, improving the feasibility of bi-level
-function selection and inference. The eigendecomposition facilitates the
+model and provides three benefits. First of all, the re-parameterization
+integrates the smoothing penalty into the design matrix, and encourages
+models to properly smooth the nonlinear function in addition to the
+sparse penalty for function selection. Secondly, the eigendecomposition
+of the smoothing penalty allows the isolation of the linear from the
+nonlinear space, improving the feasibility of bi-level function
+selection and inference. The eigendecomposition facilitates the
 construction of orthonormal design matrix, which makes imposing
@@
 
-\hypertarget{spike-and-slab-smooth-priors}{%
-\subsection{Spike-and-Slab Smooth
-Priors}\label{spike-and-slab-smooth-priors}}
-
-The family of spike-and-slab (SS) priors regression models dominates
-Bayesian high-dimensional analysis for its utility in outcome prediction
-and variable selection. We defer to Bai et al. \citep{Bai2021Review} for
-an in-depth introduction to spike-and-slab priors. To summarize,
-spike-and-slab priors are a family of mixture distributions that
-comprises a skinny spike density \(f_{\text{spike}}(\cdot)\) for weak
-signals and a flat slab density \(f_{\text{slab}}(\cdot)\) for strong
-signals, mathematically \[
+\hypertarget{spike-and-slab-spline-priors}{%
+\subsection{Spike-and-Slab Spline
+Priors}\label{spike-and-slab-spline-priors}}
+
+The family of spike-and-slab (SS) priors regression models is one of
+most commonly used models in high-dimensional data analysis for its
+utility in outcome prediction and variable selection. We defer to Bai et
+al. \citep{Bai2021Review} for an in-depth introduction to spike-and-slab
+priors. To summarize, spike-and-slab priors are a family of mixture
+distributions that comprises a skinny spike density
+\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
+\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
  \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
@@
 Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
-\citep{Clyde2004} and reference therein.
+\citep{Clyde2004} and reference therein. Two of most popular spike and
+slab priors are spike-and-slab normal prior \citep{George1993} and
+spike-and-slab double exponential prior\citep{Rockova2018}.
 
@@
 studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS \citep{Rockova2014a} and
-spike-and-slab Lasso \citep{Rockova2018b, Rockova2018}. Particularly,
-the development of spike-and-slab Lasso substantially improves the
-scalability of SS models, setting up the theoretical foundation for
-generalized models in -omics data analysis
-\citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The spike-and-slab
-Lasso (SSL) prior is composed of two double exponential distributions
-with mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
-mathematically, \begin{equation} 
+algorithms bear, which include EMVS based on spike-and-slab normal
+prior\citep{Rockova2014a} and spike-and-slab Lasso
+\citep{Rockova2018b, Rockova2018}. Particularly, the development of
+spike-and-slab Lasso substantially improves the scalability of SS
+models, setting up the theoretical foundation for generalized models in
+-omics data analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}.
+The spike-and-slab Lasso (SSL) prior is composed of two double
+exponential distributions with mean 0 and different dispersion
+parameters, \(0 < s_0 < s_1\), mathematically, \begin{equation} 
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
@@
 
-\hypertarget{spike-and-slab-lasso-prior}{%
-\subsubsection{Spike-and-Slab Lasso
-Prior}\label{spike-and-slab-lasso-prior}}
+\hypertarget{two-part-spike-and-slab-lasso-prior}{%
+\subsubsection{Two-part Spike-and-Slab Lasso
+Prior}\label{two-part-spike-and-slab-lasso-prior}}
 
@@
 from the spike-and-slab lasso prior described in Equation
-(\ref{eq:ssl}). Given the re-parameterized deign matrix
+(\ref{eq:ssl}). Given the re-parameterized design matrix
 \(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
@@
 validity of the proposed prior. First of all, the smoothing function
-bases are linear dependent and ask for extra attention. The
+bases are linear dependent and necessitate extra attention. The
 eigeondecomposition remedies the problem and hence our prior can be set
@@
 algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can extract the variance-covariance matrix and be used to
-construct Wald type inference. Moreover, the EM-IWLS is a more general
-model fitting algorithm that can be used for fitting not only SSL and
-continuous SS priors but also Student's t-priors and double exponential
-priors. To note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD
-algorithm. The main difference between proposed EM-CD algorithm and that
-in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group
-prior, while the proposed prior is pairwise independent requiring no
-special treatment in the CD algorithm.
+the EM-IWLS can estimate the variance-covariance matrix of the
+coefficients. Moreover, the EM-IWLS is a more general model fitting
+algorithm that can be used for fitting not only SSL and continuous SS
+priors but also Student's t-priors and double exponential priors. To
+note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
+main difference between proposed EM-CD algorithm and that in SB-GAM is
+that SB-GAM uses a blocked CD algorithm for their group prior, while the
+proposed prior is pairwise independent requiring no special treatment in
+the CD algorithm.
 
@@
 where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density.
+distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
+non-informative prior for the intercept \(\beta_0\) and the dispersion
+parameter \(\phi\); for example, \(f(\beta_0|\tau_0^2)=N(0,\tau_0^2)\)
+with \(\tau^2_0\) set to a large value and \(f(\log \phi) \propto 1\).
 
@@
 
-To recall, the density function of spike-and-slab mixture double
-exponential prior can be written as \[
+The density function of spike-and-slab mixture double exponential prior
+can be written as \[
 f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
@@
 using IWLS in combination with EM algorithm. In this work, we adapt the
-EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior, with the
-full potential to adopt Student's t, double exponential and mixture
-Student's t priors when sparse assumption is not necessary.
+EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior .
 
@@
 \(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
-\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}) via
-Bayes' Theorem. On the other hand, \(\tau^{2}\) following a gamma
-distribution is a conjugate prior for the normal variance, and the
-conditional posterior density of \(\tau^{-2}\) is an inverse Gaussian
-distribution. \(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are
-calculated using the closed form equation \begin{align*}
+\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}). On
+the other hand, \(\tau^{2}\) follows a gamma distribution is a conjugate
+prior for the normal variance, and the conditional posterior density of
+\(\tau^{-2}\) is an inverse Gaussian distribution.
+\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
+using the closed form equation \begin{align*}
  E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
@@
   \text{Var}(\boldsymbol{\beta}^{(t)}) = (\boldsymbol{X}_*^T\boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\phi^{(t)}.
-\] The rest of the algorithm remains the same as described in Section
-\ref{sec:EMCD}.
+\]
+
+Totally, the proposed EM-iterative weighted least square algorithm is
+summarized as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
+\item
+  Iterate over the E-step and M-step until convergence
+
+  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
+  \(E(\tau^{-2}_{j})\), \(E({\tau^*}^{-2}_{jk})\) with estimates of
+  \(\Theta^{(t-1)}\) from the previous iteration
+
+  M-step:
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii})}
+  \tightlist
+  \item
+    Based on the current value of \(\beta\), calculate the pseudo-data
+    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
+  \item
+    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
+    least squared
+  \item
+    If \(\phi\) is present, update \(\phi\)
+  \end{enumerate}
+\end{enumerate}
+
+Similar to EM-CD, we assess convergence by the criterion,
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where \(\epsilon\) is
+a small value (say \(10^{-5}\)).
 
@@
 We first generated \(n=500\) training data points with
-\(p=4, 10, 50, 200\) predictors respectively, where the predictors \(X\)
-are simulated from a multivariate normal distribution
+\(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
+\(X\) are simulated from a multivariate normal distribution
 \(\text{MVN}_{n\times p}(0, I_{P})\). We then simulated the outcome
@@
 implemented in \texttt{mgcv}, and chose the the top 200 features with
-smallest p-values. 5-Knot spline additive models for binary outcome are
-constructed using two different models, the proposed BHAM-SSL and the
+smallest p-values. We used 5-knot spline additive models for binary
+outcome using two different models, the proposed BHAM-SSL and the
 SB-GAM. 10-Fold cross-validation(CV) are used to choose the optimal
@@
 selection criterion implemented in the software. Out-of-bag samples were
-used for prediction performance evaluation, where deviance, area under
-the ROC curve (AUC), mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassication error defined as
+used for prediction performance evaluation, where deviance, AUC, Brier
+score defined as \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\),
+and misclassication error defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) were
-calculated and tabulated in Table \ref{tab:ECB_res}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrrr}
-  \hline
-Methods & deviance & auc & mse & mae & misclassification \\ 
-  \hline
-BHAM-SSL & 455.69 & 0.74 & 0.16 & 0.28 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.26 & 0.24 \\ 
-   \hline
-\end{tabular}
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation} 
-\label{tab:ECB_res}
-\end{table}
+calculated. BHAM-CD obtained superior AUC, Brier score, and
+misclassification error in the out-of-bag samples compared to SB-GAM
+(Table \ref{tab:ECB_res}).
+
+\begin{center}
+Table \ref{tab:ECB_res} here
+\end{center}
 
@@
 between metabolic biomarkers and weight loss, where the dataset is
-publically available \citep{Bihlmeyer2021_data}. In this analysis, we
+publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
 primarily focused on the analysis of one of the three studies included,
@@
 resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there were 237
-remaining in the analysis. 5-Knot spline additive models for the
-gaussian outcome are constructed using two different models, the
-proposed BHAM-SSL and the SB-GAM. 10-Fold cross-validation(CV) are used
-to choose the optimal tuning parameters of each framework with respect
-to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation,
-where deviance, mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-abosolute error (MAE)defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) were calculated and
-tabulated in Table \ref{tab:WLM_res}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & deviance & R2 & mse & mae \\ 
-  \hline
-BHAM-SSL & 669.38 & 0.08 & 0.92 & 0.74 \\ 
-  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
-   \hline
-\end{tabular}
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation} 
-\label{tab:WLM_res}
-\end{table}
+missing datapoints and addressing outliers in the data, there were
+\(p\)=237 features remaining in the analysis. 5-Knot spline additive
+models for the gaussian outcome are constructed using two different
+models, the proposed BHAM-SSL and the SB-GAM. 10-Fold
+cross-validation(CV) are used to choose the optimal tuning parameters of
+each framework with respect to the default selection criterion
+implemented in the software. Out-of-bag samples were used for prediction
+performance evaluation, where deviance, \(R^2\), mean squared error
+(MSE) defined as \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\),
+and mean absolute error (MAE) defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) were calculated.
+BHAM-CD obtained superior R2, MSE, and MAE in the out-of-bag samples
+compared to SB-GAM (Table \ref{tab:WLM_res}).
+
+\begin{center}
+Table \ref{tab:WLM_res} here
+\end{center}
 
@@
 
-For transnational science purpose, we implemented the proposed model in
+For translational science purpose, we implemented the proposed model in
 a R package \texttt{BHAM}, deposited at
@@
 selected, but the linear component is not. We currently address it
-analytically by forcing to include the linear component when non-linear
-component is included. Another possible solution is to impose an
-dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
+analytically by including the linear component in the model when
+non-linear component is selected. Another possible solution is to impose
+a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
 i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, while the
@@
 expanding GAM to Cox proportional hazard models. An naive approach would
-be convert an time-to-event outcome to a poisson outcome following
+be convert a time-to-event outcome to a Poisson outcome following
 Whitehead \citep{Whitehead1980}. However, it would be more efficient to
-directly fitting Cox models via penalized pseudo likelihood function.
-Last but not least, we plan to expand the proposed model to integrate
-external biology information. The main motivation for the integrative
-model is that biologically informed grouping of weak effects increases
-the power of detecting true associations between features and the
-outcome \citep{Peterson2016}, and stabilizes the analysis results for
-reproducibility purpose. Such integration can be achieved by setting up
-a structural hyperprior on the inclusion indicator of the smoothing
-function null space \(\boldsymbol{\gamma}^0\). The similar strategy has
-been used in Ferrari and Dunson \citep{Ferrari2020}.
+directly fit Cox models via penalized pseudo likelihood function
+\citep{Simon2011}. Last but not least, we plan to expand the proposed
+model to integrate external biology information. The main motivation for
+the integrative model is that biologically informed grouping of weak
+effects increases the power of detecting true associations between
+features and the outcome \citep{Peterson2016}, and stabilizes the
+analysis results for reproducibility purpose. Such integration can be
+achieved by setting up a structural hyperprior on the inclusion
+indicator of the smoothing function null space
+\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
+and Dunson \citep{Ferrari2020}.
 
@@
 
+\newpage
+
+\begin{figure}
+\centering
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
+};
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
+};
+
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
+
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (Theta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+\begin{pgfonlayer}{background}
+  \node [background,
+   fit=(linearGamma) (linearBeta),
+   label=above:Linear Space:] {};
+  \node [background,
+    fit=(penGamma) (penBeta),
+    label=above:Nonlinear Space:] {};
+\end{pgfonlayer}
+
+\end{tikzpicture}
+
+\caption{Directed acyclic graph of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{ccccccc}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
+  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
+  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.}
+\label{tab:gaus}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllllll}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
+   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
+   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
+  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
+  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.}
+\label{tab:bin_auc}
+\end{table}
+
+\newpage
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.\label{tab:time_sim}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Distribution}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{P}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-IWLS}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{COSSO}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Adaptive\ COSSO}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mgcv}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-8}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Distribution}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{P}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-IWLS}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{COSSO}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Adaptive\ COSSO}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{mgcv}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-8}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{5.88\ (1.01)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{6.96\ (1.76)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.97\ (0.82)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4.60\ (2.32)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.18\ (0.04)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{343.24\ (88.37)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{14.01\ (1.86)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.65\ (1.25)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.95\ (2.70)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{10.01\ (3.56)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.51\ (12.03)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{543.13\ (133.51)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{50}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{185.26\ (17.44)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{53.83\ (6.26)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{109.72\ (24.31)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{128.13\ (23.26)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{670.01\ (151.28)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1630.42\ (193.61)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{100}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1106.93\ (356.73)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{48.41\ (5.84)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{715.15\ (170.55)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{718.63\ (193.53)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2783.99\ (235.31)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-5}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Binomial}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{200}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{6923.62\ (1911.59)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{99.98\ (7.05)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{5572.51\ (1295.18)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4958.69\ (1970.46)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4780.10\ (488.66)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{4.22\ (1.51)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2.51\ (0.20)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.79\ (0.11)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.67\ (0.07)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.05\ (0.00)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{36.02\ (3.02)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{10}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{17.59\ (5.77)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{33.32\ (3.50)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.49\ (0.70)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.47\ (0.71)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{0.32\ (0.34)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{72.93\ (9.52)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{50}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{329.83\ (44.57)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{334.53\ (22.33)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{35.54\ (9.15)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{35.44\ (9.02)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{73.33\ (71.24)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{373.42\ (34.93)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{100}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1706.19\ (225.92)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{446.98\ (79.31)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{150.58\ (46.72)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{152.82\ (45.07)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{684.05\ (67.80)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-5}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Gaussian}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{200}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{16379.27\ (4825.56)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{303.73\ (87.27)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{591.11\ (147.88)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{569.92\ (107.45)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{-}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{1314.12\ (137.85)}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-8}
+
+\end{longtable}
+
+\newpage
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{5.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{488.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\end{longtable}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Methods & Deviance & AUC & Brier & Misclass \\ 
+  \hline
+BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
+  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
+   \hline
+\end{tabular}
+\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
+\label{tab:ECB_res}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Methods & Deviance & $R^2$ & MSE & MAE \\ 
+  \hline
+BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
+  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
+   \hline
+\end{tabular}
+\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
+\label{tab:WLM_res}
+\end{table}
+
 \end{document}

### 2021-10-18 Manuscript/01-Intro.Rmd
Remove github links

--- 2021-10-01 Manuscript/01-Intro.Rmd
@@
 -->
-Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via \texttt{BHAM} at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making transnational science more accessible.
+Lastly, the R implementation of the Bayesian hierarchical additive model is available publicly via \texttt{BHAM}, making transnational science more accessible.
 

### 2021-10-18 _targets.R

@@ -153,14 +153,14 @@ tar_plan(
   ## Prepare dat
   WLM_bgam_dat = make_bgam_dat(WLM_train_dat, test_dat = WLM_test_dat),
   WLM_bamlasso_raw = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
-                              ss=c(0.005, 0.05),
+                              ss=c(0.005, 0.5),
                               group = WLM_bgam_dat$group),
   
   WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw,s0 = seq(0.005, 0.1, 0.01)),
 
   
   WLM_bamlasso_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
-                          ss=c(0.025, 0.05),
+                          ss=c(0.025, 0.5),
                           group = WLM_bgam_dat$group),
   # measure.bh(bglm_mdl_fnl),
   WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),

### 2021-10-18 Manuscript/00-main.Rmd
Revise Section 2

@@
 However, the sparse reuglarization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance.
-Moreover, most of these GAMs consider an “all-in-all-out” approach for function selection, rendering them difficult to answer if nonlinear effects are necessary.
+Moreover, most of these GAMs consider an “all-in-all-out” approach for functional selection, rendering them difficult to answer if nonlinear effects are necessary.
 While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability.

### 2021-10-18 Manuscript/02-Method.Rmd
Revise Section 2

@@
 the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
-The dimension of the design matrix depends on the choice of the smoothing function.
-For example, a cubic spline function with $k$ knots produces a $k+3$ elements column vector.
+The dimension of the design matrix depends on the choice of the smoothing function, and is denoted as $K_j$ for $x_j$.
 Its corresponding smoothing penalty matrix is denoted as $\bs S_j$.
 $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
-With slight abuse of notation, we denote vectors and matrices in bold fonts $\bs \beta, \bs X$ with conformable dimensions, where scalar and random variables are denoted in with unbold fonts $\beta, X$.
+With slight abuse of notation, we denote vectors and matrices in bold fonts $\bs \beta, \bs X$ with conformable dimensions, where scalar and random variables are denoted in unbold fonts $\beta, X$.
 The matrix transposing operation is denoted with a superscript $^T$.
@@
 
-To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models.
+To encourage proper smoothing of additive functions, we adopt the idea of smoothing penalties from smoothing spline models.
 The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
@@
 $$
-and the generalized additive model in Equation (\ref{eq:gam}) be
+and the generalized additive model in Equation (\ref{eq:gam}) now is
 \begin{equation}\label{eq:gam-repa}
@@
 The re-parameterization step sets up the foundation of the proposed model and provides three benefits.
-First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection.
-Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear from the nonlinear space, improving the feasibility of bi-level function selection and inference.
+First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for functional selection.
+Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear space from the nonlinear space, improving the feasibility of bi-level functional selection.
 The eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
@@
 
-The family of spike-and-slab (SS) priors regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
+The family of spike-and-slab (SS) regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
 We defer to Bai et al.
@@
 The major criticism of early spike-and-slab models is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS based on spike-and-slab normal prior[@Rockova2014a] and spike-and-slab Lasso [@Rockova2018b; @Rockova2018].
-Particularly, the development of spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
-The spike-and-slab Lasso (SSL) prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
+[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS based on spike-and-slab normal prior[@Rockova2014a] and spike-and-slab Lasso (SSL) [@Rockova2018b; @Rockova2018].
+Particularly, the development of SSL model  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
 \begin{equation} 
@@
 The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is preferred in high-dimensional data analysis.
-We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015].
-Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.
+We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015] and limits to an "all-in-all-out" strategy for functional selection.
 
@@
 A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}.
-To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
+To note, this prior differs from previous group SSL priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
 It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
@@
 This converges to the SB-GAM [@Bai2021].
-Conversely, it is also possible to relax the assumption such that each coefficient $\beta_{jk} \in \bs \beta^\tp_j$ has its own latent indicator $\gamma_{jk}$, but at the cost of complicating the bi-level function selection.
-This reduces the proposed prior to the classic spike-and-slab lasso prior.
+Conversely, it is also possible to relax the assumption such that each coefficient $\beta_{jk} \in \bs \beta^\tp_j$ has its own latent indicator $\gamma_{jk}$, but at the cost of complicating the bi-level functional selection.
+This reduces the proposed prior to the classic SSL prior.
 
@@
 The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
-Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables.
+Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions.
 
-The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
+The rest of the hierarchical prior follows the traditional SSL prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
 The two indicators of the $j$th predictor, $\gamma_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
@@
 This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
-The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
+The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the functional selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
 Meanwhile, we specify that $\gamma_{j}$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
@@
 $a = 1, b = 1$.
-When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}.
+When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases.
+Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab spline prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
 \begin{center}
@@
 \end{align*}
-Similar to the BHAM-SSL prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
+Similar to the spike-and-slab spline prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
 One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
@@
 
-As alternatives to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of uncertainty measures.
-The earlier work for fitting spike-and-slab models using optimization algorithms includes EMVS [@Rockova2014a].
+As an alternative to sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of posterior inference.
+The earlier work for fitting SS models using optimization algorithms includes EMVS [@Rockova2014a].
 Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
@@
 To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
-The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
+The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a block CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
 
@@
 * E-step: to calculate the expectation of the posterior density function with respect to some "missing" data
-* M-step: to maximize the expectation derived in the E-step to calculate the parameter of interest
+* M-step: to maximize the expectation derived in the E-step and update parameters of interest
 
@@
 We use the EM algorithm to find the MAP estimate of $\Theta$.
-Since the latent binary indicators $\bs \gamma$ are not of our primary interest, we treat them as the "missing data" in the EM algorithm.
-Naturally, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
+This is, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
 $$
@@
 $$
-and $E(Q_1)$ can be expressed as a likelihood function with $l_1$ penalty
+and $E(Q_1)$ can be expressed as a log-likelihood function with $l_1$ penalty
 \begin{equation}\label{eq:Q1_CD}
@@
 As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}).
-On the other hand, $\tau^{2}$ follows a gamma distribution is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
+On the other hand, $\tau^{2}$, following gamma distributions, is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
 $E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
@@
 \end{align*}
-To note, the variance-covariance matrixof the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
+To note, the variance-covariance matrix of the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
 $$
@@
 
-Totally, the proposed EM-iterative weighted least square algorithm is summarized as follows:
+Totally, the proposed EM-IWLS algorithm is summarized as follows:
 
@@
 
-    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E(\tau^{-2}_{j})$, $E({\tau^\tp}^{-2}_{jk})$ with estimates of $\Theta^{(t-1)}$ from the previous iteration
+    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E(\tau^{-2}_{j})$, $E({\tau^\tp}^{-2}_{jk})$ with the estimates $\Theta^{(t-1)}$ from the previous iteration
 
@@
 Increasing the spike scale $s_0$ tends to include more non-zero coefficients in the model.
-A measure of preference calculated with cross-validations, e.g.
+A measure of preference calculated with cross-validations (CV), e.g.
 deviance, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.

### 2021-10-18 Manuscript/00-main.Rmd
Revise the rest of paper

@@
   - \usepackage{pgfplots}
+  - \usepackage{pdflscape}
 ---
@@
 
+\newpage
+
 \bibliography{bibfile.bib}
@@
 
-\caption{Directed acyclic graph of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\caption{Directed acyclic graph of the proposed Bayesian hierarchical additive model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
 \label{fig:SSprior}
@@
 
-\newpage
-
 ```{r, results = "asis"}
@@
 \newpage
-```{r}
+<!-- \begin{landscape} -->
+```{r results = "asis"}
 tab_time %>% 
@@
          `Adaptive COSSO` = acosso) %>% 
+   xtable(
+    caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.",
+    label = "tab:time_sim",
+    booktabs = TRUE
+  ) %>% 
+  print(comment = FALSE, include.rownames = FALSE,
+       floating.environment = "sidewaystable",# caption.placement = "top"
+       table.placement = "!h"
+        )
 
-  flextable::flextable() %>% 
-    merge_v(j = 1) %>% 
-  flextable::set_caption(caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
+  # flextable::flextable() %>% 
+    # merge_v(j = 1) %>% 
+  # width(j=1, width=1) %>% 
+  # autofit() %>%
+  # flextable::set_caption(caption = "The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
 The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
@@
 ```
-
+<!-- \end{landscape} -->
 \newpage
@@
   # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
+  # flextable::set_caption(caption = "Measures of optimal BHAM and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation\\label{tab:WLM_res}")
 ```

### 2021-10-18 Manuscript/01-Intro.Rmd
Revise the rest of paper

--- 2021-10-15 Manuscript/01-Intro.Rmd
@@
 The prior setup encourages a flexible solution, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-In addition, two scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented in an publicly available R pacakge \textt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
+In addition, two scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented in an publicly available R package \texttt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
 

### 2021-10-18 Manuscript/03-Simulation.Rmd
Revise the rest of paper

@@
 SB-GAM is the first spike-and-slab lasso GAM.
-We implemented COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \texttt{sparseGAM 1.0}.
-COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence used the default choices.
-Both mgcv and proposed models allow customized smoothing functions and we chose the cubic regression spline.
-We controlled the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
-5-fold cross-validation were used for COSSO models, SB-GAM and the proposed models for tuning parameter selection based on the default selection criteria.
-20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) were examined for SB-GAM and the proposed models that allow user-specification of tuning candidates.
+We implement COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \texttt{sparseGAM 1.0}.
+COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices.
+Both mgcv and proposed models allow customized smoothing functions and we choose the cubic regression spline.
+We controll the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
+We use 5-fold CV with the default selection criteria to select the final model for COSSO models, SB-GAM and the proposed models.
+20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed models which allow user-specification of tuning candidates.
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
@@
 Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al.
-[@Ravikumar2009], stochastic search term selection for GAM [@Scheipl2012].
+[@Ravikumar2009] and stochastic search term selection for GAM [@Scheipl2012].
 However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increased number of predictors.
@@
 We follow the data generating process described in Bai [@Bai2021].
-We first generated $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
-We then simulated the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
+We first generate $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
+We then simulate the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
 The mean of each outcome were simulated via the following function
@@
 $$
-for gaussian and binomial outcomes.
+for Gaussian and binomial outcomes.
 <!-- and shrinked effects -->
@@
 <!-- for  poisson outcomes.  -->
-Gaussian outcomes required specification of dispersion, where we set the dispersion parameter to be 1.
+Gaussian outcomes requires specification of dispersion, where we set the dispersion parameter to be 1.
 In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest covariates are inactive, i.e.
 $f_j(x_j) = 0$ for $j = 4, \dots, p$.
-Another set of independent sample of size $n_{test}=1000$,($x_{new} ,y_{new}$), are created following the same data generating process, serving as the testing data.
-We generated 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
+Another set of independent sample of size $n_{test}=1000$, are created following the same data generating process, serving as the testing data.
+We generate 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
 <!-- The dimensional of the smoothing functions are controlled as k=10. Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.  -->
@@
 
-To evaluate the  predictive performance of the models, the statistics calculated based on the testing dataset, $R^2$ for Gaussian model, area under the curve (AUC) for binomial model,  are calculated <!--and deviance are calculated for Poisson model,--> averaging across 50 simulations.
+To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing dataset, are averaged <!--and deviance are calculated for Poisson model,--> across 50 simulations.
+Computation time for model selection, final model fitting and prediction are recorded for all simulations.
+
 The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.

### 2021-10-18 Manuscript/04-Real_Data.Rmd
Revise the rest of paper

--- 2021-10-13 Manuscript/04-Real_Data.Rmd
@@
 
-In this section, we apply the proposed models BHAM-SSL, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+In this section, we apply the proposed models BHAM, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
@@
 ## Emory Cardiovascular Biobank
-We use the proposed models BHAM-SSL to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
+\label{sec:ECB}
+We use the proposed models BHAM to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
 The dataset is publicly available via \textit{Dryad} [@Mehta2020_data].
@@
 As there is a large number of non-overlapping features among the two cohorts, we use the cohort with larger sample size (N=454).
-There were initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
-Hence, we performed a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and chose the the top 200 features with smallest p-values.
-We used 5-knot spline additive models for binary outcome using two different models, the proposed BHAM-SSL and the SB-GAM.
-10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ were calculated and tabulated in Table \ref{tab:ECB_res}.
+There are initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
+Hence, we perform a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and choose the the top 200 features with smallest p-values.
+We use 5-knot spline additive models for binary outcome using two different models, the proposed BHAM and the SB-GAM.
+10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
+BHAM-CD obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (Table \ref{tab:ECB_res}).
+
 \begin{center}
@@
 ## Weight Loss Maintenance Cohort 
-We use the proposed models BHAM-SSL to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
-In this analysis, we primarily focused on the analysis of one of the three studies included, weight loss maintenance (WLM) cohort [@Svetkey2008], due to the drastically different intervention effects.
+We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
+In this analysis, we primarily focuse on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects.
 In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
-Quality control and natural log transformation were performed during metabolites data preparation.
-The outcome of interest was standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
-After removing missing datapoints and addressing outliers in the data, there were $p$=`r nrow(WLM_train_dat)` features remaining in the analysis.
-5-Knot spline additive models for the gaussian outcome are constructed using two different models, the proposed BHAM-SSL and the SB-GAM.
-10-Fold cross-validation(CV) are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation, where deviance, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ were calculated and tabulated in Table \ref{tab:WLM_res}.
+Quality control and natural log transformation are performed during metabolites data preparation.
+The outcome of interest are standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
+After removing missing datapoints and addressing outliers in the data, there are $p$=`r nrow(WLM_train_dat)` features remaining in the analysis.
+5-Knot spline additive models for the Gaussian outcome are constructed using two different models, the proposed BHAM and the SB-GAM.
+10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
+Out-of-bag samples are used for prediction performance evaluation, where deviance, $R^2$, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ are calculated.
+BHAM-CD obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (Table \ref{tab:WLM_res}).
 \begin{center}

### 2021-10-18 Manuscript/05-Discussion.Rmd
Remove github links
Revise the rest of paper

@@
 Meanwhile, we introduced two optimization based algorithms for model fitting.
-The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner, but also provide naive inferential supports for many commonly used hierarchical priors, e.g.
-t-distribution, double exponential distribution, mixture t distribution, and mixture double exponential distribution, in the analyses that the sparse assumption is weak or not necessary.
+The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner.
 Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.
@@
 Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.
+In addition, in \texttt{BHAM}, we streamline the model fitting algorithms to support other popular Bayesian hierarchical prior for the smoothing functions, such as Student's T distribution, double exponential distribution, mixture normal and T distribution's.
+These priors could be helpful when the sparse assumption is weak or not necessary.,
 
@@
 $\gamma_j^\tp|\gamma_{j}^{0}, \theta_j$.
-Secondly, while the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
-Our current effort resides on expanding GAM to Cox proportional hazard models.
+Secondly, the computational time for fitting a BHAM model with EM-IWLS algorithms can be improved.
+Current implementation of the EM-IWLS algorithm jointly updates all coefficients in each iteration, which requires a lot computation resources.
+This jointly updating procedure can be enhanced by adapting a backfitting step [@Hastie1987] where each smoothing function are updated individually.
+Thirdly, when using the proposed model in real data analysis where a screening procedure is implemented before joint predictive modeling.
+we recommend to include the screening procedure in the cross-validation during model selection.
+[@Friedman2017] For the sake of model comparison, we fit the models using the same pool of predictors in Section \ref{sec:ECB}.
+
+Our future efforts direct to modeling survival outcomes and integrative analysis.
+While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
 An naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].
 However, it would be more efficient to directly fit Cox models via penalized pseudo likelihood function [@Simon2011].
-Last but not least, we plan to expand the proposed model to integrate external biology information.
-The main motivation for the integrative model is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
+Meanwhile, with growing understanding of biological structure within -omics field, it is appealing to integrate external biology information in the modeling process.
+The main motivation for integrative models is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
 Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.

### 2021-10-18 Manuscript/SS_GAM.tex
Revise the rest of paper

--- 2021-09-30 Manuscript/SS_GAM.tex
@@
 \usepackage{pgfplots}
+\usepackage{pdflscape}
 \usepackage{multirow}
@@
 \abstract{There are proposals that extend the classical generalized
-additive models (GAMs) to accommodate the high-dimensional (\(p>>n\))
-problem using group penalties and priors. These models rarely smooth the
-nonlinear effect properly as the sparse regularization leads to excess
-shrinkage on coefficients and inaccurate prediction. Moreover, most of
-these GAMs consider an ``all-in-all-out'' approach for function
-selection, rendering them difficult to answer if nonlinear effects are
-necessary. While some Bayesian models can address these shortcomings,
-using Markov chain Monte Carlo (MCMC) algorithms for model fitting
-creates a new challenge, scalability. Hence, we propose Bayesian
-hierarchical generalized additive models as a solution: we propose a
-novel spike-and-slab spline prior for smoothing functions, where a
-reparameterization step facilitates smoothing penalty and separate
-selection of linear and nonlinear spaces. Two scalable and deterministic
+additive models (GAMs) to accommodate high-dimensional data (\(p>>n\))
+using group sparse regularization. However, the sparse reuglarization
+may induce excess shrinkage when estimating smoothing functions,
+damaging predictive performance. Moreover, most of these GAMs consider
+an ``all-in-all-out'' approach for functional selection, rendering them
+difficult to answer if nonlinear effects are necessary. While some
+Bayesian models can address these shortcomings, using Markov chain Monte
+Carlo algorithms for model fitting creates a new challenge, scalability.
+Hence, we propose Bayesian hierarchical generalized additive models as a
+solution: we consider the smoothing penalty for proper shrinkage of
+curve interpolation and separation of smoothing function linear and
+nonlinear spaces. A novel spike-and-slab spline prior is proposed to
+select components of smoothing functions. Two scalable and deterministic
 algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
-Squares, are developed to fit the proposed model. Simulation studies and
-metabolomics data analyses demonstrate improved performance against
-state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM. The
-software implementation of the proposed models is freely available via
-an R package BHAM.}
+Squares, are developed for different utilities. Simulation studies and
+metabolomics data analyses demonstrate improved predictive or
+computational performance against state-of-the-art models, mgcv, COSSO
+and sparse Bayesian GAM. The software implementation of the proposed
+models is freely available via an R package BHAM.}
 
@@
 respectively applied grouped SCAD penalty \citep{Fan2001} to additive
-models. Recent Bayesian hierarchical models are also used in the context
-of high-dimensional additive models. Various group spike-and-slab priors
-combining with computationally intensive Markov chain Monte Carlo (MCMC)
-algorithms \citep{Xu2015, Yang2020} are proposed, where the application
-on AMs are by-products. Bai et al. \citep{Bai2020} was the first to
-apply group spike-and-slab lasso prior to Gaussian AMs using a fast
-optimization algorithm, and further generalized the framework to
-GAMs\citep{Bai2021}. As these methods focus on addressing the sparsity,
-it is very likely to overly penalize the basis function coefficients and
-produce inaccurate predictions and curve interpolation, particularly
-when complex signals are assumed and large number of knots are used.
-\citep{Scheipl2013} In addition, these methods adapt an `all-in-all-out'
-strategy, i.e.~either including or excluding the variable completely,
-rendering no space for bi-level selection. Scheipl et al.
-\citep{Scheipl2012} proposed a spike-and-slab structure prior,
-normal-mixture-of-inverse gamma prior, that address the previous
-challenges. But the model fitting relies on computational intensive MCMC
-algorithmic and creates scalability concern. It would be of special
-interest to develop a fast, flexible and accurate generalized additive
-model framework.
+models. Recently Bayesian hierarchical models are also used in the
+context of high-dimensional additive models. Various group
+spike-and-slab priors combining with computationally intensive Markov
+chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
+proposed, where the application on AMs are by-products. Bai et al.
+\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
+to Gaussian AMs using a fast optimization algorithm, and further
+generalized the framework to GAMs\citep{Bai2021}. Focus on addressing
+the sparsity, these methods can overly penalize the basis function
+coefficients and produce inaccurate predictions and curve interpolation,
+particularly when complex signals are assumed and large number of knots
+are used. \citep{Scheipl2013} In addition, these methods adapt an
+`all-in-all-out' strategy, i.e.~either including or excluding the
+variable completely, rendering no space for bi-level selection. Scheipl
+et al. \citep{Scheipl2012} proposed a spike-and-slab structure prior
+that address the previous challenges. But the model fitting relies on
+computational intensive MCMC algorithms and creates scalability concern.
+It would be of special interest to develop a fast, flexible and accurate
+generalized additive model framework.
 
@@
 penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of smoothing functions from
-the nonlinear space. We then impose a new bi-part spike-and-slab spline
+allows the separation of the linear space of a smoothing function from
+the nonlinear space. We then impose a new two-part spike-and-slab spline
 prior on the smoothing functions for bi-level selection such that the
 linear and nonlinear spaces of smoothing functions can be selected
-separately. The prior setup encourages a flexible solution, rending one
-of three possibilities: no effect, only linear effect, or nonlinear
-effect of a predictor. In addition, two scalable optimization-based
-algorithms are developed for high and
-ultrahigh-dimensional\citep{Fan2009} settings respectively.
+separately. The prior setup encourages a flexible solution, rendering
+one of three possibilities for each predictor: no effect, only linear
+effect, or linear and nonlinear effects. In addition, two scalable
+optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm
+and EM-Iterative weighted least square (EM-IWLS) algorithm, are
+developed and implemented in an publicly available R package
+\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}, making
+translational science more accessible.
 
@@
 GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in many ways. First of all, the proposed spike-and-slab
+\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
 spline prior is a spike-and-slab lasso type prior using independent
-mixture double exponential distribution, compared to spike-and-slab GAM.
-Spike-and-slab lasso priors provide computational convenience during
-model fitting by using optimization algorithms instead of computational
-intensive sampling algorithms. They make fitting high- and
-ultrahigh-dimensional models more feasible without sacrificing
-performance in prediction and variable selection. Secondly, the
-optimization algorithms introduced in this article include EM-Coordinate
-Descent (EM-CD) algorithm and EM-Iterative weighted least square
-(EM-IWLS) algorithm for different utilities. The applications of the two
-aforementioned algorithms are fruitful in Bayesian hierarchical models
-for high-dimensional data. \citep{Yi2012, Rockova2014a, Rockova2018}
-Even though SB-GAM with a group spike-and-slab lasso prior also uses
-EM-CD for model fitting, the proposed EM-IWLS provides uncertainty
-measures that could be further used for confidence bound calculation and
-hypothesis testing. EM-IWLS complements the lack of inference potential
-from EM-CD in a median and high dimensional setting. Furthermore, the
-proposed model addresses the incapability of bi-level selection in
-SB-GAM. Lastly, the R implementation of the Bayesian hierarchical
-additive model is available publicly via BHAM at
-\url{https://github.com/boyiguo1/BHAM}, making transnational science
-more accessible.
+mixture double exponential distribution, compared to spike-and-slab GAM
+that uses normal-mixture-of-inverse gamma prior. Spike-and-slab lasso
+priors provide computational convenience during model fitting by using
+optimization algorithms instead of intensive sampling algorithms. They
+make fitting high-dimensional models more feasible without sacrificing
+performance in prediction and variable selection. Secondly, SB-GAM uses
+a group spike-and-slab lasso prior with an EM-CD algorithm to fit the
+model. While both methods use the combination of expectation
+maximization algorithm and coordient descent algorithm, there are subtle
+difference in the implementation due to the difference in prior
+specification. The proposed model sets up independent priors among basis
+function coefficients after the re-parameterization step, which provides
+some advantage in computation. In addition, we offer an alternative
+algorithm, EM-IWLS, that can provide variance-covariance matrix of the
+coefficients. Last but not least, the proposed model addresses the
+incapability of bi-level selection in SB-GAM.
 
@@
 Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model and its matrix form,
-\begin{equation}\label{eq:gam}
-E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
-\end{equation} and the data distribution is \begin{equation}
+have a generalized additive model with link function \(g(\cdot)\) and
+linear predictor \begin{equation}\label{eq:gam}
+\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
+\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
+\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
+distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
+and dispersion parameter \(\phi\), and the data distribution is
+\begin{equation}
 f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The basis function matrix, i.e.~the design
-matrix derived from the smoothing function \(B_j(x_j)\), is denoted
+\end{equation} The basis function matrix, i.e.~the design matrix derived
+from the smoothing function \(B_j(x_j)\), is denoted
 \(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
-design matrix depends on the choice of the smoothing function. For
-example, a cubic spline function with \(k\) knots produces a \(k+3\)
-elements column vector. Its corresponding smoothing penalty matrix is
-denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\) denotes the
-basis function coefficients for the \(j\)th variable such that
-\(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
+design matrix depends on the choice of the smoothing function, and is
+denoted as \(K_j\) for \(x_j\). Its corresponding smoothing penalty
+matrix is denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\)
+denotes the basis function coefficients for the \(j\)th variable such
+that \(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
 abuse of notation, we denote vectors and matrices in bold fonts
 \(\boldsymbol{\beta}, \boldsymbol{X}\) with conformable dimensions,
-where scalar and random variables are denoted in with unbold fonts
+where scalar and random variables are denoted in unbold fonts
 \(\beta, X\). The matrix transposing operation is denoted with a
@@
 
-To encourage proper smoothing of the functions, we adopt the idea of
-smoothing penalties from smoothing spline models. The basic idea is to
-set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}),
-in the prior density function. However, the direct integration of
-smoothing penalty with sparsity penalty is not obvious. Marra and Wood
-\citep{Marra2011} proposed a re-parameterization procedure to
-accommodate the smoothing penalty implicitly. Given the smoothing
-penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
+To encourage proper smoothing of additive functions, we adopt the idea
+of smoothing penalties from smoothing spline models. The basic idea is
+to set up a smoothing penalty, described in Equation
+(\ref{eq:smoothpen}), in the prior density function. However, the direct
+integration of smoothing penalty with sparsity penalty is not obvious.
+Marra and Wood \citep{Marra2011} proposed a re-parameterization
+procedure to accommodate the smoothing penalty implicitly. Given the
+smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
 semi-definite for univariate smoothing functions, we apply
@@
 B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \beta_j X^0_j + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
-\] and the generalized additive model in Equation (\ref{eq:gam}) be
+\] and the generalized additive model in Equation (\ref{eq:gam}) now is
 \begin{equation}\label{eq:gam-repa}
@@
 The re-parameterization step sets up the foundation of the proposed
-model and provides three-fold benefits. First of all, the
-re-parameterization integrates the smoothing penalty into the design
-matrix, and encourages models to properly smooth the nonlinear function
-in addition to the sparse penalty for function selection. Secondly, the
+model and provides three benefits. First of all, the re-parameterization
+integrates the smoothing penalty into the design matrix, and encourages
+models to properly smooth the nonlinear function in addition to the
+sparse penalty for functional selection. Secondly, the
 eigendecomposition of the smoothing penalty allows the isolation of the
-linear from the nonlinear space, improving the feasibility of bi-level
-function selection and inference. The eigendecomposition facilitates the
+linear space from the nonlinear space, improving the feasibility of
+bi-level functional selection. The eigendecomposition facilitates the
 construction of orthonormal design matrix, which makes imposing
@@
 
-\hypertarget{spike-and-slab-smooth-priors}{%
-\subsection{Spike-and-Slab Smooth
-Priors}\label{spike-and-slab-smooth-priors}}
-
-The family of spike-and-slab (SS) priors regression models dominates
-Bayesian high-dimensional analysis for its utility in outcome prediction
-and variable selection. We defer to Bai et al. \citep{Bai2021Review} for
-an in-depth introduction to spike-and-slab priors. To summarize,
-spike-and-slab priors are a family of mixture distributions that
-comprises a skinny spike density \(f_{\text{spike}}(\cdot)\) for weak
-signals and a flat slab density \(f_{\text{slab}}(\cdot)\) for strong
-signals, mathematically \[
+\hypertarget{spike-and-slab-spline-priors}{%
+\subsection{Spike-and-Slab Spline
+Priors}\label{spike-and-slab-spline-priors}}
+
+The family of spike-and-slab (SS) regression models is one of most
+commonly used models in high-dimensional data analysis for its utility
+in outcome prediction and variable selection. We defer to Bai et al.
+\citep{Bai2021Review} for an in-depth introduction to spike-and-slab
+priors. To summarize, spike-and-slab priors are a family of mixture
+distributions that comprises a skinny spike density
+\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
+\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
  \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
@@
 studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS \citep{Rockova2014a} and
-spike-and-slab Lasso \citep{Rockova2018b, Rockova2018}. Particularly,
-the development of spike-and-slab Lasso substantially improves the
-scalability of SS models, setting up the theoretical foundation for
-generalized models in -omics data analysis
-\citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The spike-and-slab
-Lasso (SSL) prior is composed of two double exponential distributions
-with mean 0 and different dispersion parameters, \(0 < s_0 < s_1\),
-mathematically, \begin{equation} 
+algorithms bear, which include EMVS based on spike-and-slab normal
+prior\citep{Rockova2014a} and spike-and-slab Lasso (SSL)
+\citep{Rockova2018b, Rockova2018}. Particularly, the development of SSL
+model substantially improves the scalability of SS models, setting up
+the theoretical foundation for generalized models in -omics data
+analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The SSL prior
+is composed of two double exponential distributions with mean 0 and
+different dispersion parameters, \(0 < s_0 < s_1\), mathematically,
+\begin{equation} 
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
@@
 GAM framework, where the densities of the spike and slab components take
-the group lasso density \citep{Xu2015}. Bai \citep{Bai2021} takes an
-``all-in-all-out'' strategy for function selection.
-
-\hypertarget{spike-and-slab-lasso-prior}{%
-\subsubsection{Spike-and-Slab Lasso
-Prior}\label{spike-and-slab-lasso-prior}}
+the group lasso density \citep{Xu2015} and limits to an
+``all-in-all-out'' strategy for functional selection.
+
+\hypertarget{two-part-spike-and-slab-lasso-prior}{%
+\subsubsection{Two-part Spike-and-Slab Lasso
+Prior}\label{two-part-spike-and-slab-lasso-prior}}
 
@@
 parameters comes in Section \ref{sec:tune}. To note, this prior differs
-from previous group spike-and-slab lasso priors
-\citep{Tang2018, Tang2019}, as the \(\beta_j\) and
-\(\boldsymbol{\beta}^*_j\) have different indicator variables
-\(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to add a more
-restrictive assumption on the priors, assuming that one indicator
-variable decides the inclusion of both the linear effect and nonlinear
-effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the SB-GAM
-\citep{Bai2021}. Conversely, it is also possible to relax the assumption
-such that each coefficient \(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has
-its own latent indicator \(\gamma_{jk}\), but at the cost of
-complicating the bi-level function selection. This reduces the proposed
-prior to the classic spike-and-slab lasso prior.
+from previous group SSL priors \citep{Tang2018, Tang2019}, as the
+\(\beta_j\) and \(\boldsymbol{\beta}^*_j\) have different indicator
+variables \(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to
+add a more restrictive assumption on the priors, assuming that one
+indicator variable decides the inclusion of both the linear effect and
+nonlinear effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the
+SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
+assumption such that each coefficient
+\(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has its own latent indicator
+\(\gamma_{jk}\), but at the cost of complicating the bi-level functional
+selection. This reduces the proposed prior to the classic SSL prior.
 
@@
 provides a panacea to allow unified scale parameters for all bases of
-all smoothing functions of variables.
-
-The rest of the hierarchical prior follows the traditional
-spike-and-slab lasso prior: we set up hyper-priors on
-\(\gamma_j, \gamma^*_j\) to allow local adaption of the shrinkage using
-a Bernoulli distribution, written as binomial distribution of one trial.
-The two indicators of the \(j\)th predictor, \(\gamma_j\) and
-\(\gamma^*_j\), shares the same probability parameter \(\theta_j\), \[
+all smoothing functions.
+
+The rest of the hierarchical prior follows the traditional SSL prior: we
+set up hyper-priors on \(\gamma_j, \gamma^*_j\) to allow local adaption
+of the shrinkage using a Bernoulli distribution, written as binomial
+distribution of one trial. The two indicators of the \(j\)th predictor,
+\(\gamma_j\) and \(\gamma^*_j\), shares the same probability parameter
+\(\theta_j\), \[
 \begin{aligned}
@@
 smoothing function. The hyper prior of \(\gamma_{j}\) decides the
-sparsity of the model at the function selection level, while that of
+sparsity of the model at the functional selection level, while that of
 \(\gamma_{j}^*\) decides the smoothness of the spline function at basis
@@
 any of the bases, the parameter \(\theta_j\) will be estimated large,
-which in turn encourages the model to include the rest of bases,
-achieving the local adaption among spline bases. Hereafter, we refer
-Bayesian hierarchical generalized additive models with the
-spike-and-slab lasso prior as the BHAM-SSL, and visually presented in
-Figure \ref{fig:SSprior}.
-
-\begin{figure}
-\centering
-\begin{tikzpicture} [
-staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
-PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
-background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
-]
-
-\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
-};
-\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
-};
-
-\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
-\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
-\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
-
-\draw[->] (Theta) -- (linearGamma);
-\draw[->] (Theta) -- (penGamma);
-\draw[->] (linearGamma) -- (linearBeta) ;
-\draw[->] (penGamma) -- (penBeta);
-\draw[->] (ab) -- (Theta);
-\draw[->] (s) -- (linearBeta) ;
-\draw[->] (s) -- (penBeta);
-\draw[->] (linearBeta) -- (Beta);
-\draw[->] (penBeta) -- (Beta);
-\draw[->] (Beta) --  (Y);
-
-\begin{pgfonlayer}{background}
-  \node [background,
-   fit=(linearGamma) (linearBeta),
-   label=above:Linear Space:] {};
-  \node [background,
-    fit=(penGamma) (penBeta),
-    label=above:Nonlinear Space:] {};
-\end{pgfonlayer}
-
-\end{tikzpicture}
-
-\caption{Directed acyclic grpah of BHAM-SSL model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
-\end{figure}
+which in turn encourages the model to include the rest of bases.
+Hereafter, we refer Bayesian hierarchical generalized additive models
+with the spike-and-slab spline prior as the BHAM, and visually presented
+in Figure \ref{fig:SSprior}.
+
+\begin{center}
+Figure \ref{fig:SSprior} here
+\end{center}
 
@@
   \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
-\end{align*} Similar to the BHAM-SSL prior in Equation
+\end{align*} Similar to the spike-and-slab spline prior in Equation
 (\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
@@
 
-As alternatives to the sampling algorithms for fitting Bayesian models,
+As an alternative to sampling algorithms for fitting Bayesian models,
 optimization algorithms focus on the maximum a posteriori (MAP)
 estimates and speed up the model fitting process at the cost of
-uncertainty measures. The earlier work for fitting spike-and-slab models
-using optimization algorithms includes EMVS \citep{Rockova2014a}.
-Rockova and George\citep{Rockova2014a} proposed an
-expectation-maximization (EM) based algorithm to fit models that use
-continuous mixture normal priors. In the E step, the latent binary
-indicators \(\gamma\)s are treated as the missing data, and the
-posterior means are calculated conditioning on the current value of
-other parameters; in the M step, a ridge estimator was used to update
-the coefficients, followed by updating \(\phi, \theta\). The same
-authors \citep{Rockova2018b, Rockova2018} further combined the EM
-algorithm with coordinate descent algorithm to fit SSL models. Yi and
-his group independently developed the EM-Iterative Weighted Least Square
-and EM-cyclic coordinate descent algorithms to fit models with broader
-class of priors, SSL included.\citep{Yi2019} These algorithms were
-implemented for generalized linear models\citep{Tang2017a}, Cox
-proportional hazards models \citep{Tang2017} and their grouped
+posterior inference. The earlier work for fitting SS models using
+optimization algorithms includes EMVS \citep{Rockova2014a}. Rockova and
+George\citep{Rockova2014a} proposed an expectation-maximization (EM)
+based algorithm to fit models that use continuous mixture normal priors.
+In the E step, the latent binary indicators \(\gamma\)s are treated as
+the missing data, and the posterior means are calculated conditioning on
+the current value of other parameters; in the M step, a ridge estimator
+was used to update the coefficients, followed by updating
+\(\phi, \theta\). The same authors \citep{Rockova2018b, Rockova2018}
+further combined the EM algorithm with coordinate descent algorithm to
+fit SSL models. Yi and his group independently developed the
+EM-Iterative Weighted Least Square and EM-cyclic coordinate descent
+algorithms to fit models with broader class of priors, SSL
+included.\citep{Yi2019} These algorithms were implemented for
+generalized linear models\citep{Tang2017a}, Cox proportional hazards
+models \citep{Tang2017} and their grouped
 counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
@@
 provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM-SSL with an expedited performance, recommending to use in
-high and ultra-high dimensional setting. A specific concern of EM-CD
+fitting BHAM with an expedited performance, recommending to use in high
+and ultra-high dimensional setting. A specific concern of EM-CD
 algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can extract the variance-covariance matrix and be used to
-construct Wald type inference. Moreover, the EM-IWLS is a more general
-model fitting algorithm that can be used for fitting not only SSL and
-continuous SS priors but also Student's t-priors and double exponential
-priors. To note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD
-algorithm. The main difference between proposed EM-CD algorithm and that
-in SB-GAM is that SB-GAM uses a blocked CD algorithm for their group
-prior, while the proposed prior is pairwise independent requiring no
-special treatment in the CD algorithm.
+the EM-IWLS can estimate the variance-covariance matrix of the
+coefficients. Moreover, the EM-IWLS is a more general model fitting
+algorithm that can be used for fitting not only SSL and continuous SS
+priors but also Student's t-priors and double exponential priors. To
+note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
+main difference between proposed EM-CD algorithm and that in SB-GAM is
+that SB-GAM uses a block CD algorithm for their group prior, while the
+proposed prior is pairwise independent requiring no special treatment in
+the CD algorithm.
 
@@
 \item
-  M-step: to maximize the expectation derived in the E-step to calculate
-  the parameter of interest
+  M-step: to maximize the expectation derived in the E-step and update
+  parameters of interest
 \end{itemize}
@@
 where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density.
-
-We use the EM algorithm to find the MAP estimate of \(\Theta\). Since
-the latent binary indicators \(\boldsymbol{\gamma}\) are not of our
-primary interest, we treat them as the ``missing data'' in the EM
-algorithm. Naturally, in the E-step, we calculate the expectation of
-posterior density function of
+distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
+non-informative prior for the intercept \(\beta_0\) and the dispersion
+parameter \(\phi\); for example, \(f(\beta_0|\tau_0^2)=N(0,\tau_0^2)\)
+with \(\tau^2_0\) set to a large value and \(f(\log \phi) \propto 1\).
+
+We use the EM algorithm to find the MAP estimate of \(\Theta\). This is,
+in the E-step, we calculate the expectation of posterior density
+function of
 \(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
@@
 
-To recall, the density function of spike-and-slab mixture double
-exponential prior can be written as \[
+The density function of spike-and-slab mixture double exponential prior
+can be written as \[
 f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
-\] and \(E(Q_1)\) can be expressed as a likelihood function with \(l_1\)
-penalty \begin{equation}\label{eq:Q1_CD}
+\] and \(E(Q_1)\) can be expressed as a log-likelihood function with
+\(l_1\) penalty \begin{equation}\label{eq:Q1_CD}
 E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
@@
 \(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
-\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}) via
-Bayes' Theorem. On the other hand, \(\tau^{2}\) following a gamma
-distribution is a conjugate prior for the normal variance, and the
-conditional posterior density of \(\tau^{-2}\) is an inverse Gaussian
-distribution. \(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are
-calculated using the closed form equation \begin{align*}
+\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}). On
+the other hand, \(\tau^{2}\), following gamma distributions, is a
+conjugate prior for the normal variance, and the conditional posterior
+density of \(\tau^{-2}\) is an inverse Gaussian distribution.
+\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
+using the closed form equation \begin{align*}
  E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
@@
 & \boldsymbol{\beta}^{(t)} = (\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{z}_* && \phi^{(t)} = \frac{1}{n}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)})^T\boldsymbol{\Sigma}^{-1}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)}).&
-\end{align*} To note, the variance-covariance matrixof the coefficient
+\end{align*} To note, the variance-covariance matrix of the coefficient
 estimates variance-covariance matrix can be derived in the EM-IWLS
@@
   \text{Var}(\boldsymbol{\beta}^{(t)}) = (\boldsymbol{X}_*^T\boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\phi^{(t)}.
-\] The rest of the algorithm remains the same as described in Section
-\ref{sec:EMCD}.
+\]
+
+Totally, the proposed EM-IWLS algorithm is summarized as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
+\item
+  Iterate over the E-step and M-step until convergence
+
+  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
+  \(E(\tau^{-2}_{j})\), \(E({\tau^*}^{-2}_{jk})\) with the estimates
+  \(\Theta^{(t-1)}\) from the previous iteration
+
+  M-step:
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii})}
+  \tightlist
+  \item
+    Based on the current value of \(\beta\), calculate the pseudo-data
+    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
+  \item
+    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
+    least squared
+  \item
+    If \(\phi\) is present, update \(\phi\)
+  \end{enumerate}
+\end{enumerate}
+
+Similar to EM-CD, we assess convergence by the criterion,
+\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where \(\epsilon\) is
+a small value (say \(10^{-5}\)).
 
@@
 spike scale \(s_0\) tends to include more non-zero coefficients in the
-model. A measure of preference calculated with cross-validations,
+model. A measure of preference calculated with cross-validations (CV),
 e.g.~deviance, area under the curve (AUC), mean squared error, can be
@@
 effect interpolation and prediction. SB-GAM is the first spike-and-slab
-lasso GAM. We implemented COSSO and adaptive COSSO with R package
+lasso GAM. We implement COSSO and adaptive COSSO with R package
 \texttt{cosso 2.1-1}, generalized additive models with automatic
@@
 \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence used the default
+flexibility to define smoothing functions, and hence use the default
 choices. Both mgcv and proposed models allow customized smoothing
-functions and we chose the cubic regression spline. We controlled the
+functions and we choose the cubic regression spline. We controll the
 dimensionality of each smoothing function, 10 bases, for all different
-choices of smoothing functions. 5-fold cross-validation were used for
-COSSO models, SB-GAM and the proposed models for tuning parameter
-selection based on the default selection criteria. 20 default candidates
-of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) were
-examined for SB-GAM and the proposed models that allow
-user-specification of tuning candidates. All computation was conducted
-on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz
-eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
-3.6.2 \citep{R}.
+choices of smoothing functions. We use 5-fold CV with the default
+selection criteria to select the final model for COSSO models, SB-GAM
+and the proposed models. 20 default candidates of tuning parameters
+(\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for SB-GAM and
+the proposed models which allow user-specification of tuning candidates.
+All computation was conducted on a high-performance 64-bit Linux
+platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
+processors and 24G of RAM per core and R 3.6.2 \citep{R}.
 
@@
 methods of sparse additive models by Ravikumar et al.
-\citep{Ravikumar2009}, stochastic search term selection for GAM
+\citep{Ravikumar2009} and stochastic search term selection for GAM
 \citep{Scheipl2012}. However, we exclude these methods from current
@@
 We follow the data generating process described in Bai \citep{Bai2021}.
-We first generated \(n=500\) training data points with
-\(p=4, 10, 50, 200\) predictors respectively, where the predictors \(X\)
-are simulated from a multivariate normal distribution
-\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulated the outcome
-\(y\) from two distributions, Gaussian and binomial with the identity
-link and logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The
-mean of each outcome were simulated via the following function \[
+We first generate \(n=500\) training data points with
+\(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
+\(X\) are simulated from a multivariate normal distribution
+\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(y\)
+from two distributions, Gaussian and binomial with the identity link and
+logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The mean of each
+outcome were simulated via the following function \[
 \mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for gaussian and binomial outcomes. Gaussian outcomes required
+\] for Gaussian and binomial outcomes. Gaussian outcomes requires
 specification of dispersion, where we set the dispersion parameter to be
@@
 i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
-independent sample of size \(n_{test}=1000\),(\(x_{new} ,y_{new}\)), are
-created following the same data generating process, serving as the
-testing data. We generated 50 independent pairs of training and testing
-datasets to evaluate the prediction performance of the chosen models,
-where training datasets are used to fit the models and testing datasets
-used to calculate assessment measures.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{ccccccc}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.73 (0.09) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.16) & 0.76 (0.04) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.81 (0.05) & 0.43 (0.19) & 0.54 (0.19) & 0.86 (0.02) & 0.81 (0.07) \\ 
-  200 & 0.73 (0.10) & 0.60 (0.08) & 0.39 (0.17) & 0.45 (0.19) & - & 0.81 (0.05) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations} 
-\label{tab:gaus}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.89 (0.04) & 0.90 (0.02) & 0.90 (0.02) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.87 (0.03) & 0.87 (0.03) & 0.85 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.87 (0.02) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.81 (0.06) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measure for binomial outcomes over 50 iterations} 
-\label{tab:bin_auc}
-\end{table}
-
-To evaluate the predictive performance of the models, the statistics
-caluclated based on the testing dataset, \(R^2\) for Gaussian model,
-area under the curve (AUC) for binomial model, are calculated averaging
-across 50 simulations. The results are tabulated in Table
-\ref{tab:gaus}, \ref{tab:bin_auc}.
+independent sample of size \(n_{test}=1000\), are created following the
+same data generating process, serving as the testing data. We generate
+50 independent pairs of training and testing datasets to evaluate the
+prediction performance of the chosen models, where training datasets are
+used to fit the models and testing datasets used to calculate assessment
+measures.
+
+To evaluate the predictive performance of the models, the statistics,
+\(R^2\) for Gaussian model and AUC for binomial model calculated based
+on the testing dataset, are averaged across 50 simulations. Computation
+time for model selection, final model fitting and prediction are
+recorded for all simulations.
+
+The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
+
+\begin{center}
+Table \ref{tab:gaus} here\\
+Table \ref{tab:bin_auc} here\\
+\end{center}
 
@@
 
-In this section, we apply the proposed models BHAM-SSL, fitted with the
+In this section, we apply the proposed models BHAM, fitted with the
 EM-CD algorithm, to analyze two real-world metabolomics datasets where
@@
 
-We use the proposed models BHAM-SSL to analyze a metabolic dataset from
-a recently published research \citep{Mehta2020} studying plasma
-metabolomic profile on the three-year all-cause mortality among patients
-undergoing cardiac catheterization. The dataset is publicly available
-via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
-subjects from two cohorts. As there is a large number of non-overlapping
-features among the two cohorts, we use the cohort with larger sample
-size (N=454). There were initially 6796 features in the dataset, which
-is too large to be practically meaningful to analyze. Hence, we
-performed a univariate screening procedure on the features, via GAM
-implemented in \texttt{mgcv}, and chose the the top 200 features with
-smallest p-values. 5-Knot spline additive models for binary outcome are
-constructed using two different models, the proposed BHAM-SSL and the
-SB-GAM. 10-Fold cross-validation(CV) are used to choose the optimal
-tuning parameters of each framework with respect to the default
-selection criterion implemented in the software. Out-of-bag samples were
-used for prediction performance evaluation, where deviance, area under
-the ROC curve (AUC), mean squared error (MSE) defined as
+\label{sec:ECB} We use the proposed models BHAM to analyze a metabolic
+dataset from a recently published research \citep{Mehta2020} studying
+plasma metabolomic profile on the three-year all-cause mortality among
+patients undergoing cardiac catheterization. The dataset is publicly
+available via \textit{Dryad} \citep{Mehta2020_data}. It contains in
+total of 776 subjects from two cohorts. As there is a large number of
+non-overlapping features among the two cohorts, we use the cohort with
+larger sample size (N=454). There are initially 6796 features in the
+dataset, which is too large to be practically meaningful to analyze.
+Hence, we perform a univariate screening procedure on the features, via
+GAM implemented in \texttt{mgcv}, and choose the the top 200 features
+with smallest p-values. We use 5-knot spline additive models for binary
+outcome using two different models, the proposed BHAM and the SB-GAM.
+10-Fold CV are used to choose the optimal tuning parameters of each
+framework with respect to the default selection criterion implemented in
+the software. Out-of-bag samples are used for prediction performance
+evaluation, where deviance, AUC, Brier score defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
 misclassication error defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) were
-calculated and tabulated in Table \ref{tab:ECB_res}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrrr}
-  \hline
-Methods & deviance & auc & mse & mae & misclassification \\ 
-  \hline
-BHAM-SSL & 455.69 & 0.74 & 0.16 & 0.28 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.26 & 0.24 \\ 
-   \hline
-\end{tabular}
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation} 
-\label{tab:ECB_res}
-\end{table}
+\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
+calculated. BHAM-CD obtains superior AUC, Brier score, and
+misclassification error in the out-of-bag samples compared to SB-GAM
+(Table \ref{tab:ECB_res}).
+
+\begin{center}
+Table \ref{tab:ECB_res} here
+\end{center}
 
@@
 
-We use the proposed models BHAM-SSL to analyze metabolomics data from a
+We use the proposed models BHAM to analyze metabolomics data from a
 recently published study \citep{Bihlmeyer2021} on the association
 between metabolic biomarkers and weight loss, where the dataset is
-publically available \citep{Bihlmeyer2021_data}. In this analysis, we
-primarily focused on the analysis of one of the three studies included,
-weight loss maintenance (WLM) cohort \citep{Svetkey2008}, due to the
+publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
+primarily focuse on the analysis of one of the three studies included,
+weight loss maintenance cohort \citep{Svetkey2008}, due to the
 drastically different intervention effects. In the dataset, 765
@@
 chromatography mass spectrometry. Quality control and natural log
-transformation were performed during metabolites data preparation. The
-outcome of interest was standardized percent change in insulin
+transformation are performed during metabolites data preparation. The
+outcome of interest are standardized percent change in insulin
 resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there were 237
-remaining in the analysis. 5-Knot spline additive models for the
-gaussian outcome are constructed using two different models, the
-proposed BHAM-SSL and the SB-GAM. 10-Fold cross-validation(CV) are used
-to choose the optimal tuning parameters of each framework with respect
-to the default selection criterion implemented in the software.
-Out-of-bag samples were used for prediction performance evaluation,
-where deviance, mean squared error (MSE) defined as
+missing datapoints and addressing outliers in the data, there are
+\(p\)=237 features remaining in the analysis. 5-Knot spline additive
+models for the Gaussian outcome are constructed using two different
+models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
+the optimal tuning parameters of each framework with respect to the
+default selection criterion implemented in the software. Out-of-bag
+samples are used for prediction performance evaluation, where deviance,
+\(R^2\), mean squared error (MSE) defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-abosolute error (MAE)defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) were calculated and
-tabulated in Table \ref{tab:WLM_res}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & deviance & R2 & mse & mae \\ 
-  \hline
-BHAM-SSL & 669.38 & 0.08 & 0.92 & 0.74 \\ 
-  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
-   \hline
-\end{tabular}
-\caption{Measures of optimal BHAM-SSL and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation} 
-\label{tab:WLM_res}
-\end{table}
+absolute error (MAE) defined as
+\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
+BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
+compared to SB-GAM (Table \ref{tab:WLM_res}).
+
+\begin{center}
+Table \ref{tab:WLM_res} here
+\end{center}
 
@@
 fitting. The algorithms can be easily scale up to address
-high-dimensional data analysis in a computational efficient manner, but
-also provide naive inferential supports for many commonly used
-hierarchical priors, e.g.~t-distribution, double exponential
-distribution, mixture t distribution, and mixture double exponential
-distribution, in the analyses that the sparse assumption is weak or not
-necessary. Via simulations, we demonstrated that the proposed model
-provides as good, if not better, prediction performance compared to some
-state-of-the-art non-linear modeling devices.
+high-dimensional data analysis in a computational efficient manner. Via
+simulations, we demonstrated that the proposed model provides as good,
+if not better, prediction performance compared to some state-of-the-art
+non-linear modeling devices.
 
@@
 ones. Ancillary functions are provided for model specification in
-high-dimensional settings, curve plotting and functional selection.
+high-dimensional settings, curve plotting and functional selection. In
+addition, in \texttt{BHAM}, we streamline the model fitting algorithms
+to support other popular Bayesian hierarchical prior for the smoothing
+functions, such as Student's T distribution, double exponential
+distribution, mixture normal and T distribution's. These priors could be
+helpful when the sparse assumption is weak or not necessary.,
 
@@
 selected, but the linear component is not. We currently address it
-analytically by forcing to include the linear component when non-linear
-component is included. Another possible solution is to impose an
-dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, while the
-proposed model addresses a great deal of analytic problem, analyzing the
-time-to-event outcome remains unsolved. Our current effort resides on
-expanding GAM to Cox proportional hazard models. An naive approach would
-be convert an time-to-event outcome to a poisson outcome following
-Whitehead \citep{Whitehead1980}. However, it would be more efficient to
-directly fitting Cox models via penalized pseudo likelihood function.
-Last but not least, we plan to expand the proposed model to integrate
-external biology information. The main motivation for the integrative
-model is that biologically informed grouping of weak effects increases
-the power of detecting true associations between features and the
-outcome \citep{Peterson2016}, and stabilizes the analysis results for
-reproducibility purpose. Such integration can be achieved by setting up
-a structural hyperprior on the inclusion indicator of the smoothing
-function null space \(\boldsymbol{\gamma}^0\). The similar strategy has
-been used in Ferrari and Dunson \citep{Ferrari2020}.
+analytically by including the linear component in the model when
+non-linear component is selected. Another possible solution is to impose
+a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
+i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, the
+computational time for fitting a BHAM model with EM-IWLS algorithms can
+be improved. Current implementation of the EM-IWLS algorithm jointly
+updates all coefficients in each iteration, which requires a lot
+computation resources. This jointly updating procedure can be enhanced
+by adapting a backfitting step \citep{Hastie1987} where each smoothing
+function are updated individually. Thirdly, when using the proposed
+model in real data analysis where a screening procedure is implemented
+before joint predictive modeling. we recommend to include the screening
+procedure in the cross-validation during model selection.
+\citep{Friedman2017} For the sake of model comparison, we fit the models
+using the same pool of predictors in Section \ref{sec:ECB}.
+
+Our future efforts direct to modeling survival outcomes and integrative
+analysis. While the proposed model addresses a great deal of analytic
+problem, analyzing the time-to-event outcome remains unsolved. An naive
+approach would be convert a time-to-event outcome to a Poisson outcome
+following Whitehead \citep{Whitehead1980}. However, it would be more
+efficient to directly fit Cox models via penalized pseudo likelihood
+function \citep{Simon2011}. Meanwhile, with growing understanding of
+biological structure within -omics field, it is appealing to integrate
+external biology information in the modeling process. The main
+motivation for integrative models is that biologically informed grouping
+of weak effects increases the power of detecting true associations
+between features and the outcome \citep{Peterson2016}, and stabilizes
+the analysis results for reproducibility purpose. Such integration can
+be achieved by setting up a structural hyperprior on the inclusion
+indicator of the smoothing function null space
+\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
+and Dunson \citep{Ferrari2020}.
+
+\newpage
 
@@
 
+\newpage
+
+\begin{figure}
+\centering
+\begin{tikzpicture} [
+staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
+mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
+PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
+background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
+]
+
+\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
+  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
+};
+\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
+  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
+};
+
+\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
+\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
+\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
+\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
+
+\draw[->] (Theta) -- (linearGamma);
+\draw[->] (Theta) -- (penGamma);
+\draw[->] (linearGamma) -- (linearBeta) ;
+\draw[->] (penGamma) -- (penBeta);
+\draw[->] (ab) -- (Theta);
+\draw[->] (s) -- (linearBeta) ;
+\draw[->] (s) -- (penBeta);
+\draw[->] (linearBeta) -- (Beta);
+\draw[->] (penBeta) -- (Beta);
+\draw[->] (Beta) --  (Y);
+
+\begin{pgfonlayer}{background}
+  \node [background,
+   fit=(linearGamma) (linearBeta),
+   label=above:Linear Space:] {};
+  \node [background,
+    fit=(penGamma) (penBeta),
+    label=above:Nonlinear Space:] {};
+\end{pgfonlayer}
+
+\end{tikzpicture}
+
+\caption{Directed acyclic graph of the proposed Bayesian hierarchical additive model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
+\label{fig:SSprior}
+\end{figure}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{ccccccc}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
+   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
+  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
+  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.}
+\label{tab:gaus}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllllll}
+  \hline
+P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
+   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
+   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
+  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
+  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.}
+\label{tab:bin_auc}
+\end{table}
+
+\newpage
+
+\begin{sidewaystable}[!h]
+\centering
+\begin{tabular}{lrllllll}
+  \hline
+Distribution & P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+  \hline
+Binomial & 4.00 & 5.88 (1.01) & 6.96 (1.76) & 2.97 (0.82) & 4.60 (2.32) & 0.18 (0.04) & 343.24 (88.37) \\ 
+  Binomial & 10.00 & 14.01 (1.86) & 7.65 (1.25) & 7.95 (2.70) & 10.01 (3.56) & 3.51 (12.03) & 543.13 (133.51) \\ 
+  Binomial & 50.00 & 185.26 (17.44) & 53.83 (6.26) & 109.72 (24.31) & 128.13 (23.26) & 670.01 (151.28) & 1630.42 (193.61) \\ 
+  Binomial & 100.00 & 1106.93 (356.73) & 48.41 (5.84) & 715.15 (170.55) & 718.63 (193.53) & - & 2783.99 (235.31) \\ 
+  Binomial & 200.00 & 6923.62 (1911.59) & 99.98 (7.05) & 5572.51 (1295.18) & 4958.69 (1970.46) & - & 4780.10 (488.66) \\ 
+  Gaussian & 4.00 & 4.22 (1.51) & 2.51 (0.20) & 0.79 (0.11) & 0.67 (0.07) & 0.05 (0.00) & 36.02 (3.02) \\ 
+  Gaussian & 10.00 & 17.59 (5.77) & 33.32 (3.50) & 3.49 (0.70) & 3.47 (0.71) & 0.32 (0.34) & 72.93 (9.52) \\ 
+  Gaussian & 50.00 & 329.83 (44.57) & 334.53 (22.33) & 35.54 (9.15) & 35.44 (9.02) & 73.33 (71.24) & 373.42 (34.93) \\ 
+  Gaussian & 100.00 & 1706.19 (225.92) & 446.98 (79.31) & 150.58 (46.72) & 152.82 (45.07) & - & 684.05 (67.80) \\ 
+  Gaussian & 200.00 & 16379.27 (4825.56) & 303.73 (87.27) & 591.11 (147.88) & 569.92 (107.45) & - & 1314.12 (137.85) \\ 
+   \hline
+\end{tabular}
+\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
+mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
+p = 100, 200.}
+\label{tab:time_sim}
+\end{sidewaystable}
+
+\newpage
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{2pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
+
+\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\endfirsthead
+
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{490.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+
+\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+
+\end{longtable}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Methods & Deviance & AUC & Brier & Misclass \\ 
+  \hline
+BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
+  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
+   \hline
+\end{tabular}
+\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
+\label{tab:ECB_res}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Methods & Deviance & $R^2$ & MSE & MAE \\ 
+  \hline
+BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
+  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
+   \hline
+\end{tabular}
+\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
+\label{tab:WLM_res}
+\end{table}
+
 \end{document}

### 2021-10-21 Manuscript/03-Simulation.Rmd
Revise Simulation

@@
 
-The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
 \begin{center}
@@
 
-The predictive performances have a consistent pattern across the three distributions of outcomes.
-Across all the scenarios, COSSO and adaptive COSSO have the least favorable performance among the applicable methods examined.
-To note, mgcv doesn't support high-dimensional analysis when the number of coefficients are greater than the sample size.
-mgcv predicts well when $p$ is small or moderate (p = 4, 10), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of are small, moderate, and high.
-However, in the hyper high dimensional case, SB-GAM has better performance.
-Among the proposed methods, BHAM-IWLS performs consistently better than BHAM-CD.
+The predictive performances have a consistent pattern across the two distributions of outcomes.
+Across all the scenarios, COSSO and adaptive COSSO have the least favorable performance among the applicable methods examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}).
+To note, mgcv doesn't support high-dimensional analysis, i.e.
+the number of coefficients are greater than the sample size, and hence not evaluated when $p=100, 200$.
+mgcv predicts well when $p$ is small or moderate ($p = 4, 10$), and deteriorate when the number of predictors increase.
+Among the three fast-computing Bayesian hierarchical models, the proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of are moderate (p=4, 10, 50).
+Particularly, BHAM-IWLS performs as good as mgcv if not better.
+However, in high dimensional case where we mimic the situation the signals are extremely sparse, SB-GAM has better performance than the proposed method.
+However, the BHAM-CD has extreme computational advantage over SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
 

### 2021-10-21 Manuscript/04-Real_Data.Rmd
Revise Simulation

@@
 In this section, we apply the proposed models BHAM, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
-We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021].
-Computation time for cross-validation and optimal model fitting steps are recorded and presented in Table \ref{tab:time_real_data}.
+We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}).
 \begin{center}
@@
 Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
-BHAM-CD obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (Table \ref{tab:ECB_res}).
+BHAM-CD obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).
 
@@
 Out-of-bag samples are used for prediction performance evaluation, where deviance, $R^2$, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ are calculated.
-BHAM-CD obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (Table \ref{tab:WLM_res}).
+BHAM-CD obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:WLM_res}).
 \begin{center}

### 2021-10-21 Manuscript/SS_GAM.tex
Revise Simulation

@@
 
-The results are tabulated in Table \ref{tab:gaus}, \ref{tab:bin_auc}.
-
 \begin{center}
@@
 
-The predictive performances have a consistent pattern across the three
+The predictive performances have a consistent pattern across the two
 distributions of outcomes. Across all the scenarios, COSSO and adaptive
 COSSO have the least favorable performance among the applicable methods
-examined. To note, mgcv doesn't support high-dimensional analysis when
-the number of coefficients are greater than the sample size. mgcv
-predicts well when \(p\) is small or moderate (p = 4, 10), and
-deteriorate when the number of predictors increase. Among the three
-fast-computing Bayesian hierarchical models, the proposed models,
-BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of
-are small, moderate, and high. However, in the hyper high dimensional
-case, SB-GAM has better performance. Among the proposed methods,
-BHAM-IWLS performs consistently better than BHAM-CD.
+examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}). To note, mgcv
+doesn't support high-dimensional analysis, i.e.~the number of
+coefficients are greater than the sample size, and hence not evaluated
+when \(p=100, 200\). mgcv predicts well when \(p\) is small or moderate
+(\(p = 4, 10\)), and deteriorate when the number of predictors increase.
+Among the three fast-computing Bayesian hierarchical models, the
+proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when
+the dimension of are moderate (p=4, 10, 50). Particularly, BHAM-IWLS
+performs as good as mgcv if not better. However, in high dimensional
+case where we mimic the situation the signals are extremely sparse,
+SB-GAM has better performance than the proposed method. However, the
+BHAM-CD has extreme computational advantage over SB-GAM (see Table
+\ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
 
@@
 improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}. Computation time
-for cross-validation and optimal model fitting steps are recorded and
-presented in Table \ref{tab:time_real_data}.
+hierarchical additive model, SB-GAM \citep{Bai2021}, while being
+computationally efficient (see Table \ref{tab:time_real_data}).
 
@@
 misclassification error in the out-of-bag samples compared to SB-GAM
-(Table \ref{tab:ECB_res}).
+(see Table \ref{tab:ECB_res}).
 
@@
 BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
-compared to SB-GAM (Table \ref{tab:WLM_res}).
+compared to SB-GAM (see Table \ref{tab:WLM_res}).
 

### 2021-10-21 packages.R
Set up renv for reproducible workflow

@@ -1,31 +1,0 @@
-## library() calls go here
-library(conflicted)
-library(dotenv)
-library(targets)
-library(tarchetypes)
-
-
-library(dplyr)
-library(tidyverse)
-library(rticles)
-library(gtsummary)
-library(here)
-library(mgcv)
-library(broom)
-library(unglue)
-library(knitr)
-library(ggpubr)
-library(xtable)
-library(janitor)
-library(flextable)
-
-
-library(BHAM)
-library(BhGLM)
-library(sparseGAM)
-library(cosso)
-library(glmnet)
-
-conflict_prefer("filter", "dplyr")
-conflict_prefer("mutate", "dplyr")
-conflict_prefer("select", "dplyr") 

### 2021-10-21 Manuscript/SS_GAM.tex
Update Simulation Restuls

@@
 \documentclass[AMA,STIX1COL,]{WileyNJD-v2}
+
+% For Pandoc highlighting needs
 
@@
 additive models (GAMs) to accommodate high-dimensional data (\(p>>n\))
-using group sparse regularization. However, the sparse reuglarization
+using group sparse regularization. However, the sparse regularization
 may induce excess shrinkage when estimating smoothing functions,
@@
 \caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\label{tab:time_real_data}}\\
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\textbackslash{}label\{tab:time\_real\_data\}
+}\\
 
@@
 
-\noalign{\global\setlength{\arrayrulewidth}{1pt}}\arrayrulecolor[HTML]{666666}\cline{2-7}
+\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
 
@@
 
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
 
@@
 
-\noalign{\global\setlength{\arrayrulewidth}{2pt}}\arrayrulecolor[HTML]{666666}\cline{1-7}\endhead
+\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}\endhead
 

### 2021-10-21 _targets.R

@@ -23,7 +23,7 @@ tar_plan(
   #  Simulations Studies -----------------------------------------------------
   #* Main Simulation Study ####
   tar_files(sim_main_path,
-            list.files("~/Downloads/main/", recursive = T, full.names = T)),
+            list.files("Simulation/main/", recursive = T, full.names = T)),
 
   success_rate = create_success_rate_table(sim_main_path),
   # 

### 2021-10-21 _targets.R
Remove IWLS analysis

@@ -166,14 +166,14 @@ tar_plan(
   # measure.bh(bglm_mdl_fnl),
   WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),
   
-  # ** BHAM-IWLS --------------------------------------------------------------
-  WLM_bgam_raw = bgam(out_HOMA_std~.-out_HOMA_std, 
-                      data = data.frame(out_HOMA_std =  WLM_train_dat$out_HOMA_std, WLM_bgam_dat$train_dat),
-                      family = "gaussian",
-                              prior = mde(),
-                              group = WLM_bgam_dat$group),
-  
-  WLM_bgam_cv = tune.bgam(WLM_bgam_raw,s0 = seq(0.005, 0.1, 0.01)),
+  # # ** BHAM-IWLS --------------------------------------------------------------
+  # WLM_bgam_raw = bgam(out_HOMA_std~.-out_HOMA_std, 
+  #                     data = data.frame(out_HOMA_std =  WLM_train_dat$out_HOMA_std, WLM_bgam_dat$train_dat),
+  #                     family = "gaussian",
+  #                             prior = mde(),
+  #                             group = WLM_bgam_dat$group),
+  # 
+  # WLM_bgam_cv = tune.bgam(WLM_bgam_raw,s0 = seq(0.005, 0.1, 0.01)),
   
   # ** SB_GAM ---------------------------------------------------------------
   

### 2022-01-07 Manuscript/R&R/00-response_letter.Rmd
Add Review Comments

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\hrule
\vspace*{0.02cm}
\hrule
\vspace*{1cm}
\Large __Response to Reviewers' Comments__
\vspace*{1cm}
\hrule
\vspace*{0.5cm}
\begin{tabular}{llp{11.5cm}}
Manuscript No. & : & SIM-21-0956 \\
Title & : & Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data\\
\end{tabular}
\vspace*{0.5cm}
\hrule
\vspace*{0.02cm}
\hrule

\vspace*{1cm}

First, we would like to thank you for handling this paper and thank the AE and both reviewers for their valuable comments. We have carefully addressed all the comments. Point-by-point responses are given below, where each comment is quoted in _italics_.

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item xxx
    \item yyy
    \item zzz
\end{itemize}

\newpage
# Response to Reviewer \#1
```{r reviewer1, child="01-reviewer_1.Rmd"}
```

\newpage
# Response to Reviewer \#2
```{r reviewer2, child="02-reviewer_2.Rmd"}
```

\newpage
# Response to Reviewer \#4
```{r reviewer4, child="03-reviewer_4.Rmd"}
```

\newpage
# Reference

### 2022-01-07 Manuscript/R&R/01-reviewer_1.Rmd
Add Review Comments

We sincerely thank you for the valuable comments. We have carefully addressed all the comments. A point-by-point response is given below, where each of your comments is quoted in _italics_. __For a summary of major changes, please see our response to the Associate Editor.__

\bigskip

_Comments to the Authors_

_In this paper, the authors proposed a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis. Specifically, they incorporated smoothing penalties in the model via re-parameterization of smoothing function to avoid overly shrinking basis function coefficients. Incorporating the smoothing penalty allows the separation of the linear space of a smoothing function from the nonlinear space. Then they added a new two-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately. Two scalable optimization algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented._

_The proposed framework, BHAM, provides computational convenience using independent mixture double exponential distribution during model fitting by using optimization algorithms instead of intensive sampling algorithms and addresses the incapability of bi-level selection._

_Here are some comments._

_1. In the real data study, it is very important to establish the correlation between response variables and explanatory variables with appropriate models. The authors should add the comparison of linear models to illustrate that whether using non-linear was better. In addition, except expertise knowledge, are there metrics to judge whether a model is suitable for building model with non-linear part?_

_2. In Weight Loss Maintenance Cohort, R2 was very small by the two GAM methods.  I may think whether the data pretreatment and noise reduction process is appropriate._

_3. In the paper, several priors were mentioned. How to choose the prior in different situations? What is the advantage of the spike-and-slab double exponential prior?_

_4. How to detect the necessary before using the proposed model._

### 2022-01-07 Manuscript/R&R/02-reviewer_2.Rmd
Add Review Comments

We sincerely thank you for the valuable comments. We have carefully addressed all the comments. A point-by-point response is given below, where each of your comments is quoted in _italics_. __For a summary of major changes, please see our response to the Associate Editor.__

\bigskip

_Comments to the Authorss_

_This paper introduces a new way to impose group sparse regularization in high-dimensional generalized additive models (GAMs). Specifically, this work proposes a re-parametrization, which allows separating the predictor space into linear/nonlinear ones. A spike-and-slab LASSO prior is then imposed on the coefficients for the re-parametrized predictors. The authors further uses EM algorithm to fit this model, where the M-step can be solved with either coordinate-descent or iteratively reweighted least squares. Simulation results and metabolomics data analysis are presented._

_The authors claim that the proposed model has the advantage of_
_(1) does not induce excess shrinkage while still estimating a smooth function,_
_(2) allow us to deal with linear/non-linear terms separately and understand if nonlinear terms are necessary_
_(3) more scalable than previous algorithms which require MCMC for model fitting_

_While targeting an important question, the novelty of the solution provided by the authors seem a little bit limited. In addition, the key and most significant step is the re-parametrization, yet there are a few points which I do not completely understand and might need further clarification from the authors:_
_(1) It seems that after re-parametrization, the authors do not further consider the smoothing penalty (Equation (1)) and only include the spike-and-slab LASSO prior. I wonder how does this affect the smoothness of the fitted function and how does it resolves the issue of excess shrinkage._
_(2) I was wondering why the authors choose spike-and-slab LASSO prior in particular. It is claimed in the paper that this prior yields a fast coordinate-descent algorithm, yet coordinate-descent seems also available to other priors as well. I think more discussion on the properties of SSL should be helpful here._
_(3) Since the authors have already separated linear and nonlinear effects, and we might expect linear effects to enter the model first, I wonder why the authors does not incorporate this explicitly into their model by, say, changing the prior for gamma’s to impose different sparsity._

_Besides, some writing of this paper should better be polished and some typos corrected. For example:
(1) In page 3, line 31, what does "in the prior density function" mean?
(2) Same page, line 43 to line 45, the notation here is a little bit confusing.
(3) Page 5, line 17-18, how shall we understand this statement and any explanations?
(4) Page 5 line 37 and page 6 line 21-22, the implication seems to be ‘SSL is not a continuous prior’, which is not true.
(5) Page 6, line 16, missing a period at the end.
(6) Page 6, line 30, there are duplicate "the’".
(7) Page 6, line 56, ‘converge’ should be ‘converges’.
(8) Page 9, line 24-25, there seems to be duplicated "variance-covariance matrix".
(9) page 11, line 8, there are some problems with the table reference.
(10) In section 4, what is definition for the "deviance"?_

### 2022-01-07 Manuscript/R&R/03-reviewer_4.Rmd
Add Review Comments

We sincerely thank you for the valuable comments. We have carefully addressed all the comments. A point-by-point response is given below, where each of your comments is quoted in _italics_. __For a summary of major changes, please see our response to the Associate Editor.__

\bigskip

_Comments to the authors_

## _1. Summary_

_The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs). The BHAM framework allows for variable selection and estimation in GAMs. However, in contrast to previous works on Bayesian GAMs, bilevel selection is possible under BHAM, and BHAM allows practitioners to more easily distinguish linear effects from nonlinear effects. Two deterministic algorithms, EM-Coordinate Descent (EM-CD) and EM-Iterative Weighted Least Squares (EM-IWLS), are introduced to fit BHAM. The EM-IWLS algorithm allows for inference, which is its main advantage over EM-CD. The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty._

_The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches. However, in order to be suitable for publication in Statistics in Medicine, I have a few concerns that I think the authors should address. Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [1, 2] or the spikeslabGAM of [3]). I detail my main concerns below._

## _2. Main Comments and Questions_
_The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space. This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones. However, one concern is that the method might select the nonlinear effects but fail to select the linear component. The authors brie
y allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "\[include] the linear component in the model when non-linear component is selected."_

\em 
Could the authors elaborate on how they do this? I think this point should also be
stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced. One potential downside of
using the double exponential priors on individual basis coefficients, as opposed to the
group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are
regularized. So it is conceivable that BHAM would end up including the nonlinear
component but not the linear one, which is not sensible. How can strong hierarchy
be enforced to ensure that the linear component is always included if the nonlinear
one is selected (but not necessarily the other way around)? This should be detailed in
the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce
strong hierarchy.

For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
state in their algorithm that they first check whether or not the group is selected,
and only if the group is nonzero do they then proceed with regularizing the individual
coordinates in the nonzero groups. Is a similar check being done for BHAM? This
needs to be explained further and made explicit. If necessary, the authors of the
present manuscript could include an additional subsection in Section 2 that explains
how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
nonlinear effect for the jth covariate is selected).

2. Although the authors claim that BHAM is beneficial because it can distinguish linear
effects from nonlinear effects (therefore, we can tell if nonlinear effects are necessary),
this touted benefit is not well-illustrated in the Simulation Study (Section 3) or the
metabolomics data (Section 4). In particular, the simulation setting in Section 3.1
contains only one linear function. The authors should expand their simulation study
to include a greater number of linear functions (in addition to a few nonlinear ones)
to show how well BHAM is truly distinguishing the linear effects from the nonlinear
effects.

It would also be useful to report things like FDR, FNR, Matthews correlation coef-
ficient, etc. for the function selection to see how well the proposed BHAM model
is performing with respect to selection of the nonzero effects, especially compared to
other methods like SBGAM or spikeslabGAM. In addition, if the authors could re-
port a performance metric of how well BHAM is distinguishing nonlinear from linear
functions, e.g. possibly reporting the proportion of simulations where the linear func-
tion (resp. nonlinear) was correctly identified as linear (resp. nonlinear), that would
provide empirical support for BHAM over existing methods.

3. The authors claim that BHAM is computationally beneficial because it does not rely
on MCMC. The authors illustrate the computational benefit of the double exponential
prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
should also compare it to the method spikeslabGAM of [3]. There is an R package
spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
slab prior on the basis coefficients and is implemented using MCMC. In order to assess
the benefit of using double exponential priors in BHAM vs. Gaussian spike-and-slab
priors, as well as illustrating the computational efficiency of the EM-CD and EM-
IWLS algorithms compared to MCMC, the authors should include comparisons to
spikeslabGAM.

4. The authors introduce the EM-IWLS method which can be used for inference of the
regression coefficients. Does this method work for p > n? Or is it only limited to the
case where p $\leq$ n? Even if this algorithm is limited to only the p $\leq$ n case, the benefit
of inference is not illustrated anywhere in the manuscript. Could the authors report
how to use the standard errors from the error variance-covariance matrix returned from
EM-IWLS to construct pointwise confidence intervals? There should be a section in
the manuscript for Uncertainty Quantification which explains how to use the results
from the EM-IWLS algorithm for the regression coefficients to construct pointwise
confidence intervals for the univariate functions themselves.

In addition, the coverage probability of the 95% pointwise intervals constructed by
BHAM in simulations should also be reported. Furthermore, the methods mgcv and
spikeslabGAM also return interval estimates, so the authors would do well to compare
the coverage probability of the confidence intervals constructed by BHAM vs. the
confidence/credible intervals returned by mgcv and spikeslabGAM.

5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
as a DAG. However, there are no other figures in the paper, e.g. plots of function
estimates from BHAM. The authors may consider adding a plot of the function esti-
mates vs. the ground truth in the simulation study (Section 3) for one of the nonlinear
functions and one of the linear functions (these could be, for instance, the left panel
and right panel of a figure). This will illustrate that BHAM can accurately recover
both nonlinear functions and linear functions.

The authors may also give one or two plots from the real data analysis in Section 4 of
some of the significant effects that they discovered. These plots should contain not just
the point estimates but also the the pointwise confidence intervals from the EM-IWLS
method. This will illustrate the benefit of EM-IWLS since it allows practitioners to
assess the uncertainty, in addition to an estimate of the function itself.

## _3. Minor Comments_
1. In Section 1, the authors say that nonparametric methods are an alternative to "black
box" methods like random forest. This is a bit misleading, since the "black box"
methods (random forest, deep learning, etc.) are als nonparametric methods. The
authors might consider rephrasing it (e.g. they could say that semiparametric methods
like GAMs allow for easier interpretation and selection of the significant effects than
the fully nonparametric "black box" methods).

### 2022-01-09 Manuscript/R&R/02-reviewer_2.Rmd
Update R&R

@@
 _The authors claim that the proposed model has the advantage of_
-_(1) does not induce excess shrinkage while still estimating a smooth function,_
-_(2) allow us to deal with linear/non-linear terms separately and understand if nonlinear terms are necessary_
-_(3) more scalable than previous algorithms which require MCMC for model fitting_
+
+  1. _does not induce excess shrinkage while still estimating a smooth function,_
+  2. _allow us to deal with linear/non-linear terms separately and understand if nonlinear terms are necessary_
+  3. _more scalable than previous algorithms which require MCMC for model fitting_
 
@@
 In addition, the key and most significant step is the re-parametrization, yet there are a few points which I do not completely understand and might need further clarification from the authors:_
+
 _(1) It seems that after re-parametrization, the authors do not further consider the smoothing penalty (Equation (1)) and only include the spike-and-slab LASSO prior.
 I wonder how does this affect the smoothness of the fitted function and how does it resolves the issue of excess shrinkage._
+
+__Response__:
+\bigskip
+
 _(2) I was wondering why the authors choose spike-and-slab LASSO prior in particular.
@@
 I think more discussion on the properties of SSL should be helpful here._
+
+__Response__:
+\bigskip
+
 _(3) Since the authors have already separated linear and nonlinear effects, and we might expect linear effects to enter the model first, I wonder why the authors does not incorporate this explicitly into their model by, say, changing the prior for gamma’s to impose different sparsity._
 
-_Besides, some writing of this paper should better be polished and some typos corrected. For example:
-(1) In page 3, line 31, what does "in the prior density function" mean?
-(2) Same page, line 43 to line 45, the notation here is a little bit confusing.
-(3) Page 5, line 17-18, how shall we understand this statement and any explanations?
-(4) Page 5 line 37 and page 6 line 21-22, the implication seems to be ‘SSL is not a continuous prior’, which is not true.
-(5) Page 6, line 16, missing a period at the end.
-(6) Page 6, line 30, there are duplicate "the’".
-(7) Page 6, line 56, ‘converge’ should be ‘converges’.
-(8) Page 9, line 24-25, there seems to be duplicated "variance-covariance matrix".
-(9) page 11, line 8, there are some problems with the table reference.
-(10) In section 4, what is definition for the "deviance"?_
+__Response__:
+\bigskip
+
+\em
+Besides, some writing of this paper should better be polished and some typos corrected. For example:
+\begin{enumerate}
+\item In page 3, line 31, what does "in the prior density function" mean?
+\item Same page, line 43 to line 45, the notation here is a little bit confusing.
+\item Page 5, line 17-18, how shall we understand this statement and any explanations?
+\item Page 5 line 37 and page 6 line 21-22, the implication seems to be ‘SSL is not a continuous prior’, which is not true.
+\item Page 6, line 16, missing a period at the end.
+\item Page 6, line 30, there are duplicate "the’".
+\item Page 6, line 56, ‘converge’ should be ‘converges’.
+\item Page 9, line 24-25, there seems to be duplicated "variance-covariance matrix".
+\item page 11, line 8, there are some problems with the table reference.
+\item In section 4, what is definition for the "deviance"?
+\end{enumerate}
+\em
+
+__Response__:
+\bigskip

### 2022-01-09 Manuscript/R&R/03-reviewer_4.Rmd
Update R&R

@@
 ## _2. Main Comments and Questions_
-_The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space.
+\em 1.
+The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space.
 This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones.
 However, one concern is that the method might select the nonlinear effects but fail to select the linear component.
-The authors brie
-y allude to this in the Conclusion (p. 13 of the manuscript). The authors say that they "\[include] the linear component in the model when non-linear component is selected."_
+The authors briefly allude to this in the Conclusion (p.
+13 of the manuscript).
+The authors say that they "[include] the linear component in the model when non-linear component is selected."
 
-\em 
-Could the authors elaborate on how they do this? I think this point should also be
-stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced. One potential downside of
-using the double exponential priors on individual basis coefficients, as opposed to the
-group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are
-regularized. So it is conceivable that BHAM would end up including the nonlinear
-component but not the linear one, which is not sensible. How can strong hierarchy
-be enforced to ensure that the linear component is always included if the nonlinear
-one is selected (but not necessarily the other way around)? This should be detailed in
-the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce
-strong hierarchy.
+Could the authors elaborate on how they do this?
+I think this point should also be stressed and made more explicit earlier in the manuscript, possibly in Section 2 where the EM-CD and EM-IWLS algorithms are introduced.
+One potential downside of using the double exponential priors on individual basis coefficients, as opposed to the group spike-and-slab prior in SBGAM, is the fact that only individual coefficients are regularized.
+So it is conceivable that BHAM would end up including the nonlinear component but not the linear one, which is not sensible.
+How can strong hierarchy be enforced to ensure that the linear component is always included if the nonlinear one is selected (but not necessarily the other way around)?
+This should be detailed in the EM-CD and EM-IWLS algorithms, and the authors should explain how to enforce strong hierarchy.
 
@@
 how to enforce strong hierarchy (i.e. linear effect is always selected if the corresponding
-nonlinear effect for the jth covariate is selected).
+nonlinear effect for the jth covariate is selected).\em
 
-2. Although the authors claim that BHAM is beneficial because it can distinguish linear
+__Response__:
+\bigskip
+
+\em 2. Although the authors claim that BHAM is beneficial because it can distinguish linear
 effects from nonlinear effects (therefore, we can tell if nonlinear effects are necessary),
@@
 tion (resp. nonlinear) was correctly identified as linear (resp. nonlinear), that would
-provide empirical support for BHAM over existing methods.
+provide empirical support for BHAM over existing methods.\em
 
-3. The authors claim that BHAM is computationally beneficial because it does not rely
+__Response__:
+\bigskip
+
+\em 3. The authors claim that BHAM is computationally beneficial because it does not rely
 on MCMC. The authors illustrate the computational benefit of the double exponential
@@
 IWLS algorithms compared to MCMC, the authors should include comparisons to
-spikeslabGAM.
+spikeslabGAM.\em
 
-4. The authors introduce the EM-IWLS method which can be used for inference of the
+__Response__:
+\bigskip
+
+\em 4. The authors introduce the EM-IWLS method which can be used for inference of the
 regression coefficients. Does this method work for p > n? Or is it only limited to the
@@
 the coverage probability of the confidence intervals constructed by BHAM vs. the
-confidence/credible intervals returned by mgcv and spikeslabGAM.
+confidence/credible intervals returned by mgcv and spikeslabGAM.\em
 
-5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
+__Response__:
+\bigskip
+
+\em 5. The authors only provide one figure (Figure 1) which illustrates the hierarchical model
 as a DAG. However, there are no other figures in the paper, e.g. plots of function
@@
 method. This will illustrate the benefit of EM-IWLS since it allows practitioners to
-assess the uncertainty, in addition to an estimate of the function itself.
+assess the uncertainty, in addition to an estimate of the function itself.\em
+
+__Response__:
+\bigskip
 
 ## _3. Minor Comments_
-1. In Section 1, the authors say that nonparametric methods are an alternative to "black
+\em 1. In Section 1, the authors say that nonparametric methods are an alternative to "black
 box" methods like random forest. This is a bit misleading, since the "black box"
@@
 like GAMs allow for easier interpretation and selection of the significant effects than
-the fully nonparametric "black box" methods).
+the fully nonparametric "black box" methods).\em
+
+__Response__:
+\bigskip

### 2022-01-10 Manuscript/00-main.Rmd
typo
Derivation for including linear & nonlinear selection hierarchy

@@
 abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate high-dimensional data ($p>>n$) using group sparse regularization.
-However, the sparse reuglarization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance.
+However, the sparse regularization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance.
 Moreover, most of these GAMs consider an “all-in-all-out” approach for functional selection, rendering them difficult to answer if nonlinear effects are necessary.
@@
   - \usepackage{pdflscape}
+  - \usepackage{color}
+  - \usepackage{soul}
 ---
@@
 \newcommand{\simiid}{\overset{\text{iid}}{\sim}}
+\newcommand{\bg}[1]{\textcolor{red}{#1}}
+<!-- \newcommand{\st}[1]{} -->
 
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \theta_j)$ };\\
+  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \bg{\gamma_{j}}\theta_j)$ };\\
   \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
@@
 \draw[->] (penGamma) -- (penBeta);
+\draw[->, draw = red] (linearGamma) -- (penGamma);
 \draw[->] (ab) -- (Theta);

### 2022-01-10 Manuscript/02-Method.Rmd
Revise the rest of paper
Derivation for including linear & nonlinear selection hierarchy

@@
 &\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j).
+&\gamma_{j}^\tp | \bg{\gamma_{j}, } \theta_j \sim Bin(1, \bg{\gamma_{j}}\theta_j).
 \end{aligned}
 $$
+\bg{Equivalently, we can derive the marginal distribution of $\gamma_{j}^\tp$ by integrate out $\gamma_j$ (Supplement 1), 
+$$
+\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2).
+$$
+}
+
 This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
@@
 To note, the two proposed algorithms provides different utilities.
-The EM-CD algorithm is specifically for fitting BHAM-SSL with an expedited performance, recommending to use in high and ultra-high dimensional setting.
+The EM-CD algorithm is specifically for fitting BHAM with an expedited performance, recommending to use in high and ultra-high dimensional setting.
 A specific concern of EM-CD algorithm is that it provides no information for inference.
@@
 \end{align*}
-and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = \pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show

### 2022-01-10 Manuscript/10-Supplementary.Rmd
Derivation for including linear & nonlinear selection hierarchy

---
title: "Supplementary File"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}

## Supplementary 1: Marginal Distribution of $\gamma_j^\tp$
Given that $\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)$ where $\gamma_{j} | \theta_j \sim Bin(1, \theta_j)$, we can derive the the marginal distribution of $\gamma_{j}^\tp$ with the following manipulation.
\begin{align*}
& \pr(\gamma_j^\tp = 1 | \theta_j)  = \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 1 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^\tp \text{ and }\gamma \text{.]}\\
= & \pr(\gamma_j^\tp = 1| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j)\\
= & \theta_j^2\\
& \pr(\gamma_j^\tp = 0 | \theta_j)  = \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0, \gamma_j = 0 | \theta_j)\\
= & \pr(\gamma_j^\tp = 0| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j) + \pr(\gamma_j^\tp = 0| \gamma_j = 0,  \theta_j)\pr(\gamma_j = 0| \theta_j)\\
= & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
\end{align*}

### 2022-01-10 Manuscript/SS_GAM.tex
Derivation for including linear & nonlinear selection hierarchy

@@
 \usepackage{pdflscape}
+\usepackage{color}
+\usepackage{soul}
 \usepackage{multirow}
@@
 \newcommand{\simiid}{\overset{\text{iid}}{\sim}}
+\newcommand{\bg}[1]{\textcolor{red}{#1}}
 
@@
 &\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j).
+&\gamma_{j}^*| \textcolor{red}{\gamma_{j}, } \theta_j \sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j).
 \end{aligned}
-\] This is to leverage the fact that the probability of selecting the
-bases of a smoothing function should be similar, while allowing
-different penalty on the linear space and non-linear space of the
-smoothing function. The hyper prior of \(\gamma_{j}\) decides the
-sparsity of the model at the functional selection level, while that of
-\(\gamma_{j}^*\) decides the smoothness of the spline function at basis
-function level. Meanwhile, we specify that \(\gamma_{j}\) and
-\(\gamma_{j}^*\) are independently distributed for analytic simplicity.
-We further specify the parameter \(\theta_j\) follows a beta
-distribution with given shape parameters \(a\) and \(b\), \[
+\]
+\textcolor{red}{Equivalently, we can derive the marginal distribution of $\gamma_{j}^*$ by integrate out $\gamma_j$ (Supplement 1), 
+$$
+\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2).
+$$
+}
+
+This is to leverage the fact that the probability of selecting the bases
+of a smoothing function should be similar, while allowing different
+penalty on the linear space and non-linear space of the smoothing
+function. The hyper prior of \(\gamma_{j}\) decides the sparsity of the
+model at the functional selection level, while that of \(\gamma_{j}^*\)
+decides the smoothness of the spline function at basis function level.
+Meanwhile, we specify that \(\gamma_{j}\) and \(\gamma_{j}^*\) are
+independently distributed for analytic simplicity. We further specify
+the parameter \(\theta_j\) follows a beta distribution with given shape
+parameters \(a\) and \(b\), \[
 \theta_j \sim Beta(a, b).
@@
 p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{align*} and
-\(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = \text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta_j\),
+\end{align*} and \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \theta_j\),
+\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j\),
+\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
+\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j}\),
 \(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \theta_j)$ };\\
+  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j)$ };\\
   \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\

### 2022-01-11 GAM Effect Heredity.Rmd
add Email reply to Dr. Yi

---
title: "GAM Effect Heredity with Spike-and-slab prior"
output: pdf_document
---
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}

# Prior of Selection Indicator

We assume that the linear and non-linear coefficients $\beta_{j}$ and $\bs \beta_{j\cdot}$ follows the spike and slab lasso prior with selection indicator $\gamma_j$ and $\gamma_j^\tp$ respectively. We further assume both $\gamma_j$ and $\gamma_j^\tp$ follow binomial distribution of trial 1,
\begin{align*}
\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\\
\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^\tp),
\end{align*}
  where $\theta_j^\tp = w_j\theta_j$ and $w_j \in [0,1]$ is a dynamic weight implementing the effect heredity (Chipman, 2004). Depends on the definition of effect heredity , $w_j$ can be a constant (hyperparameter $\pi_4$ for quadratic terms defined in Chipman (2004)), a function of $Pr(\gamma_j=1)$ (implemented in `BhGLM`). If we define effect heredity to be strict, i.e. effect forcing (Chipman, 2004), we can have $w_j = \gamma_j$ such that $\gamma_j^\tp = 1$ enters the model only when $\gamma_j$ enters the model. At last, we assume $\theta_j \sim Beta(1,1)$ for calculation simplicity.

With this formulation,

### 2022-01-25 Simulation/Code/Main.R
Change Naming Convention

@@ -190,7 +190,7 @@ names(mdls) <- c(
   # "bglm_t", "bglm_de", "blasso",
   # "bglm_t_group", "bglm_de_group", "blasso_group",
   # "bglm_spline_t", 
-  "bglm_spline_de", "blasso_spline"#,
+  "bgam", "bamlasso"#,
   # "bglm_share_t"
 )
 

### 2022-01-25 Simulation/Code/start_sim.R

@@ -5,8 +5,8 @@ sim_prmt <- expand.grid(
   # n_train = c(500),
   # n_test = c(10000),
   dis = c("gaussian", "binomial")
-) %>% slice(1)
-  # data.frame()
+) %>% filter(p==200)
+   # data.frame()
 
 # A wrapper function to set up each job
 start.sim <- function(
@@ -24,7 +24,7 @@ start.sim <- function(
   #                   # ntest.name, 
   #                   sep="-")
   
-  job.name <- paste("bgam_", dis.name, p.name,
+  job.name <- paste("bgam", dis.name, p.name,
                     # ntrain.name,
                     # ntest.name, 
                     sep="-")
@@ -47,8 +47,8 @@ start.sim <- function(
 }
 
 # Delete Previous Log Files
-unlink("/data/user/boyiguo1/bgam/log/", recursive = TRUE)
-unlink("/data/user/boyiguo1/bgam/sim_res/main/", recursive = TRUE)
+# unlink("/data/user/boyiguo1/bgam/log/", recursive = TRUE)
+# unlink("/data/user/boyiguo1/bgam/sim_res/main/", recursive = TRUE)
 
 
 

### 2022-01-25 R/Sim_Res/make_sim_main_table.R
Reorganize the result summary section

@@ -2,26 +2,19 @@ make_sim_main_table <- function(success_rate, dist, section = "test" , measures)
   # browser()
   
   total_dat <- success_rate %>% 
-    # filter(dist==dist, n_success!=0) %>% 
     filter(dist=={{dist}}, n_success!=0) %>%
     pull(path) %>% 
     map_dfr( .f = function(sim){
-      
-      
+      # browser()
       sim.df <- unglue_data(sim, 
-                            "{}/bgam_{study}_-dis_{dist}-p_{p}") %>% 
+                            "{}/{}-dis_{dist}-p_{p}") %>% 
         mutate( p = as.numeric(p))
-      fls <- list.files(sim, full.names = TRUE) 
-      fls <- fls[grep(".rds", x=fls)]
-      n <- length(fls)
       
-      
-      if(n == 0)
-        return(data.frame(NULL))
-      
-      # .file <- fls[1]
-      ret <- fls %>%
+      # fls <- 
+        list.files(sim, full.names = TRUE) %>% 
+        grep(".rds", x=., value = TRUE) %>%
         map_dfr(.f = function(.file){
+          # browser()
           it <- unglue_data(.file, "{whatever}/it_{it}.rds") %>% pull(it) %>% as.numeric
           ret <- data.frame(
             it = it,
@@ -33,16 +26,15 @@ make_sim_main_table <- function(success_rate, dist, section = "test" , measures)
           )
         }) %>%
         arrange(it) %>% 
-        mutate(p = sim.df$p)#  %>% 
-      
+        mutate(p = sim.df$p)
     }) %>% 
     mutate(method = factor(
-      method,
-      levels = c("bglm_t", "bglm_de", "blasso",
-                 "bglm_t_group", "bglm_de_group", "blasso_group",
-                 "bglm_spline_t", "bglm_spline_de", "blasso_spline",
-                 "cosso", "acosso", "mgcv", "SB_GAM"
-      )
+      method#,
+      # levels = c("bglm_t", "bglm_de", "blasso",
+      #            "bglm_t_group", "bglm_de_group", "blasso_group",
+      #            "bglm_spline_t", "bglm_spline_de", "blasso_spline",
+      #            "cosso", "acosso", "mgcv", "SB_GAM"
+      # )
     )
     )
   
@@ -50,8 +42,8 @@ make_sim_main_table <- function(success_rate, dist, section = "test" , measures)
     select(p, method, {{measures}}
[…]

### 2022-01-25 Manuscript/00-main.Rmd

@@
 
-tar_load(tab_binom)
-tar_load(tab_gaus)
+tar_load(sim_binom_tab)
+tar_load(sim_gaus_tab)
 
@@
 tar_load(tbl_real_time)
-tar_load(tab_time)
+tar_load(sim_tim_tab)
 ```
@@
 ```{r, results = "asis"}
-tab_gaus$mgcv[4] <- "-"
-tab_gaus$mgcv[5] <- "-"
-tab_gaus %>% 
+sim_gaus_tab$mgcv[4] <- "-"
+sim_gaus_tab$mgcv[5] <- "-"
+sim_gaus_tab %>% 
   mutate(p = as.integer(p)) %>% 
@@
 ```{r, results = "asis"}
-tab_binom$mgcv[4] <- "-"
-tab_binom$mgcv[5] <- "-"
-tab_binom %>% 
+sim_binom_tab$mgcv[4] <- "-"
+sim_binom_tab$mgcv[5] <- "-"
+sim_binom_tab %>% 
   mutate(p = as.integer(p)) %>% 
@@
 ```{r results = "asis"}
-tab_time %>% 
+sim_tim_tab %>% 
   select(dist, p, `BHAM-IWLS`,`BHAM-CD`, cosso, acosso, mgcv, `SB-GAM`) %>% 

### 2022-01-26 Simulation/Code/Main.R
Add variable selection to the simulation (#14)

@@ -38,7 +38,7 @@ source("~/GitHub/Manuscript-BHAM/Simulation/Code/sim_pars_funs.R")
 
 # Using Array ID as seed ID
 it <- Sys.getenv('SLURM_ARRAY_TASK_ID') %>% as.numeric
-# it <- 1
+# it <- 2
 
 set.seed(it)
 fam_fun <- dis()  
@@ -67,8 +67,16 @@ lasso_mdl <- cv.glmnet(x = data.matrix(dat %>% select(-y)),
                        y = data.matrix(dat %>% pull(y)),
                        nfolds = 5, family = fam_fun$family
                       )
+lasso_fnl_mdl <- glmnet(x = data.matrix(dat %>% select(-y)),
+                        y = data.matrix(dat %>% pull(y)),
+                        family = fam_fun$family, lambda = lasso_mdl$lambda.min)
+
+# predict(lasso_fnl_mdl, type = "nonzero")
+
 lasso_train <- measure.glm(dat$y, predict(lasso_mdl, newx = data.matrix(dat %>% select(-y)), type = "response", s = "lambda.min"), family = fam_fun$family)
 lasso_test <- measure.glm(test_dat$y, predict(lasso_mdl, newx = data.matrix(test_dat %>% select(-y)), type = "response", s = "lambda.min"), family = fam_fun$family)
+
+lasso_vs <- ((lasso_fnl_mdl$beta %>% as.vector())!=0) %>% `names<-`(setdiff(names(dat), "y"))
 
 #### mgcv_mdl ####
 
@@ -91,10 +99,11 @@ error = function(err) {
 mgcv_sum <- NULL
 mgcv_train <- make_null_res(fam_fun$family)
 mgcv_test <- make_null_res(fam_fun$family)
+mgcv_vs <- rep(NA, length(lasso_vs)) %>% `names<-`(setdiff(names(dat), "y"))
 
 
 if(!is.null(mgcv_mdl)){
-  mgcv_sum <- summary(mgcv_mdl)$s.table[,"p-value"]
+  mgcv_vs <- (summary(mgcv_mdl)$s.table[,"p-value"] < 0.05) %>% `names<-`(setdiff(names(dat), "y"))
   
   
   mgcv_train <- measure.glm(mgcv_mdl$y, predict(mgcv_mdl, type = "response"), family = fam_fun$family)
@@ -210,7 +219,13 @@ bham_time <- imap_dfr(mdls, .f = function(.mdl, .y){
[…]

### 2022-02-11 Manuscript/R&R/01-reviewer_1.Rmd
Update R&R
Add R&R to the target pipeline; start to draft response.

@@
 
+\bigskip
+
 _Here are some comments._
@@
 
+__Response__: Thank you for the comment.
+We expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds.
+Please see the changes and conclusion in \emph{Section 3: SIMULATION STUDY}.
+To conclude, we observed [TODO: REPLACE HERE WITH CONCLUSION].
+Hence, the proposed models perform better than linear models when functional are non-linear and as good as linear model when functions are linear.
+The expanded simulation study also demonstrates that the proposed model is flexible to model both linear and non-linear effects and broadly applicable in high-dimensional modeling without much constraints, which deems to be an additional strength.
+
+\bigskip
+
 _2. In Weight Loss Maintenance Cohort, R2 was very small by the two GAM methods.  I may think whether the data pretreatment and noise reduction process is appropriate._
+
+__Response__:
+\bigskip
 
@@
 
+__Response__:
+\bigskip
+
 _4. How to detect the necessary before using the proposed model._
+
+__Response__:

### 2022-02-11 _targets.R

@@ -235,8 +235,16 @@ tar_plan(
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
-  # tar_render(manu_ENAR, "Manuscript/00-main_ENAR.Rmd",
-  #            output_file = "SS_GAM-ENAR.pdf"),
+  # R & R Letter --------------------------------------------------------------
+  tar_target(RR_path,
+             c("Manuscript/R&R/01-reviewer_1.Rmd",
+               "Manuscript/R&R/02-reviewer_2.Rmd",
+               "Manuscript/R&R/03-reviewer_4.Rmd",
+               "Manuscript/R&R/reply_ref.bib"
+             ),
+             format = "file"),
   
+  tar_render(RR, "Manuscript/R&R/00-response_letter.Rmd",
+             output_file = "response_SS_GAM.pdf"),
 )
 

### 2022-02-11 Manuscript/R&R/01-reviewer_1.Rmd
Update response

@@
 
-__Response__:
+__Response__: We don't think any preliminary checkings or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed.
+Such rules of thumb includes .[Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.] As a supporting evidence, our added simulation demonstrates the proposed model performs well, in reference to the linear lasso model, even when signals are linear and sparse (see response to the Comment 1).
+
+We acknowledge the reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge.
+However, the discussion of modeling strategy, e.g.
+starting with more flexible models or linear models, or using statistical testings to examine assumptions, is out of the scope of the current manuscript.
+The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals.

### 2022-02-11 Manuscript/R&R/00-response_letter.Rmd
Update R&R
Add R&R to the target pipeline; start to draft response.
Added Changes

@@
 ---
-output: pdf_document
+output: 
+  pdf_document:
+    keep_tex: FALSE
+    citation_package: biblatex
+bibliography: reply_ref.bib
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(targets)
+tar_load(RR_path)
 ```
+
+\newcommand{\tp}{*}
 
@@
 \begin{itemize}
-    \item xxx
-    \item yyy
-    \item zzz
+    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects inclusion indicator such that the strong hierarchy for bi-level selection is enforced.
+    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear lasso model and spikeSlabGAM model among the models to be compared. 
+    \item According to Reviewer \#1's and Reviewer \#4's comments, we added a new set of simulation scenarios where the functions of predictors are linear.  
+    \item According to Reviewer \#4's comment, we summarized the variable selection performance for all models of comparison and bi-level selection performance for the models that is capable, i.e.
+the proposed model and spikeSlabGAM.
+    \item According to Reviewer \#4's comment, we recorded the computation performance of spikeSlabGAM.
+    \item All existing figure and tables were updated to reflect the methodology change and simulation design change.
+Figures were added according to Reviewer \#4's comment.
+[TODO: think what figures are needed]
+    \item We fixed the additional typos and grammar errors according to Review \#2.The revision of the manuscript is marked in red [TODO: change this color to red]. Deletions are not highlighted.
 \end{itemize}

### 2022-02-11 Manuscript/R&R/01-reviewer_1.Rmd
Added Changes

@@
 __Response__: We don't think any preliminary checkings or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed.
-Such rules of thumb includes .[Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.] As a supporting evidence, our added simulation demonstrates the proposed model performs well, in reference to the linear lasso model, even when signals are linear and sparse (see response to the Comment 1).
+Such rules of thumb includes [TODO: Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.].
+As a supporting evidence, our added simulation demonstrates the proposed model performs well, in reference to the linear lasso model, even when signals are linear and sparse (see response to the Comment 1).
 

### 2022-02-11 Manuscript/R&R/02-reviewer_2.Rmd
Added Changes

@@
 
-__Response__:
+__Response__: Thank you for the comment.
+We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$ (See Equation ([TODO: add the equation number])), in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
+This prior formulation is motivated by effect forcing discussed in (Chipman, 2004).
+Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
+The modification of the prior doesn't complicate the model fitting algorithm.
+The $\gamma$ can be analytically integrated out of the density function (See Equation ([TODO: add the equation number]) and appendix) and requires minimum change of the proposed algorithms.
+In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
+
 \bigskip

### 2022-02-15 Manuscript/R&R/00-response_letter.Rmd
Move EM-IWLS to supproting Information

@@
 bibliography: reply_ref.bib
+header-includes: 
+ - \usepackage{color}
 ---
+
+\newcommand{\bg}[1]{\textcolor{red}{#1}}
 
@@
     \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects inclusion indicator such that the strong hierarchy for bi-level selection is enforced.
+    \item We moved the EM-IWLS fitting algorithm to the supporting information for two reasons.
+First of all, we think moving the EM-IWLS to the supporting information can help the audience focus on the model and EM-Coordinate Descent algorithm, avoiding confusing the audience.
+Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g.
+constructing hypothesis testings and confidence band, requires a more dedicated discussion.
+We are currently preparing another manuscript discussing those.
+Lastly, by moving the EM-IWLS to the supporting information would not hurt the integrity of the current manuscript, that emphasize the scalability and bi-level selection.
     \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear lasso model and spikeSlabGAM model among the models to be compared. 
@@
 [TODO: think what figures are needed]
-    \item We fixed the additional typos and grammar errors according to Review \#2.The revision of the manuscript is marked in red [TODO: change this color to red]. Deletions are not highlighted.
+    \item We fixed the additional typos and grammar errors according to Review \#2. The revision of the manuscript is marked in \bg{red}. Deletions are not highlighted.
 \end{itemize}

### 2022-02-15 Manuscript/02-Method.Rmd
Move EM-IWLS to supproting Information

@@
 
-### EM-IWLS
-
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
-The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
-In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior
-<!-- , with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary -->
-.
-
-<!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
-
-A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
-\begin{align*}
-  \beta|\tau^2 &\sim N(0, \tau^2)\\
-  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
-\end{align*}
-For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
-The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
-\begin{equation}\label{eq:Q1_IWLS}
-Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
-\end{equation}
-Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
-To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
-$E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
-As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}).
-On the other hand, $\tau^{2}$, following gamma distributions, is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
-$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
-\begin{align*}
- E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
-\end{align*}
-where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
-\begin{equation}\label{eq:EQ1_IWLS}
-E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
-\end{equation}
-<!-- $E(Q_1)$ can be seen as a $l_2$ penalized likelihood function where the regularization parameter $\lambda = 2E({\tau}^{-2})$. Equivalently, we have  -->
-$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$, and we can formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal,
-$$
-\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
-$$
-
-Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood:
-$$
-f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
-$$
-where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. The pseudo-response and the pseudo-weight are calculated by
-$$
-\begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
-\end{aligned}
-$$
-where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
-
-With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
-We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*}
-& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
-  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
-  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
-\end{align*}
- such that 
-$$
-\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
-$$
-Using the least squares estimators to update $\bs\beta$ and $\phi$, we have 
-\begin{align*}
-& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
-\end{align*}
-To note, the variance-covariance matrix of the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
-$$
-  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
-$$
-
-Totally, the proposed EM-IWLS algorithm is summarized as follows:
-
-1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$
-
-2) Iterate over the E-step and M-step until convergence
-
-    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E(\tau^{-2}_{j})$, $E({\tau^\tp}^{-2}_{jk})$ with the estimates $\Theta^{(t-1)}$ from the previous iteration
-
-    M-step:
-
-    a) Based on the current value of $\beta$, calculate the pseudo-data $z_i^{(t)}$ and the pseudo-weights $w_i^{(t)}$
-    b) Update $\bs \beta^{(t)}$ by runing the augmented weighted least squared
-    c) If $\phi$ is present, update $\phi$
-
-Similar to EM-CD, we assess convergence by the criterion, $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $\epsilon$ is a small value (say $10^{-5}$).
-
 <!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.

### 2022-02-15 Manuscript/10-Supplementary.Rmd
Move EM-IWLS to supproting Information

@@
 ---
-title: "Supplementary File"
+title: "Supporting Infromation"
 output: pdf_document
@@
 
-## Supplementary 1: Marginal Distribution of $\gamma_j^\tp$
+\subsection{Supplementary Information 1: Marginal Distribution of $\gamma_j^\tp$}
 Given that $\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)$ where $\gamma_{j} | \theta_j \sim Bin(1, \theta_j)$, we can derive the the marginal distribution of $\gamma_{j}^\tp$ with the following manipulation.
@@
 \end{align*}
+
+\clearpage
+
+\subsection{Supplementary Information 1: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}
+
+Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
+The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
+In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior
+<!-- , with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary -->
+.
+
+<!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
+
+A double exponential prior, $\beta|S \sim DE(0, S)$ can be formulated as a hierarchical normal prior with unknown variance $\tau^2$ integrated out:
+\begin{align*}
+  \beta|\tau^2 &\sim N(0, \tau^2)\\
+  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
+\end{align*}
+For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
+The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
+\begin{equation}\label{eq:Q1_IWLS}
+Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
+\end{equation}
+Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
+To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
+$E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}).
+On the other hand, $\tau^{2}$, following gamma distributions, is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
+$E({\tau}^{-2}_{j})$ and $E({\tau^\tp}^{-2}_{jk})$ are calculated using the closed form equation
+\begin{align*}
+ E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
+\end{align*}
+where $S_j$ and $S_j^\tp$ are replaced by the expectation and $\beta$s are replaced with $\beta^{(t-1)}$. With simplification (up to constant additive terms), we have 
+\begin{equation}\label{eq:EQ1_IWLS}
+E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
+\end{equation}
+<!-- $E(Q_1)$ can be seen as a $l_2$ penalized likelihood function where the regularization parameter $\lambda = 2E({\tau}^{-2})$. Equivalently, we have  -->
+$2E({\tau}^{-2})\beta^2$ can be seen as the kernel of a normal density with mean 0 and variance $E(\tau^{2})$, and we can formulate the coefficients $\bs \beta$ as a multivariate normal distribution with means $\bs 0$ and variance covariance matrix $\bs \Sigma_{\tau^2}$, where $\bs \Sigma_{\tau^2}$ is a diagonal matrix with $E(\tau^2)$s on the diagonal,
+$$
+\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
+$$
+
+Meanwhile, following the classical IWLS, we can approximate the generalized model likelihood at each iteration with a weighted normal likelihood:
+$$
+f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
+$$
+where the ‘normal response’ $z_i$ and ‘weight’ $w_i$ are called the pseudo-response and pseudo-weight respectively. The pseudo-response and the pseudo-weight are calculated by
+$$
+\begin{aligned}
+z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
+\end{aligned}
+$$
+where $\hat\eta_i = (\bs X {\hat{\bs\beta}})_i$, $L^{'}(y_i|\hat\eta_i, \hat \phi)$ and $L^{''}(y_i|\hat\eta_i, \hat \phi)$ are the first and second derivative of the log density, $\log f(\textbf{y}_i|\bs \beta, \phi)$ with respect to $\eta_i$.
+
+With $\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)$ and $\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})$, we can augment the two multivariate normal distributions and update the estimates for $\bs \beta$ and $\phi$ via least squares in each iteration of the EM algorithm.
+We create the augmented response, augmented data, and augmented variance-covariance matrix following \begin{align*}
+& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
+  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
+  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
+\end{align*}
+ such that 
+$$
+\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
+$$
+Using the least squares estimators to update $\bs\beta$ and $\phi$, we have 
+\begin{align*}
+& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
+\end{align*}
+To note, the variance-covariance matrix of the coefficient estimates variance-covariance matrix can be derived in the EM-IWLS algorithm and in turn can be used for statistical inferences,
+$$
+  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
+$$
+
+Totally, the proposed EM-IWLS algorithm is summarized as follows:
+
+1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ for $\bs \beta$ and $\bs \theta$. For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$
+
+2) Iterate over the E-step and M-step until convergence
+
+    E-step: calculate $E(\gamma_{j})$, $E(\gamma^\tp_{j})$ and $E(\tau^{-2}_{j})$, $E({\tau^\tp}^{-2}_{jk})$ with the estimates $\Theta^{(t-1)}$ from the previous iteration
+
+    M-step:
+
+    a) Based on the current value of $\beta$, calculate the pseudo-data $z_i^{(t)}$ and the pseudo-weights $w_i^{(t)}$
+    b) Update $\bs \beta^{(t)}$ by runing the augmented weighted least squared
+    c) If $\phi$ is present, update $\phi$
+
+Similar to EM-CD, we assess convergence by the criterion, $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $\epsilon$ is a small value (say $10^{-5}$).

### 2022-02-15 Manuscript/00-main.Rmd
Update Section Headers to Latex

--- 2022-01-10 Manuscript/00-main.Rmd
@@
 
-\newpage
+\clearpage
 
@@
 ```
+\clearpage
 
@@
 
-\newpage
+\clearpage
 <!-- \begin{landscape} -->
@@
 <!-- \end{landscape} -->
-\newpage
+\clearpage
 

### 2022-02-15 Manuscript/03-Simulation.Rmd
Update Section Headers to Latex

@@
-# Simulation Study
+\section{Simulation Study}
 \label{sec:sim}
@@
 
-## Monte Carlo Simulation Study
+\subsection{Monte Carlo Simulation Study}
 We follow the data generating process described in Bai [@Bai2021].
@@
 ```
-
+\subsection{Simulation Results}
+\subsubsection{Prediction Performance}
 The predictive performances have a consistent pattern across the two distributions of outcomes.
@@
 However, the BHAM-CD has extreme computational advantage over SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
+
+\subsubsection{Variable Selection Performance}
 

### 2022-02-15 Manuscript/04-Real_Data.Rmd
Update Section Headers to Latex

@@
-# Metabolomics Data Analysis
+\section{Metabolomics Data Analysis}
 \label{sec:real_data}
@@
 
-## Emory Cardiovascular Biobank
+\subsection{Emory Cardiovascular Biobank}
 \label{sec:ECB}
@@
 
-## Weight Loss Maintenance Cohort 
+\subsection{Weight Loss Maintenance Cohort}
 We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].

### 2022-02-15 Manuscript/SS_GAM.tex
Update Section Headers to Latex

@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
+\section{Introduction}
 \label{sec:intro}
@@
 
-\hypertarget{bayesian-hierarchical-additive-models-bhams}{%
-\section{Bayesian Hierarchical Additive Models
-(BHAMs)}\label{bayesian-hierarchical-additive-models-bhams}}
-
+\section{Bayesian Hierarchical Additive Models (BHAMs)}
 \label{sec:BHAM}
@@
 
-\hypertarget{spike-and-slab-spline-priors}{%
-\subsection{Spike-and-Slab Spline
-Priors}\label{spike-and-slab-spline-priors}}
+\subsection{Spike-and-Slab Spline Priors}
 
@@
 
-\hypertarget{two-part-spike-and-slab-lasso-prior}{%
-\subsubsection{Two-part Spike-and-Slab Lasso
-Prior}\label{two-part-spike-and-slab-lasso-prior}}
+\subsubsection{Two-part Spike-and-Slab Lasso Prior}
 
@@
 
-\hypertarget{other-priors}{%
-\subsubsection{Other Priors}\label{other-priors}}
+\subsubsection{Other Priors}
 
@@
 
-\hypertarget{algorithms-for-fitting-bhams}{%
-\subsection{Algorithms for Fitting
-BHAMs}\label{algorithms-for-fitting-bhams}}
+\subsection{Algorithms for Fitting BHAMs}
 
@@
 
-\hypertarget{em-algorithms}{%
-\subsubsection{EM algorithms}\label{em-algorithms}}
+\subsubsection{EM algorithms}
 
@@
 
-\hypertarget{em-coordinate-descent}{%
-\subsubsection{EM-Coordinate Descent}\label{em-coordinate-descent}}
-
+\subsubsection{EM-Coordinate Descent}
 \label{sec:EMCD}
@@
 
-\hypertarget{em-iwls}{%
-\subsubsection{EM-IWLS}\label{em-iwls}}
-
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative
-EM-based algorithm where the iterative weighted least squares algorithm
-is used to find the estimate of \(\boldsymbol{\beta}, \phi\) that
-maximizes \(E(Q_1)\). The iterative weighted least squares algorithm was
-originally proposed to fit the classical generalized linear models, and
-generalized to fit some Bayesian hierarchical models.\citep{Gelman2013}
-Yi and Ma \citep{Yi2012} formulated Student's t-distribution and double
-exponential distribution as hierarchical normal distributions such that
-generalized linear models with shrinkage priors can be easily fitted
-using IWLS in combination with EM algorithm. In this work, we adapt the
-EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior .
-
-A double exponential prior, \(\beta|S \sim DE(0, S)\) can be formulated
-as a hierarchical normal prior with unknown variance \(\tau^2\)
-integrated out: \begin{align*}
-  \beta|\tau^2 &\sim N(0, \tau^2)\\
-  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
-\end{align*} For the mixture double exponential priors, we can define
-the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following
-Equation (\ref{eq:ssl}). The change in the prior formulation in turn
-leads to the change in the log posterior density function, as \(Q_1\)
-needs to account for the hyperprior of \(\tau^2\):
-\begin{equation}\label{eq:Q1_IWLS}
-Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
-\end{equation} Since \(\boldsymbol{\tau}^2\) are not of our primary
-interest, we treat them as the ``missing'' data in addition to the
-latent indicators \(\boldsymbol{\gamma}\), and hence construct the
-expectation
-\(E_{\boldsymbol{\gamma}, \boldsymbol{\tau}^2|\Theta^{(t-1)}}(Q_1)\) in
-the E-step. To note, unlike the same latent indicator \(\gamma^*_j\)
-which is shared by the coefficients of the non-linear terms
-\(\beta^*_{jk}\) for \(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is
-coefficient specific for \(\beta^*_{jk}\).
-\(E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
-needs to be calculated to formulate \(E(Q_1)\). As neither
-\(E({S_j}^{-1}|\beta_j, s_0, s_1)\) nor
-\(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
-\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}). On
-the other hand, \(\tau^{2}\), following gamma distributions, is a
-conjugate prior for the normal variance, and the conditional posterior
-density of \(\tau^{-2}\) is an inverse Gaussian distribution.
-\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
-using the closed form equation \begin{align*}
- E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
-\end{align*} where \(S_j\) and \(S_j^*\) are replaced by the expectation
-and \(\beta\)s are replaced with \(\beta^{(t-1)}\). With simplification
-(up to constant additive terms), we have
-\begin{equation}\label{eq:EQ1_IWLS}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^*}^{-2})}{\beta_{jk}^*}^2\right].
-\end{equation} \(2E({\tau}^{-2})\beta^2\) can be seen as the kernel of a
-normal density with mean 0 and variance \(E(\tau^{2})\), and we can
-formulate the coefficients \(\boldsymbol{\beta}\) as a multivariate
-normal distribution with means \(\boldsymbol{0}\) and variance
-covariance matrix \(\boldsymbol{\Sigma}_{\tau^2}\), where
-\(\boldsymbol{\Sigma}_{\tau^2}\) is a diagonal matrix with
-\(E(\tau^2)\)s on the diagonal, \[
-\boldsymbol{\beta }\sim \text{MVN}(0, \boldsymbol{\Sigma}_{\tau^2}).
-\]
-
-Meanwhile, following the classical IWLS, we can approximate the
-generalized model likelihood at each iteration with a weighted normal
-likelihood: \[
-f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx \text{MVN}(\textbf{z}|\boldsymbol{X} \boldsymbol{\beta}, \phi\boldsymbol{\Sigma })
-\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
-the pseudo-response and pseudo-weight respectively. The pseudo-response
-and the pseudo-weight are calculated by \[
-\begin{aligned}
-z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
-\end{aligned}
-\] where \(\hat\eta_i = (\boldsymbol{X} {\hat{\boldsymbol{\beta}}})_i\),
-\(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
-\(L^{''}(y_i|\hat\eta_i, \hat \phi)\) are the first and second
-derivative of the log density,
-\(\log f(\textbf{y}_i|\boldsymbol{\beta}, \phi)\) with respect to
-\(\eta_i\).
-
-With
-\(\boldsymbol{z}\sim \text{MVN}(\boldsymbol{X} \boldsymbol{\beta}, \phi \boldsymbol{\Sigma})\)
-and
-\(\boldsymbol{\beta }\sim \text{MVN}(0, \phi \boldsymbol{\Sigma}_{\tau^2})\),
-we can augment the two multivariate normal distributions and update the
-estimates for \(\boldsymbol{\beta}\) and \(\phi\) via least squares in
-each iteration of the EM algorithm. We create the augmented response,
-augmented data, and augmented variance-covariance matrix following
-\begin{align*}
-& \boldsymbol{z}_* = \begin{bmatrix} \boldsymbol{z}\\ \boldsymbol{0}\end{bmatrix} &&
-  \boldsymbol{X}_* = \begin{bmatrix} \boldsymbol{X} \\ \boldsymbol{I} \end{bmatrix} &&
-  \boldsymbol{\Sigma}_* = \begin{bmatrix} \boldsymbol{\Sigma }& \boldsymbol{0}  \\ \boldsymbol{0} & \boldsymbol{\Sigma}_{\tau^2}/\phi \end{bmatrix}, &
-\end{align*} such that \[
-\boldsymbol{z}_* \sim \text{MVN}(\boldsymbol{X}_* \boldsymbol{\beta }, \phi \Sigma_*).
-\] Using the least squares estimators to update \(\boldsymbol{\beta}\)
-and \(\phi\), we have \begin{align*}
-& \boldsymbol{\beta}^{(t)} = (\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{z}_* && \phi^{(t)} = \frac{1}{n}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)})^T\boldsymbol{\Sigma}^{-1}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)}).&
-\end{align*} To note, the variance-covariance matrix of the coefficient
-estimates variance-covariance matrix can be derived in the EM-IWLS
-algorithm and in turn can be used for statistical inferences, \[
-  \text{Var}(\boldsymbol{\beta}^{(t)}) = (\boldsymbol{X}_*^T\boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\phi^{(t)}.
-\]
-
-Totally, the proposed EM-IWLS algorithm is summarized as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
-  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
-  \(\boldsymbol{\theta}\). For example, we can initialize
-  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
-  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
-\item
-  Iterate over the E-step and M-step until convergence
-
-  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
-  \(E(\tau^{-2}_{j})\), \(E({\tau^*}^{-2}_{jk})\) with the estimates
-  \(\Theta^{(t-1)}\) from the previous iteration
-
-  M-step:
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii})}
-  \tightlist
-  \item
-    Based on the current value of \(\beta\), calculate the pseudo-data
-    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
-  \item
-    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
-    least squared
-  \item
-    If \(\phi\) is present, update \(\phi\)
-  \end{enumerate}
-\end{enumerate}
-
-Similar to EM-CD, we assess convergence by the criterion,
-\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where \(\epsilon\) is
-a small value (say \(10^{-5}\)).
-
-\hypertarget{selecting-optimal-scale-values}{%
-\subsection{Selecting Optimal Scale
-Values}\label{selecting-optimal-scale-values}}
-
-\label{sec:tune} Our proposed models, BHAM, require two preset scale
-parameters (\(s_0\), \(s_1\)). Hence, we need to find the optimal values
-for the scale parameters such that the model reaches its best prediction
-performance regarding a criteria of preference. This would be achieved
-by constructing a two dimensional grid, consists of different pairs of
+\subsection{Selecting Optimal Scale Values}
+\label{sec:tune}
+
+Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
+\(s_1\)). Hence, we need to find the optimal values for the scale
+parameters such that the model reaches its best prediction performance
+regarding a criteria of preference. This would be achieved by
+constructing a two dimensional grid, consists of different pairs of
 (\(s_0\), \(s_1\)) value. However, previous research suggested the value
@@
 
-\hypertarget{simulation-study}{%
-\section{Simulation Study}\label{simulation-study}}
-
+\section{Simulation Study}
 \label{sec:sim}
@@
 
-\hypertarget{monte-carlo-simulation-study}{%
-\subsection{Monte Carlo Simulation
-Study}\label{monte-carlo-simulation-study}}
+\subsection{Monte Carlo Simulation Study}
 
@@
 
+\subsection{Simulation Results}
+\subsubsection{Prediction Performance}
+
 The predictive performances have a consistent pattern across the two
@@
 
+\subsubsection{Variable Selection Performance}
+
 ~
 
-\hypertarget{metabolomics-data-analysis}{%
-\section{Metabolomics Data Analysis}\label{metabolomics-data-analysis}}
-
+\section{Metabolomics Data Analysis}
 \label{sec:real_data}
@@
 
-\hypertarget{emory-cardiovascular-biobank}{%
-\subsection{Emory Cardiovascular
-Biobank}\label{emory-cardiovascular-biobank}}
-
-\label{sec:ECB} We use the proposed models BHAM to analyze a metabolic
-dataset from a recently published research \citep{Mehta2020} studying
-plasma metabolomic profile on the three-year all-cause mortality among
-patients undergoing cardiac catheterization. The dataset is publicly
-available via \textit{Dryad} \citep{Mehta2020_data}. It contains in
-total of 776 subjects from two cohorts. As there is a large number of
-non-overlapping features among the two cohorts, we use the cohort with
-larger sample size (N=454). There are initially 6796 features in the
-dataset, which is too large to be practically meaningful to analyze.
-Hence, we perform a univariate screening procedure on the features, via
-GAM implemented in \texttt{mgcv}, and choose the the top 200 features
-with smallest p-values. We use 5-knot spline additive models for binary
-outcome using two different models, the proposed BHAM and the SB-GAM.
-10-Fold CV are used to choose the optimal tuning parameters of each
-framework with respect to the default selection criterion implemented in
-the software. Out-of-bag samples are used for prediction performance
-evaluation, where deviance, AUC, Brier score defined as
+\subsection{Emory Cardiovascular Biobank}
+\label{sec:ECB}
+
+We use the proposed models BHAM to analyze a metabolic dataset from a
+recently published research \citep{Mehta2020} studying plasma
+metabolomic profile on the three-year all-cause mortality among patients
+undergoing cardiac catheterization. The dataset is publicly available
+via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
+subjects from two cohorts. As there is a large number of non-overlapping
+features among the two cohorts, we use the cohort with larger sample
+size (N=454). There are initially 6796 features in the dataset, which is
+too large to be practically meaningful to analyze. Hence, we perform a
+univariate screening procedure on the features, via GAM implemented in
+\texttt{mgcv}, and choose the the top 200 features with smallest
+p-values. We use 5-knot spline additive models for binary outcome using
+two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are
+used to choose the optimal tuning parameters of each framework with
+respect to the default selection criterion implemented in the software.
+Out-of-bag samples are used for prediction performance evaluation, where
+deviance, AUC, Brier score defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
@@
 
-\hypertarget{weight-loss-maintenance-cohort}{%
-\subsection{Weight Loss Maintenance
-Cohort}\label{weight-loss-maintenance-cohort}}
+\subsection{Weight Loss Maintenance Cohort}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\label{sec:concl} In the paper, we described a novel generalized
-additive model using Bayesian hierarchical priors, particularly the
-proposed spike-and-slab spline prior for bi-level functional selection.
-Meanwhile, we introduced two optimization based algorithms for model
-fitting. The algorithms can be easily scale up to address
-high-dimensional data analysis in a computational efficient manner. Via
-simulations, we demonstrated that the proposed model provides as good,
-if not better, prediction performance compared to some state-of-the-art
-non-linear modeling devices.
+\section{Discussion}
+\label{sec:concl}
+
+In the paper, we described a novel generalized additive model using
+Bayesian hierarchical priors, particularly the proposed spike-and-slab
+spline prior for bi-level functional selection. Meanwhile, we introduced
+two optimization based algorithms for model fitting. The algorithms can
+be easily scale up to address high-dimensional data analysis in a
+computational efficient manner. Via simulations, we demonstrated that
+the proposed model provides as good, if not better, prediction
+performance compared to some state-of-the-art non-linear modeling
+devices.
 
@@
 \end{table}
+\clearpage
 

### 2022-02-15 Manuscript/R&R/00-response_letter.Rmd
Add simulaiton Settings

@@
     \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects inclusion indicator such that the strong hierarchy for bi-level selection is enforced.
-    \item We moved the EM-IWLS fitting algorithm to the supporting information for two reasons.
-First of all, we think moving the EM-IWLS to the supporting information can help the audience focus on the model and EM-Coordinate Descent algorithm, avoiding confusing the audience.
+    \item We moved the EM-IWLS fitting algorithm to the supporting information for three reasons.
+First of all, moving the EM-IWLS to the supporting information can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction.
 Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g.
 constructing hypothesis testings and confidence band, requires a more dedicated discussion.
-We are currently preparing another manuscript discussing those.
-Lastly, by moving the EM-IWLS to the supporting information would not hurt the integrity of the current manuscript, that emphasize the scalability and bi-level selection.
+We are preparing another manuscript discussing those.
+Lastly, relocating the EM-IWLS section would not hurt the integrity of the current manuscript that emphasize the scalability and bi-level selection.
     \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear lasso model and spikeSlabGAM model among the models to be compared. 

### 2022-02-15 Manuscript/03-Simulation.Rmd
Add simulaiton Settings

@@
 
-In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021].
+In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021]\bg{}, linear lasso model and spikeSlabGAM [TODO: add citation here]}.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
@@
 We generate 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
+\bg{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
+These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
 <!-- The dimensional of the smoothing functions are controlled as k=10. Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.  -->
@@
 To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing dataset, are averaged <!--and deviance are calculated for Poisson model,--> across 50 simulations.
-Computation time for model selection, final model fitting and prediction are recorded for all simulations.
+\bg{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
+For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
 
@@
 ```
-\subsection{Simulation Results}
-\subsubsection{Prediction Performance}
+\bg{\subsection{Simulation Results}}
+\bg{\subsubsection{Prediction Performance}}
 The predictive performances have a consistent pattern across the two distributions of outcomes.
@@
 
-\subsubsection{Variable Selection Performance}
+\bg{\subsubsection{Variable Selection Performance}}
 
@@
 <!-- where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively. -->
-
- \
 

### 2022-02-15 Manuscript/SS_GAM.tex
Add simulaiton Settings

@@
 generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM \citep{Bai2021}. COSSO is one of the earliest smoothing spline
-models that consider sparsity-smoothness penalty. Adaptive COSSO
-improved upon COSSO by using adaptive weight for penalties such that the
-penalty of each functional component are different for extra
-flexibility. Generalized additive models with automatic smoothing,
-hereafter \textit{mgcv}, is one of the most popular models for nonlinear
-effect interpolation and prediction. SB-GAM is the first spike-and-slab
-lasso GAM. We implement COSSO and adaptive COSSO with R package
-\texttt{cosso 2.1-1}, generalized additive models with automatic
-smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package
-\texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
+SB-GAM \citep{Bai2021}\textcolor{red}{}, linear lasso model and
+spikeSlabGAM {[}TODO: add citation here{]}\}. COSSO is one of the
+earliest smoothing spline models that consider sparsity-smoothness
+penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for
+penalties such that the penalty of each functional component are
+different for extra flexibility. Generalized additive models with
+automatic smoothing, hereafter \textit{mgcv}, is one of the most popular
+models for nonlinear effect interpolation and prediction. SB-GAM is the
+first spike-and-slab lasso GAM. We implement COSSO and adaptive COSSO
+with R package \texttt{cosso 2.1-1}, generalized additive models with
+automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R
+package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
 flexibility to define smoothing functions, and hence use the default
@@
 measures.
+\textcolor{red}{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
+These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
 
@@
 \(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing dataset, are averaged across 50 simulations. Computation
-time for model selection, final model fitting and prediction are
-recorded for all simulations.
+on the testing dataset, are averaged across 50 simulations.
+\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
+For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
 
@@
 
-\subsection{Simulation Results}
-\subsubsection{Prediction Performance}
+\textcolor{red}{\subsection{Simulation Results}}
+\textcolor{red}{\subsubsection{Prediction Performance}}
 
@@
 
-\subsubsection{Variable Selection Performance}
-
-~
+\textcolor{red}{\subsubsection{Variable Selection Performance}}
 

### 2022-02-18 R/Sim_Res/sim_var_select.R
create formated empty df when bamlasso_raw contain nothing

@@ -48,9 +48,15 @@ sim_comp_select_raw <- function(success_rate,
                         Component == "sm" ~ "Nonlinear",
                         TRUE ~ NA_character_),
                       ssGAM = prob>=0.5)
-          bamlasso_raw <- comp_res_raw$bamlasso$`Non-parametric` %>% 
-            pivot_longer(cols = ends_with("Linear"), names_to = "bamlasso") %>% 
+          # browser()
+          if(nrow(comp_res_raw$bamlasso$`Non-parametric`)!=0)
+            bamlasso_raw <- comp_res_raw$bamlasso$`Non-parametric` %>% 
+            pivot_longer(cols = ends_with("linear"), names_to = "bamlasso") %>% 
             transmute(Variable, Component = bamlasso, bamlasso = value )
+          else
+            bamlasso_raw <- data.frame(Variable = character(),
+                                       Component = character(),
+                                       bamlasso = logical())
           
           ssGAM_raw %>% left_join(bamlasso_raw) %>% 
             mutate(bamlasso = replace_na(bamlasso, FALSE)) %>% 

### 2022-02-18 Manuscript/03-Simulation.Rmd
Add bi-level selection section

@@
 
+\bg{\subsubsection{Bi-level Selection}
+Among the methods of comparison, there are two methods is possible for bi-level selection, i.e.
+the proposed BHAM and spikeSlabGAM.
+Hence, in this section, we further investigate the bi-level selection of the proposed models.
+}
 <!-- %are calculated to assess out-of-sample estimation accuracy, -->

### 2022-02-18 R/functions.R
Update screening of ECB data analysis from GAM based to variance based (#27)

@@ -5,29 +5,34 @@
 #'
 #' @return
 #'
-gam_screen <- function(dat, nrow = 200){
+var_screen <- function(dat, nrow = 200){
   dat %>% 
     select(starts_with("mz")) %>% 
-    names() %>% 
-    map_dfr(#feature_name[1:2],
-      .f = function(name, .dat){
-        # name <- feature_name[1]
-        y <- .dat %>% pull(death3yr)
-        x <- .dat %>% pull({{name}})
-        mgcv::gam(y ~ s(x, bs = "cr", k = 10), family = binomial(),
-                  data = data.frame(x, y)) %>% 
-          tidy() %>% 
-          filter(term == "s(x)") %>% 
-          select(p.value) %>% 
-          mutate(var = name,
-                 p.value)
-        
-      },
-      .dat = dat) %>% data.frame(
-        .,
-        p.adj = p.adjust(.$p.value, "fdr")
-      ) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
-    head(nrow)
+    apply(,MARGIN = 2, FUN = var) %>% 
+    sort(decreasing = TRUE) %>% 
+    head(nrow) %>% 
+    names
+  # browser()
+    # names() %>% 
+    # map_dfr(#feature_name[1:2],
+    #   .f = function(name, .dat){
+    #     # name <- feature_name[1]
+    #     y <- .dat %>% pull(death3yr)
+    #     x <- .dat %>% pull({{name}})
+    #     mgcv::gam(y ~ s(x, bs = "cr", k = 10), family = binomial(),
+    #               data = data.frame(x, y)) %>% 
+    #       tidy() %>% 
+    #       filter(term == "s(x)") %>% 
+    #       select(p.value) %>% 
+    #       mutate(var = name,
+    #              p.value)
+    #     
+    #   },
+    #   .dat = dat) %>% data.frame(
+    #     .,
+    #     p.adj = p.adjust(.$p.value, "fdr")
+    #   ) %>% arrange(p.adj) %>% #filter(p.adj<0.2)
+    # head(nrow)
 }
 
 

### 2022-02-18 _targets.R

@@ -53,21 +53,20 @@ tar_plan(
              readr::read_csv(ECB_train_path)),
   
   # ** GAM Screening ####
-  tar_target(ECB_gam_screen,
-             gam_screen(ECB_train_dat)),
+  tar_target(ECB_var_screen,
+             var_screen(ECB_train_dat)),
   
   tar_target(ECB_cov,
              ECB_train_dat %>% 
-               select(ECB_gam_screen %>% pull(var)) %>% 
+               select(all_of(ECB_var_screen)) %>% 
                data.matrix),
   
   tar_target(ECB_outcome,
              ECB_train_dat %>% pull(death3yr)),
   
   # ** BHAM --------------------------------------------------------------- 
-  ECB_mz_names = ECB_cov %>% colnames,
   ECB_sm_df = data.frame(
-    Var = ECB_mz_names,
+    Var = ECB_cov %>% colnames,
     Func = "s",
     Args ="bs='cr', k=5"
   ),
@@ -82,7 +81,7 @@ tar_plan(
   
   ECB_bamlasso_fnl = bamlasso(ECB_dsn_mat, ECB_outcome, family = "binomial",
                           group = make_group(names(ECB_dsn_mat)),
-                          ss = c(0.095, 0.5)),
+                          ss = c(0.025, 0.5)),
   
   ECB_bamlasso_insample_msr = measure.bh(ECB_bamlasso_fnl),
   ECB_bamlasso_var = bamlasso_var_selection(ECB_bamlasso_fnl),
@@ -102,8 +101,7 @@ tar_plan(
   
   ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 24, a = 1, b = 1),
 
-    tar_target(ECB_SBGAM_insample_msr,
-             measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial")),
+  ECB_SBGAM_insample_msr = measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial"),
   
   ECB_SBGAM_var = (ECB_cov %>% colnames())[ECB_SBGAM_fnl$classifications!=0],
   

### 2022-02-18 Manuscript/03-Simulation.Rmd
Initiate Var Select Section

@@
 
-In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021]\bg{}, linear lasso model and spikeSlabGAM [TODO: add citation here]}.
+In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021]\bg{, linear lasso model and spikeSlabGAM [TODO: add citation here].[TODO: explain about linear lasso model and spikeSlabGAM model]}.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
@@
 for Gaussian and binomial outcomes.
-<!-- and shrinked effects -->
-<!-- $$ -->
-<!-- \mathbb{E}(Y) = g^{-1}(\sin(2\pi x_1) - 2\cos(2\pi x_2 -0.5) + 0.4(x_3-0.5) - 2(x_4^2 -0.3)) -->
-<!-- $$ -->
-<!-- for  poisson outcomes.  -->
 Gaussian outcomes requires specification of dispersion, where we set the dispersion parameter to be 1.
@@
 These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
-<!-- The dimensional of the smoothing functions are controlled as k=10. Hence, the corresponding number of coefficient parameters to be estimated is 37, 82, 451 and 1801.  -->
-<!-- We also run the simulation where the number of bases are increased to 25 deliberately. -->
 
@@
 For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-```{r, fig.height=8, eval = FALSE}
-targets::tar_load(binom_plot)
-binom_plot
-```
-
-```{r, fig.height=8, eval = FALSE}
-targets::tar_load(gaussian_plot)
-gaussian_plot
-```
-
-```{r, fig.height=8, eval = FALSE}
-targets::tar_load(poisson_plot)
-poisson_plot
-```
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.
+MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
+$$
+where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.
+$$
+}
 \bg{\subsection{Simulation Results}}
@@
 
-\bg{\subsubsection{Variable Selection Performance}}
+\bg{
+
+}
+
+\bg{\subsubsection{Variable Selection Performance}
+\begin{itemize}
+\item If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
+The pattern is consitent regardless of the type of outcomes.
+To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
+Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.
+\item If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
+While the recall is lower than SBGAM.
+This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.
+\item We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while. On the contrary the SBGAM is not so well.
+\end{itemize}
+
+}
 
@@
 }
-<!-- %are calculated to assess out-of-sample estimation accuracy, -->
-
-<!-- We also calculated the area under the curve (AUC) of the receiver operating characteristic curve for $y_{new}$.
-We will also visually display the estimated smooth function with the smooth function of truth to compare if the models can infer the smooth function correctly, despite the prediction accuracy.
--->
-
-<!-- The accuracy of variable selection is measured using Matthews correlation coefficient (MCC), -->
-<!-- $$ -->
-<!-- MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}, -->
-<!-- $$ -->
-<!-- where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively. -->
-
-<!-- ### Simulation Results -->
-
-<!-- Overall, the proposed priors outperforms the state-of-the-art GAM model `mgcv` when the dimensionality is low, median, and high.
-The proposed priors predicts better than `mgcv` and infers better of the underlying smooth function.
-Most importantly, `mgcv` is infeasible in the high dimensional setting, i.e.
-when $p=50$ where the number of parameters are greater than the sample size.
--->
-
-<!-- #### Prediction Performace -->
-
-<!-- The prediction performance is tabulated in the table below for AUC and Supplementary tables for MSE and misclassification rate.
-In summary, the proposed Bayesian models outperforms the benchmark model `mgcv` regardless the choice of prior or fitting algorithm.
-The models with IRLS fitting algorithm give slightly better performance than the the model with cyclic coordinate descent algorithm.
-The advantage diminishes in the high dimensional setting (p=50).
-This is possible because the SS normal spline prior provides a more smooth solution, which would work well in a low to median dimension setting.
-However, in the high dimensional setting, there are not enough sparsity in the model, i.e.
-SS normal prior will provides a more complex model than necessary.
-This is confirmed with the visualization in the following section.
--->
-
-```{r tab.id="sim_GAM_AUC", eval=F}
-glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")
-
-glm_spline_pred_dat$auc %>% 
-        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
-        flextable() %>% 
-  set_caption(caption = "Averaged AUC of out-of-bag samples",
-              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_AUC", pre_label = "Table "))
-```
-
-```{r  eval=F}
-glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")
-
-glm_spline_pred_dat$misclass %>% 
-        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
-        flextable() %>% 
-  set_caption(caption = "Averaged misclassification rate of out-of-bag samples",
-              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_Misclass", pre_label = "Table "))
-```
-
-```{r  eval=F}
-
-glm_spline_pred_dat <-readr::read_rds("Simulation_Result/predict_res.rds")
-
-glm_spline_pred_dat$mse %>% 
-        select(-c(n_train, n_test, "bglm_t", "bglm_de", "blasso")) %>% 
-        flextable() %>% 
-  set_caption(caption = "Averaged mean squared area of out-of-bag samples",
-              autonum = run_autonum(seq_id = "tab", bkm="sim_GAM_MSE", pre_label = "Table "))
-```
 

### 2022-02-18 Manuscript/SS_GAM.tex
Add bi-level selection section
Initiate Var Select Section

@@
 generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM \citep{Bai2021}\textcolor{red}{}, linear lasso model and
-spikeSlabGAM {[}TODO: add citation here{]}\}. COSSO is one of the
-earliest smoothing spline models that consider sparsity-smoothness
-penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for
-penalties such that the penalty of each functional component are
-different for extra flexibility. Generalized additive models with
-automatic smoothing, hereafter \textit{mgcv}, is one of the most popular
-models for nonlinear effect interpolation and prediction. SB-GAM is the
-first spike-and-slab lasso GAM. We implement COSSO and adaptive COSSO
-with R package \texttt{cosso 2.1-1}, generalized additive models with
-automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R
-package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence use the default
-choices. Both mgcv and proposed models allow customized smoothing
-functions and we choose the cubic regression spline. We controll the
-dimensionality of each smoothing function, 10 bases, for all different
-choices of smoothing functions. We use 5-fold CV with the default
-selection criteria to select the final model for COSSO models, SB-GAM
-and the proposed models. 20 default candidates of tuning parameters
-(\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for SB-GAM and
-the proposed models which allow user-specification of tuning candidates.
-All computation was conducted on a high-performance 64-bit Linux
-platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
+SB-GAM
+\citep{Bai2021}\textcolor{red}{, linear lasso model and spikeSlabGAM [TODO: add citation here]}.
+COSSO is one of the earliest smoothing spline models that consider
+sparsity-smoothness penalty. Adaptive COSSO improved upon COSSO by using
+adaptive weight for penalties such that the penalty of each functional
+component are different for extra flexibility. Generalized additive
+models with automatic smoothing, hereafter \textit{mgcv}, is one of the
+most popular models for nonlinear effect interpolation and prediction.
+SB-GAM is the first spike-and-slab lasso GAM. We implement COSSO and
+adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive
+models with automatic smoothing with R package \texttt{mgcv 1.8-31},
+SB-GAM with R package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do
+not provide flexibility to define smoothing functions, and hence use the
+default choices. Both mgcv and proposed models allow customized
+smoothing functions and we choose the cubic regression spline. We
+controll the dimensionality of each smoothing function, 10 bases, for
+all different choices of smoothing functions. We use 5-fold CV with the
+default selection criteria to select the final model for COSSO models,
+SB-GAM and the proposed models. 20 default candidates of tuning
+parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for
+SB-GAM and the proposed models which allow user-specification of tuning
+candidates. All computation was conducted on a high-performance 64-bit
+Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
 processors and 24G of RAM per core and R 3.6.2 \citep{R}.
@@
 \textcolor{red}{\subsubsection{Variable Selection Performance}}
+
+\textcolor{red}{\subsubsection{Bi-level Selection}
+Among the methods of comparison, there are two methods is possible for bi-level selection, i.e.
+the proposed BHAM and spikeSlabGAM.
+Hence, in this section, we further investigate the bi-level selection of the proposed models.
+}
 
@@
 
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{490.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{100.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.5}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{104.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2,659.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{20.9}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{2,679.9}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{365.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{6.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{371.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
 
@@
   \hline
-BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
+BHAM-CD.1 & 510.99 & 0.61 & 0.19 & 0.24 \\ 
+  BHAM-CD.2 & 510.99 & 0.61 & 0.19 & 0.24 \\ 
+  SB-GAM & 636.56 & 0.56 & 0.22 & 0.30 \\ 
    \hline
@@
   \hline
-BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
+BHAM-CD & 668.01 & 0.07 & 0.93 & 0.76 \\ 
   SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 

### 2022-02-18 Manuscript/R&R/03-reviewer_4.Rmd
Add citations (#25)

@@
 
-_Comments to the authors_
+\em Comments to the authors\em
 
@@
 
-_The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs).
+\em The authors introduce the Bayesian hierarchical generalized additive model (BHAM) for high-dimensional data analysis with generalized additive models (GAMs).
 The BHAM framework allows for variable selection and estimation in GAMs.
@@
 The EM-IWLS algorithm allows for inference, which is its main advantage over EM-CD.
-The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty._
+The algorithms appear to be faster than GAMs based on group regularization since fitting a model with a lasso penalty is typically faster than fitting one with a group lasso penalty.
 
-_The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches.
+The paper is coherent and introduces a potentially promising methodology that addresses some of the shortcomings of existing Bayesian GAM approaches.
 However, in order to be suitable for publication in Statistics in Medicine, I have a few concerns that I think the authors should address.
-Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [1, 2] or the spikeslabGAM of [3]).
-I detail my main concerns below._
+Namely, the authors need to do a better job illustrating the benefits of the BHAM approach over existing methods (such as the sparse Bayesian GAM method of [@Bai2020; @Bai2021] or the spikeslabGAM of [@Scheipl2012]).
+I detail my main concerns below.\em
 
 ## _2. Main Comments and Questions_
+
 \em 1.
-The incorporation of the smoothing penalty matrix Sj in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space.
+The incorporation of the smoothing penalty matrix $S_j$ in BHAM allows for separation of the linear space of a smoothing function from the nonlinear space.
 This is one of the touted benefits of BHAM - this separation allows for bilevel selection rather than "all-in-all-out" selection, so that the practitioner can distinguish linear effects from nonlinear ones.
@@
 
-For example, in bilevel selection with the sparse group lasso [4], Simon et al. (2012)
+For example, in bilevel selection with the sparse group lasso [@simon2013], Simon et al. (2012)
 state in their algorithm that they first check whether or not the group is selected,
@@
 
-\em 3. The authors claim that BHAM is computationally beneficial because it does not rely
+_3. The authors claim that BHAM is computationally beneficial because it does not rely
 on MCMC. The authors illustrate the computational benefit of the double exponential
-prior over the group spike-and-slab prior of SBGAM [1, 2], but in my view, they
-should also compare it to the method spikeslabGAM of [3]. There is an R package
+prior over the group spike-and-slab prior of [@Bai2020; @Bai2021], but in my view, they
+should also compare it to the method spikeslabGAM of @Scheipl2012. There is an R package
 spikeslabGAM to implement this method. This method uses a Gaussian spike-and-
@@
 IWLS algorithms compared to MCMC, the authors should include comparisons to
-spikeslabGAM.\em
+spikeslabGAM._
 

### 2022-02-21 _targets.R
Add linear simulation sections

@@ -24,7 +24,7 @@ tar_plan(
   #  Simulations Studies -----------------------------------------------------
   #* Main Simulation Study ####
   tar_files(sim_main_path,
-            list.files("/data/user/boyiguo1/bgam/sim_res/main", recursive = T, full.names = T)),
+            list.files("/data/user/boyiguo1/bgam/sim_res_20220220/main", recursive = T, full.names = T)),
   sim_success_rate = create_success_rate_table(sim_main_path),
   sim_mdl_fail = create_mdl_fail_rate(sim_success_rate),
   
@@ -37,6 +37,16 @@ tar_plan(
   #*** Variable Selection  ####
   sim_binom_var_select_raw = sim_var_select_raw(sim_success_rate,
                                                        dist = "binomial"),
+  
+  # Table Presentation of Var Selection
+  # make_sim_var_metric_raw(sim_binom_var_select_raw[[5]]) %>%
+  #   group_by(Method, Metric) %>%
+  #   summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
+  #             ) %>%
+  #   pivot_wider(names_from = Method, values_from = mean),
+  
+  
+  
   # plot_var_select(sim_binom_var_select_raw[[1]])
   sim_binom_comp_select_raw = sim_comp_select_raw(sim_success_rate,
                                                 dist = "binomial"),
@@ -51,12 +61,7 @@ tar_plan(
   #   theme(axis.text.x = element_text( angle = 90))+
   #   facet_wrap(~Metric, ncol = 3, nrow = 1, scales = "free"),
 
-  # Table Presentation of Var Selection
-  make_sim_var_metric_raw(sim_binom_var_select_raw[[4]]) %>%
-    group_by(Method, Metric) %>%
-    summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
-              ) %>%
-    pivot_wider(names_from = Method, values_from = mean),
+
   
   #** Gaussian Outcome  ####
   #*** Prediction  ####
@@ -72,7 +77,7 @@ tar_plan(
   # sim_gaus_var_select = make_sim_var_select_table(sim_success_rate,
   #                                                  dist = "gaussian"),
   
-  make_sim_var_metric_raw(sim_gaus_var_select_raw[[2]]) %>%
+  make_sim_var_metric_raw(sim_gaus_var_select_raw[[4]]) %>%
[…]

### 2022-02-21 _targets.R
tuning real data analysis for BHAM v0.2.0

@@ -81,7 +81,7 @@ tar_plan(
   
   ECB_bamlasso_fnl = bamlasso(ECB_dsn_mat, ECB_outcome, family = "binomial",
                           group = make_group(names(ECB_dsn_mat)),
-                          ss = c(0.025, 0.5)),
+                          ss = c(0.065 , 0.5)),
   
   ECB_bamlasso_insample_msr = measure.bh(ECB_bamlasso_fnl),
   ECB_bamlasso_var = bamlasso_var_selection(ECB_bamlasso_fnl),
@@ -99,7 +99,7 @@ tar_plan(
   # ggplot(ECB_SBGAM_cv_per_fold)+
   #   geom_violin(aes(x = lambda0, y = deviance)),
   
-  ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 24, a = 1, b = 1),
+  ECB_SBGAM_fnl = SBGAM(X = ECB_cov, y = ECB_outcome, family = "binomial", lambda0 = 40, a = 1, b = 1),
 
   ECB_SBGAM_insample_msr = measure.glm(y = ECB_outcome, ECB_SBGAM_fnl$mu.pred, family = "binomial"),
   
@@ -156,13 +156,13 @@ tar_plan(
                               ss=c(0.005, 0.5),
                               group = WLM_bgam_dat$group),
   
-  WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw, s0 = seq(0.005, 0.1, 0.01)),
+  WLM_bamlasso_cv = tune.bgam(WLM_bamlasso_raw, s0 = seq(0.001, 0.1, 0.01)),
 
   
   WLM_bamlasso_fnl = bamlasso(WLM_bgam_dat$train_dat, WLM_train_dat$out_HOMA_std, family = "gaussian",
-                          ss=c(0.025, 0.5),
+                          ss=c(0.001, 0.5),
                           group = WLM_bgam_dat$group),
-  # measure.bh(bglm_mdl_fnl),
+  # measure.bh(WLM_bamlasso_fnl),
   WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),
   
   # # ** BHAM-IWLS --------------------------------------------------------------

### 2022-02-21 Manuscript/03-Simulation.Rmd
Add prediction result first draft

@@
 For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.
+$$
 MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
@@
 where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.
-$$
+
 }
@@
 \bg{
+\begin{itemize}
+  \item When the linear assumptions are violated, the linear lasso model performs extremely bad. adaptive cosso have improvement upon cosso.
+  \item To note, the drastic deteriation in the prediction of mgcv is caused by the model would not converge in those scenarios. A null model were returned as a final model. 
+  \item Within the simulations where the functional form of the predictors are linear, we see that the proposed model similar to the lasso model regardless of the type outcome, and better moderately better than the model of comparison.
+(See Table Lnr Gause, and supporting table Lnr_binom.) Particularly, in gaussina simulations, the proposed model shows great advantage compared to SB_GAM (numeralize the advantage), and slightly improvement compare to ssGAM (numeralize the advantage).
+This shows that the model can be applied without constraining to the assumption where all functional form are non-linear.
+  \item
 
+\end{itemize}
 }
@@
 \bg{\subsubsection{Variable Selection Performance}
+While the main utility of the proposed work is prediction, the proposed model can also serve as variable selection/functional selection as a by-product.
+Hence We evaluate the variable selection performance.
+
 \begin{itemize}

### 2022-02-21 Manuscript/SS_GAM.tex
Add prediction result first draft

@@
 SB-GAM
-\citep{Bai2021}\textcolor{red}{, linear lasso model and spikeSlabGAM [TODO: add citation here]}.
+\citep{Bai2021}\textcolor{red}{, linear lasso model and spikeSlabGAM [TODO: add citation here].[TODO: explain about linear lasso model and spikeSlabGAM model]}.
 COSSO is one of the earliest smoothing spline models that consider
@@
 For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-
-\textcolor{red}{\subsection{Simulation Results}}
-\textcolor{red}{\subsubsection{Prediction Performance}}
-
-The predictive performances have a consistent pattern across the two
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.
+MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
+$$
+where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.
+$$
+} The predictive performances have a consistent pattern across the two
 distributions of outcomes. Across all the scenarios, COSSO and adaptive
@@
 
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
+\textcolor{red}{
+
+}
+
+\textcolor{red}{\subsubsection{Variable Selection Performance}
+\begin{itemize}
+\item If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
+The pattern is consitent regardless of the type of outcomes.
+To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
+Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.
+\item If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
+While the recall is lower than SBGAM.
+This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.
+\item We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while. On the contrary the SBGAM is not so well.
+\end{itemize}
+
+}
 

### 2022-02-22 Manuscript/00-main.Rmd
Organize prediction Table to latex output (#23)

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(eval = FALSE, echo = FALSE, warning = FALSE, message = FALSE)
 
@@
 
-tar_load(intro_path)
-tar_load(method_path)
-tar_load(sim_path)
-tar_load(real_data_path)
-tar_load(disc_path)
-
-tar_load(sim_binom_tab)
-tar_load(sim_gaus_tab)
+tar_load(manu_path)
+tar_load(manu_tbs)
 

### 2022-02-22 Manuscript/10-Supplementary.tex
Organize prediction Table to latex output (#23)

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Supporting Infromation},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Supporting Infromation}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\newcommand{\tp}{*}
\newcommand{\pr}{\text{Pr}}
\newcommand{\bs}[1]{\boldsymbol{#1}}

\subsection{Supplementary Information 1: Marginal Distribution of $\gamma_j^\tp$}

Given that
\(\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)\)
where \(\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\), we can derive the
the marginal distribution of \(\gamma_{j}^\tp\) with the following
manipulation.

\begin{align*}
& \pr(\gamma_j^\tp = 1 | \theta_j)  = \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 1 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^\tp \text{ and }\gamma \text{.]}\\
= & \pr(\gamma_j^\tp = 1| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j)\\
= & \theta_j^2\\
& \pr(\gamma_j^\tp = 0 | \theta_j)  = \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0 , \gamma_j = 0| \theta_j)\\
= & \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0, \gamma_j = 0 | \theta_j)\\
= & \pr(\gamma_j^\tp = 0| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j) + \pr(\gamma_j^\tp = 0| \gamma_j = 0,  \theta_j)\pr(\gamma_j = 0| \theta_j)\\
= & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
\end{align*}

\clearpage

\subsection{Supplementary Information 2: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}

Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative
EM-based algorithm where the iterative weighted least squares algorithm
is used to find the estimate of \(\bs \beta, \phi\) that maximizes
\(E(Q_1)\). The iterative weighted least squares algorithm was
originally proposed to fit the classical generalized linear models, and
generalized to fit some Bayesian hierarchical models.{[}@Gelman2013{]}
Yi and Ma {[}@Yi2012{]} formulated Student's t-distribution and double
exponential distribution as hierarchical normal distributions such that
generalized linear models with shrinkage priors can be easily fitted
using IWLS in combination with EM algorithm. In this work, we adapt the
EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior .

A double exponential prior, \(\beta|S \sim DE(0, S)\) can be formulated
as a hierarchical normal prior with unknown variance \(\tau^2\)
integrated out:

\begin{align*}
  \beta|\tau^2 &\sim N(0, \tau^2)\\
  \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
\end{align*}

For the mixture double exponential priors, we can define the scale
parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following Equation
(\ref{eq:ssl}). The change in the prior formulation in turn leads to the
change in the log posterior density function, as \(Q_1\) needs to
account for the hyperprior of \(\tau^2\):

\begin{equation}\label{eq:Q1_IWLS}
Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
\end{equation}

Since \(\bs \tau^2\) are not of our primary interest, we treat them as
the ``missing'' data in addition to the latent indicators
\(\bs \gamma\), and hence construct the expectation
\(E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)\) in the E-step. To
note, unlike the same latent indicator \(\gamma^\tp_j\) which is shared
by the coefficients of the non-linear terms \(\beta^\tp_{jk}\) for
\(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is coefficient specific for
\(\beta^\tp_{jk}\).
\(E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})\)
needs to be calculated to formulate \(E(Q_1)\). As neither
\(E({S_j}^{-1}|\beta_j, s_0, s_1)\) nor
\(E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)\) depends on \(\tau^2\)s,
they can be derived using Equation (\ref{eq:exp_scale}). On the other
hand, \(\tau^{2}\), following gamma distributions, is a conjugate prior
for the normal variance, and the conditional posterior density of
\(\tau^{-2}\) is an inverse Gaussian distribution.
\(E({\tau}^{-2}_{j})\) and \(E({\tau^\tp}^{-2}_{jk})\) are calculated
using the closed form equation

\begin{align*}
 E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
\end{align*}

where \(S_j\) and \(S_j^\tp\) are replaced by the expectation and
\(\beta\)s are replaced with \(\beta^{(t-1)}\). With simplification (up
to constant additive terms), we have

\begin{equation}\label{eq:EQ1_IWLS}
E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
\end{equation}

\(2E({\tau}^{-2})\beta^2\) can be seen as the kernel of a normal density
with mean 0 and variance \(E(\tau^{2})\), and we can formulate the
coefficients \(\bs \beta\) as a multivariate normal distribution with
means \(\bs 0\) and variance covariance matrix \(\bs \Sigma_{\tau^2}\),
where \(\bs \Sigma_{\tau^2}\) is a diagonal matrix with \(E(\tau^2)\)s
on the diagonal, \[
\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
\]

Meanwhile, following the classical IWLS, we can approximate the
generalized model likelihood at each iteration with a weighted normal
likelihood: \[
f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
\] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
the pseudo-response and pseudo-weight respectively. The pseudo-response
and the pseudo-weight are calculated by \[
\begin{aligned}
z_i &= \hat\eta_i - \frac{L^{'}(y_i|\hat\eta_i)}{L^{''}(y_i|\hat\eta_i)}& w_i &= - L^{''}(y_i|\hat\eta_i),
\end{aligned}
\] where \(\hat\eta_i = (\bs X {\hat{\bs\beta}})_i\),
\(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
\(L^{''}(y_i|\hat\eta_i, \hat \phi)\) are the first and second
derivative of the log density, \(\log f(\textbf{y}_i|\bs \beta, \phi)\)
with respect to \(\eta_i\).

With \(\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)\) and
\(\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})\), we can
augment the two multivariate normal distributions and update the
estimates for \(\bs \beta\) and \(\phi\) via least squares in each
iteration of the EM algorithm. We create the augmented response,
augmented data, and augmented variance-covariance matrix following

\begin{align*}
& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
\end{align*}

such that \[
\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
\] Using the least squares estimators to update \(\bs\beta\) and
\(\phi\), we have

\begin{align*}
& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
\end{align*}

To note, the variance-covariance matrix of the coefficient estimates
variance-covariance matrix can be derived in the EM-IWLS algorithm and
in turn can be used for statistical inferences, \[
  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
\]

Totally, the proposed EM-IWLS algorithm is summarized as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Choose a starting value \(\bs \beta^{(0)}\) and \(\bs \theta^{(0)}\)
  for \(\bs \beta\) and \(\bs \theta\). For example, we can initialize
  \(\bs \beta^{(0)} = \bs 0\) and \(\bs \theta^{(0)} = \bs 0.5\)
\item
  Iterate over the E-step and M-step until convergence

  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^\tp_{j})\) and
  \(E(\tau^{-2}_{j})\), \(E({\tau^\tp}^{-2}_{jk})\) with the estimates
  \(\Theta^{(t-1)}\) from the previous iteration

  M-step:

  \begin{enumerate}
  \def\labelenumii{\alph{enumii})}
  \tightlist
  \item
    Based on the current value of \(\beta\), calculate the pseudo-data
    \(z_i^{(t)}\) and the pseudo-weights \(w_i^{(t)}\)
  \item
    Update \(\bs \beta^{(t)}\) by runing the augmented weighted least
    squared
  \item
    If \(\phi\) is present, update \(\phi\)
  \end{enumerate}
\end{enumerate}

Similar to EM-CD, we assess convergence by the criterion,
\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where \(\epsilon\) is
a small value (say \(10^{-5}\)).

\subsection{Supplementary Information 3: Predictive Performance of Linear Simulations}

\input{Manuscript/Tabs/sim_lnr_gaus_tab.tex} \clearpage
\input{Manuscript/Tabs/sim_lnr_binom_tab.tex}

\end{document}

### 2022-02-22 Manuscript/Tabs/sim_binom_tab.tex
Organize prediction Table to latex output (#23)

--- 2022-02-15 Manuscript/SS_GAM.tex
@@
-\documentclass[AMA,STIX1COL,]{WileyNJD-v2}
-
-% For Pandoc highlighting needs
-
-% Pandoc citation processing
-
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pdflscape}
-\usepackage{color}
-\usepackage{soul}
-\usepackage{multirow}
-\usepackage{multicol}
-\usepackage{colortbl}
-\usepackage{hhline}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{hyperref}
-
-\articletype{Research article}
-
-\received{2021-01-01}
-
-\revised{2021-02-01}
-
-\accepted{2021-03-01}
-
-\raggedbottom
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\begin{document}
-
-\title{Spike-and-Slab Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data}
-
-\author[a]{Boyi Guo*}
-\author[b]{Byron C. Jaeger}
-\author[a]{AKM Fazlur Rahman}
-\author[a]{D. Leann Long}
-\author[a]{Nengjun Yi*}
-
-\authormark{Guo et al.}
-
-\address[a]{Department of Biostatistics, University of Alabama at
-Birmingham, Birmingham, USA}
-\address[b]{Department of Biostatistics and Data Science, Wake Forest
-School of Medicine, Winston-Salem, USA}
-
-\corres{Boyi Guo and Nengjun Yi, Department of Biostatistics, University
-of Alabama at Birmingham, Birmingham, USA. \email{boyiguo1@uab.edu},
-\email{nyi@uab.edu}}
-
-\presentaddress{This is sample for present address text this is sample
-for present address text}
-
-\abstract{There are proposals that extend the classical generalized
-additive models (GAMs) to accommodate high-dimensional data (\(p>>n\))
-using group sparse regularization. However, the sparse regularization
-may induce excess shrinkage when estimating smoothing functions,
-damaging predictive performance. Moreover, most of these GAMs consider
-an ``all-in-all-out'' approach for functional selection, rendering them
-difficult to answer if nonlinear effects are necessary. While some
-Bayesian models can address these shortcomings, using Markov chain Monte
-Carlo algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a
-solution: we consider the smoothing penalty for proper shrinkage of
-curve interpolation and separation of smoothing function linear and
-nonlinear spaces. A novel spike-and-slab spline prior is proposed to
-select components of smoothing functions. Two scalable and deterministic
-algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
-Squares, are developed for different utilities. Simulation studies and
-metabolomics data analyses demonstrate improved predictive or
-computational performance against state-of-the-art models, mgcv, COSSO
-and sparse Bayesian GAM. The software implementation of the proposed
-models is freely available via an R package BHAM.}
-
-\keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
-Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility}
-
-\maketitle
-
-\pgfplotsset{compat=1.18}
-\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\tp}{*}
-\newcommand{\pr}{\text{Pr}}
-\newcommand{\repa}{\text{repa}}
-\newcommand{\simiid}{\overset{\text{iid}}{\sim}}
-\newcommand{\bg}[1]{\textcolor{red}{#1}}
-
-\section{Introduction}
-\label{sec:intro}
-
-Many modern biomedical research, e.g.~sequencing data analysis, electric
-health record data analysis, require special treatment of
-high-dimensionality, commonly known as \(p >> n\) problem. There is
-extensive literature on high-dimensional linear models via penalized
-models or Bayesian hierarchical models, see Mallick and Yi
-\citep{Mallick2013} for review. These models are built upon a
-restrictive and unrealistic assumption, linearity. In classical
-statistical modeling, many strategies and models are proposed to relax
-the linearity assumption with various degrees of complexity. For
-example, variable categorization is a simple and common practice in
-epidemiology, but suffers from power and interpretation issues. More
-complex models to address nonlinear effects include random forest and
-other so-called ``black box'' models \citep{Breiman2001}. These models
-are useful for statistical prediction but do not estimate parameters
-relevant to the data generation process that one can draw inferences
-from. In addition, how to generalize these models to the
-high-dimensional setting remains unclear.
-
-Nonparametric regression models are appropriate alternatives to the
-``black-box'' models thanks to their balance between model flexibility
-and interpretability. Among those, generalized additive models (GAMs),
-proposed in the seminal work of Hastie and Tibshirani
-\citep{Hastie1987}, grew to be one of the most popular modeling tools.
-In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution with mean \(\mu\) and dispersion
-\(\phi\), can be modeled with the summation of smoothing functions,
-\(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
-\] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
-The smoothing functions can take many forms and are estimated using a
-pseudo-weighted version of the backfitting algorithm
-\citep{Breiman1985}. Nevertheless, the classical GAMs cannot fulfill the
-increasing analytic demands for high-dimensional data analysis in
-biomedical studies.
-
-There exists some proposals to generalize the classical GAM to
-accommodate high-dimensional applications. The regularized models,
-branching out from group regularized linear models, are used to fit GAMs
-by accounting for the structure introduced when expanding smoothing
-functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
-lasso \citep{Yuan2006} to additive models (AMs); Huang et al.
-\citep{Huang2010} further developed adaptive grouped lasso for additive
-models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
-respectively applied grouped SCAD penalty \citep{Fan2001} to additive
-models. Recently Bayesian hierarchical models are also used in the
-context of high-dimensional additive models. Various group
-spike-and-slab priors combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Bai et al.
-\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
-to Gaussian AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs\citep{Bai2021}. Focus on addressing
-the sparsity, these methods can overly penalize the basis function
-coefficients and produce inaccurate predictions and curve interpolation,
-particularly when complex signals are assumed and large number of knots
-are used. \citep{Scheipl2013} In addition, these methods adapt an
-`all-in-all-out' strategy, i.e.~either including or excluding the
-variable completely, rendering no space for bi-level selection. Scheipl
-et al. \citep{Scheipl2012} proposed a spike-and-slab structure prior
-that address the previous challenges. But the model fitting relies on
-computational intensive MCMC algorithms and creates scalability concern.
-It would be of special interest to develop a fast, flexible and accurate
-generalized additive model framework.
-
-To address these challenges, we propose a novel Bayesian hierarchical
-generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via
-re-parameterization of the smoothing function to avoid overly shrinking
-basis function coefficients. Smoothing penalties are commonly
-implemented in the classical GAMs through smoothing regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive
-penalties on each smoothing function. A smoothing penalty conditioning
-on a smoothing parameter \(\lambda_j\) is a function of the integration
-of the second derivative of the spline function, expressed
-mathematically as \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
-matrix, and \(\boldsymbol{\beta}_j\) are the basis function
-coefficients. Smoothing penalties were also previously used in the
-spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
-penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of a smoothing function from
-the nonlinear space. We then impose a new two-part spike-and-slab spline
-prior on the smoothing functions for bi-level selection such that the
-linear and nonlinear spaces of smoothing functions can be selected
-separately. The prior setup encourages a flexible solution, rendering
-one of three possibilities for each predictor: no effect, only linear
-effect, or linear and nonlinear effects. In addition, two scalable
-optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm
-and EM-Iterative weighted least square (EM-IWLS) algorithm, are
-developed and implemented in an publicly available R package
-\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}, making
-translational science more accessible.
-
-The proposed framework, BHAM, differs from previous spike-and-slab based
-GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
-spline prior is a spike-and-slab lasso type prior using independent
-mixture double exponential distribution, compared to spike-and-slab GAM
-that uses normal-mixture-of-inverse gamma prior. Spike-and-slab lasso
-priors provide computational convenience during model fitting by using
-optimization algorithms instead of intensive sampling algorithms. They
-make fitting high-dimensional models more feasible without sacrificing
-performance in prediction and variable selection. Secondly, SB-GAM uses
-a group spike-and-slab lasso prior with an EM-CD algorithm to fit the
-model. While both methods use the combination of expectation
-maximization algorithm and coordient descent algorithm, there are subtle
-difference in the implementation due to the difference in prior
-specification. The proposed model sets up independent priors among basis
-function coefficients after the re-parameterization step, which provides
-some advantage in computation. In addition, we offer an alternative
-algorithm, EM-IWLS, that can provide variance-covariance matrix of the
-coefficients. Last but not least, the proposed model addresses the
-incapability of bi-level selection in SB-GAM.
-
-In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical
-generalized additive model, introduce the proposed spike-and-slab spline
-priors, and describe the two fast-fitting algorithms. In Section
-\ref{sec:sim}, we compare the proposed framework to state-of-the-art
-models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation
-studies. Analyses of two metabolomics datasets are presented in Section
-\ref{sec:real_data}. Conclusion and discussions are given in Section
-\ref{sec:concl}.
-
-\section{Bayesian Hierarchical Additive Models (BHAMs)}
-\label{sec:BHAM}
-
-Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model with link function \(g(\cdot)\) and
-linear predictor \begin{equation}\label{eq:gam}
-\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
-distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
-and dispersion parameter \(\phi\), and the data distribution is
-\begin{equation}
-f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} The basis function matrix, i.e.~the design matrix derived
-from the smoothing function \(B_j(x_j)\), is denoted
-\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
-design matrix depends on the choice of the smoothing function, and is
-denoted as \(K_j\) for \(x_j\). Its corresponding smoothing penalty
-matrix is denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\)
-denotes the basis function coefficients for the \(j\)th variable such
-that \(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
-abuse of notation, we denote vectors and matrices in bold fonts
-\(\boldsymbol{\beta}, \boldsymbol{X}\) with conformable dimensions,
-where scalar and random variables are denoted in unbold fonts
-\(\beta, X\). The matrix transposing operation is denoted with a
-superscript \(^T\). To note, the proposed model can include parametric
-forms of variables in the model, treating the parametric function as a
-special case of the smoothing function, e.g.~\(B_j(x_j) = x_j\) with the
-smoothing penalty matrix defined as
-\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
-
-To encourage proper smoothing of additive functions, we adopt the idea
-of smoothing penalties from smoothing spline models. The basic idea is
-to set up a smoothing penalty, described in Equation
-(\ref{eq:smoothpen}), in the prior density function. However, the direct
-integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and Wood \citep{Marra2011} proposed a re-parameterization
-procedure to accommodate the smoothing penalty implicitly. Given the
-smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
-semi-definite for univariate smoothing functions, we apply
-eigendecomposition on the penalty matrix
-\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
-where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
-arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
-elements of zeros on the diagonal, where the zeros are associated with
-the linear space of the smoothing function. For the most popular
-smoothing function, cubic splines, the dimension of the linear space is
-one. Hereafter, we focus on discussing a uni-dimensional linear space
-for simplicity; however, it generalizes easily to the cases where the
-linear space is multidimensional. We further write the orthonormal
-matrix
-\(\boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
-containing the eigenvectors as columns in the corresponding order to
-\(\boldsymbol{D}\). That is, \(\boldsymbol{U}\) contains the
-eigenvectors \(U^0\) with zero eigenvalues for the linear space and
-\(\boldsymbol{U}^{*}\) contains the eigenvectors (as columns) for the
-non-zero eigenvalues, i.e.~the non-linear space. We multiply the basis
-function matrix \(\boldsymbol{X}\) with the orthonormal matrix
-\(\boldsymbol{U}\) for the new design matrix
-\({\boldsymbol{X}}^\text{repa}= \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} X^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
-An additional scaling step is imposed on \(\boldsymbol{X}^{*}\) by the
-non-zero eigenvalues of \(\boldsymbol{D}\) such that the new basis
-function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
-each of its dimensions. With slight abuse of the notation, we drop the
-superscript \(^\text{repa}\) and denote
-\(\boldsymbol{X}_j \equiv \begin{bmatrix} X_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
-as the basis function matrix for the \(j\)th variable after the
-re-parameterization. A spline function can be expressed in the matrix
-form \[
-B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \beta_j X^0_j + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
-\] and the generalized additive model in Equation (\ref{eq:gam}) now is
-\begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
-\end{equation} where the coefficients
-\(\boldsymbol{\beta}_j \equiv \begin{bmatrix} \beta_j : \boldsymbol{\beta}^*_j \end{bmatrix}\)
-is an augmentation of the coefficient scalar \(\beta_j\) of linear space
-and the coefficient vector \(\boldsymbol{\beta}^*_j\) of non-linear
-space.
-
-The re-parameterization step sets up the foundation of the proposed
-model and provides three benefits. First of all, the re-parameterization
-integrates the smoothing penalty into the design matrix, and encourages
-models to properly smooth the nonlinear function in addition to the
-sparse penalty for functional selection. Secondly, the
-eigendecomposition of the smoothing penalty allows the isolation of the
-linear space from the nonlinear space, improving the feasibility of
-bi-level functional selection. The eigendecomposition facilitates the
-construction of orthonormal design matrix, which makes imposing
-independent priors on the coefficients possible. This reduces the
-computational complexity compared to using a multivariate priors, and
-greatly broadens the choices of priors and further model choices. Last
-but not least, the scaling step of \(\boldsymbol{X}^*\) further
-simplifies the choice of priors, from column-specific priors to a
-unified prior.
-
-\subsection{Spike-and-Slab Spline Priors}
-
-The family of spike-and-slab (SS) regression models is one of most
-commonly used models in high-dimensional data analysis for its utility
-in outcome prediction and variable selection. We defer to Bai et al.
-\citep{Bai2021Review} for an in-depth introduction to spike-and-slab
-priors. To summarize, spike-and-slab priors are a family of mixture
-distributions that comprises a skinny spike density
-\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
-\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-\] The most distinct feature of SS priors is that it is conditioned on a
-latent binary variable \(\gamma \in \{0,1\}\) that indicates whether the
-variable \(x\) is included in the model. There are various
-spike-and-slab priors depending on the choice for the spike density
-\(f_{\text{spike}}(\cdot)\) and the slab density
-\(f_{\text{slab}}(\cdot)\), see George and McCulloch
-\citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
-variables; Brown et al. \citep{Brown1998} for multivariate outcomes;
-Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
-\citep{Clyde2004} and reference therein. Two of most popular spike and
-slab priors are spike-and-slab normal prior \citep{George1993} and
-spike-and-slab double exponential prior\citep{Rockova2018}.
-
-The major criticism of early spike-and-slab models is being
-computationally prohibitive. \citep{Bai2021Review} Since then, many
-studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS based on spike-and-slab normal
-prior\citep{Rockova2014a} and spike-and-slab Lasso (SSL)
-\citep{Rockova2018b, Rockova2018}. Particularly, the development of SSL
-model substantially improves the scalability of SS models, setting up
-the theoretical foundation for generalized models in -omics data
-analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The SSL prior
-is composed of two double exponential distributions with mean 0 and
-different dispersion parameters, \(0 < s_0 < s_1\), mathematically,
-\begin{equation} 
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation} Given that both double exponential distributions have a
-mean of 0 and the latent indicator \(\gamma\) can only take the value of
-0 or 1, the mixture double exponential distribution can be formulated as
-one single double exponential density, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
-\end{equation} with the scale parameter
-\(S = (1-\gamma)s_0 + \gamma s_1\). The SSL also mitigates the problem
-of EMVS where the weak signals are not shrink to zero, and hence is
-preferred in high-dimensional data analysis. We notice that Bai
-\citep{Bai2021} is the first to apply spike-and-slab lasso prior in the
-GAM framework, where the densities of the spike and slab components take
-the group lasso density \citep{Xu2015} and limits to an
-``all-in-all-out'' strategy for functional selection.
-
-\subsubsection{Two-part Spike-and-Slab Lasso Prior}
-
-We introduce a novel prior for GAMs, particularly for high-dimensional
-nonlinear modeling with bi-level selection. The proposed prior extends
-from the spike-and-slab lasso prior described in Equation
-(\ref{eq:ssl}). Given the re-parameterized design matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
-for the \(j\)th variable, we impose a two-part SSL prior to the
-coefficients
-\(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
-Specifically, we impose independent group priors on the linear space
-coefficients and on the nonlinear space coefficients respectively,
-\begin{align}\label{eq:bham_ssl}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j
-\end{align} where \(\gamma_{j}\in\{0,1\}\) and
-\(\gamma^*_{j}\in \{0,1\}\) are two latent indicator variables,
-indicating if the model includes the linear effect and the nonlinear
-effect of the \(j\)th variable respectively. \(s_0\) and \(s_1\) are
-scale parameters, assuming \(0 < s_0 < s_1\) and given. These scale
-parameters \(s_0\) and \(s_1\) can be treated as tuning parameters and
-optimized via cross-validation. A discussion of how to choose the scale
-parameters comes in Section \ref{sec:tune}. To note, this prior differs
-from previous group SSL priors \citep{Tang2018, Tang2019}, as the
-\(\beta_j\) and \(\boldsymbol{\beta}^*_j\) have different indicator
-variables \(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to
-add a more restrictive assumption on the priors, assuming that one
-indicator variable decides the inclusion of both the linear effect and
-nonlinear effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the
-SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
-assumption such that each coefficient
-\(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has its own latent indicator
-\(\gamma_{jk}\), but at the cost of complicating the bi-level functional
-selection. This reduces the proposed prior to the classic SSL prior.
-
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the
-validity of the proposed prior. First of all, the smoothing function
-bases are linear dependent and necessitate extra attention. The
-eigeondecomposition remedies the problem and hence our prior can be set
-to be conditionally independent. Secondly, the eigenvalue scaling
-provides a panacea to allow unified scale parameters for all bases of
-all smoothing functions.
-
-The rest of the hierarchical prior follows the traditional SSL prior: we
-set up hyper-priors on \(\gamma_j, \gamma^*_j\) to allow local adaption
-of the shrinkage using a Bernoulli distribution, written as binomial
-distribution of one trial. The two indicators of the \(j\)th predictor,
-\(\gamma_j\) and \(\gamma^*_j\), shares the same probability parameter
-\(\theta_j\), \[
-\begin{aligned}
-&\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^*| \textcolor{red}{\gamma_{j}, } \theta_j \sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j).
-\end{aligned}
-\]
-\textcolor{red}{Equivalently, we can derive the marginal distribution of $\gamma_{j}^*$ by integrate out $\gamma_j$ (Supplement 1), 
-$$
-\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2).
-$$
-}
-
-This is to leverage the fact that the probability of selecting the bases
-of a smoothing function should be similar, while allowing different
-penalty on the linear space and non-linear space of the smoothing
-function. The hyper prior of \(\gamma_{j}\) decides the sparsity of the
-model at the functional selection level, while that of \(\gamma_{j}^*\)
-decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that \(\gamma_{j}\) and \(\gamma_{j}^*\) are
-independently distributed for analytic simplicity. We further specify
-the parameter \(\theta_j\) follows a beta distribution with given shape
-parameters \(a\) and \(b\), \[
-\theta_j \sim Beta(a, b).
-\] The beta distribution is a conjugate prior for the binomial
-distribution and hence provides some computation convenience. For
-simplicity, we focus on a special case of beta distribution, uniform
-(0,1), i.e.~\(a = 1, b = 1\). When the variable have large effects in
-any of the bases, the parameter \(\theta_j\) will be estimated large,
-which in turn encourages the model to include the rest of bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models
-with the spike-and-slab spline prior as the BHAM, and visually presented
-in Figure \ref{fig:SSprior}.
-
-\begin{center}
-Figure \ref{fig:SSprior} here
-\end{center}
-
-\subsubsection{Other Priors}
-
-With the re-parameterization step of the basis function matrix
-\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
-priors, for example normal priors for ridge-type regularization and
-mixture normal prior for spike-and-slab regularization. These priors
-would work better in low and medium dimensional settings where the
-sparse assumption is not necessary. Here we elaborate the mixture normal
-prior as a demonstration of applying continuous spike-and-slab prior in
-BHAM.
-
-A spike-and-slab mixture normal spline prior can be expressed as
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
-\end{align*} Similar to the spike-and-slab spline prior in Equation
-(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
-optimized via cross-validation. One of the critics received by the
-spike-and-slab mixture normal prior is that the tails of a normal
-distribution diminishes to zero too fast, which causes problems when
-estimating the large effects. Distributions with heavier tails can be
-used as an alternative, for example mixture Student's \(t\) distribution
-with small degree of freedom.
-
-\subsection{Algorithms for Fitting BHAMs}
-
-The proposed models can be fitted with MCMC algorithms. Nevertheless,
-the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch\citep{George1997} examined the computation speed
-for various MCMC algorithms with spike-and-slab mixture normal priors,
-and suggested MCMC algorithms works well for medium size (\(p\)=25) of
-predictors with only linear effects. However, it is not feasible for
-high-dimensional data analysis where the number of predictors easily
-exceeds 100. Specific to additive models, each predictor would expand to
-multiple new ``predictors'' via basis functions, creating greater
-computational demands. Scheipl et al. \citep{Scheipl2013} demonstrated
-the computational demands of a MCMC algorithm for fitting spike-and-slab
-GAM grow exponentially as \(p\) increases modestly via simulation
-studies. Hence, we feel compelled to develop scalable algorithms for
-fitting Bayesian hierarchical additive models in high-dimensional
-settings.
-
-As an alternative to sampling algorithms for fitting Bayesian models,
-optimization algorithms focus on the maximum a posteriori (MAP)
-estimates and speed up the model fitting process at the cost of
-posterior inference. The earlier work for fitting SS models using
-optimization algorithms includes EMVS \citep{Rockova2014a}. Rockova and
-George\citep{Rockova2014a} proposed an expectation-maximization (EM)
-based algorithm to fit models that use continuous mixture normal priors.
-In the E step, the latent binary indicators \(\gamma\)s are treated as
-the missing data, and the posterior means are calculated conditioning on
-the current value of other parameters; in the M step, a ridge estimator
-was used to update the coefficients, followed by updating
-\(\phi, \theta\). The same authors \citep{Rockova2018b, Rockova2018}
-further combined the EM algorithm with coordinate descent algorithm to
-fit SSL models. Yi and his group independently developed the
-EM-Iterative Weighted Least Square and EM-cyclic coordinate descent
-algorithms to fit models with broader class of priors, SSL
-included.\citep{Yi2019} These algorithms were implemented for
-generalized linear models\citep{Tang2017a}, Cox proportional hazards
-models \citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a popular property for
-reproducible research
-
-In this section, we extend the two EM-based algorithms, EM-CD and
-EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
-provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM with an expedited performance, recommending to use in high
-and ultra-high dimensional setting. A specific concern of EM-CD
-algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can estimate the variance-covariance matrix of the
-coefficients. Moreover, the EM-IWLS is a more general model fitting
-algorithm that can be used for fitting not only SSL and continuous SS
-priors but also Student's t-priors and double exponential priors. To
-note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
-main difference between proposed EM-CD algorithm and that in SB-GAM is
-that SB-GAM uses a block CD algorithm for their group prior, while the
-proposed prior is pairwise independent requiring no special treatment in
-the CD algorithm.
-
-\subsubsection{EM algorithms}
-
-EM algorithm is an iterative algorithm to find MAP estimates or the
-maximum likelihood estimates. It is commonly used when some necessary
-data to establish the likelihood function are missing. Instead of
-maximizing the the likelihood function, the algorithm maximizes the
-expectation of the likelihood function with respect to the ``missing''
-data.
-
-The recursive algorithm consists of two steps:
-
-\begin{itemize}
-\tightlist
-\item
-  E-step: to calculate the expectation of the posterior density function
-  with respect to some ``missing'' data
-\item
-  M-step: to maximize the expectation derived in the E-step and update
-  parameters of interest
-\end{itemize}
-
-For BHAMs, we define the parameters of interest as
-\(\Theta = \{\boldsymbol{\beta}, \boldsymbol{\theta}, \phi\}\) and
-consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model, in other words the ``missing'' data.
-Our objective is to find the parameters \(\Theta\) that maximize the
-posterior density function, or equivalently, the logarithm of the
-density function, \[
-\begin{aligned}
-& \text{argmax}_{\Theta}
-\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{j})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j),
-\end{aligned}
-\]\\
-where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
-non-informative prior for the intercept \(\beta_0\) and the dispersion
-parameter \(\phi\); for example, \(f(\beta_0|\tau_0^2)=N(0,\tau_0^2)\)
-with \(\tau^2_0\) set to a large value and \(f(\log \phi) \propto 1\).
-
-We use the EM algorithm to find the MAP estimate of \(\Theta\). This is,
-in the E-step, we calculate the expectation of posterior density
-function of
-\(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
-respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
-the values from previous iteration \(\Theta^{(t-1)}\), \[
-E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
-\] Hereafter, we use the shorthand notation
-\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}(\cdot)\). In the
-M-step, we find the \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The E-
-and M- steps are iterated until the algorithm converge.
-
-To note here, the log-posterior density of BHAMs (up to additive
-constants) can be written as a two-part equation
-\[ \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
-where
-\[ Q_1\equiv Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
-and
-\[Q_2 \equiv Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j).\]
-\(Q_1\) and \(Q_2\) are respectively the log posterior density of the
-coefficients \(\boldsymbol{\beta}\) and the log posterior density of the
-probability parameters \(\boldsymbol{\theta}\) conditioning on
-\(\boldsymbol{\gamma}\). Meanwhile, conditioning on
-\(\boldsymbol{\gamma}\), \(Q_1\) and \(Q_2\) are independent and can be
-maximized separately for \(\boldsymbol{\beta}, \phi\) and
-\(\boldsymbol{\theta}\). Depending on the choice of coefficient priors,
-\(Q_1\) can be treated as penalized likelihood function and maximization
-of \(E(Q_1)\) can be solved via CD algorithm or IWLS algorithm in each
-iteration. Maximization of \(E(Q_2)\) can be solved via closed form
-equations following the beta-binomial conjugate relationship.
-
-\subsubsection{EM-Coordinate Descent}
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double
-exponential, coordinate descent algorithm can be used to estimate the
-parameters in the M-step. Coordinate descent is an optimization
-algorithm that offers extreme computational advantage, and famous for
-its application in optimizing the \(l_1\) penalized likelihood function.
-
-The density function of spike-and-slab mixture double exponential prior
-can be written as \[
-f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
-\] and \(E(Q_1)\) can be expressed as a log-likelihood function with
-\(l_1\) penalty \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
-\end{equation} where
-\(S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
-unknown quantities \(E({S_j}^{-1})\) and \(E({S^*}^{-1}_j)\), the
-posterior probability
-\(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
-\(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
-necessary, which can be derived via Bayes' theorem. The calculation of
-\(p_j^*\) is slightly different from that of \(p_j\), as \(p_j^*\)
-depends on the value of the vector \(\boldsymbol{\beta}^*_{j}\) and
-\(p_j\) only depends on the scalar \(\beta_j\). The calculation follows
-the equations below. \begin{align*}
-p_{j} &= \frac{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \text{Pr}(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
-p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{align*} and \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j}\),
-\(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
-\(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
-show \begin{align}\label{eq:exp_scale}
-&E(\gamma_{j})  = p_{j} & &E(\gamma^{*}_{j}) = p_{j}^{*}\nonumber\\
-&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^*}^{-1}_{j}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
-\end{align} After replacing withe calculated quantities, \(E(Q_1)\) can
-be seen as a \(l_1\) penalized likelihood function with the
-regularization parameter \(\lambda = E(S^{-1})\), and hence be optimized
-via coordinate descent algorithm \citep{Friedman2010}. Independently,
-the remaining parameters of interest \(\boldsymbol{\theta}\) can be
-updated by maximizing \(E(Q_2)\). As the beta distribution is a
-conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
-be easily updated with a closed form equation,
-\begin{equation}\label{eq:update_theta}
-\theta_j = \frac{p_j + p^*_{j} + a - 1 }{a + b}.
-\end{equation}
-
-Totally, the proposed EM-coordinate descent algorithm is summarized as
-follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
-  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
-  \(\boldsymbol{\theta}\). For example, we can initialize
-  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
-  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
-\item
-  Iterate over the E-step and M-step until convergence
-
-  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
-  \(E({S}^{-1}_{j})\), \(E({S^*}^{-1}_{j})\) with estimates of
-  \(\Theta^{(t-1)}\) from previous iteration
-
-  M-step:
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii})}
-  \item
-    Update \(\boldsymbol{\beta}^{(t)}\), and the dispersion parameter
-    \(\phi^{(t)}\) if exists, using the coordinate descent algorithm
-    with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
-  \item
-    Update \(\boldsymbol{\theta}^{(t)}\) using Equation
-    (\ref{eq:update_theta})
-  \end{enumerate}
-\end{enumerate}
-
-We assess convergence by the criterion:
-\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
-\(d^{(t)} = -2\log f(\textbf{y}| \textbf{X}, \boldsymbol{\beta}^{(t)},\phi^{(t)})\)
-is the estimate of deviance at the \(t\)th iteration, and \(\epsilon\)
-is a small value (say \(10^{-5}\)).
-
-\subsection{Selecting Optimal Scale Values}
-\label{sec:tune}
-
-Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
-\(s_1\)). Hence, we need to find the optimal values for the scale
-parameters such that the model reaches its best prediction performance
-regarding a criteria of preference. This would be achieved by
-constructing a two dimensional grid, consists of different pairs of
-(\(s_0\), \(s_1\)) value. However, previous research suggested the value
-of slab scale \(s_1\) have less impact on the final model and is
-recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
-provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
-on examining different values of spike scale \(s_0\). Instead of the 2-D
-grid, We consider a sequence of \(L\) decreasing values
-\(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
-spike scale \(s_0\) tends to include more non-zero coefficients in the
-model. A measure of preference calculated with cross-validations (CV),
-e.g.~deviance, area under the curve (AUC), mean squared error, can be
-used to facilitate the selection of a final model. The procedure is
-similar to the Lasso implemented in the widely used R package
-\texttt{glmnet}, which quickly fits Lasso models over a list of values
-of regularization parameters \(\lambda\) , giving a sequence of models
-for users to choose from.
-
-\section{Simulation Study}
-\label{sec:sim}
-
-In this section, we compare the performance of the proposed models to
-four alternative models: component selection and smoothing operator
-(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
-generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM \citep{Bai2021}\textcolor{red}{}, linear lasso model and
-spikeSlabGAM {[}TODO: add citation here{]}\}. COSSO is one of the
-earliest smoothing spline models that consider sparsity-smoothness
-penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for
-penalties such that the penalty of each functional component are
-different for extra flexibility. Generalized additive models with
-automatic smoothing, hereafter \textit{mgcv}, is one of the most popular
-models for nonlinear effect interpolation and prediction. SB-GAM is the
-first spike-and-slab lasso GAM. We implement COSSO and adaptive COSSO
-with R package \texttt{cosso 2.1-1}, generalized additive models with
-automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R
-package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence use the default
-choices. Both mgcv and proposed models allow customized smoothing
-functions and we choose the cubic regression spline. We controll the
-dimensionality of each smoothing function, 10 bases, for all different
-choices of smoothing functions. We use 5-fold CV with the default
-selection criteria to select the final model for COSSO models, SB-GAM
-and the proposed models. 20 default candidates of tuning parameters
-(\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for SB-GAM and
-the proposed models which allow user-specification of tuning candidates.
-All computation was conducted on a high-performance 64-bit Linux
-platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
-processors and 24G of RAM per core and R 3.6.2 \citep{R}.
-
-Other related methods for high-dimensional GAMs also exist, notably the
-methods of sparse additive models by Ravikumar et al.
-\citep{Ravikumar2009} and stochastic search term selection for GAM
-\citep{Scheipl2012}. However, we exclude these methods from current
-simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increased number of
-predictors. \citep{Scheipl2013}
-
-\subsection{Monte Carlo Simulation Study}
-
-We follow the data generating process described in Bai \citep{Bai2021}.
-We first generate \(n=500\) training data points with
-\(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
-\(X\) are simulated from a multivariate normal distribution
-\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(y\)
-from two distributions, Gaussian and binomial with the identity link and
-logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The mean of each
-outcome were simulated via the following function \[
-\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for Gaussian and binomial outcomes. Gaussian outcomes requires
-specification of dispersion, where we set the dispersion parameter to be
-1. In this data generating process, we have \(x_1, x_2, x_3, x_4\) as
-the active covariates, while the rest covariates are inactive,
-i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
-independent sample of size \(n_{test}=1000\), are created following the
-same data generating process, serving as the testing data. We generate
-50 independent pairs of training and testing datasets to evaluate the
-prediction performance of the chosen models, where training datasets are
-used to fit the models and testing datasets used to calculate assessment
-measures.
-\textcolor{red}{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
-These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
-
-To evaluate the predictive performance of the models, the statistics,
-\(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing dataset, are averaged across 50 simulations.
-\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
-For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-
-\textcolor{red}{\subsection{Simulation Results}}
-\textcolor{red}{\subsubsection{Prediction Performance}}
-
-The predictive performances have a consistent pattern across the two
-distributions of outcomes. Across all the scenarios, COSSO and adaptive
-COSSO have the least favorable performance among the applicable methods
-examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}). To note, mgcv
-doesn't support high-dimensional analysis, i.e.~the number of
-coefficients are greater than the sample size, and hence not evaluated
-when \(p=100, 200\). mgcv predicts well when \(p\) is small or moderate
-(\(p = 4, 10\)), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the
-proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when
-the dimension of are moderate (p=4, 10, 50). Particularly, BHAM-IWLS
-performs as good as mgcv if not better. However, in high dimensional
-case where we mimic the situation the signals are extremely sparse,
-SB-GAM has better performance than the proposed method. However, the
-BHAM-CD has extreme computational advantage over SB-GAM (see Table
-\ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
-
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
-
-\section{Metabolomics Data Analysis}
-\label{sec:real_data}
-
-In this section, we apply the proposed models BHAM, fitted with the
-EM-CD algorithm, to analyze two real-world metabolomics datasets where
-the outcomes are binary and continuous respectively. We demonstrate the
-improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}, while being
-computationally efficient (see Table \ref{tab:time_real_data}).
-
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
-
-\subsection{Emory Cardiovascular Biobank}
-\label{sec:ECB}
-
-We use the proposed models BHAM to analyze a metabolic dataset from a
-recently published research \citep{Mehta2020} studying plasma
-metabolomic profile on the three-year all-cause mortality among patients
-undergoing cardiac catheterization. The dataset is publicly available
-via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
-subjects from two cohorts. As there is a large number of non-overlapping
-features among the two cohorts, we use the cohort with larger sample
-size (N=454). There are initially 6796 features in the dataset, which is
-too large to be practically meaningful to analyze. Hence, we perform a
-univariate screening procedure on the features, via GAM implemented in
-\texttt{mgcv}, and choose the the top 200 features with smallest
-p-values. We use 5-knot spline additive models for binary outcome using
-two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are
-used to choose the optimal tuning parameters of each framework with
-respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where
-deviance, AUC, Brier score defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassication error defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
-calculated. BHAM-CD obtains superior AUC, Brier score, and
-misclassification error in the out-of-bag samples compared to SB-GAM
-(see Table \ref{tab:ECB_res}).
-
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
-
-\subsection{Weight Loss Maintenance Cohort}
-
-We use the proposed models BHAM to analyze metabolomics data from a
-recently published study \citep{Bihlmeyer2021} on the association
-between metabolic biomarkers and weight loss, where the dataset is
-publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
-primarily focuse on the analysis of one of the three studies included,
-weight loss maintenance cohort \citep{Svetkey2008}, due to the
-drastically different intervention effects. In the dataset, 765
-metaboliltes in baseline plasma collected were profiled using liquid
-chromatography mass spectrometry. Quality control and natural log
-transformation are performed during metabolites data preparation. The
-outcome of interest are standardized percent change in insulin
-resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there are
-\(p\)=237 features remaining in the analysis. 5-Knot spline additive
-models for the Gaussian outcome are constructed using two different
-models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
-the optimal tuning parameters of each framework with respect to the
-default selection criterion implemented in the software. Out-of-bag
-samples are used for prediction performance evaluation, where deviance,
-\(R^2\), mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-absolute error (MAE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
-BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
-compared to SB-GAM (see Table \ref{tab:WLM_res}).
-
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}
-
-\section{Discussion}
-\label{sec:concl}
-
-In the paper, we described a novel generalized additive model using
-Bayesian hierarchical priors, particularly the proposed spike-and-slab
-spline prior for bi-level functional selection. Meanwhile, we introduced
-two optimization based algorithms for model fitting. The algorithms can
-be easily scale up to address high-dimensional data analysis in a
-computational efficient manner. Via simulations, we demonstrated that
-the proposed model provides as good, if not better, prediction
-performance compared to some state-of-the-art non-linear modeling
-devices.
-
-The proposed model shares many commonality with an high-dimensional
-Bayesian GAM, SB-GAM \citep{Bai2021}, independently developed around the
-same time of this work. Both frameworks emphasize computational
-efficiency by deploying group spike-and-slab lasso type of priors and
-optimization-based fast and scalable algorithms. Bai provided the
-theoretical proof for the consistency of variable selection using group
-spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the
-bi-level functional selection and the model inference. The proposed
-model provides solutions to these remaining questions while maintaining
-the same level of prediction accuracy. Moreover, the proposed model
-renders additional flexibility to model specification, allowing various
-choices of smoothing functions and degrees of freedom.
-
-For translational science purpose, we implemented the proposed model in
-a R package \texttt{BHAM}, deposited at
-\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
-smoothing function specification, we deploy the same programming grammar
-as in the state-of-the-art package \texttt{mgcv}, in contrast to
-previous tools where smoothing functions are limited to the default
-ones. Ancillary functions are provided for model specification in
-high-dimensional settings, curve plotting and functional selection. In
-addition, in \texttt{BHAM}, we streamline the model fitting algorithms
-to support other popular Bayesian hierarchical prior for the smoothing
-functions, such as Student's T distribution, double exponential
-distribution, mixture normal and T distribution's. These priors could be
-helpful when the sparse assumption is weak or not necessary.,
-
-There are some improvements possible for the proposed models. First of
-all, the proposed model achieves a bi-level selection via the two-part
-spike-and-slab spline prior. Nevertheless, this set-up could result in a
-situation that is not theoretically sound: the non-linear component is
-selected, but the linear component is not. We currently address it
-analytically by including the linear component in the model when
-non-linear component is selected. Another possible solution is to impose
-a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, the
-computational time for fitting a BHAM model with EM-IWLS algorithms can
-be improved. Current implementation of the EM-IWLS algorithm jointly
-updates all coefficients in each iteration, which requires a lot
-computation resources. This jointly updating procedure can be enhanced
-by adapting a backfitting step \citep{Hastie1987} where each smoothing
-function are updated individually. Thirdly, when using the proposed
-model in real data analysis where a screening procedure is implemented
-before joint predictive modeling. we recommend to include the screening
-procedure in the cross-validation during model selection.
-\citep{Friedman2017} For the sake of model comparison, we fit the models
-using the same pool of predictors in Section \ref{sec:ECB}.
-
-Our future efforts direct to modeling survival outcomes and integrative
-analysis. While the proposed model addresses a great deal of analytic
-problem, analyzing the time-to-event outcome remains unsolved. An naive
-approach would be convert a time-to-event outcome to a Poisson outcome
-following Whitehead \citep{Whitehead1980}. However, it would be more
-efficient to directly fit Cox models via penalized pseudo likelihood
-function \citep{Simon2011}. Meanwhile, with growing understanding of
-biological structure within -omics field, it is appealing to integrate
-external biology information in the modeling process. The main
-motivation for integrative models is that biologically informed grouping
-of weak effects increases the power of detecting true associations
-between features and the outcome \citep{Peterson2016}, and stabilizes
-the analysis results for reproducibility purpose. Such integration can
-be achieved by setting up a structural hyperprior on the inclusion
-indicator of the smoothing function null space
-\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
-and Dunson \citep{Ferrari2020}.
-
-\clearpage
-
-\bibliography{bibfile.bib}
-
-\clearpage
-
-\begin{figure}
-\centering
-\begin{tikzpicture} [
-staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
-PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
-background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
-]
-
-\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
-};
-\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
-};
-
-\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
-\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
-\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
-
-\draw[->] (Theta) -- (linearGamma);
-\draw[->] (Theta) -- (penGamma);
-\draw[->] (linearGamma) -- (linearBeta) ;
-\draw[->] (penGamma) -- (penBeta);
-\draw[->, draw = red] (linearGamma) -- (penGamma);
-\draw[->] (ab) -- (Theta);
-\draw[->] (s) -- (linearBeta) ;
-\draw[->] (s) -- (penBeta);
-\draw[->] (linearBeta) -- (Beta);
-\draw[->] (penBeta) -- (Beta);
-\draw[->] (Beta) --  (Y);
-
-\begin{pgfonlayer}{background}
-  \node [background,
-   fit=(linearGamma) (linearBeta),
-   label=above:Linear Space:] {};
-  \node [background,
-    fit=(penGamma) (penBeta),
-    label=above:Nonlinear Space:] {};
-\end{pgfonlayer}
-
-\end{tikzpicture}
-
-\caption{Directed acyclic graph of the proposed Bayesian hierarchical additive model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
-\end{figure}
-
-\clearpage
-
 \begin{table}[ht]
 \centering
-\begin{tabular}{ccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+P & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
-  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
-  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
+  4 & 0.94 (0.01) & 0.83 (0.01) & 0.90 (0.01) & 0.90 (0.01) & 0.92 (0.01) & 0.92 (0.01) & 0.90 (0.00) \\ 
+   10 & 0.92 (0.03) & 0.83 (0.00) & 0.86 (0.04) & 0.86 (0.03) & 0.92 (0.01) & 0.92 (0.01) & 0.90 (0.00) \\ 
+   50 & 0.76 (0.03) & 0.83 (0.01) & 0.83 (0.02) & 0.84 (0.02) & 0.90 (0.01) & 0.92 (0.01) & 0.89 (0.01) \\ 
+  100 & NaN (NA) & 0.83 (0.01) & 0.83 (0.02) & 0.81 (0.09) & 0.90 (0.01) & 0.92 (0.01) & 0.88 (0.01) \\ 
+  200 & NaN (NA) & 0.83 (0.01) & 0.81 (0.06) & 0.82 (0.05) & 0.88 (0.02) & 0.92 (0.01) & 0.87 (0.02) \\ 
    \hline
 \end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:gaus}
-\end{table}
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
+\caption{The average and standard deviation of the out-of-sample area under the curve measures
+    for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
+    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
+    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO),
+    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
+    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:bin_auc}
 \end{table}
-
-\clearpage
-
-\begin{sidewaystable}[!h]
-\centering
-\begin{tabular}{lrllllll}
-  \hline
-Distribution & P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-Binomial & 4.00 & 5.88 (1.01) & 6.96 (1.76) & 2.97 (0.82) & 4.60 (2.32) & 0.18 (0.04) & 343.24 (88.37) \\ 
-  Binomial & 10.00 & 14.01 (1.86) & 7.65 (1.25) & 7.95 (2.70) & 10.01 (3.56) & 3.51 (12.03) & 543.13 (133.51) \\ 
-  Binomial & 50.00 & 185.26 (17.44) & 53.83 (6.26) & 109.72 (24.31) & 128.13 (23.26) & 670.01 (151.28) & 1630.42 (193.61) \\ 
-  Binomial & 100.00 & 1106.93 (356.73) & 48.41 (5.84) & 715.15 (170.55) & 718.63 (193.53) & - & 2783.99 (235.31) \\ 
-  Binomial & 200.00 & 6923.62 (1911.59) & 99.98 (7.05) & 5572.51 (1295.18) & 4958.69 (1970.46) & - & 4780.10 (488.66) \\ 
-  Gaussian & 4.00 & 4.22 (1.51) & 2.51 (0.20) & 0.79 (0.11) & 0.67 (0.07) & 0.05 (0.00) & 36.02 (3.02) \\ 
-  Gaussian & 10.00 & 17.59 (5.77) & 33.32 (3.50) & 3.49 (0.70) & 3.47 (0.71) & 0.32 (0.34) & 72.93 (9.52) \\ 
-  Gaussian & 50.00 & 329.83 (44.57) & 334.53 (22.33) & 35.54 (9.15) & 35.44 (9.02) & 73.33 (71.24) & 373.42 (34.93) \\ 
-  Gaussian & 100.00 & 1706.19 (225.92) & 446.98 (79.31) & 150.58 (46.72) & 152.82 (45.07) & - & 684.05 (67.80) \\ 
-  Gaussian & 200.00 & 16379.27 (4825.56) & 303.73 (87.27) & 591.11 (147.88) & 569.92 (107.45) & - & 1314.12 (137.85) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:time_sim}
-\end{sidewaystable}
-
-\clearpage
-
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
-
-\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\textbackslash{}label\{tab:time\_real\_data\}
-}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{490.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\end{longtable}
-
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & AUC & Brier & Misclass \\ 
-  \hline
-BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
-\label{tab:ECB_res}
-\end{table}
-
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & $R^2$ & MSE & MAE \\ 
-  \hline
-BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
-  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
-\label{tab:WLM_res}
-\end{table}
-
-\end{document}

### 2022-02-22 Manuscript/Tabs/sim_gaus_tab.tex
Organize prediction Table to latex output (#23)

--- 2022-02-15 Manuscript/SS_GAM.tex
@@
-\documentclass[AMA,STIX1COL,]{WileyNJD-v2}
-
-% For Pandoc highlighting needs
-
-% Pandoc citation processing
-
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pdflscape}
-\usepackage{color}
-\usepackage{soul}
-\usepackage{multirow}
-\usepackage{multicol}
-\usepackage{colortbl}
-\usepackage{hhline}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{hyperref}
-
-\articletype{Research article}
-
-\received{2021-01-01}
-
-\revised{2021-02-01}
-
-\accepted{2021-03-01}
-
-\raggedbottom
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\begin{document}
-
-\title{Spike-and-Slab Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data}
-
-\author[a]{Boyi Guo*}
-\author[b]{Byron C. Jaeger}
-\author[a]{AKM Fazlur Rahman}
-\author[a]{D. Leann Long}
-\author[a]{Nengjun Yi*}
-
-\authormark{Guo et al.}
-
-\address[a]{Department of Biostatistics, University of Alabama at
-Birmingham, Birmingham, USA}
-\address[b]{Department of Biostatistics and Data Science, Wake Forest
-School of Medicine, Winston-Salem, USA}
-
-\corres{Boyi Guo and Nengjun Yi, Department of Biostatistics, University
-of Alabama at Birmingham, Birmingham, USA. \email{boyiguo1@uab.edu},
-\email{nyi@uab.edu}}
-
-\presentaddress{This is sample for present address text this is sample
-for present address text}
-
-\abstract{There are proposals that extend the classical generalized
-additive models (GAMs) to accommodate high-dimensional data (\(p>>n\))
-using group sparse regularization. However, the sparse regularization
-may induce excess shrinkage when estimating smoothing functions,
-damaging predictive performance. Moreover, most of these GAMs consider
-an ``all-in-all-out'' approach for functional selection, rendering them
-difficult to answer if nonlinear effects are necessary. While some
-Bayesian models can address these shortcomings, using Markov chain Monte
-Carlo algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a
-solution: we consider the smoothing penalty for proper shrinkage of
-curve interpolation and separation of smoothing function linear and
-nonlinear spaces. A novel spike-and-slab spline prior is proposed to
-select components of smoothing functions. Two scalable and deterministic
-algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
-Squares, are developed for different utilities. Simulation studies and
-metabolomics data analyses demonstrate improved predictive or
-computational performance against state-of-the-art models, mgcv, COSSO
-and sparse Bayesian GAM. The software implementation of the proposed
-models is freely available via an R package BHAM.}
-
-\keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
-Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility}
-
-\maketitle
-
-\pgfplotsset{compat=1.18}
-\usetikzlibrary{shapes.geometric, arrows, positioning, calc, matrix, backgrounds, fit}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\tp}{*}
-\newcommand{\pr}{\text{Pr}}
-\newcommand{\repa}{\text{repa}}
-\newcommand{\simiid}{\overset{\text{iid}}{\sim}}
-\newcommand{\bg}[1]{\textcolor{red}{#1}}
-
-\section{Introduction}
-\label{sec:intro}
-
-Many modern biomedical research, e.g.~sequencing data analysis, electric
-health record data analysis, require special treatment of
-high-dimensionality, commonly known as \(p >> n\) problem. There is
-extensive literature on high-dimensional linear models via penalized
-models or Bayesian hierarchical models, see Mallick and Yi
-\citep{Mallick2013} for review. These models are built upon a
-restrictive and unrealistic assumption, linearity. In classical
-statistical modeling, many strategies and models are proposed to relax
-the linearity assumption with various degrees of complexity. For
-example, variable categorization is a simple and common practice in
-epidemiology, but suffers from power and interpretation issues. More
-complex models to address nonlinear effects include random forest and
-other so-called ``black box'' models \citep{Breiman2001}. These models
-are useful for statistical prediction but do not estimate parameters
-relevant to the data generation process that one can draw inferences
-from. In addition, how to generalize these models to the
-high-dimensional setting remains unclear.
-
-Nonparametric regression models are appropriate alternatives to the
-``black-box'' models thanks to their balance between model flexibility
-and interpretability. Among those, generalized additive models (GAMs),
-proposed in the seminal work of Hastie and Tibshirani
-\citep{Hastie1987}, grew to be one of the most popular modeling tools.
-In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution with mean \(\mu\) and dispersion
-\(\phi\), can be modeled with the summation of smoothing functions,
-\(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
-\] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
-The smoothing functions can take many forms and are estimated using a
-pseudo-weighted version of the backfitting algorithm
-\citep{Breiman1985}. Nevertheless, the classical GAMs cannot fulfill the
-increasing analytic demands for high-dimensional data analysis in
-biomedical studies.
-
-There exists some proposals to generalize the classical GAM to
-accommodate high-dimensional applications. The regularized models,
-branching out from group regularized linear models, are used to fit GAMs
-by accounting for the structure introduced when expanding smoothing
-functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
-lasso \citep{Yuan2006} to additive models (AMs); Huang et al.
-\citep{Huang2010} further developed adaptive grouped lasso for additive
-models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
-respectively applied grouped SCAD penalty \citep{Fan2001} to additive
-models. Recently Bayesian hierarchical models are also used in the
-context of high-dimensional additive models. Various group
-spike-and-slab priors combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Bai et al.
-\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
-to Gaussian AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs\citep{Bai2021}. Focus on addressing
-the sparsity, these methods can overly penalize the basis function
-coefficients and produce inaccurate predictions and curve interpolation,
-particularly when complex signals are assumed and large number of knots
-are used. \citep{Scheipl2013} In addition, these methods adapt an
-`all-in-all-out' strategy, i.e.~either including or excluding the
-variable completely, rendering no space for bi-level selection. Scheipl
-et al. \citep{Scheipl2012} proposed a spike-and-slab structure prior
-that address the previous challenges. But the model fitting relies on
-computational intensive MCMC algorithms and creates scalability concern.
-It would be of special interest to develop a fast, flexible and accurate
-generalized additive model framework.
-
-To address these challenges, we propose a novel Bayesian hierarchical
-generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via
-re-parameterization of the smoothing function to avoid overly shrinking
-basis function coefficients. Smoothing penalties are commonly
-implemented in the classical GAMs through smoothing regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive
-penalties on each smoothing function. A smoothing penalty conditioning
-on a smoothing parameter \(\lambda_j\) is a function of the integration
-of the second derivative of the spline function, expressed
-mathematically as \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
-matrix, and \(\boldsymbol{\beta}_j\) are the basis function
-coefficients. Smoothing penalties were also previously used in the
-spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
-penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of a smoothing function from
-the nonlinear space. We then impose a new two-part spike-and-slab spline
-prior on the smoothing functions for bi-level selection such that the
-linear and nonlinear spaces of smoothing functions can be selected
-separately. The prior setup encourages a flexible solution, rendering
-one of three possibilities for each predictor: no effect, only linear
-effect, or linear and nonlinear effects. In addition, two scalable
-optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm
-and EM-Iterative weighted least square (EM-IWLS) algorithm, are
-developed and implemented in an publicly available R package
-\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}, making
-translational science more accessible.
-
-The proposed framework, BHAM, differs from previous spike-and-slab based
-GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
-spline prior is a spike-and-slab lasso type prior using independent
-mixture double exponential distribution, compared to spike-and-slab GAM
-that uses normal-mixture-of-inverse gamma prior. Spike-and-slab lasso
-priors provide computational convenience during model fitting by using
-optimization algorithms instead of intensive sampling algorithms. They
-make fitting high-dimensional models more feasible without sacrificing
-performance in prediction and variable selection. Secondly, SB-GAM uses
-a group spike-and-slab lasso prior with an EM-CD algorithm to fit the
-model. While both methods use the combination of expectation
-maximization algorithm and coordient descent algorithm, there are subtle
-difference in the implementation due to the difference in prior
-specification. The proposed model sets up independent priors among basis
-function coefficients after the re-parameterization step, which provides
-some advantage in computation. In addition, we offer an alternative
-algorithm, EM-IWLS, that can provide variance-covariance matrix of the
-coefficients. Last but not least, the proposed model addresses the
-incapability of bi-level selection in SB-GAM.
-
-In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical
-generalized additive model, introduce the proposed spike-and-slab spline
-priors, and describe the two fast-fitting algorithms. In Section
-\ref{sec:sim}, we compare the proposed framework to state-of-the-art
-models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation
-studies. Analyses of two metabolomics datasets are presented in Section
-\ref{sec:real_data}. Conclusion and discussions are given in Section
-\ref{sec:concl}.
-
-\section{Bayesian Hierarchical Additive Models (BHAMs)}
-\label{sec:BHAM}
-
-Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model with link function \(g(\cdot)\) and
-linear predictor \begin{equation}\label{eq:gam}
-\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
-distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
-and dispersion parameter \(\phi\), and the data distribution is
-\begin{equation}
-f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} The basis function matrix, i.e.~the design matrix derived
-from the smoothing function \(B_j(x_j)\), is denoted
-\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
-design matrix depends on the choice of the smoothing function, and is
-denoted as \(K_j\) for \(x_j\). Its corresponding smoothing penalty
-matrix is denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\)
-denotes the basis function coefficients for the \(j\)th variable such
-that \(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
-abuse of notation, we denote vectors and matrices in bold fonts
-\(\boldsymbol{\beta}, \boldsymbol{X}\) with conformable dimensions,
-where scalar and random variables are denoted in unbold fonts
-\(\beta, X\). The matrix transposing operation is denoted with a
-superscript \(^T\). To note, the proposed model can include parametric
-forms of variables in the model, treating the parametric function as a
-special case of the smoothing function, e.g.~\(B_j(x_j) = x_j\) with the
-smoothing penalty matrix defined as
-\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
-
-To encourage proper smoothing of additive functions, we adopt the idea
-of smoothing penalties from smoothing spline models. The basic idea is
-to set up a smoothing penalty, described in Equation
-(\ref{eq:smoothpen}), in the prior density function. However, the direct
-integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and Wood \citep{Marra2011} proposed a re-parameterization
-procedure to accommodate the smoothing penalty implicitly. Given the
-smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
-semi-definite for univariate smoothing functions, we apply
-eigendecomposition on the penalty matrix
-\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
-where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
-arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
-elements of zeros on the diagonal, where the zeros are associated with
-the linear space of the smoothing function. For the most popular
-smoothing function, cubic splines, the dimension of the linear space is
-one. Hereafter, we focus on discussing a uni-dimensional linear space
-for simplicity; however, it generalizes easily to the cases where the
-linear space is multidimensional. We further write the orthonormal
-matrix
-\(\boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
-containing the eigenvectors as columns in the corresponding order to
-\(\boldsymbol{D}\). That is, \(\boldsymbol{U}\) contains the
-eigenvectors \(U^0\) with zero eigenvalues for the linear space and
-\(\boldsymbol{U}^{*}\) contains the eigenvectors (as columns) for the
-non-zero eigenvalues, i.e.~the non-linear space. We multiply the basis
-function matrix \(\boldsymbol{X}\) with the orthonormal matrix
-\(\boldsymbol{U}\) for the new design matrix
-\({\boldsymbol{X}}^\text{repa}= \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} X^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
-An additional scaling step is imposed on \(\boldsymbol{X}^{*}\) by the
-non-zero eigenvalues of \(\boldsymbol{D}\) such that the new basis
-function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
-each of its dimensions. With slight abuse of the notation, we drop the
-superscript \(^\text{repa}\) and denote
-\(\boldsymbol{X}_j \equiv \begin{bmatrix} X_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
-as the basis function matrix for the \(j\)th variable after the
-re-parameterization. A spline function can be expressed in the matrix
-form \[
-B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \beta_j X^0_j + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
-\] and the generalized additive model in Equation (\ref{eq:gam}) now is
-\begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
-\end{equation} where the coefficients
-\(\boldsymbol{\beta}_j \equiv \begin{bmatrix} \beta_j : \boldsymbol{\beta}^*_j \end{bmatrix}\)
-is an augmentation of the coefficient scalar \(\beta_j\) of linear space
-and the coefficient vector \(\boldsymbol{\beta}^*_j\) of non-linear
-space.
-
-The re-parameterization step sets up the foundation of the proposed
-model and provides three benefits. First of all, the re-parameterization
-integrates the smoothing penalty into the design matrix, and encourages
-models to properly smooth the nonlinear function in addition to the
-sparse penalty for functional selection. Secondly, the
-eigendecomposition of the smoothing penalty allows the isolation of the
-linear space from the nonlinear space, improving the feasibility of
-bi-level functional selection. The eigendecomposition facilitates the
-construction of orthonormal design matrix, which makes imposing
-independent priors on the coefficients possible. This reduces the
-computational complexity compared to using a multivariate priors, and
-greatly broadens the choices of priors and further model choices. Last
-but not least, the scaling step of \(\boldsymbol{X}^*\) further
-simplifies the choice of priors, from column-specific priors to a
-unified prior.
-
-\subsection{Spike-and-Slab Spline Priors}
-
-The family of spike-and-slab (SS) regression models is one of most
-commonly used models in high-dimensional data analysis for its utility
-in outcome prediction and variable selection. We defer to Bai et al.
-\citep{Bai2021Review} for an in-depth introduction to spike-and-slab
-priors. To summarize, spike-and-slab priors are a family of mixture
-distributions that comprises a skinny spike density
-\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
-\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-\] The most distinct feature of SS priors is that it is conditioned on a
-latent binary variable \(\gamma \in \{0,1\}\) that indicates whether the
-variable \(x\) is included in the model. There are various
-spike-and-slab priors depending on the choice for the spike density
-\(f_{\text{spike}}(\cdot)\) and the slab density
-\(f_{\text{slab}}(\cdot)\), see George and McCulloch
-\citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
-variables; Brown et al. \citep{Brown1998} for multivariate outcomes;
-Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
-\citep{Clyde2004} and reference therein. Two of most popular spike and
-slab priors are spike-and-slab normal prior \citep{George1993} and
-spike-and-slab double exponential prior\citep{Rockova2018}.
-
-The major criticism of early spike-and-slab models is being
-computationally prohibitive. \citep{Bai2021Review} Since then, many
-studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS based on spike-and-slab normal
-prior\citep{Rockova2014a} and spike-and-slab Lasso (SSL)
-\citep{Rockova2018b, Rockova2018}. Particularly, the development of SSL
-model substantially improves the scalability of SS models, setting up
-the theoretical foundation for generalized models in -omics data
-analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The SSL prior
-is composed of two double exponential distributions with mean 0 and
-different dispersion parameters, \(0 < s_0 < s_1\), mathematically,
-\begin{equation} 
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation} Given that both double exponential distributions have a
-mean of 0 and the latent indicator \(\gamma\) can only take the value of
-0 or 1, the mixture double exponential distribution can be formulated as
-one single double exponential density, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
-\end{equation} with the scale parameter
-\(S = (1-\gamma)s_0 + \gamma s_1\). The SSL also mitigates the problem
-of EMVS where the weak signals are not shrink to zero, and hence is
-preferred in high-dimensional data analysis. We notice that Bai
-\citep{Bai2021} is the first to apply spike-and-slab lasso prior in the
-GAM framework, where the densities of the spike and slab components take
-the group lasso density \citep{Xu2015} and limits to an
-``all-in-all-out'' strategy for functional selection.
-
-\subsubsection{Two-part Spike-and-Slab Lasso Prior}
-
-We introduce a novel prior for GAMs, particularly for high-dimensional
-nonlinear modeling with bi-level selection. The proposed prior extends
-from the spike-and-slab lasso prior described in Equation
-(\ref{eq:ssl}). Given the re-parameterized design matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
-for the \(j\)th variable, we impose a two-part SSL prior to the
-coefficients
-\(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
-Specifically, we impose independent group priors on the linear space
-coefficients and on the nonlinear space coefficients respectively,
-\begin{align}\label{eq:bham_ssl}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j
-\end{align} where \(\gamma_{j}\in\{0,1\}\) and
-\(\gamma^*_{j}\in \{0,1\}\) are two latent indicator variables,
-indicating if the model includes the linear effect and the nonlinear
-effect of the \(j\)th variable respectively. \(s_0\) and \(s_1\) are
-scale parameters, assuming \(0 < s_0 < s_1\) and given. These scale
-parameters \(s_0\) and \(s_1\) can be treated as tuning parameters and
-optimized via cross-validation. A discussion of how to choose the scale
-parameters comes in Section \ref{sec:tune}. To note, this prior differs
-from previous group SSL priors \citep{Tang2018, Tang2019}, as the
-\(\beta_j\) and \(\boldsymbol{\beta}^*_j\) have different indicator
-variables \(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to
-add a more restrictive assumption on the priors, assuming that one
-indicator variable decides the inclusion of both the linear effect and
-nonlinear effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the
-SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
-assumption such that each coefficient
-\(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has its own latent indicator
-\(\gamma_{jk}\), but at the cost of complicating the bi-level functional
-selection. This reduces the proposed prior to the classic SSL prior.
-
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the
-validity of the proposed prior. First of all, the smoothing function
-bases are linear dependent and necessitate extra attention. The
-eigeondecomposition remedies the problem and hence our prior can be set
-to be conditionally independent. Secondly, the eigenvalue scaling
-provides a panacea to allow unified scale parameters for all bases of
-all smoothing functions.
-
-The rest of the hierarchical prior follows the traditional SSL prior: we
-set up hyper-priors on \(\gamma_j, \gamma^*_j\) to allow local adaption
-of the shrinkage using a Bernoulli distribution, written as binomial
-distribution of one trial. The two indicators of the \(j\)th predictor,
-\(\gamma_j\) and \(\gamma^*_j\), shares the same probability parameter
-\(\theta_j\), \[
-\begin{aligned}
-&\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^*| \textcolor{red}{\gamma_{j}, } \theta_j \sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j).
-\end{aligned}
-\]
-\textcolor{red}{Equivalently, we can derive the marginal distribution of $\gamma_{j}^*$ by integrate out $\gamma_j$ (Supplement 1), 
-$$
-\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2).
-$$
-}
-
-This is to leverage the fact that the probability of selecting the bases
-of a smoothing function should be similar, while allowing different
-penalty on the linear space and non-linear space of the smoothing
-function. The hyper prior of \(\gamma_{j}\) decides the sparsity of the
-model at the functional selection level, while that of \(\gamma_{j}^*\)
-decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that \(\gamma_{j}\) and \(\gamma_{j}^*\) are
-independently distributed for analytic simplicity. We further specify
-the parameter \(\theta_j\) follows a beta distribution with given shape
-parameters \(a\) and \(b\), \[
-\theta_j \sim Beta(a, b).
-\] The beta distribution is a conjugate prior for the binomial
-distribution and hence provides some computation convenience. For
-simplicity, we focus on a special case of beta distribution, uniform
-(0,1), i.e.~\(a = 1, b = 1\). When the variable have large effects in
-any of the bases, the parameter \(\theta_j\) will be estimated large,
-which in turn encourages the model to include the rest of bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models
-with the spike-and-slab spline prior as the BHAM, and visually presented
-in Figure \ref{fig:SSprior}.
-
-\begin{center}
-Figure \ref{fig:SSprior} here
-\end{center}
-
-\subsubsection{Other Priors}
-
-With the re-parameterization step of the basis function matrix
-\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
-priors, for example normal priors for ridge-type regularization and
-mixture normal prior for spike-and-slab regularization. These priors
-would work better in low and medium dimensional settings where the
-sparse assumption is not necessary. Here we elaborate the mixture normal
-prior as a demonstration of applying continuous spike-and-slab prior in
-BHAM.
-
-A spike-and-slab mixture normal spline prior can be expressed as
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
-\end{align*} Similar to the spike-and-slab spline prior in Equation
-(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
-optimized via cross-validation. One of the critics received by the
-spike-and-slab mixture normal prior is that the tails of a normal
-distribution diminishes to zero too fast, which causes problems when
-estimating the large effects. Distributions with heavier tails can be
-used as an alternative, for example mixture Student's \(t\) distribution
-with small degree of freedom.
-
-\subsection{Algorithms for Fitting BHAMs}
-
-The proposed models can be fitted with MCMC algorithms. Nevertheless,
-the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch\citep{George1997} examined the computation speed
-for various MCMC algorithms with spike-and-slab mixture normal priors,
-and suggested MCMC algorithms works well for medium size (\(p\)=25) of
-predictors with only linear effects. However, it is not feasible for
-high-dimensional data analysis where the number of predictors easily
-exceeds 100. Specific to additive models, each predictor would expand to
-multiple new ``predictors'' via basis functions, creating greater
-computational demands. Scheipl et al. \citep{Scheipl2013} demonstrated
-the computational demands of a MCMC algorithm for fitting spike-and-slab
-GAM grow exponentially as \(p\) increases modestly via simulation
-studies. Hence, we feel compelled to develop scalable algorithms for
-fitting Bayesian hierarchical additive models in high-dimensional
-settings.
-
-As an alternative to sampling algorithms for fitting Bayesian models,
-optimization algorithms focus on the maximum a posteriori (MAP)
-estimates and speed up the model fitting process at the cost of
-posterior inference. The earlier work for fitting SS models using
-optimization algorithms includes EMVS \citep{Rockova2014a}. Rockova and
-George\citep{Rockova2014a} proposed an expectation-maximization (EM)
-based algorithm to fit models that use continuous mixture normal priors.
-In the E step, the latent binary indicators \(\gamma\)s are treated as
-the missing data, and the posterior means are calculated conditioning on
-the current value of other parameters; in the M step, a ridge estimator
-was used to update the coefficients, followed by updating
-\(\phi, \theta\). The same authors \citep{Rockova2018b, Rockova2018}
-further combined the EM algorithm with coordinate descent algorithm to
-fit SSL models. Yi and his group independently developed the
-EM-Iterative Weighted Least Square and EM-cyclic coordinate descent
-algorithms to fit models with broader class of priors, SSL
-included.\citep{Yi2019} These algorithms were implemented for
-generalized linear models\citep{Tang2017a}, Cox proportional hazards
-models \citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a popular property for
-reproducible research
-
-In this section, we extend the two EM-based algorithms, EM-CD and
-EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
-provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM with an expedited performance, recommending to use in high
-and ultra-high dimensional setting. A specific concern of EM-CD
-algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can estimate the variance-covariance matrix of the
-coefficients. Moreover, the EM-IWLS is a more general model fitting
-algorithm that can be used for fitting not only SSL and continuous SS
-priors but also Student's t-priors and double exponential priors. To
-note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
-main difference between proposed EM-CD algorithm and that in SB-GAM is
-that SB-GAM uses a block CD algorithm for their group prior, while the
-proposed prior is pairwise independent requiring no special treatment in
-the CD algorithm.
-
-\subsubsection{EM algorithms}
-
-EM algorithm is an iterative algorithm to find MAP estimates or the
-maximum likelihood estimates. It is commonly used when some necessary
-data to establish the likelihood function are missing. Instead of
-maximizing the the likelihood function, the algorithm maximizes the
-expectation of the likelihood function with respect to the ``missing''
-data.
-
-The recursive algorithm consists of two steps:
-
-\begin{itemize}
-\tightlist
-\item
-  E-step: to calculate the expectation of the posterior density function
-  with respect to some ``missing'' data
-\item
-  M-step: to maximize the expectation derived in the E-step and update
-  parameters of interest
-\end{itemize}
-
-For BHAMs, we define the parameters of interest as
-\(\Theta = \{\boldsymbol{\beta}, \boldsymbol{\theta}, \phi\}\) and
-consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model, in other words the ``missing'' data.
-Our objective is to find the parameters \(\Theta\) that maximize the
-posterior density function, or equivalently, the logarithm of the
-density function, \[
-\begin{aligned}
-& \text{argmax}_{\Theta}
-\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{j})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j),
-\end{aligned}
-\]\\
-where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
-non-informative prior for the intercept \(\beta_0\) and the dispersion
-parameter \(\phi\); for example, \(f(\beta_0|\tau_0^2)=N(0,\tau_0^2)\)
-with \(\tau^2_0\) set to a large value and \(f(\log \phi) \propto 1\).
-
-We use the EM algorithm to find the MAP estimate of \(\Theta\). This is,
-in the E-step, we calculate the expectation of posterior density
-function of
-\(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
-respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
-the values from previous iteration \(\Theta^{(t-1)}\), \[
-E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
-\] Hereafter, we use the shorthand notation
-\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}(\cdot)\). In the
-M-step, we find the \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The E-
-and M- steps are iterated until the algorithm converge.
-
-To note here, the log-posterior density of BHAMs (up to additive
-constants) can be written as a two-part equation
-\[ \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
-where
-\[ Q_1\equiv Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
-and
-\[Q_2 \equiv Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j).\]
-\(Q_1\) and \(Q_2\) are respectively the log posterior density of the
-coefficients \(\boldsymbol{\beta}\) and the log posterior density of the
-probability parameters \(\boldsymbol{\theta}\) conditioning on
-\(\boldsymbol{\gamma}\). Meanwhile, conditioning on
-\(\boldsymbol{\gamma}\), \(Q_1\) and \(Q_2\) are independent and can be
-maximized separately for \(\boldsymbol{\beta}, \phi\) and
-\(\boldsymbol{\theta}\). Depending on the choice of coefficient priors,
-\(Q_1\) can be treated as penalized likelihood function and maximization
-of \(E(Q_1)\) can be solved via CD algorithm or IWLS algorithm in each
-iteration. Maximization of \(E(Q_2)\) can be solved via closed form
-equations following the beta-binomial conjugate relationship.
-
-\subsubsection{EM-Coordinate Descent}
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double
-exponential, coordinate descent algorithm can be used to estimate the
-parameters in the M-step. Coordinate descent is an optimization
-algorithm that offers extreme computational advantage, and famous for
-its application in optimizing the \(l_1\) penalized likelihood function.
-
-The density function of spike-and-slab mixture double exponential prior
-can be written as \[
-f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
-\] and \(E(Q_1)\) can be expressed as a log-likelihood function with
-\(l_1\) penalty \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
-\end{equation} where
-\(S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
-unknown quantities \(E({S_j}^{-1})\) and \(E({S^*}^{-1}_j)\), the
-posterior probability
-\(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
-\(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
-necessary, which can be derived via Bayes' theorem. The calculation of
-\(p_j^*\) is slightly different from that of \(p_j\), as \(p_j^*\)
-depends on the value of the vector \(\boldsymbol{\beta}^*_{j}\) and
-\(p_j\) only depends on the scalar \(\beta_j\). The calculation follows
-the equations below. \begin{align*}
-p_{j} &= \frac{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \text{Pr}(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
-p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{align*} and \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j}\),
-\(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
-\(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
-show \begin{align}\label{eq:exp_scale}
-&E(\gamma_{j})  = p_{j} & &E(\gamma^{*}_{j}) = p_{j}^{*}\nonumber\\
-&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^*}^{-1}_{j}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
-\end{align} After replacing withe calculated quantities, \(E(Q_1)\) can
-be seen as a \(l_1\) penalized likelihood function with the
-regularization parameter \(\lambda = E(S^{-1})\), and hence be optimized
-via coordinate descent algorithm \citep{Friedman2010}. Independently,
-the remaining parameters of interest \(\boldsymbol{\theta}\) can be
-updated by maximizing \(E(Q_2)\). As the beta distribution is a
-conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
-be easily updated with a closed form equation,
-\begin{equation}\label{eq:update_theta}
-\theta_j = \frac{p_j + p^*_{j} + a - 1 }{a + b}.
-\end{equation}
-
-Totally, the proposed EM-coordinate descent algorithm is summarized as
-follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
-  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
-  \(\boldsymbol{\theta}\). For example, we can initialize
-  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
-  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
-\item
-  Iterate over the E-step and M-step until convergence
-
-  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
-  \(E({S}^{-1}_{j})\), \(E({S^*}^{-1}_{j})\) with estimates of
-  \(\Theta^{(t-1)}\) from previous iteration
-
-  M-step:
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii})}
-  \item
-    Update \(\boldsymbol{\beta}^{(t)}\), and the dispersion parameter
-    \(\phi^{(t)}\) if exists, using the coordinate descent algorithm
-    with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
-  \item
-    Update \(\boldsymbol{\theta}^{(t)}\) using Equation
-    (\ref{eq:update_theta})
-  \end{enumerate}
-\end{enumerate}
-
-We assess convergence by the criterion:
-\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
-\(d^{(t)} = -2\log f(\textbf{y}| \textbf{X}, \boldsymbol{\beta}^{(t)},\phi^{(t)})\)
-is the estimate of deviance at the \(t\)th iteration, and \(\epsilon\)
-is a small value (say \(10^{-5}\)).
-
-\subsection{Selecting Optimal Scale Values}
-\label{sec:tune}
-
-Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
-\(s_1\)). Hence, we need to find the optimal values for the scale
-parameters such that the model reaches its best prediction performance
-regarding a criteria of preference. This would be achieved by
-constructing a two dimensional grid, consists of different pairs of
-(\(s_0\), \(s_1\)) value. However, previous research suggested the value
-of slab scale \(s_1\) have less impact on the final model and is
-recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
-provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
-on examining different values of spike scale \(s_0\). Instead of the 2-D
-grid, We consider a sequence of \(L\) decreasing values
-\(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
-spike scale \(s_0\) tends to include more non-zero coefficients in the
-model. A measure of preference calculated with cross-validations (CV),
-e.g.~deviance, area under the curve (AUC), mean squared error, can be
-used to facilitate the selection of a final model. The procedure is
-similar to the Lasso implemented in the widely used R package
-\texttt{glmnet}, which quickly fits Lasso models over a list of values
-of regularization parameters \(\lambda\) , giving a sequence of models
-for users to choose from.
-
-\section{Simulation Study}
-\label{sec:sim}
-
-In this section, we compare the performance of the proposed models to
-four alternative models: component selection and smoothing operator
-(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
-generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM \citep{Bai2021}\textcolor{red}{}, linear lasso model and
-spikeSlabGAM {[}TODO: add citation here{]}\}. COSSO is one of the
-earliest smoothing spline models that consider sparsity-smoothness
-penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for
-penalties such that the penalty of each functional component are
-different for extra flexibility. Generalized additive models with
-automatic smoothing, hereafter \textit{mgcv}, is one of the most popular
-models for nonlinear effect interpolation and prediction. SB-GAM is the
-first spike-and-slab lasso GAM. We implement COSSO and adaptive COSSO
-with R package \texttt{cosso 2.1-1}, generalized additive models with
-automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R
-package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence use the default
-choices. Both mgcv and proposed models allow customized smoothing
-functions and we choose the cubic regression spline. We controll the
-dimensionality of each smoothing function, 10 bases, for all different
-choices of smoothing functions. We use 5-fold CV with the default
-selection criteria to select the final model for COSSO models, SB-GAM
-and the proposed models. 20 default candidates of tuning parameters
-(\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for SB-GAM and
-the proposed models which allow user-specification of tuning candidates.
-All computation was conducted on a high-performance 64-bit Linux
-platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
-processors and 24G of RAM per core and R 3.6.2 \citep{R}.
-
-Other related methods for high-dimensional GAMs also exist, notably the
-methods of sparse additive models by Ravikumar et al.
-\citep{Ravikumar2009} and stochastic search term selection for GAM
-\citep{Scheipl2012}. However, we exclude these methods from current
-simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increased number of
-predictors. \citep{Scheipl2013}
-
-\subsection{Monte Carlo Simulation Study}
-
-We follow the data generating process described in Bai \citep{Bai2021}.
-We first generate \(n=500\) training data points with
-\(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
-\(X\) are simulated from a multivariate normal distribution
-\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(y\)
-from two distributions, Gaussian and binomial with the identity link and
-logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The mean of each
-outcome were simulated via the following function \[
-\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for Gaussian and binomial outcomes. Gaussian outcomes requires
-specification of dispersion, where we set the dispersion parameter to be
-1. In this data generating process, we have \(x_1, x_2, x_3, x_4\) as
-the active covariates, while the rest covariates are inactive,
-i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
-independent sample of size \(n_{test}=1000\), are created following the
-same data generating process, serving as the testing data. We generate
-50 independent pairs of training and testing datasets to evaluate the
-prediction performance of the chosen models, where training datasets are
-used to fit the models and testing datasets used to calculate assessment
-measures.
-\textcolor{red}{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
-These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
-
-To evaluate the predictive performance of the models, the statistics,
-\(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing dataset, are averaged across 50 simulations.
-\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
-For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-
-\textcolor{red}{\subsection{Simulation Results}}
-\textcolor{red}{\subsubsection{Prediction Performance}}
-
-The predictive performances have a consistent pattern across the two
-distributions of outcomes. Across all the scenarios, COSSO and adaptive
-COSSO have the least favorable performance among the applicable methods
-examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}). To note, mgcv
-doesn't support high-dimensional analysis, i.e.~the number of
-coefficients are greater than the sample size, and hence not evaluated
-when \(p=100, 200\). mgcv predicts well when \(p\) is small or moderate
-(\(p = 4, 10\)), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the
-proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when
-the dimension of are moderate (p=4, 10, 50). Particularly, BHAM-IWLS
-performs as good as mgcv if not better. However, in high dimensional
-case where we mimic the situation the signals are extremely sparse,
-SB-GAM has better performance than the proposed method. However, the
-BHAM-CD has extreme computational advantage over SB-GAM (see Table
-\ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
-
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
-
-\section{Metabolomics Data Analysis}
-\label{sec:real_data}
-
-In this section, we apply the proposed models BHAM, fitted with the
-EM-CD algorithm, to analyze two real-world metabolomics datasets where
-the outcomes are binary and continuous respectively. We demonstrate the
-improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}, while being
-computationally efficient (see Table \ref{tab:time_real_data}).
-
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
-
-\subsection{Emory Cardiovascular Biobank}
-\label{sec:ECB}
-
-We use the proposed models BHAM to analyze a metabolic dataset from a
-recently published research \citep{Mehta2020} studying plasma
-metabolomic profile on the three-year all-cause mortality among patients
-undergoing cardiac catheterization. The dataset is publicly available
-via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
-subjects from two cohorts. As there is a large number of non-overlapping
-features among the two cohorts, we use the cohort with larger sample
-size (N=454). There are initially 6796 features in the dataset, which is
-too large to be practically meaningful to analyze. Hence, we perform a
-univariate screening procedure on the features, via GAM implemented in
-\texttt{mgcv}, and choose the the top 200 features with smallest
-p-values. We use 5-knot spline additive models for binary outcome using
-two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are
-used to choose the optimal tuning parameters of each framework with
-respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where
-deviance, AUC, Brier score defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassication error defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
-calculated. BHAM-CD obtains superior AUC, Brier score, and
-misclassification error in the out-of-bag samples compared to SB-GAM
-(see Table \ref{tab:ECB_res}).
-
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
-
-\subsection{Weight Loss Maintenance Cohort}
-
-We use the proposed models BHAM to analyze metabolomics data from a
-recently published study \citep{Bihlmeyer2021} on the association
-between metabolic biomarkers and weight loss, where the dataset is
-publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
-primarily focuse on the analysis of one of the three studies included,
-weight loss maintenance cohort \citep{Svetkey2008}, due to the
-drastically different intervention effects. In the dataset, 765
-metaboliltes in baseline plasma collected were profiled using liquid
-chromatography mass spectrometry. Quality control and natural log
-transformation are performed during metabolites data preparation. The
-outcome of interest are standardized percent change in insulin
-resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there are
-\(p\)=237 features remaining in the analysis. 5-Knot spline additive
-models for the Gaussian outcome are constructed using two different
-models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
-the optimal tuning parameters of each framework with respect to the
-default selection criterion implemented in the software. Out-of-bag
-samples are used for prediction performance evaluation, where deviance,
-\(R^2\), mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-absolute error (MAE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
-BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
-compared to SB-GAM (see Table \ref{tab:WLM_res}).
-
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}
-
-\section{Discussion}
-\label{sec:concl}
-
-In the paper, we described a novel generalized additive model using
-Bayesian hierarchical priors, particularly the proposed spike-and-slab
-spline prior for bi-level functional selection. Meanwhile, we introduced
-two optimization based algorithms for model fitting. The algorithms can
-be easily scale up to address high-dimensional data analysis in a
-computational efficient manner. Via simulations, we demonstrated that
-the proposed model provides as good, if not better, prediction
-performance compared to some state-of-the-art non-linear modeling
-devices.
-
-The proposed model shares many commonality with an high-dimensional
-Bayesian GAM, SB-GAM \citep{Bai2021}, independently developed around the
-same time of this work. Both frameworks emphasize computational
-efficiency by deploying group spike-and-slab lasso type of priors and
-optimization-based fast and scalable algorithms. Bai provided the
-theoretical proof for the consistency of variable selection using group
-spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the
-bi-level functional selection and the model inference. The proposed
-model provides solutions to these remaining questions while maintaining
-the same level of prediction accuracy. Moreover, the proposed model
-renders additional flexibility to model specification, allowing various
-choices of smoothing functions and degrees of freedom.
-
-For translational science purpose, we implemented the proposed model in
-a R package \texttt{BHAM}, deposited at
-\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
-smoothing function specification, we deploy the same programming grammar
-as in the state-of-the-art package \texttt{mgcv}, in contrast to
-previous tools where smoothing functions are limited to the default
-ones. Ancillary functions are provided for model specification in
-high-dimensional settings, curve plotting and functional selection. In
-addition, in \texttt{BHAM}, we streamline the model fitting algorithms
-to support other popular Bayesian hierarchical prior for the smoothing
-functions, such as Student's T distribution, double exponential
-distribution, mixture normal and T distribution's. These priors could be
-helpful when the sparse assumption is weak or not necessary.,
-
-There are some improvements possible for the proposed models. First of
-all, the proposed model achieves a bi-level selection via the two-part
-spike-and-slab spline prior. Nevertheless, this set-up could result in a
-situation that is not theoretically sound: the non-linear component is
-selected, but the linear component is not. We currently address it
-analytically by including the linear component in the model when
-non-linear component is selected. Another possible solution is to impose
-a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, the
-computational time for fitting a BHAM model with EM-IWLS algorithms can
-be improved. Current implementation of the EM-IWLS algorithm jointly
-updates all coefficients in each iteration, which requires a lot
-computation resources. This jointly updating procedure can be enhanced
-by adapting a backfitting step \citep{Hastie1987} where each smoothing
-function are updated individually. Thirdly, when using the proposed
-model in real data analysis where a screening procedure is implemented
-before joint predictive modeling. we recommend to include the screening
-procedure in the cross-validation during model selection.
-\citep{Friedman2017} For the sake of model comparison, we fit the models
-using the same pool of predictors in Section \ref{sec:ECB}.
-
-Our future efforts direct to modeling survival outcomes and integrative
-analysis. While the proposed model addresses a great deal of analytic
-problem, analyzing the time-to-event outcome remains unsolved. An naive
-approach would be convert a time-to-event outcome to a Poisson outcome
-following Whitehead \citep{Whitehead1980}. However, it would be more
-efficient to directly fit Cox models via penalized pseudo likelihood
-function \citep{Simon2011}. Meanwhile, with growing understanding of
-biological structure within -omics field, it is appealing to integrate
-external biology information in the modeling process. The main
-motivation for integrative models is that biologically informed grouping
-of weak effects increases the power of detecting true associations
-between features and the outcome \citep{Peterson2016}, and stabilizes
-the analysis results for reproducibility purpose. Such integration can
-be achieved by setting up a structural hyperprior on the inclusion
-indicator of the smoothing function null space
-\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
-and Dunson \citep{Ferrari2020}.
-
-\clearpage
-
-\bibliography{bibfile.bib}
-
-\clearpage
-
-\begin{figure}
-\centering
-\begin{tikzpicture} [
-staticCompo/.style = {rectangle, minimum width=1cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-outCome/.style={ellipse, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30},
-mymatrix/.style={matrix of nodes, nodes=outCome, row sep=1em},
-PriorBoarder/.style={rectangle, minimum width=5cm, minimum height=10cm, text centered, fill=lightgray!30},
-background/.style={rectangle, fill=gray!10,inner sep=0.2cm, rounded corners=5mm}
-]
-
-\matrix (linearPrior) [matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (linearGamma) [outCome] { $\gamma_j \sim Bin(1, \theta_j) $ };\\
-  \node (linearBeta) [outCome] { $\beta_j \sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)$};\\
-};
-\matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
-};
-
-\node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
-\node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
-\node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
-
-\draw[->] (Theta) -- (linearGamma);
-\draw[->] (Theta) -- (penGamma);
-\draw[->] (linearGamma) -- (linearBeta) ;
-\draw[->] (penGamma) -- (penBeta);
-\draw[->, draw = red] (linearGamma) -- (penGamma);
-\draw[->] (ab) -- (Theta);
-\draw[->] (s) -- (linearBeta) ;
-\draw[->] (s) -- (penBeta);
-\draw[->] (linearBeta) -- (Beta);
-\draw[->] (penBeta) -- (Beta);
-\draw[->] (Beta) --  (Y);
-
-\begin{pgfonlayer}{background}
-  \node [background,
-   fit=(linearGamma) (linearBeta),
-   label=above:Linear Space:] {};
-  \node [background,
-    fit=(penGamma) (penBeta),
-    label=above:Nonlinear Space:] {};
-\end{pgfonlayer}
-
-\end{tikzpicture}
-
-\caption{Directed acyclic graph of the proposed Bayesian hierarchical additive model with parameter expansion. Elliposes are stochastic nodes, rectangles and are deterministic nodes. }
-\label{fig:SSprior}
-\end{figure}
-
-\clearpage
-
 \begin{table}[ht]
 \centering
-\begin{tabular}{ccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
+P & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
-  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
-  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
+  4 & 0.90 (0.01) & 0.33 (0.01) & 0.71 (0.13) & 0.72 (0.11) & 0.90 (0.01) & 0.79 (0.04) & 0.80 (0.00) \\ 
+   10 & 0.90 (0.01) & 0.33 (0.01) & 0.66 (0.21) & 0.77 (0.02) & 0.89 (0.01) & 0.79 (0.04) & 0.79 (0.00) \\ 
+   50 & 0.86 (0.02) & 0.32 (0.01) & 0.46 (0.19) & 0.57 (0.18) & 0.80 (0.02) & 0.78 (0.05) & 0.78 (0.01) \\ 
+  100 & NaN (NA) & 0.32 (0.01) & 0.41 (0.23) & 0.48 (0.25) & 0.79 (0.01) & 0.79 (0.05) & 0.77 (0.01) \\ 
+  200 & NaN (NA) & 0.32 (0.01) & 0.39 (0.19) & 0.40 (0.17) & 0.79 (0.01) & 0.78 (0.04) & 0.75 (0.01) \\ 
    \hline
 \end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
+\caption{The average and standard deviation of the out-of-sample $R^2$ measure for
+    Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
+    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
+    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive
+    COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
+    whe number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:gaus}
 \end{table}
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:bin_auc}
-\end{table}
-
-\clearpage
-
-\begin{sidewaystable}[!h]
-\centering
-\begin{tabular}{lrllllll}
-  \hline
-Distribution & P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-Binomial & 4.00 & 5.88 (1.01) & 6.96 (1.76) & 2.97 (0.82) & 4.60 (2.32) & 0.18 (0.04) & 343.24 (88.37) \\ 
-  Binomial & 10.00 & 14.01 (1.86) & 7.65 (1.25) & 7.95 (2.70) & 10.01 (3.56) & 3.51 (12.03) & 543.13 (133.51) \\ 
-  Binomial & 50.00 & 185.26 (17.44) & 53.83 (6.26) & 109.72 (24.31) & 128.13 (23.26) & 670.01 (151.28) & 1630.42 (193.61) \\ 
-  Binomial & 100.00 & 1106.93 (356.73) & 48.41 (5.84) & 715.15 (170.55) & 718.63 (193.53) & - & 2783.99 (235.31) \\ 
-  Binomial & 200.00 & 6923.62 (1911.59) & 99.98 (7.05) & 5572.51 (1295.18) & 4958.69 (1970.46) & - & 4780.10 (488.66) \\ 
-  Gaussian & 4.00 & 4.22 (1.51) & 2.51 (0.20) & 0.79 (0.11) & 0.67 (0.07) & 0.05 (0.00) & 36.02 (3.02) \\ 
-  Gaussian & 10.00 & 17.59 (5.77) & 33.32 (3.50) & 3.49 (0.70) & 3.47 (0.71) & 0.32 (0.34) & 72.93 (9.52) \\ 
-  Gaussian & 50.00 & 329.83 (44.57) & 334.53 (22.33) & 35.54 (9.15) & 35.44 (9.02) & 73.33 (71.24) & 373.42 (34.93) \\ 
-  Gaussian & 100.00 & 1706.19 (225.92) & 446.98 (79.31) & 150.58 (46.72) & 152.82 (45.07) & - & 684.05 (67.80) \\ 
-  Gaussian & 200.00 & 16379.27 (4825.56) & 303.73 (87.27) & 591.11 (147.88) & 569.92 (107.45) & - & 1314.12 (137.85) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:time_sim}
-\end{sidewaystable}
-
-\clearpage
-
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
-
-\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\textbackslash{}label\{tab:time\_real\_data\}
-}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{490.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\end{longtable}
-
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & AUC & Brier & Misclass \\ 
-  \hline
-BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
-\label{tab:ECB_res}
-\end{table}
-
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & $R^2$ & MSE & MAE \\ 
-  \hline
-BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
-  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
-\label{tab:WLM_res}
-\end{table}
-
-\end{document}

### 2022-02-22 R/Sim_Res/format_predict_tbls.R

@@ -1,0 +1,21 @@
+format_predict_tbls <- function( dat, caption, label){
+  dat %>% 
+    mutate(p = as.integer(p)) %>% 
+    select(
+      "P" = p,
+      "mgcv" = mgcv,
+      "LASSO" = lasso,
+      "COSSO" = cosso,
+      "Adaptive COSSO" = acosso,
+      # "BHAM-IWLS" = bglm_spline_de,
+      # "BHAM-CD" = bamlasso,
+      "BHAM" = bamlasso,      
+      "SB-GAM" = SB_GAM,
+      "spikeSlabGAM" = ssGAM
+    ) %>% 
+    xtable(
+      align = "ccccccccc",
+      caption = caption,
+      label = label ) %>% 
+    print(comment = FALSE, include.rownames = FALSE)
+}

### 2022-02-22 _targets.R

@@ -4,7 +4,7 @@ library(tarchetypes)
 library(dotenv)
 
 tar_option_set(
-  packages = c("tidyverse", "dplyr", "rticles", "gtsummary", "mgcv", #"here", 
+  packages = c("tidyverse", "dplyr", "rticles", "gtsummary", "mgcv", "gt", 
                "broom", "unglue", "knitr", "ggpubr", "xtable", "janitor", 
                "flextable", "BHAM", "BhGLM", "sparseGAM",  "glmnet", "yardstick",
                "rmarkdown"#, "cosso",
@@ -31,12 +31,23 @@ tar_plan(
   #** Binomial Outcome  ####
   #*** Prediction  ####
   sim_binom_tab = make_sim_main_table(sim_success_rate,
-                                  dist = "binomial",
-                                  measures = "auc"),
+                                      dist = "binomial",
+                                      measures = "auc"),
+  
+  sim_binom_tab_latex = format_predict_tbls(
+    sim_binom_tab,
+    caption = "The average and standard deviation of the out-of-sample area under the curve measures
+    for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
+    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
+    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO),
+    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
+    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
+    label = "tab:bin_auc")%>% 
+    cat(file = "Manuscript/Tabs/sim_binom_tab.tex"),
   
   #*** Variable Selection  ####
   sim_binom_var_select_raw = sim_var_select_raw(sim_success_rate,
-                                                       dist = "binomial"),
+                                                dist = "binomial"),
   
   # Table Presentation of Var Selection
   # make_sim_var_metric_raw(sim_binom_var_select_raw[[5]]) %>%
@@ -49,8 +60,8 @@ tar_plan(
   
   # plot_var_select(sim_binom_var_select_raw[[1]])
   sim_binom_comp_select_raw = sim_comp_select_raw(sim_success_rate,
-                                                dist = "binomial"),
-
[…]

### 2022-02-22 _targets_packages.R

@@ -6,6 +6,7 @@ library(dplyr)
 library(flextable)
 library(ggpubr)
 library(glmnet)
+library(gt)
 library(gtsummary)
 library(janitor)
 library(knitr)

### 2022-02-22 Manuscript/10-Supplementary.Rmd
Update Support Info
Add lnr sim tables to supplementary

@@
 knitr::opts_chunk$set(echo = TRUE)
+
 ```
@@
 \newcommand{\pr}{\text{Pr}}
+\newcommand{\bs}[1]{\boldsymbol{#1}}
 
@@
 
-\subsection{Supplementary Information 1: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}
+\subsection{Supplementary Information 2: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}
 
@@
 Similar to EM-CD, we assess convergence by the criterion, $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where $\epsilon$ is a small value (say $10^{-5}$).
+
+\subsection{Supplementary Information 3: Predictive Performance of Linear Simulations}
+\input{Manuscript/Tabs/sim_lnr_gaus_tab.tex}
+\clearpage
+\input{Manuscript/Tabs/sim_lnr_binom_tab.tex}

### 2022-02-22 Manuscript/10-Supplementary.tex
Add latex tables to the manuscript

@@
-\documentclass[]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\title{Supporting Infromation}
+\author{}
+\date{\vspace{-2.5em}}
+
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
   pdftitle={Supporting Infromation},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\title{Supporting Infromation}
-\author{}
-\date{\vspace{-2.5em}}
 
@@
 
-\subsection{Supplementary Information 1: Marginal Distribution of $\gamma_j^\tp$}
+\subsection{Supplementary Information 1: Marginal Distribution of $\gamma_j^*$}
 
 Given that
-\(\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)\)
+\(\gamma_{j}^*| \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)\)
 where \(\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\), we can derive the
-the marginal distribution of \(\gamma_{j}^\tp\) with the following
-manipulation.
-
-\begin{align*}
-& \pr(\gamma_j^\tp = 1 | \theta_j)  = \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 1 , \gamma_j = 0| \theta_j)\\
-= & \pr(\gamma_j^\tp = 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^\tp \text{ and }\gamma \text{.]}\\
-= & \pr(\gamma_j^\tp = 1| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j)\\
+the marginal distribution of \(\gamma_{j}^*\) with the following
+manipulation. \begin{align*}
+& \text{Pr}(\gamma_j^*= 1 | \theta_j)  = \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 1 , \gamma_j = 0| \theta_j)\\
+= & \text{Pr}(\gamma_j^*= 1, \gamma_j = 1 | \theta_j) + 0 \qquad \text{[hierarchical structure between }\gamma^*\text{ and }\gamma \text{.]}\\
+= & \text{Pr}(\gamma_j^*= 1| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j)\\
 = & \theta_j^2\\
-& \pr(\gamma_j^\tp = 0 | \theta_j)  = \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0 , \gamma_j = 0| \theta_j)\\
-= & \pr(\gamma_j^\tp = 0, \gamma_j = 1 | \theta_j) + \pr(\gamma_j^\tp = 0, \gamma_j = 0 | \theta_j)\\
-= & \pr(\gamma_j^\tp = 0| \gamma_j = 1,  \theta_j)\pr(\gamma_j = 1| \theta_j) + \pr(\gamma_j^\tp = 0| \gamma_j = 0,  \theta_j)\pr(\gamma_j = 0| \theta_j)\\
+& \text{Pr}(\gamma_j^*= 0 | \theta_j)  = \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0 , \gamma_j = 0| \theta_j)\\
+= & \text{Pr}(\gamma_j^*= 0, \gamma_j = 1 | \theta_j) + \text{Pr}(\gamma_j^*= 0, \gamma_j = 0 | \theta_j)\\
+= & \text{Pr}(\gamma_j^*= 0| \gamma_j = 1,  \theta_j)\text{Pr}(\gamma_j = 1| \theta_j) + \text{Pr}(\gamma_j^*= 0| \gamma_j = 0,  \theta_j)\text{Pr}(\gamma_j = 0| \theta_j)\\
 = & (1-\theta_j)\theta_j + 1(1-\theta_j) = 1-\theta_j^2
@@
 EM-based algorithm where the iterative weighted least squares algorithm
-is used to find the estimate of \(\bs \beta, \phi\) that maximizes
-\(E(Q_1)\). The iterative weighted least squares algorithm was
+is used to find the estimate of \(\boldsymbol{\beta}, \phi\) that
+maximizes \(E(Q_1)\). The iterative weighted least squares algorithm was
 originally proposed to fit the classical generalized linear models, and
@@
 as a hierarchical normal prior with unknown variance \(\tau^2\)
-integrated out:
-
-\begin{align*}
+integrated out: \begin{align*}
   \beta|\tau^2 &\sim N(0, \tau^2)\\
   \tau^2|S & \sim Gamma(1, 1/(2S^2)), 
-\end{align*}
-
-For the mixture double exponential priors, we can define the scale
-parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following Equation
-(\ref{eq:ssl}). The change in the prior formulation in turn leads to the
-change in the log posterior density function, as \(Q_1\) needs to
-account for the hyperprior of \(\tau^2\):
-
+\end{align*} For the mixture double exponential priors, we can define
+the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following
+Equation (\ref{eq:ssl}). The change in the prior formulation in turn
+leads to the change in the log posterior density function, as \(Q_1\)
+needs to account for the hyperprior of \(\tau^2\):
 \begin{equation}\label{eq:Q1_IWLS}
-Q_1(\bs \beta, \phi) = \log f(\textbf{y}|\bs \beta, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{\tp}_{jk}|{\tau^{\tp}}^2_{jk})+\log f({\tau^\tp}^2_{jk}| S^\tp_j)\}\right].
-\end{equation}
-
-Since \(\bs \tau^2\) are not of our primary interest, we treat them as
-the ``missing'' data in addition to the latent indicators
-\(\bs \gamma\), and hence construct the expectation
-\(E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)\) in the E-step. To
-note, unlike the same latent indicator \(\gamma^\tp_j\) which is shared
-by the coefficients of the non-linear terms \(\beta^\tp_{jk}\) for
-\(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is coefficient specific for
-\(\beta^\tp_{jk}\).
-\(E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})\)
+Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
+\end{equation} Since \(\boldsymbol{\tau}^2\) are not of our primary
+interest, we treat them as the ``missing'' data in addition to the
+latent indicators \(\boldsymbol{\gamma}\), and hence construct the
+expectation
+\(E_{\boldsymbol{\gamma}, \boldsymbol{\tau}^2|\Theta^{(t-1)}}(Q_1)\) in
+the E-step. To note, unlike the same latent indicator \(\gamma^*_j\)
+which is shared by the coefficients of the non-linear terms
+\(\beta^*_{jk}\) for \(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is
+coefficient specific for \(\beta^*_{jk}\).
+\(E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^*}^2_{jk}|S_j^*, \beta^*_{jk})\)
 needs to be calculated to formulate \(E(Q_1)\). As neither
 \(E({S_j}^{-1}|\beta_j, s_0, s_1)\) nor
-\(E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)\) depends on \(\tau^2\)s,
-they can be derived using Equation (\ref{eq:exp_scale}). On the other
-hand, \(\tau^{2}\), following gamma distributions, is a conjugate prior
-for the normal variance, and the conditional posterior density of
-\(\tau^{-2}\) is an inverse Gaussian distribution.
-\(E({\tau}^{-2}_{j})\) and \(E({\tau^\tp}^{-2}_{jk})\) are calculated
-using the closed form equation
-
-\begin{align*}
- E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^\tp}^{-2}_{jk}|S_j^\tp, \beta^\tp_{jk})={S_j^\tp}^{-1}/|\beta^\tp_{jk}|,
-\end{align*}
-
-where \(S_j\) and \(S_j^\tp\) are replaced by the expectation and
-\(\beta\)s are replaced with \(\beta^{(t-1)}\). With simplification (up
-to constant additive terms), we have
-
+\(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
+\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}). On
+the other hand, \(\tau^{2}\), following gamma distributions, is a
+conjugate prior for the normal variance, and the conditional posterior
+density of \(\tau^{-2}\) is an inverse Gaussian distribution.
+\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
+using the closed form equation \begin{align*}
+ E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
+\end{align*} where \(S_j\) and \(S_j^*\) are replaced by the expectation
+and \(\beta\)s are replaced with \(\beta^{(t-1)}\). With simplification
+(up to constant additive terms), we have
 \begin{equation}\label{eq:EQ1_IWLS}
-E(Q_1) = \log f(\textbf{y}|\bs \beta, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^\tp}^{-2})}{\beta_{jk}^\tp}^2\right].
-\end{equation}
-
-\(2E({\tau}^{-2})\beta^2\) can be seen as the kernel of a normal density
-with mean 0 and variance \(E(\tau^{2})\), and we can formulate the
-coefficients \(\bs \beta\) as a multivariate normal distribution with
-means \(\bs 0\) and variance covariance matrix \(\bs \Sigma_{\tau^2}\),
-where \(\bs \Sigma_{\tau^2}\) is a diagonal matrix with \(E(\tau^2)\)s
-on the diagonal, \[
-\bs \beta \sim \text{MVN}(0, \bs \Sigma_{\tau^2}).
+E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[ {2E({\tau_j}^{-2})}{\beta_j}^2 +\sum\limits_{k=1}^{K_j} {2E({\tau_{jk}^*}^{-2})}{\beta_{jk}^*}^2\right].
+\end{equation} \(2E({\tau}^{-2})\beta^2\) can be seen as the kernel of a
+normal density with mean 0 and variance \(E(\tau^{2})\), and we can
+formulate the coefficients \(\boldsymbol{\beta}\) as a multivariate
+normal distribution with means \(\boldsymbol{0}\) and variance
+covariance matrix \(\boldsymbol{\Sigma}_{\tau^2}\), where
+\(\boldsymbol{\Sigma}_{\tau^2}\) is a diagonal matrix with
+\(E(\tau^2)\)s on the diagonal, \[
+\boldsymbol{\beta }\sim \text{MVN}(0, \boldsymbol{\Sigma}_{\tau^2}).
 \]
@@
 likelihood: \[
-f(\textbf{y}|\bs \beta, \phi) \approx \text{MVN}(\textbf{z}|\bs X \bs \beta, \phi\bs \Sigma )
+f(\textbf{y}|\boldsymbol{\beta}, \phi) \approx \text{MVN}(\textbf{z}|\boldsymbol{X} \boldsymbol{\beta}, \phi\boldsymbol{\Sigma })
 \] where the `normal response' \(z_i\) and `weight' \(w_i\) are called
@@
 \end{aligned}
-\] where \(\hat\eta_i = (\bs X {\hat{\bs\beta}})_i\),
+\] where \(\hat\eta_i = (\boldsymbol{X} {\hat{\boldsymbol{\beta}}})_i\),
 \(L^{'}(y_i|\hat\eta_i, \hat \phi)\) and
 \(L^{''}(y_i|\hat\eta_i, \hat \phi)\) are the first and second
-derivative of the log density, \(\log f(\textbf{y}_i|\bs \beta, \phi)\)
-with respect to \(\eta_i\).
+derivative of the log density,
+\(\log f(\textbf{y}_i|\boldsymbol{\beta}, \phi)\) with respect to
+\(\eta_i\).
 
-With \(\bs z\sim \text{MVN}(\bs X \bs \beta, \phi \bs \Sigma)\) and
-\(\bs \beta \sim \text{MVN}(0, \phi \bs \Sigma_{\tau^2})\), we can
-augment the two multivariate normal distributions and update the
-estimates for \(\bs \beta\) and \(\phi\) via least squares in each
-iteration of the EM algorithm. We create the augmented response,
+With
+\(\boldsymbol{z}\sim \text{MVN}(\boldsymbol{X} \boldsymbol{\beta}, \phi \boldsymbol{\Sigma})\)
+and
+\(\boldsymbol{\beta }\sim \text{MVN}(0, \phi \boldsymbol{\Sigma}_{\tau^2})\),
+we can augment the two multivariate normal distributions and update the
+estimates for \(\boldsymbol{\beta}\) and \(\phi\) via least squares in
+each iteration of the EM algorithm. We create the augmented response,
 augmented data, and augmented variance-covariance matrix following
-
 \begin{align*}
-& \bs z_* = \begin{bmatrix} \bs z\\ \bs 0\end{bmatrix} &&
-  \bs X_* = \begin{bmatrix} \bs X \\ \bs I \end{bmatrix} &&
-  \bs \Sigma_* = \begin{bmatrix} \bs \Sigma & \bs 0  \\ \bs 0 & \bs \Sigma_{\tau^2}/\phi \end{bmatrix}, &
-\end{align*}
-
-such that \[
-\bs z_* \sim \text{MVN}(\bs X_* \bs \beta , \phi \Sigma_*).
-\] Using the least squares estimators to update \(\bs\beta\) and
-\(\phi\), we have
-
-\begin{align*}
-& \bs \beta^{(t)} = (\bs X_*^T \bs \Sigma^{-1} \bs X_*)^{-1}\bs X_*^T \bs \Sigma^{-1} \bs z_* && \phi^{(t)} = \frac{1}{n}(\bs z_*-X_*\bs \beta^{(t)})^T\bs \Sigma^{-1}(\bs z_*-X_*\bs \beta^{(t)}).&
-\end{align*}
-
-To note, the variance-covariance matrix of the coefficient estimates
-variance-covariance matrix can be derived in the EM-IWLS algorithm and
-in turn can be used for statistical inferences, \[
-  \text{Var}(\bs\beta^{(t)}) = (\bs X_*^T\bs \Sigma^{-1} \bs X_*)^{-1}\phi^{(t)}.
+& \boldsymbol{z}_* = \begin{bmatrix} \boldsymbol{z}\\ \boldsymbol{0}\end{bmatrix} &&
+  \boldsymbol{X}_* = \begin{bmatrix} \boldsymbol{X} \\ \boldsymbol{I} \end{bmatrix} &&
+  \boldsymbol{\Sigma}_* = \begin{bmatrix} \boldsymbol{\Sigma }& \boldsymbol{0}  \\ \boldsymbol{0} & \boldsymbol{\Sigma}_{\tau^2}/\phi \end{bmatrix}, &
+\end{align*} such that \[
+\boldsymbol{z}_* \sim \text{MVN}(\boldsymbol{X}_* \boldsymbol{\beta }, \phi \Sigma_*).
+\] Using the least squares estimators to update \(\boldsymbol{\beta}\)
+and \(\phi\), we have \begin{align*}
+& \boldsymbol{\beta}^{(t)} = (\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\boldsymbol{X}_*^T \boldsymbol{\Sigma}^{-1} \boldsymbol{z}_* && \phi^{(t)} = \frac{1}{n}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)})^T\boldsymbol{\Sigma}^{-1}(\boldsymbol{z}_*-X_*\boldsymbol{\beta}^{(t)}).&
+\end{align*} To note, the variance-covariance matrix of the coefficient
+estimates variance-covariance matrix can be derived in the EM-IWLS
+algorithm and in turn can be used for statistical inferences, \[
+  \text{Var}(\boldsymbol{\beta}^{(t)}) = (\boldsymbol{X}_*^T\boldsymbol{\Sigma}^{-1} \boldsymbol{X}_*)^{-1}\phi^{(t)}.
 \]
@@
 \item
-  Choose a starting value \(\bs \beta^{(0)}\) and \(\bs \theta^{(0)}\)
-  for \(\bs \beta\) and \(\bs \theta\). For example, we can initialize
-  \(\bs \beta^{(0)} = \bs 0\) and \(\bs \theta^{(0)} = \bs 0.5\)
+  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
+  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
+  \(\boldsymbol{\theta}\). For example, we can initialize
+  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
+  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
 \item
@@
 
-  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^\tp_{j})\) and
-  \(E(\tau^{-2}_{j})\), \(E({\tau^\tp}^{-2}_{jk})\) with the estimates
+  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
+  \(E(\tau^{-2}_{j})\), \(E({\tau^*}^{-2}_{jk})\) with the estimates
   \(\Theta^{(t-1)}\) from the previous iteration
@@
   \item
-    Update \(\bs \beta^{(t)}\) by runing the augmented weighted least
-    squared
+    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
+    least squared
   \item
@@
 
+\clearpage
+
 \subsection{Supplementary Information 3: Predictive Performance of Linear Simulations}
-
-\input{Manuscript/Tabs/sim_lnr_gaus_tab.tex} \clearpage
-\input{Manuscript/Tabs/sim_lnr_binom_tab.tex}
+\input{Tabs/sim_lnr_gaus_tab.tex}
+\input{Tabs/sim_lnr_binom_tab.tex}
 

### 2022-02-22 Manuscript/SS_GAM.tex
Add latex tables to the manuscript

--- 2022-02-15 Manuscript/SS_GAM.tex
@@
 \usepackage{soul}
-\usepackage{multirow}
-\usepackage{multicol}
-\usepackage{colortbl}
-\usepackage{hhline}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{hyperref}
 
@@
 
-\section{Introduction}
-\label{sec:intro}
-
-Many modern biomedical research, e.g.~sequencing data analysis, electric
-health record data analysis, require special treatment of
-high-dimensionality, commonly known as \(p >> n\) problem. There is
-extensive literature on high-dimensional linear models via penalized
-models or Bayesian hierarchical models, see Mallick and Yi
-\citep{Mallick2013} for review. These models are built upon a
-restrictive and unrealistic assumption, linearity. In classical
-statistical modeling, many strategies and models are proposed to relax
-the linearity assumption with various degrees of complexity. For
-example, variable categorization is a simple and common practice in
-epidemiology, but suffers from power and interpretation issues. More
-complex models to address nonlinear effects include random forest and
-other so-called ``black box'' models \citep{Breiman2001}. These models
-are useful for statistical prediction but do not estimate parameters
-relevant to the data generation process that one can draw inferences
-from. In addition, how to generalize these models to the
-high-dimensional setting remains unclear.
-
-Nonparametric regression models are appropriate alternatives to the
-``black-box'' models thanks to their balance between model flexibility
-and interpretability. Among those, generalized additive models (GAMs),
-proposed in the seminal work of Hastie and Tibshirani
-\citep{Hastie1987}, grew to be one of the most popular modeling tools.
-In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution with mean \(\mu\) and dispersion
-\(\phi\), can be modeled with the summation of smoothing functions,
-\(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
-\] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
-The smoothing functions can take many forms and are estimated using a
-pseudo-weighted version of the backfitting algorithm
-\citep{Breiman1985}. Nevertheless, the classical GAMs cannot fulfill the
-increasing analytic demands for high-dimensional data analysis in
-biomedical studies.
-
-There exists some proposals to generalize the classical GAM to
-accommodate high-dimensional applications. The regularized models,
-branching out from group regularized linear models, are used to fit GAMs
-by accounting for the structure introduced when expanding smoothing
-functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
-lasso \citep{Yuan2006} to additive models (AMs); Huang et al.
-\citep{Huang2010} further developed adaptive grouped lasso for additive
-models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
-respectively applied grouped SCAD penalty \citep{Fan2001} to additive
-models. Recently Bayesian hierarchical models are also used in the
-context of high-dimensional additive models. Various group
-spike-and-slab priors combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Bai et al.
-\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
-to Gaussian AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs\citep{Bai2021}. Focus on addressing
-the sparsity, these methods can overly penalize the basis function
-coefficients and produce inaccurate predictions and curve interpolation,
-particularly when complex signals are assumed and large number of knots
-are used. \citep{Scheipl2013} In addition, these methods adapt an
-`all-in-all-out' strategy, i.e.~either including or excluding the
-variable completely, rendering no space for bi-level selection. Scheipl
-et al. \citep{Scheipl2012} proposed a spike-and-slab structure prior
-that address the previous challenges. But the model fitting relies on
-computational intensive MCMC algorithms and creates scalability concern.
-It would be of special interest to develop a fast, flexible and accurate
-generalized additive model framework.
-
-To address these challenges, we propose a novel Bayesian hierarchical
-generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via
-re-parameterization of the smoothing function to avoid overly shrinking
-basis function coefficients. Smoothing penalties are commonly
-implemented in the classical GAMs through smoothing regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive
-penalties on each smoothing function. A smoothing penalty conditioning
-on a smoothing parameter \(\lambda_j\) is a function of the integration
-of the second derivative of the spline function, expressed
-mathematically as \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
-matrix, and \(\boldsymbol{\beta}_j\) are the basis function
-coefficients. Smoothing penalties were also previously used in the
-spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
-penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of a smoothing function from
-the nonlinear space. We then impose a new two-part spike-and-slab spline
-prior on the smoothing functions for bi-level selection such that the
-linear and nonlinear spaces of smoothing functions can be selected
-separately. The prior setup encourages a flexible solution, rendering
-one of three possibilities for each predictor: no effect, only linear
-effect, or linear and nonlinear effects. In addition, two scalable
-optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm
-and EM-Iterative weighted least square (EM-IWLS) algorithm, are
-developed and implemented in an publicly available R package
-\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}, making
-translational science more accessible.
-
-The proposed framework, BHAM, differs from previous spike-and-slab based
-GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
-spline prior is a spike-and-slab lasso type prior using independent
-mixture double exponential distribution, compared to spike-and-slab GAM
-that uses normal-mixture-of-inverse gamma prior. Spike-and-slab lasso
-priors provide computational convenience during model fitting by using
-optimization algorithms instead of intensive sampling algorithms. They
-make fitting high-dimensional models more feasible without sacrificing
-performance in prediction and variable selection. Secondly, SB-GAM uses
-a group spike-and-slab lasso prior with an EM-CD algorithm to fit the
-model. While both methods use the combination of expectation
-maximization algorithm and coordient descent algorithm, there are subtle
-difference in the implementation due to the difference in prior
-specification. The proposed model sets up independent priors among basis
-function coefficients after the re-parameterization step, which provides
-some advantage in computation. In addition, we offer an alternative
-algorithm, EM-IWLS, that can provide variance-covariance matrix of the
-coefficients. Last but not least, the proposed model addresses the
-incapability of bi-level selection in SB-GAM.
-
-In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical
-generalized additive model, introduce the proposed spike-and-slab spline
-priors, and describe the two fast-fitting algorithms. In Section
-\ref{sec:sim}, we compare the proposed framework to state-of-the-art
-models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation
-studies. Analyses of two metabolomics datasets are presented in Section
-\ref{sec:real_data}. Conclusion and discussions are given in Section
-\ref{sec:concl}.
-
-\section{Bayesian Hierarchical Additive Models (BHAMs)}
-\label{sec:BHAM}
-
-Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model with link function \(g(\cdot)\) and
-linear predictor \begin{equation}\label{eq:gam}
-\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
-distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
-and dispersion parameter \(\phi\), and the data distribution is
-\begin{equation}
-f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} The basis function matrix, i.e.~the design matrix derived
-from the smoothing function \(B_j(x_j)\), is denoted
-\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
-design matrix depends on the choice of the smoothing function, and is
-denoted as \(K_j\) for \(x_j\). Its corresponding smoothing penalty
-matrix is denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\)
-denotes the basis function coefficients for the \(j\)th variable such
-that \(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
-abuse of notation, we denote vectors and matrices in bold fonts
-\(\boldsymbol{\beta}, \boldsymbol{X}\) with conformable dimensions,
-where scalar and random variables are denoted in unbold fonts
-\(\beta, X\). The matrix transposing operation is denoted with a
-superscript \(^T\). To note, the proposed model can include parametric
-forms of variables in the model, treating the parametric function as a
-special case of the smoothing function, e.g.~\(B_j(x_j) = x_j\) with the
-smoothing penalty matrix defined as
-\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
-
-To encourage proper smoothing of additive functions, we adopt the idea
-of smoothing penalties from smoothing spline models. The basic idea is
-to set up a smoothing penalty, described in Equation
-(\ref{eq:smoothpen}), in the prior density function. However, the direct
-integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and Wood \citep{Marra2011} proposed a re-parameterization
-procedure to accommodate the smoothing penalty implicitly. Given the
-smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
-semi-definite for univariate smoothing functions, we apply
-eigendecomposition on the penalty matrix
-\(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
-where the matrix \(\boldsymbol{D}\) is diagonal with the eigenvalues
-arranged in the ascending order. To note, \(\boldsymbol{D}\) can contain
-elements of zeros on the diagonal, where the zeros are associated with
-the linear space of the smoothing function. For the most popular
-smoothing function, cubic splines, the dimension of the linear space is
-one. Hereafter, we focus on discussing a uni-dimensional linear space
-for simplicity; however, it generalizes easily to the cases where the
-linear space is multidimensional. We further write the orthonormal
-matrix
-\(\boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
-containing the eigenvectors as columns in the corresponding order to
-\(\boldsymbol{D}\). That is, \(\boldsymbol{U}\) contains the
-eigenvectors \(U^0\) with zero eigenvalues for the linear space and
-\(\boldsymbol{U}^{*}\) contains the eigenvectors (as columns) for the
-non-zero eigenvalues, i.e.~the non-linear space. We multiply the basis
-function matrix \(\boldsymbol{X}\) with the orthonormal matrix
-\(\boldsymbol{U}\) for the new design matrix
-\({\boldsymbol{X}}^\text{repa}= \boldsymbol{X} \boldsymbol{U} \equiv \begin{bmatrix} X^0 : \boldsymbol{X}^{*} \end{bmatrix}\).
-An additional scaling step is imposed on \(\boldsymbol{X}^{*}\) by the
-non-zero eigenvalues of \(\boldsymbol{D}\) such that the new basis
-function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
-each of its dimensions. With slight abuse of the notation, we drop the
-superscript \(^\text{repa}\) and denote
-\(\boldsymbol{X}_j \equiv \begin{bmatrix} X_j^0 : \boldsymbol{X}_j^{*} \end{bmatrix}\)
-as the basis function matrix for the \(j\)th variable after the
-re-parameterization. A spline function can be expressed in the matrix
-form \[
-B_j(x_j) = B_j^0(x_j) + B_j^*(x_j) = \beta_j X^0_j + \boldsymbol{\beta_j^*}^T \boldsymbol{X}_j^*,
-\] and the generalized additive model in Equation (\ref{eq:gam}) now is
-\begin{equation}\label{eq:gam-repa}
-E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j) = g^{-1}\left[\beta_0 + \sum\limits^p_{j=1} (\beta_j X^0_j + {\boldsymbol{\beta}_j^*}^T \boldsymbol{X}_j^*)\right],
-\end{equation} where the coefficients
-\(\boldsymbol{\beta}_j \equiv \begin{bmatrix} \beta_j : \boldsymbol{\beta}^*_j \end{bmatrix}\)
-is an augmentation of the coefficient scalar \(\beta_j\) of linear space
-and the coefficient vector \(\boldsymbol{\beta}^*_j\) of non-linear
-space.
-
-The re-parameterization step sets up the foundation of the proposed
-model and provides three benefits. First of all, the re-parameterization
-integrates the smoothing penalty into the design matrix, and encourages
-models to properly smooth the nonlinear function in addition to the
-sparse penalty for functional selection. Secondly, the
-eigendecomposition of the smoothing penalty allows the isolation of the
-linear space from the nonlinear space, improving the feasibility of
-bi-level functional selection. The eigendecomposition facilitates the
-construction of orthonormal design matrix, which makes imposing
-independent priors on the coefficients possible. This reduces the
-computational complexity compared to using a multivariate priors, and
-greatly broadens the choices of priors and further model choices. Last
-but not least, the scaling step of \(\boldsymbol{X}^*\) further
-simplifies the choice of priors, from column-specific priors to a
-unified prior.
-
-\subsection{Spike-and-Slab Spline Priors}
-
-The family of spike-and-slab (SS) regression models is one of most
-commonly used models in high-dimensional data analysis for its utility
-in outcome prediction and variable selection. We defer to Bai et al.
-\citep{Bai2021Review} for an in-depth introduction to spike-and-slab
-priors. To summarize, spike-and-slab priors are a family of mixture
-distributions that comprises a skinny spike density
-\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
-\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-\] The most distinct feature of SS priors is that it is conditioned on a
-latent binary variable \(\gamma \in \{0,1\}\) that indicates whether the
-variable \(x\) is included in the model. There are various
-spike-and-slab priors depending on the choice for the spike density
-\(f_{\text{spike}}(\cdot)\) and the slab density
-\(f_{\text{slab}}(\cdot)\), see George and McCulloch
-\citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
-variables; Brown et al. \citep{Brown1998} for multivariate outcomes;
-Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
-\citep{Clyde2004} and reference therein. Two of most popular spike and
-slab priors are spike-and-slab normal prior \citep{George1993} and
-spike-and-slab double exponential prior\citep{Rockova2018}.
-
-The major criticism of early spike-and-slab models is being
-computationally prohibitive. \citep{Bai2021Review} Since then, many
-studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS based on spike-and-slab normal
-prior\citep{Rockova2014a} and spike-and-slab Lasso (SSL)
-\citep{Rockova2018b, Rockova2018}. Particularly, the development of SSL
-model substantially improves the scalability of SS models, setting up
-the theoretical foundation for generalized models in -omics data
-analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The SSL prior
-is composed of two double exponential distributions with mean 0 and
-different dispersion parameters, \(0 < s_0 < s_1\), mathematically,
-\begin{equation} 
-\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation} Given that both double exponential distributions have a
-mean of 0 and the latent indicator \(\gamma\) can only take the value of
-0 or 1, the mixture double exponential distribution can be formulated as
-one single double exponential density, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
-\end{equation} with the scale parameter
-\(S = (1-\gamma)s_0 + \gamma s_1\). The SSL also mitigates the problem
-of EMVS where the weak signals are not shrink to zero, and hence is
-preferred in high-dimensional data analysis. We notice that Bai
-\citep{Bai2021} is the first to apply spike-and-slab lasso prior in the
-GAM framework, where the densities of the spike and slab components take
-the group lasso density \citep{Xu2015} and limits to an
-``all-in-all-out'' strategy for functional selection.
-
-\subsubsection{Two-part Spike-and-Slab Lasso Prior}
-
-We introduce a novel prior for GAMs, particularly for high-dimensional
-nonlinear modeling with bi-level selection. The proposed prior extends
-from the spike-and-slab lasso prior described in Equation
-(\ref{eq:ssl}). Given the re-parameterized design matrix
-\(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
-for the \(j\)th variable, we impose a two-part SSL prior to the
-coefficients
-\(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
-Specifically, we impose independent group priors on the linear space
-coefficients and on the nonlinear space coefficients respectively,
-\begin{align}\label{eq:bham_ssl}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j
-\end{align} where \(\gamma_{j}\in\{0,1\}\) and
-\(\gamma^*_{j}\in \{0,1\}\) are two latent indicator variables,
-indicating if the model includes the linear effect and the nonlinear
-effect of the \(j\)th variable respectively. \(s_0\) and \(s_1\) are
-scale parameters, assuming \(0 < s_0 < s_1\) and given. These scale
-parameters \(s_0\) and \(s_1\) can be treated as tuning parameters and
-optimized via cross-validation. A discussion of how to choose the scale
-parameters comes in Section \ref{sec:tune}. To note, this prior differs
-from previous group SSL priors \citep{Tang2018, Tang2019}, as the
-\(\beta_j\) and \(\boldsymbol{\beta}^*_j\) have different indicator
-variables \(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to
-add a more restrictive assumption on the priors, assuming that one
-indicator variable decides the inclusion of both the linear effect and
-nonlinear effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the
-SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
-assumption such that each coefficient
-\(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has its own latent indicator
-\(\gamma_{jk}\), but at the cost of complicating the bi-level functional
-selection. This reduces the proposed prior to the classic SSL prior.
-
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the
-validity of the proposed prior. First of all, the smoothing function
-bases are linear dependent and necessitate extra attention. The
-eigeondecomposition remedies the problem and hence our prior can be set
-to be conditionally independent. Secondly, the eigenvalue scaling
-provides a panacea to allow unified scale parameters for all bases of
-all smoothing functions.
-
-The rest of the hierarchical prior follows the traditional SSL prior: we
-set up hyper-priors on \(\gamma_j, \gamma^*_j\) to allow local adaption
-of the shrinkage using a Bernoulli distribution, written as binomial
-distribution of one trial. The two indicators of the \(j\)th predictor,
-\(\gamma_j\) and \(\gamma^*_j\), shares the same probability parameter
-\(\theta_j\), \[
-\begin{aligned}
-&\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^*| \textcolor{red}{\gamma_{j}, } \theta_j \sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j).
-\end{aligned}
-\]
-\textcolor{red}{Equivalently, we can derive the marginal distribution of $\gamma_{j}^*$ by integrate out $\gamma_j$ (Supplement 1), 
-$$
-\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2).
-$$
-}
-
-This is to leverage the fact that the probability of selecting the bases
-of a smoothing function should be similar, while allowing different
-penalty on the linear space and non-linear space of the smoothing
-function. The hyper prior of \(\gamma_{j}\) decides the sparsity of the
-model at the functional selection level, while that of \(\gamma_{j}^*\)
-decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that \(\gamma_{j}\) and \(\gamma_{j}^*\) are
-independently distributed for analytic simplicity. We further specify
-the parameter \(\theta_j\) follows a beta distribution with given shape
-parameters \(a\) and \(b\), \[
-\theta_j \sim Beta(a, b).
-\] The beta distribution is a conjugate prior for the binomial
-distribution and hence provides some computation convenience. For
-simplicity, we focus on a special case of beta distribution, uniform
-(0,1), i.e.~\(a = 1, b = 1\). When the variable have large effects in
-any of the bases, the parameter \(\theta_j\) will be estimated large,
-which in turn encourages the model to include the rest of bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models
-with the spike-and-slab spline prior as the BHAM, and visually presented
-in Figure \ref{fig:SSprior}.
-
-\begin{center}
-Figure \ref{fig:SSprior} here
-\end{center}
-
-\subsubsection{Other Priors}
-
-With the re-parameterization step of the basis function matrix
-\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
-priors, for example normal priors for ridge-type regularization and
-mixture normal prior for spike-and-slab regularization. These priors
-would work better in low and medium dimensional settings where the
-sparse assumption is not necessary. Here we elaborate the mixture normal
-prior as a demonstration of applying continuous spike-and-slab prior in
-BHAM.
-
-A spike-and-slab mixture normal spline prior can be expressed as
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
-\end{align*} Similar to the spike-and-slab spline prior in Equation
-(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
-optimized via cross-validation. One of the critics received by the
-spike-and-slab mixture normal prior is that the tails of a normal
-distribution diminishes to zero too fast, which causes problems when
-estimating the large effects. Distributions with heavier tails can be
-used as an alternative, for example mixture Student's \(t\) distribution
-with small degree of freedom.
-
-\subsection{Algorithms for Fitting BHAMs}
-
-The proposed models can be fitted with MCMC algorithms. Nevertheless,
-the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch\citep{George1997} examined the computation speed
-for various MCMC algorithms with spike-and-slab mixture normal priors,
-and suggested MCMC algorithms works well for medium size (\(p\)=25) of
-predictors with only linear effects. However, it is not feasible for
-high-dimensional data analysis where the number of predictors easily
-exceeds 100. Specific to additive models, each predictor would expand to
-multiple new ``predictors'' via basis functions, creating greater
-computational demands. Scheipl et al. \citep{Scheipl2013} demonstrated
-the computational demands of a MCMC algorithm for fitting spike-and-slab
-GAM grow exponentially as \(p\) increases modestly via simulation
-studies. Hence, we feel compelled to develop scalable algorithms for
-fitting Bayesian hierarchical additive models in high-dimensional
-settings.
-
-As an alternative to sampling algorithms for fitting Bayesian models,
-optimization algorithms focus on the maximum a posteriori (MAP)
-estimates and speed up the model fitting process at the cost of
-posterior inference. The earlier work for fitting SS models using
-optimization algorithms includes EMVS \citep{Rockova2014a}. Rockova and
-George\citep{Rockova2014a} proposed an expectation-maximization (EM)
-based algorithm to fit models that use continuous mixture normal priors.
-In the E step, the latent binary indicators \(\gamma\)s are treated as
-the missing data, and the posterior means are calculated conditioning on
-the current value of other parameters; in the M step, a ridge estimator
-was used to update the coefficients, followed by updating
-\(\phi, \theta\). The same authors \citep{Rockova2018b, Rockova2018}
-further combined the EM algorithm with coordinate descent algorithm to
-fit SSL models. Yi and his group independently developed the
-EM-Iterative Weighted Least Square and EM-cyclic coordinate descent
-algorithms to fit models with broader class of priors, SSL
-included.\citep{Yi2019} These algorithms were implemented for
-generalized linear models\citep{Tang2017a}, Cox proportional hazards
-models \citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a popular property for
-reproducible research
-
-In this section, we extend the two EM-based algorithms, EM-CD and
-EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
-provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM with an expedited performance, recommending to use in high
-and ultra-high dimensional setting. A specific concern of EM-CD
-algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can estimate the variance-covariance matrix of the
-coefficients. Moreover, the EM-IWLS is a more general model fitting
-algorithm that can be used for fitting not only SSL and continuous SS
-priors but also Student's t-priors and double exponential priors. To
-note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
-main difference between proposed EM-CD algorithm and that in SB-GAM is
-that SB-GAM uses a block CD algorithm for their group prior, while the
-proposed prior is pairwise independent requiring no special treatment in
-the CD algorithm.
-
-\subsubsection{EM algorithms}
-
-EM algorithm is an iterative algorithm to find MAP estimates or the
-maximum likelihood estimates. It is commonly used when some necessary
-data to establish the likelihood function are missing. Instead of
-maximizing the the likelihood function, the algorithm maximizes the
-expectation of the likelihood function with respect to the ``missing''
-data.
-
-The recursive algorithm consists of two steps:
-
-\begin{itemize}
-\tightlist
-\item
-  E-step: to calculate the expectation of the posterior density function
-  with respect to some ``missing'' data
-\item
-  M-step: to maximize the expectation derived in the E-step and update
-  parameters of interest
-\end{itemize}
-
-For BHAMs, we define the parameters of interest as
-\(\Theta = \{\boldsymbol{\beta}, \boldsymbol{\theta}, \phi\}\) and
-consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model, in other words the ``missing'' data.
-Our objective is to find the parameters \(\Theta\) that maximize the
-posterior density function, or equivalently, the logarithm of the
-density function, \[
-\begin{aligned}
-& \text{argmax}_{\Theta}
-\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) \\
-&= \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{j})\right]\\
-& +\sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j),
-\end{aligned}
-\]\\
-where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
-non-informative prior for the intercept \(\beta_0\) and the dispersion
-parameter \(\phi\); for example, \(f(\beta_0|\tau_0^2)=N(0,\tau_0^2)\)
-with \(\tau^2_0\) set to a large value and \(f(\log \phi) \propto 1\).
-
-We use the EM algorithm to find the MAP estimate of \(\Theta\). This is,
-in the E-step, we calculate the expectation of posterior density
-function of
-\(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
-respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
-the values from previous iteration \(\Theta^{(t-1)}\), \[
-E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
-\] Hereafter, we use the shorthand notation
-\(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}(\cdot)\). In the
-M-step, we find the \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The E-
-and M- steps are iterated until the algorithm converge.
-
-To note here, the log-posterior density of BHAMs (up to additive
-constants) can be written as a two-part equation
-\[ \log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) = Q_1(\boldsymbol{\beta}, \phi) + Q_2 (\boldsymbol{\gamma},\boldsymbol{\theta}),\]
-where
-\[ Q_1\equiv Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log f(\beta^{*}_{jk}|\gamma^{*}_{jk})\right]\]
-and
-\[Q_2 \equiv Q_2(\boldsymbol{\gamma},\boldsymbol{\theta}) = \sum\limits_{j=1}^{p} \left[ (\gamma_j+\gamma_{j}^{*})\log \theta_j + (2-\gamma_j-\gamma_{j}^{*}) \log (1-\theta_j)\right] +  \sum\limits_{j=1}^{p}\log f(\theta_j).\]
-\(Q_1\) and \(Q_2\) are respectively the log posterior density of the
-coefficients \(\boldsymbol{\beta}\) and the log posterior density of the
-probability parameters \(\boldsymbol{\theta}\) conditioning on
-\(\boldsymbol{\gamma}\). Meanwhile, conditioning on
-\(\boldsymbol{\gamma}\), \(Q_1\) and \(Q_2\) are independent and can be
-maximized separately for \(\boldsymbol{\beta}, \phi\) and
-\(\boldsymbol{\theta}\). Depending on the choice of coefficient priors,
-\(Q_1\) can be treated as penalized likelihood function and maximization
-of \(E(Q_1)\) can be solved via CD algorithm or IWLS algorithm in each
-iteration. Maximization of \(E(Q_2)\) can be solved via closed form
-equations following the beta-binomial conjugate relationship.
-
-\subsubsection{EM-Coordinate Descent}
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double
-exponential, coordinate descent algorithm can be used to estimate the
-parameters in the M-step. Coordinate descent is an optimization
-algorithm that offers extreme computational advantage, and famous for
-its application in optimizing the \(l_1\) penalized likelihood function.
-
-The density function of spike-and-slab mixture double exponential prior
-can be written as \[
-f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
-\] and \(E(Q_1)\) can be expressed as a log-likelihood function with
-\(l_1\) penalty \begin{equation}\label{eq:Q1_CD}
-E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
-\end{equation} where
-\(S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
-\(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
-unknown quantities \(E({S_j}^{-1})\) and \(E({S^*}^{-1}_j)\), the
-posterior probability
-\(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
-\(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
-necessary, which can be derived via Bayes' theorem. The calculation of
-\(p_j^*\) is slightly different from that of \(p_j\), as \(p_j^*\)
-depends on the value of the vector \(\boldsymbol{\beta}^*_{j}\) and
-\(p_j\) only depends on the scalar \(\beta_j\). The calculation follows
-the equations below. \begin{align*}
-p_{j} &= \frac{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \text{Pr}(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
-p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{align*} and \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j}\),
-\(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
-\(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
-show \begin{align}\label{eq:exp_scale}
-&E(\gamma_{j})  = p_{j} & &E(\gamma^{*}_{j}) = p_{j}^{*}\nonumber\\
-&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^*}^{-1}_{j}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
-\end{align} After replacing withe calculated quantities, \(E(Q_1)\) can
-be seen as a \(l_1\) penalized likelihood function with the
-regularization parameter \(\lambda = E(S^{-1})\), and hence be optimized
-via coordinate descent algorithm \citep{Friedman2010}. Independently,
-the remaining parameters of interest \(\boldsymbol{\theta}\) can be
-updated by maximizing \(E(Q_2)\). As the beta distribution is a
-conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
-be easily updated with a closed form equation,
-\begin{equation}\label{eq:update_theta}
-\theta_j = \frac{p_j + p^*_{j} + a - 1 }{a + b}.
-\end{equation}
-
-Totally, the proposed EM-coordinate descent algorithm is summarized as
-follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  Choose a starting value \(\boldsymbol{\beta}^{(0)}\) and
-  \(\boldsymbol{\theta}^{(0)}\) for \(\boldsymbol{\beta}\) and
-  \(\boldsymbol{\theta}\). For example, we can initialize
-  \(\boldsymbol{\beta}^{(0)} = \boldsymbol{0}\) and
-  \(\boldsymbol{\theta}^{(0)} = \boldsymbol{0}.5\)
-\item
-  Iterate over the E-step and M-step until convergence
-
-  E-step: calculate \(E(\gamma_{j})\), \(E(\gamma^*_{j})\) and
-  \(E({S}^{-1}_{j})\), \(E({S^*}^{-1}_{j})\) with estimates of
-  \(\Theta^{(t-1)}\) from previous iteration
-
-  M-step:
-
-  \begin{enumerate}
-  \def\labelenumii{\alph{enumii})}
-  \item
-    Update \(\boldsymbol{\beta}^{(t)}\), and the dispersion parameter
-    \(\phi^{(t)}\) if exists, using the coordinate descent algorithm
-    with the penalized likelihood function in Equation (\ref{eq:Q1_CD})
-  \item
-    Update \(\boldsymbol{\theta}^{(t)}\) using Equation
-    (\ref{eq:update_theta})
-  \end{enumerate}
-\end{enumerate}
-
-We assess convergence by the criterion:
-\(|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon\), where
-\(d^{(t)} = -2\log f(\textbf{y}| \textbf{X}, \boldsymbol{\beta}^{(t)},\phi^{(t)})\)
-is the estimate of deviance at the \(t\)th iteration, and \(\epsilon\)
-is a small value (say \(10^{-5}\)).
-
-\subsection{Selecting Optimal Scale Values}
-\label{sec:tune}
-
-Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
-\(s_1\)). Hence, we need to find the optimal values for the scale
-parameters such that the model reaches its best prediction performance
-regarding a criteria of preference. This would be achieved by
-constructing a two dimensional grid, consists of different pairs of
-(\(s_0\), \(s_1\)) value. However, previous research suggested the value
-of slab scale \(s_1\) have less impact on the final model and is
-recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
-provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
-on examining different values of spike scale \(s_0\). Instead of the 2-D
-grid, We consider a sequence of \(L\) decreasing values
-\(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
-spike scale \(s_0\) tends to include more non-zero coefficients in the
-model. A measure of preference calculated with cross-validations (CV),
-e.g.~deviance, area under the curve (AUC), mean squared error, can be
-used to facilitate the selection of a final model. The procedure is
-similar to the Lasso implemented in the widely used R package
-\texttt{glmnet}, which quickly fits Lasso models over a list of values
-of regularization parameters \(\lambda\) , giving a sequence of models
-for users to choose from.
-
-\section{Simulation Study}
-\label{sec:sim}
-
-In this section, we compare the performance of the proposed models to
-four alternative models: component selection and smoothing operator
-(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
-generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM \citep{Bai2021}\textcolor{red}{}, linear lasso model and
-spikeSlabGAM {[}TODO: add citation here{]}\}. COSSO is one of the
-earliest smoothing spline models that consider sparsity-smoothness
-penalty. Adaptive COSSO improved upon COSSO by using adaptive weight for
-penalties such that the penalty of each functional component are
-different for extra flexibility. Generalized additive models with
-automatic smoothing, hereafter \textit{mgcv}, is one of the most popular
-models for nonlinear effect interpolation and prediction. SB-GAM is the
-first spike-and-slab lasso GAM. We implement COSSO and adaptive COSSO
-with R package \texttt{cosso 2.1-1}, generalized additive models with
-automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R
-package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence use the default
-choices. Both mgcv and proposed models allow customized smoothing
-functions and we choose the cubic regression spline. We controll the
-dimensionality of each smoothing function, 10 bases, for all different
-choices of smoothing functions. We use 5-fold CV with the default
-selection criteria to select the final model for COSSO models, SB-GAM
-and the proposed models. 20 default candidates of tuning parameters
-(\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for SB-GAM and
-the proposed models which allow user-specification of tuning candidates.
-All computation was conducted on a high-performance 64-bit Linux
-platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
-processors and 24G of RAM per core and R 3.6.2 \citep{R}.
-
-Other related methods for high-dimensional GAMs also exist, notably the
-methods of sparse additive models by Ravikumar et al.
-\citep{Ravikumar2009} and stochastic search term selection for GAM
-\citep{Scheipl2012}. However, we exclude these methods from current
-simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increased number of
-predictors. \citep{Scheipl2013}
-
-\subsection{Monte Carlo Simulation Study}
-
-We follow the data generating process described in Bai \citep{Bai2021}.
-We first generate \(n=500\) training data points with
-\(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
-\(X\) are simulated from a multivariate normal distribution
-\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(y\)
-from two distributions, Gaussian and binomial with the identity link and
-logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The mean of each
-outcome were simulated via the following function \[
-\mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for Gaussian and binomial outcomes. Gaussian outcomes requires
-specification of dispersion, where we set the dispersion parameter to be
-1. In this data generating process, we have \(x_1, x_2, x_3, x_4\) as
-the active covariates, while the rest covariates are inactive,
-i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
-independent sample of size \(n_{test}=1000\), are created following the
-same data generating process, serving as the testing data. We generate
-50 independent pairs of training and testing datasets to evaluate the
-prediction performance of the chosen models, where training datasets are
-used to fit the models and testing datasets used to calculate assessment
-measures.
-\textcolor{red}{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
-These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
-
-To evaluate the predictive performance of the models, the statistics,
-\(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing dataset, are averaged across 50 simulations.
-\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
-For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-
-\textcolor{red}{\subsection{Simulation Results}}
-\textcolor{red}{\subsubsection{Prediction Performance}}
-
-The predictive performances have a consistent pattern across the two
-distributions of outcomes. Across all the scenarios, COSSO and adaptive
-COSSO have the least favorable performance among the applicable methods
-examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}). To note, mgcv
-doesn't support high-dimensional analysis, i.e.~the number of
-coefficients are greater than the sample size, and hence not evaluated
-when \(p=100, 200\). mgcv predicts well when \(p\) is small or moderate
-(\(p = 4, 10\)), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the
-proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when
-the dimension of are moderate (p=4, 10, 50). Particularly, BHAM-IWLS
-performs as good as mgcv if not better. However, in high dimensional
-case where we mimic the situation the signals are extremely sparse,
-SB-GAM has better performance than the proposed method. However, the
-BHAM-CD has extreme computational advantage over SB-GAM (see Table
-\ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
-
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
-
-\section{Metabolomics Data Analysis}
-\label{sec:real_data}
-
-In this section, we apply the proposed models BHAM, fitted with the
-EM-CD algorithm, to analyze two real-world metabolomics datasets where
-the outcomes are binary and continuous respectively. We demonstrate the
-improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}, while being
-computationally efficient (see Table \ref{tab:time_real_data}).
-
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
-
-\subsection{Emory Cardiovascular Biobank}
-\label{sec:ECB}
-
-We use the proposed models BHAM to analyze a metabolic dataset from a
-recently published research \citep{Mehta2020} studying plasma
-metabolomic profile on the three-year all-cause mortality among patients
-undergoing cardiac catheterization. The dataset is publicly available
-via \textit{Dryad} \citep{Mehta2020_data}. It contains in total of 776
-subjects from two cohorts. As there is a large number of non-overlapping
-features among the two cohorts, we use the cohort with larger sample
-size (N=454). There are initially 6796 features in the dataset, which is
-too large to be practically meaningful to analyze. Hence, we perform a
-univariate screening procedure on the features, via GAM implemented in
-\texttt{mgcv}, and choose the the top 200 features with smallest
-p-values. We use 5-knot spline additive models for binary outcome using
-two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are
-used to choose the optimal tuning parameters of each framework with
-respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where
-deviance, AUC, Brier score defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassication error defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
-calculated. BHAM-CD obtains superior AUC, Brier score, and
-misclassification error in the out-of-bag samples compared to SB-GAM
-(see Table \ref{tab:ECB_res}).
-
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
-
-\subsection{Weight Loss Maintenance Cohort}
-
-We use the proposed models BHAM to analyze metabolomics data from a
-recently published study \citep{Bihlmeyer2021} on the association
-between metabolic biomarkers and weight loss, where the dataset is
-publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
-primarily focuse on the analysis of one of the three studies included,
-weight loss maintenance cohort \citep{Svetkey2008}, due to the
-drastically different intervention effects. In the dataset, 765
-metaboliltes in baseline plasma collected were profiled using liquid
-chromatography mass spectrometry. Quality control and natural log
-transformation are performed during metabolites data preparation. The
-outcome of interest are standardized percent change in insulin
-resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there are
-\(p\)=237 features remaining in the analysis. 5-Knot spline additive
-models for the Gaussian outcome are constructed using two different
-models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
-the optimal tuning parameters of each framework with respect to the
-default selection criterion implemented in the software. Out-of-bag
-samples are used for prediction performance evaluation, where deviance,
-\(R^2\), mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-absolute error (MAE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
-BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
-compared to SB-GAM (see Table \ref{tab:WLM_res}).
-
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}
-
-\section{Discussion}
-\label{sec:concl}
-
-In the paper, we described a novel generalized additive model using
-Bayesian hierarchical priors, particularly the proposed spike-and-slab
-spline prior for bi-level functional selection. Meanwhile, we introduced
-two optimization based algorithms for model fitting. The algorithms can
-be easily scale up to address high-dimensional data analysis in a
-computational efficient manner. Via simulations, we demonstrated that
-the proposed model provides as good, if not better, prediction
-performance compared to some state-of-the-art non-linear modeling
-devices.
-
-The proposed model shares many commonality with an high-dimensional
-Bayesian GAM, SB-GAM \citep{Bai2021}, independently developed around the
-same time of this work. Both frameworks emphasize computational
-efficiency by deploying group spike-and-slab lasso type of priors and
-optimization-based fast and scalable algorithms. Bai provided the
-theoretical proof for the consistency of variable selection using group
-spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the
-bi-level functional selection and the model inference. The proposed
-model provides solutions to these remaining questions while maintaining
-the same level of prediction accuracy. Moreover, the proposed model
-renders additional flexibility to model specification, allowing various
-choices of smoothing functions and degrees of freedom.
-
-For translational science purpose, we implemented the proposed model in
-a R package \texttt{BHAM}, deposited at
-\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
-smoothing function specification, we deploy the same programming grammar
-as in the state-of-the-art package \texttt{mgcv}, in contrast to
-previous tools where smoothing functions are limited to the default
-ones. Ancillary functions are provided for model specification in
-high-dimensional settings, curve plotting and functional selection. In
-addition, in \texttt{BHAM}, we streamline the model fitting algorithms
-to support other popular Bayesian hierarchical prior for the smoothing
-functions, such as Student's T distribution, double exponential
-distribution, mixture normal and T distribution's. These priors could be
-helpful when the sparse assumption is weak or not necessary.,
-
-There are some improvements possible for the proposed models. First of
-all, the proposed model achieves a bi-level selection via the two-part
-spike-and-slab spline prior. Nevertheless, this set-up could result in a
-situation that is not theoretically sound: the non-linear component is
-selected, but the linear component is not. We currently address it
-analytically by including the linear component in the model when
-non-linear component is selected. Another possible solution is to impose
-a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, the
-computational time for fitting a BHAM model with EM-IWLS algorithms can
-be improved. Current implementation of the EM-IWLS algorithm jointly
-updates all coefficients in each iteration, which requires a lot
-computation resources. This jointly updating procedure can be enhanced
-by adapting a backfitting step \citep{Hastie1987} where each smoothing
-function are updated individually. Thirdly, when using the proposed
-model in real data analysis where a screening procedure is implemented
-before joint predictive modeling. we recommend to include the screening
-procedure in the cross-validation during model selection.
-\citep{Friedman2017} For the sake of model comparison, we fit the models
-using the same pool of predictors in Section \ref{sec:ECB}.
-
-Our future efforts direct to modeling survival outcomes and integrative
-analysis. While the proposed model addresses a great deal of analytic
-problem, analyzing the time-to-event outcome remains unsolved. An naive
-approach would be convert a time-to-event outcome to a Poisson outcome
-following Whitehead \citep{Whitehead1980}. However, it would be more
-efficient to directly fit Cox models via penalized pseudo likelihood
-function \citep{Simon2011}. Meanwhile, with growing understanding of
-biological structure within -omics field, it is appealing to integrate
-external biology information in the modeling process. The main
-motivation for integrative models is that biologically informed grouping
-of weak effects increases the power of detecting true associations
-between features and the outcome \citep{Peterson2016}, and stabilizes
-the analysis results for reproducibility purpose. Such integration can
-be achieved by setting up a structural hyperprior on the inclusion
-indicator of the smoothing function null space
-\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
-and Dunson \citep{Ferrari2020}.
-
 \clearpage
@@
 
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{ccccccc}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
-  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
-  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:gaus}
-\end{table}
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:bin_auc}
-\end{table}
+\include{"Tabs/sim_gaus_tab.tex"}
+\include{Tabs/sim_binom_tab.tex}
 
@@
 
-\begin{sidewaystable}[!h]
-\centering
-\begin{tabular}{lrllllll}
-  \hline
-Distribution & P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-Binomial & 4.00 & 5.88 (1.01) & 6.96 (1.76) & 2.97 (0.82) & 4.60 (2.32) & 0.18 (0.04) & 343.24 (88.37) \\ 
-  Binomial & 10.00 & 14.01 (1.86) & 7.65 (1.25) & 7.95 (2.70) & 10.01 (3.56) & 3.51 (12.03) & 543.13 (133.51) \\ 
-  Binomial & 50.00 & 185.26 (17.44) & 53.83 (6.26) & 109.72 (24.31) & 128.13 (23.26) & 670.01 (151.28) & 1630.42 (193.61) \\ 
-  Binomial & 100.00 & 1106.93 (356.73) & 48.41 (5.84) & 715.15 (170.55) & 718.63 (193.53) & - & 2783.99 (235.31) \\ 
-  Binomial & 200.00 & 6923.62 (1911.59) & 99.98 (7.05) & 5572.51 (1295.18) & 4958.69 (1970.46) & - & 4780.10 (488.66) \\ 
-  Gaussian & 4.00 & 4.22 (1.51) & 2.51 (0.20) & 0.79 (0.11) & 0.67 (0.07) & 0.05 (0.00) & 36.02 (3.02) \\ 
-  Gaussian & 10.00 & 17.59 (5.77) & 33.32 (3.50) & 3.49 (0.70) & 3.47 (0.71) & 0.32 (0.34) & 72.93 (9.52) \\ 
-  Gaussian & 50.00 & 329.83 (44.57) & 334.53 (22.33) & 35.54 (9.15) & 35.44 (9.02) & 73.33 (71.24) & 373.42 (34.93) \\ 
-  Gaussian & 100.00 & 1706.19 (225.92) & 446.98 (79.31) & 150.58 (46.72) & 152.82 (45.07) & - & 684.05 (67.80) \\ 
-  Gaussian & 200.00 & 16379.27 (4825.56) & 303.73 (87.27) & 591.11 (147.88) & 569.92 (107.45) & - & 1314.12 (137.85) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:time_sim}
-\end{sidewaystable}
+\clearpage
 
@@
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
-
-\setlength{\tabcolsep}{2pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}
-
-\caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\textbackslash{}label\{tab:time\_real\_data\}
-}\\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\endfirsthead
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
-
-\hhline{~>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\multirow[c]{-2}{*}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Data}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{CV}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Final}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{Total}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{ECB}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{225.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{228.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,506.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{34.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,540.7}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{483.1}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{7.6}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{490.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}
-
-\end{longtable}
+\clearpage
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & AUC & Brier & Misclass \\ 
-  \hline
-BHAM-CD & 455.69 & 0.74 & 0.16 & 0.21 \\ 
-  SB-GAM & 1230.06 & 0.71 & 0.21 & 0.24 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
-\label{tab:ECB_res}
-\end{table}
-
-\clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Methods & Deviance & $R^2$ & MSE & MAE \\ 
-  \hline
-BHAM-CD & 665.63 & 0.07 & 0.94 & 0.75 \\ 
-  SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
-   \hline
-\end{tabular}
-\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
-\label{tab:WLM_res}
-\end{table}
-
 \end{document}

### 2022-02-22 _targets.R

@@ -320,19 +320,7 @@ tar_plan(
   
   # measure.glm(WLM_train_dat$out_HOMA_std, WLM_SBGAM_fnl$mu.pred, family = "gaussian"),  
   
-  # Manuscript --------------------------------------------------------------
-  tar_files(manu_path,
-            c(
-              "Manuscript/01-Intro.Rmd",
-              "Manuscript/02-Method.Rmd",
-              "Manuscript/03-Simulation.Rmd",
-              "Manuscript/04-Real_Data.Rmd",
-              "Manuscript/05-Discussion.Rmd",
-              "bibfile.bib"
-            )
-  ),
-  
-  # TODO: laocate to the real data analysis session.
+  #* Running Time ---------------------------------------------------------------
   # TODO: create a latex output
   tar_target(tbl_real_time,
              create_real_data_time_tbl(  # Introduce dependency
@@ -344,23 +332,27 @@ tar_plan(
                WLM_bamlasso_fnl,
                WLM_SBGAM_cv_raw,
                WLM_SBGAM_fnl)),
+  
+  # Manuscript --------------------------------------------------------------
+  tar_files(manu_path,
+            c("Manuscript/01-Intro.Rmd", "Manuscript/02-Method.Rmd",
+              "Manuscript/03-Simulation.Rmd", "Manuscript/04-Real_Data.Rmd",
+              "Manuscript/05-Discussion.Rmd", "Manuscript/bibfile.bib")
+  ),
+  
   tar_files(
     manu_tbs,
     list.files("Manuscript/Tabs/", full.names = TRUE)
   ),
   
-  #* Assemble Manuscript ####
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
   # R & R Letter --------------------------------------------------------------
-  tar_target(RR_path,
-             c("Manuscript/R&R/01-reviewer_1.Rmd",
-               "Manuscript/R&R/02-reviewer_2.Rmd",
-               "Manuscript/R&R/03-reviewer_4.Rmd",
-               "Manuscript/R&R/reply_ref.bib"
-             ),
-             format = "file"),
+  tar_files(RR_path,
+            c("Manuscript/R&R/01-reviewer_1.Rmd", "Manuscript/R&R/02-reviewer_2.Rmd",
+              "Manuscript/R&R/03-reviewer_4.Rmd", "Manuscript/R&R/reply_ref.bib")
+  ),
[…]

### 2022-02-22 R/Sim_Res/format_predict_tbls.R
replace mgcv NA values

@@ -1,6 +1,7 @@
 format_predict_tbls <- function( dat, caption, label){
   dat %>% 
-    mutate(p = as.integer(p)) %>% 
+    mutate(p = as.integer(p),
+           mgcv = replace( mgcv, p %in% c(100,200) , "-")) %>% 
     select(
       "P" = p,
       "mgcv" = mgcv,

### 2022-02-22 Manuscript/00-main.Rmd
Add latex tables to the manuscript
Replace prediction performance tables with latex tables

--- 2022-02-15 Manuscript/00-main.Rmd
@@
 
-tar_load(intro_path)
-tar_load(method_path)
-tar_load(sim_path)
-tar_load(real_data_path)
-tar_load(disc_path)
-
-tar_load(tab_binom)
-tar_load(tab_gaus)
+tar_load(manu_path)
+tar_load(manu_tbs)
 
@@
 tar_load(tbl_real_time)
-tar_load(tab_time)
+# tar_load(sim_tim_tab)
 ```
@@
 \clearpage
-
-```{r, results = "asis"}
-tab_gaus$mgcv[4] <- "-"
-tab_gaus$mgcv[5] <- "-"
-tab_gaus %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>% 
-  xtable(
-    align = "cccccccc",
-    caption = "The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.",
-    label = "tab:gaus") %>% 
-  print(comment = FALSE, include.rownames = FALSE)
-  # flextable(
-  #   # col_keys = c("P", "BHAM-IWLS", "BHAM-CD", "COSSO", "Adaptive COSSO", "mgcv", "SB-GAM")
-  # ) %>% 
-  # flextable::autofit() %>% 
-  # flextable::set_caption(caption = "The average and standard deviation of the out-of-sample $R^2$ measures for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-\\label{tab:gaus}")
-```
+\input{Tabs/sim_gaus_tab.tex}
 \clearpage
-
-```{r, results = "asis"}
-tab_binom$mgcv[4] <- "-"
-tab_binom$mgcv[5] <- "-"
-tab_binom %>% 
-  mutate(p = as.integer(p)) %>% 
-  rename(
-    "P" = p,
-    "BHAM-IWLS" = bglm_spline_de,
-    "BHAM-CD" = blasso_spline,
-    "COSSO" = cosso,
-    "Adaptive COSSO" = acosso,
-    "SB-GAM" = SB_GAM
-  ) %>%  
-  xtable(caption = "The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.", label = "tab:bin_auc") %>%
-  print(comment = FALSE, include.rownames = FALSE)
-```
+\input{Tabs/sim_binom_tab.tex}
 
@@
 <!-- \begin{landscape} -->
-```{r results = "asis"}
-tab_time %>% 
+```{r sim_time, results = "asis", eval = F}
+sim_tim_tab %>% 
   select(dist, p, `BHAM-IWLS`,`BHAM-CD`, cosso, acosso, mgcv, `SB-GAM`) %>% 

### 2022-02-22 Manuscript/03-Simulation.Rmd
Replace prediction performance tables with latex tables

--- 2022-02-18 Manuscript/03-Simulation.Rmd
@@
 For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.
-MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
 $$
-where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.
+\bg{MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},}
 $$
-}
-\bg{\subsection{Simulation Results}}
-\bg{\subsubsection{Prediction Performance}}
+\bg{where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.}
+\subsection{\bg{Simulation Results}}
+\subsubsection{\bg{Prediction Performance}}
 The predictive performances have a consistent pattern across the two distributions of outcomes.
@@
 
-\bg{
+\begin{itemize}
+\item \bg{When the linear assumptions are violated, the linear lasso model performs extremely bad. adaptive cosso have improvement upon cosso.}
+\item \bg{To note, the drastic deteriation in the prediction of mgcv is caused by the model would not converge in those scenarios. A null model were returned as a final model. }
+\item \bg{Within the simulations where the functional form of the predictors are linear, we see that the proposed model similar to the lasso model regardless of the type outcome, and better moderately better than the model of comparison.
+(See Table Lnr Gause, and supporting table Lnr-binom.) Particularly, in gaussina simulations, the proposed model shows great advantage compared to SB-GAM (numeralize the advantage), and slightly improvement compare to ssGAM (numeralize the advantage).
+This shows that the model can be applied without constraining to the assumption where all functional form are non-linear.}
+\end{itemize}
 
+\bg{\subsubsection{Variable Selection Performance}
+While the main utility of the proposed work is prediction, the proposed model can also serve as variable selection/functional selection as a by-product.
+Hence We evaluate the variable selection performance.
 }
 
-\bg{\subsubsection{Variable Selection Performance}
 \begin{itemize}
-\item If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
+\item \bg{If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
 The pattern is consitent regardless of the type of outcomes.
 To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
-Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.
-\item If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
+Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.}
+\item \bg{If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
 While the recall is lower than SBGAM.
-This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.
-\item We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while. On the contrary the SBGAM is not so well.
+This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.}
+\item \bg{We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while. On the contrary the SBGAM is not so well.}
 \end{itemize}
-
-}
 

### 2022-02-22 Manuscript/SS_GAM.tex
Replace prediction performance tables with latex tables

--- 2022-02-18 Manuscript/SS_GAM.tex
@@
 For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-
-\begin{center}
-Table \ref{tab:gaus} here\\
-Table \ref{tab:bin_auc} here\\
-\end{center}
-
-\textcolor{red}{\subsection{Simulation Results}}
-\textcolor{red}{\subsubsection{Prediction Performance}}
+\[
+\textcolor{red}{MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},}
+\]
+\textcolor{red}{where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.}
+
+\subsection{\textcolor{red}{Simulation Results}}
+\subsubsection{\textcolor{red}{Prediction Performance}}
 
@@
 
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
+\begin{itemize}
+\item \textcolor{red}{When the linear assumptions are violated, the linear lasso model performs extremely bad. adaptive cosso have improvement upon cosso.}
+\item \textcolor{red}{To note, the drastic deteriation in the prediction of mgcv is caused by the model would not converge in those scenarios. A null model were returned as a final model. }
+\item \textcolor{red}{Within the simulations where the functional form of the predictors are linear, we see that the proposed model similar to the lasso model regardless of the type outcome, and better moderately better than the model of comparison.
+(See Table Lnr Gause, and supporting table Lnr-binom.) Particularly, in gaussina simulations, the proposed model shows great advantage compared to SB-GAM (numeralize the advantage), and slightly improvement compare to ssGAM (numeralize the advantage).
+This shows that the model can be applied without constraining to the assumption where all functional form are non-linear.}
+\end{itemize}
+
+\textcolor{red}{\subsubsection{Variable Selection Performance}
+While the main utility of the proposed work is prediction, the proposed model can also serve as variable selection/functional selection as a by-product.
+Hence We evaluate the variable selection performance.
+}
+
+\begin{itemize}
+\item \textcolor{red}{If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
+The pattern is consitent regardless of the type of outcomes.
+To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
+Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.}
+\item \textcolor{red}{If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
+While the recall is lower than SBGAM.
+This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.}
+\item \textcolor{red}{We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while.
+On the contrary the SBGAM is not so well.}
+\end{itemize}
 
@@
 \clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{ccccccc}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.90 (0.01) & 0.90 (0.01) & 0.75 (0.03) & 0.71 (0.13) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   10 & 0.90 (0.01) & 0.88 (0.01) & 0.67 (0.15) & 0.76 (0.03) & 0.90 (0.01) & 0.82 (0.04) \\ 
-   50 & 0.88 (0.01) & 0.80 (0.04) & 0.43 (0.17) & 0.57 (0.19) & 0.86 (0.02) & 0.82 (0.04) \\ 
-  100 & 0.80 (0.07) & 0.73 (0.06) & 0.41 (0.19) & 0.51 (0.22) & - & 0.81 (0.04) \\ 
-  200 & 0.72 (0.10) & 0.77 (0.02) & 0.33 (0.15) & 0.44 (0.19) & - & 0.82 (0.04) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample $R^2$ measure for Gaussian outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:gaus}
-\end{table}
+\input{Tabs/sim_gaus_tab.tex}
 \clearpage
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllllll}
-  \hline
-P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-  4 & 0.94 (0.01) & 0.94 (0.02) & 0.90 (0.02) & 0.90 (0.01) & 0.94 (0.01) & 0.93 (0.01) \\ 
-   10 & 0.93 (0.01) & 0.89 (0.02) & 0.85 (0.04) & 0.86 (0.03) & 0.92 (0.04) & 0.92 (0.01) \\ 
-   50 & 0.92 (0.01) & 0.89 (0.01) & 0.83 (0.02) & 0.83 (0.02) & 0.76 (0.04) & 0.92 (0.01) \\ 
-  100 & 0.89 (0.02) & 0.86 (0.02) & 0.83 (0.02) & 0.84 (0.02) & - & 0.92 (0.01) \\ 
-  200 & 0.88 (0.01) & 0.86 (0.02) & 0.82 (0.05) & 0.81 (0.08) & - & 0.92 (0.01) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of the out-of-sample area under the curve measures for binomial outcomes over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:bin_auc}
-\end{table}
+\input{Tabs/sim_binom_tab.tex}
 
 \clearpage
-
-\begin{sidewaystable}[!h]
-\centering
-\begin{tabular}{lrllllll}
-  \hline
-Distribution & P & BHAM-IWLS & BHAM-CD & COSSO & Adaptive COSSO & mgcv & SB-GAM \\ 
-  \hline
-Binomial & 4.00 & 5.88 (1.01) & 6.96 (1.76) & 2.97 (0.82) & 4.60 (2.32) & 0.18 (0.04) & 343.24 (88.37) \\ 
-  Binomial & 10.00 & 14.01 (1.86) & 7.65 (1.25) & 7.95 (2.70) & 10.01 (3.56) & 3.51 (12.03) & 543.13 (133.51) \\ 
-  Binomial & 50.00 & 185.26 (17.44) & 53.83 (6.26) & 109.72 (24.31) & 128.13 (23.26) & 670.01 (151.28) & 1630.42 (193.61) \\ 
-  Binomial & 100.00 & 1106.93 (356.73) & 48.41 (5.84) & 715.15 (170.55) & 718.63 (193.53) & - & 2783.99 (235.31) \\ 
-  Binomial & 200.00 & 6923.62 (1911.59) & 99.98 (7.05) & 5572.51 (1295.18) & 4958.69 (1970.46) & - & 4780.10 (488.66) \\ 
-  Gaussian & 4.00 & 4.22 (1.51) & 2.51 (0.20) & 0.79 (0.11) & 0.67 (0.07) & 0.05 (0.00) & 36.02 (3.02) \\ 
-  Gaussian & 10.00 & 17.59 (5.77) & 33.32 (3.50) & 3.49 (0.70) & 3.47 (0.71) & 0.32 (0.34) & 72.93 (9.52) \\ 
-  Gaussian & 50.00 & 329.83 (44.57) & 334.53 (22.33) & 35.54 (9.15) & 35.44 (9.02) & 73.33 (71.24) & 373.42 (34.93) \\ 
-  Gaussian & 100.00 & 1706.19 (225.92) & 446.98 (79.31) & 150.58 (46.72) & 152.82 (45.07) & - & 684.05 (67.80) \\ 
-  Gaussian & 200.00 & 16379.27 (4825.56) & 303.73 (87.27) & 591.11 (147.88) & 569.92 (107.45) & - & 1314.12 (137.85) \\ 
-   \hline
-\end{tabular}
-\caption{The average and standard deviation of computation time in seconds, including cross-validation and final model fitting, over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM).
-mgcv doesn't provide estimation whe number of parameters exceeds sample size i.e.
-p = 100, 200.}
-\label{tab:time_sim}
-\end{sidewaystable}
 

### 2022-02-22 Manuscript/03-Simulation.Rmd
Revise Simulation Prediction Results (#23)

@@
 Generalized additive models with automatic smoothing, hereafter \textit{mgcv}, is one of the most popular models for nonlinear effect interpolation and prediction.
-SB-GAM is the first spike-and-slab lasso GAM.
+\bg{mgcv doesn't support analyses where the number of parameters exceeds the sample size.} SB-GAM is the first spike-and-slab lasso GAM.
 We implement COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \texttt{sparseGAM 1.0}.
@@
 \subsubsection{\bg{Prediction Performance}}
-The predictive performances have a consistent pattern across the two distributions of outcomes.
-Across all the scenarios, COSSO and adaptive COSSO have the least favorable performance among the applicable methods examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}).
-To note, mgcv doesn't support high-dimensional analysis, i.e.
-the number of coefficients are greater than the sample size, and hence not evaluated when $p=100, 200$.
-mgcv predicts well when $p$ is small or moderate ($p = 4, 10$), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when the dimension of are moderate (p=4, 10, 50).
-Particularly, BHAM-IWLS performs as good as mgcv if not better.
-However, in high dimensional case where we mimic the situation the signals are extremely sparse, SB-GAM has better performance than the proposed method.
+\bg{Among the main set of simulations where the functional forms of the predictors are mostly non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
+(See Tables \ref{tab:gaus} and \ref{tab:bin_auc}) \bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional spline models.
+Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
+When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse [TODO: re-evaluate after tuning the models] in high-dimensional cases ($p=50,100, 200$).
+As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv still fit models with increasing in large dimension case ($p=50$).
+But the algorithm fails to converge within the default number of iterations, and hence leads to bad prediction performance.} In high dimensional cases where we mimic the situation the signals are extremely sparse, SB-GAM has better performance than the proposed method.
 However, the BHAM-CD has extreme computational advantage over SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
 
-\begin{itemize}
-\item \bg{When the linear assumptions are violated, the linear lasso model performs extremely bad. adaptive cosso have improvement upon cosso.}
-\item \bg{To note, the drastic deteriation in the prediction of mgcv is caused by the model would not converge in those scenarios. A null model were returned as a final model. }
-\item \bg{Within the simulations where the functional form of the predictors are linear, we see that the proposed model similar to the lasso model regardless of the type outcome, and better moderately better than the model of comparison.
-(See Table Lnr Gause, and supporting table Lnr-binom.) Particularly, in gaussina simulations, the proposed model shows great advantage compared to SB-GAM (numeralize the advantage), and slightly improvement compare to ssGAM (numeralize the advantage).
-This shows that the model can be applied without constraining to the assumption where all functional form are non-linear.}
-\end{itemize}
+\bg{We also examine the prediction performance when the functional form of predictors are linear.
+(See support information Table 1, and 2 [TODO: add hyperlink to the github repo]) The proposed model, BHAM, has similar performance as the linear lasso model regardless of the type outcome.
+This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
+spikeSlabGAM have similar prediction performance to BHAM (numeralize the advantage).
+Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.
+(numeralize the advantage)}
 

### 2022-02-22 Manuscript/SS_GAM.tex
Revise Simulation Prediction Results (#23)

@@
 most popular models for nonlinear effect interpolation and prediction.
+\textcolor{red}{mgcv doesn't support analyses where the number of parameters exceeds the sample size.}
 SB-GAM is the first spike-and-slab lasso GAM. We implement COSSO and
@@
 
-The predictive performances have a consistent pattern across the two
-distributions of outcomes. Across all the scenarios, COSSO and adaptive
-COSSO have the least favorable performance among the applicable methods
-examined (See Table \ref{tab:gaus} and \ref{tab:bin_auc}). To note, mgcv
-doesn't support high-dimensional analysis, i.e.~the number of
-coefficients are greater than the sample size, and hence not evaluated
-when \(p=100, 200\). mgcv predicts well when \(p\) is small or moderate
-(\(p = 4, 10\)), and deteriorate when the number of predictors increase.
-Among the three fast-computing Bayesian hierarchical models, the
-proposed models, BHAM-IWLS and BHAM-CD predicts better than SB-GAM when
-the dimension of are moderate (p=4, 10, 50). Particularly, BHAM-IWLS
-performs as good as mgcv if not better. However, in high dimensional
-case where we mimic the situation the signals are extremely sparse,
-SB-GAM has better performance than the proposed method. However, the
-BHAM-CD has extreme computational advantage over SB-GAM (see Table
-\ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
-
-\begin{itemize}
-\item \textcolor{red}{When the linear assumptions are violated, the linear lasso model performs extremely bad. adaptive cosso have improvement upon cosso.}
-\item \textcolor{red}{To note, the drastic deteriation in the prediction of mgcv is caused by the model would not converge in those scenarios. A null model were returned as a final model. }
-\item \textcolor{red}{Within the simulations where the functional form of the predictors are linear, we see that the proposed model similar to the lasso model regardless of the type outcome, and better moderately better than the model of comparison.
-(See Table Lnr Gause, and supporting table Lnr-binom.) Particularly, in gaussina simulations, the proposed model shows great advantage compared to SB-GAM (numeralize the advantage), and slightly improvement compare to ssGAM (numeralize the advantage).
-This shows that the model can be applied without constraining to the assumption where all functional form are non-linear.}
-\end{itemize}
+\textcolor{red}{Among the main set of simulations where the functional forms of the predictors are mostly non-linear,}
+the predictive performances have a consistent pattern across the two
+distributions of outcomes. (See Tables \ref{tab:gaus} and
+\ref{tab:bin_auc})
+\textcolor{red}{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional spline models.
+Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
+When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse [TODO: re-evaluate after tuning the models] in high-dimensional cases ($p=50,100, 200$).
+As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv still fit models with increasing in large dimension case ($p=50$).
+But the algorithm fails to converge within the default number of iterations, and hence leads to bad prediction performance.}
+In high dimensional cases where we mimic the situation the signals are
+extremely sparse, SB-GAM has better performance than the proposed
+method. However, the BHAM-CD has extreme computational advantage over
+SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the
+prediction accuracy.
+
+\textcolor{red}{
+We also examine the prediction performance when the functional form of predictors are linear.
+(See support information Table 1, and 2 [TODO: add hyperlink to the github repo]) The proposed model, BHAM, has similar performance as the linear lasso model regardless of the type outcome.
+This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
+spikeSlabGAM have similar prediction performance to BHAM (numeralize the advantage).
+Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.
+(numeralize the advantage)}
 

### 2022-02-22 R/Sim_Res/compare_methods.R
Add functions to create percentage change, see example in _targets (percentage of change)

@@ -1,0 +1,29 @@
+compare_methods <- function(dat, method_1, method_2, measures){
+  # browser()
+  if(!all(c(method_1, method_2) %in% unique(dat$method)))
+    stop("method_1, method_2 don't exist in data. Check spelling")
+  
+
+  
+  tmp <- dat %>% filter(
+    method %in% c(method_1, method_2)
+  ) %>% 
+  select(p, dist, it, method, all_of({{measures}}))
+  
+  if(tmp %>% 
+     pull(all_of({{measures}})) %>% 
+     is.na() %>% 
+     all())
+  stop("Meausure does not exist in data")
+  
+  tmp %>% pivot_wider(
+    names_from = method,
+    values_from = {{measures}}
+  ) %>% 
+    rename(m1 := {{method_1}},
+           m2 := {{method_2}}) %>% 
+    mutate(ratio = (m1-m2)/m2) %>%
+    summarize(across(ratio, list(median = median, IQR = IQR), .names="{.fn}"))
+  
+   
+}

### 2022-02-22 R/Sim_Res/create_sim_pred_raw_dat.R

@@ -1,0 +1,64 @@
+create_raw_data <- function(success_rate, section = "test"){
+  # browser()
+  
+  total_dat <- success_rate %>% 
+    filter( n_success!=0) %>%
+    pull(path) %>% 
+    map_dfr( .f = function(sim){
+      # browser()
+      sim.df <- unglue_data(sim, 
+                            "{}/{}-dis_{dist}-p_{p}") %>% 
+        mutate(p = as.numeric(p))
+      
+      # fls <- 
+      list.files(sim, full.names = TRUE) %>% 
+        grep(".rds", x=., value = TRUE) %>%
+        map_dfr(.f = function(.file){
+          # browser()
+          it <- unglue_data(.file, "{whatever}/it_{it}.rds") %>% pull(it) %>% as.numeric
+          ret <- data.frame(
+            it = it,
+            read_rds(.file) %>% 
+              `[[`(section)%>% 
+              data.frame() %>% 
+              rownames_to_column("method")
+            
+          )
+        }) %>%
+        arrange(it) %>% 
+        mutate(dist = sim.df$dist,
+               p = sim.df$p)
+    }) %>% 
+    mutate(method = factor(
+      method#,
+      # levels = c("bglm_t", "bglm_de", "blasso",
+      #            "bglm_t_group", "bglm_de_group", "blasso_group",
+      #            "bglm_spline_t", "bglm_spline_de", "blasso_spline",
+      #            "cosso", "acosso", "mgcv", "SB_GAM"
+      # )
+    )
+    )
+  # browser()
+  # tmp <- total_dat %>% 
+  #   select(p, method, {{measures}}
+  #          # {{measures}} := paste(section, measures, sep=".")
+  #   ) #%>% #head
+  
+    # filter(method %in% c("bglm_spline_de", "blasso_spline",
+    #                      "cosso", "acosso", "mgcv", "SB_GAM")) %>% 
+    # group_by(p, method) %>% 
+    # summarize(
+      # across({{measures}}, 
+    #          list(mean = mean, sd = sd), na.rm = T,
+    #          .names = "{.fn}")#,
+    #   # n_miss = sum(is.na(auc))
+    # ) %>% 
+    # ungroup() %>% 
+    # transmute(
+    #   p, method,
+    #   {{measures}} := sprintf("%.2f (%.2f)", mean, sd)#,
+    #   # misclass = sprintf("%.2f (%.2f)", misclass_mean, misclass_sd)
+    # ) %>% 
[…]

### 2022-02-22 _targets.R

@@ -27,6 +27,7 @@ tar_plan(
             list.files("/data/user/boyiguo1/bgam/sim_res_20220220/main", recursive = T, full.names = T)),
   sim_success_rate = create_success_rate_table(sim_main_path),
   sim_mdl_fail = create_mdl_fail_rate(sim_success_rate),
+  sim_pred_raw = create_raw_data(sim_success_rate),
   
   #** Binomial Outcome  ####
   #*** Prediction  ####
@@ -44,6 +45,11 @@ tar_plan(
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:bin_auc")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_tab.tex"),
+  
+  # Percentage of change
+  sim_binom_pred_bham_lasso = sim_pred_raw %>% filter(dist == "binomial") %>%
+    compare_methods(method_1 = "bamlasso", method_2 = "lasso", measures="auc") %>% 
+    with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   
   #*** Variable Selection  ####
   sim_binom_var_select_raw = sim_var_select_raw(sim_success_rate,
@@ -90,6 +96,11 @@ tar_plan(
     whe number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:gaus") %>% 
     cat(file = "Manuscript/Tabs/sim_gaus_tab.tex"),
+  
+  # Percentage of change
+  sim_gaus_pred_bham_lasso = sim_pred_raw %>% filter(dist == "gaussian") %>%
+    compare_methods(method_1 = "bamlasso", method_2 = "lasso", measures="R2") %>% 
+    with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   #*** Variable Selection  ####
   sim_gaus_var_select_raw = sim_var_select_raw(sim_success_rate,
                                                dist = "gaussian"),

### 2022-02-23 Simulation/Code/Main.R
Expreiment with fine-tuning the proposed model (#35)

@@ -11,7 +11,7 @@ if(length(args)==0){
 # Arguments
 # n_train <- 500
 # n_test <- 1000
-# p <- 4
+# p <- 200
 # it <- 10
 # # dis <- "gaussian"
 # dis <- "binomial"
@@ -30,11 +30,7 @@ library(glmnet)
 library(spikeSlabGAM)
 
 
-# Helper functions
-source("~/GitHub/Manuscript-BHAM/Simulation/Code/helper_func.R")
 
-# Simulation Fixed Parameters and Functions
-source("~/GitHub/Manuscript-BHAM/Simulation/Code/sim_pars_funs.R")
 
 # Using Array ID as seed ID
 it <- Sys.getenv('SLURM_ARRAY_TASK_ID') %>% as.numeric
@@ -43,6 +39,12 @@ it <- Sys.getenv('SLURM_ARRAY_TASK_ID') %>% as.numeric
 set.seed(it)
 fam_fun <- dis()  
 # fam_fun <- str2expression(paste0(dis,"()")) %>% eval
+
+# Helper functions
+source("~/GitHub/Manuscript-BHAM/Simulation/Code/helper_func.R")
+
+# Simulation Fixed Parameters and Functions
+source("~/GitHub/Manuscript-BHAM/Simulation/Code/sim_pars_funs.R")
 
 # Train Data
 tmp <- sim_Bai(n_train, p, family = fam_fun)
@@ -115,14 +117,14 @@ if(!is.null(mgcv_mdl)){
 
 
 
-#### BGAMs ####
+
+#### Fit BH Models ####
 train_sm_dat <- BHAM::construct_smooth_data(mgcv_df, dat)
 train_smooth <- train_sm_dat$Smooth
 train_smooth_data <- train_sm_dat$data
 test_sm_dat <- BHAM::make_predict_dat(train_sm_dat$Smooth, dat = test_dat)
 
 
-#### Fit BH Models ####
 
 func_list <- list(
   # "bglm", "bglm", "bmlasso", 
@@ -174,6 +176,7 @@ mdls <- pmap(list(func_list, group_list, prior), .f = function(.fun, .group, .pr
   # mdl <- call(.fun, x = train_smooth_data, y = dat$y, family = fam_fun$family, group = .group,
   #             ss = c(0.0011, 0.5)) %>% eval
   
+  # browser()
   
[…]

### 2022-02-23 Simulation/Code/sim_pars_funs.R

@@ -10,14 +10,14 @@ bamlasso_seq <- bgam_seq <- seq(0.001, 0.05, length.out = 20)
 
 
 # 
-# if(fam_fun$family == "binomial"){
-#   bamlasso_seq <- case_when(
-#     p == 4 ~ seq(from = 0.01, to = 0.8, length.out = 20),
-#     p == 10 ~ seq(0.01, 0.3, length.out = 20),
-#     p == 50 ~  seq(0.01, 0.3, length.out = 20),
-#     p == 100 ~ seq(0.005, 0.10, by = 0.005),
-#     p == 200 ~ seq(0.005, by = 0.0025, length.out = 20)
-#   )
+if(fam_fun$family == "binomial"){
+  bamlasso_seq <- case_when(
+    p == 4 ~ seq(from = 0.01, to = 0.8, length.out = 20),
+    p == 10 ~ seq(0.01, 0.3, length.out = 20),
+    p == 50 ~  seq(0.005, 0.1, length.out = 20),
+    p == 100 ~ seq(0.005, 0.10, by = 0.005),
+    p == 200 ~ seq(0.001, 0.03, length.out = 20)
+  )
 #   
 #   bgam_seq <- case_when(
 #     p == 4 ~ seq(from = 0.005, to = 0.1, length.out = 20),
@@ -29,7 +29,7 @@ bamlasso_seq <- bgam_seq <- seq(0.001, 0.05, length.out = 20)
 #   
 #   
 #   
-# } else if(fam_fun$family == "gaussian"){
+} #else if(fam_fun$family == "gaussian"){
 #   bamlasso_seq <- case_when(
 #     p == 4 ~ seq(from = 0.08, to = 0.25, length.out = 20),
 #     p == 10 ~ seq(0.005, 0.03, length.out = 20),

### 2022-02-23 Manuscript/Tabs/sim_time_tab.tex
Format method simulation time table (#33)

\begin{sidewaystable}[!h]
\centering
\begin{tabular}{lrllllll}
  \hline
Distribution & P & mgcv & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
Binomial &   4 & 0.18 (0.04) & 3.16 (1.39) & 5.51 (4.07) & 2.73 (0.22) & 347.17 (89.43) & 8.41 (0.91) \\ 
  Binomial &  10 & 3.46 (11.06) & 8.30 (1.70) & 10.66 (5.30) & 4.08 (0.29) & 539.05 (135.55) & 20.36 (2.16) \\ 
  Binomial &  50 & 660.31 (141.53) & 103.41 (20.00) & 118.82 (18.45) & 14.22 (0.58) & 1590.09 (142.19) & 236.73 (14.83) \\ 
  Binomial & 100 & - & 662.61 (125.00) & 672.65 (185.09) & 31.61 (2.56) & 2720.53 (250.43) & 967.97 (186.85) \\ 
  Binomial & 200 & - & 5325.66 (995.60) & 4963.93 (1482.11) & 82.17 (3.29) & 4788.76 (420.64) & 3371.88 (194.02) \\ 
  Gaussian &   4 & 0.05 (0.01) & 0.75 (0.09) & 0.75 (0.11) & 8.78 (1.57) & 38.82 (2.74) & 1.84 (0.18) \\ 
  Gaussian &  10 & 0.32 (0.39) & 3.42 (0.24) & 3.41 (0.23) & 20.77 (3.95) & 76.12 (5.55) & 5.93 (0.57) \\ 
  Gaussian &  50 & 72.03 (57.99) & 33.98 (2.88) & 34.35 (2.86) & 285.73 (12.53) & 374.76 (23.79) & 65.18 (8.12) \\ 
  Gaussian & 100 & - & 117.79 (3.33) & 119.63 (3.66) & 372.01 (56.92) & 640.44 (21.91) & 194.14 (8.09) \\ 
  Gaussian & 200 & - & 518.86 (40.78) & 524.76 (39.15) & 471.46 (72.23) & 1300.70 (72.74) & 738.52 (62.76) \\ 
   \hline
\end{tabular}
\caption{The average and standard deviation of computation time in seconds,
      including cross-validation and final model fitting, over 50 iterations. The models
      of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted
      with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD)
      algorithms, component selection and smoothing operator (COSSO), adaptive COSSO,
      mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't 
      provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 
\label{tab:time_sim}
\end{sidewaystable}

### 2022-02-23 R/Sim_Res/make_time_table.R

@@ -1,7 +1,5 @@
 make_time_table <- function(success_rate){
   total_dat <- success_rate %>% 
-    # filter(dist==dist, n_success!=0) %>% 
-    # filter(dist=={{dist}}, n_success!=0) %>%
     filter(n_success!=0) %>% 
     pull(path) %>% 
     map_dfr( .f = function(sim){
@@ -12,14 +10,6 @@ make_time_table <- function(success_rate){
         mutate( p = as.numeric(p))
       fls <- list.files(sim, full.names = TRUE) 
       fls <- fls[grep(".rds", x=fls)]
-      # n <- length(fls)
-      # 
-      # 
-      # if(n == 0)
-      #   return(data.frame(NULL))
-      
-      # .file <- fls[1]
-      # ret <- fls 
       list.files(sim, full.names = TRUE) %>%
         grep(".rds", x=., value = TRUE) %>%
         map_dfr(.f = function(.file){
@@ -30,25 +20,25 @@ make_time_table <- function(success_rate){
               `[[`("time")%>% 
               data.frame() %>% 
               rownames_to_column("method")
-            
           )
         }) %>%
         arrange(it) %>% 
         mutate(p = sim.df$p,
-               dist = sim.df$dist)#  %>% 
+               dist = sim.df$dist)
       
     }) %>% 
     select(dist, p, it, method, elapsed) %>%
     pivot_wider(names_from = method, values_from = elapsed) %>% 
     transmute(dist = case_when(dist=="binomial"~"Binomial",
                                dist=="gaussian"~"Gaussian"),
-              p, it, mgcv, cosso, acosso,
-              `BHAM-CD` = blasso_spline_cv + blasso_spline_final,
-              `BHAM-IWLS` = bglm_spline_de_cv + bglm_spline_de_final,
-              `SB-GAM` = bai_cv + bai_final) %>% 
+              p = as.integer(p), it, mgcv, cosso, acosso,
+              `BHAM` = bamlasso_cv + bamlasso_final,
+              # `BHAM-IWLS` = bglm_spline_de_cv + bglm_spline_de_final,
+              `SB-GAM` = bai_cv + bai_final,
+              `spikeSlabGAM` = ssGAM) %>% 
     group_by(dist, p) %>%  
     summarize(
-      across(mgcv:`SB-GAM`, ~sprintf("%.2f (%.2f)", mean(.x), sd(.x))
[…]

### 2022-02-23 _targets.R

@@ -105,6 +105,8 @@ tar_plan(
   sim_gaus_var_select_raw = sim_var_select_raw(sim_success_rate,
                                                dist = "gaussian"),
   # plot_var_select(sim_gaus_var_select_raw[[3]])
+  
+  #*** Bi-level Selection  ####
   sim_gaus_comp_select_raw = sim_comp_select_raw(sim_success_rate,
                                                  dist = "gaussian"),
   # sim_gaus_var_select = make_sim_var_select_table(sim_success_rate,
@@ -115,8 +117,30 @@ tar_plan(
   #   summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
   #   ) %>%
   #   pivot_wider(names_from = Method, values_from = mean),
-  #*** Bi-level Selection  ####
-  # sim_tim_tab = make_time_table(sim_success_rate),
+
+  #** Simulation Time ####
+  sim_tim_tab = make_time_table(sim_success_rate),
+  sim_tim_tab_latex = sim_tim_tab %>% 
+    rename(Distribution = dist,
+           P = p,
+           COSSO = cosso,
+           `Adaptive COSSO` = acosso) %>% 
+    xtable(
+      caption = "The average and standard deviation of computation time in seconds,
+      including cross-validation and final model fitting, over 50 iterations. The models
+      of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted
+      with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD)
+      algorithms, component selection and smoothing operator (COSSO), adaptive COSSO,
+      mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't 
+      provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
+      label = "tab:time_sim",
+      booktabs = TRUE
+    ) %>% 
+    print(comment = FALSE, include.rownames = FALSE,
+          floating.environment = "sidewaystable",# caption.placement = "top"
+          table.placement = "!h"
+    ) %>% 
+    cat(file = "Manuscript/Tabs/sim_time_tab.tex"),
   
   
   #* Linear Simulation Study ####

### 2022-02-23 Simulation/Code/sim_pars_funs.R
Use a general value for s0 when tuning. (#35)

@@ -10,14 +10,14 @@ bamlasso_seq <- bgam_seq <- seq(0.001, 0.05, length.out = 20)
 
 
 # 
-if(fam_fun$family == "binomial"){
-  bamlasso_seq <- case_when(
-    p == 4 ~ seq(from = 0.01, to = 0.8, length.out = 20),
-    p == 10 ~ seq(0.01, 0.3, length.out = 20),
-    p == 50 ~  seq(0.005, 0.1, length.out = 20),
-    p == 100 ~ seq(0.005, 0.10, by = 0.005),
-    p == 200 ~ seq(0.001, 0.03, length.out = 20)
-  )
+# if(fam_fun$family == "binomial"){
+#   bamlasso_seq <- case_when(
+#     p == 4 ~ seq(from = 0.01, to = 0.8, length.out = 20),
+#     p == 10 ~ seq(0.01, 0.3, length.out = 20),
+#     p == 50 ~  seq(0.005, 0.1, length.out = 20),
+#     p == 100 ~ seq(0.005, 0.10, by = 0.005),
+#     p == 200 ~ seq(0.001, 0.03, length.out = 20)
+#   )
 #   
 #   bgam_seq <- case_when(
 #     p == 4 ~ seq(from = 0.005, to = 0.1, length.out = 20),
@@ -29,7 +29,7 @@ if(fam_fun$family == "binomial"){
 #   
 #   
 #   
-} #else if(fam_fun$family == "gaussian"){
+# } else if(fam_fun$family == "gaussian"){
 #   bamlasso_seq <- case_when(
 #     p == 4 ~ seq(from = 0.08, to = 0.25, length.out = 20),
 #     p == 10 ~ seq(0.005, 0.03, length.out = 20),

### 2022-02-25 Manuscript/10-Supplementary.Rmd
Add latex tables to the manuscript
Update supporting document title and authors

@@
 ---
-title: "Supporting Infromation"
-output: pdf_document
+title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
+subtitle: "Supporting Infromation"
+author: Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi
+output: 
+  pdf_document:
+    keep_tex: true
 ---
@@
 
+\clearpage
+
 \subsection{Supplementary Information 3: Predictive Performance of Linear Simulations}
-\input{Manuscript/Tabs/sim_lnr_gaus_tab.tex}
-\clearpage
-\input{Manuscript/Tabs/sim_lnr_binom_tab.tex}
+\input{Tabs/sim_lnr_gaus_tab.tex}
+\input{Tabs/sim_lnr_binom_tab.tex}

### 2022-02-25 Simulation/Code/sim_pars_funs.R
Fine tuning for Gaussian p=200 (Implemented & now commented out)(#35)

@@ -29,8 +29,12 @@ bamlasso_seq <- bgam_seq <- seq(0.001, 0.05, length.out = 20)
 #   
 #   
 #   
-# } else if(fam_fun$family == "gaussian"){
-#   bamlasso_seq <- case_when(
+# } else 
+# if(fam_fun$family == "gaussian"){
+#   if(p==200)
+#     bamlasso_seq <- seq(0.00025, to=0.0025, length.out = 20)
+# }
+      # case_when(
 #     p == 4 ~ seq(from = 0.08, to = 0.25, length.out = 20),
 #     p == 10 ~ seq(0.005, 0.03, length.out = 20),
 #     p == 50 ~ seq(0.0005, 0.01, length.out = 20),

### 2022-02-25 Manuscript/10-Supplementary.tex

@@
 ]{article}
-\title{Supporting Infromation}
-\author{}
-\date{\vspace{-2.5em}}
-
-\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
+\usepackage{amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \hypersetup{
-  pdftitle={Supporting Infromation},
+  pdftitle={Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data},
+  pdfauthor={Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi},
   hidelinks,
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
+\ifluatex
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
+
+\title{Spike-and-Slab Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Supporting Infromation}
+\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
+Nengjun Yi}
+\date{}
 

### 2022-02-25 Manuscript/Tabs/sim_binom_var_slct_tab.tex
Create Variable Selection Tables (#22)

\begin{table}[ht]
\centering
\begin{tabular}{ccccccccc}
  \hline
P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.88 & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.49 \\ 
   50 & Precision &  & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.04 \\ 
  100 & Precision &  & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.02 \\ 
  200 & Precision &  & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.03 \\ 
    4 & Recall & 1.00 & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
   10 & Recall & 0.92 & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.31 \\ 
   50 & Recall & 0.00 & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.07 \\ 
  100 & Recall &  & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.06 \\ 
  200 & Recall &  & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.05 \\ 
    4 & MCC &  &  &  &  &  &  &  \\ 
   10 & MCC & 0.89 & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.10 \\ 
   50 & MCC &  & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & -0.04 \\ 
  100 & MCC &  & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & -0.01 \\ 
  200 & MCC &  & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.01 \\ 
   \hline
\end{tabular}
\caption{} 
\label{}
\end{table}

### 2022-02-25 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
Create Variable Selection Tables (#22)

\begin{table}[ht]
\centering
\begin{tabular}{ccccccccc}
  \hline
P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.89 & 0.76 & 0.84 & 0.93 & 0.62 & 0.86 & 0.50 \\ 
   50 & Precision & 0.36 & 0.48 & 0.70 & 0.69 & 0.88 & 0.75 & 0.00 \\ 
  100 & Precision &  & 0.43 & 0.61 & 0.59 & 0.99 & 0.79 & 0.00 \\ 
  200 & Precision &  & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.00 \\ 
    4 & Recall & 1.00 & 0.53 & 0.49 & 0.53 & 0.88 & 0.99 & 0.51 \\ 
   10 & Recall & 1.00 & 0.40 & 0.52 & 0.52 & 0.83 & 1.00 & 0.25 \\ 
   50 & Recall & 1.00 & 0.35 & 0.40 & 0.48 & 0.37 & 1.00 & 0.00 \\ 
  100 & Recall &  & 0.33 & 0.36 & 0.40 & 0.30 & 0.99 & 0.00 \\ 
  200 & Recall &  & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.00 \\ 
    4 & MCC &  &  &  &  &  &  &  \\ 
   10 & MCC & 0.90 & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.10 \\ 
   50 & MCC & 0.54 & 0.31 & 0.47 & 0.53 & 0.50 & 0.83 & -0.06 \\ 
  100 & MCC &  & 0.32 & 0.41 & 0.45 & 0.53 & 0.87 & -0.03 \\ 
  200 & MCC &  & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & -0.01 \\ 
   \hline
\end{tabular}
\caption{} 
\label{}
\end{table}

### 2022-02-25 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
Create Variable Selection Tables (#22)

\begin{table}[ht]
\centering
\begin{tabular}{ccccccccc}
  \hline
P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.94 & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.69 \\ 
   50 & Precision &  & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.19 \\ 
  100 & Precision &  & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.15 \\ 
  200 & Precision &  & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.10 \\ 
    4 & Recall & 1.00 & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
   10 & Recall & 1.00 & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 0.77 \\ 
   50 & Recall & 0.00 & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 0.33 \\ 
  100 & Recall &  & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.27 \\ 
  200 & Recall &  & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.28 \\ 
    4 & MCC &  &  &  &  &  &  &  \\ 
   10 & MCC & 0.94 & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.52 \\ 
   50 & MCC &  & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.15 \\ 
  100 & MCC &  & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.15 \\ 
  200 & MCC &  & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.13 \\ 
   \hline
\end{tabular}
\caption{} 
\label{}
\end{table}

### 2022-02-25 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
Create Variable Selection Tables (#22)

\begin{table}[ht]
\centering
\begin{tabular}{ccccccccc}
  \hline
P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
  \hline
  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
   10 & Precision & 0.91 & 0.59 & 0.97 & 0.97 & 0.90 & 0.56 & 0.74 \\ 
   50 & Precision & 0.39 & 0.43 & 0.74 & 0.84 & 0.89 & 0.39 & 0.25 \\ 
  100 & Precision &  & 0.34 & 0.53 & 0.72 & 0.91 & 0.29 & 0.25 \\ 
  200 & Precision &  & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.25 \\ 
    4 & Recall & 1.00 & 1.00 & 0.87 & 0.84 & 1.00 & 1.00 & 1.00 \\ 
   10 & Recall & 1.00 & 1.00 & 0.98 & 0.96 & 1.00 & 1.00 & 0.75 \\ 
   50 & Recall & 1.00 & 1.00 & 0.85 & 0.99 & 1.00 & 1.00 & 0.25 \\ 
  100 & Recall &  & 1.00 & 0.84 & 0.97 & 1.00 & 1.00 & 0.25 \\ 
  200 & Recall &  & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 0.25 \\ 
    4 & MCC &  &  &  &  &  &  &  \\ 
   10 & MCC & 0.91 & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.58 \\ 
   50 & MCC & 0.57 & 0.59 & 0.76 & 0.90 & 0.92 & 0.57 & 0.18 \\ 
  100 & MCC &  & 0.54 & 0.61 & 0.81 & 0.94 & 0.48 & 0.22 \\ 
  200 & MCC &  & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.23 \\ 
   \hline
\end{tabular}
\caption{} 
\label{}
\end{table}

### 2022-02-25 R/Sim_Res/format_var_slct_tbls.R

@@ -1,0 +1,28 @@
+format_var_slct_tbls <- function( dat, caption, label){
+  dat %>% 
+    mutate(Metric = factor(Metric, 
+                           levels = c("precesion", "recall", "mcc"),
+                           labels = c("Precision", "Recall", "MCC")),
+           p = as.integer(p)
+    )%>% 
+    arrange(Metric, p) %>% 
+    select(
+      "P" = p,
+      Metric,
+      "mgcv" = mgcv,
+      "LASSO" = lasso,
+      "COSSO" = cosso,
+      "Adaptive COSSO" = acosso,
+      # "BHAM-IWLS" = bglm_spline_de,
+      # "BHAM-CD" = bamlasso,
+      "BHAM" = bamlasso,      
+      "SB-GAM" = SBGAM,
+      "spikeSlabGAM" = ssGAM
+    ) %>% 
+    xtable(
+      align = "cccccccccc",
+      caption = caption,
+      label = label
+    ) %>% 
+    print(comment = FALSE, include.rownames = FALSE)
+}

### 2022-02-25 R/Sim_Res/sim_var_select.R

@@ -201,7 +201,8 @@ make_sim_var_metric_raw <- function(var_select){
         pivot_longer(cols = lasso_recall:ssGAM_mcc,
                      names_to = c("Method", "Metric"),
                      names_sep = "_") %>% 
-        mutate(It = it)
+        mutate(It = it,
+               p = p)
     })
   
   

### 2022-02-25 _targets.R

@@ -24,7 +24,7 @@ tar_plan(
   #  Simulations Studies -----------------------------------------------------
   #* Main Simulation Study ####
   tar_files(sim_main_path,
-            list.files("/data/user/boyiguo1/bgam/sim_res_20220220/main", recursive = T, full.names = T)),
+            list.files("/data/user/boyiguo1/bgam/sim_res_20220220/main/", recursive = T, full.names = T)),
   sim_success_rate = create_success_rate_table(sim_main_path),
   sim_mdl_fail = create_mdl_fail_rate(sim_success_rate),
   sim_pred_raw = create_raw_data(sim_success_rate),
@@ -55,22 +55,20 @@ tar_plan(
   sim_binom_var_select_raw = sim_var_select_raw(sim_success_rate,
                                                 dist = "binomial"),
   
+  # plot_var_select(sim_binom_var_select_raw[[1]])  
+
+  
   # Table Presentation of Var Selection
-  # make_sim_var_metric_raw(sim_binom_var_select_raw[[5]]) %>%
-  #   group_by(Method, Metric) %>%
-  #   summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
-  #             ) %>%
-  #   pivot_wider(names_from = Method, values_from = mean),
+  sim_binom_var_select_tbl = map_dfr(1:5, ~make_sim_var_metric_raw(sim_binom_var_select_raw[[.x]])) %>%
+    group_by(p, Method, Metric) %>%
+    summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
+              ) %>%
+    pivot_wider(names_from = Method, values_from = mean) %>% 
+    ungroup(),
   
-  
-  
-  # plot_var_select(sim_binom_var_select_raw[[1]])
-  sim_binom_comp_select_raw = sim_comp_select_raw(sim_success_rate,
-                                                  dist = "binomial"),
-  
-  # plot_comp_select(sim_binom_comp_select_raw[[4]])
-  
-  #*** Bi-level Selection  ####
+  sim_binom_var_select_tbl_latex = sim_binom_var_select_tbl %>% 
+    format_var_slct_tbls(caption = "", label = "")%>% 
+    cat(file = "Manuscript/Tabs/sim_binom_var_slct_tab.tex"),
[…]

### 2022-02-25 Manuscript/03-Simulation.Rmd
Revise Variable Selection Results (#22)

@@
 \bg{\subsubsection{Variable Selection Performance}
-While the main utility of the proposed work is prediction, the proposed model can also serve as variable selection/functional selection as a by-product.
-Hence We evaluate the variable selection performance.
+Within the simulation settings we examined, the proposed model tends to have a more conservative bebaviour in terms of variable selection, having high precision and low recall.
+(See Table xxx for gaussian outcomes, and support information xxx for binomial informaitons.) [TODO: mention the variable selection pattern is consistent across the types of outcomes.] THe high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high matthew correlation coefficient.
+Another Bayesian model, spike-and-slab performs bad in terms of variable selection in high-dimensional situations ($p=100,200$).
+We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
+
 }

### 2022-02-25 Manuscript/03-Simulation.Rmd
Revise Variable Selection Results + Add Tables (#22)

@@
 We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
-
 }
-
-\begin{itemize}
-\item \bg{If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
-The pattern is consitent regardless of the type of outcomes.
-To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
-Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.}
-\item \bg{If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
-While the recall is lower than SBGAM.
-This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.}
-\item \bg{We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while. On the contrary the SBGAM is not so well.}
-\end{itemize}
 

### 2022-02-25 Manuscript/SS_GAM.tex
Revise Variable Selection Results + Add Tables (#22)

@@
 
-\textcolor{red}{
-We also examine the prediction performance when the functional form of predictors are linear.
+\textcolor{red}{We also examine the prediction performance when the functional form of predictors are linear.
 (See support information Table 1, and 2 [TODO: add hyperlink to the github repo]) The proposed model, BHAM, has similar performance as the linear lasso model regardless of the type outcome.
@@
 \textcolor{red}{\subsubsection{Variable Selection Performance}
-While the main utility of the proposed work is prediction, the proposed model can also serve as variable selection/functional selection as a by-product.
-Hence We evaluate the variable selection performance.
+Within the simulation settings we examined, the proposed model tends to have a more conservative bebaviour in terms of variable selection, having high precision and low recall.
+(See Table xxx for gaussian outcomes, and support information xxx for binomial informaitons.) [TODO: mention the variable selection pattern is consistent across the types of outcomes.] THe high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high matthew correlation coefficient.
+Another Bayesian model, spike-and-slab performs bad in terms of variable selection in high-dimensional situations ($p=100,200$).
+We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
 }
-
-\begin{itemize}
-\item \textcolor{red}{If we use mcc as the measure, SBGAM has the best performance among all the method of comparison, followed by the proposed model BHAM.
-The pattern is consitent regardless of the type of outcomes.
-To note here, mgcv is not included in the performance table, as the mgcv isn't compatable with hign-dimensitonal data.
-Nevertheless, when the dimensionality is low, mgcv perfoms while for variable selection.}
-\item \textcolor{red}{If we look at the precision of the variable selection, BHAM performs better than SBGAM model particularaly when the dimensionality increases.
-While the recall is lower than SBGAM.
-This translate to the selection of BHAM is more conservative that SBGAM, such that among the selected variables, many of them are signals, however, due to the conversative selection, there are few variables are selected.}
-\item \textcolor{red}{We also want to high light that in the set of linear simulation, when the outcome is cgaussina, the proposed model performed extremely while.
-On the contrary the SBGAM is not so well.}
-\end{itemize}
 
@@
 \input{Tabs/sim_binom_tab.tex}
+\clearpage
+\input{Tabs/sim_gaus_var_slct_tab.tex}
 

### 2022-02-27 Manuscript/03-Simulation.Rmd

@@
 Another Bayesian model, spike-and-slab performs bad in terms of variable selection in high-dimensional situations ($p=100,200$).
-We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
-}
+We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.}
 

### 2022-02-28 _targets.R
Turn off renv

@@ -1,12 +1,13 @@
 ## Load your packages, e.g. library(targets).
 library(targets)
 library(tarchetypes)
-library(dotenv)
+# library(dotenv)
 
 tar_option_set(
-  packages = c("tidyverse", "dplyr", "rticles", "gtsummary", "mgcv", "gt", 
+  packages = c("tidyverse", "dplyr", "rticles", #"gtsummary", "gt", 
                "broom", "unglue", "knitr", "ggpubr", "xtable", "janitor", 
-               "flextable", "BHAM", "BhGLM", "sparseGAM",  "glmnet", "yardstick",
+               # "flextable", 
+               "mgcv", "BHAM", "BhGLM", "sparseGAM",  "glmnet", "yardstick",
                "rmarkdown"#, "cosso",
   ),
   imports = c( "BHAM", "BhGLM", "sparseGAM")
@@ -82,7 +83,7 @@ tar_plan(
 
   sim_binom_comp_select_raw = sim_comp_select_raw(sim_success_rate,
                                                   dist = "binomial"),
-  # plot_comp_select(sim_binom_comp_select_raw[[4]])
+  # plot_comp_select(sim_binom_comp_select_raw[[1]])
   
 
   
@@ -128,9 +129,10 @@ tar_plan(
   #*** Bi-level Selection  ####
   sim_gaus_comp_select_raw = sim_comp_select_raw(sim_success_rate,
                                                  dist = "gaussian"),
-  # sim_gaus_var_select = make_sim_var_select_table(sim_success_rate,
-  #                                                  dist = "gaussian"),
+
+  # plot_comp_select(sim_gaus_comp_select_raw[[5]]),
   
+  sim_gaus_comp_select_raw[[5]]$comp_select %>% filter(Variable %in% paste0("x",1:4)),
 
 
   #** Simulation Time ####
@@ -195,6 +197,12 @@ tar_plan(
     format_var_slct_tbls(caption = "", label = "")%>% 
     cat(file = "Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex"),
   
+  #*** Bi-level Selection  ####
+  sim_lnr_binom_comp_select_raw = sim_comp_select_raw(sim_lnr_success_rate,
+                                                     dist = "binomial"),
+  
+  # plot_comp_select(sim_lnr_binom_comp_select_raw[[2]]),
+  
   #** Gaussian Outcome  ####
[…]

### 2022-02-28 _targets_packages.R

@@ -3,11 +3,8 @@ library(BHAM)
 library(BhGLM)
 library(broom)
 library(dplyr)
-library(flextable)
 library(ggpubr)
 library(glmnet)
-library(gt)
-library(gtsummary)
 library(janitor)
 library(knitr)
 library(mgcv)

### 2022-02-28 Simulation/Code/Main.R
Correct simulation ssGAM variable selection results recording

@@ -11,8 +11,8 @@ if(length(args)==0){
 # Arguments
 # n_train <- 500
 # n_test <- 1000
-# p <- 200
-# it <- 10
+# p <- 50
+# it <- 1
 # # dis <- "gaussian"
 # dis <- "binomial"
 
@@ -393,7 +393,12 @@ select(part, prob = `P.gamma...1.`) %>%
   unglue::unglue_unnest(part, "{}({var})", remove = FALSE) %>% 
   group_by(var) %>% 
   summarize(ssGAM = any(prob>0.5)) %>% 
-  column_to_rownames("var") %>% as.vector
+  unglue::unglue_unnest(var, "x{id}", remove = FALSE) %>% 
+  mutate(id = as.integer(id)) %>% 
+  arrange(id) %>% 
+  column_to_rownames("var") %>% 
+  select(ssGAM) %>% 
+  as.vector
 
 
 ssGAM_vs_part <- (summary(ssGAM_mdl)$trmSummary) %>% data.frame() %>% rownames_to_column("part") %>% 
@@ -403,6 +408,7 @@ ssGAM_vs_part <- (summary(ssGAM_mdl)$trmSummary) %>% data.frame() %>% rownames_t
 ssGAM_train_msr <- measure.glm(dat$y, 
                                predict(ssGAM_mdl, type = "response") %>% rescale_ssGAM_pred(family = fam_fun$family, y=ssGAM_train_dat$y),
                                family = fam_fun$family)
+
 ssGAM_test_msr <- measure.glm(test_dat$y, 
                               predict(ssGAM_mdl, newdata = test_dat, type = "response") %>% rescale_ssGAM_pred(family = fam_fun$family, y=ssGAM_train_dat$y),
                               family = fam_fun$family)

### 2022-03-01 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
Update Variable Selection Tables (#39)

@@
   4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.89 & 0.76 & 0.84 & 0.93 & 0.62 & 0.86 & 0.50 \\ 
-   50 & Precision & 0.36 & 0.48 & 0.70 & 0.69 & 0.88 & 0.75 & 0.00 \\ 
-  100 & Precision &  & 0.43 & 0.61 & 0.59 & 0.99 & 0.79 & 0.00 \\ 
-  200 & Precision &  & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.00 \\ 
+   10 & Precision & 0.89 & 0.76 & 0.84 & 0.93 & 0.62 & 0.86 & 1.00 \\ 
+   50 & Precision & 0.36 & 0.48 & 0.70 & 0.69 & 0.88 & 0.75 & 1.00 \\ 
+  100 & Precision &  & 0.43 & 0.61 & 0.59 & 0.99 & 0.79 & 0.99 \\ 
+  200 & Precision &  & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.99 \\ 
     4 & Recall & 1.00 & 0.53 & 0.49 & 0.53 & 0.88 & 0.99 & 0.51 \\ 
-   10 & Recall & 1.00 & 0.40 & 0.52 & 0.52 & 0.83 & 1.00 & 0.25 \\ 
-   50 & Recall & 1.00 & 0.35 & 0.40 & 0.48 & 0.37 & 1.00 & 0.00 \\ 
-  100 & Recall &  & 0.33 & 0.36 & 0.40 & 0.30 & 0.99 & 0.00 \\ 
-  200 & Recall &  & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.00 \\ 
+   10 & Recall & 1.00 & 0.40 & 0.52 & 0.52 & 0.83 & 1.00 & 0.50 \\ 
+   50 & Recall & 1.00 & 0.35 & 0.40 & 0.48 & 0.37 & 1.00 & 0.50 \\ 
+  100 & Recall &  & 0.33 & 0.36 & 0.40 & 0.30 & 0.99 & 0.50 \\ 
+  200 & Recall &  & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.50 \\ 
     4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.90 & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.10 \\ 
-   50 & MCC & 0.54 & 0.31 & 0.47 & 0.53 & 0.50 & 0.83 & -0.06 \\ 
-  100 & MCC &  & 0.32 & 0.41 & 0.45 & 0.53 & 0.87 & -0.03 \\ 
-  200 & MCC &  & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & -0.01 \\ 
+   10 & MCC & 0.90 & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.61 \\ 
+   50 & MCC & 0.54 & 0.31 & 0.47 & 0.53 & 0.50 & 0.83 & 0.69 \\ 
+  100 & MCC &  & 0.32 & 0.41 & 0.45 & 0.53 & 0.87 & 0.70 \\ 
+  200 & MCC &  & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & 0.70 \\ 
    \hline

### 2022-03-01 _targets.R
Add Function Plotting for real data analysis

@@ -285,6 +285,14 @@ tar_plan(
   ECB_bamlasso_var = bamlasso_var_selection(ECB_bamlasso_fnl),
   # TODO: Use cv.bh to get the measures.
   
+  # *** Plot Non-Linear Functions ###
+  # TODO: to generalize this for a list of plots.
+  ECB_bamlasso_nonlnr = ECB_bamlasso_var$`Non-parametric`[[1]][2],
+  ECB_plot = plot_smooth_term(ECB_bamlasso_fnl, ECB_bamlasso_nonlnr, ECB_sm_obj$Smooth,
+                              min = min(ECB_cov[, ECB_bamlasso_nonlnr])-0.1,
+                              max = max(ECB_cov[, ECB_bamlasso_nonlnr]) + 0.1),
+  
+  
   # ** SB-GAM ---------------------------------------------------------------
   ECB_SBGAM_cv_raw = cv.SBGAM( X = ECB_cov,
                                y = ECB_outcome,
@@ -362,6 +370,15 @@ tar_plan(
                               group = WLM_bgam_dat$group),
   # measure.bh(WLM_bamlasso_fnl),
   WLM_bamlasso_var = bamlasso_var_selection(WLM_bamlasso_fnl),
+  
+  # *** Plot Non-Linear Functions ###
+  # TODO: to generalize this for a list of plots.
+  WLM_bamlasso_nonlnr = WLM_bamlasso_var$`Non-parametric`[[1]][1],
+  WLM_plot = plot_smooth_term(WLM_bamlasso_fnl, WLM_bamlasso_nonlnr, WLM_sm_obj$Smooth,
+                              min = min(WLM_bgam_dat$train_dat[, WLM_bamlasso_nonlnr])-0.1,
+                              max = max(WLM_bgam_dat$train_dat[, WLM_bamlasso_nonlnr]) + 0.1),
+  
+  
   
   # # ** BHAM-IWLS --------------------------------------------------------------
   # WLM_bgam_raw = bgam(out_HOMA_std~.-out_HOMA_std, 

### 2022-03-01 Manuscript/Tabs/sim_binom_var_slct_tab.tex

@@
   4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.88 & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.49 \\ 
-   50 & Precision &  & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.04 \\ 
-  100 & Precision &  & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.02 \\ 
-  200 & Precision &  & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.03 \\ 
+   10 & Precision & 0.88 & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.89 \\ 
+   50 & Precision &  & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.52 \\ 
+  100 & Precision &  & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.42 \\ 
+  200 & Precision &  & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.36 \\ 
     4 & Recall & 1.00 & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
-   10 & Recall & 0.92 & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.31 \\ 
-   50 & Recall & 0.00 & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.07 \\ 
-  100 & Recall &  & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.06 \\ 
-  200 & Recall &  & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.05 \\ 
+   10 & Recall & 0.92 & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.54 \\ 
+   50 & Recall & 0.00 & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.55 \\ 
+  100 & Recall &  & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.54 \\ 
+  200 & Recall &  & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.53 \\ 
     4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.89 & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.10 \\ 
-   50 & MCC &  & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & -0.04 \\ 
-  100 & MCC &  & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & -0.01 \\ 
-  200 & MCC &  & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.01 \\ 
+   10 & MCC & 0.89 & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.55 \\ 
+   50 & MCC &  & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & 0.47 \\ 
+  100 & MCC &  & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & 0.43 \\ 
+  200 & MCC &  & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
    \hline

### 2022-03-01 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex

@@
   4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.94 & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.69 \\ 
-   50 & Precision &  & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.19 \\ 
-  100 & Precision &  & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.15 \\ 
-  200 & Precision &  & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.10 \\ 
+   10 & Precision & 0.94 & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.91 \\ 
+   50 & Precision &  & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.61 \\ 
+  100 & Precision &  & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.57 \\ 
+  200 & Precision &  & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.38 \\ 
     4 & Recall & 1.00 & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Recall & 1.00 & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 0.77 \\ 
-   50 & Recall & 0.00 & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 0.33 \\ 
-  100 & Recall &  & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.27 \\ 
-  200 & Recall &  & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.28 \\ 
+   10 & Recall & 1.00 & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 1.00 \\ 
+   50 & Recall & 0.00 & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 1.00 \\ 
+  100 & Recall &  & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.99 \\ 
+  200 & Recall &  & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.98 \\ 
     4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.94 & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.52 \\ 
-   50 & MCC &  & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.15 \\ 
-  100 & MCC &  & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.15 \\ 
-  200 & MCC &  & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.13 \\ 
+   10 & MCC & 0.94 & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.91 \\ 
+   50 & MCC &  & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.74 \\ 
+  100 & MCC &  & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.72 \\ 
+  200 & MCC &  & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
    \hline

### 2022-03-01 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex

@@
   4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.91 & 0.59 & 0.97 & 0.97 & 0.90 & 0.56 & 0.74 \\ 
-   50 & Precision & 0.39 & 0.43 & 0.74 & 0.84 & 0.89 & 0.39 & 0.25 \\ 
-  100 & Precision &  & 0.34 & 0.53 & 0.72 & 0.91 & 0.29 & 0.25 \\ 
-  200 & Precision &  & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.25 \\ 
+   10 & Precision & 0.91 & 0.59 & 0.97 & 0.97 & 0.90 & 0.56 & 0.99 \\ 
+   50 & Precision & 0.39 & 0.43 & 0.74 & 0.84 & 0.89 & 0.39 & 0.99 \\ 
+  100 & Precision &  & 0.34 & 0.53 & 0.72 & 0.91 & 0.29 & 0.99 \\ 
+  200 & Precision &  & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.99 \\ 
     4 & Recall & 1.00 & 1.00 & 0.87 & 0.84 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Recall & 1.00 & 1.00 & 0.98 & 0.96 & 1.00 & 1.00 & 0.75 \\ 
-   50 & Recall & 1.00 & 1.00 & 0.85 & 0.99 & 1.00 & 1.00 & 0.25 \\ 
-  100 & Recall &  & 1.00 & 0.84 & 0.97 & 1.00 & 1.00 & 0.25 \\ 
-  200 & Recall &  & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 0.25 \\ 
+   10 & Recall & 1.00 & 1.00 & 0.98 & 0.96 & 1.00 & 1.00 & 1.00 \\ 
+   50 & Recall & 1.00 & 1.00 & 0.85 & 0.99 & 1.00 & 1.00 & 1.00 \\ 
+  100 & Recall &  & 1.00 & 0.84 & 0.97 & 1.00 & 1.00 & 1.00 \\ 
+  200 & Recall &  & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 1.00 \\ 
     4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.91 & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.58 \\ 
-   50 & MCC & 0.57 & 0.59 & 0.76 & 0.90 & 0.92 & 0.57 & 0.18 \\ 
-  100 & MCC &  & 0.54 & 0.61 & 0.81 & 0.94 & 0.48 & 0.22 \\ 
-  200 & MCC &  & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.23 \\ 
+   10 & MCC & 0.91 & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.99 \\ 
+   50 & MCC & 0.57 & 0.59 & 0.76 & 0.90 & 0.92 & 0.57 & 1.00 \\ 
+  100 & MCC &  & 0.54 & 0.61 & 0.81 & 0.94 & 0.48 & 1.00 \\ 
+  200 & MCC &  & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.99 \\ 
    \hline

### 2022-03-08 Manuscript/R&R/03-reviewer_4.Rmd
Add response to Reviewer 4 Item 6

@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+We agree that using "nonparametric methods" to differentiate from "black box" methods would be misleading.
+However, we respectively argue that our word choice, i.e.
+"nonparametric regression models", is not misleading.
+First of all The black box models normally are not treated as _regression_ models.
+Moreover, the term "nonparametric regression models" has been consistently used in the generalized additive models literature, for example @Hastie1987.
+Last but not least, to the authors' knowledge, there is a subtle distinction between nonparametric regression models and semiparametric regression models: semiparametric regression models assume some predictors have linear effects, while nonparametric regression models assume unknown functions for all predictors.
+To give an example, see @du2010.
+The subtle definition distinction would not matter as much in this manuscript, but would be relevant to the future work, incorporating predictor structure in the additive model, e.g.
+gene network.
+We hope to establish the consistency in the word choice to avoid future confusion.
+
+Nevertheless, we edited the sentence for clarity. [TODO: check if this is done]
 \bigskip

### 2022-03-08 Manuscript/01-Intro.Rmd
Update Section Headers to Latex
Restructure of the Introduction section, Removing comments

--- 2021-10-15 Manuscript/01-Intro.Rmd
@@
-# Introduction
+\section{Introduction}
 \label{sec:intro}
@@
 These models are useful for statistical prediction but do not estimate parameters relevant to the data generation process that one can draw inferences from.
-In addition, how to generalize these models to the high-dimensional setting remains unclear.
+In addition, how to generalize these \bg{"black box"} models to the high-dimensional setting remains unclear.
 
-<!-- Using additive models to interpolate nonlinear effects has a long history.
-Early work of additive models can date back to 1979 when Cleveland [@Cleveland1979] proposed local smoothing for bivariate analysis.
-Since then, many mathematical and statistical models have been proposed.
-[@Wood2020] With the recent advancement of computational technology, many machine learning algorithms that are computationally intensive start to gain much popularity in prognostic modeling and functional selection, especially when complex signals, for example nonlinear effects, are assumed.
-Such models generally have improved prediction accuracy over classical statistical models.
--->
-
-<!-- However, these models have a fatal flaw, the lack of transparency and interpretability, and hence are referred to as "black box" models [@Breiman2001].
-Given that the primary objective of biomedical studies is to understand , "Black box" models are less helpful as they can hardly explain how the change of biomarkers is associated with the disease.
--->
-
-Nonparametric regression models are appropriate alternatives to the "black-box" models thanks to their balance between model flexibility and interpretability.
+\bg{For their easy interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in the context of prediction and variable selection}.
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
@@
 $$
-where $g^{-1}(\cdot)$ is the inverse of a monotonic link function.
-The smoothing functions can take many forms and are estimated using a pseudo-weighted version of the backfitting algorithm [@Breiman1985].
-Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis in biomedical studies.
+where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.
 
@@
 The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smoothing functions.
-<!-- These models assume sparsity among the predictors, i.e.
-most predictors have small or no effects on the outcome.
-Further development focuses on addressing the grouping structure among the predictors, where AMs and GAMs can be considered special cases.
-Various penalties are imposed on the basis function coefficients:  -->
 Ravikumar et al.
@@
 [@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
-Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models.
-Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.
-Bai et al.
+Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models.\bg{[TODO: limit the scope, "particuaraly with spike-and-slab literatures."]} Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.\bg{[TODO: adjust citation position for clarity.]} Bai et al.
 [@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
@@
 
-To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via re-parameterization of the smoothing function to avoid overly shrinking basis function coefficients.
-Smoothing penalties are commonly implemented in the classical GAMs through smoothing regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive penalties on each smoothing function.
-A smoothing penalty conditioning on a smoothing parameter $\lambda_j$ is a function of the integration of the second derivative of the spline function, expressed mathematically as
-\begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
-\end{equation}
-where $\bs S_j$ is a known smoothing penalty matrix, and $\bs \beta_j$ are the basis function coefficients.
+\bg{To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
+[TODO: add a sentences on the contribution.
+Focuse on prediction of high-dimensional data, and possibly mentioning of vairbale selection.] Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature [TODO: add Woods citation], via re-parameterization [of the smoothing function] to avoid overly shrinking basis function coefficients.
 Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-Moreover, incorporating the smoothing penalty allows the separation of the linear space of a smoothing function from the nonlinear space.
 We then impose a new two-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
-The prior setup encourages a flexible solution, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-In addition, two scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm and EM-Iterative weighted least square (EM-IWLS) algorithm, are developed and implemented in an publicly available R pacakge \textt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.
-
-<!-- Marra and Wood [@Marra2011] introduced a re-parameterization of the basis function matrix based on the decomposition of the smoothness penalty matrix $\bs S$.
-The re-parameterization allows different penalties imposed on the null and nonlinear spaces separately, and further allows the sparsity in linear space and smoothness in nonlinear effect interpolation.
--->
- <!-- In addition, the priors of the linear and nonlinear effects are independent, creating computational convenience and further improving the scalability of the model. -->
+In addition, a scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm are developed.
+While the primary focus of this model is improvement in prediction, the proposed model also provides utility in fucntional selection.
+Particuaraly, the porposed prior setup encourages a bib-level selection/flexible solution, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects, as the re-parameterization allows the separation of the linear space of a smoothing function from the nonlinear space.
+The proposed model is implemented in an publicly available R package \texttt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.}
 
@@
 They make fitting high-dimensional models more feasible without sacrificing performance in prediction and variable selection.
-Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-CD algorithm to fit the model.
+Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-Coordinate Descent (EM-CD hereafter) algorithm to fit the model.
 While both methods use the combination of expectation maximization algorithm and coordient descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.
@@
 
-In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the two fast-fitting algorithms.
-In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation studies.
+In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the fast-fitting EM-CD algorithmalgorithm.
+In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM \bg{[TODO: add other models]} via Monte Carlo simulation studies.
 Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.
 Conclusion and discussions are given in Section \ref{sec:concl}.
-
-<!-- Similar to expanding classical models to high-dimensional models, expanding the GAM framework for high-dimensional applications is not a trivial task.
-The difficulties mainly come in two-fold: grouping nature of the basis functions, and the balance between spline smoothness and variable sparsity.
-For GAM models, a smoothing function is consists of multiple basis functions.
-The basis functions of a variable is inherently related.
-The grouping structure should be properly taken care of for accurate estimation.
-Moreover, shrinkage is previously used solely for sparsity of the model in high dimensional regression models, or smoothness in GAMs.
-How to organically combine the two together remains as a troubling question.
--->
-
-<!-- To the best of our knowledge, there are few GAM models proposed for high dimensional application.
-In common practice, model matrix of spline terms are generated using base functions of preference, and then feed to a penalized model.
-Without proper handle of the grouping nature of bases functions, the estimation of the smoothing function is less accurate.
-@Bai2020Spline shows inferior performance of such practice compared to using an ad hoc high-dimensional additive model.
-Among the specialized high-dimensional spline methods, @Ravikumar2009 extended the original backfitting algorithm [@Hastie1987] to accommodate the high-dimensional setting.
-The sparsity and smoothness was handled separately with a soft threshold.
-The authors also demonstrated its similarity to Group Lasso.
-@Meier2009 proposed an a sparsity-smoothness penalty, -->
-
-<!-- $$ -->
-<!-- J(f_j) = \lambda_1 \sqrt{||f_j||^2_n + \lambda_2 \int(f^{''}_j(x))^2dx}, -->
-<!-- $$ -->
-<!-- which can re-formulated as a group lasso problem as well.
-However, the model can not estimate the smoothness of the spline functions correctly, as the smoothness of all spline models are controlled by the lambda_2.
-The inflexible of the model creates problems in prediction.
--->
-
-<!-- Among all the choices of smoothing function, one particular family stood out, splines.
-Splines are piece-wise polynomial functions to interpolate nonlinear signals.
-Most commonly used spline is cubic spline, thanks to its smoothness property.
-[@Wood2017, PP.197] Early spline research focused on regression splines, where number and location of knots are pre-determined and ordinary least squares is used to estimate the smoothing function.
-However, the specification of knots is not a trivial task, as the the positioning of knots are influential on the performance of the model.
-As an improvement, researchers proposed smoothing splines where as many knots are used and later shrunk using penalized algorithms.
-The smoothing spline concept was first introduced [@OSullivan1986; @OSullivan1988], and later generalized and made famous by @Eilers1996, called P-spline.
-Independently, @Ruppert1999 proposed similar idea.
-Later, Bayesian GAMs were introduced as an alternative to address measurement error in covariates [@Berry2002] or local adaptive smoothing [@Lang2004].
--->

### 2022-03-10 Manuscript/R&R/01-reviewer_1.Rmd
Rough draft of the response

@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+We didn't perform the data processing step ourselves.
+We used the publicly available data that has been pre-processed.
+The data processing step is described in [TODO: add citation].
+Meanwhile, we respectively argue that the small R2 is reasonable, as in many applications, genomics data can only accounts small amount of data variation.
+We think this example would be the case, as environmental factors (not accounted in these models) would be more substantial when explaining weight loss.
 \bigskip
@@
 
-__Response__:
+__Response__: Thank you for the comment.
+The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms.
+In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalably model.
+We included an additional sentence in the discussion section.
+[TODO: list the sentence here].
+
+Compared to other priors, the spike-and-slab double exponential prior provides three advantages.
+First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients.
+Hence, the smoothing functions can be estimated more accurately.
+Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+
+We revised our method section to highlight these advantages, see Pxxx.
+
 \bigskip

### 2022-03-10 Manuscript/R&R/02-reviewer_2.Rmd
Rough draft of the response

@@
 
-_Comments to the Authorss_
+_Comments to the Authors_
 
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+The purpose of the reparameterization step is to factor the smoothing penalty matrix in the design matrix, such that the smoothing penalty is no long a function of the smoothing penalty matrix.
+In other words, the smoothing penalty changes from $\lambda \bs \beta^T \bs S \bs \beta$ to $\lambda \bs {\beta^*}^T \bs {\beta^*}$, where the $\bs \beta$ and $\bs \beta^*$ are the coefficients of the smoothing function before and after the reparameterization respectively.
+The spike-and-slab LASSO prior estimates whether to include the smoothing function and its appropriate smoothing parameter $\lambda$ with local adaptivity.
+This contrasts to previous methods that don't consider smoothing penalty at all or impose a uniform sparsity penalty upon uniform smoothing penalty, where both approaches tend to induce excessive shrinkage for the smoothing function.
+
+In the updated manuscript, we clarify the utility of the reparmeterization step and emphasize how the smoothness of the additive function are properiately modeled.
+Please see the revision of the method section, highlighted in red.
+
 \bigskip
@@
 
-__Response__:
+__Response__: Compared to other priors, the spike-and-slab double exponential prior provides three advantages.
+First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients.
+Hence, the smoothing functions can be estimated more accurately.
+Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+
+We revised our method section to highlight these advantages, see Pxxx.
 \bigskip
@@
 
-__Response__: Thank you for the comment.
+__Response__: We thank the reviewer for the comment.
 We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$ (See Equation ([TODO: add the equation number])), in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.

### 2022-03-10 Manuscript/R&R/03-reviewer_4.Rmd
Rough draft of the response

@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$ (See Equation ([TODO: add the equation number])), in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
+This prior formulation is motivated by effect forcing discussed in (Chipman, 2004).
+Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
+The modification of the prior doesn't complicate the model fitting algorithm.
+The $\gamma$ can be analytically integrated out of the density function (See Equation ([TODO: add the equation number]) and appendix) and requires minimum change of the proposed algorithms.
+In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the review for the comments.
+We added a new subsection discussing the function selection result from the simulation studies.
+The suggested metrics, including aaa, bbb, ccc are described for all models of comparisons.
+Lastly,the bi-level selection results are discussed too.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the review for the comment.
+In the update manuscript, we add the suggested mdoel, spikeslabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared.
+We see that the proposed method still preserves the computation advantage over spikeSlabGAM, particularly in the high-dimensional cases.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+After careful deliberation, we decide to focusing an explaining the EM-CD method well.
+The inclusion of EM-IWLS would distract the reader from the main focus of manuscript, scalable model, good prediction and bi-level selection.
+We fully agree that the questions raised by the reviewer are very important and deserves careful discussion.
+Nevertheless, that would be out of the scope of the current work.
+We will dedicate another manuscript to detail many of the questions here, for example estimating effective degree of freedom, limiting distribution of the test statistics to the EM-IWLS method.
+
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment. We added more graphs displaying the effects of the significant effects.
 \bigskip

### 2022-03-11 Manuscript/01-Intro.Rmd
Revised the intro and method section

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution with mean $\mu$ and dispersion $\phi$, can be modeled with the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
-$$
- E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
-$$
-where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.
+In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution, can be modeled with the summation of flexible smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$.
+Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.
 

### 2022-03-11 Manuscript/02-Method.Rmd
Update Section Headers to Latex
Revised the intro and method section

@@
-# Bayesian Hierarchical Additive Models (BHAMs)
+\section{Bayesian Hierarchical Additive Models (BHAM)}
 \label{sec:BHAM}
 
-Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model with link function $g(\cdot)$ and linear predictor
-\begin{equation}\label{eq:gam}
-\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j,
+\bg{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
+The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+$$
+ E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
+$$
+where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \bs x_i\}^n_{i=1}$,} the data distribution is expressed as
+\begin{equation}
+f(\bs Y = \bs y| \bs \beta, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\bs \beta, \phi).\nonumber
 \end{equation}
-with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The outcome $Y$ follows a exponential family distribution with density function $f(y)$, mean $\mu = g^{-1}(\eta)$ and dispersion parameter $\phi$, and the data distribution is
-\begin{equation}
-f(\bs Y = \bs y| \bs \beta, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\bs \beta, \phi),\nonumber
-\end{equation}
+
 The basis function matrix, i.e.
@@
 The dimension of the design matrix depends on the choice of the smoothing function, and is denoted as $K_j$ for $x_j$.
-Its corresponding smoothing penalty matrix is denoted as $\bs S_j$.
 $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
@@
 The matrix transposing operation is denoted with a superscript $^T$.
-To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g.
-$B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+\bg{To note, the proposed model can include parametric forms of variables in the model, and hence considers general linear models and semiparametric regression models as special cases.}
 
-To encourage proper smoothing of additive functions, we adopt the idea of smoothing penalties from smoothing spline models.
-The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function.
-However, the direct integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and  Wood [@Marra2011] proposed a re-parameterization procedure to accommodate the smoothing penalty implicitly.
-Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+\subsection{Smoothing Function Reparameterization}
+
+To encourage proper smoothing of each additive function, we adopt the smoothing penalty from smoothing spline models [TODOO: add Wood citation].
+A smoothing penalty, a function of the integration of the second derivative of the spline function.
+is quadratic norms of the coefficients \bg{of what?} and allow locally adaptive penalties on \bg{each smoothing function[TODO: not accurate, bases]}, mathematically
+\begin{equation}\label{eq:smoothpen}
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
+\end{equation}
+where $\bs S_j$ is a known smoothing penalty matrix and $\lambda_j$ denotes a smoothing parameter.
+A linear function can be modeled as $B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+\bg{Unlike previous methods that embeds smoothing penalty in the sparse penalty which leads to a more restrictive solution, we considers an additional mechanism via the propsed prior (discussed in Section xxx) to address sparsity in signal.
+such that the locally adpative nature of the smoothing penalty retains.}
+
+Marra and  Wood [@Marra2011] proposed a re-parameterization procedure \bg{to factor the smoothing penalty into the design matrix[TODO: vague]}.
+Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
 To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
@@
 
-The re-parameterization step sets up the foundation of the proposed model and provides three benefits.
+To summarize, the re-parameterization step provides three benefits.
 First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for functional selection.
@@
 
-## Spike-and-Slab Spline Priors
+\subsection{Two-part Spike-and-Slab Lasso Prior for Smoothing Fucntions}
 
 The family of spike-and-slab (SS) regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
-We defer to Bai et al.
-[@Bai2021Review] for an in-depth introduction to spike-and-slab priors.
-To summarize, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{\text{spike}}(\cdot)$ for weak signals and a flat slab density $f_{\text{slab}}(\cdot)$ for strong signals, mathematically
-$$
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-$$
-The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma \in \{0,1\}$ that indicates whether the variable $x$ is included in the model.
-There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al.
-[@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.
-Two of most popular spike and slab priors are spike-and-slab normal prior [@George1993] and spike-and-slab double exponential prior[@Rockova2018].
-
-<!-- Comments -->
-<!-- Even the conjugate normal prior of $\bs \beta$ provides analytic simplicity, there is no closed form solution for the posterior density of $\bs \gamma$.
-@George1993 described a Gibbs sampling algorithm to approximate the posterior distribution of $\bs \gamma$ for variable selection.
-The same authors later compared SSVS with other Bayesian variable selection approaches, providing some empirical suggestions on the advantages and disadvantages of various formulations and estimation algorithms.
-[@George1997] Among the various problem SSVS addresses, high-dimensionality is not one of them, due to the intensive computation needs of MCMC algorithms.
-@George1997 suggested MCMC algorithms works well for medium size of predictors (p=25), which is not feasible for high-dimensional data, where the number of predictors easily exceeds 100.
--->
-
-The major criticism of early spike-and-slab models is being computationally prohibitive.
-[@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS based on spike-and-slab normal prior[@Rockova2014a] and spike-and-slab Lasso (SSL) [@Rockova2018b; @Rockova2018].
-Particularly, the development of SSL model  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+\bg{Among all the spike-and-slab priors, the spike-and-slab Lasso (SSL) prior [@Rockova2018b; @Rockova2018] is one of the most popular choices.
 The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
@@
 \end{equation}
+}
+The latent binary variable $\gamma \in \{0,1\}$ indicates whether a variable $x$ is included in the model, while the dispersion parameters $s_0$ and $s_1$ controls the shrinkage of the coefficient.
 Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as one single double exponential density,
 \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
+\beta | \gamma \sim DE(0, (1-\gamma)s_0 + \gamma s_1), 0 < s_0 < s_1.
 \end{equation}
-with the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$.
-The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is preferred in high-dimensional data analysis.
-We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015] and limits to an "all-in-all-out" strategy for functional selection.
+Compared to other priors for high-dimensional data analysis, SSL has the following advantages.First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients.
+Hence, the smoothing functions can be estimated more accurately.
+Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+We defer to Bai et al.
+[@Bai2021Review] for an detailed discussion.
+Particularly, the development of SSL model substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+We notice that Bai[TODO: add Bai arxiv paper] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [TODO: add Xu Ghosh (2015)] and limits to an “all-in-all-out” strategy for functional selection.
 
-### Two-part Spike-and-Slab Lasso Prior
-
-We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear modeling with bi-level selection.
-The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}).
+We introduce a novel SSL-based prior for smoothing functions in GAMs.
 Given the re-parameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
-Specifically, we impose independent group priors on the linear space coefficients and on the nonlinear space coefficients respectively,
+Specifically, the linear space coefficient has a SSL prior and the nonlinear space coefficients shares a group SSL prior,
 \begin{align}\label{eq:bham_ssl}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
+  \beta_{j} | \gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
   \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j
 \end{align}
-where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
-$s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
-These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation.
-A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}.
-To note, this prior differs from previous group SSL priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
-It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
-$\gamma_j = \gamma^\tp_j$.
-This converges to the SB-GAM [@Bai2021].
-Conversely, it is also possible to relax the assumption such that each coefficient $\beta_{jk} \in \bs \beta^\tp_j$ has its own latent indicator $\gamma_{jk}$, but at the cost of complicating the bi-level functional selection.
-This reduces the proposed prior to the classic SSL prior.
+where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.$s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
+These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation, discussed in Section \ref{sec:tune}.
 
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
-First of all, the smoothing function bases are linear dependent and necessitate extra attention.
-The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
-Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions.
-
-The rest of the hierarchical prior follows the traditional SSL prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
-The two indicators of the $j$th predictor, $\gamma_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
+\bg{The group SSL prior of the nonlinear space coefficients differs from previous group SSL priors [@Tang2018; @Tang2019], as effect hierarchy are preferred when selecting linear effect and non-linear effect.
+Effect hierarchy refers to the principle that lower-order effects are more likely to be active than higher-order effects.
+[TODO: add chipman paper citation.] To implement, we consider the hyper prior of the shared latent indicator $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$.
+Both latent indicators $\gamma_j and \gamma^\tp_j$ follows a Bernoulli distribution, written as binomial distribution of one trial.
+While the probability of including the linear effect is $\theta_j$, a probability parameter, the probability of including the nonlinear effect is $\bg{\gamma_{j}}\theta_j$.}
 $$
@@
 $$
-\bg{Equivalently, we can derive the marginal distribution of $\gamma_{j}^\tp$ by integrate out $\gamma_j$ (Supplement 1), 
-$$
-\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2).
-$$
+\bg{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
+For the computational convenience in the proposed EM-CD algorithm (see Section [TODO: add section ref]), we analytically integrate $\theta_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$.
 }
 
-This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
+<!-- To note, this prior differs from previous group SSL priors [@Tang2018; @Tang2019], as the $\beta_j$ and $\bs \beta^\tp_j$ have different indicator variables $\gamma_j$, $\gamma_j^\tp$ respectively.
+It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e.
+$\gamma_j = \gamma^\tp_j$.
+This converges to the SB-GAM [@Bai2021].
+Conversely, it is also possible to relax the assumption such that each coefficient $\beta_{jk} \in \bs \beta^\tp_j$ has its own latent indicator $\gamma_{jk}$, but at the cost of complicating the bi-level functional selection.
+This reduces the proposed prior to the classic SSL prior.
+-->
+
+<!-- The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
+First of all, the smoothing function bases are linear dependent and necessitate extra attention.
+The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
+Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions.
+-->
+
+<!-- The rest of the hierarchical prior follows the traditional SSL prior: we set up hyper-priors on $\gamma_j, \gamma^\tp_j$ to allow local adaption of the shrinkage using a Bernoulli distribution, written as binomial distribution of one trial.
+The two indicators of the $j$th predictor, $\gamma_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$, -->
+
+<!-- This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
 The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the functional selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that $\gamma_{j}$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity.
-We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
+-->
+
+\bg{To allow the shrinkage to self-adapt to the sparsity and smoothing pattern of the data, }we further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
 $$
@@
 $a = 1, b = 1$.
-When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases.
-Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab spline prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
+\bg{This enables the proposed prior to inherit the selective shrinkage property and self-adaptivity [TODO: add Bai review] from the classical SSL prior.
+In other words, when a smoothing function is significant, the coefficients of the smoothing function will escape the overall shrinkage and produce a more accurate estimate, particular with the smoothing penalty addressed implicitly via the reparameterization.
+Meanwhile, automatic adjustment for different level of sparsity are feasible via the information borrowing cross coordinates.}
+<!-- This informaiton borrowing helps the model to find adequate shrinkage for the linear coefficient and coefficients of the nonlinear part.
+Not that, the reparameterazation step factor the smoothing penalty matrix into the design matrix, and we don't have to further explicitly arrange a multivariate piror to consider appropriate shrinkage of those coefficient.} -->
+<!-- When the variable has large effect in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases.  -->
+Hereafter, we refer Bayesian hierarchical generalized additive models with the two-part spike-and-slab LASSO prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
 \begin{center}
@@
 
-### Other Priors
+\subsection{Scalable EM-Coordinate Descent Algorithm}
 
-With the re-parameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
-These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
-Here we elaborate the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in BHAM.
+\bg{Despite the advantage to estimate posterior densities and derive uncertainty inference, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
+Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) [@George1997], and substaintially slows as $p$ increases modestly in hte GAM context [@Scheipl2013].
+Hence, we considers optimizaiton algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
+Optimizaiton algorithms greatly speed up the model fitting process, and improve the scalability of BHAM.}
 
-A spike-and-slab mixture normal spline prior can be expressed as 
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j.
-\end{align*}
-Similar to the spike-and-slab spline prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
-One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
-Distributions with heavier tails can be used as an alternative, for example mixture Student's $t$ distribution with small degree of freedom.
+In this section, we extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
+\bg{Specifically for the SSL priors, Rochova and co-authors [@Rockova2018b; @Rockova2018] further integrate the EM algorithm and coordinate descent algorithm for fast computation.
+Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
+Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research.} To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
+The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a block CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
 
-## Algorithms for Fitting BHAMs
-
-The proposed models can be fitted with MCMC algorithms.
+<!-- The proposed models can be fitted with MCMC algorithms.
 Nevertheless, the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch[@George1997] examined the computation speed for various MCMC algorithms with spike-and-slab mixture normal priors, and suggested MCMC algorithms works well for medium size ($p$=25) of predictors with only linear effects.
+George and McCulloch examined the computation speed for various MCMC algorithms with spike-and-slab mixture normal priors, and suggested MCMC algorithms works well for medium size ($p$=25) of predictors with only linear effects.
 However, it is not feasible for high-dimensional data analysis where the number of predictors easily exceeds 100.
@@
 Scheipl et al.
-[@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab GAM grow exponentially as $p$ increases modestly via simulation studies.
+[@Scheipl2013] demonstrated the computational demands of a MCMC algorithm for fitting spike-and-slab GAM grow exponentially as $p$  via simulation studies.
 Hence, we feel compelled to develop scalable algorithms for fitting Bayesian hierarchical additive models in high-dimensional settings.
+-->
 
-As an alternative to sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of posterior inference.
+<!-- As an alternative to sampling algorithms for fitting Bayesian models, optimizatio algorithms focus on the maximum a posteriori (MAP) estimates and speed up the model fitting process at the cost of posterior inference.
 The earlier work for fitting SS models using optimization algorithms includes EMVS [@Rockova2014a].
 Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors.
+-->
 <!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$.
@@
 -->
-In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, followed by updating $\phi, \theta$.
-<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter.
-EMVS is also compatible with structured prior when dealing with grouped variables.
-The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.
--->
-The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent algorithm to fit SSL models.
-<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL).
-The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective.
-The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero.
--->
-Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research
-
-<!-- Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion. -->
-
-In this section, we extend the two EM-based algorithms, EM-CD and EM-IWLS algorithms, to fit BHAMs.
-To note, the two proposed algorithms provides different utilities.
-The EM-CD algorithm is specifically for fitting BHAM with an expedited performance, recommending to use in high and ultra-high dimensional setting.
-A specific concern of EM-CD algorithm is that it provides no information for inference.
-In contrast, the EM-IWLS can estimate the variance-covariance matrix of the coefficients.
-Moreover, the EM-IWLS is a more general model fitting algorithm that can be used for fitting not only SSL and continuous SS priors but also Student's t-priors and double exponential priors.
-To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
-The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a block CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
-
-<!-- , stemming from previous EM-coordinate descent algorithms[@Tang2018; @Tang2019] and EM-IWLS [@Rockova2014a].
-Instead of simulating the posterior distribution via MCMC, we iteratively estimate the MAP.
-Compared to the EM-IRLS algorithm, the EM-coordinate descent provides an expedited model fitting process at the cost of model uncertainty.
+<!-- In the E step, the latent binary indicators $\gamma$s are treated as the missing data, and the posterior means are calculated conditioning on the current value of other parameters; in the M step, a ridge estimator was used to update the coefficients, followed by updating $\phi, \theta$.
 -->
 
-<!-- The main difference between Bai's work and the proposed models are two-folded: 1) the proposed models are feasible to answer the question if the nonlinear effects are necessary for the predictors, while Bai's work can't due to the "all-in-all-out" approach; 2) the proposed models can provide uncertainly measures in addition to the point estimates of the coefficients, while Bai's work can't not due to the model fitting algorithm.
-One of the exception is Bai (in progress), who a grouped spike-and-slab LASSO prior [@Rockova2018] to address the grouping structure of spline bases.
+<!-- EM algorithm is an iterative algorithm to find MAP estimates or the maximum likelihood estimates.
+It is commonly used when some necessary data to establish the likelihood function are missing.
+Instead of maximizing the the likelihood function, the algorithm maximizes the expectation of the likelihood function with respect to the "missing" data.
 -->
 
-### EM algorithms
+<!-- The recursive algorithm consists of two steps:  -->
 
-EM algorithm is an iterative algorithm to find MAP estimates or the maximum likelihood estimates.
-It is commonly used when some necessary data to establish the likelihood function are missing.
-Instead of maximizing the the likelihood function, the algorithm maximizes the expectation of the likelihood function with respect to the "missing" data.
+<!-- * E-step: to calculate the expectation of the posterior density function with respect to some "missing" data -->
+<!-- * M-step: to maximize the expectation derived in the E-step and update parameters of interest -->
 
-The recursive algorithm consists of two steps:
-
-* E-step: to calculate the expectation of the posterior density function with respect to some "missing" data
-* M-step: to maximize the expectation derived in the E-step and update parameters of interest
-
-For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data.
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data\bg{} for the EM algorithm}.
 Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
@@
 $$  
-where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(1,1) density.
+where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(\bg{a},\bg{b}) density.
 We choose non-informative prior for the intercept $\beta_0$ and the dispersion parameter $\phi$; for example, $f(\beta_0|\tau_0^2)=N(0,\tau_0^2)$ with $\tau^2_0$ set to a large value and $f(\log \phi) \propto 1$.
 
-<!-- , and rewrite $\beta_j$ as $\beta_{j0}$, and hence removing the superscript $^{*}$ from the coefficients of the the penalized space.
-Using the new notation, the part component $\log p(\beta_j|\gamma_j)+\sum\limits_{k=1}^{K_j} \log p(\beta^{*}_{jk}|\gamma^{*}_{jk})$ can simplifies to $\sum\limits_{k=0}^{K_j} \log p(\beta_{jk}|\gamma_{jk})$.
-Without loos of generality, the model still work for the spline function whose dimensionality of the null space is more than 1.
--->
-
-We use the EM algorithm to find the MAP estimate of $\Theta$.
+We use the EM algorithm to find the the maximum a posteriori estimate of $\Theta$.
 This is, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
@@
 In the M-step, we find the $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
-The E- and M- steps are iterated until the algorithm converge.
+\bg{The parenthesized subscription denotes the parameter estimation at the $t$th iteration.} The E- and M- steps are iterated until the algorithm converge.
 
@@
 Meanwhile, conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
-Depending on the choice of coefficient priors, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via CD algorithm or IWLS algorithm in each iteration.
+With the proposed two-part spike-and-slab lasso prior, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via the Coordinate Descent algorithm in each iteration.
+Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
 Maximization of $E(Q_2)$ can be solved via closed form equations following the beta-binomial conjugate relationship.
 
-### EM-Coordinate Descent
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double exponential, coordinate descent algorithm can be used to estimate the parameters in the M-step.
-Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
-
-The density function of spike-and-slab mixture double exponential prior can be written as
+The density function of the mixture double exponential prior of coefficient $\beta$ can be written as
 $$
@@
 \end{equation}
-where $S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1$ and $S^\tp_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
-To calculate two unknown quantities $E({S_j}^{-1})$ and $E({S^\tp}^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma^{0}_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
+where $S_{j} = (1-\gamma_{j}) s_0 + \gamma_{j} s_1$ and $S^\tp_{j} = (1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1$.
+To calculate two unknown quantities $E({S_j}^{-1})$ and $E({S^\tp}^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
 The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the value of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
@@
 \end{align*}
-and $\pr(\gamma_{j}^{0} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+and $\pr(\gamma_{j} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
@@
 
-<!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
-The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S_j}^{-1}){\beta_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S}_1^{-1}), E(S_1^{-1}), \dots, E({S_p}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S_1}^{-1}), \dots E({S_p}^{-1})$.
-An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.
-The estimates of $\bs \beta$ and $\phi$ can be quickly updated in the IRLS algorithm.
--->
-
-<!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients.
-Moreover, to note that, we have different shrinkage $S^{-1}_{jk}$for the coefficients $\beta_{jk}$ of the variable $x_j$, and hence, we can penalize the null space of the spline differently and allow local adaption.
--->
-
-<!-- Totally, the framework of the proposed EM IRLS algorithm was summarized as follows: -->
-
-<!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ [for $\bs \beta$ and $\bs \theta$.
-For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
-
-<!-- 2) Iterate over the E-step and M-step until convergence -->
-
-<!-- E-step: calculate $E(\gamma_{j})$, $E(\gamma^{pen}_{j})$and $E({S}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
-
-<!-- M-step: -->
-
-<!-- a) Update $\bs \beta^{(t)}$ using the IRLS algorithm -->
-
-<!-- b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta} -->
-
-<!-- We assess convergence by the criterion: -->
-<!-- $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where -->
-<!-- $d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at -->
-<!-- the $t$th iteration, and $\epsilon$ is a small value (say -->
-<!-- $10^{-5}$). -->
-
-## Selecting Optimal Scale Values
+\subsection{Selecting Optimal Scale Values}
 \label{sec:tune}
@@
 Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding a criteria of preference.
-This would be achieved by constructing a two dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
+This would be achieved by constructing a two-dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
 However, previous research suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value, e.g.
@@
 [@Rockova2018] As a result, we focus on examining different values of spike scale $s_0$.
-Instead of the 2-D grid, We consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
+Instead of the two-dimensional grid, we consider a sequence of $L$ decreasing values $\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1$.
 Increasing the spike scale $s_0$ tends to include more non-zero coefficients in the model.

### 2022-03-14 Manuscript/03-Simulation.Rmd
Revise simulation section

--- 2022-02-25 Manuscript/03-Simulation.Rmd
@@
 
-In this section, we compare the performance of the proposed models to four alternative models: component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing [@Wood2011], SB-GAM [@Bai2021]\bg{, linear lasso model and spikeSlabGAM [TODO: add citation here].[TODO: explain about linear lasso model and spikeSlabGAM model]}.
+In this section, we compare the performance of the proposed model to \bg{six alternative models: linear Lasso models, component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM [TODO: add citation here]}}, and SB-GAM [@Bai2021].
+\bg{We use linear lasso model as a benchmark, examing the performance when linearity assumption doesn't hold}.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
 Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component are different for extra flexibility.
-Generalized additive models with automatic smoothing, hereafter \textit{mgcv}, is one of the most popular models for nonlinear effect interpolation and prediction.
-\bg{mgcv doesn't support analyses where the number of parameters exceeds the sample size.} SB-GAM is the first spike-and-slab lasso GAM.
-We implement COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, SB-GAM with R package \texttt{sparseGAM 1.0}.
-COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices.
-Both mgcv and proposed models allow customized smoothing functions and we choose the cubic regression spline.
-We controll the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
-We use 5-fold CV with the default selection criteria to select the final model for COSSO models, SB-GAM and the proposed models.
-20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed models which allow user-specification of tuning candidates.
+\textit{mgcv} is one of the most popular models for nonlinear effect interpolation and prediction.
+Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab lasso GAM.
+We implement linear Lasso model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
+COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices; mgcv, spikeSlabGAM and the proposed model allow customized smoothing functions and we choose the cubic regression spline.
+We control the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
+We use 5-fold CV with the default selection criteria to select the final model for linear Lasso model, COSSO models, SB-GAM and the proposed model.
+20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed model which allow user-specification of tuning candidates.
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
@@
 Other related methods for high-dimensional GAMs also exist, notably the methods of sparse additive models by Ravikumar et al.
-[@Ravikumar2009] and stochastic search term selection for GAM [@Scheipl2012].
-However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to mgcv and scalability issues with increased number of predictors.
-[@Scheipl2013]
+[@Ravikumar2009].
+However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].
 
@@
 We first generate $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
-We then simulate the outcome $y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
+We then simulate the outcome $Y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
 The mean of each outcome were simulated via the following function
@@
 Another set of independent sample of size $n_{test}=1000$, are created following the same data generating process, serving as the testing data.
-We generate 50 independent pairs of training and testing datasets to evaluate the prediction performance of the chosen models, where training datasets are used to fit the models and testing datasets used to calculate assessment measures.
-\bg{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
-These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
+We generate 50 independent pairs of training and testing datasets to evaluate the prediction and variable selection performance of the chosen models, where training datasets are used to fit the models and testing datasets are used to calculate metrics of interest.
+\bg{In addition, we consider the data generating process where all the functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
+These additional sets of linear simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
 
-To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing dataset, are averaged <!--and deviance are calculated for Poisson model,--> across 50 simulations.
-\bg{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
-For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-$$
-\bg{MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},}
-$$
-\bg{where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.}
+To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing datasets, are averaged across 50 simulations.
+\bg{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC)
+\begin{align*}
+&\text{precision} = \frac{TP}{PP}\\
+&\text{recall} = \frac{TP}{TP+FP}\\
+&\text{MCC} = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
+\end{align*}
+where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively.} For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} which relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.[TODO: verify]}
+
 \subsection{\bg{Simulation Results}}
 \subsubsection{\bg{Prediction Performance}}
-\bg{Among the main set of simulations where the functional forms of the predictors are mostly non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
+\bg{Among the set of simulations where the functional forms of the predictors are mostly non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
 (See Tables \ref{tab:gaus} and \ref{tab:bin_auc}) \bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional spline models.
 Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
-When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse [TODO: re-evaluate after tuning the models] in high-dimensional cases ($p=50,100, 200$).
+When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse in high-dimensional cases ($p=50,100, 200$).
 As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv still fit models with increasing in large dimension case ($p=50$).
-But the algorithm fails to converge within the default number of iterations, and hence leads to bad prediction performance.} In high dimensional cases where we mimic the situation the signals are extremely sparse, SB-GAM has better performance than the proposed method.
-However, the BHAM-CD has extreme computational advantage over SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the prediction accuracy.
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to setimate ($p=50$).
+This leads to bad performance.} In high dimensional cases where we mimic the situation the signals are extremely sparse, SB-GAM has slight better performance than the proposed method[TODO: quantify].
+However, the BHAM has extreme computational advantage over SB-GAM [TODO: quantify] without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}) .
 
@@
 
-\bg{\subsubsection{Variable Selection Performance}
-Within the simulation settings we examined, the proposed model tends to have a more conservative bebaviour in terms of variable selection, having high precision and low recall.
-(See Table xxx for gaussian outcomes, and support information xxx for binomial informaitons.) [TODO: mention the variable selection pattern is consistent across the types of outcomes.] THe high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
-Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high matthew correlation coefficient.
-Another Bayesian model, spike-and-slab performs bad in terms of variable selection in high-dimensional situations ($p=100,200$).
-We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
-}
+\bg{\subsubsection{Variable Selection Performance}}
+\bg{Among the set of simulations where the functional forms of the predictors are mostly non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table xxx for Gaussian outcomes, and support information xxx for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
+We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$ due to using a generic sequence of $s_0$ that doesn't contain the optimal value, discovered when analyzing the simulation result.
+Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high MCC.
+The performance of another Bayesian model, spikeSlabGAM deteriorates as sparsity grows, particularly when ($p>50$), or for binomial outcomes.
+The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.
 

### 2022-03-15 Manuscript/03-Simulation.Rmd
Add bilevel selection result (#37)

@@
 The performance of another Bayesian model, spikeSlabGAM deteriorates as sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.
+The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.}
 
-\bg{\subsubsection{Bi-level Selection}
-Among the methods of comparison, there are two methods is possible for bi-level selection, i.e.
-the proposed BHAM and spikeSlabGAM.
-Hence, in this section, we further investigate the bi-level selection of the proposed models.
-}
+\bg{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
+Among the linear simulations, both methods can accurately select the linear components while having a drastically lowered probability to include the nonlinear component in the model.
+Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM.
+However, this advantage of spikeSlabGAM over BHAM is less obvious among the mixed function form simulations: spikeSlabGAM performs better than BHAM when selecting components when the function form includes only linear or nonlinear component, e.g.
+function form for $x_3$ and $x_4$.
+However, the spikeSlabGAM inclines to complete ignore the variable that have more complex function form, e.g.
+function form for $x_1$ and $x_2$, while BHAM are more likely to include them in the model.
+This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
+We defer an in-depth discussion to the Section \ref{sec:concl}.}
 

### 2022-03-15 Manuscript/10-Supplementary.Rmd
Revise Variable Selection Results + Add Tables (#22)
Move Other Prior section to supplementary.

@@
 
+<!--     where $\eta_i = \sum\limits_{j=1}^p \hat f_j^{(t)}(x_{ij})$ is the estimated linear predictor using current estimated value of $\bs \beta^{}$, $L^{'}$ and $L^{''}$ are the first derivative and second derivative of the log-likelihood function.
+The $l_2$ penalty, $\sum\limits_{j=1}^{p}\left[E({S_j}^{-1}){\beta_j}^2+\sum\limits_{k=1}^{K_j}E(S^{-1}_{j})\beta_{jk}^2\right]$ can be seems as a multivariate normal distribution of mean $\textbf{0}$ and variance-covariance matrix $\Sigma(E({S}_1^{-1}), E(S_1^{-1}), \dots, E({S_p}^{-1}), E(S_p^{-1})$, conditioning on the value of $E({S_1}^{-1}), \dots E({S_p}^{-1})$.
+An augmentation step combines the normal approximation and multivariate normal for $\bs \beta$, similarly to Ridge regression model fitting.
+The estimates of $\bs \beta$ and $\phi$ can be quickly updated in the IRLS algorithm.
+-->
+
+<!-- From Equation^[TODO: add equation number], we can see that the estimates of $\gamma_jk, S^{-1}_{jk}$ are larger for larger coefficients $\beta_{jk}$, leading to different shrinkage for different coefficients.
+Moreover, to note that, we have different shrinkage $S^{-1}_{jk}$for the coefficients $\beta_{jk}$ of the variable $x_j$, and hence, we can penalize the null space of the spline differently and allow local adaption.
+-->
+
+<!-- Totally, the framework of the proposed EM IRLS algorithm was summarized as follows: -->
+
+<!-- 1) Choose a starting value $\bs \beta^{(0)}$ and $\bs \theta^{(0)}$ [for $\bs \beta$ and $\bs \theta$.
+For example, we can initialize $\bs \beta^{(0)} = \bs 0$ and $\bs \theta^{(0)} = \bs 0.5$ -->
+
+<!-- 2) Iterate over the E-step and M-step until convergence -->
+
+<!-- E-step: calculate $E(\gamma_{j})$, $E(\gamma^{pen}_{j})$and $E({S}^{-1}_{j})$  , $E({S}^{-1}_{j})$with estimates of $\Theta^{(t-1)}$ from previous iteration -->
+
+<!-- M-step: -->
+
+<!-- a) Update $\bs \beta^{(t)}$ using the IRLS algorithm -->
+
+<!-- b) Update $\bs \theta^{(t)}$ using Equation \ref{eq:update_theta} -->
+
+<!-- We assess convergence by the criterion: -->
+<!-- $|d^{(t)}-d^{(t-1)}|/(0.1+|d^{(t)}|)<\epsilon$, where -->
+<!-- $d^{(t)} = -2l(\beta^{(t)},\phi^{(t)})$ is the estimate of deviance at -->
+<!-- the $t$th iteration, and $\epsilon$ is a small value (say -->
+<!-- $10^{-5}$). -->
+
+\subsubsection{Other Priors}
+
+With the re-parameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
+These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
+Here we elaborate the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in BHAM.
+
+A spike-and-slab mixture normal spline prior can be expressed as 
+\begin{align*}
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
+  \beta^\tp_{jk} | \gamma^\tp_{j},s_0,s_1 &\simiid N(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1), k=1,\dots, K_j.
+\end{align*}
+Similar to the spike-and-slab spline prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
+One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
+Distributions with heavier tails can be used as an alternative, for example mixture Student's $t$ distribution with small degree of freedom.
+
 \clearpage
@@
 \input{Tabs/sim_lnr_binom_tab.tex}
+
+\clearpage
+
+\subsection{Supplementary Information 4: Variable Selection Performance of Simulations}
+\input{Tabs/sim_binom_var_slct_tab.tex}
+\input{Tabs/sim_lnr_gaus_var_slct_tab.tex}
+\input{Tabs/sim_lnr_binom_var_slct_tab.tex}

### 2022-03-15 Manuscript/02-Method.Rmd
Clarify deviance

@@
 A measure of preference calculated with cross-validations (CV), e.g.
-deviance, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
+deviance \bg{(defined as model log-likelihood times -2, $-2\log f(\bs y|\bs{\hat\beta}, \hat\phi)$)}, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
 The procedure is similar to the Lasso implemented in the widely used R package `glmnet`, which quickly fits Lasso models over a list of values of regularization parameters $\lambda$ , giving a sequence of models for users to choose from.

### 2022-03-15 Manuscript/04-Real_Data.Rmd
Revise Real Data analysis

@@
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}).
+[TODO: add a sentenec how much faster.]
 \begin{center}
@@
 There are initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
-Hence, we perform a univariate screening procedure on the features, via GAM implemented in \texttt{mgcv}, and choose the the top 200 features with smallest p-values.
+Hence, we choose the the top 200 features \bg{with largest variance}.
 We use 5-knot spline additive models for binary outcome using two different models, the proposed BHAM and the SB-GAM.
@@
 We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
-In this analysis, we primarily focuse on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects.
+In this analysis, we primarily focus on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects.
 In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
-Quality control and natural log transformation are performed during metabolites data preparation.
+Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team [@Bihlmeyer2021]}.
 The outcome of interest are standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
@@
 Out-of-bag samples are used for prediction performance evaluation, where deviance, $R^2$, mean squared error (MSE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and mean absolute error (MAE) defined as $\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|$ are calculated.
-BHAM-CD obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:WLM_res}).
+BHAM obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:WLM_res}).
 \begin{center}

### 2022-03-17 _targets.R
Add Real data functional plots

@@ -308,10 +308,25 @@ tar_plan(
   
   # *** Plot Non-Linear Functions ###
   # TODO: to generalize this for a list of plots.
-  ECB_bamlasso_nonlnr = ECB_bamlasso_var$`Non-parametric`[[1]][2],
-  ECB_plot = plot_smooth_term(ECB_bamlasso_fnl, ECB_bamlasso_nonlnr, ECB_sm_obj$Smooth,
-                              min = min(ECB_cov[, ECB_bamlasso_nonlnr])-0.1,
-                              max = max(ECB_cov[, ECB_bamlasso_nonlnr]) + 0.1),
+  ECB_bamlasso_nonlnr = ECB_bamlasso_var$`Non-parametric`[[1]],
+  
+  tar_target(
+    ECB_plot_list,
+    plot_smooth_term(ECB_bamlasso_fnl, ECB_bamlasso_nonlnr, ECB_sm_obj$Smooth,
+                     min = min(ECB_cov[, ECB_bamlasso_nonlnr])-0.1,
+                     max = max(ECB_cov[, ECB_bamlasso_nonlnr]) + 0.1)+
+      xlab(ECB_bamlasso_nonlnr)+
+      theme_pubr()+
+      theme(axis.title.y = element_blank()),
+    pattern = map(ECB_bamlasso_nonlnr),
+    iteration = "list"
+  ),
+  
+  ECB_plot = ggarrange(plotlist = ECB_plot_list),
+  
+
+  
+  
   
   
   # ** SB-GAM ---------------------------------------------------------------
@@ -394,10 +409,23 @@ tar_plan(
   
   # *** Plot Non-Linear Functions ###
   # TODO: to generalize this for a list of plots.
-  WLM_bamlasso_nonlnr = WLM_bamlasso_var$`Non-parametric`[[1]][1],
-  WLM_plot = plot_smooth_term(WLM_bamlasso_fnl, WLM_bamlasso_nonlnr, WLM_sm_obj$Smooth,
-                              min = min(WLM_bgam_dat$train_dat[, WLM_bamlasso_nonlnr])-0.1,
-                              max = max(WLM_bgam_dat$train_dat[, WLM_bamlasso_nonlnr]) + 0.1),
+
+  
+  WLM_bamlasso_nonlnr = WLM_bamlasso_var$`Non-parametric`[[1]],
+  
+  tar_target(
+    WLM_plot_list,
[…]

### 2022-03-17 Manuscript/00-main.Rmd
Add real data graphs to the manuscript

@@
 tar_load(tbl_real_time)
+tar_load(ECB_plot)
+tar_load(WLM_plot)
 # tar_load(sim_tim_tab)
@@
 ```
+
+\clearpage
+\begin{figure}[h]
+\includegraphics{Figs/ECB_plot}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[h]
+\includegraphics{Figs/WLM_plot}
+\end{figure}

### 2022-03-17 _targets.R

@@ -64,7 +64,7 @@ tar_plan(
                                                 dist = "binomial"),
   
   # plot_var_select(sim_binom_var_select_raw[[1]])  
-
+  
   
   # Table Presentation of Var Selection
   tmp = make_sim_var_metric_raw(),
@@ -73,7 +73,7 @@ tar_plan(
                                                               sim_binom_comp_select_raw[[.x]])) %>%
     group_by(p, Method, Metric) %>%
     summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
-              ) %>%
+    ) %>%
     pivot_wider(names_from = Method, values_from = mean) %>% 
     ungroup(),
   
@@ -88,9 +88,9 @@ tar_plan(
   #   facet_wrap(~Metric, ncol = 3, nrow = 1, scales = "free"),
   
   
-
   
-
+  
+  
   
   
   #** Gaussian Outcome  ####
@@ -142,8 +142,8 @@ tar_plan(
     cat(file = "Manuscript/Tabs/sim_gaus_var_slct_tab.tex"),
   
   
-
-
+  
+  
   #** Simulation Time ####
   sim_tim_tab = make_time_table(sim_success_rate),
   sim_tim_tab_latex = sim_tim_tab %>% 
@@ -205,7 +205,7 @@ tar_plan(
   sim_lnr_binom_var_select_tbl = map_dfr(1:5, 
                                          ~make_sim_var_metric_raw(sim_lnr_binom_var_select_raw[[.x]],
                                                                   sim_lnr_binom_comp_select_raw[[.x]])
-                                         ) %>%
+  ) %>%
     group_by(p, Method, Metric) %>%
     summarise(mean = mean(value, na.rm = TRUE)#, sd = sd(value, na.rm = TRUE)
     ) %>%
@@ -247,7 +247,7 @@ tar_plan(
   sim_lnr_gaus_var_select_tbl = map_dfr(1:5, 
                                         ~make_sim_var_metric_raw(sim_lnr_gaus_var_select_raw[[.x]],
                                                                  sim_lnr_gaus_comp_select_raw[[.x]])
-                                        ) %>%
+  ) %>%
     group_by(p, Method, Metric) %>%
[…]

### 2022-03-18 Manuscript/05-Discussion.Rmd
Update Section Headers to Latex
revise discussion

@@
-# Discussion
+\section{Discussion}
 \label{sec:concl}
-In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection.
-Meanwhile, we introduced two optimization based algorithms for model fitting.
-The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner.
-Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.
+\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling.
+In particular, we introduce a two-part spike-and-slab lasso prior for re-parameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
+The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
+Th EM-CD algorithm, extended from previous spike-and-slab regression models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms for model fitting.
+Meanwhile, the two-part prior motivates the bi-level selection of functional forms of the predictors, selection of linear and nonlinear component.
+In the simulation and real-data analysis, the proposed model demonstrates improvement in prediction and computational advantage when compared to some state-of-the-art models.
+When serving the purpose of variable selection, trade-offs exist among methods of comparison.} We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
+To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones.
+Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.
 
 The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work.
-Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based fast and scalable algorithms.
-Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior.
-Nevertheless, SB-GAM fails to address the bi-level functional selection and the model inference.
-The proposed model provides solutions to these remaining questions while maintaining the same level of prediction accuracy.
-Moreover, the proposed model renders additional flexibility to model specification, allowing various choices of smoothing functions and degrees of freedom.
+Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based scalable algorithms.
+Nevertheless, Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior.
+\bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
+Moreover, the proposed model can be generalized to other family of priors or other family of smoothing functions if desired.
+Not focused in this manuscript, the generalization is described in the supporting materials.}
 
-For translational science purpose, we implemented the proposed model in a R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
-To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones.
-Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.
-In addition, in \texttt{BHAM}, we streamline the model fitting algorithms to support other popular Bayesian hierarchical prior for the smoothing functions, such as Student's T distribution, double exponential distribution, mixture normal and T distribution's.
-These priors could be helpful when the sparse assumption is weak or not necessary.,
+\bg{During designing and analyzing the simulation study, we made couple interesting observations.
+First of all, variable selection is a delicate topic in the context predictive modelling.
+When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example Lasso.
+[TODO: add Denilla paper] In addition, bi-level selection is a more complex problem than variable selection.
+The complexity reflects on the validity of effect hierarchy assumption.
+While most functional forms follow effect hierarchy that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
+$x^2$.
+The proposed prior and spikeSlabGAM employ different structure: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
+The different prior setups lead to trade-offs for the purpose of bi-level selection.
+We recommend to use more judgement in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
+Secondly, we find the performance of the proposed model are more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
+Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend to fine-tune the model with granular sequence of $s_0$ for performance improvement.}
 
-<!-- the implementation of ssGAM-NM using EM-IRLS.
-Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM.
-Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g.
-mixture t-distribution, and mixture double exponential distributions.
-@Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors, -->
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- \beta_i \sim N(0, \tau^2_i),\\ -->
-<!-- \tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2) -->
-<!-- \end{aligned} -->
-<!-- $$ -->
-<!-- with $\tau^2$ integrated out analytically.
-Hence, the EM-IRLS can be easily applied to those two additional mixture priors.
-One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g.
-$v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$.
-The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates.
-[@Rockova2014a] -->
-
-There are some improvements possible for the proposed models.
+<!-- There are some improvements possible for the proposed models.
 First of all, the proposed model achieves a bi-level selection via the two-part spike-and-slab spline prior.
@@
 [@Friedman2017] For the sake of model comparison, we fit the models using the same pool of predictors in Section \ref{sec:ECB}.
+-->
 
@@
 The similar strategy has been used in Ferrari and Dunson [@Ferrari2020].
-
-  <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function.
-The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder.
-In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.
--->
-
-<!-- The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not.
-On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e.
-$\gamma^\tp_{jk}$ of $\beta^\tp_{jk}$ for $k = 1\dots, K_j$.
-This prior set-up generalized to the group spike-and-slab.
-The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity.
-The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup.
-Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary.
--->

### 2022-03-19 Manuscript/00-main.Rmd
Revise Abstract

@@
 While some Bayesian models can address these shortcomings, using Markov chain Monte Carlo algorithms for model fitting creates a new challenge, scalability.
-Hence, we propose Bayesian hierarchical generalized additive models as a solution: we consider the smoothing penalty for proper shrinkage of curve interpolation and separation of smoothing function linear and nonlinear spaces.
-A novel spike-and-slab spline prior is proposed to select components of smoothing functions.
-Two scalable and deterministic algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least Squares, are developed for different utilities.
-Simulation studies and metabolomics data analyses demonstrate improved predictive or computational performance against state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM.
-The software implementation of the proposed models is freely available via an R package BHAM."
-keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility
+Hence, we propose Bayesian hierarchical generalized additive models as a solution: we consider the smoothing penalty for proper shrinkage of curve interpolation via reparameterization.
+A novel two-part spike-and-slab LASSO prior for smoothing functions is developed to address the sparsity of signals while providing extra flexibility to select the linear or nonlinear components of smoothing functions.
+A scalable and deterministic algorithm, EM-Coordinate Descent, is implemented in an open-source R package BHAM.
+Simulation studies and metabolomics data analyses demonstrate improved predictive and computational performance against state-of-the-art models.
+Functional selection performance suggests trade-offs exist regarding the effect hierarchy assumption."
+keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-Coordinate Decent; Scalablility; Predictive Modeling
 # bibliography: bibfile.bib

### 2022-03-21 Manuscript/01-Intro.Rmd
Revise Intro and Method Section

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, $Y$, which is assumed to follow some exponential family distribution, can be modeled with the summation of flexible smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$.
+In a GAM, the response variable, which is assumed to follow some exponential family distribution, can be modeled with the summation of smoothing functions.
 Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.
@@
 [@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
-Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models.\bg{[TODO: limit the scope, "particuaraly with spike-and-slab literatures."]} Various group spike-and-slab priors combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms [@Xu2015; @Yang2020] are proposed, where the application on AMs are by-products.\bg{[TODO: adjust citation position for clarity.]} Bai et al.
-[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs[@Bai2021].
-Focus on addressing the sparsity, these methods can overly penalize the basis function coefficients and produce inaccurate predictions and curve interpolation, particularly when complex signals are assumed and large number of knots are used.
+Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models, particularly within the spike-and-slab literature.
+Various group spike-and-slab priors [@Xu2015; @Yang2020] combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms are proposed, where the application on AMs are treated as a special case.
+Bai et al.
+[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs [@Bai2021].
+Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large number of knots are used.
 [@Scheipl2013] In addition, these methods adapt an 'all-in-all-out' strategy, i.e.
@@
 Scheipl et al.
-[@Scheipl2012] proposed a spike-and-slab structure prior that address the previous challenges.
+[@Scheipl2012] proposed a spike-and-slab structure prior that address the bi-level selection.
 But the model fitting relies on computational intensive MCMC algorithms and creates scalability concern.
@@
 
-\bg{To address these challenges, we propose a novel Bayesian hierarchical generalized additive model (BHAM) for high dimensional data analysis.
-[TODO: add a sentences on the contribution.
-Focuse on prediction of high-dimensional data, and possibly mentioning of vairbale selection.] Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature [TODO: add Woods citation], via re-parameterization [of the smoothing function] to avoid overly shrinking basis function coefficients.
-Smoothing penalties were also previously used in the spike-and-slab GAM [@Scheipl2012] and the sparsity-smoothness penalty [@Meier2009].
-We then impose a new two-part spike-and-slab spline prior on the smoothing functions for bi-level selection such that the linear and nonlinear spaces of smoothing functions can be selected separately.
+\bg{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
+Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smoothing functions to avoid excessive shrinkage on the bases.
+Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
+We then impose a new two-part spike-and-slab LASSO prior to address the sparsity of the signal.
 In addition, a scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm are developed.
-While the primary focus of this model is improvement in prediction, the proposed model also provides utility in fucntional selection.
-Particuaraly, the porposed prior setup encourages a bib-level selection/flexible solution, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects, as the re-parameterization allows the separation of the linear space of a smoothing function from the nonlinear space.
-The proposed model is implemented in an publicly available R package \texttt{BHAM} via [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM), making translational science more accessible.}
+While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection.
+Particularly, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
+The proposed model is implemented in an publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
 
@@
 They make fitting high-dimensional models more feasible without sacrificing performance in prediction and variable selection.
-Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-Coordinate Descent (EM-CD hereafter) algorithm to fit the model.
-While both methods use the combination of expectation maximization algorithm and coordient descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.
-The proposed model sets up independent priors among basis function coefficients after the re-parameterization step, which provides some advantage in computation.
-In addition, we offer an alternative algorithm, EM-IWLS, that can provide variance-covariance matrix of the coefficients.
+Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-CD algorithm to fit the model.
+While both methods use the combination of expectation maximization algorithm and coordinate descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.
+The proposed model sets up independent priors among basis coefficients after the reparameterization step, which provides some advantage in computation.
 Last but not least, the proposed model addresses the incapability of bi-level selection in SB-GAM.
 
-In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the fast-fitting EM-CD algorithmalgorithm.
-In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models, mgcv, COSSO and sparse Bayesian GAM \bg{[TODO: add other models]} via Monte Carlo simulation studies.
+In Section \ref{sec:BHAM}, we establish the Bayesian hierarchical generalized additive model, introduce the proposed spike-and-slab spline priors, and describe the fast-fitting EM-CD algorithm.
+In Section \ref{sec:sim}, we compare the proposed framework to state-of-the-art models via Monte Carlo simulation studies.
 Analyses of two metabolomics datasets are presented in Section \ref{sec:real_data}.

### 2022-03-21 Manuscript/02-Method.Rmd
Revise Intro and Method Section

@@
 The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
-$$
+\begin{equation}\label{eq:gam}
  E(Y|\bs x) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
-$$
+\end{equation}
 where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \bs x_i\}^n_{i=1}$,} the data distribution is expressed as
@@
 
-To encourage proper smoothing of each additive function, we adopt the smoothing penalty from smoothing spline models [TODOO: add Wood citation].
-A smoothing penalty, a function of the integration of the second derivative of the spline function.
-is quadratic norms of the coefficients \bg{of what?} and allow locally adaptive penalties on \bg{each smoothing function[TODO: not accurate, bases]}, mathematically
-\begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,
+To encourage proper smoothing of each additive function, we adopt the smoothing penalty from smoothing spline models[@Wood2017].
+A smoothing penalty is the quadratic norm of the basis coefficients and allows different shrinkage on different bases, mathematically
+\begin{equation}
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,\nonumber
 \end{equation}
 where $\bs S_j$ is a known smoothing penalty matrix and $\lambda_j$ denotes a smoothing parameter.
-A linear function can be modeled as $B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
-\bg{Unlike previous methods that embeds smoothing penalty in the sparse penalty which leads to a more restrictive solution, we considers an additional mechanism via the propsed prior (discussed in Section xxx) to address sparsity in signal.
-such that the locally adpative nature of the smoothing penalty retains.}
+A linear function can be modeled as $B_j(x_j) = x_j$ with the smoothing penalty matrix $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
+\bg{Unlike previous regularized methods that either ignore the smoothing penalty completely or restrain the smoothing penalty as a component of sparse penalty which leads to a more restrictive solution, we consider an additional mechanism in pair with the proposed prior (described in Section \ref{sec:method_prior}) to address the smoothness and sparsity in signals such that the locally adaptive nature of the smoothing penalty retains.}
 
-Marra and  Wood [@Marra2011] proposed a re-parameterization procedure \bg{to factor the smoothing penalty into the design matrix[TODO: vague]}.
-Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+Marra and  Wood [@Marra2011] proposed a reparameterization procedure \bg{to factor the smoothing penalty into the design matrix of each smoothing function}.
+Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
 To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
@@
 An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
-With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization.
+With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the reparameterization.
 A spline function can be expressed in the matrix form
@@
 
-To summarize, the re-parameterization step provides three benefits.
-First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for functional selection.
-Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear space from the nonlinear space, improving the feasibility of bi-level functional selection.
-The eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
-This reduces the computational complexity compared to using a multivariate priors, and greatly broadens the choices of priors and further model choices.
-Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of priors, from column-specific priors to a unified prior.
+To summarize, the reparameterization step provides three benefits.
+First of all, the reparameterization integrates the smoothing penalty matrix into the design matrix, and encourages models to properly smooth the nonlinear function when sparsity penalty exists.
+Secondly, the eigendecomposition of the smoothing penalty matrix allows the isolation of the linear space from the nonlinear space, improving the feasibility of bi-level functional selection.
+Last but not least, the eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
+This reduces the computational complexity compared to using a multivariate priors, and improve the generalizability of the framework to be compatible with other choices of priors.
 
-\subsection{Two-part Spike-and-Slab Lasso Prior for Smoothing Fucntions}
+\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Fucntions}\label{sec:method_prior}
 
-The family of spike-and-slab (SS) regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
-\bg{Among all the spike-and-slab priors, the spike-and-slab Lasso (SSL) prior [@Rockova2018b; @Rockova2018] is one of the most popular choices.
+The family of spike-and-slab regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
+\bg{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
 The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
@@
 \end{equation}
-Compared to other priors for high-dimensional data analysis, SSL has the following advantages.First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients.
-Hence, the smoothing functions can be estimated more accurately.
-Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
-Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+\bg{Compared to other priors for high-dimensional data analysis, SSL has the following advantages.
+First of all, the SSL prior provides a locally adaptive shrinkage when estimating the coefficients.
+Secondly, the SSL prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the SSL prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
 We defer to Bai et al.
-[@Bai2021Review] for an detailed discussion.
-Particularly, the development of SSL model substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
+\cite{Bai2021Review} for an detailed discussion.}
+
+<!-- Particularly, the development of SSL model substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
 We notice that Bai[TODO: add Bai arxiv paper] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [TODO: add Xu Ghosh (2015)] and limits to an “all-in-all-out” strategy for functional selection.
+-->
 
 We introduce a novel SSL-based prior for smoothing functions in GAMs.
-Given the re-parameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
+Given the reparameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
 Specifically, the linear space coefficient has a SSL prior and the nonlinear space coefficients shares a group SSL prior,
@@
 \end{align}
-where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.$s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
+where $\gamma_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}$ are two latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively.
+$s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given.
 These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation, discussed in Section \ref{sec:tune}.
 
-\bg{The group SSL prior of the nonlinear space coefficients differs from previous group SSL priors [@Tang2018; @Tang2019], as effect hierarchy are preferred when selecting linear effect and non-linear effect.
-Effect hierarchy refers to the principle that lower-order effects are more likely to be active than higher-order effects.
-[TODO: add chipman paper citation.] To implement, we consider the hyper prior of the shared latent indicator $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$.
-Both latent indicators $\gamma_j and \gamma^\tp_j$ follows a Bernoulli distribution, written as binomial distribution of one trial.
-While the probability of including the linear effect is $\theta_j$, a probability parameter, the probability of including the nonlinear effect is $\bg{\gamma_{j}}\theta_j$.}
+\bg{The proposed two-part SSL prior, particularly the group SSL prior of the nonlinear space coefficients, differs from previous group SSL priors \cite{Tang2018, Tang2019}, as the proposed prior follows the effect hierarchy principle.
+Effect hierarchy refers to the principle that "lower-order effects are more likely to be active than higher-order effects" defined by Chipman\cite{chipman2006prior}.
+To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$, while both latent indicators $\gamma_j and \gamma^\tp_j$ follow a Bernoulli distribution.
+While the probability of including the linear effect is $\theta_j$, the probability of including the nonlinear effect is $\bg{\gamma_{j}}\theta_j$.
 $$
@@
 \bg{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For the computational convenience in the proposed EM-CD algorithm (see Section [TODO: add section ref]), we analytically integrate $\theta_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$.
+For the computational convenience, we analytically integrate $\theta_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$.
 }
@@
 The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience.
-For simplicity, we focus on a special case of beta distribution, uniform (0,1), i.e.
-$a = 1, b = 1$.
-\bg{This enables the proposed prior to inherit the selective shrinkage property and self-adaptivity [TODO: add Bai review] from the classical SSL prior.
-In other words, when a smoothing function is significant, the coefficients of the smoothing function will escape the overall shrinkage and produce a more accurate estimate, particular with the smoothing penalty addressed implicitly via the reparameterization.
-Meanwhile, automatic adjustment for different level of sparsity are feasible via the information borrowing cross coordinates.}
+<!-- For simplicity, we focus on a special case of beta distribution, uniform (0,1), i.e. $a = 1, b = 1$. -->
+\bg{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
+In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particular in pair with the smoothing penalty implicitly addressed via the reparameterization.
+Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different level of sparsity.}
 <!-- This informaiton borrowing helps the model to find adequate shrinkage for the linear coefficient and coefficients of the nonlinear part.
@@
 Hereafter, we refer Bayesian hierarchical generalized additive models with the two-part spike-and-slab LASSO prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
-\begin{center}
-Figure \ref{fig:SSprior} here
-\end{center}
 
@@
 
-\bg{Despite the advantage to estimate posterior densities and derive uncertainty inference, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
-Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) [@George1997], and substaintially slows as $p$ increases modestly in hte GAM context [@Scheipl2013].
-Hence, we considers optimizaiton algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
-Optimizaiton algorithms greatly speed up the model fitting process, and improve the scalability of BHAM.}
+\bg{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
+Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
+Hence, we consider the optimization algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
+Specifically, we
+extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
+Similar to the EMVS algorithm \cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is based on the the expectation-maximization (EM) algorithm, integrating Coordinate Descent algorithm in each iterative step to find the posterior mode.
+The EM-CD algorithm has been well adapted in generalized linear models \cite{Tang2017a}, Cox proportional hazards models \cite{Tang2017}, and their grouped counterparts \cite{Tang2018, Tang2019}.
+The EM-CD algorithm provides deterministic solutions, which becomes a popular property for reproducible research.}
 
-In this section, we extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
-\bg{Specifically for the SSL priors, Rochova and co-authors [@Rockova2018b; @Rockova2018] further integrate the EM algorithm and coordinate descent algorithm for fast computation.
-Yi and his group independently developed the EM-Iterative Weighted Least Square and EM-cyclic coordinate descent algorithms to fit models with broader class of priors, SSL included.[@Yi2019] These algorithms were implemented for generalized linear models[@Tang2017a], Cox proportional hazards models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019].
-Both EM based algorithms provide deterministic solutions, which becomes a popular property for reproducible research.} To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
+<!-- To note, SB-GAM[@Bai2020; @Bai2021] also used an EM-CD algorithm.
 The main difference between proposed EM-CD algorithm and that in SB-GAM is that SB-GAM uses a block CD algorithm for their group prior, while the proposed prior is pairwise independent requiring no special treatment in the CD algorithm.
+-->
 
@@
 
-For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data\bg{} for the EM algorithm}.
-Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently, the logarithm of the density function,
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data in the EM context.
+Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently the logarithm of the density function,
 $$
@@
 $$  
-where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(\bg{a},\bg{b}) density.
+where $f(\textbf{y}|\bs \beta, \phi)$ is the data distribution and $f(\theta)$ is the Beta(a, b) density.
 We choose non-informative prior for the intercept $\beta_0$ and the dispersion parameter $\phi$; for example, $f(\beta_0|\tau_0^2)=N(0,\tau_0^2)$ with $\tau^2_0$ set to a large value and $f(\log \phi) \propto 1$.
 
-We use the EM algorithm to find the the maximum a posteriori estimate of $\Theta$.
-This is, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the values from previous iteration $\Theta^{(t-1)}$,
+We use the EM algorithm to find the maximum a posteriori estimate of $\Theta$.
+This is, in the E-step, we calculate the expectation of posterior density function of $\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$ with respect to the latent indicators $\bs \gamma$ conditioning on the parameter values from previous iteration $\Theta^{(t-1)}$,
 $$
@@
 Hereafter, we use the shorthand notation $E(\cdot)\equiv E_{\bs \gamma|\Theta^{(t-1)}}(\cdot)$.
-In the M-step, we find the $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
-\bg{The parenthesized subscription denotes the parameter estimation at the $t$th iteration.} The E- and M- steps are iterated until the algorithm converge.
+In the M-step, we find the parameters $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
+\bg{The parenthesized subscription $^{(t)}$ denotes the parameter estimation at the $t$th iteration.} The E- and M- steps are iterated until the algorithm converges.
 
@@
 Meanwhile, conditioning on $\bs \gamma$, $Q_1$ and $Q_2$ are independent and can be maximized separately for $\bs \beta, \phi$ and $\bs \theta$.
-With the proposed two-part spike-and-slab lasso prior, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via the Coordinate Descent algorithm in each iteration.
+With the proposed two-part spike-and-slab LASSO prior, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via the Coordinate Descent algorithm in each iteration.
 Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
@@
 To calculate two unknown quantities $E({S_j}^{-1})$ and $E({S^\tp}^{-1}_j)$, the posterior probability $p_{j} \equiv \pr(\gamma_{j}=1|\Theta^{(t-1)})$ and $p_{j}^\tp \equiv \pr(\gamma^\tp_{j}=1|\Theta^{(t-1)})$ are necessary, which can be derived via Bayes' theorem.
-The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the value of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
-The calculation follows the equations below.
+The calculation of $p_j^\tp$ is slightly different from that of $p_j$, as $p_j^\tp$ depends on the values of the vector $\bs \beta^\tp_{j}$ and $p_j$ only depends on the scalar $\beta_j$.
+The calculation follows the equations below,
 \begin{align*}
@@
 \end{align*}
-and $\pr(\gamma_{j} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+where $\pr(\gamma_{j} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
-\begin{align}\label{eq:exp_scale}
+\begin{align*}\label{eq:exp_scale}
 &E(\gamma_{j})  = p_{j} & &E(\gamma^{\tp}_{j}) = p_{j}^{\tp}\nonumber\\
-&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^\tp}^{-1}_{j}) = \frac{1-p_{j}^{\tp}}{s_0} + \frac{p_{j}^{\tp}}{s_1}.
-\end{align}
-After replacing withe calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm [@Friedman2010].
+&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S_{j}^\tp}^{-1}) = \frac{1-p_{j}^{\tp}}{s_0} + \frac{p_{j}^{\tp}}{s_1}.
+\end{align*}
+After replacing the calculated quantities, $E(Q_1)$ can be seen as a $l_1$ penalized likelihood function with the regularization parameter $\lambda = E(S^{-1})$, and hence be optimized via coordinate descent algorithm [@Friedman2010].
 Independently, the remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
@@
 
-Totally, the proposed EM-coordinate descent algorithm is summarized as follows:
+Totally, the proposed EM-CD algorithm is summarized as follows:
 
@@
 \label{sec:tune}
-Our proposed models, BHAM, require two preset scale parameters ($s_0$, $s_1$).
+Our proposed model, BHAM, requires two preset scale parameters ($s_0$, $s_1$).
 Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding a criteria of preference.
 This would be achieved by constructing a two-dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
-However, previous research suggested the value of slab scale $s_1$ have less impact on the final model and is recommended to be set as a generally large value, e.g.
+However, previous research suggests the value of slab scale $s_1$ has less impact on the final model and is recommended to be set as a generally large value, e.g.
 $s_1 = 1$, that provides no or weak shrinkage.
@@
 deviance \bg{(defined as model log-likelihood times -2, $-2\log f(\bs y|\bs{\hat\beta}, \hat\phi)$)}, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
-The procedure is similar to the Lasso implemented in the widely used R package `glmnet`, which quickly fits Lasso models over a list of values of regularization parameters $\lambda$ , giving a sequence of models for users to choose from.
+The procedure is similar to the LASSO implementation in the widely used R package `glmnet`, which quickly fits LASSO models over a list of values of regularization parameters $\lambda$ and gives a sequence of models for users to choose from.

### 2022-03-21 Manuscript/02-Method.Rmd
Revise Discussion

@@
 \bg{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
-The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of covariates $\bs x$, written as
+The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\bs x$, written as
 \begin{equation}\label{eq:gam}

### 2022-03-21 Manuscript/05-Discussion.Rmd
Revise Discussion

@@
 \bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling.
-In particular, we introduce a two-part spike-and-slab lasso prior for re-parameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
 The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
-Th EM-CD algorithm, extended from previous spike-and-slab regression models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms for model fitting.
-Meanwhile, the two-part prior motivates the bi-level selection of functional forms of the predictors, selection of linear and nonlinear component.
-In the simulation and real-data analysis, the proposed model demonstrates improvement in prediction and computational advantage when compared to some state-of-the-art models.
+Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
+In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear component.
+In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
 When serving the purpose of variable selection, trade-offs exist among methods of comparison.} We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
@@
 
-The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work.
-Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based scalable algorithms.
-Nevertheless, Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior.
+The proposed model shares many commonality with the SB-GAM [@Bai2021], independently developed around the same time of the proposed work.
+Both frameworks emphasize computational efficiency by deploying group spike-and-slab LASSO type priors and optimization-based scalable algorithms.
+Bai provides the theoretical proof for the consistency of variable selection using group spike-and-slab LASSO prior.
 \bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can be generalized to other family of priors or other family of smoothing functions if desired.
+Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.
 Not focused in this manuscript, the generalization is described in the supporting materials.}
@@
 First of all, variable selection is a delicate topic in the context predictive modelling.
-When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example Lasso.
-[TODO: add Denilla paper] In addition, bi-level selection is a more complex problem than variable selection.
-The complexity reflects on the validity of effect hierarchy assumption.
-While most functional forms follow effect hierarchy that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
+When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example LASSO and LASSO-based models.
+\cite{Wu2019} Moreover, bi-level selection is a more complex problem than variable selection.
+The complexity reflects on the validity of the effect hierarchy principle.
+While most functional forms follow that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
 $x^2$.
@@
 
-Our future efforts direct to modeling survival outcomes and integrative analysis.
+Our future efforts direct to survival analysis and integrative analysis.
 While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.

### 2022-03-21 Manuscript/SS_GAM.tex
Revise Discussion

@@
 solution: we consider the smoothing penalty for proper shrinkage of
-curve interpolation and separation of smoothing function linear and
-nonlinear spaces. A novel spike-and-slab spline prior is proposed to
-select components of smoothing functions. Two scalable and deterministic
-algorithms, EM-Coordinate Descent and EM-Iterative Weighted Least
-Squares, are developed for different utilities. Simulation studies and
-metabolomics data analyses demonstrate improved predictive or
-computational performance against state-of-the-art models, mgcv, COSSO
-and sparse Bayesian GAM. The software implementation of the proposed
-models is freely available via an R package BHAM.}
+curve interpolation via reparameterization. A novel two-part
+spike-and-slab LASSO prior for smoothing functions is developed to
+address the sparsity of signals while providing extra flexibility to
+select the linear or nonlinear components of smoothing functions. A
+scalable and deterministic algorithm, EM-Coordinate Descent, is
+implemented in an open-source R package BHAM. Simulation studies and
+metabolomics data analyses demonstrate improved predictive and
+computational performance against state-of-the-art models. Functional
+selection performance suggests trade-offs exist regarding the effect
+hierarchy assumption.}
 
 \keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
-Additive Models; EM-IWLS; EM-Coordinate Decent; Scalablility}
+Additive Models; EM-Coordinate Decent; Scalablility; Predictive
+Modeling}
 
@@
 relevant to the data generation process that one can draw inferences
-from. In addition, how to generalize these models to the
-high-dimensional setting remains unclear.
-
-Nonparametric regression models are appropriate alternatives to the
-``black-box'' models thanks to their balance between model flexibility
-and interpretability. Among those, generalized additive models (GAMs),
-proposed in the seminal work of Hastie and Tibshirani
-\citep{Hastie1987}, grew to be one of the most popular modeling tools.
-In a GAM, the response variable, \(Y\), which is assumed to follow some
-exponential family distribution with mean \(\mu\) and dispersion
-\(\phi\), can be modeled with the summation of smoothing functions,
-\(B_j(\cdot), j = 1, \dots, p\), of a given \(p\)-dimensional vector of
-covariates \(\boldsymbol{x}\), written as \[
- E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)),
-\] where \(g^{-1}(\cdot)\) is the inverse of a monotonic link function.
-The smoothing functions can take many forms and are estimated using a
-pseudo-weighted version of the backfitting algorithm
-\citep{Breiman1985}. Nevertheless, the classical GAMs cannot fulfill the
-increasing analytic demands for high-dimensional data analysis in
-biomedical studies.
+from. In addition, how to generalize these \textcolor{red}{"black box"}
+models to the high-dimensional setting remains unclear.
+
+\textcolor{red}{For their easy interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in the context of prediction and variable selection}.
+Among those, generalized additive models (GAMs), proposed in the seminal
+work of Hastie and Tibshirani \citep{Hastie1987}, grew to be one of the
+most popular modeling tools. In a GAM, the response variable, which is
+assumed to follow some exponential family distribution, can be modeled
+with the summation of smoothing functions. Nevertheless, the classical
+GAMs cannot fulfill the increasing analytic demands for high-dimensional
+data analysis.
 
@@
 models. Recently Bayesian hierarchical models are also used in the
-context of high-dimensional additive models. Various group
-spike-and-slab priors combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms \citep{Xu2015, Yang2020} are
-proposed, where the application on AMs are by-products. Bai et al.
-\citep{Bai2020} was the first to apply group spike-and-slab lasso prior
-to Gaussian AMs using a fast optimization algorithm, and further
-generalized the framework to GAMs\citep{Bai2021}. Focus on addressing
-the sparsity, these methods can overly penalize the basis function
-coefficients and produce inaccurate predictions and curve interpolation,
-particularly when complex signals are assumed and large number of knots
-are used. \citep{Scheipl2013} In addition, these methods adapt an
-`all-in-all-out' strategy, i.e.~either including or excluding the
-variable completely, rendering no space for bi-level selection. Scheipl
-et al. \citep{Scheipl2012} proposed a spike-and-slab structure prior
-that address the previous challenges. But the model fitting relies on
-computational intensive MCMC algorithms and creates scalability concern.
-It would be of special interest to develop a fast, flexible and accurate
-generalized additive model framework.
-
-To address these challenges, we propose a novel Bayesian hierarchical
-generalized additive model (BHAM) for high dimensional data analysis.
-Specifically, we incorporate smoothing penalties in the model via
-re-parameterization of the smoothing function to avoid overly shrinking
-basis function coefficients. Smoothing penalties are commonly
-implemented in the classical GAMs through smoothing regression splines.
-They are quadratic norms of the coefficients and allow locally adaptive
-penalties on each smoothing function. A smoothing penalty conditioning
-on a smoothing parameter \(\lambda_j\) is a function of the integration
-of the second derivative of the spline function, expressed
-mathematically as \begin{equation}\label{eq:smoothpen}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,
-\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
-matrix, and \(\boldsymbol{\beta}_j\) are the basis function
-coefficients. Smoothing penalties were also previously used in the
-spike-and-slab GAM \citep{Scheipl2012} and the sparsity-smoothness
-penalty \citep{Meier2009}. Moreover, incorporating the smoothing penalty
-allows the separation of the linear space of a smoothing function from
-the nonlinear space. We then impose a new two-part spike-and-slab spline
-prior on the smoothing functions for bi-level selection such that the
-linear and nonlinear spaces of smoothing functions can be selected
-separately. The prior setup encourages a flexible solution, rendering
-one of three possibilities for each predictor: no effect, only linear
-effect, or linear and nonlinear effects. In addition, two scalable
-optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm
-and EM-Iterative weighted least square (EM-IWLS) algorithm, are
-developed and implemented in an publicly available R package
-\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}, making
-translational science more accessible.
+context of high-dimensional additive models, particularly within the
+spike-and-slab literature. Various group spike-and-slab priors
+\citep{Xu2015, Yang2020} combining with computationally intensive Markov
+chain Monte Carlo (MCMC) algorithms are proposed, where the application
+on AMs are treated as a special case. Bai et al. \citep{Bai2020} was the
+first to apply group spike-and-slab lasso prior to Gaussian AMs using a
+fast optimization algorithm, and further generalized the framework to
+GAMs \citep{Bai2021}. Focus on addressing the sparsity, these methods
+can excessively penalize the bases of a smoothing function and produce
+inaccurate predictions, particularly when complex signals are assumed
+and large number of knots are used. \citep{Scheipl2013} In addition,
+these methods adapt an `all-in-all-out' strategy, i.e.~either including
+or excluding the variable completely, rendering no space for bi-level
+selection. Scheipl et al. \citep{Scheipl2012} proposed a spike-and-slab
+structure prior that address the bi-level selection. But the model
+fitting relies on computational intensive MCMC algorithms and creates
+scalability concern. It would be of special interest to develop a fast,
+flexible and accurate generalized additive model framework.
+
+\textcolor{red}{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
+Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smoothing functions to avoid excessive shrinkage on the bases.
+Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
+We then impose a new two-part spike-and-slab LASSO prior to address the sparsity of the signal.
+In addition, a scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm are developed.
+While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection.
+Particularly, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
+The proposed model is implemented in an publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
 
@@
 model. While both methods use the combination of expectation
-maximization algorithm and coordient descent algorithm, there are subtle
-difference in the implementation due to the difference in prior
+maximization algorithm and coordinate descent algorithm, there are
+subtle difference in the implementation due to the difference in prior
 specification. The proposed model sets up independent priors among basis
-function coefficients after the re-parameterization step, which provides
-some advantage in computation. In addition, we offer an alternative
-algorithm, EM-IWLS, that can provide variance-covariance matrix of the
-coefficients. Last but not least, the proposed model addresses the
-incapability of bi-level selection in SB-GAM.
+coefficients after the reparameterization step, which provides some
+advantage in computation. Last but not least, the proposed model
+addresses the incapability of bi-level selection in SB-GAM.
 
@@
 generalized additive model, introduce the proposed spike-and-slab spline
-priors, and describe the two fast-fitting algorithms. In Section
+priors, and describe the fast-fitting EM-CD algorithm. In Section
 \ref{sec:sim}, we compare the proposed framework to state-of-the-art
-models, mgcv, COSSO and sparse Bayesian GAM via Monte Carlo simulation
-studies. Analyses of two metabolomics datasets are presented in Section
-\ref{sec:real_data}. Conclusion and discussions are given in Section
-\ref{sec:concl}.
-
-\section{Bayesian Hierarchical Additive Models (BHAMs)}
+models via Monte Carlo simulation studies. Analyses of two metabolomics
+datasets are presented in Section \ref{sec:real_data}. Conclusion and
+discussions are given in Section \ref{sec:concl}.
+
+\section{Bayesian Hierarchical Additive Models (BHAM)}
 \label{sec:BHAM}
 
-Following the GLM notation introduced in Section \ref{sec:intro}, we
-have a generalized additive model with link function \(g(\cdot)\) and
-linear predictor \begin{equation}\label{eq:gam}
-\eta = \beta_0 + \sum\limits^p_{j=1}B_j(x_j) = \beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j,
-\end{equation} with smoothing functions \(B_j(x_j)\) of the variable
-\(x_j, j = 1, \dots, p.\) The outcome \(Y\) follows a exponential family
-distribution with density function \(f(y)\), mean \(\mu = g^{-1}(\eta)\)
-and dispersion parameter \(\phi\), and the data distribution is
-\begin{equation}
-f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi),\nonumber
-\end{equation} The basis function matrix, i.e.~the design matrix derived
-from the smoothing function \(B_j(x_j)\), is denoted
-\(\boldsymbol{X}_j\) for the variable \(x_j\). The dimension of the
-design matrix depends on the choice of the smoothing function, and is
-denoted as \(K_j\) for \(x_j\). Its corresponding smoothing penalty
-matrix is denoted as \(\boldsymbol{S}_j\). \(\boldsymbol{\beta}_j\)
-denotes the basis function coefficients for the \(j\)th variable such
-that \(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
+\textcolor{red}{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
+The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\boldsymbol{x}$, written as
+\begin{equation}\label{eq:gam}
+ E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
+\end{equation}
+where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \boldsymbol{x}_i\}^n_{i=1}$,}
+the data distribution is expressed as \begin{equation}
+f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi).\nonumber
+\end{equation}
+
+The basis function matrix, i.e.~the design matrix derived from the
+smoothing function \(B_j(x_j)\), is denoted \(\boldsymbol{X}_j\) for the
+variable \(x_j\). The dimension of the design matrix depends on the
+choice of the smoothing function, and is denoted as \(K_j\) for \(x_j\).
+\(\boldsymbol{\beta}_j\) denotes the basis function coefficients for the
+\(j\)th variable such that
+\(B_j(x_j) = \boldsymbol{\beta}_j^T \boldsymbol{X}_j\). With slight
 abuse of notation, we denote vectors and matrices in bold fonts
@@
 \(\beta, X\). The matrix transposing operation is denoted with a
-superscript \(^T\). To note, the proposed model can include parametric
-forms of variables in the model, treating the parametric function as a
-special case of the smoothing function, e.g.~\(B_j(x_j) = x_j\) with the
-smoothing penalty matrix defined as
-\(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
-
-To encourage proper smoothing of additive functions, we adopt the idea
-of smoothing penalties from smoothing spline models. The basic idea is
-to set up a smoothing penalty, described in Equation
-(\ref{eq:smoothpen}), in the prior density function. However, the direct
-integration of smoothing penalty with sparsity penalty is not obvious.
-Marra and Wood \citep{Marra2011} proposed a re-parameterization
-procedure to accommodate the smoothing penalty implicitly. Given the
-smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and positive
-semi-definite for univariate smoothing functions, we apply
-eigendecomposition on the penalty matrix
+superscript \(^T\).
+\textcolor{red}{To note, the proposed model can include parametric forms of variables in the model, and hence considers general linear models and semiparametric regression models as special cases.}
+
+\subsection{Smoothing Function Reparameterization}
+
+To encourage proper smoothing of each additive function, we adopt the
+smoothing penalty from smoothing spline models\citep{Wood2017}. A
+smoothing penalty is the quadratic norm of the basis coefficients and
+allows different shrinkage on different bases, mathematically
+\begin{equation}
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,\nonumber
+\end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
+matrix and \(\lambda_j\) denotes a smoothing parameter. A linear
+function can be modeled as \(B_j(x_j) = x_j\) with the smoothing penalty
+matrix \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
+\textcolor{red}{Unlike previous regularized methods that either ignore the smoothing penalty completely or restrain the smoothing penalty as a component of sparse penalty which leads to a more restrictive solution, we consider an additional mechanism in pair with the proposed prior (described in Section \ref{sec:method_prior}) to address the smoothness and sparsity in signals such that the locally adaptive nature of the smoothing penalty retains.}
+
+Marra and Wood \citep{Marra2011} proposed a reparameterization procedure
+\textcolor{red}{to factor the smoothing penalty into the design matrix of each smoothing function}.
+Given the smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and
+positive semi-definite for the univariate smoothing functions, we
+eigendecompose the penalty matrix
 \(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
@@
 as the basis function matrix for the \(j\)th variable after the
-re-parameterization. A spline function can be expressed in the matrix
+reparameterization. A spline function can be expressed in the matrix
 form \[
@@
 
-The re-parameterization step sets up the foundation of the proposed
-model and provides three benefits. First of all, the re-parameterization
-integrates the smoothing penalty into the design matrix, and encourages
-models to properly smooth the nonlinear function in addition to the
-sparse penalty for functional selection. Secondly, the
-eigendecomposition of the smoothing penalty allows the isolation of the
-linear space from the nonlinear space, improving the feasibility of
-bi-level functional selection. The eigendecomposition facilitates the
-construction of orthonormal design matrix, which makes imposing
-independent priors on the coefficients possible. This reduces the
-computational complexity compared to using a multivariate priors, and
-greatly broadens the choices of priors and further model choices. Last
-but not least, the scaling step of \(\boldsymbol{X}^*\) further
-simplifies the choice of priors, from column-specific priors to a
-unified prior.
-
-\subsection{Spike-and-Slab Spline Priors}
-
-The family of spike-and-slab (SS) regression models is one of most
-commonly used models in high-dimensional data analysis for its utility
-in outcome prediction and variable selection. We defer to Bai et al.
-\citep{Bai2021Review} for an in-depth introduction to spike-and-slab
-priors. To summarize, spike-and-slab priors are a family of mixture
-distributions that comprises a skinny spike density
-\(f_{\text{spike}}(\cdot)\) for weak signals and a flat slab density
-\(f_{\text{slab}}(\cdot)\) for strong signals, mathematically \[
- \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
-\] The most distinct feature of SS priors is that it is conditioned on a
-latent binary variable \(\gamma \in \{0,1\}\) that indicates whether the
-variable \(x\) is included in the model. There are various
-spike-and-slab priors depending on the choice for the spike density
-\(f_{\text{spike}}(\cdot)\) and the slab density
-\(f_{\text{slab}}(\cdot)\), see George and McCulloch
-\citep{George1993, George1997}; Chipman \citep{Chipman1996} for grouped
-variables; Brown et al. \citep{Brown1998} for multivariate outcomes;
-Ishwaran and Rao \citep{Ishwaran2005}; Clyde and George
-\citep{Clyde2004} and reference therein. Two of most popular spike and
-slab priors are spike-and-slab normal prior \citep{George1993} and
-spike-and-slab double exponential prior\citep{Rockova2018}.
-
-The major criticism of early spike-and-slab models is being
-computationally prohibitive. \citep{Bai2021Review} Since then, many
-studies focus on alleviating the computational burden that sampling
-algorithms bear, which include EMVS based on spike-and-slab normal
-prior\citep{Rockova2014a} and spike-and-slab Lasso (SSL)
-\citep{Rockova2018b, Rockova2018}. Particularly, the development of SSL
-model substantially improves the scalability of SS models, setting up
-the theoretical foundation for generalized models in -omics data
-analysis \citep{Tang2017a, Tang2017, Tang2018, Tang2019}. The SSL prior
-is composed of two double exponential distributions with mean 0 and
-different dispersion parameters, \(0 < s_0 < s_1\), mathematically,
+To summarize, the reparameterization step provides three benefits. First
+of all, the reparameterization integrates the smoothing penalty matrix
+into the design matrix, and encourages models to properly smooth the
+nonlinear function when sparsity penalty exists. Secondly, the
+eigendecomposition of the smoothing penalty matrix allows the isolation
+of the linear space from the nonlinear space, improving the feasibility
+of bi-level functional selection. Last but not least, the
+eigendecomposition facilitates the construction of orthonormal design
+matrix, which makes imposing independent priors on the coefficients
+possible. This reduces the computational complexity compared to using a
+multivariate priors, and improve the generalizability of the framework
+to be compatible with other choices of priors.
+
+\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Fucntions}\label{sec:method_prior}
+
+The family of spike-and-slab regression models is one of most commonly
+used models in high-dimensional data analysis for its utility in outcome
+prediction and variable selection.
+\textcolor{red}{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
+The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
 \begin{equation} 
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation} Given that both double exponential distributions have a
-mean of 0 and the latent indicator \(\gamma\) can only take the value of
-0 or 1, the mixture double exponential distribution can be formulated as
-one single double exponential density, \begin{equation} \label{eq:ssl}
-\beta | \gamma \sim DE(0, S), 0 < s_0 < s_1,
-\end{equation} with the scale parameter
-\(S = (1-\gamma)s_0 + \gamma s_1\). The SSL also mitigates the problem
-of EMVS where the weak signals are not shrink to zero, and hence is
-preferred in high-dimensional data analysis. We notice that Bai
-\citep{Bai2021} is the first to apply spike-and-slab lasso prior in the
-GAM framework, where the densities of the spike and slab components take
-the group lasso density \citep{Xu2015} and limits to an
-``all-in-all-out'' strategy for functional selection.
-
-\subsubsection{Two-part Spike-and-Slab Lasso Prior}
-
-We introduce a novel prior for GAMs, particularly for high-dimensional
-nonlinear modeling with bi-level selection. The proposed prior extends
-from the spike-and-slab lasso prior described in Equation
-(\ref{eq:ssl}). Given the re-parameterized design matrix
+\end{equation}
+} The latent binary variable \(\gamma \in \{0,1\}\) indicates whether a
+variable \(x\) is included in the model, while the dispersion parameters
+\(s_0\) and \(s_1\) controls the shrinkage of the coefficient. Given
+that both double exponential distributions have a mean of 0 and the
+latent indicator \(\gamma\) can only take the value of 0 or 1, the
+mixture double exponential distribution can be formulated as one single
+double exponential density, \begin{equation} \label{eq:ssl}
+\beta | \gamma \sim DE(0, (1-\gamma)s_0 + \gamma s_1), 0 < s_0 < s_1.
+\end{equation}
+\textcolor{red}{Compared to other priors for high-dimensional data analysis, SSL has the following advantages.
+First of all, the SSL prior provides a locally adaptive shrinkage when estimating the coefficients.
+Secondly, the SSL prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the SSL prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+We defer to Bai et al.
+\cite{Bai2021Review} for an detailed discussion.}
+
+We introduce a novel SSL-based prior for smoothing functions in GAMs.
+Given the reparameterized design matrix
 \(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
@@
 \(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
-Specifically, we impose independent group priors on the linear space
-coefficients and on the nonlinear space coefficients respectively,
+Specifically, the linear space coefficient has a SSL prior and the
+nonlinear space coefficients shares a group SSL prior,
 \begin{align}\label{eq:bham_ssl}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
+  \beta_{j} | \gamma_{j},s_0,s_1 &\sim DE(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1) \nonumber \\
   \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j
@@
 parameters \(s_0\) and \(s_1\) can be treated as tuning parameters and
-optimized via cross-validation. A discussion of how to choose the scale
-parameters comes in Section \ref{sec:tune}. To note, this prior differs
-from previous group SSL priors \citep{Tang2018, Tang2019}, as the
-\(\beta_j\) and \(\boldsymbol{\beta}^*_j\) have different indicator
-variables \(\gamma_j\), \(\gamma_j^*\) respectively. It is possible to
-add a more restrictive assumption on the priors, assuming that one
-indicator variable decides the inclusion of both the linear effect and
-nonlinear effect, i.e.~\(\gamma_j = \gamma^*_j\). This converges to the
-SB-GAM \citep{Bai2021}. Conversely, it is also possible to relax the
-assumption such that each coefficient
-\(\beta_{jk} \in \boldsymbol{\beta}^*_j\) has its own latent indicator
-\(\gamma_{jk}\), but at the cost of complicating the bi-level functional
-selection. This reduces the proposed prior to the classic SSL prior.
-
-The re-parameterization introduced in Section \ref{sec:BHAM} grants the
-validity of the proposed prior. First of all, the smoothing function
-bases are linear dependent and necessitate extra attention. The
-eigeondecomposition remedies the problem and hence our prior can be set
-to be conditionally independent. Secondly, the eigenvalue scaling
-provides a panacea to allow unified scale parameters for all bases of
-all smoothing functions.
-
-The rest of the hierarchical prior follows the traditional SSL prior: we
-set up hyper-priors on \(\gamma_j, \gamma^*_j\) to allow local adaption
-of the shrinkage using a Bernoulli distribution, written as binomial
-distribution of one trial. The two indicators of the \(j\)th predictor,
-\(\gamma_j\) and \(\gamma^*_j\), shares the same probability parameter
-\(\theta_j\), \[
+optimized via cross-validation, discussed in Section \ref{sec:tune}.
+
+\textcolor{red}{The proposed two-part SSL prior, particularly the group SSL prior of the nonlinear space coefficients, differs from previous group SSL priors \cite{Tang2018, Tang2019}, as the proposed prior follows the effect hierarchy principle.
+Effect hierarchy refers to the principle that "lower-order effects are more likely to be active than higher-order effects" defined by Chipman\cite{chipman2006prior}.
+To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^*_j$}
+depends on the value of the linear space latent indicator \(\gamma_j\),
+while both latent indicators \(\gamma_j and \gamma^*_j\) follow a
+Bernoulli distribution. While the probability of including the linear
+effect is \(\theta_j\), the probability of including the nonlinear
+effect is \(\textcolor{red}{\gamma_{j}}\theta_j\). \[
 \begin{aligned}
@@
 \]
-\textcolor{red}{Equivalently, we can derive the marginal distribution of $\gamma_{j}^*$ by integrate out $\gamma_j$ (Supplement 1), 
-$$
-\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2).
-$$
+\textcolor{red}{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
+For the computational convenience, we analytically integrate $\theta_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$.
 }
 
-This is to leverage the fact that the probability of selecting the bases
-of a smoothing function should be similar, while allowing different
-penalty on the linear space and non-linear space of the smoothing
-function. The hyper prior of \(\gamma_{j}\) decides the sparsity of the
-model at the functional selection level, while that of \(\gamma_{j}^*\)
-decides the smoothness of the spline function at basis function level.
-Meanwhile, we specify that \(\gamma_{j}\) and \(\gamma_{j}^*\) are
-independently distributed for analytic simplicity. We further specify
-the parameter \(\theta_j\) follows a beta distribution with given shape
-parameters \(a\) and \(b\), \[
+\textcolor{red}{To allow the shrinkage to self-adapt to the sparsity and smoothing pattern of the data, }we
+further specify the parameter \(\theta_j\) follows a beta distribution
+with given shape parameters \(a\) and \(b\), \[
 \theta_j \sim Beta(a, b).
 \] The beta distribution is a conjugate prior for the binomial
-distribution and hence provides some computation convenience. For
-simplicity, we focus on a special case of beta distribution, uniform
-(0,1), i.e.~\(a = 1, b = 1\). When the variable have large effects in
-any of the bases, the parameter \(\theta_j\) will be estimated large,
-which in turn encourages the model to include the rest of bases.
+distribution and hence provides some computation convenience.
+\textcolor{red}{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
+In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particular in pair with the smoothing penalty implicitly addressed via the reparameterization.
+Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different level of sparsity.}
 Hereafter, we refer Bayesian hierarchical generalized additive models
-with the spike-and-slab spline prior as the BHAM, and visually presented
-in Figure \ref{fig:SSprior}.
-
-\begin{center}
-Figure \ref{fig:SSprior} here
-\end{center}
-
-\subsubsection{Other Priors}
-
-With the re-parameterization step of the basis function matrix
-\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
-priors, for example normal priors for ridge-type regularization and
-mixture normal prior for spike-and-slab regularization. These priors
-would work better in low and medium dimensional settings where the
-sparse assumption is not necessary. Here we elaborate the mixture normal
-prior as a demonstration of applying continuous spike-and-slab prior in
-BHAM.
-
-A spike-and-slab mixture normal spline prior can be expressed as
-\begin{align*}
-  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
-  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
-\end{align*} Similar to the spike-and-slab spline prior in Equation
-(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
-optimized via cross-validation. One of the critics received by the
-spike-and-slab mixture normal prior is that the tails of a normal
-distribution diminishes to zero too fast, which causes problems when
-estimating the large effects. Distributions with heavier tails can be
-used as an alternative, for example mixture Student's \(t\) distribution
-with small degree of freedom.
-
-\subsection{Algorithms for Fitting BHAMs}
-
-The proposed models can be fitted with MCMC algorithms. Nevertheless,
-the computational burden of MCMC algorithms creates scalability issues.
-George and McCulloch\citep{George1997} examined the computation speed
-for various MCMC algorithms with spike-and-slab mixture normal priors,
-and suggested MCMC algorithms works well for medium size (\(p\)=25) of
-predictors with only linear effects. However, it is not feasible for
-high-dimensional data analysis where the number of predictors easily
-exceeds 100. Specific to additive models, each predictor would expand to
-multiple new ``predictors'' via basis functions, creating greater
-computational demands. Scheipl et al. \citep{Scheipl2013} demonstrated
-the computational demands of a MCMC algorithm for fitting spike-and-slab
-GAM grow exponentially as \(p\) increases modestly via simulation
-studies. Hence, we feel compelled to develop scalable algorithms for
-fitting Bayesian hierarchical additive models in high-dimensional
-settings.
-
-As an alternative to sampling algorithms for fitting Bayesian models,
-optimization algorithms focus on the maximum a posteriori (MAP)
-estimates and speed up the model fitting process at the cost of
-posterior inference. The earlier work for fitting SS models using
-optimization algorithms includes EMVS \citep{Rockova2014a}. Rockova and
-George\citep{Rockova2014a} proposed an expectation-maximization (EM)
-based algorithm to fit models that use continuous mixture normal priors.
-In the E step, the latent binary indicators \(\gamma\)s are treated as
-the missing data, and the posterior means are calculated conditioning on
-the current value of other parameters; in the M step, a ridge estimator
-was used to update the coefficients, followed by updating
-\(\phi, \theta\). The same authors \citep{Rockova2018b, Rockova2018}
-further combined the EM algorithm with coordinate descent algorithm to
-fit SSL models. Yi and his group independently developed the
-EM-Iterative Weighted Least Square and EM-cyclic coordinate descent
-algorithms to fit models with broader class of priors, SSL
-included.\citep{Yi2019} These algorithms were implemented for
-generalized linear models\citep{Tang2017a}, Cox proportional hazards
-models \citep{Tang2017} and their grouped
-counterparts\citep{Tang2018, Tang2019}. Both EM based algorithms provide
-deterministic solutions, which becomes a popular property for
-reproducible research
-
-In this section, we extend the two EM-based algorithms, EM-CD and
-EM-IWLS algorithms, to fit BHAMs. To note, the two proposed algorithms
-provides different utilities. The EM-CD algorithm is specifically for
-fitting BHAM with an expedited performance, recommending to use in high
-and ultra-high dimensional setting. A specific concern of EM-CD
-algorithm is that it provides no information for inference. In contrast,
-the EM-IWLS can estimate the variance-covariance matrix of the
-coefficients. Moreover, the EM-IWLS is a more general model fitting
-algorithm that can be used for fitting not only SSL and continuous SS
-priors but also Student's t-priors and double exponential priors. To
-note, SB-GAM\citep{Bai2020, Bai2021} also used an EM-CD algorithm. The
-main difference between proposed EM-CD algorithm and that in SB-GAM is
-that SB-GAM uses a block CD algorithm for their group prior, while the
-proposed prior is pairwise independent requiring no special treatment in
-the CD algorithm.
-
-\subsubsection{EM algorithms}
-
-EM algorithm is an iterative algorithm to find MAP estimates or the
-maximum likelihood estimates. It is commonly used when some necessary
-data to establish the likelihood function are missing. Instead of
-maximizing the the likelihood function, the algorithm maximizes the
-expectation of the likelihood function with respect to the ``missing''
-data.
-
-The recursive algorithm consists of two steps:
-
-\begin{itemize}
-\tightlist
-\item
-  E-step: to calculate the expectation of the posterior density function
-  with respect to some ``missing'' data
-\item
-  M-step: to maximize the expectation derived in the E-step and update
-  parameters of interest
-\end{itemize}
+with the two-part spike-and-slab LASSO prior as the BHAM, and visually
+presented in Figure \ref{fig:SSprior}.
+
+\subsection{Scalable EM-Coordinate Descent Algorithm}
+
+\textcolor{red}{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
+Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
+Hence, we consider the optimization algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
+Specifically, we
+extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
+Similar to the EMVS algorithm \cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is based on the the expectation-maximization (EM) algorithm, integrating Coordinate Descent algorithm in each iterative step to find the posterior mode.
+The EM-CD algorithm has been well adapted in generalized linear models \cite{Tang2017a}, Cox proportional hazards models \cite{Tang2017}, and their grouped counterparts \cite{Tang2018, Tang2019}.
+The EM-CD algorithm provides deterministic solutions, which becomes a popular property for reproducible research.}
 
@@
 consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model, in other words the ``missing'' data.
-Our objective is to find the parameters \(\Theta\) that maximize the
-posterior density function, or equivalently, the logarithm of the
-density function, \[
+nuisance parameters of the model, in other words the ``missing'' data in
+the EM context. Our objective is to find the parameters \(\Theta\) that
+maximize the posterior density function, or equivalently the logarithm
+of the density function, \[
 \begin{aligned}
@@
 where \(f(\textbf{y}|\boldsymbol{\beta}, \phi)\) is the data
-distribution and \(f(\theta)\) is the Beta(1,1) density. We choose
+distribution and \(f(\theta)\) is the Beta(a, b) density. We choose
 non-informative prior for the intercept \(\beta_0\) and the dispersion
@@
 
-We use the EM algorithm to find the MAP estimate of \(\Theta\). This is,
-in the E-step, we calculate the expectation of posterior density
-function of
+We use the EM algorithm to find the maximum a posteriori estimate of
+\(\Theta\). This is, in the E-step, we calculate the expectation of
+posterior density function of
 \(\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\) with
 respect to the latent indicators \(\boldsymbol{\gamma}\) conditioning on
-the values from previous iteration \(\Theta^{(t-1)}\), \[
+the parameter values from previous iteration \(\Theta^{(t-1)}\), \[
 E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X}) .
@@
 \(E(\cdot)\equiv E_{\boldsymbol{\gamma}|\Theta^{(t-1)}}(\cdot)\). In the
-M-step, we find the \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The E-
-and M- steps are iterated until the algorithm converge.
+M-step, we find the parameters \(\Theta^{(t)}\) that maximize
+\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\).
+\textcolor{red}{The parenthesized subscription $^{(t)}$ denotes the parameter estimation at the $t$th iteration.}
+The E- and M- steps are iterated until the algorithm converges.
 
@@
 maximized separately for \(\boldsymbol{\beta}, \phi\) and
-\(\boldsymbol{\theta}\). Depending on the choice of coefficient priors,
-\(Q_1\) can be treated as penalized likelihood function and maximization
-of \(E(Q_1)\) can be solved via CD algorithm or IWLS algorithm in each
-iteration. Maximization of \(E(Q_2)\) can be solved via closed form
-equations following the beta-binomial conjugate relationship.
-
-\subsubsection{EM-Coordinate Descent}
-\label{sec:EMCD}
-
-When the prior distribution of the coefficients is set to mixture double
-exponential, coordinate descent algorithm can be used to estimate the
-parameters in the M-step. Coordinate descent is an optimization
+\(\boldsymbol{\theta}\). With the proposed two-part spike-and-slab LASSO
+prior, \(Q_1\) can be treated as penalized likelihood function and
+maximization of \(E(Q_1)\) can be solved via the Coordinate Descent
+algorithm in each iteration. Coordinate descent is an optimization
 algorithm that offers extreme computational advantage, and famous for
 its application in optimizing the \(l_1\) penalized likelihood function.
-
-The density function of spike-and-slab mixture double exponential prior
-can be written as \[
+Maximization of \(E(Q_2)\) can be solved via closed form equations
+following the beta-binomial conjugate relationship.
+
+The density function of the mixture double exponential prior of
+coefficient \(\beta\) can be written as \[
 f(\beta|\gamma, s_0, s_1) = \frac{1}{2\left[(1-\gamma)s_0 + \gamma s_1\right]}\exp(-\frac{|\beta|}{(1-\gamma)s_0 + \gamma s_1}),
@@
 E(Q_1) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) - \sum\limits_{j=1}^p\left[E({S_j}^{-1})|\beta_j|+\sum\limits_{k=1}^{K_j}E({S^{*}}^{-1}_{j})|\beta_{jk}|\right],
-\end{equation} where
-\(S_{j} = (1-\gamma^{0}_{j}) s_0 + \gamma^{0}_{j} s_1\) and
+\end{equation} where \(S_{j} = (1-\gamma_{j}) s_0 + \gamma_{j} s_1\) and
 \(S^*_{j} = (1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1\). To calculate two
@@
 posterior probability
-\(p_{j} \equiv \text{Pr}(\gamma^{0}_{j}=1|\Theta^{(t-1)})\) and
+\(p_{j} \equiv \text{Pr}(\gamma_{j}=1|\Theta^{(t-1)})\) and
 \(p_{j}^*\equiv \text{Pr}(\gamma^*_{j}=1|\Theta^{(t-1)})\) are
@@
 \(p_j^*\) is slightly different from that of \(p_j\), as \(p_j^*\)
-depends on the value of the vector \(\boldsymbol{\beta}^*_{j}\) and
+depends on the values of the vector \(\boldsymbol{\beta}^*_{j}\) and
 \(p_j\) only depends on the scalar \(\beta_j\). The calculation follows
-the equations below. \begin{align*}
+the equations below, \begin{align*}
 p_{j} &= \frac{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) }{\text{Pr}(\gamma_{j} = 1|\theta_j)f(\beta_{j}|\gamma_{j}=1, s_1) + \text{Pr}(\gamma_{j} = 0|\theta_j)f(\beta_{j}|\gamma_{j}=0, s_0)}\\
 p^*_{j} &= \frac{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) }{\text{Pr}(\gamma^{*}_{j} = 1|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=1, s_1) + \text{Pr}(\gamma^{*}_{j} = 0|\theta_j)\prod\limits_{k=1}^{K_j}f(\beta_{jk}|\gamma^{*}_{j}=0, s_0)}
-\end{align*} and \(\text{Pr}(\gamma_{j}^{0} = 1|\theta_j) = \theta_j\),
-\(\text{Pr}(\gamma_{j}^{0} = 0|\theta_j) = 1-\theta_j\),
+\end{align*} where \(\text{Pr}(\gamma_{j} = 1|\theta_j) = \theta_j\),
+\(\text{Pr}(\gamma_{j} = 0|\theta_j) = 1-\theta_j\),
 \(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
@@
 \(f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)\). It is trivial to
-show \begin{align}\label{eq:exp_scale}
+show \begin{align*}\label{eq:exp_scale}
 &E(\gamma_{j})  = p_{j} & &E(\gamma^{*}_{j}) = p_{j}^{*}\nonumber\\
-&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S^*}^{-1}_{j}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
-\end{align} After replacing withe calculated quantities, \(E(Q_1)\) can
+&E({S}^{-1}_{j}) = \frac{1-p_{j}}{s_0} + \frac{p_{j}}{s_1} & &E({S_{j}^*}^{-1}) = \frac{1-p_{j}^{*}}{s_0} + \frac{p_{j}^{*}}{s_1}.
+\end{align*} After replacing the calculated quantities, \(E(Q_1)\) can
 be seen as a \(l_1\) penalized likelihood function with the
@@
 
-Totally, the proposed EM-coordinate descent algorithm is summarized as
-follows:
+Totally, the proposed EM-CD algorithm is summarized as follows:
 
@@
 
-Our proposed models, BHAM, require two preset scale parameters (\(s_0\),
+Our proposed model, BHAM, requires two preset scale parameters (\(s_0\),
 \(s_1\)). Hence, we need to find the optimal values for the scale
@@
 regarding a criteria of preference. This would be achieved by
-constructing a two dimensional grid, consists of different pairs of
-(\(s_0\), \(s_1\)) value. However, previous research suggested the value
-of slab scale \(s_1\) have less impact on the final model and is
+constructing a two-dimensional grid, consists of different pairs of
+(\(s_0\), \(s_1\)) value. However, previous research suggests the value
+of slab scale \(s_1\) has less impact on the final model and is
 recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
 provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
-on examining different values of spike scale \(s_0\). Instead of the 2-D
-grid, We consider a sequence of \(L\) decreasing values
+on examining different values of spike scale \(s_0\). Instead of the
+two-dimensional grid, we consider a sequence of \(L\) decreasing values
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
@@
 model. A measure of preference calculated with cross-validations (CV),
-e.g.~deviance, area under the curve (AUC), mean squared error, can be
-used to facilitate the selection of a final model. The procedure is
-similar to the Lasso implemented in the widely used R package
-\texttt{glmnet}, which quickly fits Lasso models over a list of values
-of regularization parameters \(\lambda\) , giving a sequence of models
-for users to choose from.
+e.g.~deviance
+\textcolor{red}{(defined as model log-likelihood times -2, $-2\log f(\boldsymbol{y}|\boldsymbol{\hat\beta}, \hat\phi)$)},
+area under the curve (AUC), mean squared error, can be used to
+facilitate the selection of a final model. The procedure is similar to
+the LASSO implementation in the widely used R package \texttt{glmnet},
+which quickly fits LASSO models over a list of values of regularization
+parameters \(\lambda\) and gives a sequence of models for users to
+choose from.
 
@@
 
-In this section, we compare the performance of the proposed models to
-four alternative models: component selection and smoothing operator
-(COSSO) \citep{Zhang2006GAM}, adaptive COSSO \citep{Storlie2011},
-generalized additive models with automatic smoothing \citep{Wood2011},
-SB-GAM
-\citep{Bai2021}\textcolor{red}{, linear lasso model and spikeSlabGAM [TODO: add citation here].[TODO: explain about linear lasso model and spikeSlabGAM model]}.
+In this section, we compare the performance of the proposed model to
+\textcolor{red}{six alternative models: linear LASSO models,} component
+selection and smoothing operator (COSSO) \citep{Zhang2006GAM}, adaptive
+COSSO \citep{Storlie2011}, generalized additive models with automatic
+smoothing (referred as \textit{mgcv} hereafter)\citep{Wood2011},
+\textcolor{red}{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM
+\citep{Bai2021}.
+\textcolor{red}{We use linear LASSO model as the benchmark, examing the performance when linearity assumption doesn't hold}.
 COSSO is one of the earliest smoothing spline models that consider
@@
 adaptive weight for penalties such that the penalty of each functional
-component are different for extra flexibility. Generalized additive
-models with automatic smoothing, hereafter \textit{mgcv}, is one of the
-most popular models for nonlinear effect interpolation and prediction.
-\textcolor{red}{mgcv doesn't support analyses where the number of parameters exceeds the sample size.}
-SB-GAM is the first spike-and-slab lasso GAM. We implement COSSO and
-adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive
-models with automatic smoothing with R package \texttt{mgcv 1.8-31},
-SB-GAM with R package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do
-not provide flexibility to define smoothing functions, and hence use the
-default choices. Both mgcv and proposed models allow customized
+component are different for extra flexibility. \textit{mgcv} is one of
+the most popular models for nonlinear effect interpolation and
+prediction. Nevertheless,
+\textcolor{red}{mgcv doesn't support analyses when the number of parameters exceeds the sample size.}
+\textcolor{red}{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses a MCMC algorithm for model fitting.}
+SB-GAM is the first spike-and-slab LASSO GAM. We implement linear LASSO
+model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO
+with R package \texttt{cosso 2.1-1}, generalized additive models with
+automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab
+GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R
+package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
+flexibility to define smoothing functions, and hence use the default
+choices; mgcv, spikeSlabGAM and the proposed model allow customized
 smoothing functions and we choose the cubic regression spline. We
-controll the dimensionality of each smoothing function, 10 bases, for
-all different choices of smoothing functions. We use 5-fold CV with the
-default selection criteria to select the final model for COSSO models,
-SB-GAM and the proposed models. 20 default candidates of tuning
-parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are examined for
-SB-GAM and the proposed models which allow user-specification of tuning
-candidates. All computation was conducted on a high-performance 64-bit
-Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680
-processors and 24G of RAM per core and R 3.6.2 \citep{R}.
+control the dimensionality of each smoothing function, 10 bases, for all
+different choices of smoothing functions. We use 5-fold CV with the
+default selection criteria to select the final model for linear LASSO
+model, COSSO models, SB-GAM and the proposed model. 20 default
+candidates of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in
+SB-GAM) are examined for SB-GAM and the proposed model which allow
+user-specification of tuning candidates. All computation was conducted
+on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz
+eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
+3.6.2 \citep{R}.
 
@@
 methods of sparse additive models by Ravikumar et al.
-\citep{Ravikumar2009} and stochastic search term selection for GAM
-\citep{Scheipl2012}. However, we exclude these methods from current
-simulation study because of demonstrated inferior predictive performance
-compared to mgcv and scalability issues with increased number of
-predictors. \citep{Scheipl2013}
+\citep{Ravikumar2009}. However, we exclude these methods from the
+current simulation study because of their demonstrated inferior
+predictive performance compared to \textit{mgcv} \citep{Scheipl2013}.
 
@@
 
-We follow the data generating process described in Bai \citep{Bai2021}.
-We first generate \(n=500\) training data points with
+We follow the data generating process described in Bai \citep{Bai2021}:
+we first generate \(n=500\) training data points with
 \(p=4, 10, 50, 100, 200\) predictors respectively, where the predictors
-\(X\) are simulated from a multivariate normal distribution
-\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(y\)
+\(\boldsymbol{X}\) are simulated from a multivariate normal distribution
+\(\text{MVN}_{n\times p}(0, I_{P})\). We then simulate the outcome \(Y\)
 from two distributions, Gaussian and binomial with the identity link and
 logit link \(g(x) = \log(\frac{x}{1-x})\) respectively. The mean of each
-outcome were simulated via the following function \[
+outcome is derived via the following function \[
 \mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
@@
 1. In this data generating process, we have \(x_1, x_2, x_3, x_4\) as
-the active covariates, while the rest covariates are inactive,
+the active predictors, while the rest predictors are inactive,
 i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
@@
 50 independent pairs of training and testing datasets to evaluate the
-prediction performance of the chosen models, where training datasets are
-used to fit the models and testing datasets used to calculate assessment
-measures.
-\textcolor{red}{In addition, we considered the data generating process where the all the functional forms of the predictors are linear while keeping the rest of simulation parameters keep the same.
-These additional sets of simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
+prediction and variable selection performance of the chosen models,
+where training datasets are used to fit the models and testing datasets
+are used to calculate metrics of interest.
+\textcolor{red}{In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
+This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.}
 
@@
 \(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing dataset, are averaged across 50 simulations.
-\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged false discover rate (FDR), false negative rate(FNR), and Matthews correlation coefficient (MCC).
-For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for mgcv that relies on hypothesis testing for variable selection, and soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+on the testing datasets, are averaged across 50 simulations.
+\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
+\begin{align*}
+&\text{precision} = \frac{TP}{PP}\\
+&\text{recall} = \frac{TP}{TP+FP}\\
+&\text{MCC} = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
+\end{align*}
+where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively.
+For the methods that don't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
 For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
-\[
-\textcolor{red}{MCC = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},}
-\]
-\textcolor{red}{where TP, TN, FP, and FN are true positives, true negatives, false positives, and false negatives, respectively.}
 
@@
 
-\textcolor{red}{Among the main set of simulations where the functional forms of the predictors are mostly non-linear,}
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are mostly non-linear,}
 the predictive performances have a consistent pattern across the two
@@
 Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
-When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse [TODO: re-evaluate after tuning the models] in high-dimensional cases ($p=50,100, 200$).
+When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse in high-dimensional cases ($p=50,100, 200$).
 As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv still fit models with increasing in large dimension case ($p=50$).
-But the algorithm fails to converge within the default number of iterations, and hence leads to bad prediction performance.}
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to setimate ($p=50$).
+This leads to bad performance.}
 In high dimensional cases where we mimic the situation the signals are
-extremely sparse, SB-GAM has better performance than the proposed
-method. However, the BHAM-CD has extreme computational advantage over
-SB-GAM (see Table \ref{tab:time_sim}) without sacrificing much of the
-prediction accuracy.
+extremely sparse, SB-GAM has slight better performance than the proposed
+method{[}TODO: quantify{]}. However, the BHAM has extreme computational
+advantage over SB-GAM {[}TODO: quantify{]} without sacrificing much of
+the prediction accuracy (see Table \ref{tab:time_sim}) .
 
@@
 
-\textcolor{red}{\subsubsection{Variable Selection Performance}
-Within the simulation settings we examined, the proposed model tends to have a more conservative bebaviour in terms of variable selection, having high precision and low recall.
-(See Table xxx for gaussian outcomes, and support information xxx for binomial informaitons.) [TODO: mention the variable selection pattern is consistent across the types of outcomes.] THe high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
-Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high matthew correlation coefficient.
-Another Bayesian model, spike-and-slab performs bad in terms of variable selection in high-dimensional situations ($p=100,200$).
-We also want to highlight in our set of linear Gaussian simulations, the SB-GAM performs poorly (See Supproting table 1), while we didn't observe the same for linear binomial outcomes.
-}
-
-\textcolor{red}{\subsubsection{Bi-level Selection}
-Among the methods of comparison, there are two methods is possible for bi-level selection, i.e.
-the proposed BHAM and spikeSlabGAM.
-Hence, in this section, we further investigate the bi-level selection of the proposed models.
-}
+\textcolor{red}{\subsubsection{Variable Selection Performance}}
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are mostly non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table xxx for Gaussian outcomes, and support information xxx for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
+We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$ due to using a generic sequence of $s_0$ that doesn't contain the optimal value, discovered when analyzing the simulation result.
+Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high MCC.
+The performance of another Bayesian model, spikeSlabGAM deteriorates as sparsity grows, particularly when ($p>50$), or for binomial outcomes.
+The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.}
+
+\textcolor{red}{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
+Among the linear simulations, both methods can accurately select the linear components while having a drastically lowered probability to include the nonlinear component in the model.
+Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM.
+However, this advantage of spikeSlabGAM over BHAM is less obvious among the mixed function form simulations: spikeSlabGAM performs better than BHAM when selecting components when the function form includes only linear or nonlinear component, e.g.
+function form for $x_3$ and $x_4$.
+However, the spikeSlabGAM inclines to complete ignore the variable that have more complex function form, e.g.
+function form for $x_1$ and $x_2$, while BHAM are more likely to include them in the model.
+This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
+We defer an in-depth discussion to the Section \ref{sec:concl}.}
 
@@
 hierarchical additive model, SB-GAM \citep{Bai2021}, while being
-computationally efficient (see Table \ref{tab:time_real_data}).
+computationally efficient (see Table \ref{tab:time_real_data}). {[}TODO:
+add a sentenec how much faster.{]}
 
@@
 size (N=454). There are initially 6796 features in the dataset, which is
-too large to be practically meaningful to analyze. Hence, we perform a
-univariate screening procedure on the features, via GAM implemented in
-\texttt{mgcv}, and choose the the top 200 features with smallest
-p-values. We use 5-knot spline additive models for binary outcome using
-two different models, the proposed BHAM and the SB-GAM. 10-Fold CV are
-used to choose the optimal tuning parameters of each framework with
-respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where
-deviance, AUC, Brier score defined as
+too large to be practically meaningful to analyze. Hence, we choose the
+the top 200 features \textcolor{red}{with largest variance}. We use
+5-knot spline additive models for binary outcome using two different
+models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
+the optimal tuning parameters of each framework with respect to the
+default selection criterion implemented in the software. Out-of-bag
+samples are used for prediction performance evaluation, where deviance,
+AUC, Brier score defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
@@
 publicly available \citep{Bihlmeyer2021_data}. In this analysis, we
-primarily focuse on the analysis of one of the three studies included,
+primarily focus on the analysis of one of the three studies included,
 weight loss maintenance cohort \citep{Svetkey2008}, due to the
@@
 chromatography mass spectrometry. Quality control and natural log
-transformation are performed during metabolites data preparation. The
+transformation \textcolor{red}{were previously} performed and documented
+\textcolor{red}{by the study publishing team [@Bihlmeyer2021]}. The
 outcome of interest are standardized percent change in insulin
@@
 \(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
-BHAM-CD obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
+BHAM obtains superior \(R^2\), MSE, and MAE in the out-of-bag samples
 compared to SB-GAM (see Table \ref{tab:WLM_res}).
@@
 
-In the paper, we described a novel generalized additive model using
-Bayesian hierarchical priors, particularly the proposed spike-and-slab
-spline prior for bi-level functional selection. Meanwhile, we introduced
-two optimization based algorithms for model fitting. The algorithms can
-be easily scale up to address high-dimensional data analysis in a
-computational efficient manner. Via simulations, we demonstrated that
-the proposed model provides as good, if not better, prediction
-performance compared to some state-of-the-art non-linear modeling
-devices.
-
-The proposed model shares many commonality with an high-dimensional
-Bayesian GAM, SB-GAM \citep{Bai2021}, independently developed around the
-same time of this work. Both frameworks emphasize computational
-efficiency by deploying group spike-and-slab lasso type of priors and
-optimization-based fast and scalable algorithms. Bai provided the
-theoretical proof for the consistency of variable selection using group
-spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the
-bi-level functional selection and the model inference. The proposed
-model provides solutions to these remaining questions while maintaining
-the same level of prediction accuracy. Moreover, the proposed model
-renders additional flexibility to model specification, allowing various
-choices of smoothing functions and degrees of freedom.
-
-For translational science purpose, we implemented the proposed model in
-a R package \texttt{BHAM}, deposited at
-\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
-smoothing function specification, we deploy the same programming grammar
-as in the state-of-the-art package \texttt{mgcv}, in contrast to
-previous tools where smoothing functions are limited to the default
-ones. Ancillary functions are provided for model specification in
-high-dimensional settings, curve plotting and functional selection. In
-addition, in \texttt{BHAM}, we streamline the model fitting algorithms
-to support other popular Bayesian hierarchical prior for the smoothing
-functions, such as Student's T distribution, double exponential
-distribution, mixture normal and T distribution's. These priors could be
-helpful when the sparse assumption is weak or not necessary.,
-
-There are some improvements possible for the proposed models. First of
-all, the proposed model achieves a bi-level selection via the two-part
-spike-and-slab spline prior. Nevertheless, this set-up could result in a
-situation that is not theoretically sound: the non-linear component is
-selected, but the linear component is not. We currently address it
-analytically by including the linear component in the model when
-non-linear component is selected. Another possible solution is to impose
-a dependent structure of \(\gamma_{j}^*\) on \(\gamma_{j^{0}}\),
-i.e.~\(\gamma_j^*|\gamma_{j}^{0}, \theta_j\). Secondly, the
-computational time for fitting a BHAM model with EM-IWLS algorithms can
-be improved. Current implementation of the EM-IWLS algorithm jointly
-updates all coefficients in each iteration, which requires a lot
-computation resources. This jointly updating procedure can be enhanced
-by adapting a backfitting step \citep{Hastie1987} where each smoothing
-function are updated individually. Thirdly, when using the proposed
-model in real data analysis where a screening procedure is implemented
-before joint predictive modeling. we recommend to include the screening
-procedure in the cross-validation during model selection.
-\citep{Friedman2017} For the sake of model comparison, we fit the models
-using the same pool of predictors in Section \ref{sec:ECB}.
-
-Our future efforts direct to modeling survival outcomes and integrative
-analysis. While the proposed model addresses a great deal of analytic
-problem, analyzing the time-to-event outcome remains unsolved. An naive
-approach would be convert a time-to-event outcome to a Poisson outcome
-following Whitehead \citep{Whitehead1980}. However, it would be more
-efficient to directly fit Cox models via penalized pseudo likelihood
-function \citep{Simon2011}. Meanwhile, with growing understanding of
-biological structure within -omics field, it is appealing to integrate
-external biology information in the modeling process. The main
-motivation for integrative models is that biologically informed grouping
-of weak effects increases the power of detecting true associations
-between features and the outcome \citep{Peterson2016}, and stabilizes
-the analysis results for reproducibility purpose. Such integration can
-be achieved by setting up a structural hyperprior on the inclusion
+\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
+The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
+Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
+In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear component.
+In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
+When serving the purpose of variable selection, trade-offs exist among methods of comparison.}
+We implement the proposed model in an open-source R package
+\texttt{BHAM}, deposited at \url{https://github.com/boyiguo1/BHAM}. To
+maximize the flexibility of smoothing function specification, we deploy
+the same programming grammar as in the state-of-the-art package
+\texttt{mgcv}, in contrast to previous tools where smoothing functions
+are limited to the default ones. Ancillary functions are provided for
+model specification in high-dimensional settings, curve plotting and
+functional selection.
+
+The proposed model shares many commonality with the SB-GAM
+\citep{Bai2021}, independently developed around the same time of the
+proposed work. Both frameworks emphasize computational efficiency by
+deploying group spike-and-slab LASSO type priors and optimization-based
+scalable algorithms. Bai provides the theoretical proof for the
+consistency of variable selection using group spike-and-slab LASSO
+prior.
+\textcolor{red}{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
+Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.
+Not focused in this manuscript, the generalization is described in the supporting materials.}
+
+\textcolor{red}{During designing and analyzing the simulation study, we made couple interesting observations.
+First of all, variable selection is a delicate topic in the context predictive modelling.
+When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example LASSO and LASSO-based models.
+\cite{Wu2019} Moreover, bi-level selection is a more complex problem than variable selection.
+The complexity reflects on the validity of the effect hierarchy principle.
+While most functional forms follow that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
+$x^2$.
+The proposed prior and spikeSlabGAM employ different structure: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
+The different prior setups lead to trade-offs for the purpose of bi-level selection.
+We recommend to use more judgement in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
+Secondly, we find the performance of the proposed model are more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
+Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend to fine-tune the model with granular sequence of $s_0$ for performance improvement.}
+
+Our future efforts direct to survival analysis and integrative analysis.
+While the proposed model addresses a great deal of analytic problem,
+analyzing the time-to-event outcome remains unsolved. An naive approach
+would be convert a time-to-event outcome to a Poisson outcome following
+Whitehead \citep{Whitehead1980}. However, it would be more efficient to
+directly fit Cox models via penalized pseudo likelihood function
+\citep{Simon2011}. Meanwhile, with growing understanding of biological
+structure within -omics field, it is appealing to integrate external
+biology information in the modeling process. The main motivation for
+integrative models is that biologically informed grouping of weak
+effects increases the power of detecting true associations between
+features and the outcome \citep{Peterson2016}, and stabilizes the
+analysis results for reproducibility purpose. Such integration can be
+achieved by setting up a structural hyperprior on the inclusion
 indicator of the smoothing function null space
@@
 
+\clearpage
+\begin{figure}[h]
+\includegraphics{Figs/ECB_plot}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[h]
+\includegraphics{Figs/WLM_plot}
+\end{figure}
+
 \end{document}

### 2022-03-21 Manuscript/03-Simulation.Rmd
Revise simulaiton results

@@
 
-In this section, we compare the performance of the proposed model to \bg{six alternative models: linear Lasso models, component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM [TODO: add citation here]}}, and SB-GAM [@Bai2021].
-\bg{We use linear lasso model as a benchmark, examing the performance when linearity assumption doesn't hold}.
+In this section, we compare the performance of the proposed model to \bg{six alternative models: linear LASSO models,} component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM [@Bai2021].
+\bg{We use linear LASSO model as the benchmark, examing the performance when linearity assumption doesn't hold}.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
@@
 \textit{mgcv} is one of the most popular models for nonlinear effect interpolation and prediction.
-Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab lasso GAM.
-We implement linear Lasso model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
+Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses a MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab LASSO GAM.
+We implement linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
 COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices; mgcv, spikeSlabGAM and the proposed model allow customized smoothing functions and we choose the cubic regression spline.
 We control the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
-We use 5-fold CV with the default selection criteria to select the final model for linear Lasso model, COSSO models, SB-GAM and the proposed model.
+We use 5-fold CV with the default selection criteria to select the final model for linear LASSO model, COSSO models, SB-GAM and the proposed model.
 20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed model which allow user-specification of tuning candidates.
@@
 [@Ravikumar2009].
-However, we exclude these methods from current simulation study because of demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].
+However, we exclude these methods from the current simulation study because of their demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].
 
 \subsection{Monte Carlo Simulation Study}
-We follow the data generating process described in Bai [@Bai2021].
-We first generate $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
+We follow the data generating process described in Bai [@Bai2021]: we first generate $n=500$ training data points with $p=4, 10, 50, 100, 200$ predictors respectively, where the predictors $\bs X$ are simulated from a multivariate normal distribution $\text{MVN}_{n\times p}(0, I_{P})$.
 We then simulate the outcome $Y$ from two distributions, Gaussian and binomial with the identity link and logit link $g(x) = \log(\frac{x}{1-x})$ respectively.
-The mean of each outcome were simulated via the following function
+The mean of each outcome is derived via the following function
 $$
@@
 Gaussian outcomes requires specification of dispersion, where we set the dispersion parameter to be 1.
-In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active covariates, while the rest covariates are inactive, i.e.
+In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active predictors, while the rest predictors are inactive, i.e.
 $f_j(x_j) = 0$ for $j = 4, \dots, p$.
@@
 We generate 50 independent pairs of training and testing datasets to evaluate the prediction and variable selection performance of the chosen models, where training datasets are used to fit the models and testing datasets are used to calculate metrics of interest.
-\bg{In addition, we consider the data generating process where all the functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
-These additional sets of linear simulations are designed to investigate the flexibility of the proposed model when non-linear assumptions are not met.}
+\bg{In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
+This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.}
 
 To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing datasets, are averaged across 50 simulations.
-\bg{To evaluate the variable selection performance of the models, we record the set of variables each model selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC)
+\bg{To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
 \begin{align*}
@@
 \end{align*}
-where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively.} For method that doesn't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} which relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.[TODO: verify]}
+where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively.
+For the methods that don't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
 
@@
 \subsubsection{\bg{Prediction Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are mostly non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
-(See Tables \ref{tab:gaus} and \ref{tab:bin_auc}) \bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional spline models.
-Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
-When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to setimate ($p=50$).
-This leads to bad performance.} In high dimensional cases where we mimic the situation the signals are extremely sparse, SB-GAM has slight better performance than the proposed method[TODO: quantify].
-However, the BHAM has extreme computational advantage over SB-GAM [TODO: quantify] without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}) .
+\bg{Among the set of simulations where the functional forms of the predictors are non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
+For simplicity, we use Gaussian simulations to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus} and \ref{tab:bin_auc} for detailed statistics.
+\bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
+Specifically, BHAM shows greate improvement over COSSO methods, resulting a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) imporvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
+The improvement over spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
+When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
+As previously hypothesized, the linear lasso model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
 
-\bg{We also examine the prediction performance when the functional form of predictors are linear.
-(See support information Table 1, and 2 [TODO: add hyperlink to the github repo]) The proposed model, BHAM, has similar performance as the linear lasso model regardless of the type outcome.
+\bg{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
+The proposed model, BHAM, has similar performance as the linear lasso model regardless of the distribution.
 This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
-spikeSlabGAM have similar prediction performance to BHAM (numeralize the advantage).
-Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.
-(numeralize the advantage)}
+spikeSlabGAM have similar prediction performance to BHAM.
+Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
 
 \bg{\subsubsection{Variable Selection Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are mostly non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table xxx for Gaussian outcomes, and support information xxx for binomial outcomes): being conservative.
-The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+\bg{Among the set of simulations where the functional forms of the predictors are non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
 In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
-We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$ due to using a generic sequence of $s_0$ that doesn't contain the optimal value, discovered when analyzing the simulation result.
-Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high MCC.
-The performance of another Bayesian model, spikeSlabGAM deteriorates as sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.}
+We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$.
+Upon further investigation, we discover it's because the generic sequence of $s_0$ used to tune the model doesn't contain the optimal value.
+Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yield a high MCC.
+The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes.
+The variable selection performance for linear simulations match with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
 
 \bg{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
-Among the linear simulations, both methods can accurately select the linear components while having a drastically lowered probability to include the nonlinear component in the model.
+Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
 Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM.
-However, this advantage of spikeSlabGAM over BHAM is less obvious among the mixed function form simulations: spikeSlabGAM performs better than BHAM when selecting components when the function form includes only linear or nonlinear component, e.g.
-function form for $x_3$ and $x_4$.
-However, the spikeSlabGAM inclines to complete ignore the variable that have more complex function form, e.g.
-function form for $x_1$ and $x_2$, while BHAM are more likely to include them in the model.
+However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the function forms that include only linear or nonlinear component, e.g.
+function forms for $x_3$ and $x_4$.
+However, spikeSlabGAM inclines to ignore the variable that have more complex function forms, e.g.
+function forms for $x_1$ and $x_2$.
+In contrast, BHAM is more likely to include them in the model.
 This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.

### 2022-03-21 Manuscript/SS_GAM.tex
Revise simulaiton results

@@
 
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are mostly non-linear,}
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are non-linear,}
 the predictive performances have a consistent pattern across the two
-distributions of outcomes. (See Tables \ref{tab:gaus} and
-\ref{tab:bin_auc})
-\textcolor{red}{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional spline models.
-Specifically, BHAM shows greate improvement over COSSO methods [TODO: add statsitics] and the spikeSlabGAM model [TODO: add statsitics].
-When comparing to SB-GAM, BHAM performs better in lower dimensional cases ($p=4,10$), and equally good or slightly worse in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear lasso model predicts less accurate compared to other flexibly models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to setimate ($p=50$).
-This leads to bad performance.}
-In high dimensional cases where we mimic the situation the signals are
-extremely sparse, SB-GAM has slight better performance than the proposed
-method{[}TODO: quantify{]}. However, the BHAM has extreme computational
-advantage over SB-GAM {[}TODO: quantify{]} without sacrificing much of
-the prediction accuracy (see Table \ref{tab:time_sim}) .
+distributions of outcomes. For simplicity, we use Gaussian simulations
+to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus}
+and \ref{tab:bin_auc} for detailed statistics.
+\textcolor{red}{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
+Specifically, BHAM shows greate improvement over COSSO methods, resulting a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) imporvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
+The improvement over spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
+When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
+As previously hypothesized, the linear lasso model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.}
+Even though SB-GAM has slight prediction advantage over the proposed
+model in high-dimensional situations, the BHAM has extreme computational
+advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in
+computation time (measured in seconds) for Gaussian simulations, without
+sacrificing much of the prediction accuracy (see Table
+\ref{tab:time_sim}).
 
-\textcolor{red}{We also examine the prediction performance when the functional form of predictors are linear.
-(See support information Table 1, and 2 [TODO: add hyperlink to the github repo]) The proposed model, BHAM, has similar performance as the linear lasso model regardless of the type outcome.
+\textcolor{red}{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
+The proposed model, BHAM, has similar performance as the linear lasso model regardless of the distribution.
 This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
-spikeSlabGAM have similar prediction performance to BHAM (numeralize the advantage).
-Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.
-(numeralize the advantage)}
+spikeSlabGAM have similar prediction performance to BHAM.
+Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
 
 \textcolor{red}{\subsubsection{Variable Selection Performance}}
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are mostly non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table xxx for Gaussian outcomes, and support information xxx for binomial outcomes): being conservative.
-The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the positive predictions.
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
 In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
@@
 \clearpage
+\input{Tabs/sim_time_tab.tex}
+\clearpage
 \input{Tabs/sim_gaus_var_slct_tab.tex}

### 2022-03-21 R/Sim_Res/compare_methods.R
Add sections to compare performance of methods

@@ -1,3 +1,7 @@
+# Example:
+#
+
+
 compare_methods <- function(dat, method_1, method_2, measures){
   # browser()
   if(!all(c(method_1, method_2) %in% unique(dat$method)))

### 2022-03-21 _targets.R

@@ -29,6 +29,7 @@ tar_plan(
   sim_success_rate = create_success_rate_table(sim_main_path),
   sim_mdl_fail = create_mdl_fail_rate(sim_success_rate),
   sim_pred_raw = create_raw_data(sim_success_rate),
+  sim_time_raw = create_raw_data(sim_success_rate, section = "time"),
   
   #** Binomial Outcome  ####
   #*** Prediction  ####
@@ -111,8 +112,8 @@ tar_plan(
     cat(file = "Manuscript/Tabs/sim_gaus_tab.tex"),
   
   # Percentage of change
-  sim_gaus_pred_bham_lasso = sim_pred_raw %>% filter(dist == "gaussian") %>%
-    compare_methods(method_1 = "bamlasso", method_2 = "lasso", measures="R2") %>% 
+  sim_gaus_time_bham_lasso = sim_pred_raw %>% filter(dist == "gaussian", !(p %in% c(4,10))) %>%
+    compare_methods(method_1 = "bamlasso", method_2 = "SB_GAM", measures="R2") %>% 
     with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   
   
@@ -141,6 +142,16 @@ tar_plan(
     format_var_slct_tbls(caption = "", label = "")%>% 
     cat(file = "Manuscript/Tabs/sim_gaus_var_slct_tab.tex"),
   
+  #*** Time  ####
+  sim_gaus_pred_bham_lasso = sim_time_raw %>% filter(dist == "gaussian", ) %>%
+    select(-c(user.self, sys.self, user.child, sys.child)) %>% 
+    pivot_wider(names_from = method, values_from = elapsed) %>% 
+    mutate(bamlasso = bamlasso_cv + bamlasso_final,
+           SB_GAM = bai_cv + bai_final) %>% 
+    select(it, dist, p, bamlasso, SB_GAM) %>% 
+    pivot_longer(cols = c(bamlasso, SB_GAM), names_to = "method", values_to = "time") %>% 
+    compare_methods(method_1 = "bamlasso", method_2 = "SB_GAM", measures="time") %>% 
+    with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   
   
   

### 2022-03-21 _targets.R
Add sections to compare performance of methods

@@ -29,6 +29,7 @@ tar_plan(
   sim_success_rate = create_success_rate_table(sim_main_path),
   sim_mdl_fail = create_mdl_fail_rate(sim_success_rate),
   sim_pred_raw = create_raw_data(sim_success_rate),
+  sim_time_raw = create_raw_data(sim_success_rate, section = "time"),
   
   #** Binomial Outcome  ####
   #*** Prediction  ####
@@ -111,8 +112,8 @@ tar_plan(
     cat(file = "Manuscript/Tabs/sim_gaus_tab.tex"),
   
   # Percentage of change
-  sim_gaus_pred_bham_lasso = sim_pred_raw %>% filter(dist == "gaussian") %>%
-    compare_methods(method_1 = "bamlasso", method_2 = "lasso", measures="R2") %>% 
+  sim_gaus_pred_bham_lasso = sim_pred_raw %>% filter(dist == "gaussian", !(p %in% c(4,10))) %>%
+    compare_methods(method_1 = "bamlasso", method_2 = "SB_GAM", measures="R2") %>% 
     with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   
   
@@ -141,6 +142,16 @@ tar_plan(
     format_var_slct_tbls(caption = "", label = "")%>% 
     cat(file = "Manuscript/Tabs/sim_gaus_var_slct_tab.tex"),
   
+  #*** Time  ####
+  sim_gaus_pred_bham_lasso = sim_time_raw %>% filter(dist == "gaussian", ) %>%
+    select(-c(user.self, sys.self, user.child, sys.child)) %>% 
+    pivot_wider(names_from = method, values_from = elapsed) %>% 
+    mutate(bamlasso = bamlasso_cv + bamlasso_final,
+           SB_GAM = bai_cv + bai_final) %>% 
+    select(it, dist, p, bamlasso, SB_GAM) %>% 
+    pivot_longer(cols = c(bamlasso, SB_GAM), names_to = "method", values_to = "time") %>% 
+    compare_methods(method_1 = "bamlasso", method_2 = "SB_GAM", measures="time") %>% 
+    with(sprintf("%.0f%% (%.0f%%)", median*100, IQR*100)),
   
   
   

### 2022-03-21 Manuscript/00-main.Rmd
Revise simulaiton results
Revise real data analysis

@@
 tar_load(manu_tbs)
+tar_load(manu_figs)
 
 tar_load(WLM_train_dat)
-# tar_load(tbl_one)
 tar_load(ECB_bamlasso_cv)
@@
 tar_load(WLM_plot)
-# tar_load(sim_tim_tab)
 ```
@@
 \input{Tabs/sim_binom_tab.tex}
+\clearpage
+\input{Tabs/sim_time_tab.tex}
 \clearpage
@@
 \includegraphics{Figs/ECB_plot}
+\caption{TODO: change this caption}
+\label{fig:ECB_fig}
 \end{figure}

### 2022-03-21 Manuscript/04-Real_Data.Rmd
Revise real data analysis

@@
 
-In this section, we apply the proposed models BHAM, fitted with the EM-CD algorithm, to analyze two real-world metabolomics datasets where the outcomes are binary and continuous respectively.
+In this section, we apply the proposed models BHAM to analyze two published metabolomics datasets where the outcomes are binary and continuous respectively.
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}).
-[TODO: add a sentenec how much faster.]
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
+\bg{BHAM requires roughly 10\% of the computation time of SB-GAM to fit models.}
 
@@
 10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassication error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
-BHAM-CD obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).
-
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
+Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassification error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
+BHAM obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).
+\bg{We plot the 33 features included in the final BHAM model in Figure \ref{fig:ECB_fig}.}
 
@@
 In this analysis, we primarily focus on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects.
-In the dataset, 765 metaboliltes in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
-Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team [@Bihlmeyer2021]}.
+In the dataset, 765 metabolites in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
+Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team \cite{Bihlmeyer2021}}.
 The outcome of interest are standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
@@
 BHAM obtains superior $R^2$, MSE, and MAE in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:WLM_res}).
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}

### 2022-03-21 Manuscript/SS_GAM.tex
Revise real data analysis

@@
 In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
-We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$ due to using a generic sequence of $s_0$ that doesn't contain the optimal value, discovered when analyzing the simulation result.
-Overall, among all the models examined, SB-GAM have the best performance, both high precision and high recall, and yield a high MCC.
-The performance of another Bayesian model, spikeSlabGAM deteriorates as sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations match with prediction performance: BHAM performs great with the Gaussian outcome, while SB-GAM's performance deteriorates.}
+We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$.
+Upon further investigation, we discover it's because the generic sequence of $s_0$ used to tune the model doesn't contain the optimal value.
+Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yield a high MCC.
+The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes.
+The variable selection performance for linear simulations match with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
 
 \textcolor{red}{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
-Among the linear simulations, both methods can accurately select the linear components while having a drastically lowered probability to include the nonlinear component in the model.
+Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
 Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM.
-However, this advantage of spikeSlabGAM over BHAM is less obvious among the mixed function form simulations: spikeSlabGAM performs better than BHAM when selecting components when the function form includes only linear or nonlinear component, e.g.
-function form for $x_3$ and $x_4$.
-However, the spikeSlabGAM inclines to complete ignore the variable that have more complex function form, e.g.
-function form for $x_1$ and $x_2$, while BHAM are more likely to include them in the model.
+However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the function forms that include only linear or nonlinear component, e.g.
+function forms for $x_3$ and $x_4$.
+However, spikeSlabGAM inclines to ignore the variable that have more complex function forms, e.g.
+function forms for $x_1$ and $x_2$.
+In contrast, BHAM is more likely to include them in the model.
 This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
@@
 
-In this section, we apply the proposed models BHAM, fitted with the
-EM-CD algorithm, to analyze two real-world metabolomics datasets where
-the outcomes are binary and continuous respectively. We demonstrate the
-improved prediction performance compared to the other Bayesian
-hierarchical additive model, SB-GAM \citep{Bai2021}, while being
-computationally efficient (see Table \ref{tab:time_real_data}). {[}TODO:
-add a sentenec how much faster.{]}
-
-\begin{center}
-Table \ref{tab:time_real_data} here
-\end{center}
+In this section, we apply the proposed models BHAM to analyze two
+published metabolomics datasets where the outcomes are binary and
+continuous respectively. We demonstrate the improved prediction
+performance compared to the other Bayesian hierarchical additive model,
+SB-GAM \citep{Bai2021}, while being computationally efficient (see Table
+\ref{tab:time_real_data}).
+\textcolor{red}{BHAM requires roughly 10\% of the computation time of SB-GAM to fit models.}
 
@@
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassication error defined as
+misclassification error defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
-calculated. BHAM-CD obtains superior AUC, Brier score, and
+calculated. BHAM obtains superior AUC, Brier score, and
 misclassification error in the out-of-bag samples compared to SB-GAM
 (see Table \ref{tab:ECB_res}).
-
-\begin{center}
-Table \ref{tab:ECB_res} here
-\end{center}
+\textcolor{red}{We plot the 33 features included in the final BHAM model in Figure \ref{fig:ECB_fig}.}
 
@@
 drastically different intervention effects. In the dataset, 765
-metaboliltes in baseline plasma collected were profiled using liquid
+metabolites in baseline plasma collected were profiled using liquid
 chromatography mass spectrometry. Quality control and natural log
 transformation \textcolor{red}{were previously} performed and documented
-\textcolor{red}{by the study publishing team [@Bihlmeyer2021]}. The
+\textcolor{red}{by the study publishing team \cite{Bihlmeyer2021}}. The
 outcome of interest are standardized percent change in insulin
@@
 compared to SB-GAM (see Table \ref{tab:WLM_res}).
-
-\begin{center}
-Table \ref{tab:WLM_res} here
-\end{center}
 

### 2022-03-21 _targets.R

@@ -505,6 +505,12 @@ tar_plan(
     list.files("Manuscript/Tabs/", full.names = TRUE)
   ),
   
+  tar_files(
+    manu_figs,
+    list.files("Manuscript/Figs/", full.names = TRUE)
+  ),
+  
+  
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   

### 2022-03-21 Manuscript/01-Intro.Rmd
Correct LASSO spelling

@@
 Ravikumar et al.
-[@Ravikumar2009] extended the grouped lasso [@Yuan2006] to additive models (AMs); Huang et al.
-[@Huang2010] further developed adaptive grouped lasso for additive models; Wang et al.
+[@Ravikumar2009] extended the grouped LASSO [@Yuan2006] to additive models (AMs); Huang et al.
+[@Huang2010] further developed adaptive grouped LASSO for additive models; Wang et al.
 [@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
@@
 Bai et al.
-[@Bai2020] was the first to apply group spike-and-slab lasso prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs [@Bai2021].
+[@Bai2020] was the first to apply group spike-and-slab LASSO prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs [@Bai2021].
 Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large number of knots are used.
@@
 the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in three ways.
-First of all, the proposed spike-and-slab spline prior is a spike-and-slab lasso type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
-Spike-and-slab lasso priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms.
+First of all, the proposed spike-and-slab spline prior is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
+Spike-and-slab LASSO priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms.
 They make fitting high-dimensional models more feasible without sacrificing performance in prediction and variable selection.
-Secondly, SB-GAM uses a group spike-and-slab lasso prior with an EM-CD algorithm to fit the model.
+Secondly, SB-GAM uses a group spike-and-slab LASSO prior with an EM-CD algorithm to fit the model.
 While both methods use the combination of expectation maximization algorithm and coordinate descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.

### 2022-03-21 Manuscript/02-Method.Rmd
Correct LASSO spelling

@@
 <!-- Particularly, the development of SSL model substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019].
-We notice that Bai[TODO: add Bai arxiv paper] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [TODO: add Xu Ghosh (2015)] and limits to an “all-in-all-out” strategy for functional selection.
+We notice that Bai[TODO: add Bai arxiv paper] is the first to apply spike-and-slab LASSO prior in the GAM framework, where the densities of the spike and slab components take the group LASSO density [TODO: add Xu Ghosh (2015)] and limits to an “all-in-all-out” strategy for functional selection.
 -->

### 2022-03-21 Manuscript/03-Simulation.Rmd
Correct LASSO spelling

@@
 When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear lasso model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+As previously hypothesized, the linear LASSO model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
 To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
@@
 \bg{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
-The proposed model, BHAM, has similar performance as the linear lasso model regardless of the distribution.
+The proposed model, BHAM, has similar performance as the linear LASSO model regardless of the distribution.
 This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.

### 2022-03-21 Manuscript/SS_GAM.tex
Correct LASSO spelling

@@
 functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
-lasso \citep{Yuan2006} to additive models (AMs); Huang et al.
-\citep{Huang2010} further developed adaptive grouped lasso for additive
+LASSO \citep{Yuan2006} to additive models (AMs); Huang et al.
+\citep{Huang2010} further developed adaptive grouped LASSO for additive
 models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
@@
 on AMs are treated as a special case. Bai et al. \citep{Bai2020} was the
-first to apply group spike-and-slab lasso prior to Gaussian AMs using a
+first to apply group spike-and-slab LASSO prior to Gaussian AMs using a
 fast optimization algorithm, and further generalized the framework to
@@
 \citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
-spline prior is a spike-and-slab lasso type prior using independent
+spline prior is a spike-and-slab LASSO type prior using independent
 mixture double exponential distribution, compared to spike-and-slab GAM
-that uses normal-mixture-of-inverse gamma prior. Spike-and-slab lasso
+that uses normal-mixture-of-inverse gamma prior. Spike-and-slab LASSO
 priors provide computational convenience during model fitting by using
@@
 performance in prediction and variable selection. Secondly, SB-GAM uses
-a group spike-and-slab lasso prior with an EM-CD algorithm to fit the
+a group spike-and-slab LASSO prior with an EM-CD algorithm to fit the
 model. While both methods use the combination of expectation
@@
 When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear lasso model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+As previously hypothesized, the linear LASSO model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
 To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.}
@@
 \textcolor{red}{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
-The proposed model, BHAM, has similar performance as the linear lasso model regardless of the distribution.
+The proposed model, BHAM, has similar performance as the linear LASSO model regardless of the distribution.
 This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.

### 2022-03-21 Manuscript/02-Method.Rmd
Correct nonlinear spelling

@@
 That is, $\bs U$ contains the eigenvectors $U^0$ with zero eigenvalues for the linear space and $\bs U^{\tp}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e.
-the non-linear space.
+the nonlinear space.
 We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix ${\bs X}^\repa = \bs X \bs U \equiv \begin{bmatrix} X^0 : \bs X^{\tp} \end{bmatrix}$.
@@
 \end{equation}
-where the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of non-linear space.
+where the coefficients $\bs \beta_j \equiv \begin{bmatrix} \beta_j : \bs \beta^\tp_j \end{bmatrix}$ is an augmentation of the coefficient scalar $\beta_j$ of linear space and the coefficient vector $\bs \beta^\tp_j$ of nonlinear space.
 
@@
 
-<!-- This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and non-linear space of the smoothing function.
+<!-- This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and nonlinear space of the smoothing function.
 The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the functional selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.

### 2022-03-21 Manuscript/03-Simulation.Rmd
Correct nonlinear spelling

@@
 \subsubsection{\bg{Prediction Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are non-linear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
+\bg{Among the set of simulations where the functional forms of the predictors are nonlinear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
 For simplicity, we use Gaussian simulations to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus} and \ref{tab:bin_auc} for detailed statistics.
@@
 \bg{\subsubsection{Variable Selection Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
 The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
@@
 Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
-Specifically, spikeSlabGAM have smaller probability to include non-linear component in the model than BHAM.
+Specifically, spikeSlabGAM have smaller probability to include nonlinear component in the model than BHAM.
 However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the function forms that include only linear or nonlinear component, e.g.

### 2022-03-21 Manuscript/05-Discussion.Rmd
Correct nonlinear spelling

@@
 First of all, the proposed model achieves a bi-level selection via the two-part spike-and-slab spline prior.
-Nevertheless, this set-up could result in a situation that is not theoretically sound: the non-linear component is selected, but the linear component is not.
-We currently address it analytically by including the linear component in the model when non-linear component is selected.
+Nevertheless, this set-up could result in a situation that is not theoretically sound: the nonlinear component is selected, but the linear component is not.
+We currently address it analytically by including the linear component in the model when nonlinear component is selected.
 Another possible solution is to impose a dependent structure of $\gamma_{j}^\tp$ on $\gamma_{j^{0}}$, i.e.

### 2022-03-21 Manuscript/10-Supplementary.Rmd
Correct nonlinear spelling

@@
 Since $\bs \tau^2$ are not of our primary interest, we treat them as the "missing" data in addition to the latent indicators $\bs \gamma$, and hence construct the expectation $E_{\bs \gamma, \bs \tau^2|\Theta^{(t-1)}}(Q_1)$ in the E-step.
-To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the non-linear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
+To note, unlike the same latent indicator $\gamma^\tp_j$ which is shared by the coefficients of the nonlinear terms $\beta^\tp_{jk}$ for $k = 1, \dots, K_j$ , $\tau^2_{jk}$ is coefficient specific for $\beta^\tp_{jk}$.
 $E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.

### 2022-03-21 Manuscript/SS_GAM.tex
Correct nonlinear spelling

@@
 \(\boldsymbol{U}^{*}\) contains the eigenvectors (as columns) for the
-non-zero eigenvalues, i.e.~the non-linear space. We multiply the basis
+non-zero eigenvalues, i.e.~the nonlinear space. We multiply the basis
 function matrix \(\boldsymbol{X}\) with the orthonormal matrix
@@
 is an augmentation of the coefficient scalar \(\beta_j\) of linear space
-and the coefficient vector \(\boldsymbol{\beta}^*_j\) of non-linear
+and the coefficient vector \(\boldsymbol{\beta}^*_j\) of nonlinear
 space.
@@
 
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are non-linear,}
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear,}
 the predictive performances have a consistent pattern across the two
@@
 \textcolor{red}{\subsubsection{Variable Selection Performance}}
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are non-linear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
 The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.

### 2022-03-21 Manuscript/00-main.Rmd
Revise and add caption to the plot

@@
 <!-- \begin{landscape} -->
+
 ```{r sim_time, results = "asis", eval = F}
@@
 
-```{r}
+```{r }
 tbl_real_time %>% 
   flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\\label{tab:time_real_data}")
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.")
 ```
+\label{tab:time_real_data}
 
@@
 rbind(
-  'BHAM-CD' = ECB_bamlasso_cv %>% slice_max(auc) %>% select(-s0),
+  'BHAM' = ECB_bamlasso_cv %>% slice_max(auc)  %>% slice_min(s0) %>% select(-s0),
   'SB-GAM' = ECB_SBGAM_cv %>%  slice_max(auc) %>% select(-lambda)
@@
   xtable(
-    caption = "Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
+    caption = "Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
     label = "tab:ECB_res"
@@
 rbind(
-  'BHAM-CD' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
+  'BHAM' = WLM_bamlasso_cv %>% slice_min(mse) %>% select(-s0),
   'SB-GAM' = WLM_SBGAM_cv %>%  slice_min(mse) %>% select(-lambda)
@@
     xtable(
-    caption = "Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
+    caption = "Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
     label = "tab:WLM_res"
@@
 \includegraphics{Figs/ECB_plot}
-\caption{TODO: change this caption}
+\caption{Plots of the functions for the 33 metablites selected by BHAM in the Emory Cardiovascular Biobank data analysis}
 \label{fig:ECB_fig}
@@
 
-\clearpage
+<!-- \clearpage -->
 
-\begin{figure}[h]
-\includegraphics{Figs/WLM_plot}
-\end{figure}
+<!-- \begin{figure}[h] -->
+<!-- \includegraphics{Figs/WLM_plot} -->
+<!-- \end{figure} -->

### 2022-03-21 Manuscript/SS_GAM.tex
Revise and add caption to the plot

@@
 \caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.\textbackslash{}label\{tab:time\_real\_data\}
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.
 }\\
@@
 
+\label{tab:time_real_data}
+
 \clearpage
@@
   \hline
-BHAM-CD.1 & 510.99 & 0.61 & 0.19 & 0.24 \\ 
-  BHAM-CD.2 & 510.99 & 0.61 & 0.19 & 0.24 \\ 
+BHAM & 510.99 & 0.61 & 0.19 & 0.24 \\ 
   SB-GAM & 636.56 & 0.56 & 0.22 & 0.30 \\ 
@@
 \end{tabular}
-\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
+\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
 \label{tab:ECB_res}
@@
   \hline
-BHAM-CD & 668.01 & 0.07 & 0.93 & 0.76 \\ 
+BHAM & 668.01 & 0.07 & 0.93 & 0.76 \\ 
   SB-GAM & 666.83 & 0.03 & 0.98 & 0.77 \\ 
@@
 \end{tabular}
-\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM-CD)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
+\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
 \label{tab:WLM_res}
@@
 
-\clearpage
-
-\begin{figure}[h]
-\includegraphics{Figs/WLM_plot}
-\end{figure}
-
 \end{document}

### 2022-03-21 _targets.R

@@ -140,7 +140,7 @@ tar_plan(
     ungroup(),
   
   sim_gaus_var_select_tbl_latex = sim_gaus_var_select_tbl %>% 
-    format_var_slct_tbls(caption = "", label = "")%>% 
+    format_var_slct_tbls(caption = "", label = "tab:sim_gaus_var_select")%>% 
     cat(file = "Manuscript/Tabs/sim_gaus_var_slct_tab.tex"),
   
   #*** Time  ####
@@ -327,9 +327,11 @@ tar_plan(
     plot_smooth_term(ECB_bamlasso_fnl, ECB_bamlasso_nonlnr, ECB_sm_obj$Smooth,
                      min = min(ECB_cov[, ECB_bamlasso_nonlnr])-0.1,
                      max = max(ECB_cov[, ECB_bamlasso_nonlnr]) + 0.1)+
-      xlab(ECB_bamlasso_nonlnr)+
+      xlab(str_split(ECB_bamlasso_nonlnr, "[.]")[[1]][1])+
       theme_pubr()+
-      theme(axis.title.y = element_blank()),
+      theme(axis.title.y = element_blank(),
+            axis.text.x = element_blank(),
+            axis.text.y = element_blank()),
     pattern = map(ECB_bamlasso_nonlnr),
     iteration = "list"
   ),

### 2022-03-21 Manuscript/Tabs/sim_binom_tab.tex
Edit Captions

@@
    50 & 0.76 (0.03) & 0.83 (0.01) & 0.83 (0.02) & 0.84 (0.02) & 0.90 (0.01) & 0.92 (0.01) & 0.89 (0.01) \\ 
-  100 & NaN (NA) & 0.83 (0.01) & 0.83 (0.02) & 0.81 (0.09) & 0.90 (0.01) & 0.92 (0.01) & 0.88 (0.01) \\ 
-  200 & NaN (NA) & 0.83 (0.01) & 0.81 (0.06) & 0.82 (0.05) & 0.88 (0.02) & 0.92 (0.01) & 0.87 (0.02) \\ 
+  100 & - & 0.83 (0.01) & 0.83 (0.02) & 0.81 (0.09) & 0.90 (0.01) & 0.92 (0.01) & 0.88 (0.01) \\ 
+  200 & - & 0.83 (0.01) & 0.81 (0.06) & 0.82 (0.05) & 0.88 (0.02) & 0.92 (0.01) & 0.87 (0.02) \\ 
    \hline
@@
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO),
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
     adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide

### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex
Edit Captions

@@
 \centering
-\begin{tabular}{ccccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
+P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.88 & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.89 \\ 
-   50 & Precision &  & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.52 \\ 
-  100 & Precision &  & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.42 \\ 
-  200 & Precision &  & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.36 \\ 
-    4 & Recall & 1.00 & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
-   10 & Recall & 0.92 & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.54 \\ 
-   50 & Recall & 0.00 & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.55 \\ 
-  100 & Recall &  & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.54 \\ 
-  200 & Recall &  & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.53 \\ 
-    4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.89 & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.55 \\ 
-   50 & MCC &  & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & 0.47 \\ 
-  100 & MCC &  & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & 0.43 \\ 
-  200 & MCC &  & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
+  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Precision & 0.58 & 0.71 & 0.69 & 0.93 & 0.88 & 0.89 \\ 
+   50 & Precision & 0.38 & 0.60 & 0.59 & 0.77 & 0.80 & 0.52 \\ 
+  100 & Precision & 0.35 & 0.62 & 0.65 & 0.82 & 0.77 & 0.42 \\ 
+  200 & Precision & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.36 \\ 
+    4 & Recall & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
+   10 & Recall & 0.46 & 0.50 & 0.54 & 0.38 & 0.97 & 0.54 \\ 
+   50 & Recall & 0.34 & 0.30 & 0.30 & 0.32 & 0.90 & 0.55 \\ 
+  100 & Recall & 0.27 & 0.30 & 0.25 & 0.27 & 0.92 & 0.54 \\ 
+  200 & Recall & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.53 \\ 
+   10 & MCC & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.55 \\ 
+   50 & MCC & 0.25 & 0.34 & 0.36 & 0.45 & 0.83 & 0.47 \\ 
+  100 & MCC & 0.24 & 0.37 & 0.37 & 0.45 & 0.82 & 0.43 \\ 
+  200 & MCC & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
    \hline
@@
 \caption{} 
-\label{}
+\label{tab:sim_binom_var_select}
 \end{table}

### 2022-03-21 Manuscript/Tabs/sim_gaus_tab.tex
replace mgcv NA values (#23)
Edit Captions

@@
    50 & 0.86 (0.02) & 0.32 (0.01) & 0.46 (0.19) & 0.57 (0.18) & 0.80 (0.02) & 0.78 (0.05) & 0.78 (0.01) \\ 
-  100 & NaN (NA) & 0.32 (0.01) & 0.41 (0.23) & 0.48 (0.25) & 0.79 (0.01) & 0.79 (0.05) & 0.77 (0.01) \\ 
-  200 & NaN (NA) & 0.32 (0.01) & 0.39 (0.19) & 0.40 (0.17) & 0.79 (0.01) & 0.78 (0.04) & 0.75 (0.01) \\ 
+  100 & - & 0.32 (0.01) & 0.41 (0.23) & 0.48 (0.25) & 0.79 (0.01) & 0.79 (0.05) & 0.77 (0.01) \\ 
+  200 & - & 0.32 (0.01) & 0.39 (0.19) & 0.40 (0.17) & 0.79 (0.01) & 0.78 (0.04) & 0.75 (0.01) \\ 
    \hline
@@
     Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
     COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation

### 2022-03-21 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
Revise simulaiton results
Revise and add caption to the plot
Edit Captions

@@
 \centering
-\begin{tabular}{ccccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
+P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.89 & 0.76 & 0.84 & 0.93 & 0.62 & 0.86 & 1.00 \\ 
-   50 & Precision & 0.36 & 0.48 & 0.70 & 0.69 & 0.88 & 0.75 & 1.00 \\ 
-  100 & Precision &  & 0.43 & 0.61 & 0.59 & 0.99 & 0.79 & 0.99 \\ 
-  200 & Precision &  & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.99 \\ 
-    4 & Recall & 1.00 & 0.53 & 0.49 & 0.53 & 0.88 & 0.99 & 0.51 \\ 
-   10 & Recall & 1.00 & 0.40 & 0.52 & 0.52 & 0.83 & 1.00 & 0.50 \\ 
-   50 & Recall & 1.00 & 0.35 & 0.40 & 0.48 & 0.37 & 1.00 & 0.50 \\ 
-  100 & Recall &  & 0.33 & 0.36 & 0.40 & 0.30 & 0.99 & 0.50 \\ 
-  200 & Recall &  & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.50 \\ 
-    4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.90 & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.61 \\ 
-   50 & MCC & 0.54 & 0.31 & 0.47 & 0.53 & 0.50 & 0.83 & 0.69 \\ 
-  100 & MCC &  & 0.32 & 0.41 & 0.45 & 0.53 & 0.87 & 0.70 \\ 
-  200 & MCC &  & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & 0.70 \\ 
+  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Precision & 0.76 & 0.84 & 0.93 & 0.62 & 0.86 & 1.00 \\ 
+   50 & Precision & 0.48 & 0.70 & 0.69 & 0.88 & 0.75 & 1.00 \\ 
+  100 & Precision & 0.43 & 0.61 & 0.59 & 0.99 & 0.79 & 0.99 \\ 
+  200 & Precision & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.99 \\ 
+    4 & Recall & 0.53 & 0.49 & 0.53 & 0.88 & 0.99 & 0.51 \\ 
+   10 & Recall & 0.40 & 0.52 & 0.52 & 0.83 & 1.00 & 0.50 \\ 
+   50 & Recall & 0.35 & 0.40 & 0.48 & 0.37 & 1.00 & 0.50 \\ 
+  100 & Recall & 0.33 & 0.36 & 0.40 & 0.30 & 0.99 & 0.50 \\ 
+  200 & Recall & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.50 \\ 
+   10 & MCC & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.61 \\ 
+   50 & MCC & 0.31 & 0.47 & 0.53 & 0.50 & 0.83 & 0.69 \\ 
+  100 & MCC & 0.32 & 0.41 & 0.45 & 0.53 & 0.87 & 0.70 \\ 
+  200 & MCC & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & 0.70 \\ 
    \hline
@@
 \caption{} 
-\label{}
+\label{tab:sim_gaus_var_select}
 \end{table}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_tab.tex
Organize prediction Table to latex output (#23)
replace mgcv NA values (#23)
Edit Captions

--- 2022-02-22 Manuscript/Tabs/sim_binom_tab.tex
@@
   \hline
-  4 & 0.94 (0.01) & 0.83 (0.01) & 0.90 (0.01) & 0.90 (0.01) & 0.92 (0.01) & 0.92 (0.01) & 0.90 (0.00) \\ 
-   10 & 0.92 (0.03) & 0.83 (0.00) & 0.86 (0.04) & 0.86 (0.03) & 0.92 (0.01) & 0.92 (0.01) & 0.90 (0.00) \\ 
-   50 & 0.76 (0.03) & 0.83 (0.01) & 0.83 (0.02) & 0.84 (0.02) & 0.90 (0.01) & 0.92 (0.01) & 0.89 (0.01) \\ 
-  100 & NaN (NA) & 0.83 (0.01) & 0.83 (0.02) & 0.81 (0.09) & 0.90 (0.01) & 0.92 (0.01) & 0.88 (0.01) \\ 
-  200 & NaN (NA) & 0.83 (0.01) & 0.81 (0.06) & 0.82 (0.05) & 0.88 (0.02) & 0.92 (0.01) & 0.87 (0.02) \\ 
+  4 & 0.79 (0.01) & 0.79 (0.01) & 0.76 (0.04) & 0.75 (0.04) & 0.78 (0.01) & 0.76 (0.01) & 0.79 (0.01) \\ 
+   10 & 0.77 (0.01) & 0.79 (0.01) & 0.78 (0.01) & 0.78 (0.01) & 0.78 (0.01) & 0.75 (0.01) & 0.79 (0.01) \\ 
+   50 & 0.62 (0.01) & 0.78 (0.01) & 0.75 (0.03) & 0.73 (0.04) & 0.74 (0.07) & 0.75 (0.02) & 0.77 (0.01) \\ 
+  100 & - & 0.78 (0.01) & 0.73 (0.04) & 0.69 (0.05) & 0.73 (0.07) & 0.74 (0.02) & 0.76 (0.02) \\ 
+  200 & - & 0.78 (0.01) & 0.71 (0.05) & 0.67 (0.05) & 0.73 (0.06) & 0.73 (0.03) & 0.72 (0.03) \\ 
    \hline
@@
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO),
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
     adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 
-\label{tab:bin_auc}
+\label{tab:lnr_bin_auc}
 \end{table}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
Edit Captions

@@
 \centering
-\begin{tabular}{ccccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
+P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.94 & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.91 \\ 
-   50 & Precision &  & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.61 \\ 
-  100 & Precision &  & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.57 \\ 
-  200 & Precision &  & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.38 \\ 
-    4 & Recall & 1.00 & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Recall & 1.00 & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 1.00 \\ 
-   50 & Recall & 0.00 & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 1.00 \\ 
-  100 & Recall &  & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.99 \\ 
-  200 & Recall &  & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.98 \\ 
-    4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.94 & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.91 \\ 
-   50 & MCC &  & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.74 \\ 
-  100 & MCC &  & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.72 \\ 
-  200 & MCC &  & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
+  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Precision & 0.61 & 0.97 & 0.98 & 0.55 & 0.74 & 0.91 \\ 
+   50 & Precision & 0.35 & 0.59 & 0.68 & 0.26 & 0.59 & 0.61 \\ 
+  100 & Precision & 0.28 & 0.47 & 0.51 & 0.30 & 0.46 & 0.57 \\ 
+  200 & Precision & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.38 \\ 
+    4 & Recall & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Recall & 1.00 & 0.96 & 0.97 & 0.99 & 1.00 & 1.00 \\ 
+   50 & Recall & 1.00 & 0.88 & 0.74 & 0.92 & 1.00 & 1.00 \\ 
+  100 & Recall & 1.00 & 0.78 & 0.56 & 0.89 & 1.00 & 0.99 \\ 
+  200 & Recall & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.98 \\ 
+   10 & MCC & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.91 \\ 
+   50 & MCC & 0.52 & 0.66 & 0.65 & 0.37 & 0.73 & 0.74 \\ 
+  100 & MCC & 0.49 & 0.55 & 0.47 & 0.43 & 0.65 & 0.72 \\ 
+  200 & MCC & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
    \hline
@@
 \caption{} 
-\label{}
+\label{tab:sim_lnr_binom_var_select}
 \end{table}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_tab.tex
Organize prediction Table to latex output (#23)
Edit Captions

--- 2022-02-22 Manuscript/Tabs/sim_gaus_tab.tex
@@
   \hline
-  4 & 0.90 (0.01) & 0.33 (0.01) & 0.71 (0.13) & 0.72 (0.11) & 0.90 (0.01) & 0.79 (0.04) & 0.80 (0.00) \\ 
-   10 & 0.90 (0.01) & 0.33 (0.01) & 0.66 (0.21) & 0.77 (0.02) & 0.89 (0.01) & 0.79 (0.04) & 0.79 (0.00) \\ 
-   50 & 0.86 (0.02) & 0.32 (0.01) & 0.46 (0.19) & 0.57 (0.18) & 0.80 (0.02) & 0.78 (0.05) & 0.78 (0.01) \\ 
-  100 & NaN (NA) & 0.32 (0.01) & 0.41 (0.23) & 0.48 (0.25) & 0.79 (0.01) & 0.79 (0.05) & 0.77 (0.01) \\ 
-  200 & NaN (NA) & 0.32 (0.01) & 0.39 (0.19) & 0.40 (0.17) & 0.79 (0.01) & 0.78 (0.04) & 0.75 (0.01) \\ 
+  4 & 0.38 (0.01) & 0.39 (0.01) & 0.31 (0.08) & 0.29 (0.11) & 0.38 (0.01) & 0.35 (0.01) & 0.39 (0.01) \\ 
+   10 & 0.36 (0.02) & 0.38 (0.01) & 0.35 (0.03) & 0.34 (0.04) & 0.39 (0.01) & 0.33 (0.02) & 0.39 (0.01) \\ 
+   50 & 0.09 (0.09) & 0.37 (0.01) & 0.30 (0.06) & 0.30 (0.36) & 0.38 (0.01) & 0.32 (0.03) & 0.37 (0.01) \\ 
+  100 & - & 0.37 (0.01) & 0.28 (0.07) & 0.34 (0.04) & 0.38 (0.01) & 0.29 (0.07) & 0.35 (0.01) \\ 
+  200 & - & 0.36 (0.01) & 0.26 (0.08) & 0.31 (0.06) & 0.38 (0.03) & 0.28 (0.06) & 0.33 (0.02) \\ 
    \hline
@@
     Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
     COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
     whe number of parameters exceeds sample size i.e. p = 100, 200.} 
-\label{tab:gaus}
+\label{tab:lnr_gaus}
 \end{table}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
Edit Captions

@@
 \centering
-\begin{tabular}{ccccccccc}
+\begin{tabular}{cccccccc}
   \hline
-P & Metric & mgcv & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
+P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
   \hline
-  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Precision & 0.91 & 0.59 & 0.97 & 0.97 & 0.90 & 0.56 & 0.99 \\ 
-   50 & Precision & 0.39 & 0.43 & 0.74 & 0.84 & 0.89 & 0.39 & 0.99 \\ 
-  100 & Precision &  & 0.34 & 0.53 & 0.72 & 0.91 & 0.29 & 0.99 \\ 
-  200 & Precision &  & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.99 \\ 
-    4 & Recall & 1.00 & 1.00 & 0.87 & 0.84 & 1.00 & 1.00 & 1.00 \\ 
-   10 & Recall & 1.00 & 1.00 & 0.98 & 0.96 & 1.00 & 1.00 & 1.00 \\ 
-   50 & Recall & 1.00 & 1.00 & 0.85 & 0.99 & 1.00 & 1.00 & 1.00 \\ 
-  100 & Recall &  & 1.00 & 0.84 & 0.97 & 1.00 & 1.00 & 1.00 \\ 
-  200 & Recall &  & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 1.00 \\ 
-    4 & MCC &  &  &  &  &  &  &  \\ 
-   10 & MCC & 0.91 & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.99 \\ 
-   50 & MCC & 0.57 & 0.59 & 0.76 & 0.90 & 0.92 & 0.57 & 1.00 \\ 
-  100 & MCC &  & 0.54 & 0.61 & 0.81 & 0.94 & 0.48 & 1.00 \\ 
-  200 & MCC &  & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.99 \\ 
+  4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Precision & 0.59 & 0.97 & 0.97 & 0.90 & 0.56 & 0.99 \\ 
+   50 & Precision & 0.43 & 0.74 & 0.84 & 0.89 & 0.39 & 0.99 \\ 
+  100 & Precision & 0.34 & 0.53 & 0.72 & 0.91 & 0.29 & 0.99 \\ 
+  200 & Precision & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.99 \\ 
+    4 & Recall & 1.00 & 0.87 & 0.84 & 1.00 & 1.00 & 1.00 \\ 
+   10 & Recall & 1.00 & 0.98 & 0.96 & 1.00 & 1.00 & 1.00 \\ 
+   50 & Recall & 1.00 & 0.85 & 0.99 & 1.00 & 1.00 & 1.00 \\ 
+  100 & Recall & 1.00 & 0.84 & 0.97 & 1.00 & 1.00 & 1.00 \\ 
+  200 & Recall & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 1.00 \\ 
+   10 & MCC & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.99 \\ 
+   50 & MCC & 0.59 & 0.76 & 0.90 & 0.92 & 0.57 & 1.00 \\ 
+  100 & MCC & 0.54 & 0.61 & 0.81 & 0.94 & 0.48 & 1.00 \\ 
+  200 & MCC & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.99 \\ 
    \hline
@@
 \caption{} 
-\label{}
+\label{tab:sim_lnr_gaus_var_select}
 \end{table}

### 2022-03-21 R/Sim_Res/format_var_slct_tbls.R

@@ -5,11 +5,12 @@ format_var_slct_tbls <- function( dat, caption, label){
                            labels = c("Precision", "Recall", "MCC")),
            p = as.integer(p)
     )%>% 
+    filter(!(Metric == "MCC" & p==4)) %>%
     arrange(Metric, p) %>% 
     select(
       "P" = p,
       Metric,
-      "mgcv" = mgcv,
+      # "mgcv" = mgcv,
       "LASSO" = lasso,
       "COSSO" = cosso,
       "Adaptive COSSO" = acosso,
@@ -20,7 +21,7 @@ format_var_slct_tbls <- function( dat, caption, label){
       "spikeSlabGAM" = ssGAM
     ) %>% 
     xtable(
-      align = "cccccccccc",
+      align = "ccccccccc",
       caption = caption,
       label = label
     ) %>% 

### 2022-03-21 _targets.R

@@ -41,8 +41,7 @@ tar_plan(
     sim_binom_tab,
     caption = "The average and standard deviation of the out-of-sample area under the curve measures
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO),
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
     adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:bin_auc")%>% 
@@ -79,7 +78,7 @@ tar_plan(
     ungroup(),
   
   sim_binom_var_select_tbl_latex = sim_binom_var_select_tbl %>% 
-    format_var_slct_tbls(caption = "", label = "")%>% 
+    format_var_slct_tbls(caption = "", label = "tab:sim_binom_var_select")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_var_slct_tab.tex"),
   # Visual Presentation of Var Selection
   # make_sim_var_metric_raw(sim_binom_var_select_raw[[4]]) %>% 
@@ -104,8 +103,7 @@ tar_plan(
     sim_gaus_tab,
     caption = "The average and standard deviation of the out-of-sample $R^2$ measure for
     Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM) fitted with Iterative Weighted Least Square (BHAM-IWLS) and
-    Coordinate Descent (BHAM-CD) algorithms, component selection and smoothing operator (COSSO), adaptive
+    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
     COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
     whe number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:gaus") %>% 
@@ -196,8 +194,7 @@ tar_plan(
     # TODO: edit the caption
     caption = "The average and standard deviation of the out-of-sample area under the curve measures
[…]

### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex
Update var selection captions.

@@
 \end{tabular}
-\caption{} 
+\caption{The variable selection performance of binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
 \label{tab:sim_binom_var_select}

### 2022-03-21 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
Update var selection captions.

@@
 \end{tabular}
-\caption{} 
+\caption{The variable selection performance of Gaussian simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
 \label{tab:sim_gaus_var_select}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
Update var selection captions.

@@
 \end{tabular}
-\caption{} 
+\caption{The variable selection performance of linear binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
 \label{tab:sim_lnr_binom_var_select}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
Update var selection captions.

@@
 \end{tabular}
-\caption{} 
+\caption{The variable selection performance of linear Gaussian simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
+The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
 \label{tab:sim_lnr_gaus_var_select}

### 2022-03-21 _targets.R

@@ -78,7 +78,8 @@ tar_plan(
     ungroup(),
   
   sim_binom_var_select_tbl_latex = sim_binom_var_select_tbl %>% 
-    format_var_slct_tbls(caption = "", label = "tab:sim_binom_var_select")%>% 
+    format_var_slct_tbls(caption = "The variable selection performance of binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations. The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.",
+                         label = "tab:sim_binom_var_select")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_var_slct_tab.tex"),
   # Visual Presentation of Var Selection
   # make_sim_var_metric_raw(sim_binom_var_select_raw[[4]]) %>% 
@@ -138,7 +139,7 @@ tar_plan(
     ungroup(),
   
   sim_gaus_var_select_tbl_latex = sim_gaus_var_select_tbl %>% 
-    format_var_slct_tbls(caption = "", label = "tab:sim_gaus_var_select")%>% 
+    format_var_slct_tbls(caption = "The variable selection performance of Gaussian simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations. The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.", label = "tab:sim_gaus_var_select")%>% 
     cat(file = "Manuscript/Tabs/sim_gaus_var_slct_tab.tex"),
   
   #*** Time  ####
@@ -222,7 +223,8 @@ tar_plan(
     ungroup(),
   
   sim_lnr_binom_var_select_tbl_latex = sim_lnr_binom_var_select_tbl %>% 
[…]

### 2022-03-21 Manuscript/Tabs/sim_binom_tab.tex
Revise caption

@@
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
-    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
+    hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing
+    operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and
+    spikeSlabGAM mdoel. mgcv doesn't provide
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 

### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex
Revise caption

@@
 \end{tabular}
-\caption{The variable selection performance of binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
-MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
+\caption{The variable selection performance of binomial simulations,
+                         measured by positive predictive value (precision), true positive rate (recall),
+                         and Matthews correlation coefficient (MCC), for the high-dimensional methods 
+                         averaged over 50 iterations. The models of comparison include the proposed Bayesian
+                         hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and
+                         smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM),
+                         and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.} 
 \label{tab:sim_binom_var_select}

### 2022-03-21 Manuscript/Tabs/sim_gaus_tab.tex
Revise caption

@@
 \caption{The average and standard deviation of the out-of-sample $R^2$ measure for
-    Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
-    COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
+    Gaussian outcomes over 50 iterations. The models of comparison include the proposed
+    Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
+    selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
+    generalized additive model (SB-GAM), and spikeSlabGAM mdoel. mgcv doesn't provide estimation
     whe number of parameters exceeds sample size i.e. p = 100, 200.} 

### 2022-03-21 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
Revise caption

@@
 \end{tabular}
-\caption{The variable selection performance of Gaussian simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
-MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
+\caption{The variable selection performance of Gaussian simulations,
+                         measured by positive predictive value (precision), true positive rate (recall),
+                         and Matthews correlation coefficient (MCC), for the high-dimensional methods
+                         averaged over 50 iterations. The models of comparison include the proposed Bayesian
+                         hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection
+                         and smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive
+                         model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation
+                         (no true negative), and hence omitted for all methods.} 
 \label{tab:sim_gaus_var_select}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_tab.tex
Revise caption

@@
 \caption{The average and standard deviation of the out-of-sample area under the curve measures
-    for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
-    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
+    for binomial outcomes over 50 iterations. The models of comparison include the proposed
+    Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
+    selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
+    generalized additive model (SB-GAM), and spikeSlabGAM mdoel. mgcv doesn't provide
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
Revise caption

@@
 \end{tabular}
-\caption{The variable selection performance of linear binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
-MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.}
+\caption{The variable selection performance of linear binomial simulations,
+                         measured by positive predictive value (precision), true positive rate (recall),
+                         and Matthews correlation coefficient (MCC), for the high-dimensional methods
+                         averaged over 50 iterations. The models of comparison include the proposed Bayesian
+                         hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection
+                         and smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized
+                         additive model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$
+                         simulation (no true negative), and hence omitted for all methods.} 
 \label{tab:sim_lnr_binom_var_select}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
Revise caption

@@
 \end{tabular}
-\caption{The variable selection performance of linear Gaussian simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations.
-The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+\caption{The variable selection performance of linear Gaussian
+                         simulations, measured by positive predictive value (precision),
+                         true positive rate (recall), and Matthews correlation coefficient (MCC),
+                         for the high-dimensional methods averaged over 50 iterations.
+                         The models of comparison include the proposed Bayesian hierarchical additive model (BHAM),
+                         linear LASSO model (LASSO), component selection and smoothing operator (COSSO),
+                         adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
                          MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.} 

### 2022-03-21 Manuscript/Tabs/sim_time_tab.tex
Revise caption

@@
       including cross-validation and final model fitting, over 50 iterations. The models
-      of comparison include the proposed Bayesian hierarchical additive model (BHAM) fitted
-      with Iterative Weighted Least Square (BHAM-IWLS) and Coordinate Descent (BHAM-CD)
-      algorithms, component selection and smoothing operator (COSSO), adaptive COSSO,
-      mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't 
+      of comparison include the proposed Bayesian hierarchical additive model (BHAM), the linear LASSO model (LASSO),
+      component selection and smoothing operator (COSSO), adaptive COSSO,
+      mgcv, sparse Bayesian generalized additive model (SB-GAM), spikeSlabGAM. mgcv doesn't 
       provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 

### 2022-03-21 _targets.R

@@ -41,8 +41,9 @@ tar_plan(
     sim_binom_tab,
     caption = "The average and standard deviation of the out-of-sample area under the curve measures
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
-    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
+    hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing
+    operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and
+    spikeSlabGAM mdoel. mgcv doesn't provide
     estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:bin_auc")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_tab.tex"),
@@ -78,7 +79,13 @@ tar_plan(
     ungroup(),
   
   sim_binom_var_select_tbl_latex = sim_binom_var_select_tbl %>% 
-    format_var_slct_tbls(caption = "The variable selection performance of binomial simulations, measured by positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC), for the high-dimensional methods averaged over 50 iterations. The models of comparison include the proposed Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.",
+    format_var_slct_tbls(caption = "The variable selection performance of binomial simulations,
+                         measured by positive predictive value (precision), true positive rate (recall),
+                         and Matthews correlation coefficient (MCC), for the high-dimensional methods 
+                         averaged over 50 iterations. The models of comparison include the proposed Bayesian
+                         hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and
+                         smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM),
+                         and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.",
[…]

### 2022-03-21 Manuscript/00-main.Rmd
Revise caption

@@
   flextable::set_caption(caption = "Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.")
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM and the model of comparison SB-GAM.")
 ```

### 2022-03-21 Manuscript/03-Simulation.Rmd
Revise caption

@@
 \bg{\subsubsection{Variable Selection Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
 The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.

### 2022-03-21 Manuscript/SS_GAM.tex
Revise caption

@@
 \textcolor{red}{\subsubsection{Variable Selection Performance}}
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
 The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
@@
 \caption{Model fitting time in seconds for two metabolomics data analyses, from Emory Cardiovascular Biobank (ECB) and Weight Loss Maintenance Cohort (WLM).
-It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM with EM-CD algorithm (BHAM-CD) and the model of comparison SB-GAM.
+It tabulates the computation time for cross-validation step (CV) and optimal model fitting step (Final), and total computation time (Total) for the proposed model BHAM and the model of comparison SB-GAM.
 }\\
@@
 
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM-CD}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{BHAM}}} & \multicolumn{3}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 2.25in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{SB-GAM}}} \\
 
@@
 
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{365.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{6.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{371.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{WLM}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{365.4}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{6.8}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{372.2}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,116.0}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{32.7}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{11}{11}\selectfont{\textcolor[HTML]{000000}{3,148.7}}} \\
 

### 2022-03-21 R/real_data_time.R

@@ -26,7 +26,7 @@ create_real_data_time_tbl <- function(...){
       col_keys,"{line2}_{line3}",
       na = "Data",
       remove = FALSE) %>% 
-    mutate(line2 = recode(line2, bamlasso = "BHAM-CD",
+    mutate(line2 = recode(line2, bamlasso = "BHAM",
                           SBGAM = "SB-GAM"),
            line3 = recode(line3, cv = "CV", fnl = "Final",
                           total = "Total"))
@@ -34,14 +34,14 @@ create_real_data_time_tbl <- function(...){
   flextable(time_tbl_raw) %>% 
     colformat_double(digits = 1) %>% 
     set_header_df(mapping = time_tbl_header) %>% 
-    set_header_labels(bamlasso = "BHAM-CD",
+    set_header_labels(bamlasso = "BHAM",
                       SBGAM = "SB-GAM",
                       cv = "CV",
                       fnl = "Final",
                       total = "Total") %>% 
     merge_h(part = "header", i = 1) %>% 
     merge_v(part = "header", j = 1) %>% 
-    theme_booktabs %>% 
-    align(align = "center", part = "all")
+    flextable::theme_booktabs() %>% 
+    flextable::align(align = "center", part = "all")
   
 }

### 2022-03-21 _targets.R

@@ -6,7 +6,7 @@ library(tarchetypes)
 tar_option_set(
   packages = c("tidyverse", "dplyr", "rticles", #"gtsummary", "gt", 
                "broom", "unglue", "knitr", "ggpubr", "xtable", "janitor", 
-               # "flextable", 
+               "flextable", 
                "mgcv", "BHAM", "BhGLM", "sparseGAM",  "glmnet", "yardstick",
                "rmarkdown"#, "cosso",
   ),

### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex
add hlines to var select table

@@
 \begin{tabular}{cccccccc}
-  \hline
  P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
-  \hline
    4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
@@
   200 & Precision & 0.29 & 0.65 & 0.57 & 0.21 & 0.74 & 0.36 \\ 
+   \hline
   4 & Recall & 0.54 & 0.61 & 0.54 & 0.38 & 0.97 & 0.55 \\ 
@@
   200 & Recall & 0.27 & 0.25 & 0.29 & 0.56 & 0.92 & 0.53 \\ 
+   \hline
  10 & MCC & 0.18 & 0.35 & 0.36 & 0.47 & 0.86 & 0.55 \\ 
@@
   200 & MCC & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
-   \hline
   \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
add hlines to var select table

@@
   200 & Precision & 0.36 & 0.61 & 0.47 & 0.28 & 0.75 & 0.99 \\ 
+   \hline
   4 & Recall & 0.53 & 0.49 & 0.53 & 0.88 & 0.99 & 0.51 \\ 
@@
   200 & Recall & 0.32 & 0.33 & 0.35 & 0.52 & 1.00 & 0.50 \\ 
+   \hline
  10 & MCC & 0.32 & 0.49 & 0.57 & 0.46 & 0.86 & 0.61 \\ 
@@
   200 & MCC & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & 0.70 \\ 
-   \hline
   \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
add hlines to var select table

@@
   200 & Precision & 0.26 & 0.44 & 0.42 & 0.38 & 0.41 & 0.38 \\ 
+   \hline
   4 & Recall & 1.00 & 0.85 & 0.78 & 1.00 & 1.00 & 1.00 \\ 
@@
   200 & Recall & 1.00 & 0.67 & 0.48 & 0.89 & 0.99 & 0.98 \\ 
+   \hline
  10 & MCC & 0.58 & 0.95 & 0.96 & 0.49 & 0.72 & 0.91 \\ 
@@
   200 & MCC & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
-   \hline
   \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
add hlines to var select table

@@
   200 & Precision & 0.27 & 0.42 & 0.52 & 0.92 & 0.33 & 0.99 \\ 
+   \hline
   4 & Recall & 1.00 & 0.87 & 0.84 & 1.00 & 1.00 & 1.00 \\ 
@@
   200 & Recall & 1.00 & 0.78 & 0.88 & 1.00 & 0.98 & 1.00 \\ 
+   \hline
  10 & MCC & 0.59 & 0.96 & 0.94 & 0.91 & 0.54 & 0.99 \\ 
@@
   200 & MCC & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.99 \\ 
-   \hline
   \end{tabular}

### 2022-03-21 R/Sim_Res/format_var_slct_tbls.R

@@ -25,5 +25,6 @@ format_var_slct_tbls <- function( dat, caption, label){
       caption = caption,
       label = label
     ) %>% 
-    print(comment = FALSE, include.rownames = FALSE)
+    print(comment = FALSE, include.rownames = FALSE,
+          hline.after = c(5,10))
 }

### 2022-03-21 Manuscript/Tabs/sim_binom_var_slct_tab.tex
correct hlines position for var select tables

@@
 \begin{tabular}{cccccccc}
+  \hline
 P & Metric & LASSO & COSSO & Adaptive COSSO & BHAM & SB-GAM & spikeSlabGAM \\ 
+  \hline
   4 & Precision & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
@@
   200 & MCC & 0.23 & 0.36 & 0.36 & 0.23 & 0.81 & 0.40 \\ 
+   \hline
 \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
correct hlines position for var select tables

@@
   200 & MCC & 0.49 & 0.49 & 0.39 & 0.47 & 0.62 & 0.58 \\ 
+   \hline
 \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_gaus_var_slct_tab.tex
correct hlines position for var select tables

@@
   200 & MCC & 0.28 & 0.41 & 0.38 & 0.36 & 0.85 & 0.70 \\ 
+   \hline
 \end{tabular}

### 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
correct hlines position for var select tables

@@
   200 & MCC & 0.49 & 0.52 & 0.64 & 0.95 & 0.52 & 0.99 \\ 
+   \hline
 \end{tabular}

### 2022-03-21 R/Sim_Res/format_var_slct_tbls.R

@@ -26,5 +26,5 @@ format_var_slct_tbls <- function( dat, caption, label){
       label = label
     ) %>% 
     print(comment = FALSE, include.rownames = FALSE,
-          hline.after = c(5,10))
+          hline.after = c(-1, 0, nrow(.), 5,10))
 }

### 2022-03-22 Manuscript/10-Supplementary.tex
Update Supplementary File

@@
 ]{article}
+\title{Spike-and-Slab Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Supporting Infromation}
+\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
+Nengjun Yi}
+\date{}
+
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
-  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-
-\title{Spike-and-Slab Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Supporting Infromation}
-\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
-Nengjun Yi}
-\date{}
 
@@
 \newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\simiid}{\overset{\text{iid}}{\sim}}
 
@@
 the E-step. To note, unlike the same latent indicator \(\gamma^*_j\)
-which is shared by the coefficients of the non-linear terms
+which is shared by the coefficients of the nonlinear terms
 \(\beta^*_{jk}\) for \(k = 1, \dots, K_j\) , \(\tau^2_{jk}\) is
@@
 
+\subsubsection{Other Priors}
+
+With the re-parameterization step of the basis function matrix
+\(\boldsymbol{X}\), it is possible to generalized the SSL prior to other
+priors, for example normal priors for ridge-type regularization and
+mixture normal prior for spike-and-slab regularization. These priors
+would work better in low and medium dimensional settings where the
+sparse assumption is not necessary. Here we elaborate the mixture normal
+prior as a demonstration of applying continuous spike-and-slab prior in
+BHAM.
+
+A spike-and-slab mixture normal spline prior can be expressed as
+\begin{align*}
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
+\end{align*} Similar to the spike-and-slab spline prior in Equation
+(\ref{eq:bham_ssl}), \(0 < s_0 < s_1\) are tuning parameters and can be
+optimized via cross-validation. One of the critics received by the
+spike-and-slab mixture normal prior is that the tails of a normal
+distribution diminishes to zero too fast, which causes problems when
+estimating the large effects. Distributions with heavier tails can be
+used as an alternative, for example mixture Student's \(t\) distribution
+with small degree of freedom.
+
 \clearpage
@@
 
+\clearpage
+
+\subsection{Supplementary Information 4: Variable Selection Performance of Simulations}
+\input{Tabs/sim_binom_var_slct_tab.tex}
+\input{Tabs/sim_lnr_gaus_var_slct_tab.tex}
+\input{Tabs/sim_lnr_binom_var_slct_tab.tex}
+
 \end{document}

### 2022-03-22 Manuscript/R&R/00-response_letter.Rmd
Add citations (#25)
Rough draft of the response
Revise response letter

--- 2022-02-11 Manuscript/R&R/00-response_letter.Rmd
@@
     keep_tex: FALSE
-    citation_package: biblatex
 bibliography: reply_ref.bib
+header-includes: 
+ - \usepackage{color}
+link-citations: true
 ---
+
+\newcommand{\bg}[1]{\textcolor{red}{#1}}
 
@@
 \newcommand{\tp}{*}
+\newcommand{\bs}{\boldsymbol}
 
@@
 \begin{itemize}
-    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects inclusion indicator such that the strong hierarchy for bi-level selection is enforced.
-    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear lasso model and spikeSlabGAM model among the models to be compared. 
-    \item According to Reviewer \#1's and Reviewer \#4's comments, we added a new set of simulation scenarios where the functions of predictors are linear.  
-    \item According to Reviewer \#4's comment, we summarized the variable selection performance for all models of comparison and bi-level selection performance for the models that is capable, i.e.
-the proposed model and spikeSlabGAM.
-    \item According to Reviewer \#4's comment, we recorded the computation performance of spikeSlabGAM.
-    \item All existing figure and tables were updated to reflect the methodology change and simulation design change.
-Figures were added according to Reviewer \#4's comment.
-[TODO: think what figures are needed]
-    \item We fixed the additional typos and grammar errors according to Review \#2.The revision of the manuscript is marked in red [TODO: change this color to red]. Deletions are not highlighted.
+    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects to reflect the effect hierarchy principle for bi-level selection.
+We updated the simulation and real data analysis results.
+    \item We streamlined the structure of the manuscript, moving the previous EM-IWLS fitting algorithm subsection to the supporting information for three reasons.
+First of all, moving can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction.
+Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g.
+constructing hypothesis testings and confidence band, requires a more dedicated discussion.
+We are preparing another manuscript discussing those.
+Lastly, with the added variable selection performance, the manuscript would be length if keeping the EM-IWLS in the main body.
+The readability of the manuscript would be damaged.
+Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasize the scalability.
+    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study.
+In addition, we added a new set of simulation scenarios where the functions of predictors are linear.
+    \item According to Reviewer \#4's comment, we summarized the variable selection and bi-level selection performance in the simulation study section.
+The computation performance of spikeSlabGAM is recorded and addressed too.
+    \item According to Reviewer \#4's comment, we added the plot showing the estimated functions of predictors for real-data analysis.
+    \item We updated the feature screening procedure in our real-data analysis of Emory Cardiovascular Biobank dataset.
+Our previous screening method, fitting univariate GAM, made the "double-dipping" mistake that would overestimate the prediction performance of any downstream prediction model.
+We updated the screening procedure with variance-based selection, and the prediction of both methods dropped.
+Nevertheless, the proposed model still outperforms the compared method.
+See change on P9.
+    \item We fixed the additional typos and grammar errors according to Review \#2.
+The revision of the manuscript is marked in \bg{red} (roughly due to the technical difficulties associated with Latex).
+Deletions are not highlighted.
 \end{itemize}

### 2022-03-22 Manuscript/R&R/01-reviewer_1.Rmd
Revise response letter

@@
 __Response__: Thank you for the comment.
-We expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds.
+We added linear LASSO models to our simulation studies and see that the proposed model predicts as good as linear LASSO models when the underlying signals are linear, and substantially better when the underlying signals are nonlinear.
+This empirically demonstrates the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals.
+The simulation results also match with our expectation.
+The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as bi-level selection problem in the manuscript.
+When the underlying signal is linear, the nonlinear component of the smoothing function would not be included in the model due to the proposed two-part spike-and-slab lasso prior.
+
+<!-- expanded our simulation study from two perspectives: 1) we added linear lasso models and spikeSlabGAM models, where linear lasso models serve as the reference to benchmark the performance of linear models; 2) we added new scenarios where the the functional form of each predictors are linear, to demonstrate the robustness of the proposed, regarding prediction and variable selection performances, when the non-linear assumption no longer holds.
 Please see the changes and conclusion in \emph{Section 3: SIMULATION STUDY}.
@@
 The expanded simulation study also demonstrates that the proposed model is flexible to model both linear and non-linear effects and broadly applicable in high-dimensional modeling without much constraints, which deems to be an additional strength.
+-->
 
@@
 
-__Response__: We thank the reviewer for the comment.
+__Response__: Thank you for the comment.
 We didn't perform the data processing step ourselves.
-We used the publicly available data that has been pre-processed.
-The data processing step is described in [TODO: add citation].
-Meanwhile, we respectively argue that the small R2 is reasonable, as in many applications, genomics data can only accounts small amount of data variation.
+We used the publicly available data that has been pre-processed, which is clarified in the manuscript on P9.
+
+<!-- %Meanwhile, we respectively argue that the small R2 is reasonable, as in many applications, genomics data can only accounts small amount of data variation.
 We think this example would be the case, as environmental factors (not accounted in these models) would be more substantial when explaining weight loss.
+-->
 \bigskip
@@
 The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms.
-In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalably model.
-We included an additional sentence in the discussion section.
-[TODO: list the sentence here].
+In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model.
+We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.").
 
@@
 
-We revised our method section to highlight these advantages, see Pxxx.
-
 \bigskip
@@
 
-__Response__: We don't think any preliminary checkings or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed.
-Such rules of thumb includes [TODO: Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.].
-As a supporting evidence, our added simulation demonstrates the proposed model performs well, in reference to the linear lasso model, even when signals are linear and sparse (see response to the Comment 1).
+__Response__: We thank the review for the comments.
+For prediction purpose, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing.
+The two-part spike-and-slab LASSO prior automatically detects if the underlying signal is linear or complex.
+When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic spline (implemented in the package) is used to estimate the complex signal.
+As a supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see response to the Comment 1).
 
-We acknowledge the reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge.
+For the bi-level selection purpose, we agree with the reviewer that it is a more delicate problem.
+In our simulation study, we observe that trade-off exists with different underlying signals compared to the only other method, spikeSlabGAM, that is capable of bi-level selection.
+(see P8) We added a discussion about the reason for the trade-off and gave the reader suggestions (see P10).
+
+In conclusion, the proposed model do not addition detection for prediction purpose, and have different utility for bi-level selection.
+
+<!-- reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge.
 However, the discussion of modeling strategy, e.g.
@@
 The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals.
+-->
+
+<!-- We don't think any preliminary checking or testings are necessary before implementing the proposed model, as long as some general rules of thumb for high-dimensional modeling are followed. -->
+<!-- Such rules of thumb includes [TODO: Insert citaiton for ultra high-dimension thing, as the number of predictors is a function of the sample size.].  -->
+
+<!-- We acknowledge the reviewer's comment that careful deliberation are necessary when conducting analysis, which includes domain knowledge.
+However, the discussion of modeling strategy, e.g.
+starting with more flexible models or linear models, or using statistical testings to examine assumptions, is out of the scope of the current manuscript.
+The current manuscript focuses on introducing a novel model that allows flexibility when modeling complex signals.
+-->

### 2022-03-22 Manuscript/R&R/02-reviewer_2.Rmd
Revise response letter

--- 2022-01-09 Manuscript/R&R/02-reviewer_2.Rmd
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+The purpose of the reparameterization step is to factor the smoothing penalty matrix in the design matrix, such that the smoothing penalty can be addressed implicitly.
+This contrasts to some previous methods that do not consider smoothing penalty at all.
+With the reparameterization, the smoothing penalty changes from $\lambda \bs \beta^T \bs S \bs \beta$ to $\lambda \bs {\beta^*}^T \bs {\beta^*}$, where the $\bs \beta$ and $\bs \beta^*$ are the coefficients of the smoothing function before and after the reparameterization respectively.
+However, the smoothing parameter $\lambda$, equivalently the scale parameter $s$ in the proposed model, needs to be estimated/tuned.
+The proposed two-part spike-and-slab LASSO prior provides the the estimation of the smoothing parameter: when the nonlinear effect does not exists, a large shrinkage is applied to the coefficients via the spike density such that coefficients shrink to 0s; when then nonlinear exists, a small shrinkage is applied via the slab density such that the smoothing function can be estimated appropriately.
+Moreover, the local adaptivity of he spike-and-slab LASSO prior allows the shrinkage to adapt to the data, compared to previous methods that impose an uniform shrinkage.
+
+We clarified the utility of the reparameterization step and emphasized how the smoothness of the additive function are appropriately modeled (see P3).
+
 \bigskip
@@
 
-__Response__:
+__Response__: Compared to other priors, the spike-and-slab double exponential prior provides three advantages.
+First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients.
+Hence, the smoothing functions can be estimated more accurately.
+Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
+Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
+
+We revised our method section to highlight these advantages, see P4.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
+This prior formulation is motivated by effect hierarchy discussed in Chipman (2004).
+Now, the inclusion of the nonlinear effects is dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
+The modification of the prior doesn't complicate the model fitting algorithm.
+The $\gamma$ can be analytically integrated out of the density function of $\gamma^*$ and requires minimum change of the proposed algorithms.
+The updated model still shows performance advantages as before.
+
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the review for the comment. We revised the manuscript according to the comment.
 \bigskip

### 2022-03-22 Manuscript/R&R/03-reviewer_4.Rmd
Revise response letter

--- 2022-03-08 Manuscript/R&R/03-reviewer_4.Rmd
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
+This prior formulation is motivated by effect hierarchy discussed in Chipman (2004).
+Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
+The modification of the prior doesn't complicate the model fitting algorithm.
+The $\gamma$ can be analytically integrated out of the density function and requires minimum change of the proposed algorithm.
+In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the review for the comments.
+We made change to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear.
+The suggested model spikeSlabGAM and linear Lasso model as the benchmark were added to the models of comparison.
+When reporting the variable selection performance, we considered three metrics, positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient.
+Bi-level selection performance was provided for the two applicable models, the proposed model and spikeSlabGAM.
+Please see the updated simulation study on P7-9.
+
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the review for the comment.
+In the update manuscript, we add the suggested model, spikeSlabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared.
+We see that the proposed method still preserves the computation advantage over spikeSlabGAM, particularly in the high-dimensional cases.
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment.
+After deliberation, we decide to focusing explaining the EM-CD method well.
+The inclusion of EM-IWLS would distract the reader from the main focus of manuscript, scalable model, good prediction and bi-level selection.
+We fully agree that the questions raised by the reviewer are very important and deserves more careful discussion.
+Nevertheless, that would be out of the scope of the current work.
+We will dedicate another manuscript to detail many of the questions here, for example estimating effective degree of freedom, limiting distribution of the test statistics to the EM-IWLS method.
+
 \bigskip
@@
 
-__Response__:
+__Response__: We thank the reviewer for the comment. We added Figure 2 displaying the estimated functions of the predictors from BHAM.
 \bigskip
@@
 
-Nevertheless, we edited the sentence for clarity. [TODO: check if this is done]
 \bigskip

### 2022-03-22 Manuscript/02-Method.Rmd
Correct typo

@@
 \bg{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For the computational convenience, we analytically integrate $\theta_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$.
+For the computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
 }

### 2022-03-24 Manuscript/00-main.Rmd
Revise discussion, adding IWLS method. Change the title

@@
 ---
-title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
+title: Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis
 author:

### 2022-03-24 Manuscript/05-Discussion.Rmd
Revise discussion, adding IWLS method. Change the title

@@
 Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.
-Not focused in this manuscript, the generalization is described in the supporting materials.}
+Not focused in this manuscript, the generalization is described in the Supporting Information.}
 
@@
 
-Our future efforts direct to survival analysis and integrative analysis.
-While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
+Our future efforts direct to \bg{uncertainty inference of the proposed model}, survival analysis and integrative analysis.
+\bg{Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
+We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
+Instead of the Coordinate Descent algorithm, we use Iterative Weighted Least Square algorithm in the EM procedure.
+The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
+Due to the space limit, technical details will be explained in a future manuscript.} While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
 An naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].

### 2022-03-24 Manuscript/SS_GAM.tex
Revise discussion, adding IWLS method. Change the title

@@
 
-\title{Spike-and-Slab Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data}
+\title{Spike-and-Slab LASSO Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data Analysis}
 
@@
 \textcolor{red}{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For the computational convenience, we analytically integrate $\theta_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$.
+For the computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
 }
@@
 
-Our future efforts direct to survival analysis and integrative analysis.
+Our future efforts direct to
+\textcolor{red}{uncertainty inference of the proposed model}, survival
+analysis and integrative analysis.
+\textcolor{red}{Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
+We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
+Instead of the Coordinate Descent algorithm, we use Iterative Weighted Least Square algorithm in the EM procedure.
+The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
+Due to the space limit, technical details will be explained in a future manuscript.}
 While the proposed model addresses a great deal of analytic problem,

### 2022-03-24 Manuscript/10-Supplementary.Rmd
Update Supplementary File
Revise supporting informaiton

@@
 
-\subsection{Supplementary Information 2: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}
+\subsection{Supplementary Information 2: EM-Iterative Weighted Least Square Algorithm for BHAM Model Fitting}
 
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
-The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted using IWLS in combination with EM algorithm.
-In this work, we adapt the EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior
-<!-- , with the full potential to adopt Student's t, double exponential and mixture Student's t priors when sparse assumption is not necessary -->
-.
-
-<!-- The primary purpose of the EM algorithm here is mitigate the computational burden associate with the hierarchical normal formulation.  -->
+Similar to the EM-CD algorithm, the EM-Iterative Weighted Least Square (EM-IWLS) algorithm is an EM-based algorithm where the iterative weighted least squares algorithm is used to find the estimate of $\bs \beta, \phi$ that maximizes $E(Q_1)$.
+The iterative weighted least squares algorithm was originally proposed to fit the classical generalized linear models, and generalized to fit some Bayesian hierarchical models.[@Gelman2013] Yi and Ma [@Yi2012] extended the algorithm to fit Bayesian hierarchical models for high-dimensional data analysis.
+Specifically, the authors formulated Student's t-distribution and double exponential distribution as hierarchical normal distributions such that generalized linear models with shrinkage priors can be easily fitted.
+In this work, we further extend the EM-IWLS paradigm to fit the proposed BHAM method.
+COmpare to the EM-CD algorithm, EM-IWLS estimate the variance-covariance matrix of the coefficients, providing an opportunity to derive the uncertainty quantification of smoothing functions.
+We will defer this discussion to the future work, due to the delicacy of the topic.
+However, the EM-IWLS is not as computationally efficient as EM_CD, particularly in high-dimensional settings.
 
@@
 \end{align*}
-For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$ following Equation (\ref{eq:ssl}).
+For the mixture double exponential priors, we can define the scale parameter $S = (1-\gamma)s_0 + \gamma s_1$.
 The change in the prior formulation in turn leads to the change in the log posterior density function, as $Q_1$ needs to account for the hyperprior of $\tau^2$:
@@
 $E({S_j}^{-1}|\beta_j, s_0, s_1), E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1), E({\tau}^2_{j}|S_j, \beta_j) \text{ and } E({\tau^\tp}^2_{jk}|S_j^\tp, \beta^\tp_{jk})$ needs to be calculated to formulate $E(Q_1)$.
-As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived using Equation (\ref{eq:exp_scale}).
+As neither $E({S_j}^{-1}|\beta_j, s_0, s_1)$ nor $E({S^\tp}^{-1}_j|\bs \beta_j^\tp, s_0, s_1)$ depends on $\tau^2$s, they can be derived following the same derivation in the EM-CD algorithm.
 On the other hand, $\tau^{2}$, following gamma distributions, is a conjugate prior for the normal variance, and the conditional posterior density of $\tau^{-2}$ is an inverse Gaussian distribution.
@@
 
-\subsubsection{Other Priors}
+\subsection{Supplementary Information: Generalization of BHAM framework}
 
-With the re-parameterization step of the basis function matrix $\bs X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization.
-These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary.
-Here we elaborate the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in BHAM.
-
-A spike-and-slab mixture normal spline prior can be expressed as 
+In the manuscript, we describe the Bayesian hierarchical additive model with the two-part spike-and-slab LASSO prior.
+Nevertheless, the proposed model and algorithm can be easily generalized to accommodate other priors thanks to the reparameterization of smoothing functions.
+To be more specific, we can apply a regularized prior on the linear coefficient, and a group regularized prior for the nonlinear coefficients.
+For example, we can apply the same proposed framework with a spike-and-slab mixture normal prior,
 \begin{align*}
@@
 \end{align*}
-Similar to the spike-and-slab spline prior in Equation (\ref{eq:bham_ssl}), $0 < s_0 < s_1$ are tuning parameters and can be optimized via cross-validation.
-One of the critics received by the spike-and-slab mixture normal prior is that the tails of a normal distribution diminishes to zero too fast, which causes problems when estimating the large effects.
-Distributions with heavier tails can be used as an alternative, for example mixture Student's $t$ distribution with small degree of freedom.
+The algorithm derivation still follows with slight modification, replacing  $l_1$ penalization with $l_2$ penalization.
+The implementation of effect hierarchy would be more challenging for the priors that do rely on a latent indicator for variable selection.
+As a naive solution, readers can consider the linear prior and the nonlinear prior being independent at the cost of bi-level selection accuracy.
 

### 2022-03-24 Manuscript/10-Supplementary.tex
Revise supporting informaiton

--- 2022-02-25 Manuscript/10-Supplementary.tex
@@
 
-\subsection{Supplementary Information 2: EM-IWLS Algorithm for Fitting Bayesian Hierarchical Additive Models}
+\subsection{Supplementary Information 2: EM-Iterative Weighted Least Square Algorithm for BHAM Model Fitting}
 
-Similar to the EM-CD algorithm, the EM-IWLS algorithm is an iterative
-EM-based algorithm where the iterative weighted least squares algorithm
-is used to find the estimate of \(\boldsymbol{\beta}, \phi\) that
-maximizes \(E(Q_1)\). The iterative weighted least squares algorithm was
-originally proposed to fit the classical generalized linear models, and
-generalized to fit some Bayesian hierarchical models.{[}@Gelman2013{]}
-Yi and Ma {[}@Yi2012{]} formulated Student's t-distribution and double
-exponential distribution as hierarchical normal distributions such that
-generalized linear models with shrinkage priors can be easily fitted
-using IWLS in combination with EM algorithm. In this work, we adapt the
-EM-IWLS paradigm to fit BHAM with spike-and-slab spline prior .
+Similar to the EM-CD algorithm, the EM-Iterative Weighted Least Square
+(EM-IWLS) algorithm is an EM-based algorithm where the iterative
+weighted least squares algorithm is used to find the estimate of
+\(\boldsymbol{\beta}, \phi\) that maximizes \(E(Q_1)\). The iterative
+weighted least squares algorithm was originally proposed to fit the
+classical generalized linear models, and generalized to fit some
+Bayesian hierarchical models.{[}@Gelman2013{]} Yi and Ma {[}@Yi2012{]}
+extended the algorithm to fit Bayesian hierarchical models for
+high-dimensional data analysis. Specifically, the authors formulated
+Student's t-distribution and double exponential distribution as
+hierarchical normal distributions such that generalized linear models
+with shrinkage priors can be easily fitted. In this work, we further
+extend the EM-IWLS paradigm to fit the proposed BHAM method. COmpare to
+the EM-CD algorithm, EM-IWLS estimate the variance-covariance matrix of
+the coefficients, providing an opportunity to derive the uncertainty
+quantification of smoothing functions. We will defer this discussion to
+the future work, due to the delicacy of the topic. However, the EM-IWLS
+is not as computationally efficient as EM\_CD, particularly in
+high-dimensional settings.
 
@@
 \end{align*} For the mixture double exponential priors, we can define
-the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\) following
-Equation (\ref{eq:ssl}). The change in the prior formulation in turn
-leads to the change in the log posterior density function, as \(Q_1\)
-needs to account for the hyperprior of \(\tau^2\):
-\begin{equation}\label{eq:Q1_IWLS}
+the scale parameter \(S = (1-\gamma)s_0 + \gamma s_1\). The change in
+the prior formulation in turn leads to the change in the log posterior
+density function, as \(Q_1\) needs to account for the hyperprior of
+\(\tau^2\): \begin{equation}\label{eq:Q1_IWLS}
 Q_1(\boldsymbol{\beta}, \phi) = \log f(\textbf{y}|\boldsymbol{\beta}, \phi) + \sum\limits_{j=1}^p\left[\log f(\beta_j|{\tau}^2_j) + \log f({\tau}^2_j| S_j)+\sum\limits_{k=1}^{K_j} \{\log f(\beta^{*}_{jk}|{\tau^{*}}^2_{jk})+\log f({\tau^*}^2_{jk}| S^*_j)\}\right].
@@
 \(E({S^*}^{-1}_j|\boldsymbol{\beta}_j^*, s_0, s_1)\) depends on
-\(\tau^2\)s, they can be derived using Equation (\ref{eq:exp_scale}). On
-the other hand, \(\tau^{2}\), following gamma distributions, is a
-conjugate prior for the normal variance, and the conditional posterior
-density of \(\tau^{-2}\) is an inverse Gaussian distribution.
-\(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are calculated
-using the closed form equation \begin{align*}
+\(\tau^2\)s, they can be derived following the same derivation in the
+EM-CD algorithm. On the other hand, \(\tau^{2}\), following gamma
+distributions, is a conjugate prior for the normal variance, and the
+conditional posterior density of \(\tau^{-2}\) is an inverse Gaussian
+distribution. \(E({\tau}^{-2}_{j})\) and \(E({\tau^*}^{-2}_{jk})\) are
+calculated using the closed form equation \begin{align*}
  E({\tau}^{-2}_{j}|S_j, \beta_j) ={S_j}^{-1}/|\beta_j| \qquad E({\tau^*}^{-2}_{jk}|S_j^*, \beta^*_{jk})={S_j^*}^{-1}/|\beta^*_{jk}|,
@@
 
+\subsection{Supplementary Information: Generalization of BHAM framework}
+
+In the manuscript, we describe the Bayesian hierarchical additive model
+with the two-part spike-and-slab LASSO prior. Nevertheless, the proposed
+model and algorithm can be easily generalized to accommodate other
+priors thanks to the reparameterization of smoothing functions. To be
+more specific, we can apply a regularized prior on the linear
+coefficient, and a group regularized prior for the nonlinear
+coefficients. For example, we can apply the same proposed framework with
+a spike-and-slab mixture normal prior, \begin{align*}
+  \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\
+  \beta^*_{jk} | \gamma^*_{j},s_0,s_1 &\overset{\text{iid}}{\sim}N(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1), k=1,\dots, K_j.
+\end{align*} The algorithm derivation still follows with slight
+modification, replacing \(l_1\) penalization with \(l_2\) penalization.
+The implementation of effect hierarchy would be more challenging for the
+priors that do rely on a latent indicator for variable selection. As a
+naive solution, readers can consider the linear prior and the nonlinear
+prior being independent at the cost of bi-level selection accuracy.
+
 \clearpage

### 2022-04-01 _targets.R
Correct Typo

@@ -44,7 +44,7 @@ tar_plan(
     hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing
     operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and
     spikeSlabGAM mdoel. mgcv doesn't provide
-    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
+    estimation when number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:bin_auc")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_tab.tex"),
   
@@ -114,7 +114,7 @@ tar_plan(
     Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
     selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
     generalized additive model (SB-GAM), and spikeSlabGAM mdoel. mgcv doesn't provide estimation
-    whe number of parameters exceeds sample size i.e. p = 100, 200.",
+    when number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:gaus") %>% 
     cat(file = "Manuscript/Tabs/sim_gaus_tab.tex"),
   
@@ -183,7 +183,7 @@ tar_plan(
       of comparison include the proposed Bayesian hierarchical additive model (BHAM), the linear LASSO model (LASSO),
       component selection and smoothing operator (COSSO), adaptive COSSO,
       mgcv, sparse Bayesian generalized additive model (SB-GAM), spikeSlabGAM. mgcv doesn't 
-      provide estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
+      provide estimation when number of parameters exceeds sample size i.e. p = 100, 200.",
       label = "tab:time_sim",
       booktabs = TRUE
     ) %>% 
@@ -212,7 +212,7 @@ tar_plan(
     Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
     selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
     generalized additive model (SB-GAM), and spikeSlabGAM mdoel. mgcv doesn't provide
-    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.",
+    estimation when number of parameters exceeds sample size i.e. p = 100, 200.",
[…]

### 2022-04-01 Manuscript/10-Supplementary.Rmd
Correct Typo for the Appendix

@@
 ---
-title: Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data
-subtitle: "Supporting Infromation"
+title: Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis
+subtitle: "Supporting Information"
 author: Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi
+bibliography: bibfile.bib
 output: 
@@
 \subsection{Supplementary Information 1: Marginal Distribution of $\gamma_j^\tp$}
-Given that $\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)$ where $\gamma_{j} | \theta_j \sim Bin(1, \theta_j)$, we can derive the the marginal distribution of $\gamma_{j}^\tp$ with the following manipulation.
+Given that $\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j)$ where $\gamma_{j} | \theta_j \sim Bin(1, \theta_j)$, we can derive the marginal distribution of $\gamma_{j}^\tp$ with the following manipulation.
 \begin{align*}
@@
 In this work, we further extend the EM-IWLS paradigm to fit the proposed BHAM method.
-COmpare to the EM-CD algorithm, EM-IWLS estimate the variance-covariance matrix of the coefficients, providing an opportunity to derive the uncertainty quantification of smoothing functions.
-We will defer this discussion to the future work, due to the delicacy of the topic.
-However, the EM-IWLS is not as computationally efficient as EM_CD, particularly in high-dimensional settings.
+Compare to the EM-CD algorithm, EM-IWLS estimates the variance-covariance matrix of the coefficients, providing an opportunity to derive the uncertainty quantification of smoothing functions.
+We will defer this discussion to future work, due to the delicacy of the topic.
+However, the EM-IWLS is not as computationally efficient as EM-CD, particularly in high-dimensional settings.
 
@@
     a) Based on the current value of $\beta$, calculate the pseudo-data $z_i^{(t)}$ and the pseudo-weights $w_i^{(t)}$
-    b) Update $\bs \beta^{(t)}$ by runing the augmented weighted least squared
+    b) Update $\bs \beta^{(t)}$ by running the augmented weighted least squared
     c) If $\phi$ is present, update $\phi$
@@
 
-\subsection{Supplementary Information: Generalization of BHAM framework}
+\subsection{Supplementary Information 3: Generalization of BHAM framework}
 
@@
 
-\subsection{Supplementary Information 3: Predictive Performance of Linear Simulations}
+\subsection{Supplementary Information 4: Predictive Performance of Linear Simulations}
 \input{Tabs/sim_lnr_gaus_tab.tex}
@@
 
-\subsection{Supplementary Information 4: Variable Selection Performance of Simulations}
+\subsection{Supplementary Information 5: Variable Selection Performance of Simulations}
 \input{Tabs/sim_binom_var_slct_tab.tex}
@@
 \input{Tabs/sim_lnr_binom_var_slct_tab.tex}
+
+\clearpage
+\subsection{References}

### 2022-04-01 _targets.R

@@ -43,7 +43,7 @@ tar_plan(
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
     hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing
     operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and
-    spikeSlabGAM mdoel. mgcv doesn't provide
+    spikeSlabGAM model. mgcv doesn't provide
     estimation when number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:bin_auc")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_tab.tex"),
@@ -85,7 +85,7 @@ tar_plan(
                          averaged over 50 iterations. The models of comparison include the proposed Bayesian
                          hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and
                          smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM),
-                         and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.",
+                         and spikeSlabGAM model. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.",
                          label = "tab:sim_binom_var_select")%>% 
     cat(file = "Manuscript/Tabs/sim_binom_var_slct_tab.tex"),
   # Visual Presentation of Var Selection
@@ -113,7 +113,7 @@ tar_plan(
     Gaussian outcomes over 50 iterations. The models of comparison include the proposed
     Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
     selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
-    generalized additive model (SB-GAM), and spikeSlabGAM mdoel. mgcv doesn't provide estimation
+    generalized additive model (SB-GAM), and spikeSlabGAM model. mgcv doesn't provide estimation
     when number of parameters exceeds sample size i.e. p = 100, 200.",
     label = "tab:gaus") %>% 
     cat(file = "Manuscript/Tabs/sim_gaus_tab.tex"),
@@ -153,7 +153,7 @@ tar_plan(
                          averaged over 50 iterations. The models of comparison include the proposed Bayesian
                          hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection
[…]

### 2022-04-01 Manuscript/SS_GAM_Supporting_Information.tex
Add Supporting information to workflow

--- 2022-03-24 Manuscript/10-Supplementary.tex
@@
 ]{article}
-\title{Spike-and-Slab Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data}
+\title{Spike-and-Slab LASSO Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data Analysis}
 \usepackage{etoolbox}
@@
 \makeatother
-\subtitle{Supporting Infromation}
+\subtitle{Supporting Information}
 \author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
@@
 \hypersetup{
-  pdftitle={Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data},
+  pdftitle={Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis},
   pdfauthor={Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long, Nengjun Yi},
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \ifLuaTeX
@@
 where \(\gamma_{j} | \theta_j \sim Bin(1, \theta_j)\), we can derive the
-the marginal distribution of \(\gamma_{j}^*\) with the following
+marginal distribution of \(\gamma_{j}^*\) with the following
 manipulation. \begin{align*}
@@
 classical generalized linear models, and generalized to fit some
-Bayesian hierarchical models.{[}@Gelman2013{]} Yi and Ma {[}@Yi2012{]}
-extended the algorithm to fit Bayesian hierarchical models for
+Bayesian hierarchical models.(Gelman et al. 2013) Yi and Ma (Yi and Ma
+2012) extended the algorithm to fit Bayesian hierarchical models for
 high-dimensional data analysis. Specifically, the authors formulated
@@
 with shrinkage priors can be easily fitted. In this work, we further
-extend the EM-IWLS paradigm to fit the proposed BHAM method. COmpare to
-the EM-CD algorithm, EM-IWLS estimate the variance-covariance matrix of
+extend the EM-IWLS paradigm to fit the proposed BHAM method. Compare to
+the EM-CD algorithm, EM-IWLS estimates the variance-covariance matrix of
 the coefficients, providing an opportunity to derive the uncertainty
 quantification of smoothing functions. We will defer this discussion to
-the future work, due to the delicacy of the topic. However, the EM-IWLS
-is not as computationally efficient as EM\_CD, particularly in
+future work, due to the delicacy of the topic. However, the EM-IWLS is
+not as computationally efficient as EM-CD, particularly in
 high-dimensional settings.
@@
   \item
-    Update \(\boldsymbol{\beta}^{(t)}\) by runing the augmented weighted
-    least squared
+    Update \(\boldsymbol{\beta}^{(t)}\) by running the augmented
+    weighted least squared
   \item
@@
 
+\clearpage
+\subsection{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
+Gelman, A, JB Carlin, HS Stern, DB Dunson, A Vehtari, and BD Rubin.
+2013. {``Bayesian Data Analysis. 3rd Editio.''}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yi2012}{}}%
+Yi, Nengjun, and Shuangge Ma. 2012. {``{Hierarchical Shrinkage Priors
+and Model Fitting for High-dimensional Generalized Linear Models}.''}
+\emph{Statistical Applications in Genetics and Molecular Biology} 11
+(6). \url{https://doi.org/10.1515/1544-6115.1803}.
+
+\end{CSLReferences}
+
 \end{document}

### 2022-04-01 _targets.R

@@ -545,6 +545,10 @@ tar_plan(
   tar_render(manu, "Manuscript/00-main.Rmd",
              output_file = "SS_GAM.pdf"),
   
+  
+  tar_render(manu_app, "Manuscript/10-Supplementary.Rmd",
+             output_file = "SS_GAM_Supporting_Information.pdf"),
+  
   # R & R Letter --------------------------------------------------------------
   tar_files(RR_path,
             c("Manuscript/R&R/01-reviewer_1.Rmd", "Manuscript/R&R/02-reviewer_2.Rmd",

### 2022-04-02 Manuscript/R&R/00-response_letter.Rmd
correct typo

@@
 
-__Response__: We summarize the major changes of our manuscript as a response to address the comments from the referees:
+__Response__: We summarize the major changes in our manuscript as a response to address the comments from the referees:
 \begin{itemize}
@@
 First of all, moving can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction.
-Secondly, many useful comments raised by the reviewers related to the EM-IWLS algorithm, e.g.
-constructing hypothesis testings and confidence band, requires a more dedicated discussion.
+Secondly, many useful comments raised by the reviewers on the EM-IWLS algorithm, e.g.
+constructing hypothesis testings and confidence band, require a more dedicated discussion.
 We are preparing another manuscript discussing those.
-Lastly, with the added variable selection performance, the manuscript would be length if keeping the EM-IWLS in the main body.
+Lastly, with the added variable selection performance, the manuscript would be too lengthy if keeping the EM-IWLS in the main body.
 The readability of the manuscript would be damaged.
-Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasize the scalability.
-    \item According to Reviewer \#1's and Reviewer \#4's comments, we added linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study.
+Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasizes scalability.
+    \item According to Reviewer \#1's and Reviewer \#4's comments, we added the linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study.
 In addition, we added a new set of simulation scenarios where the functions of predictors are linear.
@@
     \item According to Reviewer \#4's comment, we added the plot showing the estimated functions of predictors for real-data analysis.
-    \item We updated the feature screening procedure in our real-data analysis of Emory Cardiovascular Biobank dataset.
+    \item We updated the feature screening procedure in our real-data analysis of the Emory Cardiovascular Biobank dataset.
 Our previous screening method, fitting univariate GAM, made the "double-dipping" mistake that would overestimate the prediction performance of any downstream prediction model.

### 2022-04-02 Manuscript/R&R/01-reviewer_1.Rmd
correct typo

@@
 We added linear LASSO models to our simulation studies and see that the proposed model predicts as good as linear LASSO models when the underlying signals are linear, and substantially better when the underlying signals are nonlinear.
-This empirically demonstrates the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals.
-The simulation results also match with our expectation.
-The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as bi-level selection problem in the manuscript.
+This empirically demonstrates that the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals.
+The simulation results also match our expectations.
+The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as the bi-level selection problem in the manuscript.
 When the underlying signal is linear, the nonlinear component of the smoothing function would not be included in the model due to the proposed two-part spike-and-slab lasso prior.
@@
 In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model.
-We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.").
+We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.").
 
@@
 Hence, the smoothing functions can be estimated more accurately.
-Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward.
+Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straightforward.
 Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
@@
 
-__Response__: We thank the review for the comments.
-For prediction purpose, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing.
+__Response__: We thank the reviewer for the comments.
+For prediction purposes, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing.
 The two-part spike-and-slab LASSO prior automatically detects if the underlying signal is linear or complex.
-When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic spline (implemented in the package) is used to estimate the complex signal.
-As a supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see response to the Comment 1).
+When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic splines (implemented in the package) are used to estimate the complex signal.
+As supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see the response to Comment 1).
 
@@
 
-In conclusion, the proposed model do not addition detection for prediction purpose, and have different utility for bi-level selection.
+In conclusion, the proposed model does not require addition detection when used for prediction, and has different utility for bi-level selection.
 

### 2022-04-02 Manuscript/R&R/03-reviewer_4.Rmd
correct typo

@@
 We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
-This prior formulation is motivated by effect hierarchy discussed in Chipman (2004).
+This prior formulation is motivated by the effect hierarchy discussed in Chipman (2004).
 Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
@@
 The $\gamma$ can be analytically integrated out of the density function and requires minimum change of the proposed algorithm.
-In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
+In other words, the proposed model remains feasible for high-dimensional data analysis and retains the claimed computational advantages.
 \bigskip
@@
 
-__Response__: We thank the review for the comments.
-We made change to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear.
+__Response__: We thank the reviewer for the comments.
+We made changes to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear.
 The suggested model spikeSlabGAM and linear Lasso model as the benchmark were added to the models of comparison.
@@
 
-__Response__: We thank the review for the comment.
+__Response__: We thank the reviewer for the comment.
 In the update manuscript, we add the suggested model, spikeSlabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared.
@@
 gene network.
-We hope to establish the consistency in the word choice to avoid future confusion.
+We hope to establish consistency in the word choice to avoid future confusion.
 

### 2022-04-02 Manuscript/Tabs/sim_gaus_tab.tex
Correct typos in the tables

--- 2022-03-21 Manuscript/Tabs/sim_gaus_tab.tex
@@
 \caption{The average and standard deviation of the out-of-sample $R^2$ measure for
-    Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
-    COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
-    whe number of parameters exceeds sample size i.e. p = 100, 200.} 
+    Gaussian outcomes over 50 iterations. The models of comparison include the proposed
+    Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
+    selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
+    generalized additive model (SB-GAM), and spikeSlabGAM model. mgcv doesn't provide estimation
+    when number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:gaus}

### 2022-04-02 Manuscript/Tabs/sim_lnr_binom_tab.tex
Correct typos in the tables

--- 2022-03-21 Manuscript/Tabs/sim_lnr_binom_tab.tex
@@
 \caption{The average and standard deviation of the out-of-sample area under the curve measures
-    for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
-    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
-    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 
+    for binomial outcomes over 50 iterations. The models of comparison include the proposed
+    Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
+    selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
+    generalized additive model (SB-GAM), and spikeSlabGAM model. mgcv doesn't provide
+    estimation when number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:lnr_bin_auc}

### 2022-04-02 Manuscript/Tabs/sim_lnr_binom_var_slct_tab.tex
Correct typos in the tables

@@
                          and smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized
-                         additive model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$
+                         additive model (SB-GAM), and spikeSlabGAM model. MCC is ill-defined when $p=4$
                          simulation (no true negative), and hence omitted for all methods.} 

### 2022-04-02 Manuscript/Tabs/sim_lnr_gaus_tab.tex
Correct typos in the tables

--- 2022-03-21 Manuscript/Tabs/sim_lnr_gaus_tab.tex
@@
 \caption{The average and standard deviation of the out-of-sample $R^2$ measure for
-    Gaussian outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO), adaptive
-    COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide estimation
-    whe number of parameters exceeds sample size i.e. p = 100, 200.} 
+    Gaussian outcomes over 50 iterations. The models of comparison include the proposed
+    Bayesian hierarchical additive model (BHAM), linear LASSO model (LASSO), component
+    selection and smoothing operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian
+    generalized additive model (SB-GAM), and spikeSlabGAM model. mgcv doesn't provide estimation
+    when number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:lnr_gaus}

### 2022-04-02 Manuscript/Tabs/sim_lnr_gaus_var_slct_tab.tex
Correct typos in the tables

@@
                          linear LASSO model (LASSO), component selection and smoothing operator (COSSO),
-                         adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM mdoel.
+                         adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM), and spikeSlabGAM model.
                          MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.} 

### 2022-05-15 Manuscript/00-main.Rmd

@@
 Functional selection performance suggests trade-offs exist regarding the effect hierarchy assumption."
-keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-Coordinate Decent; Scalablility; Predictive Modeling
+keywords: Spike-and-Slab Priors; High-Dimensional Data; Generalized Additive Models; EM-Coordinate Decsent; Scalablility; Predictive Modeling
 # bibliography: bibfile.bib
@@
   xtable(
-    caption = "Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
+    caption = "Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.",
     label = "tab:ECB_res"
@@
     xtable(
-    caption = "Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
+    caption = "Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).",
     label = "tab:WLM_res"

### 2022-05-15 Manuscript/01-Intro.Rmd

@@
 
-Many modern biomedical research, e.g.
-sequencing data analysis, electric health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
+Much modern biomedical research, e.g., sequencing data analysis and electronic health record data analysis, require special treatment of high-dimensionality, commonly known as $p >> n$ problem.
 There is extensive literature on high-dimensional linear models via penalized models or Bayesian hierarchical models, see Mallick and Yi [@Mallick2013] for review.
@@
 In classical statistical modeling, many strategies and models are proposed to relax the linearity assumption with various degrees of complexity.
-For example, variable categorization is a simple and common practice in epidemiology, but suffers from power and interpretation issues.
+For example, variable categorization is a simple and common practice in epidemiology but suffers from power and interpretation issues.
 More complex models to address nonlinear effects include random forest and other so-called "black box" models [@Breiman2001].
@@
 
-\bg{For their easy interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in the context of prediction and variable selection}.
+\bg{For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection}.
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
@@
 
-There exists some proposals to generalize the classical GAM to accommodate high-dimensional applications.
+There exist some proposals to generalize the classical GAM to accommodate high-dimensional applications.
 The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smoothing functions.
 Ravikumar et al.
-[@Ravikumar2009] extended the grouped LASSO [@Yuan2006] to additive models (AMs); Huang et al.
-[@Huang2010] further developed adaptive grouped LASSO for additive models; Wang et al.
-[@Wang2007] and Xue [@Xue2009] respectively applied grouped SCAD penalty [@Fan2001] to additive models.
-Recently Bayesian hierarchical models are also used in the context of high-dimensional additive models, particularly within the spike-and-slab literature.
-Various group spike-and-slab priors [@Xu2015; @Yang2020] combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms are proposed, where the application on AMs are treated as a special case.
-Bai et al.
-[@Bai2020] was the first to apply group spike-and-slab LASSO prior to Gaussian AMs using a fast optimization algorithm, and further generalized the framework to GAMs [@Bai2021].
-Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large number of knots are used.
-[@Scheipl2013] In addition, these methods adapt an 'all-in-all-out' strategy, i.e.
+[@Ravikumar2009] extended the group LASSO [@Yuan2006] to additive models (AMs); Huang et al.
+[@Huang2010] further developed adaptive group LASSO for additive models; Wang et al.
+[@Wang2007] and Xue [@Xue2009] respectively applied group SCAD penalty [@Fan2001] to additive models.
+Bayesian hierarchical models are also used in the context of high-dimensional additive models, particularly within the spike-and-slab literature.
+Various group spike-and-slab priors [@Xu2015; @Yang2020] combining with computationally intensive Markov chain Monte Carlo (MCMC) algorithms are proposed, where the application on AMs is treated as a special case.
+Bai and co-authors[@Bai2020] were the first to apply group spike-and-slab LASSO prior to Gaussian AMs using a fast optimization algorithm and further generalized the framework to GAMs [@Bai2021].
+Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large numbers of knots are used.
+[@Scheipl2013] In addition, these methods adopt an 'all-in-all-out' strategy, i.e.
 either including or excluding the variable completely, rendering no space for bi-level selection.
 Scheipl et al.
-[@Scheipl2012] proposed a spike-and-slab structure prior that address the bi-level selection.
-But the model fitting relies on computational intensive MCMC algorithms and creates scalability concern.
-It would be of special interest to develop a fast, flexible and accurate generalized additive model framework.
+[@Scheipl2012] proposed a spike-and-slab structure prior that addresses the bi-level selection.
+But the model fitting relies on computationally intensive MCMC algorithms and creates scalability concerns.
+Developing a fast, flexible and accurate generalized additive model framework would be of special interest.
 
@@
 Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
-We then impose a new two-part spike-and-slab LASSO prior to address the sparsity of the signal.
-In addition, a scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm are developed.
+We then impose a new two-part spike-and-slab LASSO prior to address the signal sparsity.
+In addition, a scalable optimization-based algorithm, EM-Coordinate Descent (EM-CD) algorithm, is developed.
 While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection.
-Particularly, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-The proposed model is implemented in an publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
+Notably, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
+The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
 
-The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e.
-the spike-and-slab GAM [@Scheipl2012] and the SB-GAM [@Bai2021] in three ways.
-First of all, the proposed spike-and-slab spline prior is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
+The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e., the spike-and-slab GAM [@Scheipl2012] and the sparse Bayesian GAM (SB-GAM) [@Bai2021], in three ways.
+Firstly, the proposed prior for smoothing functions is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
 Spike-and-slab LASSO priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms.
@@
 Secondly, SB-GAM uses a group spike-and-slab LASSO prior with an EM-CD algorithm to fit the model.
-While both methods use the combination of expectation maximization algorithm and coordinate descent algorithm, there are subtle difference in the implementation due to the difference in prior specification.
+While both methods use the combination of expectation maximization algorithm and coordinate descent algorithm, there are subtle differences in the implementation due to the difference in prior specification.
 The proposed model sets up independent priors among basis coefficients after the reparameterization step, which provides some advantage in computation.
-Last but not least, the proposed model addresses the incapability of bi-level selection in SB-GAM.
+Lastly, the proposed model addresses the incapability of bi-level selection in SB-GAM.
 

### 2022-05-15 Manuscript/02-Method.Rmd

--- 2022-03-21 Manuscript/02-Method.Rmd
@@
 Marra and  Wood [@Marra2011] proposed a reparameterization procedure \bg{to factor the smoothing penalty into the design matrix of each smoothing function}.
-Given the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+Given that the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
 To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
@@
 We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix ${\bs X}^\repa = \bs X \bs U \equiv \begin{bmatrix} X^0 : \bs X^{\tp} \end{bmatrix}$.
-An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions.
+An additional scaling step is imposed on $\bs X^{\tp}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive a uniform penalty on each of its dimensions.
 With slight abuse of the notation, we drop the superscript $^\repa$ and denote $\bs X_j \equiv \begin{bmatrix} X_j^0 : \bs X_j^{\tp} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the reparameterization.
@@
 To summarize, the reparameterization step provides three benefits.
-First of all, the reparameterization integrates the smoothing penalty matrix into the design matrix, and encourages models to properly smooth the nonlinear function when sparsity penalty exists.
+Firstly, the reparameterization integrates the smoothing penalty matrix into the design matrix, and encourages models to properly smooth the nonlinear function when sparsity penalty exists.
 Secondly, the eigendecomposition of the smoothing penalty matrix allows the isolation of the linear space from the nonlinear space, improving the feasibility of bi-level functional selection.
-Last but not least, the eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
+Lastly, the eigendecomposition facilitates the construction of an orthonormal design matrix, which makes imposing independent priors on the coefficients possible.
 This reduces the computational complexity compared to using a multivariate priors, and improve the generalizability of the framework to be compatible with other choices of priors.
 
-\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Fucntions}\label{sec:method_prior}
+\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Functions}\label{sec:method_prior}
 
-The family of spike-and-slab regression models is one of most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
+The family of spike-and-slab regression models is one of the most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
 \bg{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
@@
 }
-The latent binary variable $\gamma \in \{0,1\}$ indicates whether a variable $x$ is included in the model, while the dispersion parameters $s_0$ and $s_1$ controls the shrinkage of the coefficient.
+The latent binary variable $\gamma \in \{0,1\}$ indicates whether a variable $x$ is included in the model, while the dispersion parameters $s_0$ and $s_1$ control the shrinkage of the coefficient.
 Given that both double exponential distributions have a mean of 0 and the latent indicator $\gamma$ can only take the value of 0 or 1, the mixture double exponential distribution can be formulated as one single double exponential density,
@@
 First of all, the SSL prior provides a locally adaptive shrinkage when estimating the coefficients.
-Secondly, the SSL prior encourages a sparse solution, making variable selection straight forward.
+Secondly, the SSL prior encourages a sparse solution, making variable selection straightforward.
 Thirdly, the SSL prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
 We defer to Bai et al.
-\cite{Bai2021Review} for an detailed discussion.}
+\cite{Bai2021Review} for a detailed discussion.}
 
@@
 Given the reparameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
-Specifically, the linear space coefficient has a SSL prior and the nonlinear space coefficients shares a group SSL prior,
+Specifically, the linear space coefficient has an SSL prior and the nonlinear space coefficients share a group SSL prior,
 \begin{align}\label{eq:bham_ssl}
@@
 Effect hierarchy refers to the principle that "lower-order effects are more likely to be active than higher-order effects" defined by Chipman\cite{chipman2006prior}.
-To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$, while both latent indicators $\gamma_j and \gamma^\tp_j$ follow a Bernoulli distribution.
+To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$, while both latent indicators $\gamma_j$ and $\gamma^\tp_j$ follow a Bernoulli distribution.
 While the probability of including the linear effect is $\theta_j$, the probability of including the nonlinear effect is $\bg{\gamma_{j}}\theta_j$.
@@
 \bg{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For the computational convenience, we analytically integrate $\theta_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$.
+For computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
 }
@@
 $$
-The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience.
+The beta distribution is a conjugate prior to the binomial distribution and hence provides some computation convenience.
 <!-- For simplicity, we focus on a special case of beta distribution, uniform (0,1), i.e. $a = 1, b = 1$. -->
 \bg{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
-In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particular in pair with the smoothing penalty implicitly addressed via the reparameterization.
-Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different level of sparsity.}
+In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
+Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}
 <!-- This informaiton borrowing helps the model to find adequate shrinkage for the linear coefficient and coefficients of the nonlinear part.
@@
 <!-- When the variable has large effect in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases.  -->
-Hereafter, we refer Bayesian hierarchical generalized additive models with the two-part spike-and-slab LASSO prior as the BHAM, and visually presented in Figure \ref{fig:SSprior}.
+Hereafter, we refer to the Bayesian hierarchical generalized additive models with the two-part spike-and-slab LASSO prior as BHAM, and visually presented in Figure \ref{fig:SSprior}.
 
@@
 
-\bg{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
-Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
+\bg{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computationaly prohibited and not feasible for high-dimensional data.
+Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium-sized data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
 Hence, we consider the optimization algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
@@
 extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
-Similar to the EMVS algorithm \cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is based on the the expectation-maximization (EM) algorithm, integrating Coordinate Descent algorithm in each iterative step to find the posterior mode.
+Similar to the EMVS algorithm \cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is based on the expectation-maximization (EM) algorithm, integrating the Coordinate Descent algorithm in each iterative step to find the posterior mode.
 The EM-CD algorithm has been well adapted in generalized linear models \cite{Tang2017a}, Cox proportional hazards models \cite{Tang2017}, and their grouped counterparts \cite{Tang2018, Tang2019}.
@@
 
-For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words the "missing" data in the EM context.
+For BHAMs, we define the parameters of interest as $\Theta = \{\bs \beta, \bs \theta, \phi\}$ and consider the latent binary indicators $\bs \gamma$ as nuisance parameters of the model, in other words, the "missing" data in the EM context.
 Our objective is to find the parameters $\Theta$ that maximize the posterior density function, or equivalently the logarithm of the density function,
@@
 With the proposed two-part spike-and-slab LASSO prior, $Q_1$ can be treated as penalized likelihood function and maximization of $E(Q_1)$ can be solved via the Coordinate Descent algorithm in each iteration.
-Coordinate descent is an optimization algorithm that offers extreme computational advantage, and famous for its application in optimizing the $l_1$ penalized likelihood function.
-Maximization of $E(Q_2)$ can be solved via closed form equations following the beta-binomial conjugate relationship.
+Coordinate descent is an optimization algorithm that offers extreme computational advantages, and is famous for its application in optimizing the $l_1$ penalized likelihood function.
+Maximization of $E(Q_2)$ can be solved via closed-form equations following the beta-binomial conjugate relationship.
 
@@
 Independently, the remaining parameters of interest $\bs \theta$ can be updated by maximizing $E(Q_2)$.
-As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed form equation,
+As the beta distribution is a conjugate prior for Bernoulli distribution, $\bs \theta$ can be easily updated with a closed-form equation,
 \begin{equation}\label{eq:update_theta}
@@
 Our proposed model, BHAM, requires two preset scale parameters ($s_0$, $s_1$).
-Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding a criteria of preference.
-This would be achieved by constructing a two-dimensional grid, consists of different pairs of ($s_0$, $s_1$) value.
+Hence, we need to find the optimal values for the scale parameters such that the model reaches its best prediction performance regarding criteria of preference.
+This would be achieved by constructing a two-dimensional grid, consisting of different ($s_0$, $s_1$) pairs.
 However, previous research suggests the value of slab scale $s_1$ has less impact on the final model and is recommended to be set as a generally large value, e.g.
@@
 deviance \bg{(defined as model log-likelihood times -2, $-2\log f(\bs y|\bs{\hat\beta}, \hat\phi)$)}, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
-The procedure is similar to the LASSO implementation in the widely used R package `glmnet`, which quickly fits LASSO models over a list of values of regularization parameters $\lambda$ and gives a sequence of models for users to choose from.
+The procedure is similar to the LASSO implementation in the widely used R package \texttt{glmnet}, which quickly fits LASSO models over a list of candidate values of $\lambda$ and gives a sequence of models for users to choose from.

### 2022-05-15 Manuscript/03-Simulation.Rmd

--- 2022-03-21 Manuscript/03-Simulation.Rmd
@@
 
-In this section, we compare the performance of the proposed model to \bg{six alternative models: linear LASSO models,} component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM [@Bai2021].
-\bg{We use linear LASSO model as the benchmark, examing the performance when linearity assumption doesn't hold}.
+In this section, we compare the performance of the proposed model to \bg{six alternative models: linear LASSO models,} component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred to as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM [@Bai2021].
+\bg{We use linear LASSO models as the benchmark, examining the performance when the linearity assumption doesn't hold}.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
-Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component are different for extra flexibility.
+Adaptive COSSO improved upon COSSO by using adaptive weight for penalties such that the penalty of each functional component is different for extra flexibility.
 \textit{mgcv} is one of the most popular models for nonlinear effect interpolation and prediction.
-Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses a MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab LASSO GAM.
+Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses an MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab LASSO GAM.
 We implement linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
@@
 We control the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
-We use 5-fold CV with the default selection criteria to select the final model for linear LASSO model, COSSO models, SB-GAM and the proposed model.
-20 default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed model which allow user-specification of tuning candidates.
+We use a 5-fold CV with the default selection criteria to select the final model for linear LASSO model, COSSO models, SB-GAM and the proposed model.
+Twenty default candidates of tuning parameters ($s_0$ in BHAM, $\lambda_0$ in SB-GAM) are examined for SB-GAM and the proposed model that allow user specification of tuning candidates.
 All computation was conducted on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R 3.6.2 [@R].
@@
 [@Ravikumar2009].
-However, we exclude these methods from the current simulation study because of their demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].
+However, we exclude these methods from the current simulation study because they demonstrated inferior predictive performance compared to \textit{mgcv} [@Scheipl2013].
 
@@
 for Gaussian and binomial outcomes.
-Gaussian outcomes requires specification of dispersion, where we set the dispersion parameter to be 1.
+Gaussian outcomes require specification of dispersion, where we set the dispersion parameter to be 1.
 In this data generating process, we have $x_1, x_2, x_3, x_4$ as the active predictors, while the rest predictors are inactive, i.e.
 $f_j(x_j) = 0$ for $j = 4, \dots, p$.
-Another set of independent sample of size $n_{test}=1000$, are created following the same data generating process, serving as the testing data.
+Another set of independent sample of size $n_{test}=1000$, is created following the same data generating process, serving as the testing data.
 We generate 50 independent pairs of training and testing datasets to evaluate the prediction and variable selection performance of the chosen models, where training datasets are used to fit the models and testing datasets are used to calculate metrics of interest.
@@
 \end{align*}
-where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positive respectively.
-For the methods that don't automatically achieve variable selection, we set alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictors are selected in the models.}
+where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positives respectively.
+For the methods that don't automatically achieve variable selection, we set the alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictor are selected in the models.}
 
@@
 \bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
-Specifically, BHAM shows greate improvement over COSSO methods, resulting a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) imporvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
-The improvement over spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
+Specifically, BHAM shows great improvement over COSSO methods, resulting in a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) improvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
+The improvement over the spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
 When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear LASSO model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
+As previously hypothesized, the linear LASSO model predicts less accurate than other flexible models across all scenarios; mgcv performs extremely well in low-dimensional cases ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size approaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has an extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
 
-\bg{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
-The proposed model, BHAM, has similar performance as the linear LASSO model regardless of the distribution.
-This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
-spikeSlabGAM have similar prediction performance to BHAM.
+\bg{We also examine the prediction performance when the functional form of predictors is linear, see Supporting Information Table S1, and S2.
+The proposed model, BHAM, has a similar performance as the linear LASSO model regardless of the distribution.
+This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless of the underlying functional form of predictors.
+spikeSlabGAM has a similar prediction performance to BHAM.
 Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
@@
 \bg{\subsubsection{Variable Selection Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{sec:gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
-The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
+\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimensions and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that the model selected a small subset among all the active predictions.
 In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
@@
 Upon further investigation, we discover it's because the generic sequence of $s_0$ used to tune the model doesn't contain the optimal value.
-Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yield a high MCC.
+Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yields a high MCC.
 The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations match with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
+The variable selection performance for linear simulations matches with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
 
@@
 Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
-Specifically, spikeSlabGAM have smaller probability to include nonlinear component in the model than BHAM.
-However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the function forms that include only linear or nonlinear component, e.g.
-function forms for $x_3$ and $x_4$.
-However, spikeSlabGAM inclines to ignore the variable that have more complex function forms, e.g.
+Specifically, spikeSlabGAM have a smaller probability to include the nonlinear component in the model than BHAM.
+However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the functional forms that include only linear or nonlinear component, e.g.
+functional forms for $x_3$ and $x_4$.
+However, spikeSlabGAM inclines to ignore variables that have more complex function forms, e.g.
 function forms for $x_1$ and $x_2$.
@@
 This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
-We defer an in-depth discussion to the Section \ref{sec:concl}.}
+We defer an in-depth discussion to Section \ref{sec:concl}.}
 

### 2022-05-15 Manuscript/04-Real_Data.Rmd

@@
 
-In this section, we apply the proposed models BHAM to analyze two published metabolomics datasets where the outcomes are binary and continuous respectively.
+In this section, we apply the proposed model BHAM to analyze two published metabolomics datasets where the outcomes are binary and continuous respectively.
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}).
@@
 \label{sec:ECB}
-We use the proposed models BHAM to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
+We use the proposed model BHAM to analyze a metabolic dataset from a recently published research [@Mehta2020] studying plasma metabolomic profile on the three-year all-cause mortality among patients undergoing cardiac catheterization.
 The dataset is publicly available via \textit{Dryad} [@Mehta2020_data].
 It contains in total of 776 subjects from two cohorts.
-As there is a large number of non-overlapping features among the two cohorts, we use the cohort with larger sample size (N=454).
+As there is a large number of non-overlapping features among the two cohorts, we use the cohort with a larger sample size (N=454).
 There are initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
-Hence, we choose the the top 200 features \bg{with largest variance}.
+Hence, we choose the top 200 features \bg{with the largest variance}.
 We use 5-knot spline additive models for binary outcome using two different models, the proposed BHAM and the SB-GAM.
 10-Fold CV are used to choose the optimal tuning parameters of each framework with respect to the default selection criterion implemented in the software.
-Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassification error defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
+Out-of-bag samples are used for prediction performance evaluation, where deviance, AUC, Brier score, defined as $\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2$, and misclassification error, defined as $\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)$ are calculated.
 BHAM obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).
@@
 \subsection{Weight Loss Maintenance Cohort}
-We use the proposed models BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
+We use the proposed model BHAM to analyze metabolomics data from a recently published study [@Bihlmeyer2021] on the association between metabolic biomarkers and weight loss, where the dataset is publicly available [@Bihlmeyer2021_data].
 In this analysis, we primarily focus on the analysis of one of the three studies included, weight loss maintenance cohort [@Svetkey2008], due to the drastically different intervention effects.
@@
 Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team \cite{Bihlmeyer2021}}.
-The outcome of interest are standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
+The outcome of interest is standardized percent change in insulin resistance, and hence modeled using a Gaussian model.
 After removing missing datapoints and addressing outliers in the data, there are $p$=`r nrow(WLM_train_dat)` features remaining in the analysis.

### 2022-05-15 Manuscript/05-Discussion.Rmd

@@
 \label{sec:concl}
-\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling.
-In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
+\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeing.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing functions and derive a scalable EM-CD algorithm for model fitting.
 The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
 Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
-In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear component.
+In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear components.
 In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
@@
 
-The proposed model shares many commonality with the SB-GAM [@Bai2021], independently developed around the same time of the proposed work.
+The proposed model shares many commonalities with the SB-GAM [@Bai2021], which is independently developed around the same time as the proposed work.
 Both frameworks emphasize computational efficiency by deploying group spike-and-slab LASSO type priors and optimization-based scalable algorithms.
@@
 \bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.
+Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.
 Not focused in this manuscript, the generalization is described in the Supporting Information.}
 
-\bg{During designing and analyzing the simulation study, we made couple interesting observations.
-First of all, variable selection is a delicate topic in the context predictive modelling.
-When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example LASSO and LASSO-based models.
+\bg{During designing and analyzing the simulation study, we made couple of interesting observations.
+First of all, variable selection is a delicate topic in the context of predictive modeling.
+When prediction performance is used to tune a model, the model could possibly include noise variables in models, for example LASSO and LASSO-based models.
 \cite{Wu2019} Moreover, bi-level selection is a more complex problem than variable selection.
-The complexity reflects on the validity of the effect hierarchy principle.
-While most functional forms follow that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
+The complexity shows on the validity of the effect hierarchy principle.
+While most functional forms follow that the linear component exists in the nonlinear function, there are functions that don't follow it, e.g.
 $x^2$.
-The proposed prior and spikeSlabGAM employ different structure: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
+The proposed prior and spikeSlabGAM employ different structures: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
 The different prior setups lead to trade-offs for the purpose of bi-level selection.
-We recommend to use more judgement in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
-Secondly, we find the performance of the proposed model are more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
-Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend to fine-tune the model with granular sequence of $s_0$ for performance improvement.}
+We recommend to use more judgment in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
+Secondly, we find the performance of the proposed model is more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
+Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend fine-tuning the model with a granular sequence of $s_0$ for performance improvement.}
 
@@
 
-Our future efforts direct to \bg{uncertainty inference of the proposed model}, survival analysis and integrative analysis.
+Our future efforts are direct to \bg{uncertainty inference of the proposed model}, survival analysis and integrative analysis.
 \bg{Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
 We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
-Instead of the Coordinate Descent algorithm, we use Iterative Weighted Least Square algorithm in the EM procedure.
+Instead of the Coordinate Descent algorithm, we use the Iterative Weighted Least Square algorithm in the EM procedure.
 The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
-Due to the space limit, technical details will be explained in a future manuscript.} While the proposed model addresses a great deal of analytic problem, analyzing the time-to-event outcome remains unsolved.
-An naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].
+Due to the space limit, technical details will be explained in a future manuscript.} While the proposed model addresses a great deal of analytic problems, analyzing the time-to-event outcome remains unsolved.
+A naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].
 However, it would be more efficient to directly fit Cox models via penalized pseudo likelihood function [@Simon2011].
-Meanwhile, with growing understanding of biological structure within -omics field, it is appealing to integrate external biology information in the modeling process.
-The main motivation for integrative models is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purpose.
+Meanwhile, with the growing understanding of biological structure within -omics field, it is appealing to integrate external biology information in the modeling process.
+The main motivation for integrative models is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purposes.
 Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.
-The similar strategy has been used in Ferrari and Dunson [@Ferrari2020].
+A similar strategy has been used in Ferrari and Dunson [@Ferrari2020].

### 2022-05-15 Manuscript/R&R/01-reviewer_1.Rmd

@@
 __Response__: Thank you for the comment.
-The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms.
+The purpose of mentioning several other priors in the previous manuscript is to describe the generalization of the proposed algorithms.
 In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model.
-We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.").
+We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily accommodate other families of priors and smoothing functions if desired.").
 
@@
 When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic splines (implemented in the package) are used to estimate the complex signal.
-As supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see the response to Comment 1).
+As supporting evidence, our added simulation demonstrates that the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see the response to Comment 1).
 

### 2022-05-15 Manuscript/R&R/02-reviewer_2.Rmd

@@
 The purpose of the reparameterization step is to factor the smoothing penalty matrix in the design matrix, such that the smoothing penalty can be addressed implicitly.
-This contrasts to some previous methods that do not consider smoothing penalty at all.
+This contrasts with some previous methods that do not consider smoothing penalties at all.
 With the reparameterization, the smoothing penalty changes from $\lambda \bs \beta^T \bs S \bs \beta$ to $\lambda \bs {\beta^*}^T \bs {\beta^*}$, where the $\bs \beta$ and $\bs \beta^*$ are the coefficients of the smoothing function before and after the reparameterization respectively.
 However, the smoothing parameter $\lambda$, equivalently the scale parameter $s$ in the proposed model, needs to be estimated/tuned.
-The proposed two-part spike-and-slab LASSO prior provides the the estimation of the smoothing parameter: when the nonlinear effect does not exists, a large shrinkage is applied to the coefficients via the spike density such that coefficients shrink to 0s; when then nonlinear exists, a small shrinkage is applied via the slab density such that the smoothing function can be estimated appropriately.
-Moreover, the local adaptivity of he spike-and-slab LASSO prior allows the shrinkage to adapt to the data, compared to previous methods that impose an uniform shrinkage.
+The proposed two-part spike-and-slab LASSO prior provides the estimation of the smoothing parameter: when the nonlinear effect does not exist, a large shrinkage is applied to the coefficients via the spike density such that coefficients shrink to 0s; when then nonlinear exists, a small shrinkage is applied via the slab density such that the smoothing function can be estimated appropriately.
+Moreover, the local adaptivity of the spike-and-slab LASSO prior allows the shrinkage to adapt to the data, compared to previous methods that impose a uniform shrinkage.
 
-We clarified the utility of the reparameterization step and emphasized how the smoothness of the additive function are appropriately modeled (see P3).
+We clarified the utility of the reparameterization step and emphasized how the smoothness of the additive function is appropriately modeled (see P3).
 
@@
 We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously.
-This prior formulation is motivated by effect hierarchy discussed in Chipman (2004).
+This prior formulation is motivated by the effect hierarchy discussed in Chipman (2004).
 Now, the inclusion of the nonlinear effects is dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection.
 The modification of the prior doesn't complicate the model fitting algorithm.
-The $\gamma$ can be analytically integrated out of the density function of $\gamma^*$ and requires minimum change of the proposed algorithms.
+The $\gamma$ can be analytically integrated out of the density function of $\gamma^*$ and requires a minimum change of the proposed algorithms.
 The updated model still shows performance advantages as before.

### 2022-05-15 Manuscript/SS_GAM.tex

@@
 \keywords{Spike-and-Slab Priors; High-Dimensional Data; Generalized
-Additive Models; EM-Coordinate Decent; Scalablility; Predictive
+Additive Models; EM-Coordinate Decsent; Scalablility; Predictive
 Modeling}
@@
 
-Many modern biomedical research, e.g.~sequencing data analysis, electric
-health record data analysis, require special treatment of
+Much modern biomedical research, e.g., sequencing data analysis and
+electronic health record data analysis, require special treatment of
 high-dimensionality, commonly known as \(p >> n\) problem. There is
@@
 example, variable categorization is a simple and common practice in
-epidemiology, but suffers from power and interpretation issues. More
+epidemiology but suffers from power and interpretation issues. More
 complex models to address nonlinear effects include random forest and
@@
 
-\textcolor{red}{For their easy interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in the context of prediction and variable selection}.
+\textcolor{red}{For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection}.
 Among those, generalized additive models (GAMs), proposed in the seminal
@@
 
-There exists some proposals to generalize the classical GAM to
+There exist some proposals to generalize the classical GAM to
 accommodate high-dimensional applications. The regularized models,
@@
 by accounting for the structure introduced when expanding smoothing
-functions. Ravikumar et al. \citep{Ravikumar2009} extended the grouped
+functions. Ravikumar et al. \citep{Ravikumar2009} extended the group
 LASSO \citep{Yuan2006} to additive models (AMs); Huang et al.
-\citep{Huang2010} further developed adaptive grouped LASSO for additive
+\citep{Huang2010} further developed adaptive group LASSO for additive
 models; Wang et al. \citep{Wang2007} and Xue \citep{Xue2009}
-respectively applied grouped SCAD penalty \citep{Fan2001} to additive
-models. Recently Bayesian hierarchical models are also used in the
-context of high-dimensional additive models, particularly within the
-spike-and-slab literature. Various group spike-and-slab priors
-\citep{Xu2015, Yang2020} combining with computationally intensive Markov
-chain Monte Carlo (MCMC) algorithms are proposed, where the application
-on AMs are treated as a special case. Bai et al. \citep{Bai2020} was the
-first to apply group spike-and-slab LASSO prior to Gaussian AMs using a
-fast optimization algorithm, and further generalized the framework to
-GAMs \citep{Bai2021}. Focus on addressing the sparsity, these methods
-can excessively penalize the bases of a smoothing function and produce
+respectively applied group SCAD penalty \citep{Fan2001} to additive
+models. Bayesian hierarchical models are also used in the context of
+high-dimensional additive models, particularly within the spike-and-slab
+literature. Various group spike-and-slab priors \citep{Xu2015, Yang2020}
+combining with computationally intensive Markov chain Monte Carlo (MCMC)
+algorithms are proposed, where the application on AMs is treated as a
+special case. Bai and co-authors\citep{Bai2020} were the first to apply
+group spike-and-slab LASSO prior to Gaussian AMs using a fast
+optimization algorithm and further generalized the framework to GAMs
+\citep{Bai2021}. Focus on addressing the sparsity, these methods can
+excessively penalize the bases of a smoothing function and produce
 inaccurate predictions, particularly when complex signals are assumed
-and large number of knots are used. \citep{Scheipl2013} In addition,
-these methods adapt an `all-in-all-out' strategy, i.e.~either including
+and large numbers of knots are used. \citep{Scheipl2013} In addition,
+these methods adopt an `all-in-all-out' strategy, i.e.~either including
 or excluding the variable completely, rendering no space for bi-level
 selection. Scheipl et al. \citep{Scheipl2012} proposed a spike-and-slab
-structure prior that address the bi-level selection. But the model
-fitting relies on computational intensive MCMC algorithms and creates
-scalability concern. It would be of special interest to develop a fast,
-flexible and accurate generalized additive model framework.
+structure prior that addresses the bi-level selection. But the model
+fitting relies on computationally intensive MCMC algorithms and creates
+scalability concerns. Developing a fast, flexible and accurate
+generalized additive model framework would be of special interest.
 
@@
 Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
-We then impose a new two-part spike-and-slab LASSO prior to address the sparsity of the signal.
-In addition, a scalable optimization-based algorithms, EM-Coordinate Descent (EM-CD) algorithm are developed.
+We then impose a new two-part spike-and-slab LASSO prior to address the signal sparsity.
+In addition, a scalable optimization-based algorithm, EM-Coordinate Descent (EM-CD) algorithm, is developed.
 While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection.
-Particularly, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-The proposed model is implemented in an publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
+Notably, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
+The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
 
 The proposed framework, BHAM, differs from previous spike-and-slab based
-GAMs, i.e.~the spike-and-slab GAM \citep{Scheipl2012} and the SB-GAM
-\citep{Bai2021} in three ways. First of all, the proposed spike-and-slab
-spline prior is a spike-and-slab LASSO type prior using independent
-mixture double exponential distribution, compared to spike-and-slab GAM
-that uses normal-mixture-of-inverse gamma prior. Spike-and-slab LASSO
-priors provide computational convenience during model fitting by using
-optimization algorithms instead of intensive sampling algorithms. They
-make fitting high-dimensional models more feasible without sacrificing
-performance in prediction and variable selection. Secondly, SB-GAM uses
-a group spike-and-slab LASSO prior with an EM-CD algorithm to fit the
-model. While both methods use the combination of expectation
-maximization algorithm and coordinate descent algorithm, there are
-subtle difference in the implementation due to the difference in prior
-specification. The proposed model sets up independent priors among basis
-coefficients after the reparameterization step, which provides some
-advantage in computation. Last but not least, the proposed model
-addresses the incapability of bi-level selection in SB-GAM.
+GAMs, i.e., the spike-and-slab GAM \citep{Scheipl2012} and the sparse
+Bayesian GAM (SB-GAM) \citep{Bai2021}, in three ways. Firstly, the
+proposed prior for smoothing functions is a spike-and-slab LASSO type
+prior using independent mixture double exponential distribution,
+compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma
+prior. Spike-and-slab LASSO priors provide computational convenience
+during model fitting by using optimization algorithms instead of
+intensive sampling algorithms. They make fitting high-dimensional models
+more feasible without sacrificing performance in prediction and variable
+selection. Secondly, SB-GAM uses a group spike-and-slab LASSO prior with
+an EM-CD algorithm to fit the model. While both methods use the
+combination of expectation maximization algorithm and coordinate descent
+algorithm, there are subtle differences in the implementation due to the
+difference in prior specification. The proposed model sets up
+independent priors among basis coefficients after the reparameterization
+step, which provides some advantage in computation. Lastly, the proposed
+model addresses the incapability of bi-level selection in SB-GAM.
 
@@
 \textcolor{red}{to factor the smoothing penalty into the design matrix of each smoothing function}.
-Given the smoothing penalty matrix \(\boldsymbol{S}_j\) is symmetric and
-positive semi-definite for the univariate smoothing functions, we
-eigendecompose the penalty matrix
+Given that the smoothing penalty matrix \(\boldsymbol{S}_j\) is
+symmetric and positive semi-definite for the univariate smoothing
+functions, we eigendecompose the penalty matrix
 \(\boldsymbol{S} = \boldsymbol{U} \boldsymbol{D} \boldsymbol{U}^T\) ,
@@
 non-zero eigenvalues of \(\boldsymbol{D}\) such that the new basis
-function matrix \(\boldsymbol{X}^\ast\) can receive uniform penalty on
+function matrix \(\boldsymbol{X}^\ast\) can receive a uniform penalty on
 each of its dimensions. With slight abuse of the notation, we drop the
@@
 
-To summarize, the reparameterization step provides three benefits. First
-of all, the reparameterization integrates the smoothing penalty matrix
+To summarize, the reparameterization step provides three benefits.
+Firstly, the reparameterization integrates the smoothing penalty matrix
 into the design matrix, and encourages models to properly smooth the
@@
 of the linear space from the nonlinear space, improving the feasibility
-of bi-level functional selection. Last but not least, the
-eigendecomposition facilitates the construction of orthonormal design
-matrix, which makes imposing independent priors on the coefficients
-possible. This reduces the computational complexity compared to using a
-multivariate priors, and improve the generalizability of the framework
-to be compatible with other choices of priors.
+of bi-level functional selection. Lastly, the eigendecomposition
+facilitates the construction of an orthonormal design matrix, which
+makes imposing independent priors on the coefficients possible. This
+reduces the computational complexity compared to using a multivariate
+priors, and improve the generalizability of the framework to be
+compatible with other choices of priors.
 
-\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Fucntions}\label{sec:method_prior}
+\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Functions}\label{sec:method_prior}
 
-The family of spike-and-slab regression models is one of most commonly
-used models in high-dimensional data analysis for its utility in outcome
-prediction and variable selection.
+The family of spike-and-slab regression models is one of the most
+commonly used models in high-dimensional data analysis for its utility
+in outcome prediction and variable selection.
 \textcolor{red}{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
@@
 variable \(x\) is included in the model, while the dispersion parameters
-\(s_0\) and \(s_1\) controls the shrinkage of the coefficient. Given
-that both double exponential distributions have a mean of 0 and the
-latent indicator \(\gamma\) can only take the value of 0 or 1, the
-mixture double exponential distribution can be formulated as one single
-double exponential density, \begin{equation} \label{eq:ssl}
+\(s_0\) and \(s_1\) control the shrinkage of the coefficient. Given that
+both double exponential distributions have a mean of 0 and the latent
+indicator \(\gamma\) can only take the value of 0 or 1, the mixture
+double exponential distribution can be formulated as one single double
+exponential density, \begin{equation} \label{eq:ssl}
 \beta | \gamma \sim DE(0, (1-\gamma)s_0 + \gamma s_1), 0 < s_0 < s_1.
@@
 \(\boldsymbol{\beta}_j = \begin{bmatrix} \beta_j : \boldsymbol{\beta}_j^*\end{bmatrix}\).
-Specifically, the linear space coefficient has a SSL prior and the
-nonlinear space coefficients shares a group SSL prior,
+Specifically, the linear space coefficient has an SSL prior and the
+nonlinear space coefficients share a group SSL prior,
 \begin{align}\label{eq:bham_ssl}
@@
 depends on the value of the linear space latent indicator \(\gamma_j\),
-while both latent indicators \(\gamma_j and \gamma^*_j\) follow a
+while both latent indicators \(\gamma_j\) and \(\gamma^*_j\) follow a
 Bernoulli distribution. While the probability of including the linear
@@
 \textcolor{red}{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For the computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
+For computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
 }
@@
 \theta_j \sim Beta(a, b).
-\] The beta distribution is a conjugate prior for the binomial
+\] The beta distribution is a conjugate prior to the binomial
 distribution and hence provides some computation convenience.
 \textcolor{red}{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
-In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particular in pair with the smoothing penalty implicitly addressed via the reparameterization.
-Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different level of sparsity.}
-Hereafter, we refer Bayesian hierarchical generalized additive models
-with the two-part spike-and-slab LASSO prior as the BHAM, and visually
-presented in Figure \ref{fig:SSprior}.
+In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
+Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}
+Hereafter, we refer to the Bayesian hierarchical generalized additive
+models with the two-part spike-and-slab LASSO prior as BHAM, and
+visually presented in Figure \ref{fig:SSprior}.
 
@@
 
-\textcolor{red}{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computational prohibited and not feasible for high-dimensional data.
-Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium size data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
+\textcolor{red}{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computationaly prohibited and not feasible for high-dimensional data.
+Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium-sized data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
 Hence, we consider the optimization algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
@@
 consider the latent binary indicators \(\boldsymbol{\gamma}\) as
-nuisance parameters of the model, in other words the ``missing'' data in
-the EM context. Our objective is to find the parameters \(\Theta\) that
-maximize the posterior density function, or equivalently the logarithm
-of the density function, \[
+nuisance parameters of the model, in other words, the ``missing'' data
+in the EM context. Our objective is to find the parameters \(\Theta\)
+that maximize the posterior density function, or equivalently the
+logarithm of the density function, \[
 \begin{aligned}
@@
 algorithm in each iteration. Coordinate descent is an optimization
-algorithm that offers extreme computational advantage, and famous for
-its application in optimizing the \(l_1\) penalized likelihood function.
-Maximization of \(E(Q_2)\) can be solved via closed form equations
-following the beta-binomial conjugate relationship.
+algorithm that offers extreme computational advantages, and is famous
+for its application in optimizing the \(l_1\) penalized likelihood
+function. Maximization of \(E(Q_2)\) can be solved via closed-form
+equations following the beta-binomial conjugate relationship.
 
@@
 conjugate prior for Bernoulli distribution, \(\boldsymbol{\theta}\) can
-be easily updated with a closed form equation,
+be easily updated with a closed-form equation,
 \begin{equation}\label{eq:update_theta}
@@
 parameters such that the model reaches its best prediction performance
-regarding a criteria of preference. This would be achieved by
-constructing a two-dimensional grid, consists of different pairs of
-(\(s_0\), \(s_1\)) value. However, previous research suggests the value
-of slab scale \(s_1\) has less impact on the final model and is
-recommended to be set as a generally large value, e.g.~\(s_1 = 1\), that
-provides no or weak shrinkage. \citep{Rockova2018} As a result, we focus
-on examining different values of spike scale \(s_0\). Instead of the
-two-dimensional grid, we consider a sequence of \(L\) decreasing values
+regarding criteria of preference. This would be achieved by constructing
+a two-dimensional grid, consisting of different (\(s_0\), \(s_1\))
+pairs. However, previous research suggests the value of slab scale
+\(s_1\) has less impact on the final model and is recommended to be set
+as a generally large value, e.g.~\(s_1 = 1\), that provides no or weak
+shrinkage. \citep{Rockova2018} As a result, we focus on examining
+different values of spike scale \(s_0\). Instead of the two-dimensional
+grid, we consider a sequence of \(L\) decreasing values
 \(\{s_0^l\}: 0 < s_0^1 < s_0^2 < \dots < s_0^L < s_1\). Increasing the
@@
 the LASSO implementation in the widely used R package \texttt{glmnet},
-which quickly fits LASSO models over a list of values of regularization
-parameters \(\lambda\) and gives a sequence of models for users to
-choose from.
+which quickly fits LASSO models over a list of candidate values of
+\(\lambda\) and gives a sequence of models for users to choose from.
 
@@
 COSSO \citep{Storlie2011}, generalized additive models with automatic
-smoothing (referred as \textit{mgcv} hereafter)\citep{Wood2011},
+smoothing (referred to as \textit{mgcv} hereafter)\citep{Wood2011},
 \textcolor{red}{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM
 \citep{Bai2021}.
-\textcolor{red}{We use linear LASSO model as the benchmark, examing the performance when linearity assumption doesn't hold}.
+\textcolor{red}{We use linear LASSO models as the benchmark, examining the performance when the linearity assumption doesn't hold}.
 COSSO is one of the earliest smoothing spline models that consider
@@
 adaptive weight for penalties such that the penalty of each functional
-component are different for extra flexibility. \textit{mgcv} is one of
+component is different for extra flexibility. \textit{mgcv} is one of
 the most popular models for nonlinear effect interpolation and
@@
 \textcolor{red}{mgcv doesn't support analyses when the number of parameters exceeds the sample size.}
-\textcolor{red}{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses a MCMC algorithm for model fitting.}
+\textcolor{red}{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses an MCMC algorithm for model fitting.}
 SB-GAM is the first spike-and-slab LASSO GAM. We implement linear LASSO
@@
 control the dimensionality of each smoothing function, 10 bases, for all
-different choices of smoothing functions. We use 5-fold CV with the
+different choices of smoothing functions. We use a 5-fold CV with the
 default selection criteria to select the final model for linear LASSO
-model, COSSO models, SB-GAM and the proposed model. 20 default
+model, COSSO models, SB-GAM and the proposed model. Twenty default
 candidates of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in
-SB-GAM) are examined for SB-GAM and the proposed model which allow
-user-specification of tuning candidates. All computation was conducted
-on a high-performance 64-bit Linux platform with 48 cores of 2.70GHz
+SB-GAM) are examined for SB-GAM and the proposed model that allow user
+specification of tuning candidates. All computation was conducted on a
+high-performance 64-bit Linux platform with 48 cores of 2.70GHz
 eight-core Intel Xeon E5-2680 processors and 24G of RAM per core and R
@@
 \citep{Ravikumar2009}. However, we exclude these methods from the
-current simulation study because of their demonstrated inferior
-predictive performance compared to \textit{mgcv} \citep{Scheipl2013}.
+current simulation study because they demonstrated inferior predictive
+performance compared to \textit{mgcv} \citep{Scheipl2013}.
 
@@
 \mathbb{E}(Y) = g^{-1}(5 \sin(2\pi x_1) - 4 \cos(2\pi x_2 -0.5) + 6(x_3-0.5) - 5(x_4^2 -0.3))
-\] for Gaussian and binomial outcomes. Gaussian outcomes requires
+\] for Gaussian and binomial outcomes. Gaussian outcomes require
 specification of dispersion, where we set the dispersion parameter to be
@@
 i.e.~\(f_j(x_j) = 0\) for \(j = 4, \dots, p\). Another set of
-independent sample of size \(n_{test}=1000\), are created following the
+independent sample of size \(n_{test}=1000\), is created following the
 same data generating process, serving as the testing data. We generate
@@
 \textcolor{red}{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
-Specifically, BHAM shows greate improvement over COSSO methods, resulting a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) imporvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
-The improvement over spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
+Specifically, BHAM shows great improvement over COSSO methods, resulting in a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) improvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
+The improvement over the spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
 When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear LASSO model predicts less accurate compared to other flexible models across all scenario; mgcv performs extremely well in low-dimensional case ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size apporaches the number of coefficients to estimate ($p=50$), which leads to bad performance.}
+As previously hypothesized, the linear LASSO model predicts less accurate than other flexible models across all scenarios; mgcv performs extremely well in low-dimensional cases ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size approaches the number of coefficients to estimate ($p=50$), which leads to bad performance.}
 Even though SB-GAM has slight prediction advantage over the proposed
-model in high-dimensional situations, the BHAM has extreme computational
-advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in
-computation time (measured in seconds) for Gaussian simulations, without
-sacrificing much of the prediction accuracy (see Table
-\ref{tab:time_sim}).
+model in high-dimensional situations, the BHAM has an extreme
+computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%)
+reduction in computation time (measured in seconds) for Gaussian
+simulations, without sacrificing much of the prediction accuracy (see
+Table \ref{tab:time_sim}).
 
-\textcolor{red}{We also examine the prediction performance when the functional form of predictors are linear, see Supporting Information Table S1, and S2.
-The proposed model, BHAM, has similar performance as the linear LASSO model regardless of the distribution.
-This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless the underlying functional form of predictors.
-spikeSlabGAM have similar prediction performance to BHAM.
+\textcolor{red}{We also examine the prediction performance when the functional form of predictors is linear, see Supporting Information Table S1, and S2.
+The proposed model, BHAM, has a similar performance as the linear LASSO model regardless of the distribution.
+This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless of the underlying functional form of predictors.
+spikeSlabGAM has a similar prediction performance to BHAM.
 Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
@@
 \textcolor{red}{\subsubsection{Variable Selection Performance}}
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimension and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
-The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that, the model selected small subset among all the active predictions.
+\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimensions and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that the model selected a small subset among all the active predictions.
 In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
@@
 
-In this section, we apply the proposed models BHAM to analyze two
+In this section, we apply the proposed model BHAM to analyze two
 published metabolomics datasets where the outcomes are binary and
@@
 
-We use the proposed models BHAM to analyze a metabolic dataset from a
+We use the proposed model BHAM to analyze a metabolic dataset from a
 recently published research \citep{Mehta2020} studying plasma
@@
 subjects from two cohorts. As there is a large number of non-overlapping
-features among the two cohorts, we use the cohort with larger sample
+features among the two cohorts, we use the cohort with a larger sample
 size (N=454). There are initially 6796 features in the dataset, which is
 too large to be practically meaningful to analyze. Hence, we choose the
-the top 200 features \textcolor{red}{with largest variance}. We use
+top 200 features \textcolor{red}{with the largest variance}. We use
 5-knot spline additive models for binary outcome using two different
@@
 samples are used for prediction performance evaluation, where deviance,
-AUC, Brier score defined as
+AUC, Brier score, defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
-misclassification error defined as
+misclassification error, defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}I(|y_i - \hat y_i|>0.5)\) are
@@
 
-We use the proposed models BHAM to analyze metabolomics data from a
+We use the proposed model BHAM to analyze metabolomics data from a
 recently published study \citep{Bihlmeyer2021} on the association
@@
 \textcolor{red}{by the study publishing team \cite{Bihlmeyer2021}}. The
-outcome of interest are standardized percent change in insulin
+outcome of interest is standardized percent change in insulin
 resistance, and hence modeled using a Gaussian model. After removing
@@
 
-\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling.
-In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing function and derive a scalable EM-CD algorithm for model fitting.
+\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeing.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing functions and derive a scalable EM-CD algorithm for model fitting.
 The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
 Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
-In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear component.
+In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear components.
 In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
@@
 
-The proposed model shares many commonality with the SB-GAM
-\citep{Bai2021}, independently developed around the same time of the
-proposed work. Both frameworks emphasize computational efficiency by
-deploying group spike-and-slab LASSO type priors and optimization-based
-scalable algorithms. Bai provides the theoretical proof for the
-consistency of variable selection using group spike-and-slab LASSO
-prior.
+The proposed model shares many commonalities with the SB-GAM
+\citep{Bai2021}, which is independently developed around the same time
+as the proposed work. Both frameworks emphasize computational efficiency
+by deploying group spike-and-slab LASSO type priors and
+optimization-based scalable algorithms. Bai provides the theoretical
+proof for the consistency of variable selection using group
+spike-and-slab LASSO prior.
 \textcolor{red}{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.
+Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.
 Not focused in this manuscript, the generalization is described in the Supporting Information.}
 
-\textcolor{red}{During designing and analyzing the simulation study, we made couple interesting observations.
-First of all, variable selection is a delicate topic in the context predictive modelling.
-When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example LASSO and LASSO-based models.
+\textcolor{red}{During designing and analyzing the simulation study, we made couple of interesting observations.
+First of all, variable selection is a delicate topic in the context of predictive modeling.
+When prediction performance is used to tune a model, the model could possibly include noise variables in models, for example LASSO and LASSO-based models.
 \cite{Wu2019} Moreover, bi-level selection is a more complex problem than variable selection.
-The complexity reflects on the validity of the effect hierarchy principle.
-While most functional forms follow that linear components exists in the nonlinear function, there are functions that don't follow it, e.g.
+The complexity shows on the validity of the effect hierarchy principle.
+While most functional forms follow that the linear component exists in the nonlinear function, there are functions that don't follow it, e.g.
 $x^2$.
-The proposed prior and spikeSlabGAM employ different structure: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
+The proposed prior and spikeSlabGAM employ different structures: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
 The different prior setups lead to trade-offs for the purpose of bi-level selection.
-We recommend to use more judgement in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
-Secondly, we find the performance of the proposed model are more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
-Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend to fine-tune the model with granular sequence of $s_0$ for performance improvement.}
+We recommend to use more judgment in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
+Secondly, we find the performance of the proposed model is more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
+Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend fine-tuning the model with a granular sequence of $s_0$ for performance improvement.}
 
-Our future efforts direct to
+Our future efforts are direct to
 \textcolor{red}{uncertainty inference of the proposed model}, survival
@@
 We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
-Instead of the Coordinate Descent algorithm, we use Iterative Weighted Least Square algorithm in the EM procedure.
+Instead of the Coordinate Descent algorithm, we use the Iterative Weighted Least Square algorithm in the EM procedure.
 The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
 Due to the space limit, technical details will be explained in a future manuscript.}
-While the proposed model addresses a great deal of analytic problem,
-analyzing the time-to-event outcome remains unsolved. An naive approach
+While the proposed model addresses a great deal of analytic problems,
+analyzing the time-to-event outcome remains unsolved. A naive approach
 would be convert a time-to-event outcome to a Poisson outcome following
@@
 directly fit Cox models via penalized pseudo likelihood function
-\citep{Simon2011}. Meanwhile, with growing understanding of biological
-structure within -omics field, it is appealing to integrate external
-biology information in the modeling process. The main motivation for
-integrative models is that biologically informed grouping of weak
-effects increases the power of detecting true associations between
-features and the outcome \citep{Peterson2016}, and stabilizes the
-analysis results for reproducibility purpose. Such integration can be
-achieved by setting up a structural hyperprior on the inclusion
+\citep{Simon2011}. Meanwhile, with the growing understanding of
+biological structure within -omics field, it is appealing to integrate
+external biology information in the modeling process. The main
+motivation for integrative models is that biologically informed grouping
+of weak effects increases the power of detecting true associations
+between features and the outcome \citep{Peterson2016}, and stabilizes
+the analysis results for reproducibility purposes. Such integration can
+be achieved by setting up a structural hyperprior on the inclusion
 indicator of the smoothing function null space
-\(\boldsymbol{\gamma}^0\). The similar strategy has been used in Ferrari
+\(\boldsymbol{\gamma}^0\). A similar strategy has been used in Ferrari
 and Dunson \citep{Ferrari2020}.
@@
 \end{tabular}
-\caption{Prediction performance of BHAM fitted with Cooridnate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
+\caption{Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM) and SB-GAM models for Emory Cardiovascular Biobank by 10-fold cross-validation, including deviance, area under the curve (AUC), Brier score, and misclassification error (Misclass) where class labels are defined using threshold = 0.5.}
 \label{tab:ECB_res}
@@
 \end{tabular}
-\caption{Prediction performance of of BHAM fitted with Cooridnate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
+\caption{Prediction performance of BHAM fitted with Coordinate Descent algorithm (BHAM)  and SB-GAM models for Weight Loss Maintenance Cohort by 10-fold cross-validation, including deviance, $R^2$,  mean squared error (MSE), and mean absolute error (MAE).}
 \label{tab:WLM_res}

### 2022-05-15 Manuscript/Tabs/sim_binom_tab.tex

--- 2022-03-21 Manuscript/Tabs/sim_binom_tab.tex
@@
     for binomial outcomes over 50 iterations. The models of comparison include the proposed Bayesian
-    hierarchical additive model (BHAM), component selection and smoothing operator (COSSO),
-    adaptive COSSO, mgcv and sparse Bayesian generalized additive model (SB-GAM). mgcv doesn't provide
-    estimation whe number of parameters exceeds sample size i.e. p = 100, 200.} 
+    hierarchical additive model (BHAM), linear LASSO model (LASSO), component selection and smoothing
+    operator (COSSO), adaptive COSSO, mgcv, sparse Bayesian generalized additive model (SB-GAM), and
+    spikeSlabGAM model. mgcv doesn't provide
+    estimation when number of parameters exceeds sample size i.e. p = 100, 200.} 
 \label{tab:bin_auc}

### 2022-05-15 Manuscript/Tabs/sim_binom_var_slct_tab.tex

@@
                          smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive model (SB-GAM),
-                         and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.} 
+                         and spikeSlabGAM model. MCC is ill-defined when $p=4$ simulation (no true negative), and hence omitted for all methods.} 
 \label{tab:sim_binom_var_select}

### 2022-05-15 Manuscript/Tabs/sim_gaus_var_slct_tab.tex

@@
                          and smoothing operator (COSSO), adaptive COSSO, sparse Bayesian generalized additive
-                         model (SB-GAM), and spikeSlabGAM mdoel. MCC is ill-defined when $p=4$ simulation
+                         model (SB-GAM), and spikeSlabGAM model. MCC is ill-defined when $p=4$ simulation
                          (no true negative), and hence omitted for all methods.} 

### 2022-05-16 Manuscript/00-main.Rmd
Replace "smoothing function" with "smooth function"

@@
 abstract: "There are proposals that extend the classical generalized additive models (GAMs) to accommodate high-dimensional data ($p>>n$) using group sparse regularization.
-However, the sparse regularization may induce excess shrinkage when estimating smoothing functions, damaging predictive performance.
+However, the sparse regularization may induce excess shrinkage when estimating smooth functions, damaging predictive performance.
 Moreover, most of these GAMs consider an “all-in-all-out” approach for functional selection, rendering them difficult to answer if nonlinear effects are necessary.
@@
 Hence, we propose Bayesian hierarchical generalized additive models as a solution: we consider the smoothing penalty for proper shrinkage of curve interpolation via reparameterization.
-A novel two-part spike-and-slab LASSO prior for smoothing functions is developed to address the sparsity of signals while providing extra flexibility to select the linear or nonlinear components of smoothing functions.
+A novel two-part spike-and-slab LASSO prior for smooth functions is developed to address the sparsity of signals while providing extra flexibility to select the linear or nonlinear components of smooth functions.
 A scalable and deterministic algorithm, EM-Coordinate Descent, is implemented in an open-source R package BHAM.

### 2022-05-16 Manuscript/01-Intro.Rmd
Replace "smoothing function" with "smooth function"

@@
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
-In a GAM, the response variable, which is assumed to follow some exponential family distribution, can be modeled with the summation of smoothing functions.
+In a GAM, the response variable, which is assumed to follow some exponential family distribution, can be modeled with the summation of smooth functions.
 Nevertheless, the classical GAMs cannot fulfill the increasing analytic demands for high-dimensional data analysis.
@@
 There exist some proposals to generalize the classical GAM to accommodate high-dimensional applications.
-The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smoothing functions.
+The regularized models, branching out from group regularized linear models, are used to fit GAMs by accounting for the structure introduced when expanding smooth functions.
 Ravikumar et al.
@@
 Bai and co-authors[@Bai2020] were the first to apply group spike-and-slab LASSO prior to Gaussian AMs using a fast optimization algorithm and further generalized the framework to GAMs [@Bai2021].
-Focus on addressing the sparsity, these methods can excessively penalize the bases of a smoothing function and produce inaccurate predictions, particularly when complex signals are assumed and large numbers of knots are used.
+Focus on addressing the sparsity, these methods can excessively penalize the bases of a smooth function and produce inaccurate predictions, particularly when complex signals are assumed and large numbers of knots are used.
 [@Scheipl2013] In addition, these methods adopt an 'all-in-all-out' strategy, i.e.
@@
 \bg{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
-Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smoothing functions to avoid excessive shrinkage on the bases.
+Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smooth functions to avoid excessive shrinkage on the bases.
 Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
@@
 The proposed framework, BHAM, differs from previous spike-and-slab based GAMs, i.e., the spike-and-slab GAM [@Scheipl2012] and the sparse Bayesian GAM (SB-GAM) [@Bai2021], in three ways.
-Firstly, the proposed prior for smoothing functions is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
+Firstly, the proposed prior for smooth functions is a spike-and-slab LASSO type prior using independent mixture double exponential distribution, compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
 Spike-and-slab LASSO priors provide computational convenience during model fitting by using optimization algorithms instead of intensive sampling algorithms.

### 2022-05-16 Manuscript/02-Method.Rmd
Replace "smoothing function" with "smooth function"

@@
 \bg{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
-The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\bs x$, written as
+The mean of the response variable can be modeled as the summation of smooth functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\bs x$, written as
 \begin{equation}\label{eq:gam}
@@
 The basis function matrix, i.e.
-the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
-The dimension of the design matrix depends on the choice of the smoothing function, and is denoted as $K_j$ for $x_j$.
+the design matrix derived from the smooth function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$.
+The dimension of the design matrix depends on the choice of the smooth function, and is denoted as $K_j$ for $x_j$.
 $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable such that $B_j(x_j) = \bs \beta_j^T \bs X_j$.
@@
 
-\subsection{Smoothing Function Reparameterization}
+\subsection{smooth function Reparameterization}
 
@@
 
-Marra and  Wood [@Marra2011] proposed a reparameterization procedure \bg{to factor the smoothing penalty into the design matrix of each smoothing function}.
-Given that the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smoothing functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
-To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smoothing function.
-For the most popular smoothing function, cubic splines, the dimension of the linear space is one.
+Marra and  Wood [@Marra2011] proposed a reparameterization procedure \bg{to factor the smoothing penalty into the design matrix of each smooth function}.
+Given that the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smooth functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
+To note, $\bs D$ can contain elements of zeros on the diagonal, where the zeros are associated with the linear space of the smooth function.
+For the most popular smooth function, cubic splines, the dimension of the linear space is one.
 Hereafter, we focus on discussing a uni-dimensional linear space for simplicity; however, it generalizes easily to the cases where the linear space is multidimensional.
@@
 
-\subsection{Two-part Spike-and-Slab LASSO Prior for Smoothing Functions}\label{sec:method_prior}
+\subsection{Two-part Spike-and-Slab LASSO Prior for Smooth Functions}\label{sec:method_prior}
 
@@
 
-We introduce a novel SSL-based prior for smoothing functions in GAMs.
+We introduce a novel SSL-based prior for smooth functions in GAMs.
 Given the reparameterized design matrix $\bs X_j = \begin{bmatrix} X^0_j : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j = \begin{bmatrix} \beta_j : \bs \beta_j^\tp \end{bmatrix}$.
@@
 <!-- The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior.
-First of all, the smoothing function bases are linear dependent and necessitate extra attention.
+First of all, the smooth function bases are linear dependent and necessitate extra attention.
 The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent.
-Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions.
+Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smooth functions.
 -->
@@
 
-<!-- This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the linear space and nonlinear space of the smoothing function.
+<!-- This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the linear space and nonlinear space of the smooth function.
 The hyper prior of $\gamma_{j}$ decides the sparsity of the model at the functional selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level.
@@
 \bg{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
-In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
+In other words, when a smooth function is significant, the coefficients of the smooth function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
 Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}

### 2022-05-16 Manuscript/03-Simulation.Rmd
Replace "smoothing function" with "smooth function"

@@
 We implement linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
-COSSO models and SB-GAM do not provide flexibility to define smoothing functions, and hence use the default choices; mgcv, spikeSlabGAM and the proposed model allow customized smoothing functions and we choose the cubic regression spline.
-We control the dimensionality of each smoothing function, 10 bases, for all different choices of smoothing functions.
+COSSO models and SB-GAM do not provide flexibility to define smooth functions, and hence use the default choices; mgcv, spikeSlabGAM and the proposed model allow customized smooth functions and we choose the cubic regression spline.
+We control the dimensionality of each smooth function, 10 bases, for all different choices of smooth functions.
 We use a 5-fold CV with the default selection criteria to select the final model for linear LASSO model, COSSO models, SB-GAM and the proposed model.

### 2022-05-16 Manuscript/05-Discussion.Rmd
Replace "smoothing function" with "smooth function"

@@
 \label{sec:concl}
-\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeing.
-In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing functions and derive a scalable EM-CD algorithm for model fitting.
-The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
+\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeling.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smooth functions and derive a scalable EM-CD algorithm for model fitting.
+The proposed model provides a new angle to address the excess shrinkage of smooth functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
 Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
@@
 When serving the purpose of variable selection, trade-offs exist among methods of comparison.} We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
-To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones.
+To maximize the flexibility of smooth function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smooth functions are limited to the default ones.
 Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.
@@
 \bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.
+Moreover, the proposed model can easily generalize to other families of priors and smooth functions if desired.
 Not focused in this manuscript, the generalization is described in the Supporting Information.}
@@
 Current implementation of the EM-IWLS algorithm jointly updates all coefficients in each iteration, which requires a lot computation resources.
-This jointly updating procedure can be enhanced by adapting a backfitting step [@Hastie1987] where each smoothing function are updated individually.
+This jointly updating procedure can be enhanced by adapting a backfitting step [@Hastie1987] where each smooth function are updated individually.
 Thirdly, when using the proposed model in real data analysis where a screening procedure is implemented before joint predictive modeling.
@@
 The main motivation for integrative models is that biologically informed grouping of weak effects increases the power of detecting true associations between features and the outcome [@Peterson2016], and stabilizes the analysis results for reproducibility purposes.
-Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smoothing function null space $\bs \gamma^0$.
+Such integration can be achieved by setting up a structural hyperprior on the inclusion indicator of the smooth function null space $\bs \gamma^0$.
 A similar strategy has been used in Ferrari and Dunson [@Ferrari2020].

### 2022-05-16 Manuscript/10-Supplementary.Rmd
Replace "smoothing function" with "smooth function"

@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(targets)
+
+tar_load(manu_tbs)
 
@@
 In this work, we further extend the EM-IWLS paradigm to fit the proposed BHAM method.
-Compare to the EM-CD algorithm, EM-IWLS estimates the variance-covariance matrix of the coefficients, providing an opportunity to derive the uncertainty quantification of smoothing functions.
+Compare to the EM-CD algorithm, EM-IWLS estimates the variance-covariance matrix of the coefficients, providing an opportunity to derive the uncertainty quantification of smooth functions.
 We will defer this discussion to future work, due to the delicacy of the topic.
@@
 In the manuscript, we describe the Bayesian hierarchical additive model with the two-part spike-and-slab LASSO prior.
-Nevertheless, the proposed model and algorithm can be easily generalized to accommodate other priors thanks to the reparameterization of smoothing functions.
+Nevertheless, the proposed model and algorithm can be easily generalized to accommodate other priors thanks to the reparameterization of smooth functions.
 To be more specific, we can apply a regularized prior on the linear coefficient, and a group regularized prior for the nonlinear coefficients.

### 2022-05-16 Manuscript/SS_GAM.tex
update manuscript
Replace "smoothing function" with "smooth function"

@@
 using group sparse regularization. However, the sparse regularization
-may induce excess shrinkage when estimating smoothing functions,
-damaging predictive performance. Moreover, most of these GAMs consider
-an ``all-in-all-out'' approach for functional selection, rendering them
+may induce excess shrinkage when estimating smooth functions, damaging
+predictive performance. Moreover, most of these GAMs consider an
+``all-in-all-out'' approach for functional selection, rendering them
 difficult to answer if nonlinear effects are necessary. While some
@@
 curve interpolation via reparameterization. A novel two-part
-spike-and-slab LASSO prior for smoothing functions is developed to
-address the sparsity of signals while providing extra flexibility to
-select the linear or nonlinear components of smoothing functions. A
-scalable and deterministic algorithm, EM-Coordinate Descent, is
-implemented in an open-source R package BHAM. Simulation studies and
-metabolomics data analyses demonstrate improved predictive and
-computational performance against state-of-the-art models. Functional
-selection performance suggests trade-offs exist regarding the effect
-hierarchy assumption.}
+spike-and-slab LASSO prior for smooth functions is developed to address
+the sparsity of signals while providing extra flexibility to select the
+linear or nonlinear components of smooth functions. A scalable and
+deterministic algorithm, EM-Coordinate Descent, is implemented in an
+open-source R package BHAM. Simulation studies and metabolomics data
+analyses demonstrate improved predictive and computational performance
+against state-of-the-art models. Functional selection performance
+suggests trade-offs exist regarding the effect hierarchy assumption.}
 
@@
 assumed to follow some exponential family distribution, can be modeled
-with the summation of smoothing functions. Nevertheless, the classical
-GAMs cannot fulfill the increasing analytic demands for high-dimensional
-data analysis.
+with the summation of smooth functions. Nevertheless, the classical GAMs
+cannot fulfill the increasing analytic demands for high-dimensional data
+analysis.
 
@@
 branching out from group regularized linear models, are used to fit GAMs
-by accounting for the structure introduced when expanding smoothing
+by accounting for the structure introduced when expanding smooth
 functions. Ravikumar et al. \citep{Ravikumar2009} extended the group
@@
 \citep{Bai2021}. Focus on addressing the sparsity, these methods can
-excessively penalize the bases of a smoothing function and produce
+excessively penalize the bases of a smooth function and produce
 inaccurate predictions, particularly when complex signals are assumed
@@
 \textcolor{red}{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
-Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smoothing functions to avoid excessive shrinkage on the bases.
+Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smooth functions to avoid excessive shrinkage on the bases.
 Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
@@
 Bayesian GAM (SB-GAM) \citep{Bai2021}, in three ways. Firstly, the
-proposed prior for smoothing functions is a spike-and-slab LASSO type
-prior using independent mixture double exponential distribution,
-compared to spike-and-slab GAM that uses normal-mixture-of-inverse gamma
-prior. Spike-and-slab LASSO priors provide computational convenience
-during model fitting by using optimization algorithms instead of
-intensive sampling algorithms. They make fitting high-dimensional models
-more feasible without sacrificing performance in prediction and variable
+proposed prior for smooth functions is a spike-and-slab LASSO type prior
+using independent mixture double exponential distribution, compared to
+spike-and-slab GAM that uses normal-mixture-of-inverse gamma prior.
+Spike-and-slab LASSO priors provide computational convenience during
+model fitting by using optimization algorithms instead of intensive
+sampling algorithms. They make fitting high-dimensional models more
+feasible without sacrificing performance in prediction and variable
 selection. Secondly, SB-GAM uses a group spike-and-slab LASSO prior with
@@
 \textcolor{red}{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
-The mean of the response variable can be modeled as the summation of smoothing functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\boldsymbol{x}$, written as
+The mean of the response variable can be modeled as the summation of smooth functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\boldsymbol{x}$, written as
 \begin{equation}\label{eq:gam}
@@
 The basis function matrix, i.e.~the design matrix derived from the
-smoothing function \(B_j(x_j)\), is denoted \(\boldsymbol{X}_j\) for the
+smooth function \(B_j(x_j)\), is denoted \(\boldsymbol{X}_j\) for the
 variable \(x_j\). The dimension of the design matrix depends on the
-choice of the smoothing function, and is denoted as \(K_j\) for \(x_j\).
+choice of the smooth function, and is denoted as \(K_j\) for \(x_j\).
 \(\boldsymbol{\beta}_j\) denotes the basis function coefficients for the
@@
 Marra and Wood \citep{Marra2011} proposed a reparameterization procedure
-\textcolor{red}{to factor the smoothing penalty into the design matrix of each smoothing function}.
+\textcolor{red}{to factor the smoothing penalty into the design matrix of each smooth function}.
 Given that the smoothing penalty matrix \(\boldsymbol{S}_j\) is
-symmetric and positive semi-definite for the univariate smoothing
+symmetric and positive semi-definite for the univariate smooth
 functions, we eigendecompose the penalty matrix
@@
 elements of zeros on the diagonal, where the zeros are associated with
-the linear space of the smoothing function. For the most popular
-smoothing function, cubic splines, the dimension of the linear space is
-one. Hereafter, we focus on discussing a uni-dimensional linear space
-for simplicity; however, it generalizes easily to the cases where the
-linear space is multidimensional. We further write the orthonormal
-matrix
+the linear space of the smooth function. For the most popular smooth
+function, cubic splines, the dimension of the linear space is one.
+Hereafter, we focus on discussing a uni-dimensional linear space for
+simplicity; however, it generalizes easily to the cases where the linear
+space is multidimensional. We further write the orthonormal matrix
 \(\boldsymbol{U} \equiv \begin{bmatrix} \boldsymbol{U}^0 : \boldsymbol{U}^{*}\end{bmatrix}\)
@@
 
-We introduce a novel SSL-based prior for smoothing functions in GAMs.
-Given the reparameterized design matrix
+We introduce a novel SSL-based prior for smooth functions in GAMs. Given
+the reparameterized design matrix
 \(\boldsymbol{X}_j = \begin{bmatrix} X^0_j : \boldsymbol{X}_j^*\end{bmatrix}\)
@@
 \textcolor{red}{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
-In other words, when a smoothing function is significant, the coefficients of the smoothing function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
+In other words, when a smooth function is significant, the coefficients of the smooth function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
 Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}
@@
 package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smoothing functions, and hence use the default
+flexibility to define smooth functions, and hence use the default
 choices; mgcv, spikeSlabGAM and the proposed model allow customized
-smoothing functions and we choose the cubic regression spline. We
-control the dimensionality of each smoothing function, 10 bases, for all
-different choices of smoothing functions. We use a 5-fold CV with the
-default selection criteria to select the final model for linear LASSO
-model, COSSO models, SB-GAM and the proposed model. Twenty default
-candidates of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in
-SB-GAM) are examined for SB-GAM and the proposed model that allow user
-specification of tuning candidates. All computation was conducted on a
+smooth functions and we choose the cubic regression spline. We control
+the dimensionality of each smooth function, 10 bases, for all different
+choices of smooth functions. We use a 5-fold CV with the default
+selection criteria to select the final model for linear LASSO model,
+COSSO models, SB-GAM and the proposed model. Twenty default candidates
+of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are
+examined for SB-GAM and the proposed model that allow user specification
+of tuning candidates. All computation was conducted on a
 high-performance 64-bit Linux platform with 48 cores of 2.70GHz
@@
 
-\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeing.
-In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smoothing functions and derive a scalable EM-CD algorithm for model fitting.
-The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
+\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeling.
+In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smooth functions and derive a scalable EM-CD algorithm for model fitting.
+The proposed model provides a new angle to address the excess shrinkage of smooth functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
 Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
@@
 \texttt{BHAM}, deposited at \url{https://github.com/boyiguo1/BHAM}. To
-maximize the flexibility of smoothing function specification, we deploy
-the same programming grammar as in the state-of-the-art package
-\texttt{mgcv}, in contrast to previous tools where smoothing functions
-are limited to the default ones. Ancillary functions are provided for
-model specification in high-dimensional settings, curve plotting and
+maximize the flexibility of smooth function specification, we deploy the
+same programming grammar as in the state-of-the-art package
+\texttt{mgcv}, in contrast to previous tools where smooth functions are
+limited to the default ones. Ancillary functions are provided for model
+specification in high-dimensional settings, curve plotting and
 functional selection.
@@
 \textcolor{red}{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.
+Moreover, the proposed model can easily generalize to other families of priors and smooth functions if desired.
 Not focused in this manuscript, the generalization is described in the Supporting Information.}
@@
 be achieved by setting up a structural hyperprior on the inclusion
-indicator of the smoothing function null space
-\(\boldsymbol{\gamma}^0\). A similar strategy has been used in Ferrari
-and Dunson \citep{Ferrari2020}.
+indicator of the smooth function null space \(\boldsymbol{\gamma}^0\). A
+similar strategy has been used in Ferrari and Dunson
+\citep{Ferrari2020}.
 
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^*\sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j)$ };\\
-  \node (penBeta) [outCome] { $\beta_{jk}^*\sim  DE(0,(1-\gamma^*_{j}) s_0 + \gamma^*_{j} s_1)$};\\
+  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \bg{\gamma_{j}}\theta_j)$ };\\
+  \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
 };
@@
 \node (s) [staticCompo]  at ($(linearBeta)!0.5!(penBeta)$)  {($s_0, s_1$)};
-\node (Beta) [staticCompo, below = 1cm of s] {$\boldsymbol{\beta }= (\beta_1, \boldsymbol{\beta}^*_1, \dots,\beta_j, \boldsymbol{\beta}^*_j , \dots,\beta_p, \boldsymbol{\beta}^*_p) $};
+\node (Beta) [staticCompo, below = 1cm of s] {$\bs \beta = (\beta_1, \bs \beta^\tp_1, \dots,\beta_j, \bs \beta^\tp_j , \dots,\beta_p, \bs \beta^\tp_p) $};
 \node (Theta)[outCome, above = 2cm of s] {$\theta_{j} \sim Beta(a, b)$};
 \node (ab)[staticCompo, above = 0.5cm of Theta] {$(a, b)$};
-\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\boldsymbol{\beta}^T \boldsymbol{X}_i), \phi)$};
+\node (Y) [outCome, below = 1cm of Beta] {$y_i \sim Expo. Fam. (g^{-1}(\bs \beta^T \bs X_i), \phi)$};
 

### 2022-05-16 Manuscript/SS_GAM_Supporting_Information.tex
Replace "smoothing function" with "smooth function"

@@
 ]{article}
-\title{Spike-and-Slab LASSO Generalized Additive Models and Scalable
-Algorithms for High-Dimensional Data Analysis}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Supporting Information}
-\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
-Nengjun Yi}
-\date{}
-
 \usepackage{amsmath,amssymb}
@@
 
+\title{Spike-and-Slab LASSO Generalized Additive Models and Scalable
+Algorithms for High-Dimensional Data Analysis}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Supporting Information}
+\author{Boyi Guo, Byron C. Jaeger, AKM Fazlur Rahman, D. Leann Long,
+Nengjun Yi}
+\date{}
+
 \begin{document}
@@
 the coefficients, providing an opportunity to derive the uncertainty
-quantification of smoothing functions. We will defer this discussion to
+quantification of smooth functions. We will defer this discussion to
 future work, due to the delicacy of the topic. However, the EM-IWLS is
@@
 model and algorithm can be easily generalized to accommodate other
-priors thanks to the reparameterization of smoothing functions. To be
-more specific, we can apply a regularized prior on the linear
-coefficient, and a group regularized prior for the nonlinear
-coefficients. For example, we can apply the same proposed framework with
-a spike-and-slab mixture normal prior, \begin{align*}
+priors thanks to the reparameterization of smooth functions. To be more
+specific, we can apply a regularized prior on the linear coefficient,
+and a group regularized prior for the nonlinear coefficients. For
+example, we can apply the same proposed framework with a spike-and-slab
+mixture normal prior, \begin{align*}
   \beta_{j} |\gamma_{j},s_0,s_1 &\sim N(0,(1-\gamma_{j}) s_0 + \gamma_{j} s_1)\\

### 2022-05-16 Manuscript/01-Intro.Rmd
remove revision highlights in red

@@
 These models are useful for statistical prediction but do not estimate parameters relevant to the data generation process that one can draw inferences from.
-In addition, how to generalize these \bg{"black box"} models to the high-dimensional setting remains unclear.
+In addition, how to generalize these "black box" models to the high-dimensional setting remains unclear.
 
-\bg{For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection}.
+For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection.
 Among those, generalized additive models (GAMs), proposed in the seminal work of Hastie and Tibshirani [@Hastie1987], grew to be one of the most popular modeling tools.
@@
 
-\bg{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
+We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
 Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smooth functions to avoid excessive shrinkage on the bases.
@@
 Notably, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
+The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.
 

### 2022-05-16 Manuscript/02-Method.Rmd
* Remove useless markdowns * Correct smoothing penalty equation
remove revision highlights in red

@@
 
-\bg{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
+We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
 The mean of the response variable can be modeled as the summation of smooth functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\bs x$, written as
@@
 \end{equation}
-where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \bs x_i\}^n_{i=1}$,} the data distribution is expressed as
+where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \bs x_i\}^n_{i=1}$, the data distribution is expressed as
 \begin{equation}
@@
 The matrix transposing operation is denoted with a superscript $^T$.
-\bg{To note, the proposed model can include parametric forms of variables in the model, and hence considers general linear models and semiparametric regression models as special cases.}
+To note, the proposed model can include parametric forms of variables in the model, and hence considers general linear models and semiparametric regression models as special cases.
 
-\subsection{smooth function Reparameterization}
+\subsection{Smooth Function Reparameterization}
 
@@
 \begin{equation}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,\nonumber
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)^2dx = \lambda_j \bs \beta_j^T \bs S_j \bs \beta_j ,\nonumber
 \end{equation}
@@
 A linear function can be modeled as $B_j(x_j) = x_j$ with the smoothing penalty matrix $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.
-\bg{Unlike previous regularized methods that either ignore the smoothing penalty completely or restrain the smoothing penalty as a component of sparse penalty which leads to a more restrictive solution, we consider an additional mechanism in pair with the proposed prior (described in Section \ref{sec:method_prior}) to address the smoothness and sparsity in signals such that the locally adaptive nature of the smoothing penalty retains.}
+Unlike previous regularized methods that either ignore the smoothing penalty completely or restrain the smoothing penalty as a component of sparse penalty which leads to a more restrictive solution, we consider an additional mechanism in pair with the proposed prior (described in Section \ref{sec:method_prior}) to address the smoothness and sparsity in signals such that the locally adaptive nature of the smoothing penalty retains.
 
-Marra and  Wood [@Marra2011] proposed a reparameterization procedure \bg{to factor the smoothing penalty into the design matrix of each smooth function}.
+Marra and  Wood [@Marra2011] proposed a reparameterization procedure to factor the smoothing penalty into the design matrix of each smooth function.
 Given that the smoothing penalty matrix $\bs S_j$ is symmetric and positive semi-definite for the univariate smooth functions, we eigendecompose the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order.
@@
 The family of spike-and-slab regression models is one of the most commonly used models in high-dimensional data analysis for its utility in outcome prediction and variable selection.
-\bg{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
+Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
 The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
@@
 \end{equation}
-}
 The latent binary variable $\gamma \in \{0,1\}$ indicates whether a variable $x$ is included in the model, while the dispersion parameters $s_0$ and $s_1$ control the shrinkage of the coefficient.
@@
 \end{equation}
-\bg{Compared to other priors for high-dimensional data analysis, SSL has the following advantages.
+Compared to other priors for high-dimensional data analysis, SSL has the following advantages.
 First of all, the SSL prior provides a locally adaptive shrinkage when estimating the coefficients.
@@
 We defer to Bai et al.
-\cite{Bai2021Review} for a detailed discussion.}
+\cite{Bai2021Review} for a detailed discussion.
 
@@
 
-\bg{The proposed two-part SSL prior, particularly the group SSL prior of the nonlinear space coefficients, differs from previous group SSL priors \cite{Tang2018, Tang2019}, as the proposed prior follows the effect hierarchy principle.
+The proposed two-part SSL prior, particularly the group SSL prior of the nonlinear space coefficients, differs from previous group SSL priors \cite{Tang2018, Tang2019}, as the proposed prior follows the effect hierarchy principle.
 Effect hierarchy refers to the principle that "lower-order effects are more likely to be active than higher-order effects" defined by Chipman\cite{chipman2006prior}.
-To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^\tp_j$} depends on the value of the linear space latent indicator $\gamma_j$, while both latent indicators $\gamma_j$ and $\gamma^\tp_j$ follow a Bernoulli distribution.
-While the probability of including the linear effect is $\theta_j$, the probability of including the nonlinear effect is $\bg{\gamma_{j}}\theta_j$.
+To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^\tp_j$ depends on the value of the linear space latent indicator $\gamma_j$, while both latent indicators $\gamma_j$ and $\gamma^\tp_j$ follow a Bernoulli distribution.
+While the probability of including the linear effect is $\theta_j$, the probability of including the nonlinear effect is $\gamma_{j}\theta_j$.
 $$
@@
 &\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^\tp | \bg{\gamma_{j}, } \theta_j \sim Bin(1, \bg{\gamma_{j}}\theta_j).
+&\gamma_{j}^\tp | \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j).
 \end{aligned}
 $$
-\bg{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
+This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
 For computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^\tp | \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
-}
 
@@
 
-\bg{To allow the shrinkage to self-adapt to the sparsity and smoothing pattern of the data, }we further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
+To allow the shrinkage to self-adapt to the sparsity and smoothing pattern of the data, we further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
 $$
@@
 <!-- For simplicity, we focus on a special case of beta distribution, uniform (0,1), i.e. $a = 1, b = 1$. -->
-\bg{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
+Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
 In other words, when a smooth function is significant, the coefficients of the smooth function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
-Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}
+Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.
 <!-- This informaiton borrowing helps the model to find adequate shrinkage for the linear coefficient and coefficients of the nonlinear part.
@@
 
-\bg{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computationaly prohibited and not feasible for high-dimensional data.
+Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computationaly prohibited and not feasible for high-dimensional data.
 Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium-sized data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
@@
 The EM-CD algorithm has been well adapted in generalized linear models \cite{Tang2017a}, Cox proportional hazards models \cite{Tang2017}, and their grouped counterparts \cite{Tang2018, Tang2019}.
-The EM-CD algorithm provides deterministic solutions, which becomes a popular property for reproducible research.}
+The EM-CD algorithm provides deterministic solutions, which becomes a popular property for reproducible research.
 
@@
 In the M-step, we find the parameters $\Theta^{(t)}$ that maximize $E\log f(\Theta, \bs \gamma| \textbf{y}, \textbf{X})$.
-\bg{The parenthesized subscription $^{(t)}$ denotes the parameter estimation at the $t$th iteration.} The E- and M- steps are iterated until the algorithm converges.
+The parenthesized subscription $^{(t)}$ denotes the parameter estimation at the $t$th iteration.
+The E- and M- steps are iterated until the algorithm converges.
 
@@
 \end{align*}
-where $\pr(\gamma_{j} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j} = 0|\theta_j) = 1-\theta_j$, $\bg{\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2}$, $\bg{\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j}$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
+where $\pr(\gamma_{j} = 1|\theta_j) = \theta_j$, $\pr(\gamma_{j} = 0|\theta_j) = 1-\theta_j$, $\pr(\gamma_{j}^\tp = 1|\theta_j) = \theta_j^2$, $\pr(\gamma_{j}^\tp = 0|\theta_j) = 1-\theta^2_j$, $f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)$, $f(\beta|\gamma=0, s_0) = \text{DE}(\beta|0 , s_0)$.
 It is trivial to show
@@
 A measure of preference calculated with cross-validations (CV), e.g.
-deviance \bg{(defined as model log-likelihood times -2, $-2\log f(\bs y|\bs{\hat\beta}, \hat\phi)$)}, area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
+deviance (defined as model log-likelihood times -2, $-2\log f(\bs y|\bs{\hat\beta}, \hat\phi)$), area under the curve (AUC), mean squared error, can be used to facilitate the selection of a final model.
 The procedure is similar to the LASSO implementation in the widely used R package \texttt{glmnet}, which quickly fits LASSO models over a list of candidate values of $\lambda$ and gives a sequence of models for users to choose from.

### 2022-05-16 Manuscript/03-Simulation.Rmd
remove revision highlights in red

@@
 
-In this section, we compare the performance of the proposed model to \bg{six alternative models: linear LASSO models,} component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred to as \textit{mgcv} hereafter)[@Wood2011], \bg{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM [@Bai2021].
-\bg{We use linear LASSO models as the benchmark, examining the performance when the linearity assumption doesn't hold}.
+In this section, we compare the performance of the proposed model to six alternative models: linear LASSO models, component selection and smoothing operator (COSSO) [@Zhang2006GAM], adaptive COSSO [@Storlie2011], generalized additive models with automatic smoothing (referred to as \textit{mgcv} hereafter)[@Wood2011], spike-and-slab GAM \cite{Scheipl2012}, and SB-GAM [@Bai2021].
+We use linear LASSO models as the benchmark, examining the performance when the linearity assumption doesn't hold.
 COSSO is one of the earliest smoothing spline models that consider sparsity-smoothness penalty.
@@
 \textit{mgcv} is one of the most popular models for nonlinear effect interpolation and prediction.
-Nevertheless, \bg{mgcv doesn't support analyses when the number of parameters exceeds the sample size.} \bg{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses an MCMC algorithm for model fitting.} SB-GAM is the first spike-and-slab LASSO GAM.
+Nevertheless, mgcv doesn't support analyses when the number of parameters exceeds the sample size.
+Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses an MCMC algorithm for model fitting.
+SB-GAM is the first spike-and-slab LASSO GAM.
 We implement linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive models with automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R package \texttt{sparseGAM 1.0}.
@@
 We generate 50 independent pairs of training and testing datasets to evaluate the prediction and variable selection performance of the chosen models, where training datasets are used to fit the models and testing datasets are used to calculate metrics of interest.
-\bg{In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
-This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.}
+In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
+This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.
 
 To evaluate the predictive performance of the models, the statistics, $R^2$ for Gaussian model and AUC for binomial model calculated based on the testing datasets, are averaged across 50 simulations.
-\bg{To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
+To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
 \begin{align*}
@@
 For the methods that don't automatically achieve variable selection, we set the alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictor are selected in the models.}
+For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictor are selected in the models.
 
-\subsection{\bg{Simulation Results}}
-\subsubsection{\bg{Prediction Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are nonlinear,} the predictive performances have a consistent pattern across the two distributions of outcomes.
+\subsection{Simulation Results}
+\subsubsection{Prediction Performance}
+Among the set of simulations where the functional forms of the predictors are nonlinear, the predictive performances have a consistent pattern across the two distributions of outcomes.
 For simplicity, we use Gaussian simulations to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus} and \ref{tab:bin_auc} for detailed statistics.
-\bg{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
+The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
 Specifically, BHAM shows great improvement over COSSO methods, resulting in a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) improvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
@@
 As previously hypothesized, the linear LASSO model predicts less accurate than other flexible models across all scenarios; mgcv performs extremely well in low-dimensional cases ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size approaches the number of coefficients to estimate ($p=50$), which leads to bad performance.} Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has an extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
+To note, mgcv fits models but fails to converge within the default number of iterations when the sample size approaches the number of coefficients to estimate ($p=50$), which leads to bad performance.
+Even though SB-GAM has slight prediction advantage over the proposed model in high-dimensional situations, the BHAM has an extreme computational advantage over SB-GAM, resulting median (IRT) 64\% (39\%) reduction in computation time (measured in seconds) for Gaussian simulations, without sacrificing much of the prediction accuracy (see Table \ref{tab:time_sim}).
 
-\bg{We also examine the prediction performance when the functional form of predictors is linear, see Supporting Information Table S1, and S2.
+We also examine the prediction performance when the functional form of predictors is linear, see Supporting Information Table S1, and S2.
 The proposed model, BHAM, has a similar performance as the linear LASSO model regardless of the distribution.
@@
 spikeSlabGAM has a similar prediction performance to BHAM.
-Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
+Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.
 
-\bg{\subsubsection{Variable Selection Performance}}
-\bg{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimensions and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
+\subsubsection{Variable Selection Performance}
+Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimensions and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
 The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that the model selected a small subset among all the active predictions.
@@
 The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations matches with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
+The variable selection performance for linear simulations matches with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.
 
-\bg{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
+Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
 Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
@@
 This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
-We defer an in-depth discussion to Section \ref{sec:concl}.}
+We defer an in-depth discussion to Section \ref{sec:concl}.
 

### 2022-05-16 Manuscript/04-Real_Data.Rmd
remove revision highlights in red

@@
 We demonstrate the improved prediction performance compared to the other Bayesian hierarchical additive model, SB-GAM [@Bai2021], while being computationally efficient (see Table \ref{tab:time_real_data}).
-\bg{BHAM requires roughly 10\% of the computation time of SB-GAM to fit models.}
+BHAM requires roughly 10\% of the computation time of SB-GAM to fit models.
 
@@
 There are initially 6796 features in the dataset, which is too large to be practically meaningful to analyze.
-Hence, we choose the top 200 features \bg{with the largest variance}.
+Hence, we choose the top 200 features with the largest variance.
 We use 5-knot spline additive models for binary outcome using two different models, the proposed BHAM and the SB-GAM.
@@
 BHAM obtains superior AUC, Brier score, and misclassification error in the out-of-bag samples compared to SB-GAM (see Table \ref{tab:ECB_res}).
-\bg{We plot the 33 features included in the final BHAM model in Figure \ref{fig:ECB_fig}.}
+We plot the 33 features included in the final BHAM model in Figure \ref{fig:ECB_fig}.
 
@@
 In the dataset, 765 metabolites in baseline plasma collected were profiled using liquid chromatography mass spectrometry.
-Quality control and natural log transformation \bg{were previously} performed and documented \bg{by the study publishing team \cite{Bihlmeyer2021}}.
+Quality control and natural log transformation were previously performed and documented by the study publishing team \cite{Bihlmeyer2021}.
 The outcome of interest is standardized percent change in insulin resistance, and hence modeled using a Gaussian model.

### 2022-05-16 Manuscript/05-Discussion.Rmd
remove revision highlights in red

@@
 \label{sec:concl}
-\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeling.
+In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeling.
 In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smooth functions and derive a scalable EM-CD algorithm for model fitting.
@@
 In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
-When serving the purpose of variable selection, trade-offs exist among methods of comparison.} We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
+When serving the purpose of variable selection, trade-offs exist among methods of comparison.
+We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM).
 To maximize the flexibility of smooth function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smooth functions are limited to the default ones.
@@
 Bai provides the theoretical proof for the consistency of variable selection using group spike-and-slab LASSO prior.
-\bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
+The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
 Moreover, the proposed model can easily generalize to other families of priors and smooth functions if desired.
-Not focused in this manuscript, the generalization is described in the Supporting Information.}
+Not focused in this manuscript, the generalization is described in the Supporting Information.
 
-\bg{During designing and analyzing the simulation study, we made couple of interesting observations.
+During designing and analyzing the simulation study, we made couple of interesting observations.
 First of all, variable selection is a delicate topic in the context of predictive modeling.
@@
 Secondly, we find the performance of the proposed model is more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
-Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend fine-tuning the model with a granular sequence of $s_0$ for performance improvement.}
+Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend fine-tuning the model with a granular sequence of $s_0$ for performance improvement.
 
@@
 
-Our future efforts are direct to \bg{uncertainty inference of the proposed model}, survival analysis and integrative analysis.
-\bg{Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
+Our future efforts are direct to uncertainty inference of the proposed model, survival analysis and integrative analysis.
+Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
 We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
@@
 The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
-Due to the space limit, technical details will be explained in a future manuscript.} While the proposed model addresses a great deal of analytic problems, analyzing the time-to-event outcome remains unsolved.
+Due to the space limit, technical details will be explained in a future manuscript.
+While the proposed model addresses a great deal of analytic problems, analyzing the time-to-event outcome remains unsolved.
 A naive approach would be convert a time-to-event outcome to a Poisson outcome following Whitehead [@Whitehead1980].

### 2022-05-16 Manuscript/SS_GAM.tex
* Remove useless markdowns * Correct smoothing penalty equation
remove revision highlights in red

@@
 \newcommand{\simiid}{\overset{\text{iid}}{\sim}}
-\newcommand{\bg}[1]{\textcolor{red}{#1}}
 
@@
 relevant to the data generation process that one can draw inferences
-from. In addition, how to generalize these \textcolor{red}{"black box"}
-models to the high-dimensional setting remains unclear.
-
-\textcolor{red}{For their straightforward interpretation and flexibility, nonparametric regression models serve as great alternatives to the "black-box" models in prediction and variable selection}.
-Among those, generalized additive models (GAMs), proposed in the seminal
-work of Hastie and Tibshirani \citep{Hastie1987}, grew to be one of the
-most popular modeling tools. In a GAM, the response variable, which is
-assumed to follow some exponential family distribution, can be modeled
-with the summation of smooth functions. Nevertheless, the classical GAMs
-cannot fulfill the increasing analytic demands for high-dimensional data
+from. In addition, how to generalize these ``black box'' models to the
+high-dimensional setting remains unclear.
+
+For their straightforward interpretation and flexibility, nonparametric
+regression models serve as great alternatives to the ``black-box''
+models in prediction and variable selection. Among those, generalized
+additive models (GAMs), proposed in the seminal work of Hastie and
+Tibshirani \citep{Hastie1987}, grew to be one of the most popular
+modeling tools. In a GAM, the response variable, which is assumed to
+follow some exponential family distribution, can be modeled with the
+summation of smooth functions. Nevertheless, the classical GAMs cannot
+fulfill the increasing analytic demands for high-dimensional data
 analysis.
@@
 
-\textcolor{red}{We propose a novel Bayesian hierarchical generalized additive model (BHAM) for outcome prediction in the context of high-dimensional data analysis.
-Specifically, we incorporate smoothing penalties, derived from the smoothing spline literature \cite{Wood2017}, via reparameterization of smooth functions to avoid excessive shrinkage on the bases.
-Smoothing penalties were also previously used in the spike-and-slab GAM \cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
-We then impose a new two-part spike-and-slab LASSO prior to address the signal sparsity.
-In addition, a scalable optimization-based algorithm, EM-Coordinate Descent (EM-CD) algorithm, is developed.
-While the primary focus of this model is to improve prediction, the proposed model also provides utility in functional selection.
-Notably, the two-part prior that follows the effect hierarchy principle motivates a bi-level selection, rendering one of three possibilities for each predictor: no effect, only linear effect, or linear and nonlinear effects.
-The proposed model is implemented in a publicly available R package \texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.}
+We propose a novel Bayesian hierarchical generalized additive model
+(BHAM) for outcome prediction in the context of high-dimensional data
+analysis. Specifically, we incorporate smoothing penalties, derived from
+the smoothing spline literature \cite{Wood2017}, via reparameterization
+of smooth functions to avoid excessive shrinkage on the bases. Smoothing
+penalties were also previously used in the spike-and-slab GAM
+\cite{Scheipl2012} and the sparsity-smoothness penalty \cite{Meier2009}.
+We then impose a new two-part spike-and-slab LASSO prior to address the
+signal sparsity. In addition, a scalable optimization-based algorithm,
+EM-Coordinate Descent (EM-CD) algorithm, is developed. While the primary
+focus of this model is to improve prediction, the proposed model also
+provides utility in functional selection. Notably, the two-part prior
+that follows the effect hierarchy principle motivates a bi-level
+selection, rendering one of three possibilities for each predictor: no
+effect, only linear effect, or linear and nonlinear effects. The
+proposed model is implemented in a publicly available R package
+\texttt{BHAM} via \url{https://github.com/boyiguo1/BHAM}.
 
@@
 
-\textcolor{red}{We assume the response variable, $Y$, follows an exponential family distribution with density function $f(y)$, mean $\mu$ and dispersion parameter $\phi$.
-The mean of the response variable can be modeled as the summation of smooth functions, $B_j(\cdot), j = 1, \dots, p$, of a given $p$-dimensional vector of predictors $\boldsymbol{x}$, written as
-\begin{equation}\label{eq:gam}
+We assume the response variable, \(Y\), follows an exponential family
+distribution with density function \(f(y)\), mean \(\mu\) and dispersion
+parameter \(\phi\). The mean of the response variable can be modeled as
+the summation of smooth functions, \(B_j(\cdot), j = 1, \dots, p\), of a
+given \(p\)-dimensional vector of predictors \(\boldsymbol{x}\), written
+as \begin{equation}\label{eq:gam}
  E(Y|\boldsymbol{x}) = g^{-1}(\beta_0 + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(\beta_0 + \sum\limits^p_{j=1} \boldsymbol{\beta}_j^T \boldsymbol{X}_j),
-\end{equation}
-where $g^{-1}(\cdot)$ is the inverse of a monotonic link function. Given $n$ data points $\{y_i, \boldsymbol{x}_i\}^n_{i=1}$,}
-the data distribution is expressed as \begin{equation}
+\end{equation} where \(g^{-1}(\cdot)\) is the inverse of a monotonic
+link function. Given \(n\) data points
+\(\{y_i, \boldsymbol{x}_i\}^n_{i=1}\), the data distribution is
+expressed as \begin{equation}
 f(\boldsymbol{Y} = \boldsymbol{y}| \boldsymbol{\beta}, \phi) = \prod\limits^n_{i=1}f( Y = y_i|\boldsymbol{\beta}, \phi).\nonumber
@@
 \(\beta, X\). The matrix transposing operation is denoted with a
-superscript \(^T\).
-\textcolor{red}{To note, the proposed model can include parametric forms of variables in the model, and hence considers general linear models and semiparametric regression models as special cases.}
-
-\subsection{smooth function Reparameterization}
+superscript \(^T\). To note, the proposed model can include parametric
+forms of variables in the model, and hence considers general linear
+models and semiparametric regression models as special cases.
+
+\subsection{Smooth Function Reparameterization}
 
@@
 \begin{equation}
-  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,\nonumber
+  \text{pen}\left[B_j(x)\right] = \lambda_j \int B^{\prime\prime}_j(x)^2dx = \lambda_j \boldsymbol{\beta}_j^T \boldsymbol{S}_j \boldsymbol{\beta}_j ,\nonumber
 \end{equation} where \(\boldsymbol{S}_j\) is a known smoothing penalty
@@
 function can be modeled as \(B_j(x_j) = x_j\) with the smoothing penalty
-matrix \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\).
-\textcolor{red}{Unlike previous regularized methods that either ignore the smoothing penalty completely or restrain the smoothing penalty as a component of sparse penalty which leads to a more restrictive solution, we consider an additional mechanism in pair with the proposed prior (described in Section \ref{sec:method_prior}) to address the smoothness and sparsity in signals such that the locally adaptive nature of the smoothing penalty retains.}
+matrix \(\boldsymbol{S}_j = \begin{bmatrix}0\end{bmatrix}\). Unlike
+previous regularized methods that either ignore the smoothing penalty
+completely or restrain the smoothing penalty as a component of sparse
+penalty which leads to a more restrictive solution, we consider an
+additional mechanism in pair with the proposed prior (described in
+Section \ref{sec:method_prior}) to address the smoothness and sparsity
+in signals such that the locally adaptive nature of the smoothing
+penalty retains.
 
 Marra and Wood \citep{Marra2011} proposed a reparameterization procedure
-\textcolor{red}{to factor the smoothing penalty into the design matrix of each smooth function}.
-Given that the smoothing penalty matrix \(\boldsymbol{S}_j\) is
-symmetric and positive semi-definite for the univariate smooth
+to factor the smoothing penalty into the design matrix of each smooth
+function. Given that the smoothing penalty matrix \(\boldsymbol{S}_j\)
+is symmetric and positive semi-definite for the univariate smooth
 functions, we eigendecompose the penalty matrix
@@
 commonly used models in high-dimensional data analysis for its utility
-in outcome prediction and variable selection.
-\textcolor{red}{Among all the spike-and-slab priors, the spike-and-slab LASSO (SSL) prior \cite{Rockova2018b, Rockova2018} is one of the most popular choices because it's highly scalable.
-The SSL prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
-\begin{equation} 
+in outcome prediction and variable selection. Among all the
+spike-and-slab priors, the spike-and-slab LASSO (SSL) prior
+\cite{Rockova2018b, Rockova2018} is one of the most popular choices
+because it's highly scalable. The SSL prior is composed of two double
+exponential distributions with mean 0 and different dispersion
+parameters, \(0 < s_0 < s_1\), mathematically, \begin{equation} 
 \beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.\nonumber
-\end{equation}
-} The latent binary variable \(\gamma \in \{0,1\}\) indicates whether a
-variable \(x\) is included in the model, while the dispersion parameters
-\(s_0\) and \(s_1\) control the shrinkage of the coefficient. Given that
-both double exponential distributions have a mean of 0 and the latent
-indicator \(\gamma\) can only take the value of 0 or 1, the mixture
-double exponential distribution can be formulated as one single double
-exponential density, \begin{equation} \label{eq:ssl}
+\end{equation} The latent binary variable \(\gamma \in \{0,1\}\)
+indicates whether a variable \(x\) is included in the model, while the
+dispersion parameters \(s_0\) and \(s_1\) control the shrinkage of the
+coefficient. Given that both double exponential distributions have a
+mean of 0 and the latent indicator \(\gamma\) can only take the value of
+0 or 1, the mixture double exponential distribution can be formulated as
+one single double exponential density, \begin{equation} \label{eq:ssl}
 \beta | \gamma \sim DE(0, (1-\gamma)s_0 + \gamma s_1), 0 < s_0 < s_1.
-\end{equation}
-\textcolor{red}{Compared to other priors for high-dimensional data analysis, SSL has the following advantages.
-First of all, the SSL prior provides a locally adaptive shrinkage when estimating the coefficients.
-Secondly, the SSL prior encourages a sparse solution, making variable selection straightforward.
-Thirdly, the SSL prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.
-We defer to Bai et al.
-\cite{Bai2021Review} for a detailed discussion.}
+\end{equation} Compared to other priors for high-dimensional data
+analysis, SSL has the following advantages. First of all, the SSL prior
+provides a locally adaptive shrinkage when estimating the coefficients.
+Secondly, the SSL prior encourages a sparse solution, making variable
+selection straightforward. Thirdly, the SSL prior motivates a scalable
+algorithm, the EM-CD algorithm, for model fitting, and hence is more
+feasible for high-dimensional data analysis. We defer to Bai et
+al.~\cite{Bai2021Review} for a detailed discussion.
 
@@
 
-\textcolor{red}{The proposed two-part SSL prior, particularly the group SSL prior of the nonlinear space coefficients, differs from previous group SSL priors \cite{Tang2018, Tang2019}, as the proposed prior follows the effect hierarchy principle.
-Effect hierarchy refers to the principle that "lower-order effects are more likely to be active than higher-order effects" defined by Chipman\cite{chipman2006prior}.
-To implement, we consider the shared latent indicator of nonlinear coefficients $\gamma^*_j$}
-depends on the value of the linear space latent indicator \(\gamma_j\),
-while both latent indicators \(\gamma_j\) and \(\gamma^*_j\) follow a
-Bernoulli distribution. While the probability of including the linear
-effect is \(\theta_j\), the probability of including the nonlinear
-effect is \(\textcolor{red}{\gamma_{j}}\theta_j\). \[
+The proposed two-part SSL prior, particularly the group SSL prior of the
+nonlinear space coefficients, differs from previous group SSL priors
+\cite{Tang2018, Tang2019}, as the proposed prior follows the effect
+hierarchy principle. Effect hierarchy refers to the principle that
+``lower-order effects are more likely to be active than higher-order
+effects'' defined by Chipman\cite{chipman2006prior}. To implement, we
+consider the shared latent indicator of nonlinear coefficients
+\(\gamma^*_j\) depends on the value of the linear space latent indicator
+\(\gamma_j\), while both latent indicators \(\gamma_j\) and
+\(\gamma^*_j\) follow a Bernoulli distribution. While the probability of
+including the linear effect is \(\theta_j\), the probability of
+including the nonlinear effect is \(\gamma_{j}\theta_j\). \[
 \begin{aligned}
 &\gamma_{j} | \theta_j \sim Bin(1, \theta_j) & & 
-&\gamma_{j}^*| \textcolor{red}{\gamma_{j}, } \theta_j \sim Bin(1, \textcolor{red}{\gamma_{j}}\theta_j).
+&\gamma_{j}^*| \gamma_{j}, \theta_j \sim Bin(1, \gamma_{j}\theta_j).
 \end{aligned}
-\]
-\textcolor{red}{This is, when the linear effect is not selected, the probability of including the nonlinear effect drops from $\theta_j$ to 0.
-For computational convenience, we analytically integrate $\gamma_j$ out such that $\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)$ (see the derivation in the Supporting Information).
-}
-
-\textcolor{red}{To allow the shrinkage to self-adapt to the sparsity and smoothing pattern of the data, }we
-further specify the parameter \(\theta_j\) follows a beta distribution
-with given shape parameters \(a\) and \(b\), \[
+\] This is, when the linear effect is not selected, the probability of
+including the nonlinear effect drops from \(\theta_j\) to 0. For
+computational convenience, we analytically integrate \(\gamma_j\) out
+such that \(\gamma_{j}^*| \theta_j \sim Bin(1, \theta_j^2)\) (see the
+derivation in the Supporting Information).
+
+To allow the shrinkage to self-adapt to the sparsity and smoothing
+pattern of the data, we further specify the parameter \(\theta_j\)
+follows a beta distribution with given shape parameters \(a\) and \(b\),
+\[
 \theta_j \sim Beta(a, b).
 \] The beta distribution is a conjugate prior to the binomial
-distribution and hence provides some computation convenience.
-\textcolor{red}{Having a prior distribution of $\theta_j$ enables the proposed prior to inherit the selective shrinkage property and self-adaptivity \cite{Bai2021Review} from the classical SSL prior.
-In other words, when a smooth function is significant, the coefficients of the smooth function escape the overall shrinkage and produce a more accurate estimate, particularly in pair with the smoothing penalty implicitly addressed via the reparameterization.
-Meanwhile, the hyper prior encourages information borrowing across coordinates and hence automatic adjust for different levels of sparsity.}
-Hereafter, we refer to the Bayesian hierarchical generalized additive
-models with the two-part spike-and-slab LASSO prior as BHAM, and
-visually presented in Figure \ref{fig:SSprior}.
+distribution and hence provides some computation convenience. Having a
+prior distribution of \(\theta_j\) enables the proposed prior to inherit
+the selective shrinkage property and self-adaptivity
+\cite{Bai2021Review} from the classical SSL prior. In other words, when
+a smooth function is significant, the coefficients of the smooth
+function escape the overall shrinkage and produce a more accurate
+estimate, particularly in pair with the smoothing penalty implicitly
+addressed via the reparameterization. Meanwhile, the hyper prior
+encourages information borrowing across coordinates and hence automatic
+adjust for different levels of sparsity. Hereafter, we refer to the
+Bayesian hierarchical generalized additive models with the two-part
+spike-and-slab LASSO prior as BHAM, and visually presented in Figure
+\ref{fig:SSprior}.
 
@@
 
-\textcolor{red}{Despite the advantage to estimate posterior densities, using MCMC algorithms to fit the proposed model is computationaly prohibited and not feasible for high-dimensional data.
-Previous research shows the computation performance of MCMC algorithms for spike-and-slab models is bottlenecked for medium-sized data ($p$=25) \cite{George1997}, and substantially slows as $p$ increases modestly in the GAM context \cite{Scheipl2013}.
-Hence, we consider the optimization algorithms that focus on the maximum a posteriori estimates at the cost of posterior inference.
-Specifically, we
-extend the EM-Coordinate Descent (EM-CD) algorithm to fit BHAMs.
-Similar to the EMVS algorithm \cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is based on the expectation-maximization (EM) algorithm, integrating the Coordinate Descent algorithm in each iterative step to find the posterior mode.
-The EM-CD algorithm has been well adapted in generalized linear models \cite{Tang2017a}, Cox proportional hazards models \cite{Tang2017}, and their grouped counterparts \cite{Tang2018, Tang2019}.
-The EM-CD algorithm provides deterministic solutions, which becomes a popular property for reproducible research.}
+Despite the advantage to estimate posterior densities, using MCMC
+algorithms to fit the proposed model is computationaly prohibited and
+not feasible for high-dimensional data. Previous research shows the
+computation performance of MCMC algorithms for spike-and-slab models is
+bottlenecked for medium-sized data (\(p\)=25) \cite{George1997}, and
+substantially slows as \(p\) increases modestly in the GAM context
+\cite{Scheipl2013}. Hence, we consider the optimization algorithms that
+focus on the maximum a posteriori estimates at the cost of posterior
+inference. Specifically, we extend the EM-Coordinate Descent (EM-CD)
+algorithm to fit BHAMs. Similar to the EMVS algorithm
+\cite{Rockova2014a} for spike-and-slab models, the EM-CD algorithm is
+based on the expectation-maximization (EM) algorithm, integrating the
+Coordinate Descent algorithm in each iterative step to find the
+posterior mode. The EM-CD algorithm has been well adapted in generalized
+linear models \cite{Tang2017a}, Cox proportional hazards models
+\cite{Tang2017}, and their grouped counterparts
+\cite{Tang2018, Tang2019}. The EM-CD algorithm provides deterministic
+solutions, which becomes a popular property for reproducible research.
 
@@
 M-step, we find the parameters \(\Theta^{(t)}\) that maximize
-\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\).
-\textcolor{red}{The parenthesized subscription $^{(t)}$ denotes the parameter estimation at the $t$th iteration.}
-The E- and M- steps are iterated until the algorithm converges.
+\(E\log f(\Theta, \boldsymbol{\gamma}| \textbf{y}, \textbf{X})\). The
+parenthesized subscription \(^{(t)}\) denotes the parameter estimation
+at the \(t\)th iteration. The E- and M- steps are iterated until the
+algorithm converges.
 
@@
 \(\text{Pr}(\gamma_{j} = 0|\theta_j) = 1-\theta_j\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2}\),
-\(\textcolor{red}{\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j}\),
+\(\text{Pr}(\gamma_{j}^*= 1|\theta_j) = \theta_j^2\),
+\(\text{Pr}(\gamma_{j}^*= 0|\theta_j) = 1-\theta^2_j\),
 \(f(\beta|\gamma=1, s_1) = \text{DE}(\beta|0 , s_1)\),
@@
 model. A measure of preference calculated with cross-validations (CV),
-e.g.~deviance
-\textcolor{red}{(defined as model log-likelihood times -2, $-2\log f(\boldsymbol{y}|\boldsymbol{\hat\beta}, \hat\phi)$)},
-area under the curve (AUC), mean squared error, can be used to
-facilitate the selection of a final model. The procedure is similar to
-the LASSO implementation in the widely used R package \texttt{glmnet},
-which quickly fits LASSO models over a list of candidate values of
-\(\lambda\) and gives a sequence of models for users to choose from.
+e.g.~deviance (defined as model log-likelihood times -2,
+\(-2\log f(\boldsymbol{y}|\boldsymbol{\hat\beta}, \hat\phi)\)), area
+under the curve (AUC), mean squared error, can be used to facilitate the
+selection of a final model. The procedure is similar to the LASSO
+implementation in the widely used R package \texttt{glmnet}, which
+quickly fits LASSO models over a list of candidate values of \(\lambda\)
+and gives a sequence of models for users to choose from.
 
@@
 
-In this section, we compare the performance of the proposed model to
-\textcolor{red}{six alternative models: linear LASSO models,} component
-selection and smoothing operator (COSSO) \citep{Zhang2006GAM}, adaptive
-COSSO \citep{Storlie2011}, generalized additive models with automatic
+In this section, we compare the performance of the proposed model to six
+alternative models: linear LASSO models, component selection and
+smoothing operator (COSSO) \citep{Zhang2006GAM}, adaptive COSSO
+\citep{Storlie2011}, generalized additive models with automatic
 smoothing (referred to as \textit{mgcv} hereafter)\citep{Wood2011},
-\textcolor{red}{spike-and-slab GAM \cite{Scheipl2012}}, and SB-GAM
-\citep{Bai2021}.
-\textcolor{red}{We use linear LASSO models as the benchmark, examining the performance when the linearity assumption doesn't hold}.
-COSSO is one of the earliest smoothing spline models that consider
-sparsity-smoothness penalty. Adaptive COSSO improved upon COSSO by using
-adaptive weight for penalties such that the penalty of each functional
-component is different for extra flexibility. \textit{mgcv} is one of
-the most popular models for nonlinear effect interpolation and
-prediction. Nevertheless,
-\textcolor{red}{mgcv doesn't support analyses when the number of parameters exceeds the sample size.}
-\textcolor{red}{Spike-and-slab GAM employs a spike-and-slab prior for GAM and uses an MCMC algorithm for model fitting.}
-SB-GAM is the first spike-and-slab LASSO GAM. We implement linear LASSO
-model with R package \texttt{glmnet 4.1-2}, COSSO and adaptive COSSO
-with R package \texttt{cosso 2.1-1}, generalized additive models with
-automatic smoothing with R package \texttt{mgcv 1.8-31}, spike-and-slab
-GAM with R package \texttt{spikeSlabGAM 1.1-15}, and SB-GAM with R
-package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do not provide
-flexibility to define smooth functions, and hence use the default
-choices; mgcv, spikeSlabGAM and the proposed model allow customized
-smooth functions and we choose the cubic regression spline. We control
-the dimensionality of each smooth function, 10 bases, for all different
-choices of smooth functions. We use a 5-fold CV with the default
-selection criteria to select the final model for linear LASSO model,
-COSSO models, SB-GAM and the proposed model. Twenty default candidates
-of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in SB-GAM) are
-examined for SB-GAM and the proposed model that allow user specification
-of tuning candidates. All computation was conducted on a
+spike-and-slab GAM \cite{Scheipl2012}, and SB-GAM \citep{Bai2021}. We
+use linear LASSO models as the benchmark, examining the performance when
+the linearity assumption doesn't hold. COSSO is one of the earliest
+smoothing spline models that consider sparsity-smoothness penalty.
+Adaptive COSSO improved upon COSSO by using adaptive weight for
+penalties such that the penalty of each functional component is
+different for extra flexibility. \textit{mgcv} is one of the most
+popular models for nonlinear effect interpolation and prediction.
+Nevertheless, mgcv doesn't support analyses when the number of
+parameters exceeds the sample size. Spike-and-slab GAM employs a
+spike-and-slab prior for GAM and uses an MCMC algorithm for model
+fitting. SB-GAM is the first spike-and-slab LASSO GAM. We implement
+linear LASSO model with R package \texttt{glmnet 4.1-2}, COSSO and
+adaptive COSSO with R package \texttt{cosso 2.1-1}, generalized additive
+models with automatic smoothing with R package \texttt{mgcv 1.8-31},
+spike-and-slab GAM with R package \texttt{spikeSlabGAM 1.1-15}, and
+SB-GAM with R package \texttt{sparseGAM 1.0}. COSSO models and SB-GAM do
+not provide flexibility to define smooth functions, and hence use the
+default choices; mgcv, spikeSlabGAM and the proposed model allow
+customized smooth functions and we choose the cubic regression spline.
+We control the dimensionality of each smooth function, 10 bases, for all
+different choices of smooth functions. We use a 5-fold CV with the
+default selection criteria to select the final model for linear LASSO
+model, COSSO models, SB-GAM and the proposed model. Twenty default
+candidates of tuning parameters (\(s_0\) in BHAM, \(\lambda_0\) in
+SB-GAM) are examined for SB-GAM and the proposed model that allow user
+specification of tuning candidates. All computation was conducted on a
 high-performance 64-bit Linux platform with 48 cores of 2.70GHz
@@
 where training datasets are used to fit the models and testing datasets
-are used to calculate metrics of interest.
-\textcolor{red}{In addition, we consider the data generating process where all functional forms of the predictors are linear while keeping the rest of simulation parameters the same.
-This additional set of linear simulations is designed to investigate the flexibility of the proposed model when nonlinear assumptions are not met.}
+are used to calculate metrics of interest. In addition, we consider the
+data generating process where all functional forms of the predictors are
+linear while keeping the rest of simulation parameters the same. This
+additional set of linear simulations is designed to investigate the
+flexibility of the proposed model when nonlinear assumptions are not
+met.
 
@@
 \(R^2\) for Gaussian model and AUC for binomial model calculated based
-on the testing datasets, are averaged across 50 simulations.
-\textcolor{red}{To evaluate the variable selection performance of the models, we record the set of variables each method selects and calculate the averaged positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient (MCC),
-\begin{align*}
+on the testing datasets, are averaged across 50 simulations. To evaluate
+the variable selection performance of the models, we record the set of
+variables each method selects and calculate the averaged positive
+predictive value (precision), true positive rate (recall), and Matthews
+correlation coefficient (MCC), \begin{align*}
 &\text{precision} = \frac{TP}{PP}\\
@@
 &\text{MCC} = \frac{TP\times TN - FP \times FN}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}},
-\end{align*}
-where TP, TN, FP, FN, and PP are true positives, true negatives, false positives, false negatives, and predicted positives respectively.
-For the methods that don't automatically achieve variable selection, we set the alpha level at 0.05 for \textit{mgcv} that relies on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM given the marginal inclusion probabilities.
-For the two methods, BHAM and spikeSlabGAM, that are capable of bi-level selection, we record the probability that the linear and nonlinear components of each predictor are selected in the models.}
-
-\subsection{\textcolor{red}{Simulation Results}}
-\subsubsection{\textcolor{red}{Prediction Performance}}
-
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear,}
-the predictive performances have a consistent pattern across the two
-distributions of outcomes. For simplicity, we use Gaussian simulations
-to exemplify the improvement of BHAM and defer to Tables \ref{tab:gaus}
-and \ref{tab:bin_auc} for detailed statistics.
-\textcolor{red}{The proposed model, BHAM, predicts as good as, if not better than, other high dimensional additive models.
-Specifically, BHAM shows great improvement over COSSO methods, resulting in a median (interquartile range, IRT) 31\% (131\%) and 20\% (129\%) improvement over COSSO and adaptive COSSO in $R^2$ statistics respectively.
-The improvement over the spikeSlabGAM model is moderate, resulting in a median (IRT) 6\% (10\%) improvement in $R^2$.
-When comparing to SB-GAM, BHAM performs better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional cases ($p=4,10$), and equally good or slightly worse (median (IRT) 1\% (9\%) improvement in R2: ) in high-dimensional cases ($p=50,100, 200$).
-As previously hypothesized, the linear LASSO model predicts less accurate than other flexible models across all scenarios; mgcv performs extremely well in low-dimensional cases ($p = 4, 10$), and deteriorates as the dimensionality increases until not applicable.
-To note, mgcv fits models but fails to converge within the default number of iterations when the sample size approaches the number of coefficients to estimate ($p=50$), which leads to bad performance.}
+\end{align*} where TP, TN, FP, FN, and PP are true positives, true
+negatives, false positives, false negatives, and predicted positives
+respectively. For the methods that don't automatically achieve variable
+selection, we set the alpha level at 0.05 for \textit{mgcv} that relies
+on hypothesis testing, and a soft-threshold at 0.5 for spikeSlabGAM
+given the marginal inclusion probabilities. For the two methods, BHAM
+and spikeSlabGAM, that are capable of bi-level selection, we record the
+probability that the linear and nonlinear components of each predictor
+are selected in the models.
+
+\subsection{Simulation Results}
+\subsubsection{Prediction Performance}
+
+Among the set of simulations where the functional forms of the
+predictors are nonlinear, the predictive performances have a consistent
+pattern across the two distributions of outcomes. For simplicity, we use
+Gaussian simulations to exemplify the improvement of BHAM and defer to
+Tables \ref{tab:gaus} and \ref{tab:bin_auc} for detailed statistics. The
+proposed model, BHAM, predicts as good as, if not better than, other
+high dimensional additive models. Specifically, BHAM shows great
+improvement over COSSO methods, resulting in a median (interquartile
+range, IRT) 31\% (131\%) and 20\% (129\%) improvement over COSSO and
+adaptive COSSO in \(R^2\) statistics respectively. The improvement over
+the spikeSlabGAM model is moderate, resulting in a median (IRT) 6\%
+(10\%) improvement in \(R^2\). When comparing to SB-GAM, BHAM performs
+better (median (IRT) 13\% (8\%) improvement in R2) in lower dimensional
+cases (\(p=4,10\)), and equally good or slightly worse (median (IRT) 1\%
+(9\%) improvement in R2: ) in high-dimensional cases
+(\(p=50,100, 200\)). As previously hypothesized, the linear LASSO model
+predicts less accurate than other flexible models across all scenarios;
+mgcv performs extremely well in low-dimensional cases (\(p = 4, 10\)),
+and deteriorates as the dimensionality increases until not applicable.
+To note, mgcv fits models but fails to converge within the default
+number of iterations when the sample size approaches the number of
+coefficients to estimate (\(p=50\)), which leads to bad performance.
 Even though SB-GAM has slight prediction advantage over the proposed
@@
 
-\textcolor{red}{We also examine the prediction performance when the functional form of predictors is linear, see Supporting Information Table S1, and S2.
-The proposed model, BHAM, has a similar performance as the linear LASSO model regardless of the distribution.
-This observation demonstrates that BHAM is a flexible model, and has good prediction performance regardless of the underlying functional form of predictors.
-spikeSlabGAM has a similar prediction performance to BHAM.
-Surprisingly, SB-GAM doesn't perform well in high-dimensional Gaussian outcome scenarios.}
-
-\textcolor{red}{\subsubsection{Variable Selection Performance}}
-\textcolor{red}{Among the set of simulations where the functional forms of the predictors are nonlinear, the proposed model, BHAM, has a consistent performance across different dimensions and distribution settings (See Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support Information Table S3 for binomial outcomes): being conservative.
-The symptoms of conservative variable selection are high precision and low recall, where high precision means that among all the selected variables, high percentage of them are true signals; low recall means that the model selected a small subset among all the active predictions.
-In other words, BHAM tends to select a smaller set of variables that are truly effective to the outcome.
-We want to note, the variable selection performance of BHAM is plummeted and not optimized when $p=200$.
-Upon further investigation, we discover it's because the generic sequence of $s_0$ used to tune the model doesn't contain the optimal value.
-Overall, among all the models examined, SB-GAM has the best performance, both high precision and high recall, and yields a high MCC.
-The performance of another Bayesian model, spikeSlabGAM deteriorates as the sparsity grows, particularly when ($p>50$), or for binomial outcomes.
-The variable selection performance for linear simulations matches with prediction performance: BHAM performs great among the Gaussian scenarios, while the performance of SB-GAM deteriorates.}
-
-\textcolor{red}{Among the high-dimensional methods of comparison, there are two methods that are capable to achieve bi-level selection, the proposed BHAM and spikeSlabGAM.
-Among the linear simulations, both methods can accurately select the linear components and have a drastically lowered probability, close to 0, to include the nonlinear component, as anticipated.
-Specifically, spikeSlabGAM have a smaller probability to include the nonlinear component in the model than BHAM.
-However, this advantage of spikeSlabGAM over BHAM is less obvious among the nonlinear simulations: spikeSlabGAM performs better than BHAM when selecting components of the functional forms that include only linear or nonlinear component, e.g.
-functional forms for $x_3$ and $x_4$.
-However, spikeSlabGAM inclines to ignore variables that have more complex function forms, e.g.
-function forms for $x_1$ and $x_2$.
-In contrast, BHAM is more likely to include them in the model.
-This trade-off is determined by the assumption implicitly reflected via the prior hierarchy.
-We defer an in-depth discussion to Section \ref{sec:concl}.}
+We also examine the prediction performance when the functional form of
+predictors is linear, see Supporting Information Table S1, and S2. The
+proposed model, BHAM, has a similar performance as the linear LASSO
+model regardless of the distribution. This observation demonstrates that
+BHAM is a flexible model, and has good prediction performance regardless
+of the underlying functional form of predictors. spikeSlabGAM has a
+similar prediction performance to BHAM. Surprisingly, SB-GAM doesn't
+perform well in high-dimensional Gaussian outcome scenarios.
+
+\subsubsection{Variable Selection Performance}
+
+Among the set of simulations where the functional forms of the
+predictors are nonlinear, the proposed model, BHAM, has a consistent
+performance across different dimensions and distribution settings (See
+Table \ref{tab:sim_gaus_var_select} for Gaussian outcomes, and Support
+Information Table S3 for binomial outcomes): being conservative. The
+symptoms of conservative variable selection are high precision and low
+recall, where high precision means that among all the selected
+variables, high percentage of them are true signals; low recall means
+that the model selected a small subset among all the active predictions.
+In other words, BHAM tends to select a smaller set of variables that are
+truly effective to the outcome. We want to note, the variable selection
+performance of BHAM is plummeted and not optimized when \(p=200\). Upon
+further investigation, we discover it's because the generic sequence of
+\(s_0\) used to tune the model doesn't contain the optimal value.
+Overall, among all the models examined, SB-GAM has the best performance,
+both high precision and high recall, and yields a high MCC. The
+performance of another Bayesian model, spikeSlabGAM deteriorates as the
+sparsity grows, particularly when (\(p>50\)), or for binomial outcomes.
+The variable selection performance for linear simulations matches with
+prediction performance: BHAM performs great among the Gaussian
+scenarios, while the performance of SB-GAM deteriorates.
+
+Among the high-dimensional methods of comparison, there are two methods
+that are capable to achieve bi-level selection, the proposed BHAM and
+spikeSlabGAM. Among the linear simulations, both methods can accurately
+select the linear components and have a drastically lowered probability,
+close to 0, to include the nonlinear component, as anticipated.
+Specifically, spikeSlabGAM have a smaller probability to include the
+nonlinear component in the model than BHAM. However, this advantage of
+spikeSlabGAM over BHAM is less obvious among the nonlinear simulations:
+spikeSlabGAM performs better than BHAM when selecting components of the
+functional forms that include only linear or nonlinear component,
+e.g.~functional forms for \(x_3\) and \(x_4\). However, spikeSlabGAM
+inclines to ignore variables that have more complex function forms,
+e.g.~function forms for \(x_1\) and \(x_2\). In contrast, BHAM is more
+likely to include them in the model. This trade-off is determined by the
+assumption implicitly reflected via the prior hierarchy. We defer an
+in-depth discussion to Section \ref{sec:concl}.
 
@@
 SB-GAM \citep{Bai2021}, while being computationally efficient (see Table
-\ref{tab:time_real_data}).
-\textcolor{red}{BHAM requires roughly 10\% of the computation time of SB-GAM to fit models.}
+\ref{tab:time_real_data}). BHAM requires roughly 10\% of the computation
+time of SB-GAM to fit models.
 
@@
 too large to be practically meaningful to analyze. Hence, we choose the
-top 200 features \textcolor{red}{with the largest variance}. We use
-5-knot spline additive models for binary outcome using two different
-models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
-the optimal tuning parameters of each framework with respect to the
-default selection criterion implemented in the software. Out-of-bag
-samples are used for prediction performance evaluation, where deviance,
-AUC, Brier score, defined as
+top 200 features with the largest variance. We use 5-knot spline
+additive models for binary outcome using two different models, the
+proposed BHAM and the SB-GAM. 10-Fold CV are used to choose the optimal
+tuning parameters of each framework with respect to the default
+selection criterion implemented in the software. Out-of-bag samples are
+used for prediction performance evaluation, where deviance, AUC, Brier
+score, defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and
@@
 misclassification error in the out-of-bag samples compared to SB-GAM
-(see Table \ref{tab:ECB_res}).
-\textcolor{red}{We plot the 33 features included in the final BHAM model in Figure \ref{fig:ECB_fig}.}
+(see Table \ref{tab:ECB_res}). We plot the 33 features included in the
+final BHAM model in Figure \ref{fig:ECB_fig}.
 
@@
 chromatography mass spectrometry. Quality control and natural log
-transformation \textcolor{red}{were previously} performed and documented
-\textcolor{red}{by the study publishing team \cite{Bihlmeyer2021}}. The
-outcome of interest is standardized percent change in insulin
-resistance, and hence modeled using a Gaussian model. After removing
-missing datapoints and addressing outliers in the data, there are
-\(p\)=237 features remaining in the analysis. 5-Knot spline additive
-models for the Gaussian outcome are constructed using two different
-models, the proposed BHAM and the SB-GAM. 10-Fold CV are used to choose
-the optimal tuning parameters of each framework with respect to the
-default selection criterion implemented in the software. Out-of-bag
-samples are used for prediction performance evaluation, where deviance,
-\(R^2\), mean squared error (MSE) defined as
-\(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\), and mean
-absolute error (MAE) defined as
+transformation were previously performed and documented by the study
+publishing team \cite{Bihlmeyer2021}. The outcome of interest is
+standardized percent change in insulin resistance, and hence modeled
+using a Gaussian model. After removing missing datapoints and addressing
+outliers in the data, there are \(p\)=237 features remaining in the
+analysis. 5-Knot spline additive models for the Gaussian outcome are
+constructed using two different models, the proposed BHAM and the
+SB-GAM. 10-Fold CV are used to choose the optimal tuning parameters of
+each framework with respect to the default selection criterion
+implemented in the software. Out-of-bag samples are used for prediction
+performance evaluation, where deviance, \(R^2\), mean squared error
+(MSE) defined as \(\frac{1}{n}\sum\limits^{n}_{i=1}(y_i - \hat y_i)^2\),
+and mean absolute error (MAE) defined as
 \(\frac{1}{n}\sum\limits^{n}_{i=1}|y_i - \hat y_i|\) are calculated.
@@
 
-\textcolor{red}{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modeling.
-In particular, we introduce a two-part spike-and-slab LASSO prior for reparameterized smooth functions and derive a scalable EM-CD algorithm for model fitting.
-The proposed model provides a new angle to address the excess shrinkage of smooth functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance.
-Th EM-CD algorithm, extended from previous spike-and-slab LASSO models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms, enhancing the scalability of spike-and-slab models.
-In addition, the two-part prior motivates the bi-level selection of predictors, selection of linear and nonlinear components.
-In the simulation study and real-data analyses, the proposed model demonstrates improvement in prediction and computational advantage when compared to the state-of-the-art models.
-When serving the purpose of variable selection, trade-offs exist among methods of comparison.}
-We implement the proposed model in an open-source R package
-\texttt{BHAM}, deposited at \url{https://github.com/boyiguo1/BHAM}. To
-maximize the flexibility of smooth function specification, we deploy the
-same programming grammar as in the state-of-the-art package
-\texttt{mgcv}, in contrast to previous tools where smooth functions are
-limited to the default ones. Ancillary functions are provided for model
-specification in high-dimensional settings, curve plotting and
-functional selection.
+In the paper, we described a novel high-dimensional generalized additive
+model with Bayesian hierarchical prior for the purpose of predictive
+modeling. In particular, we introduce a two-part spike-and-slab LASSO
+prior for reparameterized smooth functions and derive a scalable EM-CD
+algorithm for model fitting. The proposed model provides a new angle to
+address the excess shrinkage of smooth functions that is commonly
+vulnerable to previous regularized high-dimensional GAMs, and hence
+improves the predictive performance. Th EM-CD algorithm, extended from
+previous spike-and-slab LASSO models, provides a computationally
+efficient alternative to the computational prohibitive MCMC algorithms,
+enhancing the scalability of spike-and-slab models. In addition, the
+two-part prior motivates the bi-level selection of predictors, selection
+of linear and nonlinear components. In the simulation study and
+real-data analyses, the proposed model demonstrates improvement in
+prediction and computational advantage when compared to the
+state-of-the-art models. When serving the purpose of variable selection,
+trade-offs exist among methods of comparison. We implement the proposed
+model in an open-source R package \texttt{BHAM}, deposited at
+\url{https://github.com/boyiguo1/BHAM}. To maximize the flexibility of
+smooth function specification, we deploy the same programming grammar as
+in the state-of-the-art package \texttt{mgcv}, in contrast to previous
+tools where smooth functions are limited to the default ones. Ancillary
+functions are provided for model specification in high-dimensional
+settings, curve plotting and functional selection.
 
@@
 proof for the consistency of variable selection using group
-spike-and-slab LASSO prior.
-\textcolor{red}{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection.
-Moreover, the proposed model can easily generalize to other families of priors and smooth functions if desired.
-Not focused in this manuscript, the generalization is described in the Supporting Information.}
-
-\textcolor{red}{During designing and analyzing the simulation study, we made couple of interesting observations.
-First of all, variable selection is a delicate topic in the context of predictive modeling.
-When prediction performance is used to tune a model, the model could possibly include noise variables in models, for example LASSO and LASSO-based models.
-\cite{Wu2019} Moreover, bi-level selection is a more complex problem than variable selection.
-The complexity shows on the validity of the effect hierarchy principle.
-While most functional forms follow that the linear component exists in the nonlinear function, there are functions that don't follow it, e.g.
-$x^2$.
-The proposed prior and spikeSlabGAM employ different structures: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent.
-The different prior setups lead to trade-offs for the purpose of bi-level selection.
-We recommend to use more judgment in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist.
-Secondly, we find the performance of the proposed model is more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings.
-Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend fine-tuning the model with a granular sequence of $s_0$ for performance improvement.}
-
-Our future efforts are direct to
-\textcolor{red}{uncertainty inference of the proposed model}, survival
-analysis and integrative analysis.
-\textcolor{red}{Using EM-CD algorithm to fit the proposed BHAM is incapable of conducting uncertainty inference.
-We derive the EM-Iterative Weighted Least Square algorithm (EM-IWLS, see the Supporting Information) as an alternative.
-Instead of the Coordinate Descent algorithm, we use the Iterative Weighted Least Square algorithm in the EM procedure.
-The EM-IWLS algorithm is previously used to fit Bayesian high-dimensional generalized linear models \cite{Yi2012}, and deliver estimates of the coefficient variance-covariance matrix.
-Due to the space limit, technical details will be explained in a future manuscript.}
-While the proposed model addresses a great deal of analytic problems,
-analyzing the time-to-event outcome remains unsolved. A naive approach
-would be convert a time-to-event outcome to a Poisson outcome following
-Whitehead \citep{Whitehead1980}. However, it would be more efficient to
-directly fit Cox models via penalized pseudo likelihood function
+spike-and-slab LASSO prior. The proposed model focuses on improving
+prediction performance for high-dimensional GAM, with the capability of
+bi-level selection. Moreover, the proposed model can easily generalize
+to other families of priors and smooth functions if desired. Not focused
+in this manuscript, the generalization is described in the Supporting
+Information.
+
+During designing and analyzing the simulation study, we made couple of
+interesting observations. First of all, variable selection is a delicate
+topic in the context of predictive modeling. When prediction performance
+is used to tune a model, the model could possibly include noise
+variables in models, for example LASSO and LASSO-based models.
+\cite{Wu2019} Moreover, bi-level selection is a more complex problem
+than variable selection. The complexity shows on the validity of the
+effect hierarchy principle. While most functional forms follow that the
+linear component exists in the nonlinear function, there are functions
+that don't follow it, e.g.~\(x^2\). The proposed prior and spikeSlabGAM
+employ different structures: the proposed prior imposes effect hierarchy
+while spikeSlabGAM treats the selection of linear and nonlinear
+components independent. The different prior setups lead to trade-offs
+for the purpose of bi-level selection. We recommend to use more judgment
+in bi-level selection, either relying on heuristic knowledge to choose
+appropriate prior or exploring multiple models when heuristic knowledge
+doesn't exist. Secondly, we find the performance of the proposed model
+is more sensitive to the granularity of \(s_0\) sequence in the
+high-dimensional settings than in the lower dimension settings. Even
+though the current default sequence of \(s_0\) can result in reasonable
+performance shown in the simulation studies, we recommend fine-tuning
+the model with a granular sequence of \(s_0\) for performance
+improvement.
+
+Our future efforts are direct to uncertainty inference of the proposed
+model, survival analysis and integrative analysis. Using EM-CD algorithm
+to fit the proposed BHAM is incapable of conducting uncertainty
+inference. We derive the EM-Iterative Weighted Least Square algorithm
+(EM-IWLS, see the Supporting Information) as an alternative. Instead of
+the Coordinate Descent algorithm, we use the Iterative Weighted Least
+Square algorithm in the EM procedure. The EM-IWLS algorithm is
+previously used to fit Bayesian high-dimensional generalized linear
+models \cite{Yi2012}, and deliver estimates of the coefficient
+variance-covariance matrix. Due to the space limit, technical details
+will be explained in a future manuscript. While the proposed model
+addresses a great deal of analytic problems, analyzing the time-to-event
+outcome remains unsolved. A naive approach would be convert a
+time-to-event outcome to a Poisson outcome following Whitehead
+\citep{Whitehead1980}. However, it would be more efficient to directly
+fit Cox models via penalized pseudo likelihood function
 \citep{Simon2011}. Meanwhile, with the growing understanding of
@@
 \matrix (penPrior) [right = 2cm of linearPrior, matrix of nodes, column sep = 0mm, row sep = 0.7cm] {
-  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \bg{\gamma_{j}}\theta_j)$ };\\
+  \node (penGamma) [outCome] { $\gamma_{j}^\tp \sim Bin(1, \gamma_{j}\theta_j)$ };\\
   \node (penBeta) [outCome] { $\beta_{jk}^\tp \sim  DE(0,(1-\gamma^\tp_{j}) s_0 + \gamma^\tp_{j} s_1)$};\\
@@
 \draw[->] (penGamma) -- (penBeta);
-\draw[->, draw = red] (linearGamma) -- (penGamma);
+\draw[->] (linearGamma) -- (penGamma);
 \draw[->] (ab) -- (Theta);
